### 2018-06-27 Sexual Selection and Extinction.Rmd
Initial commit

---
title: 'Sexual selection and extinction: a macro-evolutionary approach'
author: "Justin G. Cally^1^, Iliana Medina Guzmán^1^ Luke Holman^1^, Devi Stuart-Fox^1^ <br></br> <br></br> ^1^The University of Melbourne"
subtitle: Supplementary Material
bibliography: References/SS_Extinction.bib
csl: References/Proc_b.csl
output:
  html_document:
    toc: true # table of content true
    toc_float: true # make 
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    theme: cosmo
    code_folding: hide # awesome buttons to show/hide the code
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(pander)
library(knitr)
library(tidyr)
library(mgcv)
library(diversitree)
library(repmis)
library(lme4)
library(ape)
library(geiger)
library(nlme)
library(phytools)
library(brms)
library(ggridges)
library(caper)
source("bamm_extinction/functions/check_and_fix_ultrametric.R")
source("functions/essim.R")
```

Here we provide the supplementary material for our study on sexual selection and extinction risk in passerine birds.

#Compiling the data

For this project we aimed to compile data on a large set of species in the Passerine bird order that would allow us to account for many evolutionary and environmental predictors of extinction. The following details the source and use of the data.

##Extinction Risk

To obtain extinction risk estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM). We document the control file and process for obtaining extinction rates from BAMM in an accompanying R markdown in the ``BAMM`` folder. Additionally, we compared our BAMM output files (event data) against a pre-existing extinction rate dataset [@Harvey_2017].

Rather than obtaining one estimate of extinction rate, we ran BAMM XXXX times to obtain uncertainty estimates of the extinction risk generated from the phylogeny created by @Jetz_2012.

@Jetz_2012 also provides us with a compiled IUCN threat level listing for all species from 2010 standings.

##Sexual Selection

Our proxy for sexual selection was based on measures of sexual dichromatism. Notably, we use two existing datasets; a complete set of male and female plumage scores in Passerines from @Dale_2015 and reflectance data from 1000 birds from @Armenta_2008.

Briefly, @Dale_2015 obtained male and female plumage score, which can be divided to obtain a sexual dimorphism index (SDi). The plumage score was scored using the _Handbook of the Birds of the World_ @delhoyo2011. By scanning images of males and females in this book across multiple patches @Dale_2015 were able to obtain mean plumage scores from RGB values. Notably concern about the accuracy of these measurements were raised, thus this dataset was previously compared with a more thorough reflectance dataset of Australian birds, the two datasets correlated but was expectedly noisy (R2 = 0.67, P < 0.0001). Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008 which has colour discriminability, segment classification and PCA.

Alongside sexual dichromatism data there is also partial data for other metrics analysed in @Dale_2015 such as: Body size, tropical life history, Sexual selection, Cooperative breeding and Migration.

##Environmental predictors

We obtained species range distribution maps from Birdlife International (through IUCN). These species range maps cover ~ 7,000 species of birds. From these ranges we were able to obtain estimates of:

* Range Size
* Average and standard deviation in 19 bioclimatic variables (each range randomly sampled 1000 times)
* Average and standard deviation in 19 bioclimatic variables from the last glaciam maximum (LGM)
* Net primary productivity (NPP) estimates and variability
* Average and standard deviation of population density in the species ranges.

The code for extracting the environmental data from the species is provided in another document XXXX.

#Generating biologically relevent predictors of extinction

Here we assess the relationship between sexual selection and extinction risk. However, in doing so we attempt to take into account as many other predictors of extinction as possible, primarily through environmental variables. Basically, our model structure seeks to contain:

Extinction rate ~ Sexual selection
                   + Range size
                   + Short temporal variability of temperature (mean BIOCLIM4)
                   + Short temporal variability of precipitation (mean BIOCLIM15)
                   + Spatial variability of temperature (PCA1)
                   + Spatial variability of precipitation (PCA2)
                   + Long-term variability of temperature (BIOCLIM1)
                   + Long-term variability of precipitation (BIOCLIM12)
                   + Variability of NPP

These predictors can be generated from the compiled datasets seen here:

```{r}
complete.dataframe <- read.csv('data/complete.dataframe.csv')
```

##Spatial Variability PCAs

To obtain estimates of spatial variability we can obtain PCAs of the standard errors of the bioclimatic variables (excluding seasonality in temperature and precipitation). From there we can run GAM models with the PCA components and range size and extract residuals. We do this as we expect variability will increase alongside increased range size, thus taking the residuals allows us to correct for this association.

```{r}
restricted.data <- complete.dataframe %>% drop_na(bioclim1) #Drop species without environmental data
PCA.bioclim <- prcomp(restricted.data[c(34,35,36,38:47,49:52)], #Select se.bioclim
                      scale = TRUE, center = TRUE)
PCA.predictions <- predict(PCA.bioclim)
restricted.data <- cbind(restricted.data, PCA.predictions)

#run GAM
PC1.gam <- mgcv::gam(PC1 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")

#take residuals
restricted.data$residuals.PC1 <- residuals.gam(PC1.gam)

#plot
plot.gam(PC1.gam, residuals = T)
```

We can do the same for PC2 (precipitation)

```{r}
PC2.gam <- mgcv::gam(PC2 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")

#Take residuals
restricted.data$residuals.PC2 <- residuals.gam(PC2.gam)

plot.gam(PC2.gam, residuals = T)
```

##Long term climate variability (LGM anomaly)

To gain estimates for change in climate over the past 22,000 we can use the difference in annual mean temperature (BIO1) and annual precipitation (BIO12) between the LGM (CCSM4 models) and present values. Additionally we can calculate the difference in variability based on the standard deviation.

```{r}
#Change in mean temp
restricted.data$BIO1.LGM.Diff <- restricted.data$bioclim1 - restricted.data$cclgmbi1
#Change in variation temp
restricted.data$BIO1.LGM.var.Diff <- restricted.data$se.bioclim1 - restricted.data$se.cclgmbi1

#Change in mean precipitation
restricted.data$BIO12.LGM.Diff <- restricted.data$bioclim12 - restricted.data$cclgmbi12
#Change in variation precipitation
restricted.data$BIO12.LGM.var.Diff <- restricted.data$se.bioclim12 - restricted.data$se.cclgmbi12

#Inspect these changes 
restricted.data %>% ggplot(aes(x=BIO1.LGM.Diff, y = BIO1.LGM.var.Diff))+
  geom_point()

restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = BIO12.LGM.var.Diff))+
  geom_point()
```

From these plots we can see that on average temperature has increased but the variation in tempere has decresed (narrowed). Alternatively, precipitation sometimes increases and sometimes decreases. When precipitation increases there isw also an increase in variation in that range. However, when annual precipitation decreases there is a decrease in variation.

##Variation in NPP

Similar to the PCA's we expect NPP variation to be correlated with range size. To account for this we can run a GAM and obtain residuals.

```{r}
#run gam
NPP.gam <- mgcv::gam(NPP.se ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")

#obtain residuals
restricted.data$NPP.res <- residuals.gam(NPP.gam)

#plot
plot(NPP.gam, residuals = T)
```

_______________

##Preliminary Analysis of data

Now that we have all our relevent predictors we can combine these with the tip extinction rates from BAMM (code for this is provided in the repository). First, let us inspect the tip rates that @Harvey_2017 obtained in BAMM with our predictors.

```{r}
complete.dataframe %>% ggplot(aes(x=rab.ex.rate.lumped, y=rab.ex.rate.split))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()
```

The extinction rates of the lumped and split datasets show high similarity, insisting that we could use either.

Let us create a relatively simple model
```{r}
model.1 <- lm(rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                data = restricted.data)
summary(model.1)
```

Now we can try a PGLS model 
```{r}
hackett.consensus.tree <- read.nexus('bamm_extinction/data/trees/hackett_dale_sub.consensus.nex')
restricted.data <- restricted.data %>% drop_na(rab.ex.rate.split)
rownames(restricted.data) <- restricted.data$TipLabel
pruned.tree<-drop.tip(hackett.consensus.tree,hackett.consensus.tree$tip.label[-match(restricted.data$TipLabel, hackett.consensus.tree$tip.label)])
pgls.model.1 <- gls(rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                correlation = corBrownian(phy = pruned.tree), 
                data = restricted.data, 
                method = "REML")
summary(pgls.model.1)
```

Or perhaps OU model rather than Brownian:

```{r}
pgls.model.3 <- gls(rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                correlation = corMartins(1, phy = pruned.tree), 
                data = restricted.data, 
                method = "REML")
summary(pgls.model.3)
```

Now a Bayesian model that uses a covariance matrix from the phylogeny:

```{r}
pruned.tree$edge.length[pruned.tree$edge.length<0]<-1e-9
pruned.tree$edge.length[pruned.tree$edge.length==0]<-1e-9
pruned.tree <- check_and_fix_ultrametric(pruned.tree)
inv.phylo <- MCMCglmm::inverseA(pruned.tree, nodes = "TIPS", scale = TRUE)
A <- solve(inv.phylo$Ainv)
rownames(A) <- rownames(inv.phylo$Ainv)
```

```{r, eval=FALSE}
model_bayesian <- brm(
  rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res, data = restricted.data, 
  family = gaussian(), 
  cov_ranef = list(phylo = A),
                seed = 1,
                cores = 4, chains = 4, iter = 4000, #Run 4 chains in parallel for 4000 iterations (2000 are burn in)
                control = list(adapt_delta = 0.99, max_treedepth = 15))
```

```{r}
summary(model_bayesian)
```

Try looking at speciation rate using ES-Sim adapted by @Harvey_essim_2017 and found here https://github.com/mgharvey/ES-sim

```{r}
Essim.frame <- restricted.data$SDi
names(Essim.frame) <- restricted.data$TipLabel

essim.results <- essim(pruned.tree, Essim.frame, nsim = 1000, return.es = TRUE)
```

We can then add the essim.results and run a similar PGLS model
```{r}
restricted.data$ESsim <- essim.results[["es"]]

restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = (ESsim))) + 
  geom_point()+
  geom_smooth(method = "lm")
```

 Once we have our EsSim results we can use them in a linear model to assess the relationship between speciation and our predictors

```{r}
model.2 <- lm(ESsim ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                data = restricted.data)
summary(model.2)
```

Have a look at how these would look if we obtained model predictions for each data point.

```{r, fig.width=10, fig.height=5}
predict.frame <- cbind(predict(model.2),restricted.data)
predict.frame %>% ggplot(aes(y = `predict(model.2)`, x = SDi)) + 
  geom_point(aes(color = BLFamilyLatin))+
  scale_color_hue(h = c(0,100), l = 50) +
  geom_smooth(color = 'black')+
  theme_minimal()
```

It looks like some of the higher SDi's have slightly higher speciation rates. Perhaps we can look at those obviously dimorphic groups as category of its own.

```{r}
restricted.data %>% ggplot(aes(SDi))+
  geom_histogram(binwidth = 0.015)+
  theme_minimal()
```

### 2018-06-27 Sexual_Selection_and_Extinction.Rmd
Initial commit

--- 2018-06-27 Sexual Selection and Extinction.Rmd
@@
 library(brms)
-library(ggridges)
-library(caper)
 source("bamm_extinction/functions/check_and_fix_ultrametric.R")
-source("functions/essim.R")
 ```
@@
                    + Range size
-                   + Short temporal variability of temperature (mean BIOCLIM4)
-                   + Short temporal variability of precipitation (mean BIOCLIM15)
-                   + Spatial variability of temperature (PCA1)
-                   + Spatial variability of precipitation (PCA2)
-                   + Long-term variability of temperature (BIOCLIM1)
-                   + Long-term variability of precipitation (BIOCLIM12)
+                   + Short temporal variability of temperature (mean BIO4)
+                   + Short temporal variability of precipitation (mean BIO15)
+                   + Spatial variability of temperature
+                   + Spatial variability of precipitation
+                   + Long-term variability of temperature
+                   + Long-term variability of temperature
                    + Variability of NPP
@@
 
-Or perhaps OU model rather than Brownian:
+How about a PGLS model that uses the IUCN rankings as a categorical response:
 
 ```{r}
-pgls.model.3 <- gls(rab.ex.rate.split ~ SDi 
+pgls.model.2 <- gls(X2010.IUCN.Red.List.category ~ SDi 
                 + range.size.m2 
@@
                 + NPP.res,
-                correlation = corMartins(1, phy = pruned.tree), 
+                correlation = corBrownian(phy = pruned.tree), 
                 data = restricted.data, 
                 method = "REML")
-summary(pgls.model.3)
+summary(pgls.model.2)
 ```
@@
 
-```{r, eval=FALSE}
-model_bayesian <- brm(
+```{r}
+model_simple <- brm(
   rab.ex.rate.split ~ SDi 
@@
 ```{r}
-summary(model_bayesian)
+summary(model_simple)
 ```
-
-Try looking at speciation rate using ES-Sim adapted by @Harvey_essim_2017 and found here https://github.com/mgharvey/ES-sim
-
-```{r}
-Essim.frame <- restricted.data$SDi
-names(Essim.frame) <- restricted.data$TipLabel
-
-essim.results <- essim(pruned.tree, Essim.frame, nsim = 1000, return.es = TRUE)
-```
-
-We can then add the essim.results and run a similar PGLS model
-```{r}
-restricted.data$ESsim <- essim.results[["es"]]
-
-restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = (ESsim))) + 
-  geom_point()+
-  geom_smooth(method = "lm")
-```
-
- Once we have our EsSim results we can use them in a linear model to assess the relationship between speciation and our predictors
-
-```{r}
-model.2 <- lm(ESsim ~ SDi 
-                + range.size.m2 
-                + bioclim4
-                + bioclim15
-                + residuals.PC1
-                + residuals.PC2
-                + BIO1.LGM.Diff
-                + BIO12.LGM.Diff
-                + NPP.res,
-                data = restricted.data)
-summary(model.2)
-```
-
-Have a look at how these would look if we obtained model predictions for each data point.
-
-```{r, fig.width=10, fig.height=5}
-predict.frame <- cbind(predict(model.2),restricted.data)
-predict.frame %>% ggplot(aes(y = `predict(model.2)`, x = SDi)) + 
-  geom_point(aes(color = BLFamilyLatin))+
-  scale_color_hue(h = c(0,100), l = 50) +
-  geom_smooth(color = 'black')+
-  theme_minimal()
-```
-
-It looks like some of the higher SDi's have slightly higher speciation rates. Perhaps we can look at those obviously dimorphic groups as category of its own.
-
-```{r}
-restricted.data %>% ggplot(aes(SDi))+
-  geom_histogram(binwidth = 0.015)+
-  theme_minimal()
-```

### 2018-06-27 bamm_extinction/BirdExtinction.Rmd
Initial commit

---
title: "Bird Extinction"
output: html_notebook
---

```{r}
library(ape)
library(phytools)
library(dplyr)
library(coda)
library(ggplot2)
library(ggridges)
library(phytools)
library(BAMMtools)
library(diversitree)
source("functions/R_test_BAMM_functions.R")
source("functions/check_and_fix_ultrametric.R")
# bird.tree.eric <- read.nexus('data/ErisonGeneticFull.nex')
# averageTree(bird.tree.eric)
```

Let's read in our consensus trees for Hackett and Ericson backbones for all the Passerines with genetic data:
```{r}
consensus.eric <- file.choose()
  # read.nexus('data/trees/consensus_trees/ericson_dale_sub.consensus.nex')
consensus.hackett <- read.nexus('data/trees/consensus_trees/hackett_dale_sub.consensus.nex')
mcc.trees.eric <- read.nexus('data/trees/ericson_dale_sub.nex')
```

For now, let's use the Ericson tree for speciation/extinction analysis. We want to have a look at the species in the tree and how they match up with the full Dale dataset:

```{r}
consensus.eric$tip.label
```

##BAMM Setup

###Consensus Tree

We need branches to be ultrametric, binary and all edge lengths > 0. We can check that with the following:

```{r}
# is.ultrametric(consensus.eric)
# is.binary.tree(consensus.eric)
# # Now to check min branch length:
# min(consensus.eric$edge.length)
# 
# #try resolving negative branch lengths
# 
# #So in this case we ruin our tree for BAMM. So, instead let us manually change branch lengths to slightly positive values
# consensus.eric$edge.length[consensus.eric$edge.length<0]<-0
# 
# #And now lets see if our tree is hunky dory
# is.ultrametric(consensus.eric)
# is.binary.tree(consensus.eric)
# # Now to check min branch length:
# min(consensus.eric$edge.length)
# 
# #Now it is no longer ultrametric 
# consensus.eric <- force.ultrametric(consensus.eric, method = "nnls")
# 
# #Change non-zero to slight positive
# consensus.eric$edge.length[consensus.eric$edge.length==0]<-1e-9
# 
# #And now once again
# is.ultrametric(consensus.eric)
# is.binary.tree(consensus.eric)
# # Now to check min branch length:
# min(consensus.eric$edge.length)
```
So our minimum branch length is zero... hmmm BAMM says that it should be greater than zero, but chaging this will make the tree no longer ultrametric.

Instead, let's use a previously written function that might handle the consensus tree better

```{r}
consensus.eric <- read.nexus('data/consensus_trees/ericson_dale_sub.consensus.nex')

is.ultrametric(consensus.eric)
is.binary.tree(consensus.eric)
# Now to check min branch length:
min(consensus.eric$edge.length)

consensus.eric$edge.length[consensus.eric$edge.length<0]<-1e-9

#No longer ultrametric
consensus.eric <- check_and_fix_ultrametric(consensus.eric)

is.ultrametric(consensus.eric)
is.binary.tree(consensus.eric)
# Now to check min branch length:
min(consensus.eric$edge.length)
```

Create a control file for BAMM analysis. Try and make them as close to Huang and Rabosky 2014 as possible

```{r}
priors <- setBAMMpriors(consensus.eric, outfile = NULL)
write.tree(consensus.eric, 'consensus_eric.tre')

# Here is a block of parameters for the control file
params = list(treefile = 'consensus_eric.tre',
            sampleFromPriorOnly = 0,
            runMCMC = 1,
            initializeModel = 1,
            useGlobalSamplingProbability = 1,
            globalSamplingFraction = 0.67,
            updateLambdaInitScale = 2.0,
            updateMuInitScale = 2.0,
            updateLambdaShiftScale = 0.05,
            updateEventRateScale = 4.0,
            localGlobalMoveRatio = 10.0,
            modeltype = 'speciationextinction', #rather than trait

            mcmcWriteFreq = '100000', #turns to one
            eventDataWriteFreq = '100000',
            printFreq = '100000',
            acceptanceResetFreq = '100000',

            lambdaInitPrior = as.numeric(priors['lambdaInitPrior']),
            lambdaShiftPrior = '0.05',
            muInitPrior = as.numeric(priors['muInitPrior']),

            numberOfGenerations = '150000000', #Command line runs 150 million times
            runMCMC = 1,
            overwrite = 1,
            seed = -1,
            expectedNumberOfShifts = '24', #Huang and Rabosky had 40 with a full bird tree of 6,670 species. So with 4000 we would expect slightly less

            lambdaIsTimeVariablePrior = '0',
            updateRateLambdaShift = '1',
            updateRateEventPosition = '1',
            numberOfChains = '1',
            outName = 'consensus.eric')

bammcontrolfile <- paste("control_", "consensus_eric", ".txt", sep="")

# Now writing control parameters to file
generateControlFile(file = bammcontrolfile, type = "diversification",
        params = params)
```

We now run the BAMM through command line. Although the source of BAMM is in another directory we have made a sudo link to it so it can be run from anywhere (e.g. working directory of this project). The command line code is:

``bamm -c control_consensus_eric.txt``

> cool

###MCC Tree

```{r}
is.ultrametric(mcc.trees.eric[["tree_2241"]])
is.binary.tree(mcc.trees.eric[["tree_2241"]])
# Now to check min branch length:
min(mcc.trees.eric[["tree_2241"]]$edge.length)
```

Create a control file with same parameters as the consensus tree:

```{r}
priors <- setBAMMpriors(mcc.trees.eric[["tree_2241"]], outfile = NULL)
write.tree(mcc.trees.eric[["tree_2241"]], 'mcc_eric_2241.tre')

# Here is a block of parameters for the control file
params = list(treefile = 'mcc_eric_2241.tre',
            sampleFromPriorOnly = 0,
            runMCMC = 1,
            initializeModel = 1,
            useGlobalSamplingProbability = 1,
            globalSamplingFraction = 0.67,
            updateLambdaInitScale = 2.0,
            updateMuInitScale = 2.0,
            updateLambdaShiftScale = 0.05,
            updateEventRateScale = 4.0,
            localGlobalMoveRatio = 10.0,
            modeltype = 'speciationextinction', #rather than trait

            mcmcWriteFreq = '100000', #turns to one
            eventDataWriteFreq = '100000',
            printFreq = '100000',
            acceptanceResetFreq = '100000',

            lambdaInitPrior = as.numeric(priors['lambdaInitPrior']),
            lambdaShiftPrior = '0.05',
            muInitPrior = as.numeric(priors['muInitPrior']),

            numberOfGenerations = '150000000', #Command line runs 150 million times
            runMCMC = 1,
            overwrite = 1,
            seed = -1,
            expectedNumberOfShifts = '24', #Huang and Rabosky had 40 with a full bird tree of 6,670 species. So with 4000 we would expect slightly less

            lambdaIsTimeVariablePrior = '0',
            updateRateLambdaShift = '1',
            updateRateEventPosition = '1',
            numberOfChains = '1',
            outName = 'consensus.eric')

bammcontrolfile <- paste("control_", "mcc_eric_2241", ".txt", sep="")

# Now writing control parameters to file
generateControlFile(file = bammcontrolfile, type = "diversification",
        params = params)
```

##QUASSE test

##Analysis

Now we have a series of trees and data we can read in:Let's check for convergence

```{r}
# Read in tree, mcmc data, and BAMM output
tree.BAMM.eric.bird  <- read.tree("consensus_eric.tre")
mcmc.BAMM.eric.bird <- read.csv( "consensus.eric_mcmc_out.txt" , stringsAsFactors=F)
ed.BAMM.eric.bird <- getEventData(tree.BAMM.eric.bird ,  "consensus.eric_event_data.txt", burnin=0.1, nsamples=200)
```

```{r}
plot(mcmc.BAMM.eric.bird$logLik ~ mcmc.BAMM.eric.bird$generation)
```

--> Looks like it has converged so let's discard burn in:

```{r}
burnstart <- floor(0.1 * nrow(mcmc.BAMM.eric.bird))
postburn <- mcmc.BAMM.eric.bird[burnstart:nrow(mcmc.BAMM.eric.bird), ]

#We can also check effective population sizes of the log-likelihhod and number of shift events in each sample
#We want at least 200 (although that's on the low side)

effectiveSize(postburn$N_shifts)
effectiveSize(postburn$logLik)
```

###How many rate shifts?

Compute the posterior probabilities of models
```{r}
post_probs <- table(postburn$N_shifts) / nrow(postburn)
names(post_probs)
```

And to compute posterior odds ratio for two models:

```{r}
post_probs['11'] / post_probs['13']
```

We can obtain bayes factor matrix as such: (The Bf can help us determine the best model, with the overall best model from a BAMM analysis is the model with the highest Bayes factor relative to the null model, M_0.(ususally))
```{r}
bfmat <- computeBayesFactors(mcmc.BAMM.eric.bird, expectedNumberOfShifts=24, burnin=0.1)
```
This suggests that it is most likely that there are 20 rate shifts as Bf ~ 235

we can plot the prior via:

```{r}
plotPrior(mcmc.BAMM.eric.bird, expectedNumberOfShifts=24)
```

```{r}
summary(ed.BAMM.eric.bird)
```

Plot phylorate plot, good way of visualizing diversification rates
```{r, fig.width = 6.5, fig.height = 6.5}
pdf("figures/BAMMplot.pdf", width = 9, height = 9)
BAMMplot <- plot.bammdata(ed.BAMM.eric.bird, breaksmethod = "jenks", spex="e", method = "polar", lwd=0.5, tau=0.003, logcolor = T)
mtext("BAMM estimated extinction rates", side=3, cex=1.5)
addBAMMlegend(BAMMplot, location="bottomright")
dev.off()
```

Get histogram of extinction rates

```{r}
BAMMplot[["colordens"]][["kde.y"]] <- log1p(BAMMplot[["colordens"]][["kde.y"]] )
pdf("figures/BAMMplot_hist.pdf", width = 8, height = 4)
#create phylorate plot to generate output
ratesHistogram(BAMMplot, plotBrks = F, xlab = 'Extinction rate', ylab = 'Frequency (log-transformed)', lwd = 0.05)
dev.off()
```

##Credible number of shifts

To plot the credible shift set, we need the prior distribution on the number of rate shifts (this is generated internally by BAMMtools). We can then estimate the credible set of rate shifts using the BAMMtools function credibleShiftSet:

```{r}
css <- credibleShiftSet(ed.BAMM.eric.bird, expectedNumberOfShifts=1, threshold=5, set.limit = 0.95)
```

Now we obtain the number of distinct shifts:

```{r}
css$number.distinct
```

--> We get a lot more shifts than we expect under a traditional (AIC) approach

```{r}
summary(css)
```

```{r}
plot.credibleshiftset(css)
```

##Hackett Backbone

Let's make sure our tree is good to go:

```{r}
is.ultrametric(consensus.hackett)
is.binary.tree(consensus.hackett)
# Now to check min branch length:
min(consensus.hackett$edge.length)

#try resolving negative branch lengths

#So in this case we ruin our tree for BAMM. So, instead let us manually change branch lengths to slightly positive values
consensus.hackett$edge.length[consensus.hackett$edge.length<0]<-0

#And now lets see if our tree is hunky dory
is.ultrametric(consensus.hackett)
is.binary.tree(consensus.hackett)
# Now to check min branch length:
min(consensus.hackett$edge.length)

#Now it is no longer ultrametric 
consensus.hackett <- force.ultrametric(consensus.hackett, method = "nnls")

#Change non-zero to slight positive
consensus.hackett$edge.length[consensus.hackett$edge.length==0]<-1e-9

#And now once again
is.ultrametric(consensus.hackett)
is.binary.tree(consensus.hackett)
# Now to check min branch length:
min(consensus.hackett$edge.length)
```
So our minimum branch length is zero... hmmm BAMM says that it should be greater than zero, but chaging this will make the tree no longer ultrametric.

#Let's move forward and create a control file for the

Create a control file for BAMM analysis. Try and make them as close to Huang and Rabosky 2014 as possible

```{r}
priors <- setBAMMpriors(consensus.hackett, outfile = NULL)
write.tree(consensus.hackett, 'consensus_hackett.tre')

# Here is a block of parameters for the control file
params = list(treefile = 'consensus_hackett.tre',
            sampleFromPriorOnly = 0,
            runMCMC = 1,
            initializeModel = 1,
            useGlobalSamplingProbability = 1,
            globalSamplingFraction = 0.67,
            updateLambdaInitScale = 2.0,
            updateMuInitScale = 2.0,
            updateLambdaShiftScale = 0.05,
            updateEventRateScale = 4.0,
            localGlobalMoveRatio = 10.0,
            modeltype = 'speciationextinction', #rather than trait

            lambdaInitPrior = as.numeric(priors['lambdaInitPrior']),
            lambdaShiftPrior = '0.05',
            muInitPrior = as.numeric(priors['muInitPrior']),

            numberOfGenerations = '100000000', #Command line runs 100 million times
            runMCMC = 1,
            overwrite = '0',
            seed = -1,
            expectedNumberOfShifts = '24', #Huang and Rabosky had 40 with a full bird tree of 6,670 species. So with 4000 we would expect slightly less

            lambdaIsTimeVariablePrior = '0',
            updateRateLambdaShift = '1',
            updateRateEventPosition = '1',
            numberOfChains = '1',
            outName = 'consensus.hackett')

bammcontrolfile <- paste("control_", "consensus_hackett", ".txt", sep="")

# Now writing control parameters to file
generateControlFile(file = bammcontrolfile, type = "diversification",
        params = params)
```

``bamm -c control_consensus_hackett.txt``

> Cool, now let's run the same checks on it as the

```{r}
# Read in tree, mcmc data, and BAMM output
tree.BAMM.hackett.bird  <- read.tree("hackett_files/consensus_hackett.tre")
mcmc.BAMM.hackett.bird   <- read.csv( "hackett_files/consensus.hackett_mcmc_out.txt" , stringsAsFactors=F)
ed.BAMM.hackett.bird     <- getEventData(tree.BAMM.hackett.bird ,  "hackett_files/consensus.hackett_event_data.txt", burnin=0.1, nsamples=200)
```

Let's check for convergence
```{r}
plot(mcmc.BAMM.hackett.bird$logLik ~ mcmc.BAMM.hackett.bird$generation)
```

--> Looks like it has converged so let's discard burn in:

```{r}
hackett_burnstart <- floor(0.1 * nrow(mcmc.BAMM.hackett.bird))
hackett_postburn <- mcmc.BAMM.hackett.bird[burnstart:nrow(mcmc.BAMM.hackett.bird), ]

#We can also check effective population sizes of the log-likelihhod and number of shift events in each sample
#We want at least 200 (although that's on the low side)

effectiveSize(hackett_postburn$N_shifts)
effectiveSize(hackett_postburn$logLik)
```

```{r}
hackett_bfmat <- computeBayesFactors(mcmc.BAMM.hackett.bird, expectedNumberOfShifts=24, burnin=0.1)
```

```{r}
plotPrior(mcmc.BAMM.hackett.bird, expectedNumberOfShifts=24)
```

```{r}
summary(ed.BAMM.hackett.bird)
```

Plot phylorate plot, good way of visualizing diversification rates
```{r, fig.width = 7, fig.height = 7}
BAMMplot_hackett <- plot.bammdata(ed.BAMM.hackett.bird, breaksmethod = "jenks", spex="e", method = "polar", lwd=2, tau=0.003)
mtext("BAMM estimated rates", side=3, cex=1.5)
addBAMMlegend(BAMMplot_hackett, location="left")
```

##Tip Extinction Rates

We can get tip extinction rates to use in our dataset.

```{r}
hackett.extinction.frame <-data.frame(ed.BAMM.hackett.bird$tip.label, ed.BAMM.hackett.bird$meanTipMu)

colnames(hackett.extinction.frame) <- c("binomial", "extinction.rate.hackett")
hackett.extinction.frame$binomial <- gsub("_", " ", hackett.extinction.frame$binomial)
```

Now for Ericson rates
```{r}
ericson.extinction.frame <-data.frame(ed.BAMM.eric.bird$tip.label, ed.BAMM.eric.bird$meanTipMu)

colnames(ericson.extinction.frame) <- c("binomial", "extinction.rate.ericson")
ericson.extinction.frame$binomial <- gsub("_", " ", ericson.extinction.frame$binomial)
```

Ok we have our extinction data now let's merge SS data

```{r}
colnames(dale.data)[1] <- "binomial"
prelim.dataframe <- left_join(hackett.extinction.frame, ericson.extinction.frame, by = "binomial")
prelim.dataframe <- left_join(prelim.dataframe, dale.data, by = "binomial")

#New col of male plumage vs female plumage

prelim.dataframe$SDi <- prelim.dataframe$Male_plumage_score/prelim.dataframe$Female_plumage_score
```

Get a ggplot
```{r}
prelim.dataframe %>% ggplot(aes(x = (SDi), y = log(extinction.rate)))+
  geom_point()+
  geom_smooth()
```

Let's have a look at range size

```{r}
range.size <- read.csv('data/environmental/bird.range.size.csv')

prelim.dataframe <- left_join(prelim.dataframe, range.size)
```

```{r}
prelim.dataframe %>% ggplot(aes(y = range.size.m2, x = SDi))+
geom_point(aes(color = extinction.rate))+
  scale_color_gradient()
```

```{r}
summary(lm(extinction.rate ~ SDi, data = prelim.dataframe))
```

Let's add all the dataframes together
```{r}
#Bioclim.means
bioclim.means <- read.csv('data/environmental/bird.means.csv')
prelim.dataframe <- left_join(prelim.dataframe, bioclim.means)

#Bioclim.se
bioclim.se <- read.csv('data/environmental/bird.se.csv')
prelim.dataframe <- left_join(prelim.dataframe, bioclim.se, by = 'binomial')

#LGM means
LGM.means <- read.csv('data/environmental/LGM.bird.means.csv')
prelim.dataframe <- left_join(prelim.dataframe, LGM.means, by = 'binomial')

#LGM se
LGM.se <- read.csv('data/environmental/LGM.bird.se.csv')
prelim.dataframe <- left_join(prelim.dataframe, LGM.se, by = 'binomial')

#NPP means
NPP.means <- read.csv('data/environmental/NPP.bird.means.csv')
prelim.dataframe <- left_join(prelim.dataframe, NPP.means, by = 'binomial')

#NPP se
NPP.se <- read.csv('data/environmental/NPP.bird.se.csv')
prelim.dataframe <- left_join(prelim.dataframe, NPP.se, by = 'binomial')

#Pop Density means
Pop.density.means <- read.csv('data/environmental/Pop.density.means.csv')
prelim.dataframe <- left_join(prelim.dataframe, Pop.density.means, by = 'binomial')

#Pop density se
Pop.density.se <- read.csv('data/environmental/Pop.density.se.csv')
prelim.dataframe <- left_join(prelim.dataframe, Pop.density.se, by = 'binomial')

#IUCN rankings and Passerines from Jetz/birdtree master
IUCN.family <- read.csv('data/IUCN_Family.csv')
prelim.dataframe <- left_join(prelim.dataframe, IUCN.family, by = 'binomial')

write.csv(prelim.dataframe, 'data/complete.dataframe.csv')
```

Let's have a look at the extinction rate of Hackett vs Ericson

```{r}
prelim.dataframe %>% ggplot(aes(x = (extinction.rate.ericson), y = (extinction.rate.hackett)))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()+
  geom_smooth(method = "lm")
```

```{r}
summary(lm(extinction.rate.hackett ~ extinction.rate.ericson, data = prelim.dataframe))
```

How does IUCN rankings correlate with BAMM extinction rate

```{r}
prelim.dataframe %>% ggplot(aes(x = X2010.IUCN.Red.List.category, y = extinction.rate.hackett)) +
  geom_jitter()+
  geom_boxplot()+
  scale_y_log10()
```

--> Not really much correlation

How about IUCN and pop density

```{r}
prelim.dataframe %>% ggplot(aes(y = X2010.IUCN.Red.List.category, x = (Pop_Density))) +
  geom_density_ridges()
```

Let's try ES-Sim with our phylogeny and our traits

```{r}
#Try with the consensus eric tree and the SDi

source('ES-sim/R/essim.R')
essim(consensus.eric, prelim.dataframe, nsim = 1000)

```

Phylogenetic models can be fitted with brms: https://rdrr.io/cran/brms/f/vignettes/brms_phylogenetics.Rmd

```{r}
prelim.dataframe %>% ggplot(aes(y = log(extinction.rate.hackett), x = Sexual_selection_ppca))+
  geom_point()
```

```{r}
prelim.dataframe %>% ggplot(aes(x=SDi, y = PassNonPass))+
  geom_density_ridges()
```

Test paired t test

```{r}
x = c(1,4,5,3,7,7,8)
y = c(13,14,20,15,20,22,25)
t.test.test <- pairedSamplesTTest(~ x + y)
```

```{r}
load("hackett_files/Hackett_lumped_eventsample.rda")
rabosky.ext <- data.frame(ed[["tip.label"]], ed[["meanTipMu"]])
colnames(rabosky.ext) <- c("binomial", "rab.ext.rate")
rabosky.ext$binomial <- gsub("_", " ", rabosky.ext$binomial)
comparison.ext.rate <- right_join(rabosky.ext, hackett.extinction.frame, by = "binomial")

comparison.ext.rate %>% ggplot(aes(x=rab.ext.rate, y=extinction.rate.hackett))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()
```

```{r}
summary(lm(extinction.rate.hackett ~ rab.ext.rate, data = comparison.ext.rate))
```

###PCA

```{r}
restricted.data <- prelim.dataframe %>% drop_na(bioclim1)
PCA.bioclim <- prcomp(restricted.data[c(34,35,36,38:47,49:52)], scale = TRUE, center = TRUE)
PCA.predictions <- predict(PCA.bioclim)
restricted.data <- cbind(restricted.data, PCA.predictions)
```

```{r}
#PC1 is correlated with range size, looks exponential
restricted.data %>% ggplot(aes(x=range.size.m2, y = PC1))+
  geom_point()+
  scale_x_log10()+
  geom_smooth(method = 'auto')

#PC2 seems less correlated with range size
restricted.data %>% ggplot(aes( y = se.bioclim12, x=bioclim12))+
  geom_point()
```

```{r}
PC1.gam <- mgcv::gam(PC1 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")
plot.gam(PC1.gam, residuals = T)
```

This essentially gives me the same sort of thing as ggplot, although I am not too sure if log(range size) is appropriate.

We can then take residuals using:

```{r}
restricted.data$residuals.PC1 <- residuals.gam(PC1.gam)
```

We can do the same for PC2 (precipitation)

```{r}
PC2.gam <- mgcv::gam(PC2 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")
plot.gam(PC2.gam, residuals = T)
```

We can then take residuals using:

```{r}
restricted.data$residuals.PC2 <- residuals.gam(PC2.gam)
```

```{r}
restricted.data %>% ggplot(aes(x=residuals.PC1, y=residuals.PC2))+
  geom_point()
```

###Now get LGM Difference for mean and SE

```{r}
#Change in mean temp
restricted.data$BIO1.LGM.Diff <- restricted.data$bioclim1 - restricted.data$cclgmbi1
#Change in variation temp
restricted.data$BIO1.LGM.var.Diff <- restricted.data$se.bioclim1 - restricted.data$se.cclgmbi1

#Change in mean precipitation
restricted.data$BIO12.LGM.Diff <- restricted.data$bioclim12 - restricted.data$cclgmbi12
#Change in variation precipitation
restricted.data$BIO12.LGM.var.Diff <- restricted.data$se.bioclim12 - restricted.data$se.cclgmbi12

#Inspect these changes 
restricted.data %>% ggplot(aes(x=BIO1.LGM.Diff, y = BIO1.LGM.var.Diff))+
  geom_point()

restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = BIO12.LGM.var.Diff))+
  geom_point()
```

```{r}
css <- credibleShiftSet(ed, expectedNumberOfShifts=1, threshold=5, set.limit = 0.95)
summary(css)
css$number.distinct
plot.credibleshiftset(css)
```

```{r}
css2 <- credibleShiftSet(ed.BAMM.hackett.bird, expectedNumberOfShifts=1, threshold=5, set.limit = 0.95)
summary(css2)
css2$number.distinct
plot.credibleshiftset(css2)
```

```{r}
rabosky.spc <- data.frame(ed[["tip.label"]], ed[["meanTipLambda"]])
colnames(rabosky.spc) <- c("binomial", "rab.spc.rate")
my.spc <- data.frame(ed.BAMM.hackett.bird[["tip.label"]], ed.BAMM.hackett.bird[["meanTipLambda"]])
colnames(my.spc) <- c("binomial", "my.spc.rate")
comparison.spc.rate <- right_join(rabosky.spc, my.spc, by = "binomial")

comparison.spc.rate %>% ggplot(aes(x=rab.spc.rate, y=my.spc.rate))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()
```

###Armenta Data Vs Dale

```{r}
Armenta.dat <- read.csv('data/Armenta_2008.csv')
prelim.dataframe <- left_join(prelim.dataframe, Armenta.dat, by = 'binomial')
prelim.dataframe$Colour.discriminability <- as.numeric(prelim.dataframe$Colour.discriminability)

prelim.dataframe %>% ggplot(aes(x=Colour.discriminability, y = SDi))+
  geom_point()+
  geom_smooth(method = "lm")
```

```{r}
summary(lm(Colour.discriminability ~ SDi, data = prelim.dataframe))
```

### 2018-06-27 bamm_extinction/functions/check_and_fix_ultrametric.R

@@ -1,0 +1,17 @@
+check_and_fix_ultrametric <- function(phy){
+  
+  if (!is.ultrametric(phy)){
+    
+    vv <- vcv.phylo(phy)
+    dx <- diag(vv)
+    mxx <- max(dx) - dx
+    for (i in 1:length(mxx)){
+      phy$edge.length[phy$edge[,2] == i] <- phy$edge.length[phy$edge[,2] == i] + mxx[i]
+    }
+    if (!is.ultrametric(phy)){
+      stop("Ultrametric fix failed\n")
+    }	
+  }
+  
+  return(phy)
+}

### 2018-06-27 functions/essim.R

@@ -1,0 +1,59 @@
+essim <- function(phy, trait, nsim = 1000, es, return.es=FALSE) {
+  
+  require(ape)
+  require(mvtnorm)
+  
+  if(missing(es)) { # If inverse equal splits statistics not provided, calculate it
+    rootnode <- length(phy$tip.label) + 1
+    es <- numeric(length(phy$tip.label))
+    for (i in 1:length(es)){
+      node <- i
+      index <- 1
+      qx <- 0
+      while (node != rootnode){
+        el <- phy$edge.length[phy$edge[,2] == node]
+        node <- phy$edge[,1][phy$edge[,2] == node]			
+        qx <- qx + el* (1 / 2^(index-1))			
+        index <- index + 1
+      }
+      es[i] <- 1/qx
+    }		
+    names(es) <- phy$tip.label
+  }
+  
+  es <- log(es[phy$tip.label]) # log transform
+  trait <- trait[phy$tip.label]
+  
+  # Pearson's correlation between log inverse equal splits statistic and trait
+  res <- cor.test(es, trait, method="pearson")
+  
+  # Fit Brownian motion model to get diffusion rate and root state estimates
+  vv <- vcv.phylo(as.phylo(phy))
+  onev <- matrix(rep(1, length(trait)), nrow=length(trait), ncol=1)
+  root <- as.vector(solve(t(onev)%*% solve(vv) %*% onev) %*% (t(onev)%*% solve(vv) %*% trait))
+  rate <- as.vector((t(trait-root) %*% solve(vv) %*% (trait-root))/length(trait))
+  
+  # Brownian simulations 
+  sims <- t(rmvnorm(nsim, sigma=rate*vv))
+  rownames(sims) <- rownames(vv)
+  
+  # Pearson's correlations of simulated datasets
+  sim.r <- sapply(1:nsim, function(x) cor.test(es[as.vector(rownames(sims))], sims[,x], method="pearson")$estimate)
+  
+  # Calculate the two-tailed p value
+  corr <- res$estimate
+  upper <- (length(sim.r[sim.r >= corr])+1)/(nsim+1)
+  lower <- (length(sim.r[sim.r <= corr])+1)/(nsim+1)
+  pval <- 2*min(c(upper,lower)) # Remove "2" for one-tailed
+  
+  if(missing(return.es)) { # output just rho and p value
+    result <- as.vector(c(corr, pval))
+    names(result) <- c("rho", "P Value")
+    return(result)
[…]

### 2018-06-27 Sexual Selection and Extinction.Rmd
27/6 Update

@@
 library(geiger)
+library(gridExtra)
 library(nlme)
@@
 source("functions/essim.R")
+```
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(cache=TRUE)
 ```
@@
 To gain estimates for change in climate over the past 22,000 we can use the difference in annual mean temperature (BIO1) and annual precipitation (BIO12) between the LGM (CCSM4 models) and present values.
-Additionally we can calculate the difference in variability based on the standard deviation.
+Additionally we can calculate the difference in variability based on the standard deviation, although these will be correlated and thus maybe using only one of them (either mean or variability difference) in the analysis may be better.
 
-```{r}
+```{r, fig.width=7, fig.height=3.5}
 #Change in mean temp
@@
 #Inspect these changes 
-restricted.data %>% ggplot(aes(x=BIO1.LGM.Diff, y = BIO1.LGM.var.Diff))+
-  geom_point()
+bio1.plot <- restricted.data %>% ggplot(aes(x=BIO1.LGM.Diff, y = BIO1.LGM.var.Diff))+
+  geom_point()+
+  theme_minimal()
 
-restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = BIO12.LGM.var.Diff))+
-  geom_point()
+bio12.plot <- restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = BIO12.LGM.var.Diff))+
+  geom_point()+
+  theme_minimal()
+
+grid.arrange(bio1.plot, bio12.plot, ncol=2)
 ```
@@
 
-##Preliminary Analysis of data
+##Preliminary analysis of data
+
+###Examining BAMM rates
 
@@
   scale_x_log10()+
-  scale_y_log10()
+  scale_y_log10()+
+  theme_minimal()
 ```
 
-The extinction rates of the lumped and split datasets show high similarity, insisting that we could use either.
+###Fixed-effect models
 
-Let us create a relatively simple model
+The extinction rates of the lumped and split datasets show high similarity and we could arguably use either. Let's ude 'split':
+
+First with a simple linear model that is not phylogenetically corrected:
+
 ```{r}
@@
 
-Now we can try a PGLS model 
+Secondly, we can use a PGLS model:
+
 ```{r}
@@
 
-Or perhaps OU model rather than Brownian:
+It is likely that given the underestimation of rate shifts by BAMM we obtain groups of similar speciation/extinction rates and thus when used in PGLS the highly correlated species will have near exact extinction/speciation rates.
+Giving less weight to these results.
+
+Additionally, we can run a PGLS with an OU model rather than Brownian:
 
@@
 
+###Bayesian models
+
 Now a Bayesian model that uses a covariance matrix from the phylogeny:
@@
 ```{r, eval=FALSE}
-model_bayesian <- brm(
-  rab.ex.rate.split ~ SDi 
-                + range.size.m2 
-                + bioclim4
-                + bioclim15
-                + residuals.PC1
-                + residuals.PC2
-                + BIO1.LGM.Diff
-                + BIO12.LGM.Diff
-                + NPP.res, data = restricted.data, 
-  family = gaussian(), 
-  cov_ranef = list(phylo = A),
-                seed = 1,
-                cores = 4, chains = 4, iter = 4000, #Run 4 chains in parallel for 4000 iterations (2000 are burn in)
-                control = list(adapt_delta = 0.99, max_treedepth = 15))
+# model_bayesian <- brm(
+#   rab.ex.rate.split ~ SDi 
+#                 + range.size.m2 
+#                 + bioclim4
+#                 + bioclim15
+#                 + residuals.PC1
+#                 + residuals.PC2
+#                 + BIO1.LGM.Diff
+#                 + BIO12.LGM.Diff
+#                 + NPP.res, data = restricted.data, 
+#   family = gaussian(), 
+#   cov_ranef = list(phylo = A),
+#                 seed = 1,
+#                 cores = 4, chains = 4, iter = 4000, #Run 4 chains in parallel for 4000 iterations (2000 are burn in)
+#                 control = list(adapt_delta = 0.99, max_treedepth = 15))
 ```
@@
 ```{r}
-summary(model_bayesian)
+#summary(model_bayesian)
 ```
 
-Try looking at speciation rate using ES-Sim adapted by @Harvey_essim_2017 and found here https://github.com/mgharvey/ES-sim
+###Diversification rate (DR) estimates as opposed to BAMM
 
-```{r}
+Diversification rates can be simply estimated from thew 'tippiness of the phylogeny'.
+Here an equal splits (ES) value can be achieved that is weighted to more recent diversification.
+It was used by @Jetz_2012 and recently (in comparison with BAMM) in a study ivestigation bird diversity alongside elevation [@quintero_2018].WE CAN look at speciation rate using ES-Sim adapted by @Harvey_essim_2017 and found here https://github.com/mgharvey/ES-sim.
+
+```{r, fig.width=8, fig.height = 4}
 Essim.frame <- restricted.data$SDi
@@
 essim.results <- essim(pruned.tree, Essim.frame, nsim = 1000, return.es = TRUE)
+#Independent correlation results for ES-sim 
+essim.stats <- cbind(essim.results[["rho"]], essim.results[["P Value"]])
+colnames(essim.stats) <- c("rho", "P value")
+essim.stats %>% pander(digits = 2)
 ```
 
-We can then add the essim.results and run a similar PGLS model
+These results suggest there is not a correlation between the diversification rate (log(ES)) and sexual dichromatism.
+However.
+We should look at this in a larger model and with phylogenetic correction.
+To do that we can obtain the ES values (tip rates) for each species, the log of these is the diversification rate used by @Jetz_2012 and @quintero_2018.
+
 ```{r}
@@
 
-restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = (ESsim))) + 
-  geom_point()+
-  geom_smooth(method = "lm")
-```
+#OU model using ES-sim
 
- Once we have our EsSim results we can use them in a linear model to assess the relationship between speciation and our predictors
-
-```{r}
-model.2 <- lm(ESsim ~ SDi 
+pgls.model.4 <- gls(ESsim ~ SDi 
                 + range.size.m2 
@@
                 + NPP.res,
-                data = restricted.data)
-summary(model.2)
+                correlation = corMartins(1, phy = pruned.tree), 
+                data = restricted.data, 
+                method = "REML")
+summary(pgls.model.4)
 ```
 
-Have a look at how these would look if we obtained model predictions for each data point.
-
+We can now have a look at how these would look if we obtained model predictions for each data point.
 ```{r, fig.width=10, fig.height=5}
-predict.frame <- cbind(predict(model.2),restricted.data)
-predict.frame %>% ggplot(aes(y = `predict(model.2)`, x = SDi)) + 
+predict.frame <- cbind(predict(pgls.model.4),restricted.data)
+predict.frame %>% ggplot(aes(y = `predict(pgls.model.4)`, x = SDi)) + 
   geom_point(aes(color = BLFamilyLatin))+
@@
 ```
+
+###The sexual dimorphism dataset has an overdispersed distribution
 

### 2018-06-27 Sexual Selection and Extinction.Rmd

@@
 
-It is likely that given the underestimation of rate shifts by BAMM we obtain groups of similar speciation/extinction rates and thus when used in PGLS the highly correlated species will have near exact extinction/speciation rates.
+> My gut feeling: It is likely that given the underestimation of rate shifts by BAMM we obtain groups of similar speciation/extinction rates and thus when used in PGLS the highly correlated species will have near exact extinction/speciation rates.
 Giving less weight to these results.

### 2018-09-14 ComparativePlots.Rmd
Long time no update

---
title: "ComparativePlots"
author: "Justin Cally"
date: "03/09/2018"
output: html_document
---

#Phylo Plots

```{r, fig.width = 6.5, fig.height = 6.5}
pdf("Figures/BAMMplot.spec.pdf", width = 20, height = 20)
BAMMplot.spec <- plot.bammdata(MCC.BAMM.ED, 
                          breaksmethod = "jenks", 
                          spex="s", 
                          method = "polar", 
                          lwd=0.5, 
                          tau=0.003, 
                          logcolor = T, 
                          vtheta = 90, 
                          rbf = 0.02,
            pal= c("#ffffe5","#f7fcb9","#d9f0a3","#addd8e","#78c679","#41ab5d","#238443","#006837","#004529"))
dev.off()
```

Plot phylorate plot, good way of visualizing diversification rates
```{r, fig.width = 6.5, fig.height = 6.5}
pdf("Figures/BAMMplot.spec.tips.pdf", width = 20, height = 20)
BAMMplot.spec.tips <- plot.bammdata(MCC.BAMM.ED, 
                          breaksmethod = "jenks", 
                          spex="s", 
                          method = "polar", 
                          lwd=0.5, 
                          tau=0.003, 
                          logcolor = T, 
                          vtheta = 90, 
                          rbf = 0.02,
                          labels = T,
                          cex = 0.075,
            pal= c("#ffffe5","#f7fcb9","#d9f0a3","#addd8e","#78c679","#41ab5d","#238443","#006837","#004529"))
addBAMMlegend(BAMMplot.spec, location="bottom", direction = "horizontal", side = 1)

dev.off()
```

Get histogram of extinction rates

```{r}
#BAMMplot.spec[["colordens"]][["kde.y"]] <- log1p(BAMMplot.spec[["colordens"]][["kde.y"]] )
#pdf("figures/BAMMplot_hist.pdf", width = 8, height = 4)
#create phylorate plot to generate output
ratesHistogram(BAMMplot.spec, plotBrks = F, xlab = 'Speciation rate', ylab = 'Frequency', lwd = 0.05)
#dev.off()

```

##Plot for DR mean rate

```{r}
cols <- c("#543005", "#8c510a", "#bf812d", "#dfc27d", "#c0d884",
           brewer.pal(n = 9, name = "BuGn")[4:9])
cols <- c('#B76643','#B76643','#B76643','#B76643','#B47339','#AC8032','#A08E33','#909B3C','#7DA64D', brewer.pal(n = 9, name = "Greens")[5:9], "#00441B", "#00441B", "#00441B")

cols <- c(
'#543005',
'#8c510a',
'#fd8d3c',
'#feb24c',
'#fed976',
'#ffeda0',
#'#f7fcb9',
'#d9f0a3',
'#addd8e',
'#78c679',
'#41ab5d',
'#238443',
'#006837',
'#006837',
'#004529',
'#004529',
'#004529')

cols <- c(
'#211F29',
'#262B37',
'#2A3945',
'#2C4751',
'#2D565C',
'#2F6565',
'#34746C',
'#3E8470',
'#4D9371',
'#60A171',
'#78B06F',
'#92BD6D',
'#B0C96B',
'#D0D46B'
)

restricted.data %>% ggplot(aes(x = exp(MCC.DR), y = 0, fill = ..x..))+
  geom_density_ridges_gradient(scale = 3, size = 0.05, gradient_lwd = 0.75)+
  scale_fill_gradientn(colours = cols, guide = F, breaks = c(0.02, 0.1, 0.5, 2.5))+
  theme_minimal()+
  scale_x_continuous(trans = "log", breaks = c(0.02, 0.1, 0.5, 2.5))+
  scale_y_discrete(expand = c(0,0))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.line.y = element_blank(), 
        panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x.bottom = element_line())

MCC.BAMM.model.data %>% ggplot(aes(x = SDi, y = 0, fill = ..x..))+
  geom_density_ridges_gradient(scale = 3, size = 0.05, gradient_lwd = 0.75)+
  scale_fill_distiller(palette = "OrRd", direction = 1, guide = F)+
  theme_minimal()+
  scale_y_discrete(expand = c(0,0))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.line.y = element_blank(), 
        panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x.bottom = element_line())

MCC.BAMM.model.data %>% ggplot(aes(x = range.size.m2, y = 0, fill = ..x..))+
  geom_density_ridges_gradient(scale = 3, size = 0.05, gradient_lwd = 0.75)+
  scale_fill_distiller(palette = "BuPu", direction = 1, guide = F)+
  scale_x_log10()+
  theme_minimal()+
  scale_y_discrete(expand = c(0,0))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.line.y = element_blank(), 
        panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x.bottom = element_line())
```

```{r}
#Get more leasonable intercept for gls slope
SDi.intercept.BAMM <- coef(gls(log(mean.lambda) ~ SDi, data = MCC.BAMM.model.data))[1]

BAMM.scatter <- MCC.BAMM.model.data %>% ggplot(aes(y = mean.lambda, x = SDi, fill = ..x..))+
  # geom_point(data = restricted.data, aes(x = SDi, y = exp(MCC.DR)), color = "gray70")+
  geom_point(shape = 21, stroke = 0.35)+
  scale_fill_distiller(palette = "OrRd", direction = 1, guide = F)+
  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
  theme_bw()+
  xlab("Sexual Dichromatism")+
  ylab("Speciation Rate (BAMM)")+
    geom_abline(slope = MCC.Lambda.top[["coefficients"]][["SDi"]],
             intercept =  SDi.intercept.BAMM, 
             size = 1)+
  geom_abline(slope = min(BAMM.lambda.pgls.summary %>% 
                            filter(Parameter == "SDi") %>% 
                            select(Estimate)),
             intercept =  SDi.intercept.BAMM, 
             size = .75, 
             linetype = 2)+
  geom_abline(slope = max(BAMM.lambda.pgls.summary %>% 
                            filter(Parameter == "SDi") %>% 
                            select(Estimate)),
             intercept =  SDi.intercept.BAMM, 
             size = .75, 
             linetype = 2)+
    scale_x_continuous(expand = c(0,0.3))+
    theme(panel.grid = element_blank(),
          axis.line = element_line(),
          panel.border = element_blank(),
          plot.margin = unit(c(.5,.5,.5,.5), "cm"))

#Get more leasonable intercept for gls slope
SDi.intercept <- coef(gls(MCC.DR ~ SDi, data = restricted.data))[1]

DR.scatter <- restricted.data %>% ggplot(aes(y = exp(MCC.DR), x = SDi, fill = ..x..))+
  geom_point(shape = 21, stroke = 0.35)+
  scale_fill_distiller(palette = "OrRd", direction = 1, guide = F)+
  theme_bw()+
  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
  scale_x_continuous(expand = c(0,0.3))+
  xlab("Sexual Dichromatism")+
  ylab("Speciation Rate")+
  geom_smooth(method = "glm", se = F, color = "black", linetype = 5)+ 
    theme(panel.grid = element_blank(),
          axis.line = element_line(),
          panel.border = element_blank(),
          plot.margin = unit(c(.5,.5,.5,.5), "cm"))

#Get more leasonable intercept for gls slope
range.model <- coef(gls(MCC.DR ~ log(range.size.m2), 
                            data = restricted.data))

newRange <- data.frame(`range.size.m2` = restricted.data$range.size,
                         SDi = mean(restricted.data$SDi),
                         NPP = mean(restricted.data$NPP),
                         bioclim4 = mean(restricted.data$bioclim4),
                         residuals.PC1 = mean(restricted.data$residuals.PC1),
                         PC1.LIG = mean(restricted.data$PC1.LIG))
restricted.data$Range.Predict <- predict(MCC.DR.top, newRange)

Range.scatter <- restricted.data %>% ggplot(aes(y = exp(MCC.DR), x = range.size.m2, fill = ..x..))+
  geom_point(shape = 21, stroke = 0.35)+
  scale_fill_distiller(palette = "BuPu", direction = 1, guide = F, trans = "log")+
  theme_bw()+
  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
  scale_x_continuous(expand = c(0,0.3), trans = "log", breaks = c(1, 1*(10^7), 1*(10^10), 1*(10^13)))+
  xlab("Range Size")+
  ylab("Speciation Rate")+
    geom_smooth(method = "glm", se = F, color = "black", linetype = 5)+
    theme(panel.grid = element_blank(),
          axis.line = element_line(),
          panel.border = element_blank(),
          plot.margin = unit(c(.5,.5,.5,.5), "cm"))

BAMM.scatter
DR.scatter
Range.scatter
```

```{r}
grid.arrange(DR.scatter, Range.scatter, nrow = 2)
```

```{r}
newRange <- data.frame(`range.size.m2` = restricted.data$log.range.size,
                         SDi = mean(restricted.data$SDi),
                         NPP = mean(restricted.data$NPP),
                         bioclim4 = mean(restricted.data$bioclim4),
                         residuals.PC1 = mean(restricted.data$residuals.PC1),
                         PC1.LIG = mean(restricted.data$PC1.LIG))
restricted.data$Range.Predict <- predict(MCC.DR.top, newRange)
```

Plotting a tree showing the estimated BAMM speciation rates

```{r}
#edge.rates <- getMeanBranchLengthTree(MCC.BAMM.ED, rate = "speciation")
#saveRDS(edge.rates, "data/edge.rates.lambda.rds")
edge.rates <- readRDS("data/edge.rates.lambda.rds")
edge.rates.df <- data.frame(edge.rates[["phy"]][["edge"]], edge.rates[["phy"]][["edge.length"]]) %>% `colnames<-`(c("parent", "node", "rate"))

fortified.MCC.tree <- fortify(MCC.BAMM.tree)
fortified.MCC.tree <- left_join(fortified.MCC.tree, edge.rates.df,  by = c("parent", "node"))
```

Creating other plots:

This is the order of species on the big speciation plot going anti-clockwise:

+ Lepidothrix coeruleocapilla
+ Muscisaxicola capistratus 
+ Cercomacra manu
+ Cranioleuca curtata
+ Malurus pulcherrimus
+ Pachycephala pectoralis pectoralis
+ Paradisaea rubra
+ Mayrornis schistaceus
+ Progne cryptoleuca
+ Acrocephalus australis
+ Zosterops erythropleurus
+ Phainopepla nitens
+ Promerops gurneyi
+ Euplectes franciscanus
+ Euphonia rufiventris
+ Atlapetes leucopterus
+ Geospiza magnirostris
+ Sporophila bouvronides

Plot the SDi val
```{r}
# library(RColorBrewer)
# col <- c(scale_color_distiller(palette = "OrRd", direction = 1), scale_color_distiller(palette = "BuPu", direction = 1))
ggtree.plot <- ggtree(fortified.MCC.tree, aes(color =  rate),
                      layout="fan", open.angle = 90, size = 0.175) + 
  xlim(-75, NA)+
  scale_color_distiller(palette = "YlGn", direction = 1, guide = F, trans = "log", breaks=c(0.05, 0.15, 0.5, 1.5))+
  geom_tiplab(size=0.1, color = "black")

    # theme(legend.position = "bottom",
    #     legend.text = element_text(size = 18))

# ggtree.plot <-gheatmap2(ggtree.plot, data = MCC.BAMM.model.data %>% select(SDi), width=0.05, colnames = FALSE, low = "orange", high = "red")

#  scale_fill_distiller(palette = "OrRd", direction = 1, na.value = "white", guide = guide_colorbar(direction = "horizontal", label.position = "bottom", barheight = 1.5, barwidth = 25, label.vjust = -0.5))

  # theme(legend.position = "bottom",
  #       legend.text = element_text(size = 18))

# ggtree.plot <-gheatmap2(ggtree.plot, data = log(MCC.BAMM.model.data %>% select(range.size.m2)), width=0.05, colnames = FALSE, offset = 3.5)+
  # scale_fill_distiller(palette = "BuPu", direction = 1, na.value = "white", guide = guide_colorbar(direction = "horizontal", label.position = "bottom", barheight = 1.5, barwidth = 25, label.vjust = -0.5))+

```

Plot using DR

```{r}
DR.tips <- MCC.dr.df$es %>% `names<-`(MCC.dr.df$binomial)
# tree.rates <- contMap(MCC.BAMM.tree, DR.tips)
tree.rates.node <- fastAnc(MCC.BAMM.tree, DR.tips)
tree.rates.node.df <- as.list(tree.rates.node)
tree.rates.node.df <- bind_rows(tree.rates.node.df) %>% t()
colnames(tree.rates.node.df) <- "es"

DR.tips.df <- MCC.dr.df %>% select(es, binomial) %>% rename(label = binomial)
node.rates <- rbind(DR.tips.df %>% select(es), tree.rates.node.df)

fortified.DR.tree <- fortify(MCC.BAMM.tree)
fortified.DR.tree <- cbind(fortified.DR.tree, node.rates) %>% rename(rate = es)
```

```{r}
cols <- c("#543005", "#8c510a", "#bf812d", "#dfc27d", "#d9ef8b",
           brewer.pal(n = 9, name = "YlGn")[4:9])
display.brewer.pal(n = 11, name = "RdYlGn")[4:11]
#cols <- c("#045a8d", "#74a9cf", "#ffd92f", "#d9f0a3", "#78c679", "#005a32")
#20 x 20 plots
ggtree.plot.dr <- ggtree(fortified.DR.tree, aes(color =  rate),
                      layout="fan", open.angle = 90, size = 0.175) + 
  xlim(-75, NA)+
  scale_colour_gradientn(colours = cols, 
                         #breaks = c(-4, -2, 1.75, 1.2),
                         guide = F) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 18))

# ggtree.plot.dr <- gheatmap2(ggtree.plot.dr, data = MCC.BAMM.model.data %>% select(SDi), width=0.05, colnames = FALSE) + 
#   scale_fill_distiller(palette = "OrRd", direction = 1, na.value = "white", guide = guide_colorbar(direction = "horizontal", label.position = "bottom", barheight = 1.5, barwidth = 25, label.vjust = -0.5))+
#     theme(legend.position = "bottom",
#         legend.text = element_text(size = 18))
```

### 2018-09-14 Sexual Selection and Extinction.Rmd
Long time no update

--- 2018-06-27 Sexual Selection and Extinction.Rmd
@@
 ---
-title: 'Sexual selection and extinction: a macro-evolutionary approach'
+title: 'Sexual selection and speciation: a macro-evolutionary approach'
 author: "Justin G. Cally^1^, Iliana Medina Guzmán^1^ Luke Holman^1^, Devi Stuart-Fox^1^ <br></br> <br></br> ^1^The University of Melbourne"
@@
 bibliography: References/SS_Extinction.bib
-csl: References/Proc_b.csl
+csl: References/AmNat.csl
 output:
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-library(dplyr)
+```{r, message=FALSE, warning=FALSE}
 library(ggplot2)
@@
 library(geiger)
+library(grid)
+library(gridExtra)
 library(nlme)
@@
 library(caper)
+library(purrr)
+library(reshape2)
+library(ggExtra)
+library(grid)
+library(dplyr)
+library(BAMMtools)
+library(gridExtra)
+library(car)
+library(coda)
+library(MuMIn)
+library(parallel)
+library(HDInterval)
+library(phangorn)
+library(kableExtra)
+library(RColorBrewer)
+library(tibble)
+library("EBImage") # for images
+library("ggtree")
 source("bamm_extinction/functions/check_and_fix_ultrametric.R")
 source("functions/essim.R")
+source("functions/diversification_rate_calculation.R")
+source("functions/Vdodge_function.R")
+source('functions/symmetrise_scale.R')
+source('functions/Correlation_matrix.R')
+source('functions/gheatmap2.R') #Gets rid of spaces between tiles
+```
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(cache=TRUE)
 ```
@@
 
-For this project we aimed to compile data on a large set of species in the Passerine bird order that would allow us to account for many evolutionary and environmental predictors of extinction.
+For this project we aimed to compile data on a large set of species in the Passerine bird order that would allow us to account for many evolutionary and environmental predictors of speciation.
 The following details the source and use of the data.
 
-##Extinction Risk
-
-To obtain extinction risk estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM).
-We document the control file and process for obtaining extinction rates from BAMM in an accompanying R markdown in the ``BAMM`` folder.
+##Speciation Rate
+
+To obtain speciation rate estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM).
 Additionally, we compared our BAMM output files (event data) against a pre-existing extinction rate dataset [@Harvey_2017].
 
-Rather than obtaining one estimate of extinction rate, we ran BAMM XXXX times to obtain uncertainty estimates of the extinction risk generated from the phylogeny created by @Jetz_2012.
-
-@Jetz_2012 also provides us with a compiled IUCN threat level listing for all species from 2010 standings.
+Rather than obtaining one estimate of speciation rate from one tree, we ran BAMM 100 times on 100 trees plus an MCC tree to obtain uncertainty estimates of the speciation rate generated from the phylogeny created by @Jetz_2012.
+We also estimated extinction rate from BAMM.
 
@@
 
-Briefly, @Dale_2015 obtained male and female plumage score, which can be divided to obtain a sexual dimorphism index (SDi).
+Briefly, @Dale_2015 obtained male and female plumage score, which can be divided to obtain a sexual dichromatism index (SDi).
 The plumage score was scored using the _Handbook of the Birds of the World_ @delhoyo2011.
 By scanning images of males and females in this book across multiple patches @Dale_2015 were able to obtain mean plumage scores from RGB values.
-Notably concern about the accuracy of these measurements were raised, thus this dataset was previously compared with a more thorough reflectance dataset of Australian birds, the two datasets correlated but was expectedly noisy (R2 = 0.67, P < 0.0001).
-Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008 which has colour discriminability, segment classification and PCA.
+Notably concern about the accuracy of these measurements were raised, thus this dataset was previously compared with a more thorough reflectance dataset of Australian birds, the two datasets correlated but was expectedly noisy.
+Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008 using colour discriminability as the parameter.
 
 Alongside sexual dichromatism data there is also partial data for other metrics analysed in @Dale_2015 such as: Body size, tropical life history, Sexual selection, Cooperative breeding and Migration.
+We use a phylogenetic PCA (PC1) that estimates male-bias sexual selection from a species level of sexual size dimorphism, social polygyny and paternal care.
 
@@
 
-We obtained species range distribution maps from Birdlife International (through IUCN). These species range maps cover ~ 7,000 species of birds. From these ranges we were able to obtain estimates of:
+We obtained species range distribution maps from Birdlife International (through IUCN) [@birdlife_2017].
+These species range maps cover nearly all species of birds.
+From these spatial information files we were able to obtain estimates of the following (note that not all the data extracted was subsequently used in the analysis):
 
@@
 * Average and standard deviation in 19 bioclimatic variables (each range randomly sampled 1000 times)
-* Average and standard deviation in 19 bioclimatic variables from the last glaciam maximum (LGM)
+* Average and standard deviation in 19 bioclimatic variables from the last-glacial maximum (LGM) and the last-inter-glacial (LIG)
 * Net primary productivity (NPP) estimates and variability
-* Average and standard deviation of population density in the species ranges.
-
-The code for extracting the environmental data from the species is provided in another document XXXX.
-
-#Generating biologically relevent predictors of extinction
+* Average and standard deviation of human population density in the species ranges.
+
+The code for extracting the environmental data from the species is provided in another R-markdown document **Github link**, however no raw data is provided alongside this file due to the large file size of the shapefile and raster information.
+We have put the extracted environmental variables in a csv filed called ``complete.dataframe.csv`` for convenience.
+Briefly for each of the ~ 6,000 species, we:
+
+**1)** Randomly sampled the range polygon 1000 times:
+
+```{r, eval = FALSE}
+#Increse the iterations (defaukt is 4) so we can obtain complete samples of each range
+bird.points <- lapply(bird.ranges, 
+                       function(x) {spsample(x, n=1000, type="random", iter = 30)})
+```
+
+**2)** Extracted the bioclim or other (e.g. NPP) data from each of the points:
+```{r, eval = FALSE}
+bird.values <- lapply(bird.points, function(x) {raster::extract(bioclim_data, x)})
+bird.values <- as.list(data.frame((bird.values)))
+```
+
+**3)** Summarised the extracted data across the 1,000 points into a summary value of interest (means and se) for each species and exported that data.
+
+```{r, eval = FALSE}
+#Obtain means
+bird.frame <- lapply(bird.values, function(x) {as.data.frame(x)})
+bird.summary <- lapply(bird.frame, function(x) {
+(as.data.frame(apply(x,2,mean, na.rm =T )))})
+
+#transpose
+bird.means <- t(as.data.frame(bird.summary))
+bird.means <- (split(bird.means, 1:19))
+
+#Now add column of species name: Same order carried through
+bird.means <- cbind.data.frame(shps.jetz, bird.means)
+
+rownames(bird.means) <- NULL
+colnames(bird.means) <- c("binomial","bioclim1", "bioclim2", "bioclim3", "bioclim4", "bioclim5", "bioclim6", "bioclim7", "bioclim8", "bioclim9", "bioclim10", "bioclim11", "bioclim12", "bioclim13", "bioclim14", "bioclim15", "bioclim16", "bioclim17", "bioclim18", "bioclim19")
+
+#Write csv
+write.csv(bird.means, 'data/bird.means.csv')
+```
+
+**4)** This process was repeated when we used gridded data for the LGM, LIG, NPP and human population density, while range size was obtained from the following code:
+
+```{r, eval = FALSE}
+bird.range.size <- sapply(bird.ranges,
+                       function(x) {(area(x))})
+bird.range.size <- sapply(bird.range.size, function(x) {sum(x)})
+bird.range.size <- as.data.frame(bird.range.size)
+bird.range.size <- cbind.data.frame(shps.jetz, bird.range.size)
+rownames(bird.range.size) <- NULL
+colnames(bird.range.size) <- c("binomial", "range.size.m2")
+
+#write.csv
+write.csv(bird.range.size, 'data/bird.range.size.csv')
+```
+
+#Generating biologically relevent predictors
 
@@
 
-Extinction rate ~ Sexual selection
+ Extinction/Diversification ~ Sexual selection  
                             + Range size  
                             + Short temporal variability of temperature (mean BIOCLIM4)  
-                   + Short temporal variability of precipitation (mean BIOCLIM15)
-                   + Spatial variability of temperature (PCA1)
-                   + Spatial variability of precipitation (PCA2)
-                   + Long-term variability of temperature (BIOCLIM1)
-                   + Long-term variability of precipitation (BIOCLIM12)
-                   + Variability of NPP
-
-These predictors can be generated from the compiled datasets seen here:
-
-```{r}
+                            + Spatial variability of temperature (PCA1) [residual.PCA1]  
+                            + Long-term variability of temperature (LIG)  
+                            + NPP
+
+These predictors can be obtained from the compiled datasets seen here:
+
+```{r, warning=FALSE}
+plumage.scores <- read.csv('data/plumage_scores.csv')
+#Generate sexual dichromatism score: 
+plumage.scores$SDi <- abs(plumage.scores$Male_plumage_score - plumage.scores$Female_plumage_score)
+
+#Make DF with enviro variables and plumage scores
 complete.dataframe <- read.csv('data/complete.dataframe.csv')
+complete.dataframe <- left_join(plumage.scores %>% dplyr::select(binomial, Male_plumage_score, Female_plumage_score, SDi), complete.dataframe, by = "binomial")
 ```
@@
 
+**Table S1:** Loadings for the first two PCs, when we conduct a PCA on the variation (standard error, where n ~ 1,000) pf each bioclim variable except the standard error of temperature seasonality (se.bioclim4) and the standard error of precipitation seasonality (se.bioclim15), as these are already measures of variability and based on other bioclim variables.
+We use the standard error as, while we attempted to sample each species range 1,000 times, some points returned NA values and would thus affect the standard deviation, this inconvenience should not affect standard error.
 ```{r}
 restricted.data <- complete.dataframe %>% drop_na(bioclim1) #Drop species without environmental data
-PCA.bioclim <- prcomp(restricted.data[c(34,35,36,38:47,49:52)], #Select se.bioclim
+PCA.bioclim <- prcomp(restricted.data[c('se.bioclim1', 'se.bioclim2', 'se.bioclim3', 'se.bioclim5', 'se.bioclim6', 'se.bioclim7', 'se.bioclim8', 'se.bioclim9', 'se.bioclim10', 'se.bioclim11', 'se.bioclim12', 'se.bioclim13', 'se.bioclim14', 'se.bioclim16', 'se.bioclim17', 'se.bioclim18', 'se.bioclim19')], #Select se.bioclim
                       scale = TRUE, center = TRUE)
@@
 
+#Create table with PC1 and PC2
+as.data.frame(PCA.bioclim$rotation[,1:2]) %>% `colnames<-`(c("PC1", "PC2")) %>% pander()
+```
+
+```{r, fig.width=10, fig.height=5}
 #run GAM
-PC1.gam <- mgcv::gam(PC1 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")
+PC1.gam <- mgcv::gam(PC1*(-1) ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian") #Inverse as the PC1 loads to the negative (COUNTER-NTUITIVE)
 
@@
 
-#plot
-plot.gam(PC1.gam, residuals = T)
-```
-
-We can do the same for PC2 (precipitation)
-
-```{r}
+#We can do the same for PC2
 PC2.gam <- mgcv::gam(PC2 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")
@@
 
-plot.gam(PC2.gam, residuals = T)
-```
-
-##Long term climate variability (LGM anomaly)
-
-To gain estimates for change in climate over the past 22,000 we can use the difference in annual mean temperature (BIO1) and annual precipitation (BIO12) between the LGM (CCSM4 models) and present values.
-Additionally we can calculate the difference in variability based on the standard deviation.
+#plot
+par(mfrow = c(1,2))
+plot.gam(PC1.gam, residuals = T, main = 'PC1 residuls (temp)')
+plot.gam(PC2.gam, residuals = T, main = 'PC2 residuals (precip)')
+```
+
+**Figure S1:** The relationship between spatial variability in the temperature components (PC1) and log-range size is relatively strong.
+But not as strong for spatial variability in precipitation (PC2).
+In the analysis we only used the residuals from PC1
+
+##Long term climate variability (LIG anomaly)
+
+To gain estimates for change in climate over the past ~130,000 years we can use the difference in bioclim variables between the LIG and present values.
+The plots show the two PCs, broadly representing temperature and prescipitation.
+Data has also been provided for the last-glacial maximum (LGM) but was not used in the analysis.
+
+**Table S2:** Loadings for the first two PCs of each of the PCAs for the difference in bioclimatic variables between today and the LIG.
+Here, PC1 is more heavily loaded for absoluted temperature difference, while PC2 is more heavily loaded for absolute difference in precipitation.
 
 ```{r}
-#Change in mean temp
-restricted.data$BIO1.LGM.Diff <- restricted.data$bioclim1 - restricted.data$cclgmbi1
-#Change in variation temp
-restricted.data$BIO1.LGM.var.Diff <- restricted.data$se.bioclim1 - restricted.data$se.cclgmbi1
-
-#Change in mean precipitation
-restricted.data$BIO12.LGM.Diff <- restricted.data$bioclim12 - restricted.data$cclgmbi12
-#Change in variation precipitation
-restricted.data$BIO12.LGM.var.Diff <- restricted.data$se.bioclim12 - restricted.data$se.cclgmbi12
-
-#Inspect these changes 
-restricted.data %>% ggplot(aes(x=BIO1.LGM.Diff, y = BIO1.LGM.var.Diff))+
-  geom_point()
-
-restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = BIO12.LGM.var.Diff))+
-  geom_point()
-```
-
-From these plots we can see that on average temperature has increased but the variation in tempere has decresed (narrowed).
-Alternatively, precipitation sometimes increases and sometimes decreases.
-When precipitation increases there isw also an increase in variation in that range.
-However, when annual precipitation decreases there is a decrease in variation.
-
-##Variation in NPP
-
-Similar to the PCA's we expect NPP variation to be correlated with range size. To account for this we can run a GAM and obtain residuals.
+historical.variation.data <- as.data.frame(restricted.data[1])
+
+#FOR LIG
+
+historical.variation.data$bio1.LIG.diff <- abs(restricted.data$bioclim1 - restricted.data$LIG.bi1)
+historical.variation.data$bio2.LIG.diff <- abs(restricted.data$bioclim2 - restricted.data$LIG.bi2)
+historical.variation.data$bio3.LIG.diff <- abs(restricted.data$bioclim3 - restricted.data$LIG.bi3)
+historical.variation.data$bio4.LIG.diff <- abs(restricted.data$bioclim4 - restricted.data$LIG.bi4)
+historical.variation.data$bio5.LIG.diff <- abs(restricted.data$bioclim5 - restricted.data$LIG.bi5)
+historical.variation.data$bio6.LIG.diff <- abs(restricted.data$bioclim6 - restricted.data$LIG.bi6)
+historical.variation.data$bio7.LIG.diff <- abs(restricted.data$bioclim7 - restricted.data$LIG.bi7)
+historical.variation.data$bio8.LIG.diff <- abs(restricted.data$bioclim8 - restricted.data$LIG.bi8)
+historical.variation.data$bio9.LIG.diff <- abs(restricted.data$bioclim9 - restricted.data$LIG.bi9)
+historical.variation.data$bio10.LIG.diff <- abs(restricted.data$bioclim10 - restricted.data$LIG.bi10)
+historical.variation.data$bio11.LIG.diff <- abs(restricted.data$bioclim11 - restricted.data$LIG.bi11)
+historical.variation.data$bio12.LIG.diff <- abs(restricted.data$bioclim12 - restricted.data$LIG.bi12)
+historical.variation.data$bio13.LIG.diff <- abs(restricted.data$bioclim13 - restricted.data$LIG.bi13)
+historical.variation.data$bio14.LIG.diff <- abs(restricted.data$bioclim14 - restricted.data$LIG.bi14)
+historical.variation.data$bio15.LIG.diff <- abs(restricted.data$bioclim15 - restricted.data$LIG.bi15)
+historical.variation.data$bio16.LIG.diff <- abs(restricted.data$bioclim16 - restricted.data$LIG.bi16)
+historical.variation.data$bio17.LIG.diff <- abs(restricted.data$bioclim17 - restricted.data$LIG.bi17)
+historical.variation.data$bio18.LIG.diff <- abs(restricted.data$bioclim18 - restricted.data$LIG.bi18)
+historical.variation.data$bio19.LIG.diff <- abs(restricted.data$bioclim19 - restricted.data$LIG.bi19)
+
+historical.variation.data <- historical.variation.data %>% drop_na(bio1.LIG.diff)
+
+#Run a PCA of the difference removing the variation variables (4 and 15)
+
+#For LIG
+PCA.LIG.bioclim <- prcomp(historical.variation.data[c(
+'bio1.LIG.diff',
+'bio2.LIG.diff',
+'bio3.LIG.diff',
+'bio5.LIG.diff',
+'bio6.LIG.diff',
+'bio7.LIG.diff',
+'bio8.LIG.diff',
+'bio9.LIG.diff',
+'bio10.LIG.diff',
+'bio11.LIG.diff',
+'bio12.LIG.diff',
+'bio13.LIG.diff',
+'bio14.LIG.diff',
+'bio16.LIG.diff',
+'bio17.LIG.diff',
+'bio18.LIG.diff',
+'bio19.LIG.diff')],
+                      scale = TRUE, center = TRUE)
+
+#Create table with PC1 and PC2
+as.data.frame(PCA.LIG.bioclim$rotation[,1:2]) %>% `colnames<-`(c("PC1.LIG", "PC2.LIG")) %>% pander()
+```
+
+```{r, fig.width=8, fig.height=5}
+#Now we can predict the PCA results: 
+PCA.LIG.predictions <- as.data.frame(predict(PCA.LIG.bioclim)*(-1)) #So that higher numbers mean more variation
+PCA.LIG.predictions <- rename(PCA.LIG.predictions, PC1.LIG = PC1,
+       PC2.LIG = PC2)
+
+#Bind them to dataframe, taking only the first two PCAs
+historical.variation.data <- cbind(historical.variation.data, PCA.LIG.predictions[1:2])
+
+#Now to the restricted dataframe
+restricted.data <- right_join(restricted.data, historical.variation.data %>% dplyr::select(binomial, PC1.LIG, PC2.LIG), by = 'binomial')
+
+PCA.plot <- historical.variation.data %>% ggplot(aes(x = PC1.LIG, y= PC2.LIG))+
+  geom_point(shape = 21)+
+  theme_minimal()
+
+PCA.plot.m <- ggExtra::ggMarginal(
+  p = PCA.plot,
+  type = 'density',
+  margins = 'both',
+  size = 5,
+  colour = 'black',
+  fill = 'gray'
+)
+
+grid.arrange(PCA.plot.m)
+```
+
+**Figure S2:** The distribution of the two PCA compontents for the absolute difference in bioclimatic variables between today and the LIG.
+
+##Correlations between environmental predictors
+
+We can check the correlations between environmental predictors that we will use in our model. Specifically we want to know whether the following environmental variables are correlated:
+
+ + Range size  
+ + Short temporal variability of temperature (mean BIOCLIM4)  
+ + Spatial variability of temperature (PCA1) [residual.PCA1]  
+ + Long-term variability of temperature (LIG)  
+ + NPP
+
+To check some of the correlations between the environmental predictors to be used in the PGLS models we can inspect the following correlation plot with the correlation value plotted:
+
+```{r, fig.width=10, fig.height=10}
+#Draw plot
+restricted.data$log.range.size <- log(restricted.data$range.size.m2)
+pairs(restricted.data[,c("log.range.size", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP")], lower.panel=panel.smooth, upper.panel=panel.cor)
+```
+
+**Figure S3:** The highest correlation is between Long term temperature variation and seasonal variation.
+However this correlation is not excessively high (0.53) so we can be satisfied in including these predictors within our model.
+
+##Sexual Dimorphism
+
+The sexual dimorphism dataset looks to have an overdispersed distribution (see phylogenic plot in manuscript).
+Unfortunately, transformations do not greatly improve the distribution.
+Our model fit is expected to be reduced by this but further transformations are not obvious and would risk problems of interperetation.
+
+_______________
+
+#Analysis
+
+##Simple speciation measures: Diversification rate (DR) and Node Density (ND)
+
+Diversification rates can be simply estimated from thew 'tippiness of the phylogeny'.
+Here a log equal splits (also called DR) value can be achieved that is weighted to more recent diversification.
+It was used by @Jetz_2012 and recently (in comparison with BAMM) in a study ivestigation bird diversity alongside elevation [@quintero_2018].
+We can look at speciation rate using ES-Sim adapted by @Harvey_essim_2017 and found here https://github.com/mgharvey/ES-sim.
+
+The functions that generate these tip rates estimates are: ``calculate.log.es``, ``calculate.nd`` and ``calculate.tb``.
+Here they are run on 100 trees plus a full MCC tree of passerine birds based on 2500 samples of the posterior [@Jetz_2012].
+This was done using the ``phanghorn`` [@phangorn_2010].
 
 ```{r}
-#run gam
-NPP.gam <- mgcv::gam(NPP.se ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")
-
-#obtain residuals
-restricted.data$NPP.res <- residuals.gam(NPP.gam)
-
-#plot
-plot(NPP.gam, residuals = T)
-```
-
-_______________
-
-##Preliminary Analysis of data
-
-Now that we have all our relevent predictors we can combine these with the tip extinction rates from BAMM (code for this is provided in the repository).
-First, let us inspect the tip rates that @Harvey_2017 obtained in BAMM with our predictors.
+#Obtain DR (log(es)) estimates, by calling the calculate.dr function
+passerine.trees <- read.nexus('bamm_extinction/data/trees/Passerine_Trees_Full.nex')
+
+#DR, ND and TB can be calculated on the 100 trees by:
+
+# es.list <- sapply(passerine.trees, calculate.log.es)
+# nd.list <- sapply(passerine.trees, calculate.nd)
+# tb.list <- sapply(passerine.trees, calculate.tb)
+
+#To avoid re-running the time-consuming rates, we can save them and load them: 
+es.list <- readRDS("data/es.list.rds")
+nd.list <- readRDS("data/nd.list.rds")
+tb.list <- readRDS("data/tb.list.rds")
+
+#Also load in our MCC tree
+MCC.passerine <- read.tree('data/MCC.passerine.tre') #Tree based on 2500 samples of posterior
+```
+
+To gain an estimate of variation between trees we can take the values and obtain a mean, variance and coeeficient of variation (CV).
+We obtain a CV value to account for unequal variation at higher estimates.
+The CV is expressed as the ratio of the standard deviation ($\sigma$) to the mean ($\mu$) (converting to a percentage not needed):
+
+$C_V = \frac{\sigma}{\mu} \times 100$
+
+Furthermore, we can account for overdispersion by obtaining a transformed version of CV. In such a case we can calculate CV for a log-normal distribution:
+
+${\widehat {c_{\rm {v}}}}_{\rm {ln}}={\sqrt {\mathrm {e} ^{s_{\rm {ln}}^{2}}-1}}$
+
+Given that our _ES_ value is log-transformed, the calculation of ${\widehat {c_{\rm {v}}}}_{\rm {ln}}$ is perhaps more ideal.
+As for _ND_ we can see from the a ``qqnorm`` plot that value of ND is normally distributed, therefore the regular CV is more appropriate for this distribution.
+
+```{r, eval = FALSE}
+#For ND
+#transpose the list so that the elements are the species
+t.nd.list <- nd.list %>% purrr::transpose()
+#not too sure why this works but turn the nested list into a dataframe
+nd.values <- as.data.frame(t(sapply(t.nd.list, function(x) sapply(x, function(x) (x)))))
+nd.values <- nd.values %>% tibble::rownames_to_column("binomial")
+nd.summary <- melt(nd.values, id.vars = "binomial") %>% group_by(binomial) %>% summarise(
+  mean.nd = mean(value), #use normal mean, not a rate
+  var.nd = var(value), #Don't really need this if we use CV
+  CV.nd = (sd(value)/mean(value)) #Given normal distribution of ND, standard CV formula is appropriate
+)
+
+#For ES
+#transpose the list so that the elements are the species
+t.es.list <- es.list %>% purrr::transpose()
+#not too sure why this works but turn the nested list into a dataframe
+es.values <- as.data.frame(t(sapply(t.es.list, function(x) sapply(x, function(x) (x)))))
+es.values <- es.values %>% tibble::rownames_to_column("binomial")
+es.summary <- melt(es.values, id.vars = "binomial") %>% group_by(binomial) %>% summarise(
+  mean.loges = log(1/mean(1/exp(value))), #use harmonic mean for rates and log-transform post calculation of HM
+  var.loges = var(value), #Don't really need this if we use CV
+  CV.loges = sqrt(exp(var(value)) - 1) #CV for log-normal distribution of ES
+)
+
+#For TB
+#transpose the list so that the elements are the species
+t.tb.list <- tb.list %>% purrr::transpose()
+#not too sure why this works but turn the nested list into a dataframe
+tb.values <- as.data.frame(t(sapply(t.tb.list, function(x) sapply(x, function(x) (x)))))
+tb.values <- tb.values %>% tibble::rownames_to_column("binomial")
+tb.summary <- melt(tb.values, id.vars = "binomial") %>% group_by(binomial) %>% summarise(
+  mean.tb = mean(value), #use normal mean; not a rate
+  var.tb = var(value), #Don't really need this if we use CV
+  CV.tb = (sd(value)/mean(value)) #Given normal distribution of tb, standard CV formula is appropriate
+)
+
+MCC.nd.df <- as.data.frame(calculate.nd(MCC.passerine)) %>% tibble::rownames_to_column("binomial")
+MCC.dr.df <- as.data.frame(calculate.log.es(MCC.passerine)) %>% tibble::rownames_to_column("binomial")
+
+#Join the dataframes
+dr.summary <- full_join(nd.summary, es.summary, by = "binomial")
+dr.summary <- full_join(dr.summary, tb.summary, by = "binomial")
+dr.summary <- full_join(dr.summary, MCC.nd.df, by = "binomial")
+dr.summary <- full_join(dr.summary, MCC.dr.df, by = "binomial")
+
+dr.summary <- dr.summary %>% rename(TipLabel = binomial, MCC.ND = nd, MCC.DR = es)
+
+saveRDS(dr.summary, 'data/dr.summary.rds')
+```
 
 ```{r}
-complete.dataframe %>% ggplot(aes(x=rab.ex.rate.lumped, y=rab.ex.rate.split))+
+dr.summary <- readRDS('data/dr.summary.rds')
+#Plot a summary of logES and TB with the DRs with their weightings
+dr.plot <- dr.summary %>% ggplot(aes(x = mean.loges, y = mean.nd, 
+                          fill = (1/(CV.nd))/sum((1/(CV.nd)))*100, 
+                      size = (1/(CV.loges))/sum((1/(CV.loges)))*100))+
+  geom_point(shape = 21, alpha = 0.5)+
+  theme_minimal()+
+  geom_errorbarh(aes(xmin = mean.loges - 0.674*sqrt(var.loges), xmax = mean.loges + 0.674*sqrt(var.loges)), 
+                 size = 0.015)+
+  geom_errorbar(aes(ymin = mean.nd - 0.674*sqrt(var.nd), ymax = mean.nd + 0.674*sqrt(var.nd)), 
+                size = 0.015)+
+  theme(legend.position="bottom")+
+  labs(size = 'Weight (%) [using DR]', y='Node Density [ND]', x= 'Log Diversification Rate [DR]', fill = 'Weight (%) [using ND] \n')+
+  scale_fill_distiller(guide = "colorbar", palette = "RdPu", direction = 1, breaks=c(0,0.015,.03),
+                           limits=c(0,0.03))
+marginal.dr <-ggMarginal(
+  p = dr.plot,
+  type = 'density',
+  margins = 'both',
+  size = 5,
+  colour = '#0000009C',
+  fill = '#D12E6769'
+)
+```
+
+```{r, fig.width=8, fig.height=6}
+grid.newpage()
+grid.draw(marginal.dr)
+```
+**Figure S4:** $\lambda_{DR}$ and $\lambda_{ND}$ are correlated in their estimation of species rate.
+Furthermore, we see similar variation across each species in the estimates across 100 trees as the inverse of the coefficeint of variation (weight) is approximately equivalent across all values of ND and DR (colours and size).
+Notably, these estimates speciation rate are relatively uncertain across all species, with 50 % CIs in both axes plotted.
+
+Combine the estimates of DR and ND with the restricted dataframe containing all the other variables. We can inspect the trend in SDi in relation to the speciation rates.
+
+```{r, warning=FALSE, fig.width=8, fig.height=10}
+restricted.data <- left_join(restricted.data, dr.summary, by = 'TipLabel')
+
+es.plot <- restricted.data %>% ggplot(aes(y = mean.loges, x = SDi, fill = mean.loges))+
+geom_point(shape = 21, size = 1.5)+
+geom_smooth(show.legend = FALSE, color = "grey20", method = "lm")+
+scale_fill_distiller(palette = "YlOrBr", direction = 1, guide = FALSE)+
+ylab("mean.DR")+
+theme_minimal()
+
+nd.plot <- restricted.data %>% ggplot(aes(y = mean.nd, x = SDi, fill = mean.nd))+
+geom_point(shape = 21, size = 1.5)+
+geom_smooth(show.legend = FALSE, color = "grey20", method = "lm")+
+scale_fill_distiller(palette = "Greens", direction = 1, guide = FALSE)+
+theme_minimal()
+
+tb.plot <- restricted.data %>% ggplot(aes(y = mean.tb, x = SDi, fill = mean.tb))+
+geom_point(shape = 21, size = 1.5)+
+geom_smooth(show.legend = FALSE, color = "grey20", method = "lm")+
+scale_fill_distiller(palette = "PuBu", direction = 1, guide = FALSE)+
+theme_minimal()
+
+grid.arrange(es.plot, nd.plot, tb.plot, nrow = 3)
+```
+**Figure S5:** Scatter plots showing the raw relationship between sexual dichromatism (SDi) and speciation rates.
+Across all three measures the pattern and spread is similar, with no obvious relationship.
+Note that terminal branch lenght (mean.tb) was not used in the analysis.
+
+##BAMM measures of speciation and extinction
+
+###Set up BAMM parameters
+
+For the use of BAMM we used the following code to generate the parameters across the 100 trees.
+Each paramaeter value is specified in this code chunk to ensure reproducibility when re-running on any of the trees.
+These same parameters were also used for the MCC tree (with the seed set at 2500).
+
+```{r, eval = FALSE, results = 'hide'}
+name.passerine.tree <- names(passerine.trees)
+
+priors <- sapply(name.passerine.tree, function(x) {
+  setBAMMpriors(passerine.trees[[x]], outfile = NULL)
+})
+
+sapply(name.passerine.tree, function(x) {
+  write.tree(passerine.trees[[x]], paste("data/bamm_files/", x, ".tre", sep=""))
+})
+
+# Here is a block of parameters for the control file. We can make a control file for each tree:
+params <- list()
+for (x in name.passerine.tree) {
+
+# GENERAL SETUP AND DATA INPUT
+
+params[[x]] <- list(modeltype = 'speciationextinction',
+# Specify "speciationextinction" or "trait" analysis
+
+treefile = paste(x, ".tre", sep=""),
+# File name of the phylogenetic tree to be analyzed
+
+runInfoFilename = 'run_info.txt',
+# File name to output general information about this run
+
+sampleFromPriorOnly = 0,
+# Whether to perform analysis sampling from prior only (no likelihoods computed)
+
+runMCMC = 1,
+# Whether to perform the MCMC simulation. If runMCMC = 0, the program will only
+# check whether the data file can be read and the initial likelihood computed
+
+loadEventData = 0,                       
+# Whether to load a previous event data file
+
+eventDataInfile = 'event_data_in.txt',
+# File name of the event data file to load, used only if loadEventData = 1
+
+initializeModel = 1,
+# Whether to initialize (but not run) the MCMC. If initializeModel = 0, the
+# program will only ensure that the data files (e.g., treefile) can be read
+
+useGlobalSamplingProbability = 1,
+# Whether to use a "global" sampling probability. If False (0), expects a file
+# name for species-specific sampling probabilities (see sampleProbsFilename)
+
+globalSamplingFraction = 1,
+# The sampling probability. If useGlobalSamplingProbability = 0, this is ignored
+# and BAMM looks for a file name with species-specific sampling fractions
+
+sampleProbsFilename = 'sample_probs.txt',
+# File name containing species-specific sampling fractions
+
+seed = as.numeric(gsub("tree_", "", x, perl = TRUE)),
+# Seed for the random number generator. Set for reproducibility to the number of the treefile
+
+overwrite = 1,
+# If True (1), the program will overwrite any output files in the current
+# directory (if present)
+
+# PRIORS
+
+expectedNumberOfShifts = 100,
+# prior on the number of shifts in diversification
+# Suggested values: 
+#     expectedNumberOfShifts = 1.0 for small trees (< 500 tips)
+#  expectedNumberOfShifts = 10 or even 50 for large trees (> 5000 tips)
+
+lambdaInitPrior = as.numeric(priors['lambdaInitPrior', x]),
+# Prior (rate parameter of exponential) on the initial lambda value for rate
+# regimes
+
+lambdaShiftPrior = 0.05,
+# Prior (std dev of normal) on lambda shift parameter for rate regimes
+# You cannot adjust the mean of this distribution (fixed at zero, which is
+# equal to a constant rate diversification process)
+
+muInitPrior = as.numeric(priors['muInitPrior', x]),
+# Prior (rate parameter of exponential) on extinction rates
+
+lambdaIsTimeVariablePrior = 1,
+# Prior (probability) of the time mode being time-variable (vs. time-constant)
+
+# MCMC SIMULATION SETTINGS & OUTPUT OPTIONS
+
+numberOfGenerations = '100000000',
+# Number of generations to perform MCMC simulation
+
+mcmcOutfile = 'mcmc_out.txt',
+# File name for the MCMC output, which only includes summary information about
+# MCMC simulation (e.g., log-likelihoods, log-prior, number of processes)
+
+mcmcWriteFreq = 1000,
+# Frequency in which to write the MCMC output to a file
+
+eventDataOutfile = 'event_data.txt',
+# The raw event data (these are the main results). ALL of the results are
+# contained in this file, and all branch-specific speciation rates, shift
+# positions, marginal distributions etc can be reconstructed from this output.
+# See R package BAMMtools for working with this output
+
+eventDataWriteFreq = 1000,
+# Frequency in which to write the event data to a file
+
+printFreq = 10000,
+# Frequency in which to print MCMC status to the screen
+
+acceptanceResetFreq = 1000,
+# Frequency in which to reset the acceptance rate calculation
+# The acceptance rate is output to both the MCMC data file and the screen
+
+outName = x,
+# Optional name that will be prefixed on all output files (separated with "_")
+# If commented out, no prefix will be used
+
+# OPERATORS: MCMC SCALING OPERATORS
+
+updateLambdaInitScale = 2,
+# Scale parameter for updating the initial speciation rate for each process
+
+updateLambdaShiftScale = 0.1,
+# Scale parameter for the exponential change parameter for speciation
+
+updateMuInitScale = 2,
+# Scale parameter for updating initial extinction rate for each process
+
+updateEventLocationScale = 0.1,
+# Scale parameter for updating LOCAL moves of events on the tree
+# This defines the width of the sliding window proposal
+
+updateEventRateScale = 4,
+# Scale parameter (proportional shrinking/expanding) for updating
+# the rate parameter of the Poisson process
+
+# OPERATORS: MCMC MOVE FREQUENCIES
+
+updateRateEventNumber = 1,
+# Relative frequency of MCMC moves that change the number of events
+
+updateRateEventPosition = 0.25,
+# Relative frequency of MCMC moves that change the location of an event on the
+# tree
+
+updateRateEventRate = 1,
+# Relative frequency of MCMC moves that change the rate at which events occur
+
+updateRateLambda0 = 1,
+# Relative frequency of MCMC moves that change the initial speciation rate
+# associated with an event
+
+updateRateLambdaShift = 1,
+# Relative frequency of MCMC moves that change the exponential shift parameter
+# of the speciation rate associated with an event
+
+updateRateMu0 = 1,
+# Relative frequency of MCMC moves that change the extinction rate for a given
+# event
+
+updateRateLambdaTimeMode = 0,
+# Relative frequency of MCMC moves that flip the time mode
+# (time-constant <=> time-variable)
+
+localGlobalMoveRatio = 10,
+# Ratio of local to global moves of events
+
+# INITIAL PARAMETER VALUES
+
+lambdaInit0 = 0.032,
+# Initial speciation rate (at the root of the tree)
+
+lambdaShift0 = 0,
+# Initial shift parameter for the root process
+
+muInit0 = 0.005,
+# Initial value of extinction (at the root)
+
+initialNumberEvents = 0,
+# Initial number of non-root processes
+
+# METROPOLIS COUPLED MCMC
+
+numberOfChains = 1,
+# Number of Markov chains to run
+
+deltaT = 0.01,
+# Temperature increment parameter. This value should be > 0
+# The temperature for the i-th chain is computed as 1 / [1 + deltaT * (i - 1)]
+
+swapPeriod = 1000,
+# Number of generations in which to propose a chain swap
+
+chainSwapFileName = 'chain_swap.txt',
+# File name in which to output data about each chain swap proposal.
+# The format of each line is [generation],[rank_1],[rank_2],[swap_accepted]
+# where [generation] is the generation in which the swap proposal was made,
+# [rank_1] and [rank_2] are the chains that were chosen, and [swap_accepted] is
+# whether the swap was made. The cold chain has a rank of 1.
+
+# NUMERICAL AND OTHER PARAMETERS
+
+minCladeSizeForShift = 3,
+# Allows you to constrain location of possible rate-change events to occur
+# only on branches with at least this many descendant tips. A value of 1
+# allows shifts to occur on all branches.
+
+segLength = 0.025,
+# Controls the "grain" of the likelihood calculations. Approximates the
+# continuous-time change in diversification rates by breaking each branch into
+# a constant-rate diversification segments, with each segment given a length
+# determined by segLength. segLength is in units of the root-to-tip distance of
+# the tree. So, if the segLength parameter is 0.01, and the crown age of your
+# tree is 50, the "step size" of the constant rate approximation will be 0.5.
+# If the value is greater than the branch length (e.g., you have a branch of
+# length < 0.5 in the preceding example) BAMM will not break the branch into
+# segments but use the mean rate across the entire branch.
+
+outName = x)
+  }
+
+bammcontrolfile <- list()
+for (x in name.passerine.tree) {
+  bammcontrolfile[x] <- paste("data/bamm_files/control_", x, ".txt", sep="")
+}
+
+# Now writing control parameters to file
+
+for (x in name.passerine.tree) {generateControlFile(file = bammcontrolfile[[x]], type = "diversification", params = params[[x]])}
+```
+
+###Run analysis
+
+BAMM can be run through the terminal through the following syntax: ``bamm -c control_tree_xxxx.txt``. To generate these commands we can use a loop function, from which we get:
+
+```{r, eval = FALSE}
+bamm.commands <- list()
+for (x in name.passerine.tree) {
+  bamm.commands[x] <- paste("bamm -c control_", x, ".txt", sep="")
+}
+```
+
+The analysis was run over multiple CPU's, each generating a respective MCMC and EventData output.
+Due to the size of the event data file (~ 50 Gb in total) they are not included as supplementary material here.
+However we have simplified the event data objects into tip rate estimates of the mean and variance across 100 trees + MCC.
+
+###Read in the event data and extract tip data
+
+Now we have a series of trees and event data we can read in. Firstly and most importantly let's check for convergence in the MCC tree through the following:
+
+**Table S3:** Effective sample size (ESS) for the two key BAMM parameters (number of evolutionary shifts and log-Likelihood) for the run on the MCC indicate that BAMM converges (ESS > 200).
+
+```{r}
+#Read in the tree and MCMC to check for convergence 
+MCC.BAMM.tree  <- read.tree("data/BAMM_MCC/MCC.passerine.tre") #Same as other MCC tree already loaded
+MCC.BAMM.mcmc <- read.csv( "data/BAMM_MCC/tree_MCC_mcmc_out.txt" , stringsAsFactors=F)
+
+#Plot of convergence can be generated by:
+#plot(MCC.BAMM.mcmc$logLik ~ MCC.BAMM.mcmc$generation)
+
+#Looks like it has converged so let's discard burn in: 
+burnstart <- floor(0.1 * nrow(MCC.BAMM.mcmc))
+postburn <- MCC.BAMM.mcmc[burnstart:nrow(MCC.BAMM.mcmc), ]
+
+#We can also check effective population sizes of the log-likelihhod and number of shift events in each sample
+#We want at least 200 (although that's on the low side)
+
+cbind(effectiveSize(postburn$N_shifts), effectiveSize(postburn$logLik)) %>% `colnames<-`(c("N_Shifts", "logLik")) %>% `rownames<-`("Effective Sample Size") %>% pander()
+```
+
+Those values are well above 200, so we are good to use the MCC data in analysis!
+We can also check the convergence of BAMM across 100 runs of BAMM.
+The Raw MCMCs are not included in this file or repository but we can read in a dataframe that has extracted the ESS for each of the runs.
+
+**Table S4:** For the 100 trees that BAMM was run on effective sample size (ESS) for the two key BAMM parameters (number of evolutionary shifts and log-Likelihood) also indicates that BAMM converges with the minimum for each parameter across the 100 trees being over 200.
+```{r}
+ESS <- readRDS('data/ESS.rds')
+summary(ESS) %>% pander()
+```
+
+Given that it appears BAMM converges we can we can make a data frame with the mean and variance for extinction and speciation tip rates from the large event data set for the MCC with the following code and then plot the variation we see in tip-rates.
+
+```{r}
+# Read in Event Data
+MCC.BAMM.ED <- getEventData(MCC.BAMM.tree,  "data/BAMM_MCC/tree_MCC_event_data.txt", burnin=0.1, nsamples=1000)
+saveRDS(MCC.BAMM.ED, 'data/MCC.BAMM.ED.rds')
+#From the Event Data we can extract
+
+library(purrr)
+#BAMM.EventData <- readRDS('data/BAMM.EventData.rds')
+
+#Big lapply over each tree in BAMM event data
+BAMM.extraction.function <- function(x) {
+######Get mean and var for lambda
+#Transpose list so each element in the list is a species
+transposed.lambda <- lapply(purrr::transpose(x$tipLambda), unlist)
+
+#Now turn it into a df with mean and variance
+lambda <- sapply(transposed.lambda, function(x) {
+  mean.lambda = mean(log(x))
+  var.lambda = var(log(x))
+  return(c(mean.lambda, var.lambda))
+}) %>% t() %>% as.data.frame() %>% `colnames<-`(c("mean.lambda", "var.lambda"))
+
+lambda$TipLabel <- x[["tip.label"]]
+
+#####NOW FOR Extinction
+
+#Transpose list so each element in the list is a species
+transposed.mu <- lapply(purrr::transpose(x$tipMu), unlist)
+
+#Now turn it into a df with mean and variance
+mu <- sapply(transposed.mu, function(x) {
+  mean.mu = mean(log(x))
+  var.mu = var(log(x))
+  return(c(mean.mu, var.mu))
+}) %>% t() %>% as.data.frame() %>% `colnames<-`(c("mean.mu", "var.mu"))
+
+mu$TipLabel <- x[["tip.label"]]
+
+left_join(lambda, mu, by = "TipLabel")
+}
+
+MCC.BAMM.df <- BAMM.extraction.function(MCC.BAMM.ED)
+
+#Save df for later use
+saveRDS(MCC.BAMM.df, 'data/MCC.BAMM.df.rds')
+```
+
+```{r}
+MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')
+#Plot a summary of logES and TB with the DRs with their weightings
+BAMM.MCC.plot <- MCC.BAMM.df %>% ggplot(aes(x = mean.lambda, y = mean.mu, 
+                          fill = 1/var.lambda, 
+                      size = 1/var.mu))+
+  geom_point(shape = 21, alpha = 0.5)+
+  theme_minimal()+
+  geom_errorbarh(aes(xmin = mean.lambda - 0.674*sqrt(var.lambda), xmax = mean.lambda + 0.674*sqrt(var.lambda)),
+                 size = 0.0025)+
+  geom_errorbar(aes(ymin = mean.mu - 0.674*sqrt(var.mu), ymax = mean.mu + 0.674*sqrt(var.mu)),
+                size = 0.0025)+
+  # scale_y_continuous(trans = "log")+
+  # scale_x_continuous(trans = "log")+
+  # xlim(-8, 3)+
+  # ylim(-10,2)+
+  # scale_x_continuous(trans = "log")+
+  # scale_y_continuous(trans = "log")+
+  theme(legend.position="bottom")+
+  labs(size = 'Inverse log(var) / Weight [using Lambda]', y='Log Extinction [Mu]', x= 'Log Speciation [Lambda]', fill = 'Inverse log(var) / Weight [using Mu] \n')+
+  scale_fill_distiller(guide = "colorbar", palette = "Reds", direction = 1)
+
+BAMM.variance <- ggExtra::ggMarginal(
+  p = BAMM.MCC.plot,
+  type = 'density',
+  margins = 'both',
+  size = 5,
+  colour = 'black',
+  fill = '#BA3B1C91'
+)
+```
+
+```{r, fig.width=10, fig.height=6}
+grid.newpage()
+grid.draw(BAMM.variance)
+```
+**Figure S5:** The tip-rate estimate for BAMM are highly variable within each run of BAMM.
+Across most species there is high variability in the posterior distribution of tip-rate estimates.
+Here we show mean values, weights based on the variance and 50 % CIs.
+
+###Analysis of BAMM results
+
+Given the high variability in tip-rate estimates from the above plot, below we peform some diagnostics on BAMM to demonstrate that the variability is unlikely an error in sampling or parameters, rather an inherent aspect of BAMM.
+The following is not inherently necessary to understandfor the conclusions drawn from the paper, however they do raise a set of methodological questions about BAMM that warrant further investigation.
+
+####Credible number of shifts
+
+To plot the credible shift set, we need the prior distribution on the number of rate shifts (this is generated internally by BAMMtools).
+We can then estimate the credible set of rate shifts using the BAMMtools function credibleShiftSet:
+
+```{r}
+css <- credibleShiftSet(MCC.BAMM.ED, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95)
+
+#Now we obtain the number of distinct shifts: (Out of 1000 samples this is super high, essentially each one is distinct)
+
+css$number.distinct
+```
+
+This value indicates that from a posterior sample of 1000 we have 950 unique rate shifts, which is very high.
+Arguably, across such a large phylogenetic tree, even if the number of shifts reaches convergence the location of those > 50 shifts will be quite variable, with the variation likely impacting tip-rate estimates.
+Even if we increase the sampling of the posterior to 2,500 instead of 1,000 we still see that the CSS is high (close to 2500).
+We can have a look what our top few shifts are:
+
+```{r}
+summary(css)
+```
+
+Notably, each of the top shifts has low probability, indicating that out of the entire posterior sample we cannot differentiate which shift configuration is more likely.
+Based on the follwing eastimates of the BayesFactor, we are convinced that the number of shifts is non-zero, however it is still quite a wide distribution, sourced from potentially weakly non-informative priors.
+But, this is at odds with the ESS and the following plot:
+
+```{r}
+round(computeBayesFactors(MCC.BAMM.mcmc, expectedNumberOfShifts=100, burnin=0.1)[,1], digits = 2)
+```
+
+```{r, fig.width = 8, fig.height = 5}
+plotPrior(MCC.BAMM.mcmc, expectedNumberOfShifts=100)
+```
+**Figure S6:** The apparant convergence of the number of shifts in the posterior is at odds with the variability seen in the CSS and although there is greater certainty in the number of shifts being within the range above, the position of those shifts remain highly variable.
+
+We can compare our run of BAMM against @Harvey_2017 who used BAMM on a genetic-only MCC tree with different parameters
+
+```{r}
+load('data/Hackett_split_eventsample.rda')
+css3 <- credibleShiftSet(ed, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95)
+css3$number.distinct
+```
+
+Based on Their posterior sample of 1250, 1188 are unique. Like our run on the MCC, we find high uncertainity in the position of shift-configurations.
+
+Let's run a model based on genetic data from the Harvey MCC
+
+```{r, fig.width=10, fig.height=6}
+Harvey.BAMM.df <- BAMM.extraction.function(ed)
+
+Harvey.BAMM.df %>% ggplot(aes(x = mean.lambda, y = mean.mu, 
+                          fill = 1/(var.lambda), 
+                      size = 1/(var.mu)))+
+  geom_point(shape = 21, alpha = 0.5)+
+  theme_minimal()+
+  geom_errorbarh(aes(xmin = mean.lambda - 0.674*sqrt(var.lambda), xmax = mean.lambda + 0.674*sqrt(var.lambda)), 
+                 size = 0.0025)+
+  geom_errorbar(aes(ymin = mean.mu - 0.674*sqrt(var.mu), ymax = mean.mu + 0.674*sqrt(var.mu)), 
+                size = 0.0025)+
+  # scale_x_continuous(trans = "log")+
+  # scale_y_continuous(trans = "log")+
+  theme(legend.position="bottom")+
+  labs(size = 'Weight [using Lambda]', y='Log Extinction [Mu]', x= 'Log Speciation [Lambda]', fill = 'Weight [using Mu] \n')+
+  scale_fill_distiller(guide = "colorbar", palette = "Reds", direction = 1)
+
+# BAMM.variance <- ggExtra::ggMarginal(
+#   p = BAMM.MCC.plot,
+#   type = 'density',
+#   margins = 'both',
+#   size = 5,
+#   colour = 'black',
+#   fill = '#BA3B1C91'
+# )
+```
+**Figure S7:** Based on the event data from the BAMM run by @Harvey_2017 we find that in both our case and theirs the tip-rate estimates for many species are extrememly variable.
+here we present mean estimates with 50 % CIs.
+
+##PGLS Models
+
+The method behind running the PGLS models is as follows:
+
+**1:** Estimate the phylogenetic signal ($\lambda$) by running a model without interactions and all six predictor variables.
+The value of $\lambda$ obtained here will be fixed in all successive models.
+The value is fixed for subsequent models as independently estimating it in each case becomes computationally intensive.
+Based on preliminary analysis even slight differences in lambda lead to large reductions in AIC.
+
+**2:** Create a global model of six predictor variables plus the fiver interactions between sexual dichromatism and the other variables.
+
+**3:** Dredge the global model but fix the six independent predictors, hence conducting model selection on the interaction terms.
+
+**4:** Take the top model in the MCC model and run it on the 100 phylogenetic trees.
+
+**5:** Repeat this step for DR, ND, BAMM-speciation and BAMM-extinction.
+
+###PGLS Models on DR and ND
+
+Using ``corPagel`` we can estimate the phylogenetic signal for a model with all predictors (interactions do not appear to affect the estimate of $\lambda$):
+
+```{r, eval = FALSE}
+#Prune tree
+pruned.MCC.tree <- drop.tip(MCC.passerine,MCC.passerine$tip.label[-match(restricted.data$TipLabel, MCC.passerine$tip.label)])
+
+#Set rownames to match tree
+rownames(restricted.data) <- restricted.data$TipLabel
+
+#Run a corPagel model to estimate lambda for DR
+MCC.DR.corPagel <- gls(MCC.DR ~ SDi
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP,
+                correlation = corPagel(1, phy = pruned.MCC.tree, fixed = FALSE), 
+                data = restricted.data, 
+                method = "REML")
+saveRDS(MCC.DR.corPagel, 'data/MCC.DR.corPagel.rds')
+
+#Run a corPagel model to estimate lambda for ND
+MCC.ND.corPagel <- gls(MCC.ND ~ SDi
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP,
+                correlation = corPagel(1, phy = pruned.MCC.tree, fixed = FALSE), 
+                data = restricted.data, 
+                method = "REML")
+saveRDS(MCC.ND.corPagel, 'data/MCC.ND.corPagel.rds')
+```
+
+Inspect the $\lambda$ value, which can then be fixed for successive models.
+```{r, warning=FALSE}
+MCC.DR.corPagel <- readRDS('data/MCC.DR.corPagel.rds')
+MCC.ND.corPagel <- readRDS('data/MCC.ND.corPagel.rds')
+MCC.DR.corPagel[["modelStruct"]][["corStruct"]] %>% `names<-`("DR lambda") %>% pander()
+MCC.ND.corPagel[["modelStruct"]][["corStruct"]] %>% `names<-`("ND lambda") %>% pander()
+```
+
+The lambda is high and not too dissimilar than if we assume Brownian Motion $\lambda = 1$.
+However, given the large sample size this difference may have an effect on the results so, to play safe we included it as a fixed value for $\lambda$ in all successive models.
+
+```{r, eval = FALSE}
+#Run model for DR
+MCC.DR <- gls(MCC.DR ~ SDi
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + SDi*log(range.size.m2)
+                         + SDi*bioclim4
+                         + SDi*residuals.PC1
+                         + SDi*PC1.LIG
+                         + SDi*NPP,
+                correlation = corPagel(0.985, phy = pruned.MCC.tree, fixed = TRUE), 
+                data = restricted.data, 
+                method = "REML")
+
+#Run model for ND
+MCC.ND <- gls(MCC.ND ~ SDi
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP
+                      + SDi*log(range.size.m2)
+                      + SDi*bioclim4
+                      + SDi*residuals.PC1
+                      + SDi*PC1.LIG
+                      + SDi*NPP,
+                correlation = corPagel(0.9996, phy = pruned.MCC.tree, fixed = TRUE), 
+                data = restricted.data, 
+                method = "REML")
+
+#Set up cluster
+cores<-8
+clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
+clust <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
+myclust<-clust
+
+#Export data and packages to cluster
+clusterExport(myclust, c("restricted.data"), envir=environment())
+clusterExport(myclust, c("pruned.MCC.tree"), envir=environment())
+clusterEvalQ(myclust, library(nlme))
+clusterEvalQ(myclust, library(ape))
+clusterEvalQ(myclust, library(MuMIn))
+
+#Dredged models:
+
+dredged.ND.model <- pdredge(MCC.ND, fixed = c("SDi", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust, trace = TRUE)
+
+saveRDS(dredged.ND.model, "data/dredged.ND.model.rds")
+
+dredged.DR.model <- pdredge(MCC.DR, fixed = c("SDi", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust, trace = TRUE)
+
+saveRDS(dredged.DR.model, "data/dredged.DR.model.rds")
+```
+
+**Table S5:** The dredged models both show the top model is one with no interactions, with $\delta AICc > 4$ in both cases.
+We can be resonably confident that interactions are unlikely to affect the pattern of speciation we see in passerine birds.
+```{r}
+dredged.DR.model <- readRDS("data/dredged.DR.model.rds")
+dredged.ND.model <- readRDS("data/dredged.ND.model.rds")
+kable(dredged.DR.model, "html", caption = "DR Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+kable(dredged.ND.model, "html", caption = "ND Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+```
+
+From the dredged model list we can see that the top model is one with no interactions.
+Run this model on the MCC and 100 trees.
+Noting that each model uses a unique set of values for DR/ND and a unique tree in the correlation structure.
+
+```{r, eval=FALSE}
+#In both cases the top model is 1/2/3/4/5/6 no interaction terms. With no models within delta < 4:
+
+#Run model for DR
+MCC.DR.top <- gls(MCC.DR ~ SDi
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP,
+                correlation = corPagel(0.985, phy = pruned.MCC.tree, fixed = TRUE), 
+                data = restricted.data, 
+                method = "REML")
+
+saveRDS(MCC.DR.top, 'data/MCC.DR.top.rds')
+
+#Run model for ND
+MCC.ND.top <- gls(MCC.ND ~ SDi
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP,
+                correlation = corPagel(0.9996, phy = pruned.MCC.tree, fixed = TRUE), 
+                data = restricted.data, 
+                method = "REML")
+
+saveRDS(MCC.ND.top, 'data/MCC.ND.top.rds')
+#Run the 100 models for DR and ND using the best model:
+
+#Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
+DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
+  left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ES")), 
+            by = "TipLabel")
+})
+
+#PGLS needs tiplabel as rowname
+DR.model.data <- lapply(DR.model.data, function(x) {
+  tibble::column_to_rownames(x, "TipLabel")})
+
+#Prune the trees
+pruned.trees<-lapply(passerine.trees, function(x) {
+  drop.tip(x,x$tip.label[-match(restricted.data$TipLabel, x$tip.label)])
+})
+
+#Use mapply to create a list of PGLS global models
+DR.pgls.models <- mcmapply(function(x,y) {
+  gls(ES ~ SDi 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    corPagel(0.985, phy = y, fixed = TRUE), 
+    data = x, 
+    method = "REML")
+}, x = DR.model.data, y = pruned.trees,
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(DR.pgls.models, "data/DR.pgls.models.rds")
+
+#Now for Node Density:
+ND.model.data <- lapply(nd.list, function(x) {
+  left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
+            by = "TipLabel")
+})
+
+#PGLS needs tiplabel as rowname
+ND.model.data <- lapply(ND.model.data, function(x) {
+  tibble::column_to_rownames(x, "TipLabel")})
+
+#Use mapply to create a list of PGLS global models
+ND.pgls.models <- mcmapply(function(x,y) {
+gls(ND ~ SDi 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    corPagel(0.9996, phy = y, fixed = TRUE), 
+    data = x, 
+    method = "REML")
+}, x = ND.model.data, y = pruned.trees, 
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(ND.pgls.models, "data/ND.pgls.models.rds")
+```
+
+From these models we can get a distribution of the estimates for the models alongside the 95 % CIs of the MCC model, this will show the variation beween trees, and the variation associated with the top MCC model.
+
+```{r, fig.height = 8, fig.width = 4, message = FALSE, warning = FALSE}
+DR.pgls.models <- readRDS("data/DR.pgls.models.rds")
+MCC.DR.top <- readRDS('data/MCC.DR.top.rds')
+
+DR.pgls.summary <- lapply(DR.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+MCC.DR.summary <- data.frame(MCC.DR.top$coefficients, confint(MCC.DR.top)) %>% tibble::rownames_to_column()
+
+DR.pgls.summary <- bind_rows(DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+parameter_names <- c(
+                    `bioclim4` = "Temperature Seasonality",
+                    `log(range.size.m2)` = "Range Size (log-transformed)",
+                    `NPP` = "NPP",
+                    `PC1.LIG` = "Long-term Temperature Variation",
+                    `residuals.PC1` = "Spatial Temperature Variation",
+                    `SDi` = "Sexual Dichromatism"
+                    )
+
+DR.plot <-DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+DR.plot <- DR.plot + geom_errorbarh(data = MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_brewer(palette = "Dark2")+
+  scale_color_brewer(palette = "Dark2")+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("DR Models")
+
+#Now plot this for Node Density (ND):
+
+ND.pgls.models <- readRDS("data/ND.pgls.models.rds")
+MCC.ND.top <- readRDS('data/MCC.ND.top.rds')
+
+ND.pgls.summary <- lapply(ND.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+MCC.ND.summary <- data.frame(MCC.ND.top$coefficients, confint(MCC.ND.top)) %>% tibble::rownames_to_column()
+
+ND.pgls.summary <- bind_rows(ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+ND.plot <-ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+ND.plot <- ND.plot + geom_errorbarh(data = MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,0.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_brewer(palette = "Dark2")+
+  scale_color_brewer(palette = "Dark2")+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("ND Models")
+```
+
+##PGLS Models on BAMM estimates
+
+Using the previously generated BAMM dataframe we can now undertake model selection for speciation and extinction.
+The process is similar to that used on DR and ND, except given the use of a Bayesian approach in BAMM we can make use of varying levels of uncertainty between tips (species) by constructing a weighted model, where the weight is the inverse of the variance, meaning more precise estimates of speciation or extinction at a given tip (species) holds higher weight in the model.
+
+Based on preliminary findings we found that Pagal's lambda was = 1 and running ``corPagel`` lead to problemd of convergence.
+Therefore the following models are run assuming Brownian Motion with ``corBrownian``.
+
+```{r, eval = FALSE}
+MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')
+
+#Create model dataframe for use in models
+MCC.BAMM.model.data <- left_join(restricted.data %>% 
+                                   dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+                                 MCC.BAMM.df %>% as.data.frame(), 
+                                 by = "TipLabel")
+
+saveRDS(MCC.BAMM.model.data, 'data/MCC.BAMM.model.data.rds')
+#Prune tree
+pruned.MCC.tree <- drop.tip(MCC.passerine,MCC.passerine$tip.label[-match(MCC.BAMM.model.data$TipLabel, MCC.passerine$tip.label)])
+
+#Set rownames to match tree
+rownames(MCC.BAMM.model.data) <- MCC.BAMM.model.data$TipLabel
+
+#Run model for DR
+MCC.BAMM.lambda <- gls(mean.lambda ~ SDi
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + SDi*log(range.size.m2)
+                         + SDi*bioclim4
+                         + SDi*residuals.PC1
+                         + SDi*PC1.LIG
+                         + SDi*NPP,
+                weights = ~ sqrt(var.lambda), #sqrt to account for overdispersedskewed variance distribution
+                correlation = corBrownian(phy = pruned.MCC.tree), 
+                data = MCC.BAMM.model.data, 
+                method = "REML")
+
+#Run model for DR
+MCC.BAMM.mu <- gls(mean.mu ~ SDi
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + SDi*log(range.size.m2)
+                         + SDi*bioclim4
+                         + SDi*residuals.PC1
+                         + SDi*PC1.LIG
+                         + SDi*NPP,
+                weights = ~ sqrt(var.mu),
+                correlation = corBrownian(phy = pruned.MCC.tree), 
+                data = MCC.BAMM.model.data, 
+                method = "REML")
+
+#Dredge the global MCC models
+
+#Set up cluster
+cores<-8
+clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
+clust <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
+myclust<-clust
+
+#Export data and packages to cluster
+clusterExport(myclust, c("MCC.BAMM.model.data"), envir=environment())
+clusterExport(myclust, c("pruned.MCC.tree"), envir=environment())
+clusterEvalQ(myclust, library(nlme))
+clusterEvalQ(myclust, library(ape))
+clusterEvalQ(myclust, library(MuMIn))
+
+#Dredged models:
+
+dredged.MCC.lambda <- pdredge(MCC.BAMM.lambda, fixed = c("SDi", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(dredged.MCC.lambda, "data/dredged.MCC.lambda.rds")
+
+dredged.MCC.mu <- pdredge(MCC.BAMM.mu, fixed = c("SDi", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(dredged.MCC.mu, "data/dredged.MCC.mu.rds")
+```
+
+**Table S6:** The dredged models for BAMM speciation and BAMM extinction both show the top model is one with no interactions , with $\delta AICc > 4$.
+This is the same situation as DR/ND models (see above).
+```{r}
+dredged.MCC.lambda <- readRDS("data/dredged.MCC.lambda.rds")
+dredged.MCC.mu <- readRDS("data/dredged.MCC.mu.rds")
+kable(dredged.MCC.lambda, "html", caption = "BAMM-Speciation Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+kable(dredged.MCC.mu, "html", caption = "BAMM-Extinction Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+```
+
+Given that we are running models weighted according to the variance of the response ($\lambda_{BAMM}$ and $\mu_{BAMM}$), we can check to see whether a weighted model isw favourable to an unweighted model using an ``anova`` to compare AIC values.
+
+```{r, eval=FALSE}
+#Run the top model for the MCC
+#Run model for ND
+MCC.Lambda.top <- gls(mean.lambda ~ SDi
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP,
+                weights = ~ sqrt(var.lambda),
+                correlation = corBrownian(phy = pruned.MCC.tree), 
+                data = MCC.BAMM.model.data, 
+                method = "REML")
+saveRDS(MCC.Lambda.top, 'data/MCC.Lambda.top.rds')
+
+MCC.Mu.top <- gls(mean.mu ~ SDi
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP,
+                weights = ~ sqrt(var.mu),
+                correlation = corBrownian(phy = pruned.MCC.tree), 
+                data = MCC.BAMM.model.data, 
+                method = "REML")
+
+saveRDS(MCC.Mu.top, 'data/MCC.Mu.top.rds')
+
+#We can also see how the models look without the weightings: 
+MCC.Lambda.top.unweighted <- gls(mean.lambda ~ SDi
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP,
+                correlation = corBrownian(phy = pruned.MCC.tree), 
+                data = MCC.BAMM.model.data, 
+                method = "REML")
+saveRDS(MCC.Lambda.top.unweighted, 'data/MCC.Lambda.top.unweighted.rds')
+
+MCC.Mu.top.unweighted <- gls(mean.mu ~ SDi
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP,
+                correlation = corBrownian(phy = pruned.MCC.tree), 
+                data = MCC.BAMM.model.data, 
+                method = "REML")
+saveRDS(MCC.Mu.top.unweighted, 'data/MCC.Mu.top.unweighted.rds')
+```
+
+**Table S6:** Models with a set of weights for the response based on the inverse of the variance of the posterior distribution in $\lambda_{BAMM}$ and $\mu_{BAMM}$ have much lower AIC values, indicating that the weighting scheme improves the fit of the model.
+In any case, we found that an unweighted model did not change the main conclusions drawn in that they did not generate any statistically significant relationships.
+```{r}
+MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
+MCC.Lambda.top.unweighted <- readRDS('data/MCC.Lambda.top.unweighted.rds')
+MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')
+MCC.Mu.top.unweighted <- readRDS('data/MCC.Mu.top.unweighted.rds')
+
+anova(MCC.Lambda.top, MCC.Lambda.top.unweighted) %>% pander(split.table = Inf)
+anova(MCC.Mu.top, MCC.Mu.top.unweighted) %>% pander(split.table = Inf)
+```
+
+We can run the top (no interactions) model on the 100 trees, each with unique estimates of speciation and extinction from the BAMM runs.
+
+```{r, eval=FALSE}
+#Read in the BAMM data for the 100 trees
+BAMM.df <- readRDS('data/BAMM.df.rds')
+
+#Take the restricted data and make it simpler with just responses and predictors.Note that we join the BAMM for the 100 trees
+
+BAMM.model.data <- lapply(BAMM.df, function(x) { #es.list is a list of ES values calculated earlier
+  left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+            x %>% as.data.frame(), 
+            by = "TipLabel")
+})
+
+#PGLS needs tiplabel as rowname
+BAMM.model.data <- lapply(BAMM.model.data, function(x) {
+  tibble::column_to_rownames(x, "TipLabel")})
+
+#Prune the trees
+pruned.trees<-lapply(passerine.trees, function(x) {
+  drop.tip(x,x$tip.label[-match(restricted.data$TipLabel, x$tip.label)])
+})
+
+#Use mapply to create a list of PGLS global models
+BAMM.lambda.pgls.models <- mcmapply(function(x,y) {
+  gls(mean.lambda ~ SDi 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    weights = ~ sqrt(var.lambda),      
+    corBrownian(phy = y), 
+    data = x, 
+    method = "REML")
+}, x = BAMM.model.data, y = pruned.trees,
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(BAMM.lambda.pgls.models, "data/BAMM.lambda.pgls.models.rds")
+
+#Use mapply to create a list of PGLS global models
+BAMM.mu.pgls.models <- mcmapply(function(x,y) {
+  gls(mean.mu ~ SDi 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    weights = ~ sqrt(var.mu),      
+    corBrownian(phy = y), 
+    data = x, 
+    method = "REML")
+}, x = BAMM.model.data, y = pruned.trees,
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(BAMM.mu.pgls.models, "data/BAMM.mu.pgls.models.rds")
+```
+
+```{r, fig.height = 8, fig.width = 4, warning=FALSE}
+BAMM.lambda.pgls.models <- readRDS("data/BAMM.lambda.pgls.models.rds")
+MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
+
+BAMM.lambda.pgls.summary <- lapply(BAMM.lambda.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+parameter_names <- c(
+                    `bioclim4` = "Temperature Seasonality",
+                    `log(range.size.m2)` = "Range Size (log-transformed)",
+                    `NPP` = "NPP",
+                    `PC1.LIG` = "Long-term Temperature Variation",
+                    `residuals.PC1` = "Spatial Temperature Variation",
+                    `SDi` = "Sexual Dichromatism"
+                    )
+
+MCC.lambda.summary <- data.frame(MCC.Lambda.top$coefficients, confint(MCC.Lambda.top)) %>% tibble::rownames_to_column()
+
+BAMM.lambda.pgls.summary <- bind_rows(BAMM.lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(MCC.lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+BAMM.lambda.plot <-BAMM.lambda.pgls.summary %>% filter(Parameter != "(Intercept)") %>%
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+BAMM.lambda.plot <- BAMM.lambda.plot + geom_errorbarh(data = MCC.lambda.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = MCC.lambda.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,0.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 7, labeller = as_labeller(parameter_names))+
+  scale_fill_brewer(palette = "Dark2")+
+  scale_color_brewer(palette = "Dark2")+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("BAMM Speciation")
+
+#Figure for extinction
+
+BAMM.mu.pgls.models <- readRDS("data/BAMM.mu.pgls.models.rds")
+MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')
+
+BAMM.mu.pgls.summary <- lapply(BAMM.mu.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+MCC.mu.summary <- data.frame(MCC.Mu.top$coefficients, confint(MCC.Mu.top)) %>% tibble::rownames_to_column()
+
+BAMM.mu.pgls.summary <- bind_rows(BAMM.mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(MCC.mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+BAMM.mu.plot <-BAMM.mu.pgls.summary %>% filter(Parameter != "(Intercept)") %>%
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+BAMM.mu.plot <- BAMM.mu.plot + geom_errorbarh(data = MCC.mu.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = MCC.mu.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,0.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 7, labeller = as_labeller(parameter_names))+
+  scale_fill_brewer(palette = "Dark2")+
+  scale_color_brewer(palette = "Dark2")+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,0.5,.5),"cm"))+
+  ggtitle("BAMM Extinction")
+```
+
+###PGLS Model Summary
+
+```{r, fig.width=12, fig.height=9}
+grid.arrange(
+  symmetrise_scale(DR.plot, "x"),
+  symmetrise_scale(ND.plot, "x"),
+  symmetrise_scale(BAMM.lambda.plot, "x"),
+  symmetrise_scale(BAMM.mu.plot, "x"),
+  nrow = 1
+)
+```
+**Figure S8:** This figure is the same basic figure as seen in the manuscript (FIgure 1), it provides model estimates for four response variables across 100 random trees alongside the MCC tree.
+
+**Table S7:** The estimates of the MCC models plotted above are based on the following data tables.
+```{r}
+MCC.DR.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates")
+MCC.ND.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates")
+MCC.lambda.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates")
+MCC.mu.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates")
+```
+
+From the estimates of 100 trees we can calculate HPD intervals.
+Note that these do no account for the CI of each estimate.
+Thus it is a more a representation of tree uncertainty than model uncertainty.
+
+**Table S8:** Highest posterior density intervals are calculated for the above figure (ie; models using RGB values of sexual dichromatism).
+The HPD range is determined using the ``hdi`` function with a 95 % credible interval.
+These intervals do not take into account the variance associated with each interval and thus are not an estimate of model precision.
+Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that in the same direction (+ or -).
+```{r, warning = FALSE}
+hpd.DR.top <- list()
+for(x in unique(DR.pgls.summary$Parameter)) {
+hpd.DR.top[[x]] = hdi(DR.pgls.summary %>% filter(Parameter == x) %>% dplyr::select("Estimate"))
+}
+hpd.DR.top <- bind_rows(hpd.DR.top) %>% `rownames<-`(c("Lower", "Upper")) %>% dplyr::select(-"(Intercept)")
+
+saveRDS(hpd.DR.top, 'data/hpd.DR.top.rds')
+
+#For ND
+hpd.ND.top <- list()
+for(x in unique(ND.pgls.summary$Parameter)) {
+hpd.ND.top[[x]] = hdi(ND.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+hpd.ND.top <- bind_rows(hpd.ND.top) %>% `rownames<-`(c("Lower", "Upper")) %>% dplyr::select(-"(Intercept)")
+
+saveRDS(hpd.ND.top, 'data/hpd.ND.top.rds')
+
+hpd.Lambda.top <- list()
+for(x in unique(BAMM.lambda.pgls.summary$Parameter)) {
+hpd.Lambda.top[[x]] = hdi(BAMM.lambda.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+hpd.Lambda.top <- bind_rows(hpd.Lambda.top) %>% `rownames<-`(c("Lower", "Upper")) %>% dplyr::select(-"(Intercept)")
+
+saveRDS(hpd.Lambda.top, 'data/hpd.Lambda.top.rds')
+
+hpd.Mu.top <- list()
+for(x in unique(BAMM.mu.pgls.summary$Parameter)) {
+hpd.Mu.top[[x]] = hdi(BAMM.mu.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+hpd.Mu.top <- bind_rows(hpd.Mu.top) %>% `rownames<-`(c("Lower", "Upper")) %>% dplyr::select(-"(Intercept)")
+
+saveRDS(hpd.Mu.top, 'data/hpd.Mu.top.rds')
+
+hpd.DR.top %>% pander(split.table = Inf, digits = 3, caption = "DR HPD Interval")
+hpd.ND.top %>% pander(split.table = Inf, digits = 3, caption = "ND HPD Interval")
+hpd.Lambda.top %>% pander(split.table = Inf, digits = 3, caption = "BAMM Speciation HPD Interval")
+hpd.Mu.top %>% pander(split.table = Inf, digits = 3, caption = "BAMM Extinction HPD Interval")
+```
+
+From the HPD 95 % intervals and the intervals for the 95 % CI we can gain an estimate of how much variability there is within a model in comparison to variability between trees.
+To do this we calculate the width of the HPD interval relativel to the 95 % CI interval from the MCC.
+
+**Table S9:** Model estimates from $\lambda_{BAMM}$ and $\mu_{BAMM}$ tip-rate estimates show high levels of variability between trees relative to estimates from $\lambda_{DR}$ and $\lambda_{ND}$.
+The interval ratio is calculated as the wisth of the HPD 95 % CI relative to the MCC 95 % CI.
+A value of 1 suggests that model estimates from 95 % of the trees fall within the MCC 95 % CI.
+Here, the large values for the estimates from BAMM tip-rates suggest that there is high variation in tip-rate estimates across trees that is not see within the DR and ND models.
+```{r}
+hpd.DR.top2 <- hpd.DR.top %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Parameter")
+hpd.DR.top2$HPD.Interval <- hpd.DR.top2$Upper - hpd.DR.top2$Lower
+DR.intervals <- left_join(hpd.DR.top2, MCC.DR.summary, by = "Parameter")
+DR.intervals$MCC.Interval <- DR.intervals$UCI - DR.intervals$LCI
+DR.intervals$DR.IntervalRatio <- hpd.DR.top2$HPD.Interval/DR.intervals$MCC.Interval
+
+hpd.ND.top2 <- hpd.ND.top %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Parameter")
+hpd.ND.top2$HPD.Interval <- hpd.ND.top2$Upper - hpd.ND.top2$Lower
+ND.intervals <- left_join(hpd.ND.top2, MCC.ND.summary, by = "Parameter")
+ND.intervals$MCC.Interval <- ND.intervals$UCI - ND.intervals$LCI
+ND.intervals$ND.IntervalRatio <- hpd.ND.top2$HPD.Interval/ND.intervals$MCC.Interval
+
+hpd.Lambda.top2 <- hpd.Lambda.top %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Parameter")
+hpd.Lambda.top2$HPD.Interval <- hpd.Lambda.top2$Upper - hpd.Lambda.top2$Lower
+Lambda.intervals <- left_join(hpd.Lambda.top2, MCC.lambda.summary, by = "Parameter")
+Lambda.intervals$MCC.Interval <- Lambda.intervals$UCI - Lambda.intervals$LCI
+Lambda.intervals$Lambda.IntervalRatio <- hpd.Lambda.top2$HPD.Interval/Lambda.intervals$MCC.Interval
+
+hpd.Mu.top2 <- hpd.Mu.top %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Parameter")
+hpd.Mu.top2$HPD.Interval <- hpd.Mu.top2$Upper - hpd.Mu.top2$Lower
+Mu.intervals <- left_join(hpd.Mu.top2, MCC.mu.summary, by = "Parameter")
+Mu.intervals$MCC.Interval <- Mu.intervals$UCI - Mu.intervals$LCI
+Mu.intervals$Mu.IntervalRatio <- hpd.Mu.top2$HPD.Interval/Mu.intervals$MCC.Interval
+
+plyr::join_all(list(DR.intervals %>% select(Parameter, DR.IntervalRatio),
+          ND.intervals %>% select(Parameter, ND.IntervalRatio),
+          Lambda.intervals %>% select(Parameter, Lambda.IntervalRatio),
+          Mu.intervals %>% select(Parameter, Mu.IntervalRatio)), by = "Parameter", type = "left") %>% `colnames<-`(c("Parameter", "λDR.Interval.Ratio", "λND.Interval.Ratio", "λBAMM.Interval.Ratio", "μBAMM.Interval.Ratio")) %>%
+  pander(digits = 3, split.table = Inf) 
+```
+
+##Subsetted analysis with Armenta data
+
+Using the dataset from @Armenta_2008 we can conduct a subsetted analysis where dichromatism values are measured from spectrophotmetry and synthesised into a colour discriminability measure, which is a difference measure between the sexes.
+To make this dataset comprable with ours we use the absolute difference between the sexes.
+Therefore making the scale from monochromatism to dichromatism rather than female colouration to male colouration.
+
+```{r, warning = FALSE, fig.width=10, fig.height=6}
+#Read in Armenta data
+Armenta.data <- read.csv('data/Armenta_2008.csv', stringsAsFactors = F)
+MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')
+#A couple of rows have "-", remove from dataset
+Armenta.data <- Armenta.data %>% dplyr::select(binomial, Colour.discriminability) %>% mutate(Colour.discriminability = replace(Colour.discriminability, Colour.discriminability == "-", "NA")) %>% filter(Colour.discriminability != "NA")
+Armenta.data$Colour.discriminability <- Armenta.data$Colour.discriminability %>% as.numeric()
+
+MCC.Armenta.model.data <- inner_join(restricted.data %>% dplyr::select(TipLabel, binomial, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP, MCC.DR, MCC.ND), Armenta.data, by = "binomial")
+
+MCC.Armenta.model.data <- inner_join(MCC.Armenta.model.data, MCC.BAMM.df, by = "TipLabel") %>% filter(Colour.discriminability != "NA")
+
+#Plot the correlation
+p1 <- MCC.Armenta.model.data %>% ggplot(aes(x = SDi, y = Colour.discriminability))+
   geom_point()+
-  scale_x_log10()+
-  scale_y_log10()
-```
-
-The extinction rates of the lumped and split datasets show high similarity, insisting that we could use either.
-
-Let us create a relatively simple model
+  geom_smooth(method = 'loess')+
+  theme_minimal()
+
+grid.newpage()
+ggExtra::ggMarginal(
+  p = p1,
+  type = 'density',
+  margins = 'both',
+  size = 5,
+  colour = 'black',
+  fill = 'gray'
+)
+```
+
+**Figure S9:** There is a correlation between the RGB measures and the Spectophotometry measures, The RGB measures seem to be more noisy around the lower values.
+
+R-squared for relationship between RGB and Spectrophotometry Data:
 ```{r}
-model.1 <- lm(rab.ex.rate.split ~ SDi 
-                + range.size.m2 
-                + bioclim4
-                + bioclim15
-                + residuals.PC1
-                + residuals.PC2
-                + BIO1.LGM.Diff
-                + BIO12.LGM.Diff
-                + NPP.res,
-                data = restricted.data)
-summary(model.1)
-```
-
-Now we can try a PGLS model 
+summary(lm(SDi ~ Colour.discriminability,
+   data = MCC.Armenta.model.data))$r.squared %>% pander()
+
+cor(MCC.Armenta.model.data$SDi, MCC.Armenta.model.data$Colour.discriminability)
+```
+
+Run PGLS model for the data
+```{r, eval = FALSE}
+#Prune tree
+pruned.MCC.Armenta.tree <- drop.tip(MCC.passerine,MCC.passerine$tip.label[-match(MCC.Armenta.model.data$TipLabel, MCC.passerine$tip.label)])
+
+#Set rownames to match tree
+rownames(MCC.Armenta.model.data) <- MCC.Armenta.model.data$TipLabel
+
+#Run a corPagel model to estimate lambda for DR
+MCC.DR.Armenta.global <- gls(MCC.DR ~ Colour.discriminability
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + Colour.discriminability*log(range.size.m2)
+                         + Colour.discriminability*bioclim4
+                         + Colour.discriminability*residuals.PC1
+                         + Colour.discriminability*PC1.LIG
+                         + Colour.discriminability*NPP,
+                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = FALSE), 
+                data = MCC.Armenta.model.data, 
+                method = "REML")
+
+#Run a corPagel model to estimate lambda for DR
+MCC.ND.Armenta.global <- gls(MCC.ND ~ Colour.discriminability
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + Colour.discriminability*log(range.size.m2)
+                         + Colour.discriminability*bioclim4
+                         + Colour.discriminability*residuals.PC1
+                         + Colour.discriminability*PC1.LIG
+                         + Colour.discriminability*NPP,
+                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), #lambda = 1
+                data = MCC.Armenta.model.data, 
+                method = "REML")
+
+#Run a corPagel model to estimate lambda for DR
+MCC.Lambda.Armenta.global <- gls(log(mean.lambda) ~ Colour.discriminability
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + Colour.discriminability*log(range.size.m2)
+                         + Colour.discriminability*bioclim4
+                         + Colour.discriminability*residuals.PC1
+                         + Colour.discriminability*PC1.LIG
+                         + Colour.discriminability*NPP,
+                weights = ~ I(1/log(var.lambda)),
+                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), 
+                data = MCC.Armenta.model.data, 
+                method = "REML")
+
+#Run a corPagel model to estimate lambda for DR
+MCC.Extinction.Armenta.global <- gls(log(mean.mu) ~ Colour.discriminability
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + Colour.discriminability*log(range.size.m2)
+                         + Colour.discriminability*bioclim4
+                         + Colour.discriminability*residuals.PC1
+                         + Colour.discriminability*PC1.LIG
+                         + Colour.discriminability*NPP,
+                weights = ~ I(1/log(var.mu)),
+                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), 
+                data = MCC.Armenta.model.data, 
+                method = "REML")
+```
+
+```{r, eval = FALSE}
+#Set up cluster
+cores<-8
+clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
+clust <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
+myclust<-clust
+
+#Export data and packages to cluster
+clusterExport(myclust, c("MCC.Armenta.model.data"), envir=environment())
+clusterExport(myclust, c("pruned.MCC.Armenta.tree"), envir=environment())
+clusterEvalQ(myclust, library(nlme))
+clusterEvalQ(myclust, library(ape))
+clusterEvalQ(myclust, library(MuMIn))
+
+#Dredged models:
+
+Armenta.dredged.ND.model <- pdredge(MCC.ND.Armenta.global, fixed = c("Colour.discriminability", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(Armenta.dredged.ND.model, 'data/Armenta.dredged.ND.model.rds')
+
+Armenta.dredged.DR.model <- pdredge(MCC.DR.Armenta.global, fixed = c("Colour.discriminability", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(Armenta.dredged.DR.model, 'data/Armenta.dredged.DR.model.rds')
+
+Armenta.dredged.spec.model <- pdredge(MCC.Lambda.Armenta.global, fixed = c("Colour.discriminability", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(Armenta.dredged.spec.model, 'data/Armenta.dredged.spec.model.rds')
+
+Armenta.dredged.extinct.model <- pdredge(MCC.Extinction.Armenta.global, fixed = c("Colour.discriminability", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(Armenta.dredged.extinct.model, 'data/Armenta.dredged.extinct.model.rds')
+```
+
+**Table S10:** The dredged models for Models using a subset dataset where sexual dichromatism was measured using spectrophotometry, all four show the top model is one with no interactions , with $\delta AICc > 4$.
 ```{r}
-hackett.consensus.tree <- read.nexus('bamm_extinction/data/trees/hackett_dale_sub.consensus.nex')
-restricted.data <- restricted.data %>% drop_na(rab.ex.rate.split)
-rownames(restricted.data) <- restricted.data$TipLabel
-pruned.tree<-drop.tip(hackett.consensus.tree,hackett.consensus.tree$tip.label[-match(restricted.data$TipLabel, hackett.consensus.tree$tip.label)])
-pgls.model.1 <- gls(rab.ex.rate.split ~ SDi 
-                + range.size.m2 
-                + bioclim4
-                + bioclim15
-                + residuals.PC1
-                + residuals.PC2
-                + BIO1.LGM.Diff
-                + BIO12.LGM.Diff
-                + NPP.res,
-                correlation = corBrownian(phy = pruned.tree), 
-                data = restricted.data, 
-                method = "REML")
-summary(pgls.model.1)
-```
-
-Or perhaps OU model rather than Brownian:
-
+Armenta.dredged.ND.model <- readRDS("data/Armenta.dredged.ND.model.rds")
+Armenta.dredged.DR.model <- readRDS("data/Armenta.dredged.DR.model.rds")
+Armenta.dredged.spec.model <- readRDS("data/Armenta.dredged.spec.model.rds")
+Armenta.dredged.extinct.model <- readRDS("data/Armenta.dredged.extinct.model.rds")
+
+kable(Armenta.dredged.ND.model, "html", caption = "ND Spectrophotometry Dichromatism Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+
+kable(Armenta.dredged.DR.model, "html", caption = "DR Spectrophotometry Dichromatism Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+
+kable(Armenta.dredged.spec.model, "html", caption = "BAMM Speciation Spectrophotometry Dichromatism Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+
+kable(Armenta.dredged.extinct.model, "html", caption = "BAMM Extinction Spectrophotometry Dichromatism Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+```
+
+Run the top model and 100 models for each one:
+
+```{r, eval = FALSE}
+#In both cases the top model is 1/2/3/4/5/6 no interaction terms. With no models within delta < 4:
+
+#Run model for DR
+Armenta.MCC.DR.top <- gls(MCC.DR ~ Colour.discriminability
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP,
+                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = FALSE), 
+                data = MCC.Armenta.model.data, 
+                method = "REML")
+saveRDS(Armenta.MCC.DR.top, 'data/Armenta.MCC.DR.top.rds')
+
+#Run model for ND
+Armenta.MCC.ND.top <- gls(MCC.ND ~ Colour.discriminability
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP,
+                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), #lambda = 1
+                data = MCC.Armenta.model.data, 
+                method = "REML")
+saveRDS(Armenta.MCC.ND.top, 'data/Armenta.MCC.ND.top.rds')
+
+Armenta.data$Colour.discriminability <- abs(Armenta.data$Colour.discriminability)
+
+#Run the 100 models for DR and ND using the best model:
+Armenta.data.noMCC <- inner_join(restricted.data %>% dplyr::select(TipLabel, binomial, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), Armenta.data, by = "binomial") %>% filter(Colour.discriminability != "NA")
+#Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
+Armenta.DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
+  left_join(Armenta.data.noMCC %>% dplyr::select(binomial, TipLabel, SDi, Colour.discriminability, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "DR")), 
+            by = "TipLabel")
+})
+
+#PGLS needs tiplabel as rowname
+Armenta.DR.model.data <- lapply(Armenta.DR.model.data, function(x) {
+  tibble::column_to_rownames(x, "TipLabel")})
+
+#Prune the trees
+Armenta.pruned.trees<-lapply(passerine.trees, function(x) {
+  drop.tip(x,x$tip.label[-match(MCC.Armenta.model.data$TipLabel, x$tip.label)])
+})
+
+#Use mapply to create a list of PGLS global models
+Armenta.DR.pgls.models <- mcmapply(function(x,y) {
+  gls(DR ~ Colour.discriminability 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    correlation = corPagel(1, phy = y, fixed = FALSE), 
+    data = x, 
+    method = "REML")
+}, x = Armenta.DR.model.data, y = Armenta.pruned.trees,
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(Armenta.DR.pgls.models, "data/Armenta.DR.pgls.models.rds")
+
+#Now for Node Density:
+Armenta.ND.model.data <- lapply(nd.list, function(x) {
+  left_join(Armenta.data.noMCC %>% dplyr::select(binomial, TipLabel, SDi, Colour.discriminability, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
+            by = "TipLabel")
+})
+
+#PGLS needs tiplabel as rowname
+Armenta.ND.model.data <- lapply(Armenta.ND.model.data, function(x) {
+  tibble::column_to_rownames(x, "TipLabel")})
+
+#Use mapply to create a list of PGLS global models
+Armenta.ND.pgls.models <- mcmapply(function(x,y) {
+gls(ND ~ Colour.discriminability 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    corPagel(1, phy = y, fixed = TRUE), 
+    data = x, 
+    method = "REML")
+}, x = Armenta.ND.model.data, y = Armenta.pruned.trees, 
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(Armenta.ND.pgls.models, "data/Armenta.ND.pgls.models.rds")
+
+#Run the BAMM models
+
+Armenta.MCC.Lambda.top <- gls(log(mean.lambda) ~ Colour.discriminability
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP,
+                weights = ~ I(1/log(var.lambda)),
+                correlation = corBrownian(phy = pruned.MCC.Armenta.tree), 
+                data = MCC.Armenta.model.data, 
+                method = "REML")
+saveRDS(Armenta.MCC.Lambda.top, 'data/Armenta.MCC.Lambda.top.rds')
+
+Armenta.MCC.Mu.top <- gls(log(mean.mu) ~ Colour.discriminability
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP,
+                weights = ~ I(1/log(var.mu)),
+                correlation = corBrownian(phy = pruned.MCC.Armenta.tree), 
+                data = MCC.Armenta.model.data, 
+                method = "REML")
+saveRDS(Armenta.MCC.Mu.top, 'data/Armenta.MCC.Mu.top.rds')
+
+Armenta.BAMM.model.data <- lapply(BAMM.df, function(x) { #es.list is a list of ES values calculated earlier
+  left_join(Armenta.data.noMCC %>% dplyr::select(binomial, TipLabel, SDi, Colour.discriminability, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+            x %>% as.data.frame(), 
+            by = "TipLabel")
+})
+
+#PGLS needs tiplabel as rowname
+Armenta.BAMM.model.data <- lapply(Armenta.BAMM.model.data, function(x) {
+  tibble::column_to_rownames(x, "TipLabel")})
+
+#Use mapply to create a list of PGLS global models
+Armenta.BAMM.lambda.pgls.models <- mcmapply(function(x,y) {
+  gls(log(mean.lambda) ~ Colour.discriminability 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    weights = ~ I(1/log(var.lambda)),      
+    corBrownian(phy = y), #lambda = 1
+    data = x, 
+    method = "REML")
+}, x = Armenta.BAMM.model.data, y = Armenta.pruned.trees,
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(Armenta.BAMM.lambda.pgls.models, "data/Armenta.BAMM.lambda.pgls.models.rds")
+
+#Use mapply to create a list of PGLS global models
+Armenta.BAMM.mu.pgls.models <- mcmapply(function(x,y) {
+  gls(log(mean.mu) ~ Colour.discriminability
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    weights = ~ I(1/log(var.mu)),      
+    corBrownian(phy = y), #lambda = 1
+    data = x, 
+    method = "REML")
+}, x = Armenta.BAMM.model.data, y = Armenta.pruned.trees,
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(Armenta.BAMM.mu.pgls.models, "data/Armenta.BAMM.mu.pgls.models.rds")
+```
+
+Using the MCC models as well as the 100 PGLS models we can generate the plots similar to Supp Fig S8:
+
+```{r, fig.height = 8, fig.width = 4, warning=FALSE}
+
+Armenta.DR.pgls.models <- readRDS("data/Armenta.DR.pgls.models.rds")
+Armenta.MCC.DR.top <- readRDS('data/Armenta.MCC.DR.top.rds')
+
+Armenta.DR.pgls.summary <- lapply(Armenta.DR.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+Armenta.MCC.DR.summary <- data.frame(Armenta.MCC.DR.top$coefficients, confint(Armenta.MCC.DR.top)) %>% tibble::rownames_to_column()
+
+Armenta.DR.pgls.summary <- bind_rows(Armenta.DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(Armenta.MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+parameter_names <- c(
+                    `bioclim4` = "Temperature Seasonality",
+                    `log(range.size.m2)` = "Range Size (log-transformed)",
+                    `NPP` = "NPP",
+                    `PC1.LIG` = "Long-term Temperature Variation",
+                    `residuals.PC1` = "Spatial Temperature Variation",
+                    `Colour.discriminability` = "Sexual Dichromatism"
+                    )
+
+Armenta.DR.pgls.summary$Parameter = factor(Armenta.DR.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))
+
+Armenta.MCC.DR.summary$Parameter = factor(Armenta.MCC.DR.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))
+
+Armenta.DR.plot <-Armenta.DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Armenta.DR.plot <- Armenta.DR.plot + geom_errorbarh(data = Armenta.MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Armenta.MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_brewer(palette = "Dark2")+
+  scale_color_brewer(palette = "Dark2")+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("Armenta DR")
+
+####################################
+#For ND
+
+Armenta.ND.pgls.models <- readRDS("data/Armenta.ND.pgls.models.rds")
+Armenta.MCC.ND.top <- readRDS('data/Armenta.MCC.ND.top.rds')
+
+Armenta.ND.pgls.summary <- lapply(Armenta.ND.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+Armenta.MCC.ND.summary <- data.frame(Armenta.MCC.ND.top$coefficients, confint(Armenta.MCC.ND.top)) %>% tibble::rownames_to_column()
+
+Armenta.ND.pgls.summary <- bind_rows(Armenta.ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(Armenta.MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+Armenta.ND.pgls.summary$Parameter = factor(Armenta.ND.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))
+
+Armenta.MCC.ND.summary$Parameter = factor(Armenta.MCC.ND.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))
+
+Armenta.ND.plot <-Armenta.ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Armenta.ND.plot <- Armenta.ND.plot + geom_errorbarh(data = Armenta.MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Armenta.MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_brewer(palette = "Dark2")+
+  scale_color_brewer(palette = "Dark2")+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("Armenta ND")
+
+######################
+
+#For Lambda
+Armenta.BAMM.lambda.pgls.models <- readRDS("data/Armenta.BAMM.lambda.pgls.models.rds")
+Armenta.MCC.Lambda.top <- readRDS('data/Armenta.MCC.Lambda.top.rds')
+
+Armenta.Lambda.pgls.summary <- lapply(Armenta.BAMM.lambda.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+Armenta.MCC.Lambda.summary <- data.frame(Armenta.MCC.Lambda.top$coefficients, confint(Armenta.MCC.Lambda.top)) %>% tibble::rownames_to_column()
+
+Armenta.Lambda.pgls.summary <- bind_rows(Armenta.Lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(Armenta.MCC.Lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+Armenta.Lambda.pgls.summary$Parameter = factor(Armenta.Lambda.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))
+
+Armenta.MCC.Lambda.summary$Parameter = factor(Armenta.MCC.Lambda.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))
+
+Armenta.Lambda.plot <-Armenta.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Armenta.Lambda.plot <- Armenta.Lambda.plot + geom_errorbarh(data = Armenta.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Armenta.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_brewer(palette = "Dark2")+
+  scale_color_brewer(palette = "Dark2")+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("Armenta BAMM Speciaton")
+
+######################
+
+#For Mu
+
+Armenta.BAMM.Mu.pgls.models <- readRDS("data/Armenta.BAMM.mu.pgls.models.rds")
+Armenta.MCC.Mu.top <- readRDS('data/Armenta.MCC.Mu.top.rds')
+
+Armenta.Mu.pgls.summary <- lapply(Armenta.BAMM.Mu.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+Armenta.MCC.Mu.summary <- data.frame(Armenta.MCC.Mu.top$coefficients, confint(Armenta.MCC.Mu.top)) %>% tibble::rownames_to_column()
+
+Armenta.Mu.pgls.summary <- bind_rows(Armenta.Mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(Armenta.MCC.Mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+Armenta.Mu.pgls.summary$Parameter = factor(Armenta.Mu.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))
+
+Armenta.MCC.Mu.summary$Parameter = factor(Armenta.MCC.Mu.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))
+
+Armenta.Mu.plot <-Armenta.Mu.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Armenta.Mu.plot <- Armenta.Mu.plot + geom_errorbarh(data = Armenta.MCC.Mu.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Armenta.MCC.Mu.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_brewer(palette = "Dark2")+
+  scale_color_brewer(palette = "Dark2")+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("Armenta BAMM Extinction")
+```
+
+Plot the results
+
+```{r, fig.width=12, fig.height=9}
+grid.arrange(symmetrise_scale(Armenta.DR.plot, "x"),
+             symmetrise_scale(Armenta.ND.plot, "x"),
+             symmetrise_scale(Armenta.Lambda.plot, "x"), 
+             symmetrise_scale(Armenta.Mu.plot, "x"), 
+             nrow = 1)
+```
+**Figure S9:** Measures of sexual dichromatism using a more precise measure in spectrophotometry do not change the main patterns seen within the PGLS models analysed here.
+This dataset is a subset of the complete dataset (n = 581), thus drawing conclusions for the other predictors (e.g.
+borderline long term temperature variation and spatial temperature variation) potentially risks type I error; additionally these environmental predictors were not measured differently in this analysis compared to the previous above so we have no obvious reason to make conclusions from this subset for the effects of the environmental predictors.
+
+**Table S11:** MCC estimates from the above plot are presented as numerical values below. 
 ```{r}
-pgls.model.3 <- gls(rab.ex.rate.split ~ SDi 
-                + range.size.m2 
-                + bioclim4
-                + bioclim15
-                + residuals.PC1
-                + residuals.PC2
-                + BIO1.LGM.Diff
-                + BIO12.LGM.Diff
-                + NPP.res,
-                correlation = corMartins(1, phy = pruned.tree), 
-                data = restricted.data, 
-                method = "REML")
-summary(pgls.model.3)
-```
-
-Now a Bayesian model that uses a covariance matrix from the phylogeny:
-
+Armenta.MCC.DR.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates from Spec measures")
+Armenta.MCC.ND.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates from Spec measures")
+Armenta.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates from Spec measures")
+Armenta.MCC.Mu.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates from Spec measures")
+```
+
+**Table S12:** Highest posterior density intervals are calculated for the above figure (ie; models using Spectrophotometry values of sexual dichromatism).
+The HPD range is determined using the ``hdi`` function with a 95 % credible interval.
+These intervals do not take into account the variance associated with each interval and thus are not an estimate of model precision.
+Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that in the same direction (+ or -).
+These intervals are calculated in the same way as in **Table S7**.
+```{r, warning = FALSE}
+Armenta.hpd.DR.top <- list()
+Armenta.DR.pgls.summary <- na.omit(Armenta.DR.pgls.summary)
+for(x in unique(Armenta.DR.pgls.summary$Parameter)) {
+Armenta.hpd.DR.top[[x]] = hdi(Armenta.DR.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+Armenta.hpd.DR.top <- bind_rows(Armenta.hpd.DR.top) %>% `rownames<-`(c("Lower", "Upper"))
+
+saveRDS(Armenta.hpd.DR.top, 'data/Armenta.hpd.DR.top.rds')
+
+#For ND
+Armenta.hpd.ND.top <- list()
+Armenta.ND.pgls.summary <- na.omit(Armenta.ND.pgls.summary)
+for(x in unique(Armenta.ND.pgls.summary$Parameter)) {
+Armenta.hpd.ND.top[[x]] = hdi(Armenta.ND.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate)Colour.discriminability
+}
+Armenta.hpd.ND.top <- bind_rows(Armenta.hpd.ND.top) %>% `rownames<-`(c("Lower", "Upper")Colour.discriminability
+
+saveRDS(Armenta.hpd.ND.top, 'data/Armenta.hpd.ND.top.rds')
+
+###############
+Armenta.hpd.Lambda.top <- list()
+Armenta.Lambda.pgls.summary <- na.omit(Armenta.Lambda.pgls.summary)
+for(x in unique(Armenta.Lambda.pgls.summary$Parameter)) {
+Armenta.hpd.Lambda.top[[x]] = hdi(Armenta.Lambda.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+Armenta.hpd.Lambda.top <- bind_rows(Armenta.hpd.Lambda.top) %>% `rownames<-`(c("Lower", "Upper")) 
+saveRDS(Armenta.hpd.Lambda.top, 'data/Armenta.hpd.Lambda.top.rds')
+
+################
+Armenta.hpd.Mu.top <- list()
+Armenta.Mu.pgls.summary <- na.omit(Armenta.Mu.pgls.summary)
+for(x in unique(Armenta.Mu.pgls.summary$Parameter)) {
+Armenta.hpd.Mu.top[[x]] = hdi(Armenta.Mu.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+Armenta.hpd.Mu.top <- bind_rows(Armenta.hpd.Mu.top) %>% `rownames<-`(c("Lower", "Upper"))
+
+saveRDS(Armenta.hpd.Mu.top, 'data/Armenta.hpd.Mu.top.rds')
+
+Armenta.hpd.DR.top %>% pander(split.table = Inf, digits = 3, caption = "Armenta DR HPD Interval")
+Armenta.hpd.ND.top %>% pander(split.table = Inf, digits = 3, caption = "Armenta ND HPD Interval")
+Armenta.hpd.Lambda.top %>% pander(split.table = Inf, digits = 3, caption = "Armenta BAMM Speciation HPD Interval")
+Armenta.hpd.Mu.top %>% pander(split.table = Inf, digits = 3, caption = "Armenta BAMM Extinction HPD Interval")
+```
+
+##Analysis using male-bias measure of sexual selection
+
+Sexual dichromatism is expected to be a good measure of sexual selection strength in birds.
+However the relationship between sexual dichromatism and male-biased measures of sexual selection (social mating system, sexual size dimorphism, and paternal care) is expected to be relatively noisy given the precision of the measurment used [@Dale_2015].
+Here we use a dataset of sexual selection for 2,465 species within the orginal dataset of sexual dichromatism scores from RGB measures.
+This male-bias sexual selection score is based on three components, combined in a phylogenetic pca (ppca) with the following loadings:
+
++ **sexual size dimorphism:** 0.37
++ **social polygyny:** 0.57
++ **paternal care:** -0.57
+
+As such, high values for this score are expected to have high sexual selection (high dimorphism, high polygyny and low paternal care).
+PGLS models using this dataset are calculated in the same process as above.
+
+```{r, warning = FALSE, fig.width=10, fig.height=6}
+SS.subset <- plumage.scores %>% filter(Sexual_selection_ppca != "NA")
+SS.subset <- inner_join(SS.subset, restricted.data %>% dplyr::select(TipLabel, binomial, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP, MCC.DR, MCC.ND), by = "TipLabel")
+SS.subset <- inner_join(SS.subset, MCC.BAMM.df, by = "TipLabel")
+
+SS.subset %>% ggplot(aes(x = SDi, y = Sexual_selection_ppca))+
+  geom_point()+
+  geom_smooth(method = "lm")+
+  theme_minimal()
+```
+**Figure S10:** The relationship between absolute sexual dichromatism and the sexual selection.
+Although positively correlated the distribution is very noisy with species having high sexual dichromatism and low sexual selection and visa versa.Colour.discriminability
 ```{r}
-pruned.tree$edge.length[pruned.tree$edge.length<0]<-1e-9
-pruned.tree$edge.length[pruned.tree$edge.length==0]<-1e-9
-pruned.tree <- check_and_fix_ultrametric(pruned.tree)
-inv.phylo <- MCMCglmm::inverseA(pruned.tree, nodes = "TIPS", scale = TRUE)
-A <- solve(inv.phylo$Ainv)
-rownames(A) <- rownames(inv.phylo$Ainv)
-```
-
-```{r, eval=FALSE}
-model_bayesian <- brm(
-  rab.ex.rate.split ~ SDi 
-                + range.size.m2 
-                + bioclim4
-                + bioclim15
-                + residuals.PC1
-                + residuals.PC2
-                + BIO1.LGM.Diff
-                + BIO12.LGM.Diff
-                + NPP.res, data = restricted.data, 
-  family = gaussian(), 
-  cov_ranef = list(phylo = A),
-                seed = 1,
-                cores = 4, chains = 4, iter = 4000, #Run 4 chains in parallel for 4000 iterations (2000 are burn in)
-                control = list(adapt_delta = 0.99, max_treedepth = 15))
-```
+# summary(lm(SDi ~ Sexual_selection_ppcaColour.discriminability
+#    data = SS.subset))$r.squared %>% pander()
+
+cor(SS.subset$SDi, SS.subset$Sexual_selection_ppca)
+```
+
+```{r, eval = FALSE}
+#Prune tree
+pruned.MCC.Subset.tree <- drop.tip(MCC.passerine,MCC.passerine$tip.label[-match(SS.subset$TipLabel, MCC.passerine$tip.label)])
+
+#Set rownames to match tree
+rownames(SS.subset) <- SS.subset$TipLabel
+
+#Run a corPagel model to estimate lambda for DR
+MCC.DR.Subset.global <- gls(MCC.DR ~ Sexual_selection_ppca
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + Sexual_selection_ppca*log(range.size.m2)
+                         + Sexual_selection_ppca*bioclim4
+                         + Sexual_selection_ppca*residuals.PC1
+                         + Sexual_selection_ppca*PC1.LIG
+                         + Sexual_selection_ppca*NPP,
+                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = FALSE), 
+                data = SS.subset, 
+                method = "REML")
+
+#Run a corPagel model to estimate lambda for DR
+MCC.ND.Subset.global <- gls(MCC.ND ~ Sexual_selection_ppca
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + Sexual_selection_ppca*log(range.size.m2)
+                         + Sexual_selection_ppca*bioclim4
+                         + Sexual_selection_ppca*residuals.PC1
+                         + Sexual_selection_ppca*PC1.LIG
+                         + Sexual_selection_ppca*NPP,
+                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), #lambda = 1
+                data = SS.subset, 
+                method = "REML")
+
+#Run a corPagel model to estimate lambda for DR
+MCC.Lambda.Subset.global <- gls(log(mean.lambda) ~ Sexual_selection_ppca
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variatioColour.discriminability
+                         + NPP
+                         + Sexual_selection_ppca*log(range.size.m2)
+                         + Sexual_selection_ppca*bioclim4
+                         + Sexual_selection_ppca*residuals.PC1
+                         + Sexual_selection_ppca*PC1.LIG
+                         + Sexual_selection_ppca*NPP,
+                weights = ~ I(1/log(var.lambda)),
+                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), 
+                data = SS.subset, 
+                method = "REML")
+
+#Run a corPagel model to estimate lambda for DR
+MCC.Extinction.Subset.global <- gls(log(mean.mu) ~ Sexual_selection_ppca
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP
+                         + Sexual_selection_ppca*log(range.size.m2)
+                         + Sexual_selection_ppca*bioclim4
+                         + Sexual_selection_ppca*residuals.PC1
+                         + Sexual_selection_ppca*PC1.LIG
+                         + Sexual_selection_ppca*NPP,
+                weights = ~ I(1/log(var.mu)),
+                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), 
+                data = SS.subset, 
+                method = "REML")
+```
+
+```{r, eval = FALSE}
+#Set up cluster
+cores<-8
+clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
+clust <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
+myclust<-clust
+
+#Export data and packages to cluster
+clusterExport(myclust, c("SS.subset"), envir=environment())
+clusterExport(myclust, c("pruned.MCC.Subset.tree"), envir=environment())
+clusterEvalQ(myclust, library(nlme))
+clusterEvalQ(myclust, library(ape))
+clusterEvalQ(myclust, library(MuMIn))
+
+#Dredged models:
+
+Subset.dredged.ND.model <- pdredge(MCC.ND.Subset.global, fixed = c("Sexual_selection_ppca", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(Subset.dredged.ND.model, 'data/Subset.dredged.ND.model.rds')
+
+Subset.dredged.DR.model <- pdredge(MCC.DR.Subset.global, fixed = c("Sexual_selection_ppca", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(Subset.dredged.DR.model, 'data/Subset.dredged.DR.model.rds')
+
+Subset.dredged.spec.model <- pdredge(MCC.Lambda.Subset.global, fixed = c("Sexual_selection_ppca", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(Subset.dredged.spec.model, 'data/Subset.dredged.spec.model.rds')
+
+Subset.dredged.extinct.model <- pdredge(MCC.Extinction.Subset.global, fixed = c("Sexual_selection_ppca", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)
+
+saveRDS(Subset.dredged.spec.model, 'data/Subset.dredged.extinct.model.rds')
+```
+
+**Table S13:** All top models ($\delta AICc > 4$) using male-biased sexual selection measures do not contain interactions between sexual selection and the environmental variables.
+Thus interaction terms are not included in further analysis.
 
 ```{r}
-summary(model_bayesian)
-```
-
-Try looking at speciation rate using ES-Sim adapted by @Harvey_essim_2017 and found here https://github.com/mgharvey/ES-sim
+Subset.dredged.ND.model <- readRDS("data/Subset.dredged.ND.model.rds")
+Subset.dredged.DR.model <- readRDS("data/Subset.dredged.DR.model.rds")
+Subset.dredged.spec.model <- readRDS("data/Subset.dredged.spec.model.rds")
+Subset.dredged.extinct.model <- readRDS("data/Subset.dredged.extinct.model.rds")
+
+kable(Subset.dredged.ND.model, "html", caption = "ND Male-bias Sexual Selection Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+
+kable(Subset.dredged.DR.model, "html", caption = "DR Male-bias Sexual Selection Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+
+kable(Subset.dredged.spec.model, "html", caption = "BAMM Speciation Male-bias Sexual Selection Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+
+kable(Subset.dredged.extinct.model, "html", caption = "BAMM Extinction Male-bias Sexual Selection Dredge Table") %>%
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+```
+
+```{r, eval = FALSE}
+#Run model for DR
+Subset.MCC.DR.top <- gls(MCC.DR ~ Sexual_selection_ppca
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP,
+                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = FALSE), 
+                data = SS.subset, 
+                method = "REML")
+saveRDS(Subset.MCC.DR.top, 'data/Subset.MCC.DR.top.rds')
+
+#Run model for ND
+Subset.MCC.ND.top <- gls(MCC.ND ~ Sexual_selection_ppca
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP,
+                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), #lambda = 1
+                data = SS.subset, 
+                method = "REML")
+saveRDS(Subset.MCC.ND.top, 'data/Subset.MCC.ND.top.rds')
+
+#Run the 100 models for DR and ND using the best model:
+Subset.data.noMCC <- SS.subset %>% dplyr::select(TipLabel, Sexual_selection_ppca, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP)
+
+#Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
+Subset.DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
+  left_join(Subset.data.noMCC, 
+            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "DR")), 
+            by = "TipLabel")
+})
+
+#PGLS needs tiplabel as rowname
+Subset.DR.model.data <- lapply(Subset.DR.model.data, function(x) {
+  tibble::column_to_rownames(x, "TipLabel")})
+
+#Prune the trees
+Subset.pruned.trees<-lapply(passerine.trees, function(x) {
+  drop.tip(x,x$tip.label[-match(SS.subset$TipLabel, x$tip.label)])
+})
+
+#Use mapply to create a list of PGLS global models
+Subset.DR.pgls.models <- mcmapply(function(x,y) {
+  gls(DR ~ Sexual_selection_ppca 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    correlation = corPagel(0.9711, phy = y, fixed = TRUE), 
+    data = x, 
+    method = "REML")
+}, x = Subset.DR.model.data, y = Subset.pruned.trees,
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(Subset.DR.pgls.models, "data/Subset.DR.pgls.models.rds")
+
+#Now for Node Density:
+Subset.ND.model.data <- lapply(nd.list, function(x) { #es.list is a list of ES values calculated earlier
+  left_join(Subset.data.noMCC, 
+            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
+            by = "TipLabel")
+})
+
+#PGLS needs tiplabel as rowname
+Subset.ND.model.data <- lapply(Subset.ND.model.data, function(x) {
+  tibble::column_to_rownames(x, "TipLabel")})
+
+#Use mapply to create a list of PGLS global models
+Subset.ND.pgls.models <- mcmapply(function(x,y) {
+gls(ND ~ Sexual_selection_ppca 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    corPagel(1, phy = y, fixed = TRUE), 
+    data = x, 
+    method = "REML")
+}, x = Subset.ND.model.data, y = Subset.pruned.trees, 
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(Subset.ND.pgls.models, "data/Subset.ND.pgls.models.rds")
+
+#BAMM Top Models
+
+Subset.MCC.Lambda.top <- gls(log(mean.lambda) ~ Sexual_selection_ppca
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP,
+                weights = ~ I(1/log(var.lambda)),
+                correlation = corBrownian(phy = pruned.MCC.Subset.tree), 
+                data = SS.subset, 
+                method = "REML")
+saveRDS(Subset.MCC.Lambda.top, 'data/Subset.MCC.Lambda.top.rds')
+
+Subset.MCC.Mu.top <- gls(log(mean.mu) ~ Sexual_selection_ppca
+                      + log(range.size.m2)
+                      + bioclim4 #Seasonal variation
+                      + residuals.PC1 #Spatial variation
+                      + PC1.LIG #Long-term climate variation
+                      + NPP,
+                weights = ~ I(1/log(var.mu)),
+                correlation = corBrownian(phy = pruned.MCC.Subset.tree), 
+                data = SS.subset, 
+                method = "REML")
+saveRDS(Subset.MCC.Mu.top, 'data/Subset.MCC.Mu.top.rds')
+
+#Now for BAMM 100 models
+
+Subset.BAMM.model.data <- lapply(BAMM.df, function(x) { #es.list is a list of ES values calculated earlier
+  left_join(Subset.data.noMCC %>% dplyr::select(TipLabel, Sexual_selection_ppca, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+            x %>% as.data.frame(), 
+            by = "TipLabel")
+})
+
+#PGLS needs tiplabel as rowname
+Subset.BAMM.model.data <- lapply(Subset.BAMM.model.data, function(x) {
+  tibble::column_to_rownames(x, "TipLabel")})
+
+#Use mapply to create a list of PGLS global models
+Subset.BAMM.lambda.pgls.models <- mcmapply(function(x,y) {
+  gls(log(mean.lambda) ~ Sexual_selection_ppca 
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    weights = ~ I(1/log(var.lambda)),      
+    corBrownian(phy = y), #lambda = 1
+    data = x, 
+    method = "REML")
+}, x = Subset.BAMM.model.data, y = Subset.pruned.trees,
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(Subset.BAMM.lambda.pgls.models, "data/Subset.BAMM.lambda.pgls.models.rds")
+
+#Use mapply to create a list of PGLS global models
+Subset.BAMM.mu.pgls.models <- mcmapply(function(x,y) {
+  gls(log(mean.mu) ~ Sexual_selection_ppca
+         + log(range.size.m2)
+         + bioclim4 #Seasonal variation
+         + residuals.PC1 #Spatial variation
+         + PC1.LIG #Long-term climate variation
+         + NPP,
+    weights = ~ I(1/log(var.mu)),      
+    corBrownian(phy = y), #lambda = 1
+    data = x, 
+    method = "REML")
+}, x = Subset.BAMM.model.data, y = Subset.pruned.trees,
+SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+mc.cores = 8) #Specify core number
+
+saveRDS(Subset.BAMM.mu.pgls.models, "data/Subset.BAMM.mu.pgls.models.rds")
+```
+
+The phylogenetic signal from the DR model is ~0.97. The other models had phylogenetic signals ~ 1, as such they were fixed at 1 to avoid problems of convergence.
+
+```{r, warning = FALSE}
+Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')
+Subset.MCC.DR.top[["modelStruct"]][["corStruct"]] %>% `names<-`("DR lambda") %>% pander()
+```
 
 ```{r}
-Essim.frame <- restricted.data$SDi
-names(Essim.frame) <- restricted.data$TipLabel
-
-essim.results <- essim(pruned.tree, Essim.frame, nsim = 1000, return.es = TRUE)
-```
-
-We can then add the essim.results and run a similar PGLS model
+Subset.DR.pgls.models <- readRDS("data/Subset.DR.pgls.models.rds")
+Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')
+
+Subset.cols <- brewer.pal(n = 7, name = "Dark2")[-6]
+
+Subset.DR.pgls.summary <- lapply(Subset.DR.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+Subset.MCC.DR.summary <- data.frame(Subset.MCC.DR.top$coefficients, confint(Subset.MCC.DR.top)) %>% tibble::rownames_to_column()
+
+Subset.DR.pgls.summary <- bind_rows(Subset.DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(Subset.MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+parameter_names <- c(
+                    `bioclim4` = "Temperature Seasonality",
+                    `log(range.size.m2)` = "Range Size (log-transformed)",
+                    `NPP` = "NPP",
+                    `PC1.LIG` = "Long-term Temperature Variation",
+                    `residuals.PC1` = "Spatial Temperature Variation",
+                    `Sexual_selection_ppca` = "Sexual Selection"
+                    )
+
+Subset.DR.pgls.summary$Parameter = factor(Subset.DR.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.MCC.DR.summary$Parameter = factor(Subset.MCC.DR.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.DR.plot <-Subset.DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Subset.DR.plot <- Subset.DR.plot + geom_errorbarh(data = Subset.MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Subset.MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = Subset.cols)+
+  scale_color_manual(values = Subset.cols)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("Male-Bias SS DR")
+
+####################################
+#For ND
+
+Subset.ND.pgls.models <- readRDS("data/Subset.ND.pgls.models.rds")
+Subset.MCC.ND.top <- readRDS('data/Subset.MCC.ND.top.rds')
+
+Subset.ND.pgls.summary <- lapply(Subset.ND.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+Subset.MCC.ND.summary <- data.frame(Subset.MCC.ND.top$coefficients, confint(Subset.MCC.ND.top)) %>% tibble::rownames_to_column()
+
+Subset.ND.pgls.summary <- bind_rows(Subset.ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(Subset.MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+Subset.ND.pgls.summary$Parameter = factor(Subset.ND.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.MCC.ND.summary$Parameter = factor(Subset.MCC.ND.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.ND.plot <-Subset.ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Subset.ND.plot <- Subset.ND.plot + geom_errorbarh(data = Subset.MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Subset.MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = Subset.cols)+
+  scale_color_manual(values = Subset.cols)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("Male-Bias SS ND")
+
+######################
+
+#For Lambda
+Subset.BAMM.lambda.pgls.models <- readRDS("data/Subset.BAMM.lambda.pgls.models.rds")
+Subset.MCC.Lambda.top <- readRDS('data/Subset.MCC.Lambda.top.rds')
+
+Subset.Lambda.pgls.summary <- lapply(Subset.BAMM.lambda.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+Subset.MCC.Lambda.summary <- data.frame(Subset.MCC.Lambda.top$coefficients, confint(Subset.MCC.Lambda.top)) %>% tibble::rownames_to_column()
+
+Subset.Lambda.pgls.summary <- bind_rows(Subset.Lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(Subset.MCC.Lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+Subset.Lambda.pgls.summary$Parameter = factor(Subset.Lambda.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.MCC.Lambda.summary$Parameter = factor(Subset.MCC.Lambda.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.Lambda.plot <-Subset.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Subset.Lambda.plot <- Subset.Lambda.plot + geom_errorbarh(data = Subset.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Subset.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = Subset.cols)+
+  scale_color_manual(values = Subset.cols)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("Male-Bias SS Speciaton")
+
+######################
+
+#For Mu
+
+Subset.BAMM.Mu.pgls.models <- readRDS("data/Subset.BAMM.mu.pgls.models.rds")
+Subset.MCC.Mu.top <- readRDS('data/Subset.MCC.Mu.top.rds')
+
+Subset.Mu.pgls.summary <- lapply(Subset.BAMM.Mu.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+Subset.MCC.Mu.summary <- data.frame(Subset.MCC.Mu.top$coefficients, confint(Subset.MCC.Mu.top)) %>% tibble::rownames_to_column()
+
+Subset.Mu.pgls.summary <- bind_rows(Subset.Mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+colnames(Subset.MCC.Mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+
+Subset.Mu.pgls.summary$Parameter = factor(Subset.Mu.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.MCC.Mu.summary$Parameter = factor(Subset.MCC.Mu.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.Mu.plot <-Subset.Mu.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Subset.Mu.plot <- Subset.Mu.plot + geom_errorbarh(data = Subset.MCC.Mu.summary %>% filter(Parameter != "(Intercept)"), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Subset.MCC.Mu.summary %>% filter(Parameter != "(Intercept)"), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = Subset.cols)+
+  scale_color_manual(values = Subset.cols)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle("Male-Bias SS Extinction")
+
+#########################################
+```
+
+```{r, fig.width=12, fig.height=9}
+grid.arrange(symmetrise_scale(Subset.DR.plot, "x"),
+             symmetrise_scale(Subset.ND.plot, "x"),
+             symmetrise_scale(Subset.Lambda.plot, "x"), 
+             symmetrise_scale(Subset.Mu.plot, "x"), 
+             nrow = 1)
+```
+**Figure S11:** Measures of sexual selection a phylogenetic pca of sexual dimorphism, social polygyny and paternal care provide some evidence to suggest sexual selection predicts extinction rate (DR and ND models).
+This figure is seen in Figure 1 within the manuscript
+
+**Table S14:** MCC model estimates from the above model are presented here as numberical values with 95 % CIs.
 ```{r}
-restricted.data$ESsim <- essim.results[["es"]]
-
-restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = (ESsim))) + 
+Subset.MCC.DR.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates from Male-bias SS")
+Subset.MCC.ND.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates from Male-bias SS")
+Subset.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates from Male-bias SS")
+Subset.MCC.Mu.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates from Male-bias SS")
+```
+
+**Table S15:** Highest posterior density intervals are calculated for the above figure (ie; models using Male-bias sexual selection measures).
+using this measure the models using $\lambda_{DR}$ have model estimates that fall on the positive side 95 % of the time in 100 trees.
+For $\lambda_{ND}$, there is a positive skew, however the 95 % HPD interval overlaps zero.
+$\lambda_{BAMM}$ also shows a lesser positive skew.
+The HPD range is determined using the ``hdi`` function with a 95 % credible interval.
+These intervals do not take into account the variance associated with each interval and thus are not an estimate of model precision.
+Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that in the same direction (+ or -).
+These intervals are calculated in the same way as in **Table S7** and **Table S9**.
+```{r, warning = FALSE}
+Subset.hpd.DR.top <- list()
+Subset.DR.pgls.summary <- na.omit(Subset.DR.pgls.summary)
+for(x in unique(Subset.DR.pgls.summary$Parameter)) {
+Subset.hpd.DR.top[[x]] = hdi(Subset.DR.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+Subset.hpd.DR.top <- bind_rows(Subset.hpd.DR.top) %>% `rownames<-`(c("Lower", "Upper"))
+
+saveRDS(Subset.hpd.DR.top, 'data/Subset.hpd.DR.top.rds')
+
+#For ND
+Subset.hpd.ND.top <- list()
+Subset.ND.pgls.summary <- na.omit(Subset.ND.pgls.summary)
+for(x in unique(Subset.ND.pgls.summary$Parameter)) {
+Subset.hpd.ND.top[[x]] = hdi(Subset.ND.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+Subset.hpd.ND.top <- bind_rows(Subset.hpd.ND.top) %>% `rownames<-`(c("Lower", "Upper"))
+
+saveRDS(Subset.hpd.ND.top, 'data/Subset.hpd.ND.top.rds')
+
+###############
+Subset.hpd.Lambda.top <- list()
+Subset.Lambda.pgls.summary <- na.omit(Subset.Lambda.pgls.summary)
+for(x in unique(Subset.Lambda.pgls.summary$Parameter)) {
+Subset.hpd.Lambda.top[[x]] = hdi(Subset.Lambda.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+Subset.hpd.Lambda.top <- bind_rows(Subset.hpd.Lambda.top) %>% `rownames<-`(c("Lower", "Upper")) 
+saveRDS(Subset.hpd.Lambda.top, 'data/Subset.hpd.Lambda.top.rds')
+
+################
+Subset.hpd.Mu.top <- list()
+Subset.Mu.pgls.summary <- na.omit(Subset.Mu.pgls.summary)
+for(x in unique(Subset.Mu.pgls.summary$Parameter)) {
+Subset.hpd.Mu.top[[x]] = hdi(Subset.Mu.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
+}
+Subset.hpd.Mu.top <- bind_rows(Subset.hpd.Mu.top) %>% `rownames<-`(c("Lower", "Upper"))
+
+saveRDS(Subset.hpd.Mu.top, 'data/Subset.hpd.Mu.top.rds')
+
+Subset.hpd.DR.top %>% pander(split.table = Inf, digits = 3, caption = "Subset DR HPD Interval")
+Subset.hpd.ND.top %>% pander(split.table = Inf, digits = 3, caption = "Subset ND HPD Interval")
+Subset.hpd.Lambda.top %>% pander(split.table = Inf, digits = 3, caption = "Subset BAMM Speciation HPD Interval")
+Subset.hpd.Mu.top %>% pander(split.table = Inf, digits = 3, caption = "Subset BAMM Extinction HPD Interval")
+```
+
+#Additional Figures and Tables
+
+```{r, fig.width = 8, fig.height = 6, eval = FALSE}
+BAMM.Rates.100Trees <- do.call(rbind, BAMM.df)
+BAMM.Rates.100Trees$mean.lambda <- log(BAMM.Rates.100Trees$mean.lambda)
+BAMM.Lambda.100Trees <- ggiraphExtra::summarySE(BAMM.Rates.100Trees, measurevar = "mean.lambda", groupvars = "TipLabel")
+
+DR.list <- plyr::llply(es.list, function(x) {
+  x %>% as.data.frame %>% rownames_to_column()
+  })
+DR.list <- do.call(rbind, DR.list)
+DR.Lambda.100Trees <- ggiraphExtra::summarySE(DR.list, measurevar = ".", groupvars = "rowname")
+DR.Lambda.100Trees <- DR.Lambda.100Trees %>% rename(DR.Lambda.100Trees, DR = ., TipLabel = rowname)
+
+BAMM.Lambda.100Trees %>% ggplot(aes(x = mean.lambda, y = DR.Lambda.100Trees$DR))+
   geom_point()+
-  geom_smooth(method = "lm")
-```
-
- Once we have our EsSim results we can use them in a linear model to assess the relationship between speciation and our predictors
+  geom_errorbarh(aes(xmin = mean.lambda - 1.96*sd, xmax = mean.lambda + 1.96*sd), size = 0.05)+
+  geom_errorbar(aes(ymin = DR.Lambda.100Trees$DR - 1.96*DR.Lambda.100Trees$sd, ymax = DR.Lambda.100Trees$DR + 1.96*DR.Lambda.100Trees$sd), size = 0.05)+
+  theme_minimal()
+
+mean(DR.Lambda.100Trees$sd)
+mean(BAMM.Lambda.100Trees$sd)
+```
+**Figure S12:**
+
+```{r, eval = FALSE}
+all.trees.join <- left_join(DR.Lambda.100Trees %>% select(TipLabel, DR), BAMM.Rates.100Trees %>% select(mean.lambda, TipLabel))
+cor(all.trees.join$DR, all.trees.join$mean.lambda)
+```
+
+```{r, eval = FALSE}
+MCC.BAMM.model.data <- readRDS('data/MCC.BAMM.model.data.rds')
+restricted.data %>% ggplot(aes(x = MCC.DR, y = MCC.BAMM.model.data$mean.lambda))+
+  geom_point()+
+  scale_y_continuous(trans = "log")+
+  theme_minimal()
+```
+**Figure S13:**
+
+```{r, eval = FALSE}
+MCC.tree.join <- left_join(restricted.data %>% select(MCC.DR, TipLabel), MCC.BAMM.model.data %>% select(mean.lambda, TipLabel))
+cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda)
+```
+
+```{r, eval = FALSE}
+MCC.BAMM.df$CV.lambda <- (sqrt(MCC.BAMM.df$var.lambda) / MCC.BAMM.df$mean.lambda)*100
+mean(MCC.BAMM.df$CV.lambda)
+min(MCC.BAMM.df$CV.lambda)
+max(MCC.BAMM.df$CV.lambda)
+
+MCC.BAMM.df %>% ggplot(aes(x = mean.lambda, y = CV.lambda)) +
+                         geom_point()+
+  scale_y_log10()+
+  scale_x_log10()
+```
+
+#R Session information
 
 ```{r}
-model.2 <- lm(ESsim ~ SDi 
-                + range.size.m2 
-                + bioclim4
-                + bioclim15
-                + residuals.PC1
-                + residuals.PC2
-                + BIO1.LGM.Diff
-                + BIO12.LGM.Diff
-                + NPP.res,
-                data = restricted.data)
-summary(model.2)
-```
-
-Have a look at how these would look if we obtained model predictions for each data point.
-
-```{r, fig.width=10, fig.height=5}
-predict.frame <- cbind(predict(model.2),restricted.data)
-predict.frame %>% ggplot(aes(y = `predict(model.2)`, x = SDi)) + 
-  geom_point(aes(color = BLFamilyLatin))+
-  scale_color_hue(h = c(0,100), l = 50) +
-  geom_smooth(color = 'black')+
-  theme_minimal()
-```
-
-It looks like some of the higher SDi's have slightly higher speciation rates. Perhaps we can look at those obviously dimorphic groups as category of its own.
-
-```{r}
-restricted.data %>% ggplot(aes(SDi))+
-  geom_histogram(binwidth = 0.015)+
-  theme_minimal()
-```
+sessionInfo() %>% pander
+```
+
+# References
+<div id="refs"></div>

### 2018-09-14 bamm_extinction/BirdExtinction.Rmd
Long time no update

@@
 
-```{r}
+#Setup
+
+```{r, warning = FALSE}
 library(ape)
@@
 ```
-
-##BAMM Setup
 
@@
 
-``bamm -c control_consensus_eric.txt``
+#BAMM Setup
+
+##Load trees and check them
+
+Read in 100 Passerine Trees from birdtree.org (5966 species) based on the Hackett backbone.
+```{r}
+passerine.trees <- read.nexus('data/trees/Passerine_Trees_Full.nex')
+```
+
+For BAMM to work, we need to check to see if trees are ultrametric, binary and edge lengths are greater than 0.
+
+We can check this on the first tree
+```{r}
+is.ultrametric(passerine.trees[["tree_8950"]])
+is.binary.tree(passerine.trees[["tree_8950"]])
+# Now to check min branch length:
+min(passerine.trees[["tree_8950"]]$edge.length)
+```
+
+> good
+
+We can check all 100 trees with this slightly messy script:
+```{r, eval=FALSE}
+is.ultrametric(passerine.trees)
+is.binary.tree(passerine.trees)
+# Now to check min branch length:
+name.passerine.tree <- names(passerine.trees)
+for(x in name.passerine.tree) {
+  if (min(passerine.trees$edge.length[[x]]) > 0)
+    print("good")
+
+  else print("bad")
+}
+```
+
+> Good, we can use any tree for BAMM
+
+##Writing the Control Files
+
+Create a control file for all the trees (100 files and 100 trees). To do this we can use the ``sapply`` function.
+
+```{r, results = 'hide'}
+name.passerine.tree <- names(passerine.trees)
+
+priors <- sapply(name.passerine.tree, function(x) {
+  setBAMMpriors(passerine.trees[[x]], outfile = NULL)
+})
+
+sapply(name.passerine.tree, function(x) {
+  write.tree(passerine.trees[[x]], paste("data/bamm_files/", x, ".tre", sep=""))
+})
+
+# Here is a block of parameters for the control file. We can make a control file for each tree:
+params <- list()
+for (x in name.passerine.tree) {
+
+# GENERAL SETUP AND DATA INPUT
+
+params[[x]] <- list(modeltype = 'speciationextinction',
+# Specify "speciationextinction" or "trait" analysis
+
+treefile = paste(x, ".tre", sep=""),
+# File name of the phylogenetic tree to be analyzed
+
+runInfoFilename = 'run_info.txt',
+# File name to output general information about this run
+
+sampleFromPriorOnly = 0,
+# Whether to perform analysis sampling from prior only (no likelihoods computed)
+
+runMCMC = 1,
+# Whether to perform the MCMC simulation. If runMCMC = 0, the program will only
+# check whether the data file can be read and the initial likelihood computed
+
+loadEventData = 0,                       
+# Whether to load a previous event data file
+
+eventDataInfile = 'event_data_in.txt',
+# File name of the event data file to load, used only if loadEventData = 1
+
+initializeModel = 1,
+# Whether to initialize (but not run) the MCMC. If initializeModel = 0, the
+# program will only ensure that the data files (e.g., treefile) can be read
+
+useGlobalSamplingProbability = 1,
+# Whether to use a "global" sampling probability. If False (0), expects a file
+# name for species-specific sampling probabilities (see sampleProbsFilename)
+
+globalSamplingFraction = 1,
+# The sampling probability. If useGlobalSamplingProbability = 0, this is ignored
+# and BAMM looks for a file name with species-specific sampling fractions
+
+sampleProbsFilename = 'sample_probs.txt',
+# File name containing species-specific sampling fractions
+
+seed = as.numeric(gsub("tree_", "", x, perl = TRUE)),
+# Seed for the random number generator. Set for reproducibility to the number of the treefile
+
+overwrite = 1,
+# If True (1), the program will overwrite any output files in the current
+# directory (if present)
+
+# PRIORS
+
+expectedNumberOfShifts = 100,
+# prior on the number of shifts in diversification
+# Suggested values: 
+#     expectedNumberOfShifts = 1.0 for small trees (< 500 tips)
+#  expectedNumberOfShifts = 10 or even 50 for large trees (> 5000 tips)
+
+lambdaInitPrior = as.numeric(priors['lambdaInitPrior', x]),
+# Prior (rate parameter of exponential) on the initial lambda value for rate
+# regimes
+
+lambdaShiftPrior = 0.05,
+# Prior (std dev of normal) on lambda shift parameter for rate regimes
+# You cannot adjust the mean of this distribution (fixed at zero, which is
+# equal to a constant rate diversification process)
+
+muInitPrior = as.numeric(priors['muInitPrior', x]),
+# Prior (rate parameter of exponential) on extinction rates
+
+lambdaIsTimeVariablePrior = 1,
+# Prior (probability) of the time mode being time-variable (vs. time-constant)
+
+# MCMC SIMULATION SETTINGS & OUTPUT OPTIONS
+
+numberOfGenerations = '100000000',
+# Number of generations to perform MCMC simulation
+
+mcmcOutfile = 'mcmc_out.txt',
+# File name for the MCMC output, which only includes summary information about
+# MCMC simulation (e.g., log-likelihoods, log-prior, number of processes)
+
+mcmcWriteFreq = 1000,
+# Frequency in which to write the MCMC output to a file
+
+eventDataOutfile = 'event_data.txt',
+# The raw event data (these are the main results). ALL of the results are
+# contained in this file, and all branch-specific speciation rates, shift
+# positions, marginal distributions etc can be reconstructed from this output.
+# See R package BAMMtools for working with this output
+
+eventDataWriteFreq = 1000,
+# Frequency in which to write the event data to a file
+
+printFreq = 10000,
+# Frequency in which to print MCMC status to the screen
+
+acceptanceResetFreq = 1000,
+# Frequency in which to reset the acceptance rate calculation
+# The acceptance rate is output to both the MCMC data file and the screen
+
+outName = x,
+# Optional name that will be prefixed on all output files (separated with "_")
+# If commented out, no prefix will be used
+
+# OPERATORS: MCMC SCALING OPERATORS
+
+updateLambdaInitScale = 2,
+# Scale parameter for updating the initial speciation rate for each process
+
+updateLambdaShiftScale = 0.1,
+# Scale parameter for the exponential change parameter for speciation
+
+updateMuInitScale = 2,
+# Scale parameter for updating initial extinction rate for each process
+
+updateEventLocationScale = 0.1,
+# Scale parameter for updating LOCAL moves of events on the tree
+# This defines the width of the sliding window proposal
+
+updateEventRateScale = 4,
+# Scale parameter (proportional shrinking/expanding) for updating
+# the rate parameter of the Poisson process
+
+# OPERATORS: MCMC MOVE FREQUENCIES
+
+updateRateEventNumber = 1,
+# Relative frequency of MCMC moves that change the number of events
+
+updateRateEventPosition = 0.25,
+# Relative frequency of MCMC moves that change the location of an event on the
+# tree
+
+updateRateEventRate = 1,
+# Relative frequency of MCMC moves that change the rate at which events occur
+
+updateRateLambda0 = 1,
+# Relative frequency of MCMC moves that change the initial speciation rate
+# associated with an event
+
+updateRateLambdaShift = 1,
+# Relative frequency of MCMC moves that change the exponential shift parameter
+# of the speciation rate associated with an event
+
+updateRateMu0 = 1,
+# Relative frequency of MCMC moves that change the extinction rate for a given
+# event
+
+updateRateLambdaTimeMode = 0,
+# Relative frequency of MCMC moves that flip the time mode
+# (time-constant <=> time-variable)
+
+localGlobalMoveRatio = 10,
+# Ratio of local to global moves of events
+
+# INITIAL PARAMETER VALUES
+
+lambdaInit0 = 0.032,
+# Initial speciation rate (at the root of the tree)
+
+lambdaShift0 = 0,
+# Initial shift parameter for the root process
+
+muInit0 = 0.005,
+# Initial value of extinction (at the root)
+
+initialNumberEvents = 0,
+# Initial number of non-root processes
+
+# METROPOLIS COUPLED MCMC
+
+numberOfChains = 1,
+# Number of Markov chains to run
+
+deltaT = 0.01,
+# Temperature increment parameter. This value should be > 0
+# The temperature for the i-th chain is computed as 1 / [1 + deltaT * (i - 1)]
+
+swapPeriod = 1000,
+# Number of generations in which to propose a chain swap
+
+chainSwapFileName = 'chain_swap.txt',
+# File name in which to output data about each chain swap proposal.
+# The format of each line is [generation],[rank_1],[rank_2],[swap_accepted]
+# where [generation] is the generation in which the swap proposal was made,
+# [rank_1] and [rank_2] are the chains that were chosen, and [swap_accepted] is
+# whether the swap was made. The cold chain has a rank of 1.
+
+# NUMERICAL AND OTHER PARAMETERS
+
+minCladeSizeForShift = 3,
+# Allows you to constrain location of possible rate-change events to occur
+# only on branches with at least this many descendant tips. A value of 1
+# allows shifts to occur on all branches.
+
+segLength = 0.025,
+# Controls the "grain" of the likelihood calculations. Approximates the
+# continuous-time change in diversification rates by breaking each branch into
+# a constant-rate diversification segments, with each segment given a length
+# determined by segLength. segLength is in units of the root-to-tip distance of
+# the tree. So, if the segLength parameter is 0.01, and the crown age of your
+# tree is 50, the "step size" of the constant rate approximation will be 0.5.
+# If the value is greater than the branch length (e.g., you have a branch of
+# length < 0.5 in the preceding example) BAMM will not break the branch into
+# segments but use the mean rate across the entire branch.
+
+outName = x)
+  }
+
+bammcontrolfile <- list()
+for (x in name.passerine.tree) {
+  bammcontrolfile[x] <- paste("data/bamm_files/control_", x, ".txt", sep="")
+}
+
+# Now writing control parameters to file
+
+for (x in name.passerine.tree) {generateControlFile(file = bammcontrolfile[[x]], type = "diversification", params = params[[x]])}
+```
+
+BAMM can be run through the terminal through the following syntax: ``bamm -c control_tree_xxxx.txt``. To generate these commands we can use a loop function, from which we get:
+
+```{r}
+bamm.commands <- list()
+for (x in name.passerine.tree) {
+  bamm.commands[x] <- paste("bamm -c control_", x, ".txt", sep="")
+}
+```
 
 > cool
-
-###MCC Tree
-
-```{r}
-is.ultrametric(mcc.trees.eric[["tree_2241"]])
-is.binary.tree(mcc.trees.eric[["tree_2241"]])
-# Now to check min branch length:
-min(mcc.trees.eric[["tree_2241"]]$edge.length)
-```
-
-Create a control file with same parameters as the consensus tree:
-
-```{r}
-priors <- setBAMMpriors(mcc.trees.eric[["tree_2241"]], outfile = NULL)
-write.tree(mcc.trees.eric[["tree_2241"]], 'mcc_eric_2241.tre')
-
-# Here is a block of parameters for the control file
-params = list(treefile = 'mcc_eric_2241.tre',
-            sampleFromPriorOnly = 0,
-            runMCMC = 1,
-            initializeModel = 1,
-            useGlobalSamplingProbability = 1,
-            globalSamplingFraction = 0.67,
-            updateLambdaInitScale = 2.0,
-            updateMuInitScale = 2.0,
-            updateLambdaShiftScale = 0.05,
-            updateEventRateScale = 4.0,
-            localGlobalMoveRatio = 10.0,
-            modeltype = 'speciationextinction', #rather than trait
-
-            mcmcWriteFreq = '100000', #turns to one
-            eventDataWriteFreq = '100000',
-            printFreq = '100000',
-            acceptanceResetFreq = '100000',
-
-            lambdaInitPrior = as.numeric(priors['lambdaInitPrior']),
-            lambdaShiftPrior = '0.05',
-            muInitPrior = as.numeric(priors['muInitPrior']),
-
-            numberOfGenerations = '150000000', #Command line runs 150 million times
-            runMCMC = 1,
-            overwrite = 1,
-            seed = -1,
-            expectedNumberOfShifts = '24', #Huang and Rabosky had 40 with a full bird tree of 6,670 species. So with 4000 we would expect slightly less
-
-            lambdaIsTimeVariablePrior = '0',
-            updateRateLambdaShift = '1',
-            updateRateEventPosition = '1',
-            numberOfChains = '1',
-            outName = 'consensus.eric')
-
-bammcontrolfile <- paste("control_", "mcc_eric_2241", ".txt", sep="")
-
-# Now writing control parameters to file
-generateControlFile(file = bammcontrolfile, type = "diversification",
-        params = params)
-```
 

### 2018-09-14 functions/Correlation_matrix.R

@@ -1,0 +1,14 @@
+#This function writes in the correlation value
+panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
+{
+  usr <- par("usr"); on.exit(par(usr))
+  par(usr = c(0, 1, 0, 1))
+  r <- abs(cor(x, y, use="pairwise.complete.obs"))
+  txt <- format(c(r, 0.123456789), digits=digits)[1]
+  txt <- paste0(prefix, txt)
+  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+  text(0.5, 0.5, txt, cex = cex.cor * r)
+}
+
+#Use this syntax to evaluate the correlation in your data. In this case we look at the 2nd, 4th, 5th, 6th, 7th and 8th columns of the mtcars data set
+#pairs(mtcars[c(2,4,5,6,7,8)], lower.panel=panel.smooth, upper.panel=panel.cor)

### 2018-09-14 functions/Vdodge_function.R

@@ -1,0 +1,127 @@
+# This code is from the Github page of Jared Lander: https://github.com/jaredlander/coefplot/blob/master/R/position.r
+
+
+# Detect and prevent collisions.
+# Powers dodging, stacking and filling.
+collidev <- function(data, height = NULL, name, strategy, check.height = TRUE) {
+  # Determine height
+  if (!is.null(height)) {
+    # height set manually
+    if (!(all(c("ymin", "ymax") %in% names(data)))) {
+      data$ymin <- data$y - height / 2
+      data$ymax <- data$y + height / 2
+    }
+  } else {
+    if (!(all(c("ymin", "ymax") %in% names(data)))) {
+      data$ymin <- data$y
+      data$ymax <- data$y
+    }
+    
+    # height determined from data, must be floating point constant
+    heights <- unique(data$ymax - data$ymin)
+    heights <- heights[!is.na(heights)]
+    height <- heights[1]
+  }
+  
+  # Reorder by x position, relying on stable sort to preserve existing
+  # ordering, which may be by group or order.
+  data <- data[order(data$ymin), ]
+  
+  # Check for overlap
+  intervals <- as.numeric(t(unique(data[c("ymin", "ymax")])))
+  intervals <- intervals[!is.na(intervals)]
+  
+  if (length(unique(intervals)) > 1 & any(diff(scale(intervals)) < -1e-6)) {
+    warning(name, " requires non-overlapping y intervals", call. = FALSE)
+    # This is where the algorithm from [L. Wilkinson. Dot plots.
+    # The American Statistician, 1999.] should be used
+  }
+  
+  if (!is.null(data$xmax)) {
+    plyr::ddply(data, "ymin", strategy, height = height)
+  } else if (!is.null(data$x)) {
+    data$xmax <- data$x
+    data <- plyr::ddply(data, "ymin", strategy, height = height)
+    data$x <- data$xmax
+    data
+  } else {
+    stop("Neither x nor xmax defined")
+  }
+}
+
+# Stack overlapping intervals.
+# Assumes that each set has the same horizontal position
+pos_stackv <- function(df, height) {
+  if (nrow(df) == 1) return(df)
+  
+  n <- nrow(df) + 1
+  x <- ifelse(is.na(df$x), 0, df$x)
+  if (all(is.na(df$y))) {
+    heights <- rep(NA, n)
+  } else {
[…]

### 2018-09-14 functions/diversification_rate_calculation.R

@@ -1,0 +1,64 @@
+#Equal splits statistic calculation ("DR Statistic")
+calculate.log.es <- function(phy) {
+rootnode <- length(phy$tip.label) + 1
+es <- numeric(length(phy$tip.label))
+for (i in 1:length(es)){
+  node <- i
+  index <- 1
+  qx <- 0
+  while (node != rootnode){
+    el <- phy$edge.length[phy$edge[,2] == node]
+    node <- phy$edge[,1][phy$edge[,2] == node]			
+    qx <- qx + el* (1 / 2^(index-1))			
+    index <- index + 1
+  }
+  es[i] <- 1/qx
+	
+names(es) <- phy$tip.label
+}
+
+es <- log(es[phy$tip.label])
+result <- as.vector(list(es))
+names(result) <- "es"
+return(result)
+}
+
+# node depth statistic calculation
+calculate.nd <- function(phy) {
+rootnode <- length(phy$tip.label) + 1
+nd <- numeric(length(phy$tip.label)) # Number of tips
+for (i in 1:length(nd)){ # For each tip
+  node <- i
+  nodecount <- 0
+  while (node != rootnode){
+    node <- phy$edge[,1][phy$edge[,2] == node] # Work back a node	
+    nodecount <- nodecount + 1
+  }
+  nd[i] <- nodecount/max(branching.times(phy))
+names(nd) <- phy$tip.label
+}
+result <- as.vector(list(nd))
+names(result) <- "nd"
+return(result)
+}
+
+#terminal branch statistic 
+calculate.tb <- function(phy) {
+  
+  require(ape)
+  require(mvtnorm)
+  
+  # Calculate terminal edge lengths
+  n <- length(phy$tip.label)
+  # based on post on Liam Revell's blog:
+  invis <- setNames(phy$edge.length[sapply(1:n, function(x,y) which(y==x), y=phy$edge[,2])], phy$tip.label)
+  tb <- 1/invis
+  
+  tb <- log(tb[phy$tip.label]) # log transform
+  names(tb) <- phy$tip.label
+  
+  result <- as.vector(list(tb))
+  names(result) <- "tb"
+  return(result)
+  
+}

### 2018-09-14 functions/gheatmap2.R

@@ -1,0 +1,68 @@
+gheatmap2<-function (p, data, offset = 0, width = 0.1, low = "green", high = "red", 
+                    border_color = NULL, colnames = TRUE, colnames_position = "bottom", 
+                    colnames_angle = 0, colnames_level = NULL, colnames_offset_x = 0, 
+                    colnames_offset_y = 0, font.size = 4, hjust = 0.5) 
+{
+#the function is modified from ggtree::gheatmap
+  require(magrittr)
+  require(tidyr)
+  colnames_position %<>% match.arg(c("bottom", "top"))
+  variable <- value <- lab <- y <- NULL
+  width <- width * (p$data$x %>% range %>% diff)/ncol(data)
+  isTip <- x <- y <- variable <- value <- from <- to <- NULL
+  df <- p$data
+  df <- df[df$isTip, ]
+  start <- max(df$x) + offset
+  dd <- as.data.frame(data)
+  lab <- df$label[order(df$y)]
+  dd <- dd[lab, , drop = FALSE]
+  dd$y <- sort(df$y)
+  dd$lab <- lab
+  dd <- gather(dd, variable, value, -c(lab, y))
+  i <- which(dd$value == "")
+  if (length(i) > 0) {
+    dd$value[i] <- NA
+  }
+  if (is.null(colnames_level)) {
+    dd$variable <- factor(dd$variable, levels = colnames(data))
+  }
+  else {
+    dd$variable <- factor(dd$variable, levels = colnames_level)
+  }
+  V2 <- start + as.numeric(dd$variable) * width
+  mapping <- data.frame(from = dd$variable, to = V2)
+  mapping <- unique(mapping)
+  dd$x <- V2
+  dd$width <- width
+  if (is.null(border_color)) {
+    p2 <- p + geom_tile(data = dd, aes(x, y, fill = value), 
+                        width = width, inherit.aes = FALSE, size=0)
+  }
+  else {
+    p2 <- p + geom_tile(data = dd, aes(x, y, fill = value), color = border_color,
+                        width = width, inherit.aes = FALSE , size=0)
+  }
+  if (is(dd$value, "numeric")) {
+    p2 <- p2 + scale_fill_gradient(low = low, high = high, 
+                                   na.value = NA)
+  }
+  else {
+    p2 <- p2 + scale_fill_discrete(na.value = NA)
+  }
+  if (colnames) {
+    if (colnames_position == "bottom") {
+      y <- 0
+    }
+    else {
+      y <- max(p$data$y) + 1
[…]

### 2018-09-14 functions/symmetrise_scale.R

@@ -1,0 +1,28 @@
+#Symmetrise Scale when using facet_wrap function
+
+symmetrise_scale <- function(p, axis = "x") {
+  gb <- ggplot2::ggplot_build(p)
+  type <- switch(axis, "x" = "x.range", "y" = "y.range")
+  
+  fname <- setdiff(names(gb$layout$panel_layout), c("PANEL", "ROW", "COL", "SCALE_X", "SCALE_Y"))
+  lims <- do.call(cbind, lapply(gb$layout$panel_ranges, "[[", type))
+  facets <- gb$layout$panel_layout[, fname, drop = FALSE]
+  
+  
+  # dev version ggplot2_2.2.1.9000 breaks everything yet again
+  if (utils::packageVersion("ggplot2") >= "2.2.1.9") {
+    fname <- setdiff(names(gb$layout$layout), c("PANEL", "ROW", "COL", "SCALE_X", "SCALE_Y"))
+    lims <- do.call(cbind, lapply(gb$layout$panel_params, "[[", type))
+    facets <- gb$layout$layout[, fname, drop = FALSE]
+  }
+  
+  
+  lims2 <- as.vector(t(tcrossprod(apply(abs(lims), 2, max), c(-1, 1))))
+  
+  dummy <- stats::setNames(data.frame(facets[rep(seq_len(nrow(facets)), each = 2), ], lims2), c(fname, axis))
+  
+  switch(axis,
+         "x" = p + geom_blank(data = dummy, aes_string(x = "x", y = "Inf"), inherit.aes = FALSE),
+         "y" = p + geom_blank(data = dummy, aes_string(x = "Inf", y = "y"), inherit.aes = FALSE)
+  )
+}

### 2018-09-18 Sexual Selection and Extinction.Rmd
18/9 Update with taxa match

@@
 
+###Time-constant
+
+We also ran a time-constant MCC of BAMM
+```{r, eval=FALSE}
+#Read in the tree and MCMC to check for convergence 
+MCC.BAMM.tree  <- read.tree("data/BAMM_MCC/MCC.passerine.tre") #Same as other MCC tree already loaded
+MCC.BAMM.Constant.mcmc <- read.csv( "data/BAMM_MCC/TimeConstant/tree_MCC_Constant_mcmc_out.txt" , stringsAsFactors=F)
+
+#Plot of convergence can be generated by:
+#plot(MCC.BAMM.mcmc$logLik ~ MCC.BAMM.mcmc$generation)
+
+#Looks like it has converged so let's discard burn in: 
+burnstart <- floor(0.1 * nrow(MCC.BAMM.Constant.mcmc))
+postburn <- MCC.BAMM.Constant.mcmc[burnstart:nrow(MCC.BAMM.Constant.mcmc), ]
+
+#We can also check effective population sizes of the log-likelihhod and number of shift events in each sample
+#We want at least 200 (although that's on the low side)
+
+cbind(effectiveSize(postburn$N_shifts), effectiveSize(postburn$logLik)) %>% `colnames<-`(c("N_Shifts", "logLik")) %>% `rownames<-`("Effective Sample Size") %>% pander()
+
+# Read in Event Data
+MCC.BAMM.ED.Constant <- getEventData(MCC.BAMM.tree,  "data/BAMM_MCC/TimeConstant/tree_MCC_Constant_event_data.txt", burnin=0.1, nsamples=1000)
+saveRDS(MCC.BAMM.ED.Constant, 'data/MCC.BAMM.ED.Constant.rds')
+#From the Event Data we can extract
+
+css2 <- credibleShiftSet(MCC.BAMM.ED.Constant, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95)
+
+#Now we obtain the number of distinct shifts: (Out of 1000 samples this is super high, essentially each one is distinct)
+
+css2$number.distinct
+
+plotPrior(MCC.BAMM.Constant.mcmc, expectedNumberOfShifts=100)
+
+MCC.BAMM.df.Constant <- BAMM.extraction.function(MCC.BAMM.ED.Constant)
+
+MCC.BAMM.df.Constant %>% ggplot(aes(x = mean.lambda, y = MCC.BAMM.df$mean.lambda))+
+  geom_point()
+cor(MCC.BAMM.df.Constant$mean.lambda, MCC.BAMM.df$mean.lambda)
+```
+
 ##PGLS Models
@@
 
-BAMM.lambda.plot <-BAMM.lambda.pgls.summary %>% filter(Parameter != "(Intercept)") %>%
+  # for (x in unique(BAMM.lambda.pgls.summary$Parameter)[2:7]){
+  #   filter(Parameter == x & between(Estimate, left = as.numeric(hpd.Lambda.top[1,x]), right = as.numeric(hpd.Lambda.top[2,x])))
+  #   }
+
+remove_outliers <- function(x, na.rm = TRUE, ...) {
+  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
+  H <- 1.5 * IQR(x, na.rm = na.rm)
+  y <- x
+  y[x < (qnt[1] - H)] <- NA
+  y[x > (qnt[2] + H)] <- NA
+  y
+}
+
+BAMM.lambda.pgls.summary.RO <- dcast(BAMM.lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
+BAMM.lambda.pgls.summary.RO$Estimate <- NULL
+BAMM.lambda.pgls.summary.RO <- sapply(BAMM.lambda.pgls.summary.RO, function(x) {
+  remove_outliers(x, na.rm = T)})
+BAMM.lambda.pgls.summary.RO <-melt(BAMM.lambda.pgls.summary.RO) %>% na.omit()
+BAMM.lambda.pgls.summary.RO$Var1 <- NULL
+colnames(BAMM.lambda.pgls.summary.RO) <- c("Parameter", "Estimate")
+
+BAMM.lambda.plot <- BAMM.lambda.pgls.summary.RO %>%
   ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
@@
 
-BAMM.mu.plot <-BAMM.mu.pgls.summary %>% filter(Parameter != "(Intercept)") %>%
+#Get rid of outliers
+BAMM.mu.pgls.summary.RO <- dcast(BAMM.mu.pgls.summary %>% select(Parameter,Estimate) %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
+BAMM.mu.pgls.summary.RO$Estimate <- NULL
+BAMM.mu.pgls.summary.RO <- sapply(BAMM.mu.pgls.summary.RO, function(x) {
+  remove_outliers(x, na.rm = T)})
+BAMM.mu.pgls.summary.RO <-melt(BAMM.mu.pgls.summary.RO) %>% na.omit()
+BAMM.mu.pgls.summary.RO$Var1 <- NULL
+colnames(BAMM.mu.pgls.summary.RO) <- c("Parameter", "Estimate")
+
+BAMM.mu.plot <-BAMM.mu.pgls.summary.RO %>%
   ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
@@
 #Run a corPagel model to estimate lambda for DR
-MCC.Lambda.Armenta.global <- gls(log(mean.lambda) ~ Colour.discriminability
+MCC.Lambda.Armenta.global <- gls(mean.lambda ~ Colour.discriminability
                          + log(range.size.m2)
@@
                          + Colour.discriminability*NPP,
-                weights = ~ I(1/log(var.lambda)),
+                weights = ~ sqrt(var.lambda),
                 correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), 
@@
 #Run a corPagel model to estimate lambda for DR
-MCC.Extinction.Armenta.global <- gls(log(mean.mu) ~ Colour.discriminability
+MCC.Extinction.Armenta.global <- gls(mean.mu ~ Colour.discriminability
                          + log(range.size.m2)
@@
                          + Colour.discriminability*NPP,
-                weights = ~ I(1/log(var.mu)),
+                weights = ~ sqrt(var.mu),
                 correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), 
@@
 saveRDS(Armenta.MCC.ND.top, 'data/Armenta.MCC.ND.top.rds')
-
-Armenta.data$Colour.discriminability <- abs(Armenta.data$Colour.discriminability)
 
@@
 
-Armenta.MCC.Lambda.top <- gls(log(mean.lambda) ~ Colour.discriminability
+Armenta.MCC.Lambda.top <- gls(mean.lambda ~ Colour.discriminability
                       + log(range.size.m2)
@@
                       + NPP,
-                weights = ~ I(1/log(var.lambda)),
+                weights = ~ sqrt(var.lambda),
                 correlation = corBrownian(phy = pruned.MCC.Armenta.tree), 
@@
 
-Armenta.MCC.Mu.top <- gls(log(mean.mu) ~ Colour.discriminability
+Armenta.MCC.Mu.top <- gls(mean.mu ~ Colour.discriminability
                       + log(range.size.m2)
@@
                       + NPP,
-                weights = ~ I(1/log(var.mu)),
+                weights = ~ sqrt(var.mu),
                 correlation = corBrownian(phy = pruned.MCC.Armenta.tree), 
@@
 Armenta.BAMM.lambda.pgls.models <- mcmapply(function(x,y) {
-  gls(log(mean.lambda) ~ Colour.discriminability 
+  gls(mean.lambda ~ Colour.discriminability 
          + log(range.size.m2)
@@
          + NPP,
-    weights = ~ I(1/log(var.lambda)),      
+    weights = ~ sqrt(var.lambda),      
     corBrownian(phy = y), #lambda = 1
@@
 Armenta.BAMM.mu.pgls.models <- mcmapply(function(x,y) {
-  gls(log(mean.mu) ~ Colour.discriminability
+  gls(mean.mu ~ Colour.discriminability
          + log(range.size.m2)
@@
          + NPP,
-    weights = ~ I(1/log(var.mu)),      
+    weights = ~ sqrt(var.mu),      
     corBrownian(phy = y), #lambda = 1
@@
 for(x in unique(Armenta.ND.pgls.summary$Parameter)) {
-Armenta.hpd.ND.top[[x]] = hdi(Armenta.ND.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate)Colour.discriminability
+Armenta.hpd.ND.top[[x]] = hdi(Armenta.ND.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
 }
-Armenta.hpd.ND.top <- bind_rows(Armenta.hpd.ND.top) %>% `rownames<-`(c("Lower", "Upper")Colour.discriminability
+Armenta.hpd.ND.top <- bind_rows(Armenta.hpd.ND.top) %>% `rownames<-`(c("Lower", "Upper"))
 
@@
 #Run a corPagel model to estimate lambda for DR
-MCC.Lambda.Subset.global <- gls(log(mean.lambda) ~ Sexual_selection_ppca
+MCC.Lambda.Subset.global <- gls(mean.lambda ~ Sexual_selection_ppca
                          + log(range.size.m2)
@@
                          + Sexual_selection_ppca*NPP,
-                weights = ~ I(1/log(var.lambda)),
+                weights = ~ sqrt(var.lambda),
                 correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), 
@@
 #Run a corPagel model to estimate lambda for DR
-MCC.Extinction.Subset.global <- gls(log(mean.mu) ~ Sexual_selection_ppca
+MCC.Extinction.Subset.global <- gls(mean.mu ~ Sexual_selection_ppca
                          + log(range.size.m2)
@@
                          + Sexual_selection_ppca*NPP,
-                weights = ~ I(1/log(var.mu)),
+                weights = ~ sqrt(var.mu),
                 correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), 
@@
 
-Subset.MCC.Lambda.top <- gls(log(mean.lambda) ~ Sexual_selection_ppca
+Subset.MCC.Lambda.top <- gls(mean.lambda ~ Sexual_selection_ppca
                       + log(range.size.m2)
@@
                       + NPP,
-                weights = ~ I(1/log(var.lambda)),
+                weights = ~ sqrt(var.lambda),
                 correlation = corBrownian(phy = pruned.MCC.Subset.tree), 
@@
 
-Subset.MCC.Mu.top <- gls(log(mean.mu) ~ Sexual_selection_ppca
+Subset.MCC.Mu.top <- gls(mean.mu ~ Sexual_selection_ppca
                       + log(range.size.m2)
@@
                       + NPP,
-                weights = ~ I(1/log(var.mu)),
+                weights = ~ sqrt(var.mu),
                 correlation = corBrownian(phy = pruned.MCC.Subset.tree), 
@@
 Subset.BAMM.lambda.pgls.models <- mcmapply(function(x,y) {
-  gls(log(mean.lambda) ~ Sexual_selection_ppca 
+  gls(mean.lambda ~ Sexual_selection_ppca 
          + log(range.size.m2)
@@
 Subset.BAMM.mu.pgls.models <- mcmapply(function(x,y) {
-  gls(log(mean.mu) ~ Sexual_selection_ppca
+  gls(mean.mu ~ Sexual_selection_ppca
          + log(range.size.m2)
@@
 
-Subset.Lambda.plot <-Subset.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+Subset.Lambda.pgls.summary.RO <- dcast(Subset.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
+Subset.Lambda.pgls.summary.RO$Estimate <- NULL
+Subset.Lambda.pgls.summary.RO <- sapply(Subset.Lambda.pgls.summary.RO, function(x) {
+  remove_outliers(x, na.rm = T)})
+Subset.Lambda.pgls.summary.RO <-melt(Subset.Lambda.pgls.summary.RO) %>% na.omit()
+Subset.Lambda.pgls.summary.RO$Var1 <- NULL
+colnames(Subset.Lambda.pgls.summary.RO) <- c("Parameter", "Estimate")
+
+Subset.Lambda.plot <-Subset.Lambda.pgls.summary.RO %>% 
   ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
@@
 
-Subset.Mu.plot <-Subset.Mu.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+Subset.Mu.pgls.summary.RO <- dcast(Subset.Mu.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
+Subset.Mu.pgls.summary.RO$Estimate <- NULL
+Subset.Mu.pgls.summary.RO <- sapply(Subset.Mu.pgls.summary.RO, function(x) {
+  remove_outliers(x, na.rm = T)})
+Subset.Mu.pgls.summary.RO <-melt(Subset.Mu.pgls.summary.RO) %>% na.omit()
+Subset.Mu.pgls.summary.RO$Var1 <- NULL
+colnames(Subset.Mu.pgls.summary.RO) <- c("Parameter", "Estimate")
+
+Subset.Mu.plot <-Subset.Mu.pgls.summary.RO %>% 
   ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
@@
 
-```{r, fig.width = 8, fig.height = 6, eval = FALSE}
+```{r, fig.width = 8, fig.height = 6}
+BAMM.df <- readRDS('data/BAMM.df.rds')
 BAMM.Rates.100Trees <- do.call(rbind, BAMM.df)
-BAMM.Rates.100Trees$mean.lambda <- log(BAMM.Rates.100Trees$mean.lambda)
+BAMM.Rates.100Trees$mean.lambda <- BAMM.Rates.100Trees$mean.lambda
 BAMM.Lambda.100Trees <- ggiraphExtra::summarySE(BAMM.Rates.100Trees, measurevar = "mean.lambda", groupvars = "TipLabel")
@@
 
-mean(DR.Lambda.100Trees$sd)
-mean(BAMM.Lambda.100Trees$sd)
+# mean(DR.Lambda.100Trees$sd)
+# mean(BAMM.Lambda.100Trees$sd)
 ```
@@
 
-```{r, eval = FALSE}
+```{r}
 MCC.BAMM.model.data <- readRDS('data/MCC.BAMM.model.data.rds')
@@
   geom_point()+
-  scale_y_continuous(trans = "log")+
   theme_minimal()
@@
 
-```{r, eval = FALSE}
-MCC.BAMM.df$CV.lambda <- (sqrt(MCC.BAMM.df$var.lambda) / MCC.BAMM.df$mean.lambda)*100
+```{r}
+MCC.BAMM.df$CV.lambda <- sqrt(exp(MCC.BAMM.df$var.lambda) - 1)*100
 mean(MCC.BAMM.df$CV.lambda)
 min(MCC.BAMM.df$CV.lambda)
+median(MCC.BAMM.df$CV.lambda)
 max(MCC.BAMM.df$CV.lambda)
 
-MCC.BAMM.df %>% ggplot(aes(x = mean.lambda, y = CV.lambda)) +
-                         geom_point()+
-  scale_y_log10()+
-  scale_x_log10()
+#saveRDS(MCC.BAMM.df, 'data/MCC.BAMM.df.rds')
+
+MCC.BAMM.df %>% ggplot(aes(y = CV.lambda)) +
+  geom_boxplot(outlier.shape = NA)+
+  scale_y_continuous(trans = "log")
+```
+
+What if we run the subset df qwith SDi
+```{r}
+#Run model for DR
+Subset.MCC.DR.SDi <- gls(MCC.DR ~ SDi
+                         + log(range.size.m2)
+                         + bioclim4 #Seasonal variation
+                         + residuals.PC1 #Spatial variation
+                         + PC1.LIG #Long-term climate variation
+                         + NPP,
+                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = FALSE), 
+                data = SS.subset, 
+                method = "REML")
+
+summary(Subset.MCC.DR.SDi)
 ```

### 2018-09-24 Manuscript.Rmd
24/9 Update (first manuscript draft, yay!)

---
title: "Sexual selection, environmental variation, range size and speciation in passerine birds"
author: 
- Justin G. Cally*\textsuperscript{§}, Iliana Medina Guzmán\textsuperscript{§}, Luke Holman\textsuperscript{§} and Devi Stuart-Fox\textsuperscript{§}
- 
- '*\textit{justin.g.cally@gmail.com}'
output:
  pdf_document: default
  html_document: default
  word_document:
      pandoc_args: ["--smart"]
bibliography: References/SS_Extinction.bib
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
  \rhead[]{Sexual Selection and Extinction} \usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot}
  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\doublespacing \usepackage[font={small}]{caption}
csl: References/AmNat.csl
abstract: Sexual selection may influence patterns of diversification. Yet their relationship remains unclear with controversy on whether sexual selection leads to a net increase or decrease in speciation rate. Sexual selection is predicted to increase speciation rate as it can facilitate adaptation to new environments by elevating the average fitness of a population. Additionally, sexual selection may work alongide divergent natural selection and swiflty promote reproductive isolation. Here we use sexual dichromatism as a proxy for sexual selection across 5,812 species of passerine birds and a multivariate male-bias measure of sexual selection (sexual dimorphism, polygyny and paternal care) across 2,465 species to investigate the association of sexual selection with speciation and extinction. Importantly, we test whether the effects of sexual selection on speciation are dependent upon ecological pressures such as temperature variability across time and space, primary productivity and range size. Our results from three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) find that male-bias sexual selection but not sexual dichromatism predicts speciation. Additionally, we find that birds with smaller ranges have higher speciation rates; however the causality of this finding remains inconclusive.
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}

\newpage
#Introduction

**Process of speciation**: Selection pressures drive speciation. As populations diverge they accumulate genetic and phenotypic changes that ultimately erect barriers against gene flow [@Presgraves_2010; *but see* @Shaw_2011]. In this view populations will speciate and adapt more readily to to new divergent environments if the selection pressures and evolvability of the traits under selection are high. Hence, the rate at which speciation occurs is likely dependent on the both the populations adaptive potential/fitness and the evolvability of phenotypic traits such as morphology or chromatism [@Rabosky_morph_2013; @Hugall_2012]. Sexual selection is an evolutionary force likely to affect fitness components as well as the evolvability of morphological and chromatic traits such as ornaments involved in mating displays or weapons used in male-male competition. It is therefore hypothesised that sexual selection is a key driver of diversification through shifting the rates of adaptation, speciation and extinction.

**Sexual selection and fitness**: Sexual selection is applied by one sex on the other - more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013]. Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations, [@Agrawal_2001; @Siller_2001] and fixate beneficial mutations [@Whitlock_2000]. Thus, sexual selection is thought to be a key process in mitigating extinction risk and facilitating adaptation. However, the beneficial aspects of sexual selection are dependent on the genotypes under sexual selection (e.g. those that increase male mating/fertilisation success) being genetically positively correlated with traits affecting fitness (e.g. female fecundity or survival). This positive correlation is unlikely to always be the case; regularly, sexual selection favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006]. For instance, in species with high levels of sexual selection mating success becomes more variable and uncertain, thus individuals must heavily invest to increase their chances of successfully reproducing. Through prioritising mating success a sex may invest in costly weaponry and sexual signals, while forgoing strategies that would improve offspring fitness (parental care). This is a potentially costly consequence that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].

**Sexual selection, heritability and mechanisms of speciation**: Alongside changes to fitness, sexual selection has been proposed to facilitate ecological diversification and speciation through increased evolvability of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982]. Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, and by itself sexual selection could lead to behavioural reproductive isolation [@Safran_2013]. But sexual selection is unlikely to be occurring without the effects of natural selection. In this situation, adaptation and speciation may be further enforced if female preference in mating signals are dually favoured by natural selection [@Maan_2008]. The extent to which sexually selected traits can evolve is dependent on the underlying genetic variation. Under sexual selection a wider phenotypic space may be available for a population to evolve into than natural selection acting alone as genetic variation can be maintained under sexual conflict arising from heightened sexual selection [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016]. Furthermore, the evolvability and transmission of fitness and sexually selected traits may vary due to different sex determination systems across taxa. In male homogametic species (ZZ/ZW or ZZ/ZO) --- such as birds --- males are anticipated to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage coloration [@Hastings_1994; @Reeve_2003]. We might therefore expect birds with higher opportunities for sexual selection to have increased evolvability and speciation rates, especially in environments where natural selection pressures are high and working concordantly to sexual selection.

**Environmental interaction**: Importantly, the effects of sexual selection, adaptation and speciation may depend on environmental conditions. Studies suggest that under stressful/changing environments sexual selection may have greater fitness benefits than under benign environments. In stressful environments, the positive correlation between male mating success and female fitness components improves, reducing the burden of negative pleiotropic effects [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017]. Conversely, under benign environments environmental stability ensures consistent selection, preferentially removing genetic variation at sexually concordant loci (i.e. loci where the optimum genotype is equivelant for both sexes). In these cases genetic variation preferentially remains at sexually antagonistic loci; causing displacement of male and female traits from their optimum. These predictions would ensue that populations subject to persistant environmental variability would have increased fitness and adaptive benefits under increased sexual selection; allowing them to adapt to novel environments [*reviewed in* @Candolin_2008]. Increased environmental variability in a species ecological and climatic niche may also work synergestically with natural selection to promote diverging phenotypic traits [@Safran_2013]. In certain circumstances sexual selection will reinforce diverging locally adaptated phenotypes through mate choice on phenotypes that advertise quality in a specific ecological context [@Maan_2011]. If mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to strengthen this positive feedback loop and promote population divergence [@Lande_1981]. Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signalling strength in stickbacks [@Wong_2007; @Candolin_2007]. In this opposing view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.

**Previous limitations**: The association between sexual selection and speciation has been investigated using many macroevolutionary approaches across various taxa. Importantly, macroevolutionary comparative studies can examine the association between sexual selection and extinction risk in more diverse taxon, over longer evolutionary time-scales and in more natural environments. A meta-analysis of 64 effect sizes from 20 papers found a small, significant association between sexual selecetion and speciation [@Kraaijeveld_2011]. However, variation in sexual selection proxies used and other methodological choices (species richness vs rate estimates) led to heterogeneity in effect sizes for the correlation between sexual selection and speciation. Given that species richness and speciation rate may be negatively correlated [@Rabosky_fish_2018] and the recent development and refinement of methodologies in obtaining tip-rate estimates of speciation and extinction [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting with methods that address several issues that have previously limited macroevolutionary comparative studies. Firstly, sexual selection strength in macroevolutionary studies are often estimated using measures of dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa, which arguably decreases the power to detect an association, especially if phylogenetic signal is strong. Secondlly, many studies use incomplete taxon sampling, with family-level analysis restricted to well sampled avian clades [@Seddon_2008; @Huang_2014]. Thirdly, until recently studies have rarely incorperated extensive bioclimatic predictors into large-scale comparative analyses [*sensu* @Jetz_env_2011; @Pigot_2016]. Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague the ability to draw widely accepted findings **I don't know if this is true**.

**Our study**: Because of the wealth of data available for passerine birds (Order: Passeriformes) we are able to conduct a large comparative analysis across nearly all species of passerines. Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset (n = 918) [@Armenta_2008]. Here we assess the relationship between sexual dichromatism, environmental variability and their interaction with three tip-rate measures of speciation and one measure of extinction ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$;) in 5,812 species of passerines. We assess the effects of sexual dichromatism measures based on RGB values alongside smaller datasets of sexual dichromatism from spectrophotometry measures (*n* = 581) and male-bias sexual selection measures (*n* = 2,465). Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is are dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species' range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).

#Materials and methods

We examined the effect of sexual selection on speciation and extinction rate in 97 % of passerines (*n* = 5,812 species; 58 % of all birds). Specifically, we (i) compiled datasets for sexual dichromatism/selection strength and environmental variability, (ii) obtained estimates of speciation and extinction rates across passerines, and (iii) conducted phylogenetic generalized least-squares (PGLS) regressions. Many of the processes undertaken in this study are documented with reproducible code in the electronic supplementary material (ESM).

##Compiling data for sexual selection and environmental stress

###Sexual dichromatism

We used sexual dichromatism as proxy of sexual selection strength. Male and female plumage scores have previously been measured across 5,983 species of passerines [@Dale_2015]. Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forhead, throat, upper breast and lower breast) from _Handbook of the Birds of the World_ (volumes 8–16) [@delhoyo2011]. The mean values of red, green and blue were calculated and provided male and female plumage scores. Here we use the absolute difference between males and female RGB scores as an estimate of sexual dichromatism. Additionally, we compiled a colour discriminability measure of sexual dichromatism from spectrophotometer measures [@Armenta_2008]. These measurements include varition in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from taxanomic samples (as opposed to colour drawings). However, this spectrophotometry data is not as extensive as the RGB measures; with only 581 passerine species available for this analysis. While there is a correlation between these two measures, there is residual variation ($r = 0.79$; **Figure S9**).

###Male-biased sexual selection

Sexual dichromatism has long been associated with levels of sexual selection [@Darwin_1871], with sexual dichromatism often used as a measure of sexual selection in avian taxa [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014]. However, sexual dichromatism is likely to account for only partial variation in sexual selection across taxa. Here we use an existing dataset [@Dale_2015] of male-bias sexual selection based on the first component of a phylogenetic principle component analysis (ppca) of three characteristics associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care). However, this measure of male-bias sexual selection is present for only 2,465 species and is weakly correlated with the absolute values for sexual dichromatism using RGB measures ($r = 0.34$; **Figure S10**).

###Environmental variables 
Using expert range maps [@birdlife_2017] we obtained estimates of species range size. Using these distributions we were able to obtain estimates for extensive climatic conditions that each species is exposed to. Because of taxanomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015]. From 1,000 random point samples in each range we extracted 19 present day bioclimatic variables (representing a variety of biologically relevent annual trends in temperature and precipitation) with 30 seconds (~1 km^2^) spatial resultion [@Fick_2017]. From these values we obtained means and standard deviations for each species. Using the same spatial sampling we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006]. Furthermore, to estimate variability in the energy available to species we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution. Estimates of NPP had 30 second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005]. Using this data, which we provide as a potentially useful data resource (see ESM), we generated five predictors of speciation associated with different patterns in environmental variability (see below).

##Estimating extinction and speciation.

Phylogenetic information was obtained from www.birdtree.org [@Jetz_2012]. We used a maximum clade credibility tree (MCC) from 2,500 samples of the posterior distribution that were subset to the passerine bird order (*n* = 5,965) as the main phylogentic tree in our comparative analysis. Additionally, a random draw of 100 full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used. These trees used a 'Hackett backbone' and were constructed using the a pure birth (Yule) model. We calculated three different tip-rate metrics of speciation and one of extinction across all trees.

Firstly, we obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny. Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [*e.g.*, @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each more basal node having the sum of lengths down-weighted. Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification [@Belmaker_2015; @Rabosky_diversification_2018]. Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events and $\lambda_{ND}$ is a measure of speciation across the root to tip path. These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE) models. Based on previous comparisons $\lambda_{DR}$ and $\lambda_{ND}$ are anticipated to achieve high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].

Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014]. The model uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate a probability distribution of evolutionary regimes with variable speciation and extinction rates. Importantly, these models provide tip-rate estimates of speciation and extinction rate. The parameters of the 100 BAMM runs are detailed in full in the ESM; briefly we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014]. Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPU's. We obtained effective sample sizes for log-likelihood and number of rate shifts over 200 for each run (Table S3, Table S4). Further details of BAMM parameters and output is available through ESM, with tip-rate means and variances included in the ESM. Additionally, given the variability in BAMM estimates we also provide some analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on an MCC from a genetic-only phylogeny across all birds [@Harvey_2017]. All analyses was conducted on log-rates.

##Statistical analysis

###Generating biologically relevent predictors for environmental variation

Using the extracted environmental variables from each species range size we developed biologically meaningful predictors of environmental variation/stress relating to (i) seasonal climate variation, (ii) spatial climate variation and (iii) long-term climate variation. To obtain seasonal climate variation we used mean values of temperature seasonality (BIO4) for each range. To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature (BIO4) and precipitation seasonality (BIO15). PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (**Table S1**). Given that species range is a potentially informative predictor of speciation and extinction we control for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; **Figure S1**). To obtain long-term variation in climates for each species range we take the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values. Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates  (**Table S2**, **Figure S2**). Despite similar theoretical underpinnings the predictors of environmental variability used in PGLS analyses are not strongly correlated (**Figure S3**). Details and code to generate these predictors can be found within the ESM.

###Phylogenetic camparative analysis

To test the association between speciation/extinction and sexual selection, environmental variability and its interaction we used phylogenetic least squares (PGLS) models in the ``nlme`` package [@nlme_2018]. We began by estimating the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using ``corPagel`` in the ``ape`` package [@ape_2004] on a MCC tree. The estimate of $\lambda$ was then fixed for model selection, which was conducted on the same MCC tree derived from 2,500 draws from the posterior distribution [@Jetz_2012]. Model selection used $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$ tip-rate estimates from the MCC (n = 5,812) as the response variable and only compared interaction terms. That is, we fixed the individual predictors of: sexual dichromatism/selection, log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation and NPP while comparing 32 models with different combinations of interactions (including none). Model selection was done in ``MuMIn`` using the ``dredge`` function [@MuMIn_2017]. Using the terms from the top-ranking model (lowest AICc) we ran the equivalent model in each of the 100 phylogenetic trees, using the unique response variables and phylogenetic tree correlation structure in each model. This method gave us the ability to present model estimates for an MCC alongside 100 random draws; acknowledging the potential role phylogenetic uncertainty has in tapering conclusions drawn from a random sample of the data. This method was conducted on models that used data from all available passerine species (*n* = 5,812). PGLS analyses were also applied to a subset dataset of passerine birds (*n* = 581), where sexual dichromatism measures were sourced from spectrophotometry measures [@Armenta_2008] and where male-bias sexual selection ppca was used as a predictor instead of sexual dichromatism (*n* = 2,465). Finally, using this data subsetted for species with measurements of male-bias sexual selection we conducted a phylogenetic path analysis using the ``phylopath`` ``R`` package [@Wouter_2018]. The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS. To minimise path complexity we use temperature seasonality (BIO4) as the single measure for environmental variability and use $\lambda_{DR}$ as the single measure of speciation. Further details of the path analysis including reasons for path directions can be found within the ESM along with all other analyses and the relevant ``R`` code to reproduce results.

#Results

```{r, echo=FALSE}
MCC.DR.top <- readRDS('data/MCC.DR.top.rds')
MCC.ND.top <- readRDS('data/MCC.ND.top.rds')
MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')

#hpds
hpd.DR.top <- readRDS('data/hpd.DR.top.rds')
hpd.ND.top <- readRDS('data/hpd.ND.top.rds')
Armenta.hpd.DR.top <- readRDS('data/Armenta.hpd.DR.top.rds')

#Male bias SS
Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')
Subset.MCC.ND.top <- readRDS('data/Subset.MCC.ND.top.rds')
Subset.MCC.Lambda.top <- readRDS('data/Subset.MCC.Lambda.top.rds')
Subset.hpd.DR.top <- readRDS('data/Subset.hpd.DR.top.rds')
Subset.hpd.ND.top <- readRDS('data/Subset.hpd.ND.top.rds')
Subset.hpd.Lambda.top <- readRDS('data/Subset.hpd.Lambda.top.rds')
MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')

#path analysis
best_model <- readRDS('data/path_model.rds')
```

##Effects of sexual dichromatism/selection on speciation are not dependent upon environmental variability

PGLS models with interaction terms between the respective measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in global models. However, in all model selection phases for different measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and different measures of sexual selection (RGB values, spectrophotometry and male-biased sexual selection) no interaction terms were significant or present in the top models ($\delta$ AICc > 4; **Table S5**, **Table S6**, **Table S10**, **Table S13**). Hence we find no evidence that the effect of sexual selection on speciation or extinction is dependent upon certain environmental conditions. Furthermore, we find no support for these individual bioclimatic factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) acting as a driver of speciation independently from sexual dichromatism/selection (\autoref{model_results}, **Figure S10**).

##No evidence that sexual dichromatism affects speciation

We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation. Using three response measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$), the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from a MCC tree. Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). PGLS analyses using subsetted data (n = 581), where sexual dichromatism was measured using spectrophotometry distance measures [@Armenta_2008] yielded results concordant to the full data set; that is, no association between sexual dichromatism and speciation or extinction (**Figure S10**). Our results from MCC models are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 100 random trees. Model estimates are normally distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, See HPD intervals in **Table S8**). For PGLS using the subsetted dataset where sexual dichromatism was determined with spectrophotmetry [@Armenta_2008] the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (`r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (**Table S12**).

##Male-bias sexual selection increases speciation

When we conducted PGLS analysis using data subsetted for values of sexual selection estimated from a phylogenetic PCA of sexual dimorphism, social polygyny and paternal care (*n* = 2,465) there was a significant positive association between sexual selection and $\lambda_{DR}$ ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b) and a non-significant positive association between sexual selection and $\lambda_{ND}$ ($\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b). However, no association was found between sexual selection and $\lambda_{BAMM}$ ($\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b). For PGLS models on 100 random trees, we find the distribution to be similar to the estimates when an MCC tree is used. Among the 100 tress we find a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a positive (but less so) association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`). Complete HPD intervals for models using male-bias sexual selection ppca as a predictor can be found within **Table S15**.

##Species with smaller ranges have increased rates of speciation

Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation we find a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$. Model estimates from PGLS models using an MCC tree for determining tip-rates and correlation structures place this negative association as small but significant ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a). This association was also evident across the estimates from models using the 100 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a). Subsetted models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- show variable evidence that range size is negatively associated with speciation. Data subsetted based on the sexual selection ppca (*n* = 2,465) show that smaller range size significantly predicted increased $\lambda_{DR}$ (\autoref{model_results}b), yet this negative association was not significant for $\lambda_{ND}$ and near-zero for $\lambda_{BAMM}$. Alternatively, models using subsetted data (*n* = 581), where sexual dichromatism was measured using spectrophotometry distance measures [@Armenta_2008] gave non-significant estimates for the effect of range size on all measures of speciation (**Figure S10, Table S11, Table S2**). Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with highest power using near-complete taxon sampling (*n* = 5,812).

##Phylogenetic path analysis

Using a phylogenetic path analysis we found there to be multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; **Figure S13**). Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from sexual selection to sexual dichromatism was relatively weak given the expectations of dichromatism being a well utilised proxy of sexual selection ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`). Additionally, given the large direct effect of temperature seasonality on range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`) and our findings from the PGLS that range size is negatively associated with $\lambda_{DR}$, there is an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}).

##Extinction rate

We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). This null result was consistent for spectrophotometry measures of sexual dichromatism (**FigureS10, Table S11, Table S12**) and for the estimates of sexual selection using the sexual selection ppca (\autoref{model_results}, **Table S14, Table S15**).

##Variability across trees and methods

We tested our hypothesis that sexual selection is associated with speciation using several methods and across 100 samples of phylogentic trees drawn from the post burn-in posterior [@Jetz_2012]. The range of methods utilised were implemented to overcome uncertainty in the power and precision of various tip-rate estimates as well as the variation between trees. We found that that the tip-rate estimates across the 100 trees were quite variable for both $\lambda_{DR}$ and $\lambda_{BAMM}$ (**Figure S13**). Despite this variation, both methods provided similar estimates for speciation in each species for the mean values across the 100 trees ($r = 0.67$) and for the tip-rate estimates from the MCC ($r = 0.68$). However in comparison to $\lambda_{DR}$, the value of $\lambda_{BAMM}$ was the _mean_ drawn from a posterior distribution (*n* = 1,000) of BAMM generations, thus estimates of $\lambda_{BAMM}$ (and $\mu_{BAMM}$) have an added level of variation. To account for this variation, weights (using the inverse of the variance) were used for the PGLS models. From 1000 posterior samples of the MCC BAMM run the coefficient of variation (CV) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* **Figures S5 and S7**). Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (**Table S3 and S4**) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (n = 5,966 species) was high. This means that although the number of rate shifts reached convergence and is more precise ($median = 59$; **Figure S6**) the locations of the rate shifts (*i.e.* the credible shift set) across the tree is highly heterogeneous. The variability in the locations for the shift configurations is a likely source of uncertainty in downstream tip-rates used as the response variable in PGLS models. Notably the model estimate BAMM produced varied a lot between trees relative to the confidence interval of the MCC tree, with 95 % HPD interval being approximately 20-fold wider than the MCC 95 % CI for models using BAMM, whereas for $\lambda_{DR}$ and $\lambda_{ND}$ the HPD 95 % interval width was equal to or less than the MCC 95 % CI (\autoref{model_results}, **Table S9**).

![Model estimates for (**a**) PGLS analyses using sexual dichromatism measures (*n* = 5,812) and (**b**) PGLS analyses using a subsetted male-bias measure of sexual selection (*n* = 2,465). Both datasets were used for anlyses with three measures of speciation $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and one measure of extinction $\mu_{BAMM}$ as response variables. When using $\lambda_{DR}$ and $\lambda_{ND}$ as the response variable, they provide relatively congruent results; with a negative relationship between $\lambda_{DR/ND}$ and range size as well as a positive relationship between $\lambda_{DR/ND}$ and sexual selection (**b**). Alternatively, when using $\lambda_{BAMM}$ and $\mu_{BAMM}$ the model estimates are more variable across trees compared to the MCC estimate, which do not significantly differ from zero. The numerical values for the model estimates using the MCC tree and hpd of estimates from 100 random trees can be found within the ESM. Density curves are based on estimates from 100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC. For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to interperet the MCC 95 % CIs. \label{model_results}](Figures/ModelEstimates.pdf)

![Speciation rate ($\lambda_{DR}$) across all passerine birds (*n* = 5,965) with estimates of sexual dichromatism and range size available for 5,812 of them. Across these species there was a small but significant negative association between $\lambda_{DR}$ and log-range size but no significant association between $\lambda_{DR}$ and sexual dichromatism based on RGB measurments. $\lambda_{DR}$ are those from the MCC tree and images of birds are from the _Handbook of the Birds of the World_. Clockwise the species are: _Sporophila bouvronides_, _Geospiza magnirostris_, _Atlapetes leucopterus_, _Euphonia rufiventris_, _Euplectes franciscanus_,_Promerops gurneyi_, _Phainopepla nitens_, _Zosterops erythropleurus_, _Acrocephalus australis_, _Progne cryptoleuca_, _Mayrornis schistaceus_, _Paradisaea rubra_, _Pachycephala pectoralis_, _Malurus pulcherrimus_, _Cranioleuca curtata_, _Cercomacra manu_, _Muscisaxicola capistratus_, _Lepidothrix coeruleocapilla_. Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation and should not be interpereted. \label{Speciation_Figure}](Figures/DRSpeciationFigure.pdf)

![Path analysis of evolutionary and ecological variables. Arrows represent direct effects with numeric values corresponding to colours (blue = positive, red = negative). The numeric values are standardised regression slopes and the asterix indicates that the 95 % confidence intervals of this estimate do not overlap with zero. The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and male-bias sexual selection measures (*n* = 2,465). \label{Path_Analysis}](Figures/Path_Analysis.pdf)

\newpage

#Discussion

We found no evidence that sexual dichromatism alters the rate of speciation in passerine birds independently or via interactions with environmental variability. This result is consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and two measures of dichromatism: a near-complete dataset of 5,812 species using RGB scores and a smaller dataset of 581 species where measurements were obtained from spectrophotometry (allowing variation in the ultraviolet and bird-visible range to be detected). However, we did find that when using a male-bias measure of sexual selection --- instead of sexual dichromatism --- there was a positive association with speciation: with significance varying across different rate estimates. Our findings suggest that the components that were used in the ppca of male-bias sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care) are better predictors of speciation than sexual dichromatism and provide discussion points for the contentious topic of sexual selection proxies. No bioclimatic measures of environmental variability (temperature seasonality, long-term temperature variation and spatial temperature variation) predicted speciation, but range size was negatively associated with two measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$). This finding suggests that species with smaller range sizes have increased speciation rate. However, small range size may be a cause or effect of speciation rate and potentially even an artifact of taxonomic classification. Interestingly we found a discrepancy in the model results obtained from tip-rate statistics ($\lambda_{DR}$, $\lambda_{ND}$) and those using tip-rate estimates from a model based approach ($\lambda_{BAMM}$); whereby BAMM produced precise null results for associations between predictors and speciation as well as extinction.

Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015] we find that sexual dichromatism does not predict speciation rate as sexual selection does. The association between sexual selection and sexual dichromatism in birds is a Darwinian trope [@Darwin_1871] that allows it to be commonly used as a proxy for sexual selection in comparative studies [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014]. However, the presence of high sexual dichromatism in bird species may be also dependent on ecological selection pressures. Ecological pressures could drive sexual dichromatism in the absence of mate choice if sex-specific niches provide opportunity for the evolution of sex-limited traits [@Kottler_1980; @Slatkin_1984; @Shine_1989]. In damselflies, differences between the sexes in their exposure to UV radiation drives dichromatism between the sexes [@Cooper_2010; @Punzalan_2010]. Ecological impacts on sexual dichromatism are supported by our path analysis; which reveals that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation) albeit by a relatively small amount. Given that speciation rate increases with sexual selection strength but not sexual dichromatism we are led to conclude that sexual dichromatism is a seemingly poor proxy of sexual selection and that the noise associated with the proxy measurement of sexual selection cancels out the obtained signal; this is despite the high power of our comparative study. The usefulness of sexual dichromatism as a proxy of sexual selection may be dependent upon the metabolic pathways that give rise to plumage colouration; with not all plumage colouration reflecting mate quality. A recent meta-analysis found that converted carotenoids (but not those sourced from diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018]. Given the importance of honest signalling in sexual selection and the variability in colour production across birds, sexual dichromatism in many species may unreliable measure of mate quality for female birds and sexual selection for researchers.

In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we find male-bias sexual selection marginally increases speciation rate (at least for $\lambda_{DR}$, $\lambda_{ND}$) yet there is no direct evidence of this association being dependent on variable or stressful environments; albeit a phylogenetic path analysis shows an indirect effect of environmental variability (through range size). When evolving under stressful conditions, sexually antagonistic selection is hypothesised to be weakened relative to selection that is conchordant between the sexes [@Connallon_SA_2015; @Connallon_Hall_2016]. While these predictions are supported in several lab (experimental evolution) studies [@Long_2012; @Berger_2014; @Punzalan_2013], other studies have not found the association [@Delcourt_2009; @Martinossi-Allibert_Savković_2017; @Holman_Jacomb_2017]. Given that we did not detect an effect of environmental variability (by itself or as an interaction with sexual selection) we are left with several possible explanations. Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of envrionmental variability the species is evolving under. Specifically, sexual antagonism may be more easily purged in environments where directional environment pressures are maintained; whereas when environmental variability is cyclical (e.g. seasonal or glacial climate cycles) sexual antagonism can remain indefinately as a fitness burden [@Connallon_Hall_2016]. Secondly, the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation. Specifically, access to dietary resources, and the impacts of predation or parasitism are unaccounted for here: arguably these are key processes affecting sexual selection and speciation [*reviewed in*, @Maan_2011]. Thirdly, it is also likly that there is no effect; at least not in the species investigated here.

<!-- **Make it about evolvability** -->
<!-- The effects of sexual selection are complex and are expected to vary amongst taxa. Although birds have a sex determination system conducive to increased father-to-son heritability of male-specific fitness traits on the Z chromosome [@Hastings_1994; @Reeve_2003] there are counteravailing arguments suggesting that deleterious recessive alleles on the Z chromosome in males (ZZ) are less exposed to selection, thus limiting the evolvability of traits [@Charlesworth_1987; @Charlesworth_2018]. In a macro-evolutionary meta-analysis on sexual selection and speciation rate @Kraaijeveld_2011 found no significant taxon based differences across a diverse sample of vertebrates and invertebrates (fish, insects, birds, spiders, reptiles, mammals). Yet sexual selection in male homogametic taxa (birds) produced noticibly faster speciation rates than in male heterogametic taxa **Would be good to get raw data off them and do a pair-wise comparison**. The evolvability of the   -->

The most convincing finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates. But drawing conclusions from this finding is fraught with challenges. Challenges arise on determining the causality of range size and speciation; as arguably small range size could either be a cause or effect of speciation. High speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated taxa. Within Himalayan songbirds range expansions are prevented by the competition for resources by related species [@Price_2014]; however this limitation on range expansion is then expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011]. Thus, if range size is a cause of speciation we might expect it to be smaller but if it is a predictor of speciation it arguably should be larger; with greater opportunities for geographic barriers to form in larger ranges [@Rosenzweig_1995]. Unlike Himalayan songbirds, species undergoing adaptive radiation into newly formed ecological environments are unlikely to be limited by competition for resources from pre-existing taxa. Across islands we expect to see a correlation between speciation and small range size, not because of any inherent properties of the range size, rather that they reflect newly formed islands with empty niches **This seems like a very established idea but not sure whose it is**. Thus our findings may be heavily dependent on species whose range size is restricted to a single island within an archipelago (*e.g.*, *Geospiza* sp.; aka Darwin's finches). Birds with large range sizes are also predicted to have greater dispersal ability; in these cases individuals may be more able to overcome geographical barriers and maintain gene flow; thus lowering speciation in these wide-ranging species [@Birand_2012; @Claramunt_2012]. Several studies have used morphological measures of wing shape, a proxy for dispersal ability (Hand-wing index/ Kipp's distance), as a predictor of speciation and coexistence [*e.g.*, @Kennedy_2016; @Pigot_2016], but these measurements across passerines is not yet extensive **I think? ~ 1000**. One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification whereby splitting of species leads to increased recent phylogenetic branching as well as a more subdivided range **Has this argument been articulated?**. The role of range size in speciation may be further investigated through ancestral geographic range reconstruction [@Quintero_2015], as ancestral ranges may be beneficial in understanding modes of speciation and the causality of range size expansions and contractions [@Losos_2003].

Here we used BAMM to assess extinction rates across passerines and whether these rates were associated with sexual dichromatism/selection or various ecological variables. We found no associations in this respect despite the wealth of theoretical predictions on the effects of sexual selection on fitness. Extinction is notoriously difficult to obtain accurate estimates for in comparative studies, as it is a process that occurs over millennia, leaving scant traces. One exception to this comes from fossil ostracods, a group of marine invertebrates. In these species there is an extensive fossil record of extinct ostracods from the Late Cretaceous epoch (84–66 million years ago) as well as identifiable sexual dimorphism. Recently, @Martins_2018 found that species of ostracods with larger sexual dimorphism (representing larger investment by males in reproductive organs) had higher extinction rates; suggesting that sexual selection increases extinction risk. Other comparative studies have assessed the relationship between sexual selection and extinction in birds (@Doherty_2003; @Mitra_1996; @Morrow_2003) and mammals [@Bro-Jorgensen_2014; @Morrow_2004] without consensus. Yet these methods often rely on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List [*e.g.*, @Morrow_2003; @Bro-Jorgensen_2014; @Morrow_2004] and may be heavily biased to species under greater anthropogenic threat, culturally valued or well researched. While estimating extinction risk from the phylogeny may not come with these biases, they are riddled with others [*reviewed in*, @Pyron_2016]. BAMM has been subject to controversy in its ability to model rate shifts and extinction rates [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], and while several tip-rate estimates exist for speciation (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$) extinction estimates are not as easily obtained. Here we anticipated extinction rates would be underestimated because we used a phylogeny constructed with a birth-only (Yule) model [@Jetz_2012]. Given these underestimates of $\mu_{BAMM}$ it is highly likely we failed to detect true associations between sexual dichromatism/selection and extinction. The power of BAMM as a model based-approach in estimating $\lambda$ and $\mu$ is dependent size of the rate regimes. BAMM often produces homogeneous $\lambda_{BAMM}$ and $\mu_{BAMM}$ for clades with few rate shifts, whereas $\lambda_{DR}$ produces higher heterogeneity in tip-rates; arguably this could reflect increased power or sampling error [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018]. Thus we are somewhat uncertain in the validity of conclusions from $\mu_{BAMM}$ estimates as well as $\lambda_{BAMM}$ and $\lambda_{DR}$ given their discrepancy.

Improvements in estimating tip-rates of speciation in the last few years has allowed us to revisit the association between sexual selection and speciation. With high power and alongside extensive ecological predictors we find that male-bias measures of sexual selection better predict speciation than sexual dichromatism. If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold with sexual selection a potentially valuable utility in conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014]. Additionally, our finding that high speciation rate is associated with smaller range size has pronounced significance to the future survival of these rapidly speciating species in a world with increased habitat loss and anthropogenic stress. Range size is the best predictor of extinction risk [@Harris_2008] and the association found here ensues that many newly speciated clades may be at greater risk of extinction due to anthropogenic effects if not evolutionary or ecological effects.

<!-- Despite the strong predictions that sexual selection is more likely to increase speciation than not **cite all those paper**, there are countervailing arguments that sexual selection may dampen speciation rate. Studies have suggested that increased selection on males through sexual selection may elevate inter-population mating rates as males seek to overcome local gene flow between populations and impede speciation [@Parker_1998; @Servedio_2014]. **How?** Sexual selection may also lead to the evolution of traits that constrain further divergence or adaptation (i.e. lead to an evolutionary dead end) **CITE**. -->

\newpage
# References
<div id="refs"></div>

### 2018-09-24 Sexual Selection and Extinction.Rmd
24/9 Update (first manuscript draft, yay!)

@@
 library(tibble)
+library(stargazer)
+library(ggraph)
+#devtools::install_github("Ax3man/phylopath")
+library(phylopath)
 library("EBImage") # for images
@@
 ```
-**Figure S9:** Measures of sexual dichromatism using a more precise measure in spectrophotometry do not change the main patterns seen within the PGLS models analysed here.
+**Figure S10:** Measures of sexual dichromatism using a more precise measure in spectrophotometry do not change the main patterns seen within the PGLS models analysed here.
 This dataset is a subset of the complete dataset (n = 581), thus drawing conclusions for the other predictors (e.g.
@@
 ```
-**Figure S10:** The relationship between absolute sexual dichromatism and the sexual selection.
+**Figure S11:** The relationship between absolute sexual dichromatism and the sexual selection.
 Although positively correlated the distribution is very noisy with species having high sexual dichromatism and low sexual selection and visa versa.Colour.discriminability
@@
 pruned.MCC.Subset.tree <- drop.tip(MCC.passerine,MCC.passerine$tip.label[-match(SS.subset$TipLabel, MCC.passerine$tip.label)])
+
+saveRDS(pruned.MCC.Subset.tree, 'data/pruned.MCC.Subset.tree.rds')
 
@@
 ```
-**Figure S11:** Measures of sexual selection a phylogenetic pca of sexual dimorphism, social polygyny and paternal care provide some evidence to suggest sexual selection predicts extinction rate (DR and ND models).
+**Figure S12:** Measures of sexual selection a phylogenetic pca of sexual dimorphism, social polygyny and paternal care provide some evidence to suggest sexual selection predicts extinction rate (DR and ND models).
 This figure is seen in Figure 1 within the manuscript
@@
 
+##Phylogenetic Path Analysis
+
+We undertook a phylogenetic path analysis using a model of causal pathways *a priori*.
+To do this we use the ``phylopath`` function.
+Our path analysis is undertaken on the dataset subset to species with measures of male bias sexual selection.
+Additionallty, the path analysis is restricted to using one measure of speciation ($\lambda_{DR}$) and the MCC tree.
+Our reasoning for the causal links are described breifly below:
+
+1) **$\lambda_{DR}$** dependent upon **Sexual Dichromatism**: Sexual dichromatism would directly impact speciation if there is an association with the evolvability of chromatism and speciation, this may more rapidly lead to niche divergence or reproductive isolation.
+2) **$\lambda_{DR}$** dependent upon **Sexual Selection**: This was a major hypothesis in our study (see introduction).  
+3) **$\lambda_{DR}$** dependent upon **Temperature Seasonality**: More variable environments are broadly expected to impact the resources, niches and fitness of species; thus there is an expectation that more variable environments may impact speciation rate.
+4) **$\lambda_{DR}$** dependent upon **Range Size**: As previously seen in the PGLS models range size shows a negative correlation with speciation rate; expected reasons for this causal link include lower dispersal ability in smaller ranges correlating with lower gene flow and greater speciation.
+It may also reflect niche specialisation and thus in line with the previous point increase speciation and reduce gene flow.
+5) **Sexual Dichromatism** dependent upon **Sexual Selection**: Higher levels of sexual selection will promote showy males and drab females (sexual dichromatism).  
+6) **Sexual Dichromatism** dependent upon **Temperature Seasonality**: Ecological aspects may facilitate sexual dimorphism as niche partitioning between the sexes causes sex-limited traits and colouration that are independent of mate choice or male-male competition.
+7) **Sexual Selection** dependent upon **Temperature Seasonality**: Greater environmental variability/stress may impact the usefulness (fitness effects) of sexual selection (see introduction).  
+8) **Range Size** dependent upon **Sexual Selection**: Sexual selection may often depend on local adaptations and the alignment of sexually selected traits with locally valuable traits, range size may limit this leading to a negative correlation.
+Alternatively if sexual selection has added fitness and evolvability benefits there may be increased range sizes in sexually selected species.
+9) **Range Size** dependent upon **Temperature Seasonality**: Anticipated to be more of a correlation than causal link.
+Yet a speecies range size may be dependent upon how variable their environment is and the availability to local resources.
+
+```{r, eval = F}
+#Set rownames to match tree
+rownames(SS.subset) <- SS.subset$TipLabel
+pruned.MCC.Subset.tree <- readRDS('data/pruned.MCC.Subset.tree.rds')
+SS.subset$log.range.size <- log(SS.subset$range.size.m2)
+
+SS.subset2 <- SS.subset %>% rename(
+  DR = MCC.DR,
+  SD = SDi,
+  TS = bioclim4,
+  RS = log.range.size,
+  SS = Sexual_selection_ppca
+)
+
+models.Subset <- define_model_set(
+  one = c(DR ~ SD, 
+          DR ~ SS,
+          DR ~ TS,
+          DR ~ RS,
+          SD ~ SS,
+          SD ~ TS,
+          SS ~ TS,
+          RS ~ SS,
+          RS ~ TS)
+)
+
+result <- phylo_path(models.Subset, data = SS.subset2, tree = pruned.MCC.Subset.tree, model = 'lambda')
+
+#best_model <- best(result, boot = 500)
+set.seed(1)
+path.plot <- plot(x = best_model,
+     type = "color",
+     algorithm = 'gem',
+     manual_layout = NULL,
+     curvature = 0.1,
+     colors = c("#b2182b", "#2166ac"),
+     show.legend = F)
+
+saveRDS(best_model, 'data/path_model.rds')
+#Inspect the coefficients with their SE
+
+# coef_plot(best_model, error_bar = "ci", reverse_order = TRUE) + 
+#   ggplot2::coord_flip()+
+#   ggplot2::theme_minimal()
+
+# pdf("Figures/Path_Plot.pdf", width=8, height=8)
+# path.plot
+# dev.off()
+
+# path.plot
+```
+
+```{r, fig.width = 6, fig.height = 4, }
+best_model <- readRDS('data/path_model.rds')
+coef_plot(best_model, error_bar = "ci", reverse_order = TRUE) + 
+  ggplot2::coord_flip()+
+  ggplot2::theme_minimal()
+```
+
+**Figure S13:** Path analysis standardized regression coefficients vary across relationships.
+Error bars are derived from confidence intervals through 500 bootstrapped iterations.
+Paths with error bars not everlapping zero are presented with an asterix in Figure 3 within the manuscript.
+
 #Additional Figures and Tables
@@
 ```
-**Figure S12:**
-
-```{r, eval = FALSE}
+**Figure S14:**
+
+```{r, eval = FALSE, echo = F}
 all.trees.join <- left_join(DR.Lambda.100Trees %>% select(TipLabel, DR), BAMM.Rates.100Trees %>% select(mean.lambda, TipLabel))
@@
 ```
-**Figure S13:**
-
-```{r, eval = FALSE}
+**Figure S15:**
+
+```{r, eval = FALSE, echo = F}
 MCC.tree.join <- left_join(restricted.data %>% select(MCC.DR, TipLabel), MCC.BAMM.model.data %>% select(mean.lambda, TipLabel))
@@
 
-```{r}
+```{r, eval = F, echo = F}
 MCC.BAMM.df$CV.lambda <- sqrt(exp(MCC.BAMM.df$var.lambda) - 1)*100
@@
 What if we run the subset df qwith SDi
-```{r}
+```{r, eval = FALSE, echo = F}
 #Run model for DR
@@
 
-summary(Subset.MCC.DR.SDi)
+#stargazer(Subset.MCC.DR.SDi, type = "html")
 ```
@@
 ```{r}
-sessionInfo() %>% pander
+sessionInfo() %>% pander()
 ```

### 2018-09-24 Manuscript.Rmd
change intro

@@
 output:
-  pdf_document: default
-  html_document: default
   word_document:
       pandoc_args: ["--smart"]
+  html_document: default
+  pdf_document: default
 bibliography: References/SS_Extinction.bib
@@
 \newpage
-#Introduction
+#Introduction xxxxxx
 

### 2018-09-26 Manuscript.Rmd
Comments first draft IMG

@@
 output:
+  html_document: default
   word_document:
       pandoc_args: ["--smart"]
-  html_document: default
   pdf_document: default
@@
 In this view populations will speciate and adapt more readily to to new divergent environments if the selection pressures and evolvability of the traits under selection are high.
-Hence, the rate at which speciation occurs is likely dependent on the both the populations adaptive potential/fitness and the evolvability of phenotypic traits such as morphology or chromatism [@Rabosky_morph_2013; @Hugall_2012].
-Sexual selection is an evolutionary force likely to affect fitness components as well as the evolvability of morphological and chromatic traits such as ornaments involved in mating displays or weapons used in male-male competition.
+Hence, the rate at which speciation occurs is likely dependent on both the populations adaptive potential/fitness and the evolvability **mmm here it becomes unclear because you are suggesting the link between speciation and the evolvability of traits, but is not clear what you mean by evolvability: a.
+the potential to evolve (e.g.
+genetic architecture) b.
+the rate of trait evolution or c.
+the magnitud of the difference in a trait between populations.
+I would either define evolvability earlier or replace the word with something easier to grasp** of phenotypic traits such as morphology or chromatism **color?
+avoid jargon when possible** [@Rabosky_morph_2013; @Hugall_2012].
+Sexual selection is an evolutionary force likely to affect fitness components as well as the evolvability **here I would just use 'evolution' rather than evolvability** of morphological and chromatic traits such as ornaments involved in mating displays or weapons used in male-male competition.
 It is therefore hypothesised that sexual selection is a key driver of diversification through shifting the rates of adaptation, speciation and extinction.
 
-**Sexual selection and fitness**: Sexual selection is applied by one sex on the other - more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
-Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations, [@Agrawal_2001; @Siller_2001] and fixate beneficial mutations [@Whitlock_2000].
+**Sexual selection and fitness**: Sexual selection is applied by one sex on the other - more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and **the strength of sexual selection**can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
+Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations, [@Agrawal_2001; @Siller_2001] and fixate beneficial mutations in a population [@Whitlock_2000].
 Thus, sexual selection is thought to be a key process in mitigating extinction risk and facilitating adaptation.
-However, the beneficial aspects of sexual selection are dependent on the genotypes under sexual selection (e.g.
-those that increase male mating/fertilisation success) being genetically positively correlated with traits affecting fitness (e.g.
+However, **sexual selection can increase the fitness of a population when ** the genotypes under sexual selection (e.g.
+those that increase male mating/fertilisation success) are genetically positively correlated with traits affecting fitness (e.g.
 female fecundity or survival).
@@
 Through prioritising mating success a sex may invest in costly weaponry and sexual signals, while forgoing strategies that would improve offspring fitness (parental care).
-This is a potentially costly consequence that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].
+This is a potentially costly consequence of sexual selection that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].
 
-**Sexual selection, heritability and mechanisms of speciation**: Alongside changes to fitness, sexual selection has been proposed to facilitate ecological diversification and speciation through increased evolvability of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982].
-Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, and by itself sexual selection could lead to behavioural reproductive isolation [@Safran_2013].
+**Sexual selection, heritability and mechanisms of speciation**: Alongside changes to fitness, sexual selection has been proposed to facilitate ecological diversification ***not sure what you mean by this, like specialization in different niches?
+or generalization?** and speciation through increased evolvability of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982].
+** I think a example illustrating the above would be good** Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, and by itself sexual selection could lead to behavioural reproductive isolation [@Safran_2013].
 But sexual selection is unlikely to be occurring without the effects of natural selection.
-In this situation, adaptation and speciation may be further enforced if female preference in mating signals are dually favoured by natural selection [@Maan_2008].
+In this situation, adaptation and speciation may be further enforced if female preferences in mating signals are dually **which are the two?** favoured by natural selection [@Maan_2008].
 The extent to which sexually selected traits can evolve is dependent on the underlying genetic variation.
-Under sexual selection a wider phenotypic space may be available for a population to evolve into than natural selection acting alone as genetic variation can be maintained under sexual conflict arising from heightened sexual selection [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
+Under sexual selection a wider phenotypic space may be available for a population to evolve into than natural selection acting alone, as genetic variation can be maintained under sexual conflict arising from heightened sexual selection [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
 Furthermore, the evolvability and transmission of fitness and sexually selected traits may vary due to different sex determination systems across taxa.
@@
 We might therefore expect birds with higher opportunities for sexual selection to have increased evolvability and speciation rates, especially in environments where natural selection pressures are high and working concordantly to sexual selection.
+*** I think that sentence needs clarification and a reference, by higher opportunities you mean stronger sex sel?
+***
 
-**Environmental interaction**: Importantly, the effects of sexual selection, adaptation and speciation may depend on environmental conditions.
+**Environmental interaction**: Importantly, the effects of sexual selection **on?** adaptation and speciation may depend on environmental conditions.
 Studies suggest that under stressful/changing environments sexual selection may have greater fitness benefits than under benign environments.
@@
 In these cases genetic variation preferentially remains at sexually antagonistic loci; causing displacement of male and female traits from their optimum.
-These predictions would ensue that populations subject to persistant environmental variability would have increased fitness and adaptive benefits under increased sexual selection; allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
-Increased environmental variability in a species ecological and climatic niche may also work synergestically with natural selection to promote diverging phenotypic traits [@Safran_2013].
+These predictions would ensue that populations **with stronger sexual selection would have increased fitness and adaptive benefits ** , allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
+Increased environmental variability in a species ecological and climatic niche may also work synergestically with natural selection to promote diverging phenotypic traits **maybe leave out this last sentence?
+we are not really interested in how nat and sex sel interact to generate phenotypic variation but how combined they can affect speciation** [@Safran_2013].
 In certain circumstances sexual selection will reinforce diverging locally adaptated phenotypes through mate choice on phenotypes that advertise quality in a specific ecological context [@Maan_2011].
-If mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to strengthen this positive feedback loop and promote population divergence [@Lande_1981].
-Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signalling strength in stickbacks [@Wong_2007; @Candolin_2007].
+**any examples?** If mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to strengthen this positive feedback loop and promote population divergence [@Lande_1981].
+**maybe mention magic traits?
+those that are the result of natural selection but end up affecting sexual selection and thus speciation?** Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signalling strength in sticklebacks [@Wong_2007; @Candolin_2007].
 In this opposing view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.
@@
 Importantly, macroevolutionary comparative studies can examine the association between sexual selection and extinction risk in more diverse taxon, over longer evolutionary time-scales and in more natural environments.
-A meta-analysis of 64 effect sizes from 20 papers found a small, significant association between sexual selecetion and speciation [@Kraaijeveld_2011].
+A meta-analysis of 64 effect sizes from 20 papers found a small, significant **positive?** association between sexual selecetion and speciation [@Kraaijeveld_2011].
 However, variation in sexual selection proxies used and other methodological choices (species richness vs rate estimates) led to heterogeneity in effect sizes for the correlation between sexual selection and speciation.
-Given that species richness and speciation rate may be negatively correlated [@Rabosky_fish_2018] and the recent development and refinement of methodologies in obtaining tip-rate estimates of speciation and extinction [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting with methods that address several issues that have previously limited macroevolutionary comparative studies.
+Given that species richness and speciation rate are not neccesarily correlated e.g.
+[@Rabosky_fish_2018] and the recent development and refinement of methodologies in obtaining tip-rate estimates of speciation and extinction [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting with methods that address several issues that have previously limited macroevolutionary comparative studies.
 Firstly, sexual selection strength in macroevolutionary studies are often estimated using measures of dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa, which arguably decreases the power to detect an association, especially if phylogenetic signal is strong.
@@
 Thirdly, until recently studies have rarely incorperated extensive bioclimatic predictors into large-scale comparative analyses [*sensu* @Jetz_env_2011; @Pigot_2016].
-Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague the ability to draw widely accepted findings **I don't know if this is true**.
+** I would give more importance to this one and frame it slightly different: e.g.
+Until recently studies on the association between sexual selection and speciation have largely ignored the contribution of environmental variables to explain variation in speciation rates.
+Furthermore, these studies haven't taken into account either the potential interaction between natural and sexual selection in the generation of new species.** Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague the ability to draw widely accepted findings **I don't know if this is true**.
 
-**Our study**: Because of the wealth of data available for passerine birds (Order: Passeriformes) we are able to conduct a large comparative analysis across nearly all species of passerines.
+**Our study**: We chose to do our study on passerine birds because of the wealth of data available for these (Order: Passeriformes).
 Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset (n = 918) [@Armenta_2008].
 Here we assess the relationship between sexual dichromatism, environmental variability and their interaction with three tip-rate measures of speciation and one measure of extinction ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$;) in 5,812 species of passerines.
-We assess the effects of sexual dichromatism measures based on RGB values alongside smaller datasets of sexual dichromatism from spectrophotometry measures (*n* = 581) and male-bias sexual selection measures (*n* = 2,465).
-Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is are dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species' range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
+Moreover, we assess the effects of sexual dichromatism measures based on RGB values alongside smaller datasets of sexual dichromatism from spectrophotometry measures (*n* = 581) and male-bias sexual selection measures (*n* = 2,465).**the bit of the sex sel index looks a bit lost in this paragraph but it's quite important, so I'd move it up and bring more attention to it** Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is are dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species' range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
 
@@
 Here we use an existing dataset [@Dale_2015] of male-bias sexual selection based on the first component of a phylogenetic principle component analysis (ppca) of three characteristics associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
-However, this measure of male-bias sexual selection is present for only 2,465 species and is weakly correlated with the absolute values for sexual dichromatism using RGB measures ($r = 0.34$; **Figure S10**).
+This measure of male-bias sexual selection is present for only 2,465 species and is weakly correlated with the absolute values for sexual dichromatism using RGB measures ($r = 0.34$; **Figure S10**).
 
@@
 Using these distributions we were able to obtain estimates for extensive climatic conditions that each species is exposed to.
-Because of taxanomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
+Because of taxanomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015] **was is literally manually?
+OMG!
+or do you have a code that does it really well?
+Can I have it!?
+:)** .
 From 1,000 random point samples in each range we extracted 19 present day bioclimatic variables (representing a variety of biologically relevent annual trends in temperature and precipitation) with 30 seconds (~1 km^2^) spatial resultion [@Fick_2017].
@@
 We used a maximum clade credibility tree (MCC) from 2,500 samples of the posterior distribution that were subset to the passerine bird order (*n* = 5,965) as the main phylogentic tree in our comparative analysis.
-Additionally, a random draw of 100 full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used.
+Additionally, a random draw of 100 full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using BAMM.
 These trees used a 'Hackett backbone' and were constructed using the a pure birth (Yule) model.
@@
 Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014].
-The model uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate a probability distribution of evolutionary regimes with variable speciation and extinction rates.
+This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate a probability distribution of evolutionary regimes with variable speciation and extinction rates.
 Importantly, these models provide tip-rate estimates of speciation and extinction rate.
@@
 We obtained effective sample sizes for log-likelihood and number of rate shifts over 200 for each run (Table S3, Table S4).
-Further details of BAMM parameters and output is available through ESM, with tip-rate means and variances included in the ESM.
+Further details of BAMM parameters and output are available through ESM, with tip-rate means and variances included in the ESM.
 Additionally, given the variability in BAMM estimates we also provide some analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on an MCC from a genetic-only phylogeny across all birds [@Harvey_2017].
-All analyses was conducted on log-rates.
+All analyses were conducted on log-rates.
 
@@
 PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (**Table S1**).
-Given that species range is a potentially informative predictor of speciation and extinction we control for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; **Figure S1**).
+Given that species range is a potentially informative predictor of speciation and extinction we controlled for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; **Figure S1**).
 To obtain long-term variation in climates for each species range we take the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
@@
 We began by estimating the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using ``corPagel`` in the ``ape`` package [@ape_2004] on a MCC tree.
-The estimate of $\lambda$ was then fixed for model selection, which was conducted on the same MCC tree derived from 2,500 draws from the posterior distribution [@Jetz_2012].
-Model selection used $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$ tip-rate estimates from the MCC (n = 5,812) as the response variable and only compared interaction terms.
+The estimate of $\lambda$ was then fixed for model selection, which was conducted on the same MCC tree drived from 2,500 draws from the posterior distribution [@Jetz_2012].
+Model selection used either $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ tip-rate estimates from the MCC (n = 5,812) as the response variable and only compared interaction terms.
 That is, we fixed the individual predictors of: sexual dichromatism/selection, log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation and NPP while comparing 32 models with different combinations of interactions (including none).
@@
 This method was conducted on models that used data from all available passerine species (*n* = 5,812).
-PGLS analyses were also applied to a subset dataset of passerine birds (*n* = 581), where sexual dichromatism measures were sourced from spectrophotometry measures [@Armenta_2008] and where male-bias sexual selection ppca was used as a predictor instead of sexual dichromatism (*n* = 2,465).
+The same proceedure decribed above was applied to a subset dataset of passerine birds (*n* = 581), where sexual dichromatism measures were sourced from spectrophotometry measures [@Armenta_2008] and where male-bias sexual selection ppca was used as a predictor instead of sexual dichromatism (*n* = 2,465).
 Finally, using this data subsetted for species with measurements of male-bias sexual selection we conducted a phylogenetic path analysis using the ``phylopath`` ``R`` package [@Wouter_2018].
-The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
-To minimise path complexity we use temperature seasonality (BIO4) as the single measure for environmental variability and use $\lambda_{DR}$ as the single measure of speciation.
+The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.**which tree did you use for this?
+or the 100 trees?** To minimise path complexity we use temperature seasonality (BIO4) as the single measure for environmental variability and use $\lambda_{DR}$ as the single measure of speciation.
 Further details of the path analysis including reasons for path directions can be found within the ESM along with all other analyses and the relevant ``R`` code to reproduce results.
@@
 
-PGLS models with interaction terms between the respective measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in global models.
+PGLS models with interaction terms between the respective measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in global models **in the top you say you chose the top ranking model but you refer here to the global, not sure what you mean by global**.
 However, in all model selection phases for different measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and different measures of sexual selection (RGB values, spectrophotometry and male-biased sexual selection) no interaction terms were significant or present in the top models ($\delta$ AICc > 4; **Table S5**, **Table S6**, **Table S10**, **Table S13**).
@@
 However, no association was found between sexual selection and $\lambda_{BAMM}$ ($\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-For PGLS models on 100 random trees, we find the distribution to be similar to the estimates when an MCC tree is used.
+** here I would say the same for ND and bamm, either that they are positive but non sig nificant or that they show no trend** For PGLS models on 100 random trees, we find the distribution to be similar to the estimates when an MCC tree is used.
 Among the 100 tress we find a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a positive (but less so) association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
@@
 Both datasets were used for anlyses with three measures of speciation $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and one measure of extinction $\mu_{BAMM}$ as response variables.
-When using $\lambda_{DR}$ and $\lambda_{ND}$ as the response variable, they provide relatively congruent results; with a negative relationship between $\lambda_{DR/ND}$ and range size as well as a positive relationship between $\lambda_{DR/ND}$ and sexual selection (**b**).
+When using $\lambda_{DR}$ and $\lambda_{ND}$ as the response variable, they provide relatively congruent results; **no need to detail the results in the legend**with a negative relationship between $\lambda_{DR/ND}$ and range size as well as a positive relationship between $\lambda_{DR/ND}$ and sexual selection (**b**).
 Alternatively, when using $\lambda_{BAMM}$ and $\mu_{BAMM}$ the model estimates are more variable across trees compared to the MCC estimate, which do not significantly differ from zero.
@@
 We found no evidence that sexual dichromatism alters the rate of speciation in passerine birds independently or via interactions with environmental variability.
-This result is consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and two measures of dichromatism: a near-complete dataset of 5,812 species using RGB scores and a smaller dataset of 581 species where measurements were obtained from spectrophotometry (allowing variation in the ultraviolet and bird-visible range to be detected).
+This result is consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and two measures of dichromatism (spectral and RGB).
 However, we did find that when using a male-bias measure of sexual selection --- instead of sexual dichromatism --- there was a positive association with speciation: with significance varying across different rate estimates.
@@
 Ecological impacts on sexual dichromatism are supported by our path analysis; which reveals that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation) albeit by a relatively small amount.
-Given that speciation rate increases with sexual selection strength but not sexual dichromatism we are led to conclude that sexual dichromatism is a seemingly poor proxy of sexual selection and that the noise associated with the proxy measurement of sexual selection cancels out the obtained signal; this is despite the high power of our comparative study.
+Given that our results show that speciation rate increases with sexual selection strength, but not sexual dichromatism, we are led to conclude that sexual dichromatism is a seemingly poor proxy of sexual selection and that the noise associated with the proxy measurement of sexual selection cancels out the obtained signal; this is despite the high power of our comparative study.
 The usefulness of sexual dichromatism as a proxy of sexual selection may be dependent upon the metabolic pathways that give rise to plumage colouration; with not all plumage colouration reflecting mate quality.
 A recent meta-analysis found that converted carotenoids (but not those sourced from diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
-Given the importance of honest signalling in sexual selection and the variability in colour production across birds, sexual dichromatism in many species may unreliable measure of mate quality for female birds and sexual selection for researchers.
+Given the importance of honest signalling in sexual selection and the variability in colour production across birds, sexual dichromatism in many species may unreliable measure of mate quality for female birds and sexual selection for researchers.**Great paragraph!!**
 
@@
 Given that we did not detect an effect of environmental variability (by itself or as an interaction with sexual selection) we are left with several possible explanations.
-Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of envrionmental variability the species is evolving under.
+Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability the species is evolving under.
 Specifically, sexual antagonism may be more easily purged in environments where directional environment pressures are maintained; whereas when environmental variability is cyclical (e.g.
-seasonal or glacial climate cycles) sexual antagonism can remain indefinately as a fitness burden [@Connallon_Hall_2016].
+seasonal or glacial climate cycles) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
 Secondly, the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
 Specifically, access to dietary resources, and the impacts of predation or parasitism are unaccounted for here: arguably these are key processes affecting sexual selection and speciation [*reviewed in*, @Maan_2011].
-Thirdly, it is also likly that there is no effect; at least not in the species investigated here.
+Thirdly, it is also likely that there is no effect; at least not in the species investigated here.
 
@@
 
-The most convincing finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates.
+The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates.
 But drawing conclusions from this finding is fraught with challenges.
 Challenges arise on determining the causality of range size and speciation; as arguably small range size could either be a cause or effect of speciation.
-High speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated taxa.
+** In the next sentence I would start the opposite way, as the model was framed with range affecting speciation instead of the opossite**High speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated taxa.
 Within Himalayan songbirds range expansions are prevented by the competition for resources by related species [@Price_2014]; however this limitation on range expansion is then expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011].
-Thus, if range size is a cause of speciation we might expect it to be smaller but if it is a predictor of speciation it arguably should be larger; with greater opportunities for geographic barriers to form in larger ranges [@Rosenzweig_1995].
-Unlike Himalayan songbirds, species undergoing adaptive radiation into newly formed ecological environments are unlikely to be limited by competition for resources from pre-existing taxa.
+Thus, if range size is a cause of speciation we might expect it to be smaller but
+if it is a predictor of speciation it arguably should be larger; with greater opportunities for geographic barriers to form in larger ranges [@Rosenzweig_1995]. 
+** this one is tricky, because yes, what you say above can be true, but also: species with larges ranges mean they have less limits to dispersal, with less restrictions on gene flow and thus lower speciation rates.
+range size can affect speciation rates just because is a proxy for dispersal abilities or something like that.
+The thing is that range size is probably the result of many processes**
+
+Unlike Himalayan songbirds, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from pre-existing taxa.
 Across islands we expect to see a correlation between speciation and small range size, not because of any inherent properties of the range size, rather that they reflect newly formed islands with empty niches **This seems like a very established idea but not sure whose it is**.
@@
 Birds with large range sizes are also predicted to have greater dispersal ability; in these cases individuals may be more able to overcome geographical barriers and maintain gene flow; thus lowering speciation in these wide-ranging species [@Birand_2012; @Claramunt_2012].
-Several studies have used morphological measures of wing shape, a proxy for dispersal ability (Hand-wing index/ Kipp's distance), as a predictor of speciation and coexistence [*e.g.*, @Kennedy_2016; @Pigot_2016], but these measurements across passerines is not yet extensive **I think?
+**ok, so yeah, like i said, maybe move this to where I mentioned it earlier**Several studies have used morphological measures of wing shape, a proxy for dispersal ability (Hand-wing index/ Kipp's distance), as a predictor of speciation and coexistence [*e.g.*, @Kennedy_2016; @Pigot_2016], but these measurements across passerines is not yet extensive **I think?
 ~ 1000**.
-One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification whereby splitting of species leads to increased recent phylogenetic branching as well as a more subdivided range **Has this argument been articulated?**.
+One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification whereby splitting of subspecies into species in clades with large ranges leads to increased recent phylogenetic branching as well as a more subdivided range **Has this argument been articulated?** **I can't remember reading it, maybe is not true and I made it up haha.
+Makes sense to me but maybe good to see what reviewers would think about it.
+I guess it would also depend on whether the taxa with large ranges are considered the same species because there is actually no strong genetic differentiation across the range and not because there isn't enough information (e.g.sampling) to split a species with massive range into smaller ranges and more species** .
 The role of range size in speciation may be further investigated through ancestral geographic range reconstruction [@Quintero_2015], as ancestral ranges may be beneficial in understanding modes of speciation and the causality of range size expansions and contractions [@Losos_2003].
 
-Here we used BAMM to assess extinction rates across passerines and whether these rates were associated with sexual dichromatism/selection or various ecological variables.
+Here we used BAMM to calculate extinction rates across passerines and test whether these rates were associated with sexual dichromatism/selection or various ecological variables.
 We found no associations in this respect despite the wealth of theoretical predictions on the effects of sexual selection on fitness.
-Extinction is notoriously difficult to obtain accurate estimates for in comparative studies, as it is a process that occurs over millennia, leaving scant traces.
+Extinction is notoriously difficult to obtain accurate estimates for in comparative studies **Cite papers, e.g.
+Rabosky: 'Extinction rates should not be estimated from molecular phylogenies'**, as it is a process that occurs over millennia, leaving scant traces.
 One exception to this comes from fossil ostracods, a group of marine invertebrates.
@@
 Here we anticipated extinction rates would be underestimated because we used a phylogeny constructed with a birth-only (Yule) model [@Jetz_2012].
-Given these underestimates of $\mu_{BAMM}$ it is highly likely we failed to detect true associations between sexual dichromatism/selection and extinction.
+Given these underestimates of $\mu_{BAMM}$ it is likely we failed to detect true associations between sexual dichromatism/selection and extinction.
 The power of BAMM as a model based-approach in estimating $\lambda$ and $\mu$ is dependent size of the rate regimes.

### 2018-10-02 ComparativePlots.Rmd
24/9 Update (first manuscript draft, yay!)
Critical Analysis and index add

@@
 cols <- c('#B76643','#B76643','#B76643','#B76643','#B47339','#AC8032','#A08E33','#909B3C','#7DA64D', brewer.pal(n = 9, name = "Greens")[5:9], "#00441B", "#00441B", "#00441B")
+
+cols <- c("#543005", "#8c510a", "#bf812d", "#dfc27d", "#c0d884",
+           brewer.pal(n = 9, name = "YlGn")[4:9])
 
@@
 ```
+
+```{r}
+Lepidothrix.coeruleocapilla.data <- restricted.data %>% filter(TipLabel == "Lepidothrix_coeruleocapilla")
+
+LC.bioclim4 <- data.frame(
+  value = rnorm(1000, mean = Lepidothrix.coeruleocapilla.data$bioclim4, sd = Lepidothrix.coeruleocapilla.data$se.bioclim4))
+
+LC.bioclim4$name <- "bioclim4"
+
+LC.bioclim4 %>% ggplot()+
+  geom_density_ridges(aes(x = value/10, y = 0), rel_min_height = 0.005, fill = "#fc8d62")+
+  scale_y_discrete(expand = c(0,0))+
+  ylab("")+
+  xlab("Temperature Seasonality")+
+
+  theme_bw()+
+
+  theme(
+        text = element_text(size=18),
+        panel.border= element_blank(),
+        axis.line.y =element_blank(),
+        axis.line.x = element_line(),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.grid.minor.x = element_blank(), 
+        legend.text = element_text(size=18), 
+        legend.title=element_text(size=18, face = "bold"),
+        axis.title.x = element_text(hjust = 0.5, size = 24),
+        axis.title.y = element_text(size = 18, hjust = 0.35, margin = margin(r=-10)),
+        axis.text.y = element_blank(),
+        plot.title = element_text(size = 18))
+```
+
+```{r}
+Lepidothrix.coeruleocapilla.data <- restricted.data %>% filter(TipLabel == "Lepidothrix_coeruleocapilla")
+
+LC.NPP <- data.frame(
+  value = rnorm(1000, mean = Lepidothrix.coeruleocapilla.data$NPP, sd = Lepidothrix.coeruleocapilla.data$NPP.se))
+
+LC.NPP$name <- "NPP"
+
+LC.NPP %>% ggplot()+
+  geom_density_ridges(aes(x = value/10, y = 0), rel_min_height = 0.005, fill = "#a6d854")+
+  scale_y_discrete(expand = c(0,0))+
+  scale_x_continuous(limits = c(1910, 2010))+
+  ylab("")+
+  xlab("NPP")+
+
+  theme_bw()+
+
+  theme(
+        text = element_text(size=18),
+        panel.border= element_blank(),
+        axis.line.y =element_blank(),
+        axis.line.x = element_line(),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.grid.minor.x = element_blank(), 
+        legend.text = element_text(size=18), 
+        legend.title=element_text(size=18, face = "bold"),
+        axis.title.x = element_text(hjust = 0.5, size = 24),
+        axis.title.y = element_text(size = 18, hjust = 0.35, margin = margin(r=-10)),
+        axis.text.y = element_blank(),
+        plot.title = element_text(size = 18))
+
+LC.NPP <- data.frame(
+  value = rnorm(1000, mean = Lepidothrix.coeruleocapilla.data$NPP, sd = Lepidothrix.coeruleocapilla.data$NPP.se))
+
+LC.NPP$name <- "NPP"
+
+LC.NPP %>% ggplot()+
+  geom_density_ridges(aes(x = value/10, y = 0), rel_min_height = 0.005, fill = "#a6d854")+
+  scale_y_discrete(expand = c(0,0))+
+  scale_x_continuous(limits = c(1910, 2010))+
+  ylab("")+
+  xlab("NPP")+
+
+  theme_bw()+
+
+  theme(
+        text = element_text(size=18),
+        panel.border= element_blank(),
+        axis.line.y =element_blank(),
+        axis.line.x = element_line(),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.grid.minor.x = element_blank(), 
+        legend.text = element_text(size=18), 
+        legend.title=element_text(size=18, face = "bold"),
+        axis.title.x = element_text(hjust = 0.5, size = 24),
+        axis.title.y = element_text(size = 18, hjust = 0.35, margin = margin(r=-10)),
+        axis.text.y = element_blank(),
+        plot.title = element_text(size = 18))
+```
+
+```{r}
+#read nexus
+manakin.trees <- read.nexus(file.choose())
+#get MCC
+manakin.tree <- maxCladeCred(manakin.trees)
+
+manakin.DR <- calculate.log.es(manakin.tree)
+```

### 2018-10-02 Manuscript.Rmd
Critical Analysis and index add

@@
 ---
-title: "Sexual selection, environmental variation, range size and speciation in passerine birds"
+title: Sexual selection, environmental variation, range size and speciation in passerine
+  birds
 author:
-- Justin G. Cally*\textsuperscript{§}, Iliana Medina Guzmán\textsuperscript{§}, Luke Holman\textsuperscript{§} and Devi Stuart-Fox\textsuperscript{§}
-- 
+- Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§} and Iliana Medina Guzmán\textsuperscript{§}
+- ~
 - '*\textit{justin.g.cally@gmail.com}'
+bibliography: References/SS_Extinction.bib
 output:
+  pdf_document: default
   html_document: default
   word_document:
-      pandoc_args: ["--smart"]
-  pdf_document: default
-bibliography: References/SS_Extinction.bib
+    pandoc_args: --smart
 header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
@@
   \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
-  \rhead[]{Sexual Selection and Extinction} \usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot}
-  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\doublespacing \usepackage[font={small}]{caption}
+  \rhead[]{Sexual Selection and Extinction} \usepackage[noabbrev,capitalise]{cleveref}
+  \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
+  \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption}
+  \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
+  \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip}
+  \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape}
+  \usepackage{hyperref}\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, urlcolor=blue}
 csl: References/AmNat.csl
-abstract: Sexual selection may influence patterns of diversification.
-Yet their relationship remains unclear with controversy on whether sexual selection leads to a net increase or decrease in speciation rate.
-Sexual selection is predicted to increase speciation rate as it can facilitate adaptation to new environments by elevating the average fitness of a population.
-Additionally, sexual selection may work alongide divergent natural selection and swiflty promote reproductive isolation.
-Here we use sexual dichromatism as a proxy for sexual selection across 5,812 species of passerine birds and a multivariate male-bias measure of sexual selection (sexual dimorphism, polygyny and paternal care) across 2,465 species to investigate the association of sexual selection with speciation and extinction.
-Importantly, we test whether the effects of sexual selection on speciation are dependent upon ecological pressures such as temperature variability across time and space, primary productivity and range size.
-Our results from three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) find that male-bias sexual selection but not sexual dichromatism predicts speciation.
-Additionally, we find that birds with smaller ranges have higher speciation rates; however the causality of this finding remains inconclusive.
+abstract: Sexual selection may influence patterns of diversification. Yet their relationship
+  remains unclear with controversy on whether sexual selection leads to a net increase
+  or decrease in speciation rate. Sexual selection is predicted to increase speciation
+  rate as it can facilitate adaptation to new environments by elevating the average
+  fitness of a population. Additionally, sexual selection may work alongide divergent
+  natural selection and swiflty promote reproductive isolation. Here we use sexual
+  dichromatism as a proxy for sexual selection across 5,812 species of passerine birds
+  and a multivariate male-bias measure of sexual selection (sexual dimorphism, polygyny
+  and paternal care) across 2,465 species to investigate the association of sexual
+  selection with speciation and extinction. Importantly, we test whether the effects
+  of sexual selection on speciation are dependent upon ecological pressures such as
+  temperature variability across time and space, primary productivity and range size.
+  Our results from three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$,
+  $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) find that male-bias
+  sexual selection but not sexual dichromatism predicts speciation. Additionally,
+  we find that birds with smaller ranges have higher speciation rates; however the
+  causality of this finding remains inconclusive.
 ---
@@
 \newpage
-#Introduction xxxxxx
+
+#Introduction
 
@@
 As populations diverge they accumulate genetic and phenotypic changes that ultimately erect barriers against gene flow [@Presgraves_2010; *but see* @Shaw_2011].
-In this view populations will speciate and adapt more readily to to new divergent environments if the selection pressures and evolvability of the traits under selection are high.
-Hence, the rate at which speciation occurs is likely dependent on both the populations adaptive potential/fitness and the evolvability **mmm here it becomes unclear because you are suggesting the link between speciation and the evolvability of traits, but is not clear what you mean by evolvability: a.
-the potential to evolve (e.g.
-genetic architecture) b.
-the rate of trait evolution or c.
-the magnitud of the difference in a trait between populations.
-I would either define evolvability earlier or replace the word with something easier to grasp** of phenotypic traits such as morphology or chromatism **color?
-avoid jargon when possible** [@Rabosky_morph_2013; @Hugall_2012].
-Sexual selection is an evolutionary force likely to affect fitness components as well as the evolvability **here I would just use 'evolution' rather than evolvability** of morphological and chromatic traits such as ornaments involved in mating displays or weapons used in male-male competition.
-It is therefore hypothesised that sexual selection is a key driver of diversification through shifting the rates of adaptation, speciation and extinction.
+In this view, populations will speciate and adapt more readily to to new divergent environments if selection pressures are high.
+Hence, the rate at which speciation occurs is likely dependent on both the populations response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour, @Hugall_2012; @Rabosky_morph_2013].
+Sexual selection is an evolutionary force likely to affect fitness components as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition.
+It is therefore hypothesised that sexual selection is a key driver of diversification through shifting the rates of adaptation, divergence, speciation and extinction.
 
-**Sexual selection and fitness**: Sexual selection is applied by one sex on the other - more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and **the strength of sexual selection**can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
-Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations, [@Agrawal_2001; @Siller_2001] and fixate beneficial mutations in a population [@Whitlock_2000].
+**Sexual selection and fitness**: Sexual selection is applied by one sex on the other - more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and the strength of sexual selection can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
+Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fixate beneficial mutations in a population [@Whitlock_2000].
 Thus, sexual selection is thought to be a key process in mitigating extinction risk and facilitating adaptation.
-However, **sexual selection can increase the fitness of a population when ** the genotypes under sexual selection (e.g.
-those that increase male mating/fertilisation success) are genetically positively correlated with traits affecting fitness (e.g.
-female fecundity or survival).
+However, sexual selection can increase the fitness of a population when the genotypes under sexual selection (*e.g.* those that increase male mating/fertilisation success) are genetically positively correlated with traits affecting fitness (*e.g.* female fecundity or survival).
 This positive correlation is unlikely to always be the case; regularly, sexual selection favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006].
@@
 Through prioritising mating success a sex may invest in costly weaponry and sexual signals, while forgoing strategies that would improve offspring fitness (parental care).
-This is a potentially costly consequence of sexual selection that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].
+This is a costly consequence of sexual selection that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].
 
-**Sexual selection, heritability and mechanisms of speciation**: Alongside changes to fitness, sexual selection has been proposed to facilitate ecological diversification ***not sure what you mean by this, like specialization in different niches?
-or generalization?** and speciation through increased evolvability of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982].
-** I think a example illustrating the above would be good** Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, and by itself sexual selection could lead to behavioural reproductive isolation [@Safran_2013].
-But sexual selection is unlikely to be occurring without the effects of natural selection.
-In this situation, adaptation and speciation may be further enforced if female preferences in mating signals are dually **which are the two?** favoured by natural selection [@Maan_2008].
-The extent to which sexually selected traits can evolve is dependent on the underlying genetic variation.
-Under sexual selection a wider phenotypic space may be available for a population to evolve into than natural selection acting alone, as genetic variation can be maintained under sexual conflict arising from heightened sexual selection [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
-Furthermore, the evolvability and transmission of fitness and sexually selected traits may vary due to different sex determination systems across taxa.
-In male homogametic species (ZZ/ZW or ZZ/ZO) --- such as birds --- males are anticipated to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage coloration [@Hastings_1994; @Reeve_2003].
-We might therefore expect birds with higher opportunities for sexual selection to have increased evolvability and speciation rates, especially in environments where natural selection pressures are high and working concordantly to sexual selection.
-*** I think that sentence needs clarification and a reference, by higher opportunities you mean stronger sex sel?
-***
+**Sexual selection, heritability and mechanisms of speciation**: Alongside changes to fitness, sexual selection has been proposed to facilitate speciation through the evolution of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982].
+For instance, in antbirds (Thamnophilidae), genera with complex songs and stark dichromatism were more species rich [@Seddon_2008].
+Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, and by itself sexual selection could lead to behavioural reproductive isolation [@Safran_2013].
+Concurrently, sexual selection can promote the extent to which populations diverge by enlarging the phenotypic space diverging species can evolve into.
+This prediction emanates from increased genetic diversity as a result of sexual conflict: a consequence of sexual selection [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
+However, the genetic architecture (*per se* heritability) of fitness and sexually selected traits may vary due to different sex determination systems across taxa.
+In male homogametic species (ZZ/ZW or ZZ/ZO) --- such as birds --- males are anticipated to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003].
+We might therefore expect birds with magnified sexual selection to have increased adaptation and speciation rates.
 
-**Environmental interaction**: Importantly, the effects of sexual selection **on?** adaptation and speciation may depend on environmental conditions.
+**Environmental interaction**: Importantly, the effects of sexual selection on adaptation and speciation may depend on environmental conditions.
 Studies suggest that under stressful/changing environments sexual selection may have greater fitness benefits than under benign environments.
 In stressful environments, the positive correlation between male mating success and female fitness components improves, reducing the burden of negative pleiotropic effects [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017].
-Conversely, under benign environments environmental stability ensures consistent selection, preferentially removing genetic variation at sexually concordant loci (i.e.
-loci where the optimum genotype is equivelant for both sexes).
+Conversely, under benign environments environmental stability ensures consistent selection, preferentially eroding genetic variation at sexually concordant loci (*i.e.* loci where the optimum genotype is equivelant for both sexes).
 In these cases genetic variation preferentially remains at sexually antagonistic loci; causing displacement of male and female traits from their optimum.
-These predictions would ensue that populations **with stronger sexual selection would have increased fitness and adaptive benefits ** , allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
-Increased environmental variability in a species ecological and climatic niche may also work synergestically with natural selection to promote diverging phenotypic traits **maybe leave out this last sentence?
-we are not really interested in how nat and sex sel interact to generate phenotypic variation but how combined they can affect speciation** [@Safran_2013].
+These predictions would ensue that populations with stronger sexual selection would have increased fitness and adaptive benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
 In certain circumstances sexual selection will reinforce diverging locally adaptated phenotypes through mate choice on phenotypes that advertise quality in a specific ecological context [@Maan_2011].
-**any examples?** If mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to strengthen this positive feedback loop and promote population divergence [@Lande_1981].
-**maybe mention magic traits?
-those that are the result of natural selection but end up affecting sexual selection and thus speciation?** Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signalling strength in sticklebacks [@Wong_2007; @Candolin_2007].
+If mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to strengthen this positive feedback loop and promote population divergence [@Lande_1981].
+In several birds, beak morphology is a trait that arose under natural selection and is now a trait under sexual selection (a.k.a.
+"magic traits").
+In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
+Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signalling strength in sticklebacks [@Wong_2007; @Candolin_2007].
 In this opposing view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.
@@
 **Previous limitations**: The association between sexual selection and speciation has been investigated using many macroevolutionary approaches across various taxa.
-Importantly, macroevolutionary comparative studies can examine the association between sexual selection and extinction risk in more diverse taxon, over longer evolutionary time-scales and in more natural environments.
-A meta-analysis of 64 effect sizes from 20 papers found a small, significant **positive?** association between sexual selecetion and speciation [@Kraaijeveld_2011].
+Importantly, macroevolutionary comparative studies can examine the association between sexual selection and speciation in more diverse taxon, over longer evolutionary time-scales and in more natural environments.
+A meta-analysis of 64 effect sizes from 20 papers found a small, but positively significant association between sexual selecetion and speciation [@Kraaijeveld_2011].
 However, variation in sexual selection proxies used and other methodological choices (species richness vs rate estimates) led to heterogeneity in effect sizes for the correlation between sexual selection and speciation.
-Given that species richness and speciation rate are not neccesarily correlated e.g.
-[@Rabosky_fish_2018] and the recent development and refinement of methodologies in obtaining tip-rate estimates of speciation and extinction [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting with methods that address several issues that have previously limited macroevolutionary comparative studies.
-Firstly, sexual selection strength in macroevolutionary studies are often estimated using measures of dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa, which arguably decreases the power to detect an association, especially if phylogenetic signal is strong.
+Given that species richness and speciation rate are not neccesarily correlated [*e.g.* @Rabosky_fish_2018] and the recent development and refinement of methodologies in obtaining tip-rate estimates of speciation and extinction [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting with methods that address several issues that have previously limited macroevolutionary comparative studies.
+Firstly, sexual selection strength in macroevolutionary studies are often estimated using measures of dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa; which arguably decreases the power to detect an association, especially if phylogenetic signal is strong.
 Secondlly, many studies use incomplete taxon sampling, with family-level analysis restricted to well sampled avian clades [@Seddon_2008; @Huang_2014].
-Thirdly, until recently studies have rarely incorperated extensive bioclimatic predictors into large-scale comparative analyses [*sensu* @Jetz_env_2011; @Pigot_2016].
-** I would give more importance to this one and frame it slightly different: e.g.
-Until recently studies on the association between sexual selection and speciation have largely ignored the contribution of environmental variables to explain variation in speciation rates.
-Furthermore, these studies haven't taken into account either the potential interaction between natural and sexual selection in the generation of new species.** Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague the ability to draw widely accepted findings **I don't know if this is true**.
+Thirdly, until recently studies on the association between sexual selection and speciation have largely ignored the contribution of environmental variables to explain variation in speciation rates.
+Furthermore, these studies haven't taken into account either the potential interaction between natural and sexual selection in the generation of new species.
+Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague our ability to draw widely accepted findings.
 
@@
 Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset (n = 918) [@Armenta_2008].
-Here we assess the relationship between sexual dichromatism, environmental variability and their interaction with three tip-rate measures of speciation and one measure of extinction ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$;) in 5,812 species of passerines.
-Moreover, we assess the effects of sexual dichromatism measures based on RGB values alongside smaller datasets of sexual dichromatism from spectrophotometry measures (*n* = 581) and male-bias sexual selection measures (*n* = 2,465).**the bit of the sex sel index looks a bit lost in this paragraph but it's quite important, so I'd move it up and bring more attention to it** Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is are dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species' range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
+But, here we assess the effects of sexual dichromatism measures based on RGB values (*n* = 5,812) alongside smaller datasets of sexual dichromatism from spectrophotometry measures (*n* = 581) and male-bias sexual selection measures (*n* = 2,465).
+Specifically, we assess the relationship between sexual dichromatism, environmental variability and their interaction with three tip-rate measures of speciation and one measure of extinction ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$).
+Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is are dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species' range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
 
@@
 We examined the effect of sexual selection on speciation and extinction rate in 97 % of passerines (*n* = 5,812 species; 58 % of all birds).
-Specifically, we (i) compiled datasets for sexual dichromatism/selection strength and environmental variability, (ii) obtained estimates of speciation and extinction rates across passerines, and (iii) conducted phylogenetic generalized least-squares (PGLS) regressions.
+Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions.
 Many of the processes undertaken in this study are documented with reproducible code in the electronic supplementary material (ESM).
@@
 Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forhead, throat, upper breast and lower breast) from _Handbook of the Birds of the World_ (volumes 8–16) [@delhoyo2011].
-The mean values of red, green and blue were calculated and provided male and female plumage scores.
+The mean RGB values were calculated and provided male and female plumage scores.
 Here we use the absolute difference between males and female RGB scores as an estimate of sexual dichromatism.
@@
 ###Environmental variables
+
 Using expert range maps [@birdlife_2017] we obtained estimates of species range size.
-Using these distributions we were able to obtain estimates for extensive climatic conditions that each species is exposed to.
-Because of taxanomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015] **was is literally manually?
-OMG!
-or do you have a code that does it really well?
-Can I have it!?
-:)** .
-From 1,000 random point samples in each range we extracted 19 present day bioclimatic variables (representing a variety of biologically relevent annual trends in temperature and precipitation) with 30 seconds (~1 km^2^) spatial resultion [@Fick_2017].
+From these distributions we were able to obtain estimates for extensive climatic conditions that each species is exposed to.
+Because of taxanomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
+From 1,000 random point samples in each range we extracted 19 present-day bioclimatic variables (representing a variety of biologically relevent annual trends in temperature and precipitation) with 30 seconds (~1 km^2^) spatial resultion [@Fick_2017].
 From these values we obtained means and standard deviations for each species.
@@
 Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPU's.
-We obtained effective sample sizes for log-likelihood and number of rate shifts over 200 for each run (Table S3, Table S4).
+We obtained effective sample sizes for log-likelihood and number of rate shifts over 200 for each run (**Table S3**, **Table S4**).
 Further details of BAMM parameters and output are available through ESM, with tip-rate means and variances included in the ESM.
@@
 
-Using the extracted environmental variables from each species range size we developed biologically meaningful predictors of environmental variation/stress relating to (i) seasonal climate variation, (ii) spatial climate variation and (iii) long-term climate variation.
+Using the extracted environmental variables from each species range size we developed biologically meaningful predictors of environmental variation/stress relating to (*i*) seasonal climate variation, (*ii*) spatial climate variation and (*iii*) long-term climate variation.
 To obtain seasonal climate variation we used mean values of temperature seasonality (BIO4) for each range.
@@
 We began by estimating the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using ``corPagel`` in the ``ape`` package [@ape_2004] on a MCC tree.
-The estimate of $\lambda$ was then fixed for model selection, which was conducted on the same MCC tree drived from 2,500 draws from the posterior distribution [@Jetz_2012].
-Model selection used either $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ tip-rate estimates from the MCC (n = 5,812) as the response variable and only compared interaction terms.
+The estimate of $\lambda$ was then fixed for model selection, which was conducted on the same MCC tree derived from 2,500 draws of the posterior distribution [@Jetz_2012].
+Model selection used either $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ tip-rate estimates from the MCC as the response variable and only compared interaction terms.
 That is, we fixed the individual predictors of: sexual dichromatism/selection, log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation and NPP while comparing 32 models with different combinations of interactions (including none).
@@
 This method was conducted on models that used data from all available passerine species (*n* = 5,812).
-The same proceedure decribed above was applied to a subset dataset of passerine birds (*n* = 581), where sexual dichromatism measures were sourced from spectrophotometry measures [@Armenta_2008] and where male-bias sexual selection ppca was used as a predictor instead of sexual dichromatism (*n* = 2,465).
+The same procedure decribed above was applied to a subset dataset of passerine birds (*n* = 581), where sexual dichromatism measures were sourced from spectrophotometry measures [@Armenta_2008] and where male-bias sexual selection ppca was used as a predictor instead of sexual dichromatism (*n* = 2,465).
 Finally, using this data subsetted for species with measurements of male-bias sexual selection we conducted a phylogenetic path analysis using the ``phylopath`` ``R`` package [@Wouter_2018].
-The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.**which tree did you use for this?
-or the 100 trees?** To minimise path complexity we use temperature seasonality (BIO4) as the single measure for environmental variability and use $\lambda_{DR}$ as the single measure of speciation.
+The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
+To minimise path complexity we use temperature seasonality (BIO4) as the single measure for environmental variability and use $\lambda_{DR}$ as the single measure of speciation.
+The phylogenetic path analysis used tip-rates and correlation structure from the MCC tree.
 Further details of the path analysis including reasons for path directions can be found within the ESM along with all other analyses and the relevant ``R`` code to reproduce results.
-
-#Results
 
@@
 ```
+#Results
 
-##Effects of sexual dichromatism/selection on speciation are not dependent upon environmental variability
+##Effects of sexual dichromatism/selection on speciation are not dependent on environmental variability
 
-PGLS models with interaction terms between the respective measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in global models **in the top you say you chose the top ranking model but you refer here to the global, not sure what you mean by global**.
+PGLS models with interaction terms between the respective measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in model selection.
 However, in all model selection phases for different measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and different measures of sexual selection (RGB values, spectrophotometry and male-biased sexual selection) no interaction terms were significant or present in the top models ($\delta$ AICc > 4; **Table S5**, **Table S6**, **Table S10**, **Table S13**).
 Hence we find no evidence that the effect of sexual selection on speciation or extinction is dependent upon certain environmental conditions.
-Furthermore, we find no support for these individual bioclimatic factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) acting as a driver of speciation independently from sexual dichromatism/selection (\autoref{model_results}, **Figure S10**).
+Furthermore, we find no support that these individual bioclimatic factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) act as drivers of speciation independently from sexual dichromatism/selection (\autoref{model_results}, **Figure S10**).
 
@@
 
-When we conducted PGLS analysis using data subsetted for values of sexual selection estimated from a phylogenetic PCA of sexual dimorphism, social polygyny and paternal care (*n* = 2,465) there was a significant positive association between sexual selection and $\lambda_{DR}$ ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b) and a non-significant positive association between sexual selection and $\lambda_{ND}$ ($\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-However, no association was found between sexual selection and $\lambda_{BAMM}$ ($\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-** here I would say the same for ND and bamm, either that they are positive but non sig nificant or that they show no trend** For PGLS models on 100 random trees, we find the distribution to be similar to the estimates when an MCC tree is used.
+When we conducted PGLS analysis using data subsetted for values of sexual selection estimated from a phylogenetic PCA of sexual dimorphism, social polygyny and paternal care (*n* = 2,465) there was a significant positive association between sexual selection and $\lambda_{DR}$ ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+However, this positive finding was non-significant between sexual selection and $\lambda_{ND}$ ($\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b) as well as between sexual selection and $\lambda_{BAMM}$ ($\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+For PGLS models on 100 random trees, we find the distribution to be similar to the estimates when an MCC tree is used.
 Among the 100 tress we find a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a positive (but less so) association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
@@
 Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from sexual selection to sexual dichromatism was relatively weak given the expectations of dichromatism being a well utilised proxy of sexual selection ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
-Additionally, given the large direct effect of temperature seasonality on range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`) and our findings from the PGLS that range size is negatively associated with $\lambda_{DR}$, there is an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}).
+Additionally, because of the large direct effect of temperature seasonality on range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`) and our findings from the PGLS that range size is negatively associated with $\lambda_{DR}$, there is an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}).
 
@@
 We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-This null result was consistent for spectrophotometry measures of sexual dichromatism (**FigureS10, Table S11, Table S12**) and for the estimates of sexual selection using the sexual selection ppca (\autoref{model_results}, **Table S14, Table S15**).
+This null result was consistent for spectrophotometry measures of sexual dichromatism (**FigureS10, Table S11, Table S12**) and for the estimates of sexual selection using the sexual selection ppca (\autoref{model_results}b, **Table S14, Table S15**).
 
@@
 The range of methods utilised were implemented to overcome uncertainty in the power and precision of various tip-rate estimates as well as the variation between trees.
-We found that that the tip-rate estimates across the 100 trees were quite variable for both $\lambda_{DR}$ and $\lambda_{BAMM}$ (**Figure S13**).
-Despite this variation, both methods provided similar estimates for speciation in each species for the mean values across the 100 trees ($r = 0.67$) and for the tip-rate estimates from the MCC ($r = 0.68$).
+We found that that the tip-rate estimates across the 100 trees were quite variable for both $\lambda_{DR}$ and $\lambda_{BAMM}$ (**Figure S14**).
+Despite this variation, both methods provided similar estimates for speciation in each species for the mean values across the 100 trees ($r = 0.75$) and for the tip-rate estimates from the MCC ($r = 0.68$; **Figure S15**).
 However in comparison to $\lambda_{DR}$, the value of $\lambda_{BAMM}$ was the _mean_ drawn from a posterior distribution (*n* = 1,000) of BAMM generations, thus estimates of $\lambda_{BAMM}$ (and $\mu_{BAMM}$) have an added level of variation.
 To account for this variation, weights (using the inverse of the variance) were used for the PGLS models.
-From 1000 posterior samples of the MCC BAMM run the coefficient of variation (CV) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* **Figures S5 and S7**).
+From 1000 posterior samples of the MCC BAMM run the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* **Figures S5 and S7**).
 Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (**Table S3 and S4**) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (n = 5,966 species) was high.
@@
 Both datasets were used for anlyses with three measures of speciation $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and one measure of extinction $\mu_{BAMM}$ as response variables.
-When using $\lambda_{DR}$ and $\lambda_{ND}$ as the response variable, they provide relatively congruent results; **no need to detail the results in the legend**with a negative relationship between $\lambda_{DR/ND}$ and range size as well as a positive relationship between $\lambda_{DR/ND}$ and sexual selection (**b**).
-Alternatively, when using $\lambda_{BAMM}$ and $\mu_{BAMM}$ the model estimates are more variable across trees compared to the MCC estimate, which do not significantly differ from zero.
 The numerical values for the model estimates using the MCC tree and hpd of estimates from 100 random trees can be found within the ESM.
@@
 However, small range size may be a cause or effect of speciation rate and potentially even an artifact of taxonomic classification.
-Interestingly we found a discrepancy in the model results obtained from tip-rate statistics ($\lambda_{DR}$, $\lambda_{ND}$) and those using tip-rate estimates from a model based approach ($\lambda_{BAMM}$); whereby BAMM produced precise null results for associations between predictors and speciation as well as extinction.
+Interestingly we found a discrepancy in the model results obtained from tip-rate statistics ($\lambda_{DR}$, $\lambda_{ND}$) and those using tip-rate estimates from a model based approach ($\lambda_{BAMM}$); whereby BAMM produced precise null results for most associations between predictors and speciation as well as extinction.
 
@@
 A recent meta-analysis found that converted carotenoids (but not those sourced from diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
-Given the importance of honest signalling in sexual selection and the variability in colour production across birds, sexual dichromatism in many species may unreliable measure of mate quality for female birds and sexual selection for researchers.**Great paragraph!!**
+Given the importance of honest signalling in sexual selection and the variability in colour production across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
 
-In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we find male-bias sexual selection marginally increases speciation rate (at least for $\lambda_{DR}$, $\lambda_{ND}$) yet there is no direct evidence of this association being dependent on variable or stressful environments; albeit a phylogenetic path analysis shows an indirect effect of environmental variability (through range size).
-When evolving under stressful conditions, sexually antagonistic selection is hypothesised to be weakened relative to selection that is conchordant between the sexes [@Connallon_SA_2015; @Connallon_Hall_2016].
+In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we find male-bias sexual selection marginally increases speciation rate (at least for $\lambda_{DR}$ and $\lambda_{ND}$) yet there is no direct evidence of this association being dependent on variable or stressful environments; albeit a phylogenetic path analysis shows an indirect effect of environmental variability (through range size).
+When evolving under stressful conditions, sexually antagonistic selection is hypothesised to be weakened relative to selection that is concordant between the sexes [@Connallon_SA_2015; @Connallon_Hall_2016].
 While these predictions are supported in several lab (experimental evolution) studies [@Long_2012; @Berger_2014; @Punzalan_2013], other studies have not found the association [@Delcourt_2009; @Martinossi-Allibert_Savković_2017; @Holman_Jacomb_2017].
@@
 Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability the species is evolving under.
-Specifically, sexual antagonism may be more easily purged in environments where directional environment pressures are maintained; whereas when environmental variability is cyclical (e.g.
-seasonal or glacial climate cycles) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
+Specifically, sexual antagonism may be more easily purged in environments where cyclic pressures (e.g.
+seasonality) are perpetuated; whereas when environmental variability is directional (e.g.
+long-term climatic change) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
 Secondly, the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
@@
 Challenges arise on determining the causality of range size and speciation; as arguably small range size could either be a cause or effect of speciation.
-** In the next sentence I would start the opposite way, as the model was framed with range affecting speciation instead of the opossite**High speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated taxa.
+Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995].
+Birds with smaller range sizes are also predicted to have lower dispersal ability; in these cases individuals might not be able to overcome geographical barriers and maintain gene flow; thus promoting reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
+Several studies have used morphological measures of wing shape, a proxy for dispersal ability (Hand-wing index/ Kipp's distance), as a predictor of speciation and coexistence [*e.g.*, @Kennedy_2016; @Pigot_2016], but these measurements across passerines is not yet extensive.
+Alternatively, reduced range size may be a consequence of speciation.
+Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives.
 Within Himalayan songbirds range expansions are prevented by the competition for resources by related species [@Price_2014]; however this limitation on range expansion is then expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011].
-Thus, if range size is a cause of speciation we might expect it to be smaller but
-if it is a predictor of speciation it arguably should be larger; with greater opportunities for geographic barriers to form in larger ranges [@Rosenzweig_1995]. 
-** this one is tricky, because yes, what you say above can be true, but also: species with larges ranges mean they have less limits to dispersal, with less restrictions on gene flow and thus lower speciation rates.
-range size can affect speciation rates just because is a proxy for dispersal abilities or something like that.
-The thing is that range size is probably the result of many processes**
-
 Unlike Himalayan songbirds, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from pre-existing taxa.
 Across islands we expect to see a correlation between speciation and small range size, not because of any inherent properties of the range size, rather that they reflect newly formed islands with empty niches **This seems like a very established idea but not sure whose it is**.
-Thus our findings may be heavily dependent on species whose range size is restricted to a single island within an archipelago (*e.g.*, *Geospiza* sp.; aka Darwin's finches).
-Birds with large range sizes are also predicted to have greater dispersal ability; in these cases individuals may be more able to overcome geographical barriers and maintain gene flow; thus lowering speciation in these wide-ranging species [@Birand_2012; @Claramunt_2012].
-**ok, so yeah, like i said, maybe move this to where I mentioned it earlier**Several studies have used morphological measures of wing shape, a proxy for dispersal ability (Hand-wing index/ Kipp's distance), as a predictor of speciation and coexistence [*e.g.*, @Kennedy_2016; @Pigot_2016], but these measurements across passerines is not yet extensive **I think?
-~ 1000**.
+Thus our findings may be heavily dependent on species whose range size is restricted to a single island within an archipelago (*e.g.* *Geospiza* sp.; a.k.a.
+Darwin's finches).
 One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification whereby splitting of subspecies into species in clades with large ranges leads to increased recent phylogenetic branching as well as a more subdivided range **Has this argument been articulated?** **I can't remember reading it, maybe is not true and I made it up haha.
@@
 
-Here we used BAMM to calculate extinction rates across passerines and test whether these rates were associated with sexual dichromatism/selection or various ecological variables.
+Here we used BAMM to calculate extinction rates and test whether these rates were associated with sexual dichromatism/selection or various ecological variables.
 We found no associations in this respect despite the wealth of theoretical predictions on the effects of sexual selection on fitness.
-Extinction is notoriously difficult to obtain accurate estimates for in comparative studies **Cite papers, e.g.
-Rabosky: 'Extinction rates should not be estimated from molecular phylogenies'**, as it is a process that occurs over millennia, leaving scant traces.
+Extinction is notoriously difficult to obtain accurate estimates for in comparative studies [*see* @Rabosky_Extinction_2016], as it is a process that occurs over millennia, leaving scant traces for use in high-power comparative studies.
 One exception to this comes from fossil ostracods, a group of marine invertebrates.
-In these species there is an extensive fossil record of extinct ostracods from the Late Cretaceous epoch (84–66 million years ago) as well as identifiable sexual dimorphism.
-Recently, @Martins_2018 found that species of ostracods with larger sexual dimorphism (representing larger investment by males in reproductive organs) had higher extinction rates; suggesting that sexual selection increases extinction risk.
+In these species there is an extensive fossil record of extinct ostracod species from the Late Cretaceous epoch (84–66 million years ago) as well as identifiable sexual dimorphism.
+Across 93 species, @Martins_2018 found that ostracod taxa with larger sexual dimorphism (representing exaggerated investment by males in reproductive organs) had higher extinction rates; insinuating that sexual selection increases extinction risk.
 Other comparative studies have assessed the relationship between sexual selection and extinction in birds (@Doherty_2003; @Mitra_1996; @Morrow_2003) and mammals [@Bro-Jorgensen_2014; @Morrow_2004] without consensus.
@@
 Given these underestimates of $\mu_{BAMM}$ it is likely we failed to detect true associations between sexual dichromatism/selection and extinction.
-The power of BAMM as a model based-approach in estimating $\lambda$ and $\mu$ is dependent size of the rate regimes.
+The power of BAMM as a model based-approach in estimating $\lambda$ and $\mu$ is dependent on the size of the rate regimes.
 BAMM often produces homogeneous $\lambda_{BAMM}$ and $\mu_{BAMM}$ for clades with few rate shifts, whereas $\lambda_{DR}$ produces higher heterogeneity in tip-rates; arguably this could reflect increased power or sampling error [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
@@
 
-Improvements in estimating tip-rates of speciation in the last few years has allowed us to revisit the association between sexual selection and speciation.
-With high power and alongside extensive ecological predictors we find that male-bias measures of sexual selection better predict speciation than sexual dichromatism.
+Improvements in estimating tip-rates of speciation recently has allowed us to revisit the association between sexual selection and speciation.
+With high power and alongside extensive ecological predictors we find that male-bias measures of sexual selection predict speciation but sexual dichromatism does not.
 If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold with sexual selection a potentially valuable utility in conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].
-Additionally, our finding that high speciation rate is associated with smaller range size has pronounced significance to the future survival of these rapidly speciating species in a world with increased habitat loss and anthropogenic stress.
-Range size is the best predictor of extinction risk [@Harris_2008] and the association found here ensues that many newly speciated clades may be at greater risk of extinction due to anthropogenic effects if not evolutionary or ecological effects.
+Furthermore, our finding that high speciation rate is associated with smaller range size has pronounced significance to the future survival of these rapidly speciating species in a world with increased habitat loss and anthropogenic stress.
+Range size is the best predictor of extinction risk [@Harris_2008] and the association found here ensues that many newly speciated clades may be at greater risk of extinction due to their reduced range being impacted by anthropogenic effects.
 
@@
 <div id="refs"></div>
+
+\newpage
+\nolinenumbers
+
+#Critical analysis
+
+\epigraph{[On Darwinism]...
+But it by no means follows that slight differences in the shape, pattern, or colours of the ornamental plumes are what lead a female to give the preference to one male over another; still less that all the females of a species, or the great majority of them, over a wide area of country, and for many successive generations, prefer exactly the same modification of the colour or ornament.\\
+--- \large\textsc{Alfred Russel Wallace}}
+
+##Strengths and limitations
+
+Perhaps one of the biggest unravelings of this project was the decoupling of male-bias sexual selection from sexual dichromatism.
+Despite the complex statistical models, monsterous sample size and modern sexual conflict theory the most fascinating finding was one that Wallace and Darwin were at odds over 130 years prior [@Kottler_1980].
+The widespread use of sexual dichromatism largely goes unchallanged as our instinctive Darwinian mind reminisces of peacocks and birds of paradise.
+My thoughts were such when I began this scientific inquiry, but now I am sceptical and somewhat disheartened.
+Of course, I was expecting sexual dichromatism from RGB measures to be a noisy measure of sexual selection, but across 5,831 species I anticipated a signal would permeate through the noise.
+And while the correlation between sexual dichromatism and male-bias sexual selection is present, it is surprisingly weak.
+With this in mind, the practical use of sexual dichromatism in macroevolutionary research is severely weakened, with these noisy sexual dichromatism measurements calling dozens of studies into question due to false negative associations.
+This finding is particularly pertenant as efficient proxy measurements are incredibly valuable in comparative analyses; they increase the accessibility, available sample size and transferability of a study making it more conducive to research synthesis.
+But, as Wallace remarks, the matter is inevitabley more complex despite the irresistability of the Darwinian argument^1^ [@Kottler_1980; @Caro_2017].
+On the bright side, the multivariate measure of sexual selection that is the ppca of sexual dimorphism, social polygyny and [lack of] paternal care is a dateset that has great value and would benefit from being expanded across birds.
+
+\maketitle\unmarkedfntext{\textsuperscript{1}Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took more interest in a view where natural selection on predator-prone nesting females generated dullness in females but not males.}
+
+Not only was the measurement of sexual selection under a cloud of uncertainty, the estimates of speciation and extinction are also a point of controversy to many.
+Macroevolutionary research is a field plagued by methedological uncertainty and conflicts.
+And during the second part of this MSc I jumped head-first into the controversy.
+Originally, I had planned to continue my meta-analysis by investigating extinction rate across a large taxanomic group.
+There were signs that extinction rate could be measured from phylogenetic methods with relative confidence **CITE**.
+However, these hopes soon dissapated, and with them the seeming \emph{raison d'\^{e}tre} of the project.
+In measuring extinction rate one can obtain the most valid measures of population fitness in evolutionary biology.
+And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured over long, evolutionarily meaningful timescales [*but see* @Martins_2018].
+We did include measurements of extinction rate within the paper, however these findings were played down.
+The reason for this movement away from extinction and towards speciation is the controversy surrounding extinction rate measures.
+Measuring extinction rates within the bird phylogeny are problematic because of the nature of phylogenetic construction; for passerine birds a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction rate.
+Furthermore, the very detection of an absence of data --- which is what such methods attempt to do --- is a hard sell both conceptually and methodically come peer review; see @Rabosky_2010 for a cautionary tale on measuring extinction rate with ``BAMM``.
+The mechanisms and rates at which speciation and exvtinction occur are an active area of research.
+In a recent pre-print @Diaz_2018 used ``BAMM`` across 104 time-calibrated molecular phylogenies and found that despite variation in ecology, biogeography and taxonomy, younger clades consistently had higher rates of speciation and extinction.
+This novel finding of time-dependency in rates of macroevolutionary diversification makes comparative analyses between challanging and calls into question countless studies.
+Similar to Wallace's view of sexual dichromatism we are left with a view of diversification rate measures where despite the advancements in the field we cannot yet account for the underlting true variation; which is vastly more complex than we can currently meausure.
+
+As with sexual dichromatism and macroevolutionary rates, the most severe limitation for environmental variation was the courseness of the meausres.
+In total, for 5,831 species we extracted 338,203,831 point measures^1^ of environmental variability, despite this our measures may not account for variation in biotic components such as food or predation.
+However, the methods employed here are arguably much better than those implemented by many studies; where the range mid-point is used to obtain a single measure of temperature, precipitation or seasonality.
+**CITE**.
+I would argue that reducing the multivariate data into several biologically meaningful predictors is a conceptually challanging exercise, but researchers should not feel limited by the amount of raw data that they can access^2^.
+In fact, previous to our study on birds the same methodology was applied to ~ 100 species of bovids with speciation, extinction and environmental variability obtained.
+This initially study was prematually halted as there was homogeneity in the response variable (speciation and extinction rates) across the smaller bovid phylogeny.
+However, this workflow was easily transferred over to passerine birds with similar methods for extracting environmental predictors.
+
+\maketitle\unmarkedfntext{\textsuperscript{1}5,831 species x 1000 samples per range x 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.}
+\unmarkedfntext{\textsuperscript{2}This statement applies to species with documented range sizes, which varies across taxa but is comprehensive for birds, mammals, amphibians and many marine groups.
+The \href{http://www.iucnredlist.org/technical-documents/spatial-data}{IUCN} has a large and accessible database for research use.
+Details of how we extracted data from these ranges is provided as a potentially useful resource in the Supplementary Material.}

### 2018-10-02 Sexual Selection and Extinction.Rmd
Critical Analysis and index add

@@
 Given the high variability in tip-rate estimates from the above plot, below we peform some diagnostics on BAMM to demonstrate that the variability is unlikely an error in sampling or parameters, rather an inherent aspect of BAMM.
-The following is not inherently necessary to understandfor the conclusions drawn from the paper, however they do raise a set of methodological questions about BAMM that warrant further investigation.
+The following is not inherently necessary to understand the conclusions drawn from the paper, however they do raise a set of methodological questions about BAMM that warrant further investigation.
 
@@
 
-```{r, fig.height = 8, fig.width = 4, warning=FALSE}
+```{r, warning=F, error = F, fig.height = 8, fig.width = 4}
 BAMM.lambda.pgls.models <- readRDS("data/BAMM.lambda.pgls.models.rds")
@@
 
-BAMM.lambda.pgls.summary.RO <- dcast(BAMM.lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
+BAMM.lambda.pgls.summary.RO <- dcast(BAMM.lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
 BAMM.lambda.pgls.summary.RO$Estimate <- NULL
@@
 #Get rid of outliers
-BAMM.mu.pgls.summary.RO <- dcast(BAMM.mu.pgls.summary %>% select(Parameter,Estimate) %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
+BAMM.mu.pgls.summary.RO <- dcast(BAMM.mu.pgls.summary %>% select(Parameter,Estimate) %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
 BAMM.mu.pgls.summary.RO$Estimate <- NULL
@@
 
-R-squared for relationship between RGB and Spectrophotometry Data:
+R-squared and correlation for relationship between RGB and Spectrophotometry Data:
 ```{r}
-summary(lm(SDi ~ Colour.discriminability,
-   data = MCC.Armenta.model.data))$r.squared %>% pander()
-
-cor(MCC.Armenta.model.data$SDi, MCC.Armenta.model.data$Colour.discriminability)
+data_frame(
+R2 = summary(lm(SDi ~ Colour.discriminability,
+   data = MCC.Armenta.model.data))$r.squared,
+r = cor(MCC.Armenta.model.data$SDi, MCC.Armenta.model.data$Colour.discriminability)) %>% pander()
 ```
@@
 ```{r}
-# summary(lm(SDi ~ Sexual_selection_ppcaColour.discriminability
-#    data = SS.subset))$r.squared %>% pander()
-
-cor(SS.subset$SDi, SS.subset$Sexual_selection_ppca)
+data_frame(
+R2 = summary(lm(SDi ~ Sexual_selection_ppca,
+   data = SS.subset))$r.squared,
+r = cor(SS.subset$SDi, SS.subset$Sexual_selection_ppca)) %>% pander(digits = 2)
 ```
@@
 
-Subset.Lambda.pgls.summary.RO <- dcast(Subset.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
+Subset.Lambda.pgls.summary.RO <- dcast(Subset.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
 Subset.Lambda.pgls.summary.RO$Estimate <- NULL
@@
 
-Subset.Mu.pgls.summary.RO <- dcast(Subset.Mu.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
+Subset.Mu.pgls.summary.RO <- dcast(Subset.Mu.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
 Subset.Mu.pgls.summary.RO$Estimate <- NULL
@@
 
-BAMM.Lambda.100Trees %>% ggplot(aes(x = mean.lambda, y = DR.Lambda.100Trees$DR))+
-  geom_point()+
-  geom_errorbarh(aes(xmin = mean.lambda - 1.96*sd, xmax = mean.lambda + 1.96*sd), size = 0.05)+
-  geom_errorbar(aes(ymin = DR.Lambda.100Trees$DR - 1.96*DR.Lambda.100Trees$sd, ymax = DR.Lambda.100Trees$DR + 1.96*DR.Lambda.100Trees$sd), size = 0.05)+
-  theme_minimal()
+all.trees.join <- left_join(DR.Lambda.100Trees %>% select(TipLabel, DR), BAMM.Lambda.100Trees %>% select(mean.lambda, TipLabel), by = "TipLabel")
+
+BAMM.Lambda.100Trees %>% ggplot(aes(x = DR.Lambda.100Trees$DR, y = mean.lambda, fill = mean.lambda))+
+  geom_point(shape = 21)+
+  geom_errorbar(aes(ymin = mean.lambda - 1.96*sd, ymax = mean.lambda + 1.96*sd), size = 0.02)+
+  geom_errorbarh(aes(xmin = DR.Lambda.100Trees$DR - 1.96*DR.Lambda.100Trees$sd, xmax = DR.Lambda.100Trees$DR + 1.96*DR.Lambda.100Trees$sd), size = 0.02)+
+  theme_minimal()+
+  ylab("100 Trees Speciation Rate (BAMM)")+
+  xlab("100 Trees Speciation Rate (DR)")+
+  scale_fill_viridis_c()
 
@@
 ```
-**Figure S14:**
-
-```{r, eval = FALSE, echo = F}
-all.trees.join <- left_join(DR.Lambda.100Trees %>% select(TipLabel, DR), BAMM.Rates.100Trees %>% select(mean.lambda, TipLabel))
-cor(all.trees.join$DR, all.trees.join$mean.lambda)
-```
-
-```{r}
+**Figure S14:** Speciation Rate means from 100 trees using either the DR statistic (x axis) or BAMM (y axis).
+While there is a clear correlation (*r* = `r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`) there is variability with BAMM showing less heterogeneity.
+95 % CIs are plotted for both axis from the 100 trees.
+Each point represents a species (*n* = 5,965).
+
+```{r, echo = F}
 MCC.BAMM.model.data <- readRDS('data/MCC.BAMM.model.data.rds')
-restricted.data %>% ggplot(aes(x = MCC.DR, y = MCC.BAMM.model.data$mean.lambda))+
-  geom_point()+
-  theme_minimal()
-```
-**Figure S15:**
-
-```{r, eval = FALSE, echo = F}
-MCC.tree.join <- left_join(restricted.data %>% select(MCC.DR, TipLabel), MCC.BAMM.model.data %>% select(mean.lambda, TipLabel))
-cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda)
-```
+
+#cor(all.trees.join$DR, all.trees.join$mean.lambda)
+
+MCC.tree.join <- left_join(restricted.data %>% select(MCC.DR, TipLabel), MCC.BAMM.model.data %>% select(mean.lambda, TipLabel), by = "TipLabel")
+#cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda)
+```
+
+```{r, fig.width = 8, fig.height = 6}
+restricted.data %>% ggplot(aes(x = MCC.DR, y = MCC.BAMM.model.data$mean.lambda, fill = MCC.DR))+
+  geom_point(shape = 21)+
+  theme_minimal()+
+  geom_smooth(method = "lm")+
+  ylab("MCC Speciation Rate (BAMM)")+
+  xlab("MCC Speciation Rate (DR)")+
+  scale_fill_viridis_c()
+
+```
+**Figure S15:** Similar to Figure S14, there is a correlation (*r* = `r round(cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda), digits = 2)`) between $\lambda_{DR}$ and $\lambda_{BAMM}$ with BAMM results showing less heterogeneity.
+Each point represents a species (*n* = 5,965).
 
@@
 
-What if we run the subset df qwith SDi
 ```{r, eval = FALSE, echo = F}
+#Run subset model with SDi
 #Run model for DR

### 2018-10-07 Manuscript.Rmd
Manuscript and Supp Headings Changed

@@
   \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
-  \rhead[]{Sexual Selection and Extinction} \usepackage[noabbrev,capitalise]{cleveref}
+  \rhead[]{Sexual Selection and Speciation} \usepackage[noabbrev,capitalise]{cleveref}
   \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
@@
   \usepackage{hyperref}\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, urlcolor=blue}
+  \raggedbottom
 csl: References/AmNat.csl
-abstract: Sexual selection may influence patterns of diversification. Yet their relationship
-  remains unclear with controversy on whether sexual selection leads to a net increase
-  or decrease in speciation rate. Sexual selection is predicted to increase speciation
-  rate as it can facilitate adaptation to new environments by elevating the average
-  fitness of a population. Additionally, sexual selection may work alongide divergent
-  natural selection and swiflty promote reproductive isolation. Here we use sexual
-  dichromatism as a proxy for sexual selection across 5,812 species of passerine birds
-  and a multivariate male-bias measure of sexual selection (sexual dimorphism, polygyny
-  and paternal care) across 2,465 species to investigate the association of sexual
-  selection with speciation and extinction. Importantly, we test whether the effects
-  of sexual selection on speciation are dependent upon ecological pressures such as
-  temperature variability across time and space, primary productivity and range size.
-  Our results from three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$,
-  $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) find that male-bias
-  sexual selection but not sexual dichromatism predicts speciation. Additionally,
-  we find that birds with smaller ranges have higher speciation rates; however the
-  causality of this finding remains inconclusive.
+abstract: 'Sexual selection may influence patterns of diversification; yet controversy remains on whether sexual selection leads to a net increase or decrease in speciation rate, or even elevates extinction.
+Sexual selection is predicted to increase speciation rate as it can facilitate adaptation to new environments, and work alongside divergent natural selection to swiftly promote reproductive isolation.
+However, sexual selection frequently favours investment in costly or harmful traits that may reduce adaptation affect diversification.
+Here, we investigate the association of sexual selection with speciation and extinction in passerine birds.
+We use two measures of sexual selection: sexual dichromatism (n = 5,812 species) and a multivariate measure of male-biased sexual selection (sexual dimorphism, polygyny and paternal care; n = 2,465), three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$).
+Importantly, we test whether the effects of sexual selection on speciation are dependent upon ecological pressures such as temperature variability across time and space, primary productivity and range size.
+We find that male-biased sexual selection but not sexual dichromatism predicts speciation.
+Additionally, we find that birds with smaller ranges have higher speciation rates; although the direction of causality is unclear.'
 ---
@@
 
-**Process of speciation**: Selection pressures drive speciation.
+Selection pressures drive speciation.
 As populations diverge they accumulate genetic and phenotypic changes that ultimately erect barriers against gene flow [@Presgraves_2010; *but see* @Shaw_2011].
-In this view, populations will speciate and adapt more readily to to new divergent environments if selection pressures are high.
-Hence, the rate at which speciation occurs is likely dependent on both the populations response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour, @Hugall_2012; @Rabosky_morph_2013].
-Sexual selection is an evolutionary force likely to affect fitness components as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition.
-It is therefore hypothesised that sexual selection is a key driver of diversification through shifting the rates of adaptation, divergence, speciation and extinction.
+In this view, populations will speciate and adapt more readily to new divergent environments if selection is strong and/or persistant.
+Hence, the rate at which speciation occurs is likely dependent on both the populations response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour @Hugall_2012; @Rabosky_morph_2013].
+Sexual selection is an evolutionary force likely to affect fitness components, such as reproductive success or viability, as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition.
+Sexual selection is therefore hypothesised to be a key driver of diversification through shifting the rates of adaptation, speciation and extinction.
 
-**Sexual selection and fitness**: Sexual selection is applied by one sex on the other - more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and the strength of sexual selection can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
-Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fixate beneficial mutations in a population [@Whitlock_2000].
+Sexual selection is applied by one sex on the other --- more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and the strength of sexual selection can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in* @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
+Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial mutations in a population [@Whitlock_2000].
 Thus, sexual selection is thought to be a key process in mitigating extinction risk and facilitating adaptation.
-However, sexual selection can increase the fitness of a population when the genotypes under sexual selection (*e.g.* those that increase male mating/fertilisation success) are genetically positively correlated with traits affecting fitness (*e.g.* female fecundity or survival).
-This positive correlation is unlikely to always be the case; regularly, sexual selection favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006].
-For instance, in species with high levels of sexual selection mating success becomes more variable and uncertain, thus individuals must heavily invest to increase their chances of successfully reproducing.
-Through prioritising mating success a sex may invest in costly weaponry and sexual signals, while forgoing strategies that would improve offspring fitness (parental care).
+Sexual selection can also increase the fitness of a population when the genotypes under sexual selection (*e.g.* those that increase male mating/fertilisation success) are genetically positively correlated with traits affecting fitness (*e.g.* female fecundity or survival).
+However, this positive correlation is not ubiquitous; indeed sexual selection frequently favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006].
+For instance, in species with high levels of sexual selection, mating success becomes more variable and uncertain; thus individuals must invest heavily to increase their chances of successfully reproducing.
+High investment into costly weaponry and sexual signals, may trade off against strategies that would improve offspring fitness (*e.g.* parental care).
 This is a costly consequence of sexual selection that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].
 
-**Sexual selection, heritability and mechanisms of speciation**: Alongside changes to fitness, sexual selection has been proposed to facilitate speciation through the evolution of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982].
-For instance, in antbirds (Thamnophilidae), genera with complex songs and stark dichromatism were more species rich [@Seddon_2008].
-Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, and by itself sexual selection could lead to behavioural reproductive isolation [@Safran_2013].
+Alongside changes to fitness, sexual selection has been proposed to facilitate speciation through the evolution of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982].
+For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism, are more species rich [@Seddon_2008].
+Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].
 Concurrently, sexual selection can promote the extent to which populations diverge by enlarging the phenotypic space diverging species can evolve into.
-This prediction emanates from increased genetic diversity as a result of sexual conflict: a consequence of sexual selection [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
+These predictions emerge as sexual selection may increase genetic diversity as a result of trade-offs between naturally- and sexually-selected traits or trade-offs under intra-locus sexual conflict between male and female trait expressions [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
 However, the genetic architecture (*per se* heritability) of fitness and sexually selected traits may vary due to different sex determination systems across taxa.
-In male homogametic species (ZZ/ZW or ZZ/ZO) --- such as birds --- males are anticipated to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003].
-We might therefore expect birds with magnified sexual selection to have increased adaptation and speciation rates.
+In male homogametic species (ZZ/ZW or ZZ/ZO) --- such as birds --- males are predicted to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003].
+We might therefore expect bird populations experiencing stronger sexual selection to have increased adaptation and speciation rates.
 
-**Environmental interaction**: Importantly, the effects of sexual selection on adaptation and speciation may depend on environmental conditions.
-Studies suggest that under stressful/changing environments sexual selection may have greater fitness benefits than under benign environments.
-In stressful environments, the positive correlation between male mating success and female fitness components improves, reducing the burden of negative pleiotropic effects [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017].
-Conversely, under benign environments environmental stability ensures consistent selection, preferentially eroding genetic variation at sexually concordant loci (*i.e.* loci where the optimum genotype is equivelant for both sexes).
-In these cases genetic variation preferentially remains at sexually antagonistic loci; causing displacement of male and female traits from their optimum.
-These predictions would ensue that populations with stronger sexual selection would have increased fitness and adaptive benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
-In certain circumstances sexual selection will reinforce diverging locally adaptated phenotypes through mate choice on phenotypes that advertise quality in a specific ecological context [@Maan_2011].
-If mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to strengthen this positive feedback loop and promote population divergence [@Lande_1981].
-In several birds, beak morphology is a trait that arose under natural selection and is now a trait under sexual selection (a.k.a.
-"magic traits").
+Importantly, the effects of sexual selection on adaptation and speciation may depend on environmental conditions.
+Both theoretical and empirical studies suggest that under stressful/changing environments, sexual selection may have greater fitness benefits than under benign environments.
+Stressful environments strengthen the positive correlation between male mating success and female fitness, reducing the burden of negative pleiotropic effects [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017].
+Conversely, benign or stable environments ensure consistent selection, preferentially eroding genetic variation at sexually concordant loci (*i.e.* loci where the optimum genotype is equivelant for both sexes).
+In this situation, genetic variation remains disproportionately at sexually antagonistic loci; causing displacement of male and female traits from their optimum.
+It follows from these predictions that populations with stronger sexual selection would have increased fitness and adaptive benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
+In certain circumstances, sexual selection will reinforce locally adaptated phenotypes through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011].
+Additionally, if mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to speed-up the evolution of divergent mating signals [@Lande_1981].
+In several bird taxa, beak morphology is a trait that arose under natural selection and is now a trait under sexual selection [a.k.a.
+"magic traits"; @Servedio_2011].
 In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signalling strength in sticklebacks [@Wong_2007; @Candolin_2007].
-In this opposing view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.
+Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signal efficacy in sticklebacks [@Wong_2007; @Candolin_2007].
+In this conflicting view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.
 
-**Previous limitations**: The association between sexual selection and speciation has been investigated using many macroevolutionary approaches across various taxa.
-Importantly, macroevolutionary comparative studies can examine the association between sexual selection and speciation in more diverse taxon, over longer evolutionary time-scales and in more natural environments.
-A meta-analysis of 64 effect sizes from 20 papers found a small, but positively significant association between sexual selecetion and speciation [@Kraaijeveld_2011].
-However, variation in sexual selection proxies used and other methodological choices (species richness vs rate estimates) led to heterogeneity in effect sizes for the correlation between sexual selection and speciation.
-Given that species richness and speciation rate are not neccesarily correlated [*e.g.* @Rabosky_fish_2018] and the recent development and refinement of methodologies in obtaining tip-rate estimates of speciation and extinction [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting with methods that address several issues that have previously limited macroevolutionary comparative studies.
-Firstly, sexual selection strength in macroevolutionary studies are often estimated using measures of dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa; which arguably decreases the power to detect an association, especially if phylogenetic signal is strong.
+The association between sexual selection and speciation has been investigated using many macroevolutionary approaches applied to various taxa.
+A meta-analysis of 64 effect sizes from 20 studies found a small, but significant positive association between sexual selection and speciation [@Kraaijeveld_2011].
+However, there was large variation in effect sizes due to variation in the measures of sexual selection used and other methodological choices (*e.g.* species richness *vs* rate estimates).
+Given that species richness and speciation rate can be inversely correlated [*e.g.* @Rabosky_fish_2018] and the recent refinement of methods to obtain tip-rate diversification estimates [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting.
+Firstly, the strength of sexual selection strength in macroevolutionary studies is often estimated using measures of sexual dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa.
 Secondlly, many studies use incomplete taxon sampling, with family-level analysis restricted to well sampled avian clades [@Seddon_2008; @Huang_2014].
-Thirdly, until recently studies on the association between sexual selection and speciation have largely ignored the contribution of environmental variables to explain variation in speciation rates.
-Furthermore, these studies haven't taken into account either the potential interaction between natural and sexual selection in the generation of new species.
-Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague our ability to draw widely accepted findings.
+Thirdly, until recently, studies on the association between sexual selection and speciation have largely ignored the contribution and interaction with environmental variables to explain variation in speciation rates.
+Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague efforts to identify drivers of speciation and extinction.
 
-**Our study**: We chose to do our study on passerine birds because of the wealth of data available for these (Order: Passeriformes).
-Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset (n = 918) [@Armenta_2008].
-But, here we assess the effects of sexual dichromatism measures based on RGB values (*n* = 5,812) alongside smaller datasets of sexual dichromatism from spectrophotometry measures (*n* = 581) and male-bias sexual selection measures (*n* = 2,465).
-Specifically, we assess the relationship between sexual dichromatism, environmental variability and their interaction with three tip-rate measures of speciation and one measure of extinction ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$).
-Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is are dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species' range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
+Here, we perform a comprehensive analysis of the relationship between sexual selection and speciation or extinction in passerine birds (Order: Passeriformes), arguably the largest group for which there exists comprehensive data.
+Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset [*n* = 918; @Armenta_2008].
+We build on these previous studies by using published datasets of sexual dichromatism measured from images [*n* = 5,812; @Dale_2015] alongside smaller datasets of sexual dichromatism measured from spectrophotometry [*n* = 581; @Armenta_2008] and a composite measure of male-biased sexual selection [*n* = 2,465; @Dale_2015].
+We assess the relationship between sexual selection, environmental variability and their interaction with three tip-rate measures of speciation and one measure of extinction ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$).
+Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species’ range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
 
@@
 Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions.
-Many of the processes undertaken in this study are documented with reproducible code in the electronic supplementary material (ESM).
+Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{electronic supplementary material (ESM)}.
 
@@
 
-We used sexual dichromatism as proxy of sexual selection strength.
-Male and female plumage scores have previously been measured across 5,983 species of passerines [@Dale_2015].
+We used a previously published measure of sexual selection for 5,983 species of passerines [@Dale_2015].
 Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forhead, throat, upper breast and lower breast) from _Handbook of the Birds of the World_ (volumes 8–16) [@delhoyo2011].
-The mean RGB values were calculated and provided male and female plumage scores.
-Here we use the absolute difference between males and female RGB scores as an estimate of sexual dichromatism.
-Additionally, we compiled a colour discriminability measure of sexual dichromatism from spectrophotometer measures [@Armenta_2008].
-These measurements include varition in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from taxanomic samples (as opposed to colour drawings).
+The relative contribution of male and female RGB colour values were averaged across body patches and provide 'male-like' and 'female-like' plumage scores.
+Here we use the absolute difference between males and female plumage scores as an estimate of sexual dichromatism.
+Additionally, we used a measure of dichromatism from spectral data, used to estimate colour distance in avian colour space [@Armenta_2008].
+These measurements include variation in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from taxonomic samples (as opposed to colour drawings).
 However, this spectrophotometry data is not as extensive as the RGB measures; with only 581 passerine species available for this analysis.
@@
 
-Sexual dichromatism has long been associated with levels of sexual selection [@Darwin_1871], with sexual dichromatism often used as a measure of sexual selection in avian taxa [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
-However, sexual dichromatism is likely to account for only partial variation in sexual selection across taxa.
-Here we use an existing dataset [@Dale_2015] of male-bias sexual selection based on the first component of a phylogenetic principle component analysis (ppca) of three characteristics associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
-This measure of male-bias sexual selection is present for only 2,465 species and is weakly correlated with the absolute values for sexual dichromatism using RGB measures ($r = 0.34$; **Figure S10**).
+Sexual dichromatism is a widely used indicator of sexual selection in birds [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014], but only partially captures variation in the strength of sexual selection across taxa.
+Therefore, we also use an existing dataset [@Dale_2015] of male-biased sexual selection based on the first component of a phylogenetic principle component analysis (PPCA) of three characteristics associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
+This measure of male-biased sexual selection is available for only 2,465 species and is weakly correlated with the absolute values of sexual dichromatism using RGB measures ($r = 0.34$; **Figure S10**).
 
@@
 
-Using expert range maps [@birdlife_2017] we obtained estimates of species range size.
-From these distributions we were able to obtain estimates for extensive climatic conditions that each species is exposed to.
+We obtained estimates of species range size using expert range maps [@birdlife_2017].
 Because of taxanomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
-From 1,000 random point samples in each range we extracted 19 present-day bioclimatic variables (representing a variety of biologically relevent annual trends in temperature and precipitation) with 30 seconds (~1 km^2^) spatial resultion [@Fick_2017].
+From these distributions we obtained estimates of climatic conditions that each species is exposed to.
+That is, we took 1,000 random point samples in each range and extracted 19 present-day bioclimatic variables (representing a variety of biologically relevant annual trends in temperature and precipitation) with 30 seconds (~1 km^2^) spatial resolution [@Fick_2017].
 From these values we obtained means and standard deviations for each species.
@@
 Using this data, which we provide as a potentially useful data resource (see ESM), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
-
-##Estimating extinction and speciation.
-
-Phylogenetic information was obtained from www.birdtree.org [@Jetz_2012].
-We used a maximum clade credibility tree (MCC) from 2,500 samples of the posterior distribution that were subset to the passerine bird order (*n* = 5,965) as the main phylogentic tree in our comparative analysis.
-Additionally, a random draw of 100 full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using BAMM.
-These trees used a 'Hackett backbone' and were constructed using the a pure birth (Yule) model.
-We calculated three different tip-rate metrics of speciation and one of extinction across all trees.
-
-Firstly, we obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
-Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [*e.g.*, @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each more basal node having the sum of lengths down-weighted.
-Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification [@Belmaker_2015; @Rabosky_diversification_2018].
-Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events and $\lambda_{ND}$ is a measure of speciation across the root to tip path.
-These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE) models.
-Based on previous comparisons $\lambda_{DR}$ and $\lambda_{ND}$ are anticipated to achieve high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
-
-Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014].
-This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate a probability distribution of evolutionary regimes with variable speciation and extinction rates.
-Importantly, these models provide tip-rate estimates of speciation and extinction rate.
-The parameters of the 100 BAMM runs are detailed in full in the ESM; briefly we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
-Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPU's.
-We obtained effective sample sizes for log-likelihood and number of rate shifts over 200 for each run (**Table S3**, **Table S4**).
-Further details of BAMM parameters and output are available through ESM, with tip-rate means and variances included in the ESM.
-Additionally, given the variability in BAMM estimates we also provide some analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on an MCC from a genetic-only phylogeny across all birds [@Harvey_2017].
-All analyses were conducted on log-rates.
-
-##Statistical analysis
 
@@
 To obtain long-term variation in climates for each species range we take the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
-Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates  (**Table S2**, **Figure S2**).
-Despite similar theoretical underpinnings the predictors of environmental variability used in PGLS analyses are not strongly correlated (**Figure S3**).
+Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates  (**Table S2**, **Figure S2**).The five predictors of environmental variability are not strongly correlated (**Figure S3**).
 Details and code to generate these predictors can be found within the ESM.
+
+##Estimating extinction and speciation.
+
+Phylogenetic information was obtained from www.birdtree.org [@Jetz_2012].
+We used a maximum clade credibility (MCC) tree from 2,500 samples of the posterior distribution that were subset to the passerine bird order (*n* = 5,965) as the main phylogentic tree in our comparative analysis.
+Additionally, a random draw of 100 full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures and BAMM [@Rabosky_2014].
+These trees used a 'Hackett backbone' [@Hackett_2008] and were constructed using the a pure birth (Yule) model.
+We calculated three different tip-rate metrics of speciation and one of extinction across all trees.
+
+Firstly, we obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
+Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [*e.g.*, @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each more basal node having the sum of lengths down-weighted.
+Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification [@Belmaker_2015; @Rabosky_diversification_2018].
+Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events and $\lambda_{ND}$ is a measure of speciation across the root to tip path.
+These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE) models.
+Based on previous simulation studies $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
+
+Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014].
+This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate a probability distribution of evolutionary rate-shift configurations with variable speciation and extinction rates.
+Importantly, these models provide tip-rate estimates of speciation and extinction rate.
+The parameters of the 100 BAMM runs are detailed in full in the ESM; briefly we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
+Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPU's.
+Each run of BAMM reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200 (an arbitrary value, above which phylogeneticists believe posterior tree distributions can be accurately inferred) for the key model parameters of log-likelihood and number of rate shifts (**Table S3**, **Table S4**).
+Further details of BAMM parameters and output are available in the ESM, with tip-rate means and variances provided.
+Additionally, given the variability in BAMM estimates, we also provide some analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on the MCC tree from a genetic-only phylogeny across all birds [@Harvey_2017].
+All analyses were conducted on log-rates.
+
+##Statistical analysis
 
@@
 
-To test the association between speciation/extinction and sexual selection, environmental variability and its interaction we used phylogenetic least squares (PGLS) models in the ``nlme`` package [@nlme_2018].
-We began by estimating the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using ``corPagel`` in the ``ape`` package [@ape_2004] on a MCC tree.
+To test the association between speciation/extinction and sexual selection, environmental variability and their interaction, we used phylogenetic least squares (PGLS) models in the ``nlme`` package [@nlme_2018].
+We began by estimating the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using ``corPagel`` in the ``ape`` package [@ape_2004] on the MCC tree.
 The estimate of $\lambda$ was then fixed for model selection, which was conducted on the same MCC tree derived from 2,500 draws of the posterior distribution [@Jetz_2012].
-Model selection used either $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ tip-rate estimates from the MCC as the response variable and only compared interaction terms.
-That is, we fixed the individual predictors of: sexual dichromatism/selection, log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation and NPP while comparing 32 models with different combinations of interactions (including none).
+Model selection used either $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ tip-rate estimates from the MCC tree as the response variable.
+For $\lambda_{BAMM}$ and $\mu_{BAMM}$ we were also able to use model weights sourced from the inverse of the variance of tip-rates for a given species from the posterior distribution (*n* = 1,000).
+Using model selection we only compared interaction terms between a measure of sexual selection (sexual dichromatism or male-biased sexual selection) and environmental measures.
+That is, we fixed the individual predictors of: dichromatism/male-biased sexual selection measures, log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation and NPP while comparing 32 models with different combinations of interactions (including none).
 Model selection was done in ``MuMIn`` using the ``dredge`` function [@MuMIn_2017].
-Using the terms from the top-ranking model (lowest AICc) we ran the equivalent model in each of the 100 phylogenetic trees, using the unique response variables and phylogenetic tree correlation structure in each model.
-This method gave us the ability to present model estimates for an MCC alongside 100 random draws; acknowledging the potential role phylogenetic uncertainty has in tapering conclusions drawn from a random sample of the data.
-This method was conducted on models that used data from all available passerine species (*n* = 5,812).
-The same procedure decribed above was applied to a subset dataset of passerine birds (*n* = 581), where sexual dichromatism measures were sourced from spectrophotometry measures [@Armenta_2008] and where male-bias sexual selection ppca was used as a predictor instead of sexual dichromatism (*n* = 2,465).
-Finally, using this data subsetted for species with measurements of male-bias sexual selection we conducted a phylogenetic path analysis using the ``phylopath`` ``R`` package [@Wouter_2018].
+Using the terms from the top-ranking model (lowest AICc), we ran the equivalent model for each of the 100 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$, using the unique response variables and phylogenetic tree correlation structure in each model.
+This method enabled us to present model estimates for an MCC tree alongside 100 trees from the posterior distribution of trees to account for phylogenetic uncertainty.
+This approach was repeated on three datasets corresponding to each measure of sexual selection: dichromatism derived from RGB values of images (*n* = 5,812); dichromatism from spectrophotometry (*n* = 581) and the multivariate measure of male-biased sexual selection (*n* = 2,465).
+
+Finally, using data subsetted for species with measurements of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
 The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
+That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection and range size to have effects on each other and not just on speciation rate.
 To minimise path complexity we use temperature seasonality (BIO4) as the single measure for environmental variability and use $\lambda_{DR}$ as the single measure of speciation.
@@
 
-PGLS models with interaction terms between the respective measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in model selection.
-However, in all model selection phases for different measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and different measures of sexual selection (RGB values, spectrophotometry and male-biased sexual selection) no interaction terms were significant or present in the top models ($\delta$ AICc > 4; **Table S5**, **Table S6**, **Table S10**, **Table S13**).
-Hence we find no evidence that the effect of sexual selection on speciation or extinction is dependent upon certain environmental conditions.
-Furthermore, we find no support that these individual bioclimatic factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) act as drivers of speciation independently from sexual dichromatism/selection (\autoref{model_results}, **Figure S10**).
+PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in model selection.
+However, no interaction terms were significant or present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; **Table S5**, **Table S6**, **Table S10**, **Table S13**).
+Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
+Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, **Figure S10**).
 
@@
 We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation.
-Using three response measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$), the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from a MCC tree.
+Using three response measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$), the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree.
 Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
 Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-PGLS analyses using subsetted data (n = 581), where sexual dichromatism was measured using spectrophotometry distance measures [@Armenta_2008] yielded results concordant to the full data set; that is, no association between sexual dichromatism and speciation or extinction (**Figure S10**).
+PGLS analyses using subsetted data (n = 581), where sexual dichromatism was measured using spectrophotometry [@Armenta_2008] yielded results concordant to the full data set; that is, no association between sexual dichromatism and speciation or extinction (**Figure S10**).
 Our results from MCC models are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 100 random trees.
-Model estimates are normally distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, See HPD intervals in **Table S8**).
-For PGLS using the subsetted dataset where sexual dichromatism was determined with spectrophotmetry [@Armenta_2008] the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (`r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (**Table S12**).
+Model estimates are normally distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, see HPD (highest posterior density) intervals in **Table S8**).
+For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (`r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (**Table S12**).
 
-##Male-bias sexual selection increases speciation
+##Male-biased sexual selection increases speciation
 
-When we conducted PGLS analysis using data subsetted for values of sexual selection estimated from a phylogenetic PCA of sexual dimorphism, social polygyny and paternal care (*n* = 2,465) there was a significant positive association between sexual selection and $\lambda_{DR}$ ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-However, this positive finding was non-significant between sexual selection and $\lambda_{ND}$ ($\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b) as well as between sexual selection and $\lambda_{BAMM}$ ($\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-For PGLS models on 100 random trees, we find the distribution to be similar to the estimates when an MCC tree is used.
-Among the 100 tress we find a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a positive (but less so) association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
-Complete HPD intervals for models using male-bias sexual selection ppca as a predictor can be found within **Table S15**.
+We found a significant positive association between male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+However, this associate was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+The distribution of estimates from PGLS models on 100 random trees was similar to the estimate from an MCC tree: among the 100 tress there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
+Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within **Table S15**.
 
@@
 
-Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation we find a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$.
-Model estimates from PGLS models using an MCC tree for determining tip-rates and correlation structures place this negative association as small but significant ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a).
+Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation we found a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$.
+This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}).
 This association was also evident across the estimates from models using the 100 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a).
-Subsetted models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- show variable evidence that range size is negatively associated with speciation.
-Data subsetted based on the sexual selection ppca (*n* = 2,465) show that smaller range size significantly predicted increased $\lambda_{DR}$ (\autoref{model_results}b), yet this negative association was not significant for $\lambda_{ND}$ and near-zero for $\lambda_{BAMM}$.
-Alternatively, models using subsetted data (*n* = 581), where sexual dichromatism was measured using spectrophotometry distance measures [@Armenta_2008] gave non-significant estimates for the effect of range size on all measures of speciation (**Figure S10, Table S11, Table S2**).
+Subsetted models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed variable evidence that range size is negatively associated with speciation.
+Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subsetted for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
+Model using data subsetted for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (**Figure S10, Table S11, Table S2**).
 Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with highest power using near-complete taxon sampling (*n* = 5,812).
@@
 
-Using a phylogenetic path analysis we found there to be multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; **Figure S13**).
-Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from sexual selection to sexual dichromatism was relatively weak given the expectations of dichromatism being a well utilised proxy of sexual selection ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
-Additionally, because of the large direct effect of temperature seasonality on range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`) and our findings from the PGLS that range size is negatively associated with $\lambda_{DR}$, there is an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}).
+Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; **Figure S13**).
+Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from male-biased sexual selection to sexual dichromatism was relatively weak, ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`), despite dichromatism being widely used as an indicator of the strength of sexual selection.
+Additionally, the large direct effect of temperature seasonality on range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`) suggests an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}), given the negative we identified between $\lambda_{DR}$ and range size in PGLS models.
 
@@
 
-We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-This null result was consistent for spectrophotometry measures of sexual dichromatism (**FigureS10, Table S11, Table S12**) and for the estimates of sexual selection using the sexual selection ppca (\autoref{model_results}b, **Table S14, Table S15**).
+We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (**FigureS10, Table S11, Table S12**) or male-biased sexual selection (\autoref{model_results}b, **Table S14, Table S15**).
 
@@
 
-We tested our hypothesis that sexual selection is associated with speciation using several methods and across 100 samples of phylogentic trees drawn from the post burn-in posterior [@Jetz_2012].
-The range of methods utilised were implemented to overcome uncertainty in the power and precision of various tip-rate estimates as well as the variation between trees.
+We tested our hypothesis that sexual selection is associated with speciation using several methods and across 100 trees drawn from the post burn-in posterior [@Jetz_2012].
+We used a range of methods to overcome uncertainty in the power and precision of various tip-rate estimates as well as the variation between trees.
 We found that that the tip-rate estimates across the 100 trees were quite variable for both $\lambda_{DR}$ and $\lambda_{BAMM}$ (**Figure S14**).
-Despite this variation, both methods provided similar estimates for speciation in each species for the mean values across the 100 trees ($r = 0.75$) and for the tip-rate estimates from the MCC ($r = 0.68$; **Figure S15**).
-However in comparison to $\lambda_{DR}$, the value of $\lambda_{BAMM}$ was the _mean_ drawn from a posterior distribution (*n* = 1,000) of BAMM generations, thus estimates of $\lambda_{BAMM}$ (and $\mu_{BAMM}$) have an added level of variation.
+Despite this, tip-rate estimates of speciation from the two methods were moderately correlated across the 100 trees ($r = 0.75$) and for the MCC tree ($r = 0.68$; **Figure S15**).
+However, in comparison to $\lambda_{DR}$, the value of $\lambda_{BAMM}$ was the _mean_ drawn from a posterior distribution (*n* = 1,000) of BAMM generations, thus estimates of $\lambda_{BAMM}$ (and $\mu_{BAMM}$) have an added level of variation.
 To account for this variation, weights (using the inverse of the variance) were used for the PGLS models.
 From 1000 posterior samples of the MCC BAMM run the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* **Figures S5 and S7**).
-Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (**Table S3 and S4**) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (n = 5,966 species) was high.
-This means that although the number of rate shifts reached convergence and is more precise ($median = 59$; **Figure S6**) the locations of the rate shifts (*i.e.* the credible shift set) across the tree is highly heterogeneous.
+Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (**Table S3 and S4**) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) was high.
+This means that although the number of rate shifts reached convergence ($median = 59$; **Figure S6**), the locations of the rate shifts (*i.e.* the credible shift set) across the tree are highly heterogeneous.
 The variability in the locations for the shift configurations is a likely source of uncertainty in downstream tip-rates used as the response variable in PGLS models.
-Notably the model estimate BAMM produced varied a lot between trees relative to the confidence interval of the MCC tree, with 95 % HPD interval being approximately 20-fold wider than the MCC 95 % CI for models using BAMM, whereas for $\lambda_{DR}$ and $\lambda_{ND}$ the HPD 95 % interval width was equal to or less than the MCC 95 % CI (\autoref{model_results}, **Table S9**).
+Notably, the 95 % HPD interval for model estimates using BAMM rates from 100 trees was about 20-fold the 95 % confidence intervals of the estimate from the MCC tree; whereas for $\lambda_{DR}$ and $\lambda_{ND}$ the HPD 95 % interval width was equal to or less than the MCC 95 % CI (\autoref{model_results}, **Table S9**).
 
-![Model estimates for (**a**) PGLS analyses using sexual dichromatism measures (*n* = 5,812) and (**b**) PGLS analyses using a subsetted male-bias measure of sexual selection (*n* = 2,465).
-Both datasets were used for anlyses with three measures of speciation $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and one measure of extinction $\mu_{BAMM}$ as response variables.
-The numerical values for the model estimates using the MCC tree and hpd of estimates from 100 random trees can be found within the ESM.
-Density curves are based on estimates from 100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC.
-For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to interperet the MCC 95 % CIs.
+![Model estimates for (**a**) PGLS analyses using sexual dichromatism (*n* = 5,812) and (**b**) PGLS analyses using measures of male-biased sexual selection (*n* = 2,465).
+Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
+The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 100 random trees can be found within the ESM.
+Density curves are based on estimates from 100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC tree.
+For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to interpret the MCC 95 % CIs.
 \label{model_results}](Figures/ModelEstimates.pdf)
@@
 ![Speciation rate ($\lambda_{DR}$) across all passerine birds (*n* = 5,965) with estimates of sexual dichromatism and range size available for 5,812 of them.
-Across these species there was a small but significant negative association between $\lambda_{DR}$ and log-range size but no significant association between $\lambda_{DR}$ and sexual dichromatism based on RGB measurments.
+Across these species there was a small but significant negative association between $\lambda_{DR}$ and log-range size but no significant association between $\lambda_{DR}$ and sexual dichromatism based on RGB.
 $\lambda_{DR}$ are those from the MCC tree and images of birds are from the _Handbook of the Birds of the World_.
 Clockwise the species are: _Sporophila bouvronides_, _Geospiza magnirostris_, _Atlapetes leucopterus_, _Euphonia rufiventris_, _Euplectes franciscanus_,_Promerops gurneyi_, _Phainopepla nitens_, _Zosterops erythropleurus_, _Acrocephalus australis_, _Progne cryptoleuca_, _Mayrornis schistaceus_, _Paradisaea rubra_, _Pachycephala pectoralis_, _Malurus pulcherrimus_, _Cranioleuca curtata_, _Cercomacra manu_, _Muscisaxicola capistratus_, _Lepidothrix coeruleocapilla_.
-Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation and should not be interpereted.
+Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation and should not be interpreted.
 \label{Speciation_Figure}](Figures/DRSpeciationFigure.pdf)
@@
 Arrows represent direct effects with numeric values corresponding to colours (blue = positive, red = negative).
-The numeric values are standardised regression slopes and the asterix indicates that the 95 % confidence intervals of this estimate do not overlap with zero.
-The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and male-bias sexual selection measures (*n* = 2,465).
+The numeric values are standardised regression slopes and the asterisks indicates that the 95 % confidence intervals of this estimate do not overlap with zero.
+The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and male-biased sexual selection measures (*n* = 2,465).
 \label{Path_Analysis}](Figures/Path_Analysis.pdf)
@@
 This result is consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and two measures of dichromatism (spectral and RGB).
-However, we did find that when using a male-bias measure of sexual selection --- instead of sexual dichromatism --- there was a positive association with speciation: with significance varying across different rate estimates.
-Our findings suggest that the components that were used in the ppca of male-bias sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care) are better predictors of speciation than sexual dichromatism and provide discussion points for the contentious topic of sexual selection proxies.
+However, we did find that when using a measure of male-biased sexual selection --- instead of sexual dichromatism --- there was a positive association with speciation, though the strength of this relationship varied across different rate estimates.
+Our findings suggest that the components of the composite measure of male-biased sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care) are better predictors of speciation than sexual dichromatism.
 No bioclimatic measures of environmental variability (temperature seasonality, long-term temperature variation and spatial temperature variation) predicted speciation, but range size was negatively associated with two measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$).
-This finding suggests that species with smaller range sizes have increased speciation rate.
-However, small range size may be a cause or effect of speciation rate and potentially even an artifact of taxonomic classification.
-Interestingly we found a discrepancy in the model results obtained from tip-rate statistics ($\lambda_{DR}$, $\lambda_{ND}$) and those using tip-rate estimates from a model based approach ($\lambda_{BAMM}$); whereby BAMM produced precise null results for most associations between predictors and speciation as well as extinction.
+This suggests that species with smaller range sizes have increased speciation rate, however, small range size may be a cause or effect of elevated speciation rate and potentially even an artifact of taxonomic classification.
+Interestingly we found a discrepancy in the model results obtained from tip-rate statistics ($\lambda_{DR}$, $\lambda_{ND}$) and those using tip-rate estimates from a model-based approach ($\lambda_{BAMM}$); whereby BAMM produced precisely null results for most associations between predictors and speciation as well as extinction.
 
-Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015] we find that sexual dichromatism does not predict speciation rate as sexual selection does.
-The association between sexual selection and sexual dichromatism in birds is a Darwinian trope [@Darwin_1871] that allows it to be commonly used as a proxy for sexual selection in comparative studies [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
-However, the presence of high sexual dichromatism in bird species may be also dependent on ecological selection pressures.
-Ecological pressures could drive sexual dichromatism in the absence of mate choice if sex-specific niches provide opportunity for the evolution of sex-limited traits [@Kottler_1980; @Slatkin_1984; @Shine_1989].
+The association between sexual selection and sexual dichromatism in birds is a Darwinian trope [@Darwin_1871] that allows it to be commonly used as a proxy for the strength of sexual selection in comparative studies [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
+Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015] we find that sexual dichromatism does not predict speciation rate, whereas male-biased sexual selection does.
+Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection, and that variance in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
+There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is enigmatic.
+Firstly, ecological pressures could drive sexual dichromatism in the absence of mate choice if sex-specific niches provide opportunity for the evolution of sex-limited traits [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989].
 In damselflies, differences between the sexes in their exposure to UV radiation drives dichromatism between the sexes [@Cooper_2010; @Punzalan_2010].
-Ecological impacts on sexual dichromatism are supported by our path analysis; which reveals that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation) albeit by a relatively small amount.
-Given that our results show that speciation rate increases with sexual selection strength, but not sexual dichromatism, we are led to conclude that sexual dichromatism is a seemingly poor proxy of sexual selection and that the noise associated with the proxy measurement of sexual selection cancels out the obtained signal; this is despite the high power of our comparative study.
-The usefulness of sexual dichromatism as a proxy of sexual selection may be dependent upon the metabolic pathways that give rise to plumage colouration; with not all plumage colouration reflecting mate quality.
+Ecological impacts on sexual dichromatism are supported by our path analysis, which reveals that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation) albeit relatively weakly.
+Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition.
+And thirdly, not all plumage colouration honestly reflects mate quality.
 A recent meta-analysis found that converted carotenoids (but not those sourced from diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
-Given the importance of honest signalling in sexual selection and the variability in colour production across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
+Given the importance of honest signaling in sexual selection and the variability in colour production across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
 
-In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we find male-bias sexual selection marginally increases speciation rate (at least for $\lambda_{DR}$ and $\lambda_{ND}$) yet there is no direct evidence of this association being dependent on variable or stressful environments; albeit a phylogenetic path analysis shows an indirect effect of environmental variability (through range size).
+In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we find that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ --- but this association was not depenendent on variable or stressful environments.
 When evolving under stressful conditions, sexually antagonistic selection is hypothesised to be weakened relative to selection that is concordant between the sexes [@Connallon_SA_2015; @Connallon_Hall_2016].
@@
 Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability the species is evolving under.
-Specifically, sexual antagonism may be more easily purged in environments where cyclic pressures (e.g.
-seasonality) are perpetuated; whereas when environmental variability is directional (e.g.
-long-term climatic change) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
+Specifically, sexual antagonism may be more easily purged in environments where cyclic pressures (*e.g.* seasonality) are perpetuated; whereas when environmental variability is directional (*e.g.* long-term climatic change) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
 Secondly, the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
-Specifically, access to dietary resources, and the impacts of predation or parasitism are unaccounted for here: arguably these are key processes affecting sexual selection and speciation [*reviewed in*, @Maan_2011].
-Thirdly, it is also likely that there is no effect; at least not in the species investigated here.
+Specifically, access to dietary resources, and the impacts of predation or parasitism are unaccounted for here: arguably these are key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
+Thirdly, it is also possible that there is no effect of environmental variability on speciation rates; at least not in the species investigated here.
 
@@
 
-The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates.
-But drawing conclusions from this finding is fraught with challenges.
-Challenges arise on determining the causality of range size and speciation; as arguably small range size could either be a cause or effect of speciation.
+The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates; however reduced range size can be a cause or consequence of speciation.
 Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995].
-Birds with smaller range sizes are also predicted to have lower dispersal ability; in these cases individuals might not be able to overcome geographical barriers and maintain gene flow; thus promoting reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
-Several studies have used morphological measures of wing shape, a proxy for dispersal ability (Hand-wing index/ Kipp's distance), as a predictor of speciation and coexistence [*e.g.*, @Kennedy_2016; @Pigot_2016], but these measurements across passerines is not yet extensive.
+However, birds with smaller range sizes are also predicted to have lower dispersal ability.
+This ensues that individuals might not be able to overcome geographical barriers and maintain gene flow; which would promote reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
+Several studies have used morphological measures of wing shape (Hand-wing index/ Kipp's distance) --- a proxy for dispersal ability --- as predictors of speciation and coexistence [*e.g.*, @Kennedy_2016; @Pigot_2016].
+But currently published datasets of these measurements are not available for many passerines.
 Alternatively, reduced range size may be a consequence of speciation.
 Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives.
-Within Himalayan songbirds range expansions are prevented by the competition for resources by related species [@Price_2014]; however this limitation on range expansion is then expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011].
-Unlike Himalayan songbirds, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from pre-existing taxa.
-Across islands we expect to see a correlation between speciation and small range size, not because of any inherent properties of the range size, rather that they reflect newly formed islands with empty niches **This seems like a very established idea but not sure whose it is**.
-Thus our findings may be heavily dependent on species whose range size is restricted to a single island within an archipelago (*e.g.* *Geospiza* sp.; a.k.a.
+Within Himalayan songbirds, range expansions are prevented by the competition for resources by related species [@Price_2014]; however, this limitation on range expansion is then expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011].
+Unlike Himalayan songbirds, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
+Across islands we expect to see a correlation between speciation and small range size, not because of any inherent properties of the range size, but because they reflect newly formed islands with empty niches.
+Thus, our findings may be heavily dependent on species whose range size is restricted to a single island within an archipelago (*e.g.* *Geospiza* sp.; a.k.a.
 Darwin's finches).
-One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification whereby splitting of subspecies into species in clades with large ranges leads to increased recent phylogenetic branching as well as a more subdivided range **Has this argument been articulated?** **I can't remember reading it, maybe is not true and I made it up haha.
-Makes sense to me but maybe good to see what reviewers would think about it.
-I guess it would also depend on whether the taxa with large ranges are considered the same species because there is actually no strong genetic differentiation across the range and not because there isn't enough information (e.g.sampling) to split a species with massive range into smaller ranges and more species** .
-The role of range size in speciation may be further investigated through ancestral geographic range reconstruction [@Quintero_2015], as ancestral ranges may be beneficial in understanding modes of speciation and the causality of range size expansions and contractions [@Losos_2003].
+One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
 
-Here we used BAMM to calculate extinction rates and test whether these rates were associated with sexual dichromatism/selection or various ecological variables.
-We found no associations in this respect despite the wealth of theoretical predictions on the effects of sexual selection on fitness.
-Extinction is notoriously difficult to obtain accurate estimates for in comparative studies [*see* @Rabosky_Extinction_2016], as it is a process that occurs over millennia, leaving scant traces for use in high-power comparative studies.
-One exception to this comes from fossil ostracods, a group of marine invertebrates.
-In these species there is an extensive fossil record of extinct ostracod species from the Late Cretaceous epoch (84–66 million years ago) as well as identifiable sexual dimorphism.
-Across 93 species, @Martins_2018 found that ostracod taxa with larger sexual dimorphism (representing exaggerated investment by males in reproductive organs) had higher extinction rates; insinuating that sexual selection increases extinction risk.
-Other comparative studies have assessed the relationship between sexual selection and extinction in birds (@Doherty_2003; @Mitra_1996; @Morrow_2003) and mammals [@Bro-Jorgensen_2014; @Morrow_2004] without consensus.
-Yet these methods often rely on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List [*e.g.*, @Morrow_2003; @Bro-Jorgensen_2014; @Morrow_2004] and may be heavily biased to species under greater anthropogenic threat, culturally valued or well researched.
-While estimating extinction risk from the phylogeny may not come with these biases, they are riddled with others [*reviewed in*, @Pyron_2016].
-BAMM has been subject to controversy in its ability to model rate shifts and extinction rates [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], and while several tip-rate estimates exist for speciation (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$) extinction estimates are not as easily obtained.
-Here we anticipated extinction rates would be underestimated because we used a phylogeny constructed with a birth-only (Yule) model [@Jetz_2012].
-Given these underestimates of $\mu_{BAMM}$ it is likely we failed to detect true associations between sexual dichromatism/selection and extinction.
-The power of BAMM as a model based-approach in estimating $\lambda$ and $\mu$ is dependent on the size of the rate regimes.
-BAMM often produces homogeneous $\lambda_{BAMM}$ and $\mu_{BAMM}$ for clades with few rate shifts, whereas $\lambda_{DR}$ produces higher heterogeneity in tip-rates; arguably this could reflect increased power or sampling error [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
-Thus we are somewhat uncertain in the validity of conclusions from $\mu_{BAMM}$ estimates as well as $\lambda_{BAMM}$ and $\lambda_{DR}$ given their discrepancy.
+In addition to speciation, sexual selection has frequently been invoked as a driver of extinction.
+Sexual selection may increase extinction risk as it often leads individuals (usually males) to invest in exaggerated mating signals as opposed to traits --- such as parental care --- that elevate offspring fitness [*reviewed in* @Kokko_Brooks_2003].
+Using BAMM, we found no association between extinction rate and sexual dichromatism, male-biased sexual selection or measures of environmental variability.
+Extinction is notoriously hard to estimate accurately [*see* @Rabosky_Extinction_2016].
+It can be estimated from phylogenetic inference, fossil record, records of recent extinctions and IUCN extinction threat status.
+Phylogenetic methods --- like BAMM --- allow for speciation and extinction rates to be measured using moderately sized phylogenies.
+However, BAMM has been subject to controversy in its ability to model rate shifts and extinction rates [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], and while several tip-rate estimates exist for speciation (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$) extinction estimates are not as easily obtained.
+Here, we found that BAMM often produced homogeneous $\lambda_{BAMM}$ and $\mu_{BAMM}$ for clades with few rate shifts; arguably this could reduce our power to detect extinction rates [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
+The ability to detect extinction rates rarely improves when fossil records are used, as many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
+But, there are exceptions, ostracods --- a group of marine invertebratesone --- have extensive fossil records and show identifiable sexual dimorphism.
+Across 93 species of ostracods, @Martins_2018 found that ostracod taxa with larger sexual dimorphism (representing exaggerated investment by males in reproductive organs) had higher extinction rates; implying that sexual selection increases extinction risk.
+Alternatively, when comparative studies have assessed the relationship between sexual selection and extinction in extant birds [@Doherty_2003; @Mitra_1996; @Morrow_2003] and mammals [@Bro-Jorgensen_2014; @Morrow_2004] there has been no consensus.
+Unfortunately, these methods often rely on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List [*e.g.*, @Morrow_2003; @Bro-Jorgensen_2014; @Morrow_2004] and may be biased to species under greater anthropogenic threat, culturally valued or well researched.
 
-Improvements in estimating tip-rates of speciation recently has allowed us to revisit the association between sexual selection and speciation.
-With high power and alongside extensive ecological predictors we find that male-bias measures of sexual selection predict speciation but sexual dichromatism does not.
+Improvements in estimating tip-rates of speciation has allowed us to revisit the association between sexual selection and speciation.
+With high power and alongside extensive ecological predictors we find that male-biased sexual selection predicts speciation but sexual dichromatism does not.
 If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold with sexual selection a potentially valuable utility in conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].
@@
 Despite the complex statistical models, monsterous sample size and modern sexual conflict theory the most fascinating finding was one that Wallace and Darwin were at odds over 130 years prior [@Kottler_1980].
-The widespread use of sexual dichromatism largely goes unchallanged as our instinctive Darwinian mind reminisces of peacocks and birds of paradise.
-My thoughts were such when I began this scientific inquiry, but now I am sceptical and somewhat disheartened.
+The widespread use of sexual dichromatism largely goes unchallenged as our instinctive Darwinian mind reminisces of peacocks and birds of paradise.
+My thoughts were such when I began this scientific inquiry, but am now skeptical and somewhat disheartened.
 Of course, I was expecting sexual dichromatism from RGB measures to be a noisy measure of sexual selection, but across 5,831 species I anticipated a signal would permeate through the noise.
-And while the correlation between sexual dichromatism and male-bias sexual selection is present, it is surprisingly weak.
-With this in mind, the practical use of sexual dichromatism in macroevolutionary research is severely weakened, with these noisy sexual dichromatism measurements calling dozens of studies into question due to false negative associations.
-This finding is particularly pertenant as efficient proxy measurements are incredibly valuable in comparative analyses; they increase the accessibility, available sample size and transferability of a study making it more conducive to research synthesis.
-But, as Wallace remarks, the matter is inevitabley more complex despite the irresistability of the Darwinian argument^1^ [@Kottler_1980; @Caro_2017].
-On the bright side, the multivariate measure of sexual selection that is the ppca of sexual dimorphism, social polygyny and [lack of] paternal care is a dateset that has great value and would benefit from being expanded across birds.
+And despite the correlation between sexual dichromatism and male-bias sexual selection is present, the relationship is surprisingly weak (see Figure 3 in main text).
+With this in mind, the practical use of sexual dichromatism in macroevolutionary research is severely weakened.
+This finding is particularly pertinent as efficient proxy measurements are incredibly valuable in comparative analyses: that's because they increase the accessibility, available sample size and transfer-ability of a study making it more conducive to research synthesis.
+But, as Wallace remarks, the matter is inevitably more complex despite the irresistibility of the Darwinian argument\footnote{Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took more interest in a view where natural selection on predator-prone nesting females generated dullness in females but not males.} [@Kottler_1980; @Caro_2017].
+On the bright side, the multivariate measure of sexual selection that is the PPCA of sexual dimorphism, social polygyny and [lack of] paternal care is a dataset that better predicts speciation and would benefit from being expanded across birds.
 
-\maketitle\unmarkedfntext{\textsuperscript{1}Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took more interest in a view where natural selection on predator-prone nesting females generated dullness in females but not males.}
+<!-- \maketitle\unmarkedfntext{\textsuperscript{1}} -->
 
-Not only was the measurement of sexual selection under a cloud of uncertainty, the estimates of speciation and extinction are also a point of controversy to many.
-Macroevolutionary research is a field plagued by methedological uncertainty and conflicts.
+Not only are the measurement of sexual selection under a cloud of uncertainty, the estimates of speciation and extinction are another point of controversy.
+Macroevolutionary research is a field plagued by methodological uncertainty and conflicts.
 And during the second part of this MSc I jumped head-first into the controversy.
-Originally, I had planned to continue my meta-analysis by investigating extinction rate across a large taxanomic group.
-There were signs that extinction rate could be measured from phylogenetic methods with relative confidence **CITE**.
-However, these hopes soon dissapated, and with them the seeming \emph{raison d'\^{e}tre} of the project.
-In measuring extinction rate one can obtain the most valid measures of population fitness in evolutionary biology.
+Originally, I had planned to continue my meta-analysis by investigating extinction rate across a large taxonomic group.
+There were signs that extinction rate could be measured from phylogenetic methods with relative confidence [@Pyron_2016].
+However, these hopes soon dissipated, and with them the seeming \emph{raison d'\^{e}tre} of the project.
+Extinction rate is arguably the best measure of population fitness.
 And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured over long, evolutionarily meaningful timescales [*but see* @Martins_2018].
-We did include measurements of extinction rate within the paper, however these findings were played down.
+We included measurements of extinction rate within the paper, however these findings were played down.
 The reason for this movement away from extinction and towards speciation is the controversy surrounding extinction rate measures.
 Measuring extinction rates within the bird phylogeny are problematic because of the nature of phylogenetic construction; for passerine birds a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction rate.
-Furthermore, the very detection of an absence of data --- which is what such methods attempt to do --- is a hard sell both conceptually and methodically come peer review; see @Rabosky_2010 for a cautionary tale on measuring extinction rate with ``BAMM``.
-The mechanisms and rates at which speciation and exvtinction occur are an active area of research.
+Furthermore, the very detection of an absence of data --- which is what such methods attempt to do --- is a hard sell both conceptually and methodically come peer review: see @Rabosky_2010 for a cautionary tale on measuring extinction rate with ``BAMM``.
+The mechanisms and rates at which speciation and extinction occur are an active area of research.
 In a recent pre-print @Diaz_2018 used ``BAMM`` across 104 time-calibrated molecular phylogenies and found that despite variation in ecology, biogeography and taxonomy, younger clades consistently had higher rates of speciation and extinction.
-This novel finding of time-dependency in rates of macroevolutionary diversification makes comparative analyses between challanging and calls into question countless studies.
-Similar to Wallace's view of sexual dichromatism we are left with a view of diversification rate measures where despite the advancements in the field we cannot yet account for the underlting true variation; which is vastly more complex than we can currently meausure.
+This novel finding of time-dependency in rates of macroevolutionary diversification makes comparative analyses between clades challenging and calls into question countless studies [@Diaz_2018].
+Similar to Wallace's view of sexual dichromatism we are left with a view of diversification rate measures where we cannot yet account for the underlying true variation; which is vastly more complex than we can currently measure.
 
-As with sexual dichromatism and macroevolutionary rates, the most severe limitation for environmental variation was the courseness of the meausres.
-In total, for 5,831 species we extracted 338,203,831 point measures^1^ of environmental variability, despite this our measures may not account for variation in biotic components such as food or predation.
-However, the methods employed here are arguably much better than those implemented by many studies; where the range mid-point is used to obtain a single measure of temperature, precipitation or seasonality.
-**CITE**.
-I would argue that reducing the multivariate data into several biologically meaningful predictors is a conceptually challanging exercise, but researchers should not feel limited by the amount of raw data that they can access^2^.
+As with sexual dichromatism and macroevolutionary rates, the most severe limitation for environmental variation was the coarseness of the measures.
+In total, for 5,831 species we extracted 338,203,831 point measures\footnote{5,831 species x 1000 samples per range x 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.} of environmental variability, but our measures may not account for variation in biotic components such as nutritional resources or predation.
+However, the methods employed here are arguably much better than those implemented by other studies; where the range mid-point is used to obtain a single measure of temperature, precipitation or seasonality or where only one bioclimatic variable is used.
+I would argue that reducing the multivariate data into several biologically meaningful predictors is a conceptually challenging exercise, but researchers should not feel limited by the amount of raw data that they can access\footnote{This statement applies to species with documented range sizes, which varies across taxa but is comprehensive for birds, mammals, amphibians and many marine groups.
+The \href{http://www.iucnredlist.org/technical-documents/spatial-data}{IUCN} has a large and accessible database for research use.
+Details of how we extracted data from these ranges is provided as a potentially useful resource in the Supplementary Material.}.
 In fact, previous to our study on birds the same methodology was applied to ~ 100 species of bovids with speciation, extinction and environmental variability obtained.
-This initially study was prematually halted as there was homogeneity in the response variable (speciation and extinction rates) across the smaller bovid phylogeny.
-However, this workflow was easily transferred over to passerine birds with similar methods for extracting environmental predictors.
+This study was prematurely halted as there was homogeneity in the response variable (speciation and extinction rates) across the smaller bovid phylogeny.
+However, this workflow was easily transferred over to passerine birds with similar methods used for extracting environmental predictors.
+The use of extensive bioclimatic predictors and spatial data in macroevolutionary research is imperative as we scale up studies.
+At very least, if bioclimatic variables do not encapsulate key ecological parameters (*e.g.*, resources, predation, habitat structure) these predictors will account for some of the noise in the dataset and may help isolate the effect of the key moderator variables of interest.
 
-\maketitle\unmarkedfntext{\textsuperscript{1}5,831 species x 1000 samples per range x 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.}
-\unmarkedfntext{\textsuperscript{2}This statement applies to species with documented range sizes, which varies across taxa but is comprehensive for birds, mammals, amphibians and many marine groups.
-The \href{http://www.iucnredlist.org/technical-documents/spatial-data}{IUCN} has a large and accessible database for research use.
-Details of how we extracted data from these ranges is provided as a potentially useful resource in the Supplementary Material.}
+<!-- \maketitle\unmarkedfntext{\textsuperscript{1}} -->
+<!-- \unmarkedfntext{\textsuperscript{2}} -->
+
+##Future Directions
+
+I had initially envisaged that this comparative study would allow for the reliable measure of extinction rate, utilise multiple taxonomic groups and use more robust measures of sexual selection [*see* @Anthes_2016].
+As such, many questions regarding this topic remain unanswered.
+Here I discuss the directions that we may follow in order to consolidate our understanding of sexual selection and its effects on fitness and diversification.
+I will draw heavily on the discussions from a Royal Society meeting on _'Sexual selection: patterns in the history of life'_\footnote{Specifically, the following two sections are spurred from dialogue with \href{https://paleobiology.si.edu/staff/individuals/hunt.cfm}{Dr Gene Hunt} and \href{https://www.earth.ox.ac.uk/people/erin-saupe/}{Associate Professor Erin E.
+Saupe}} in the UK this year.
+This meeting spurred further inquiry into the subject and how interdisciplinary approaches can be harnessed to liberate ourselves from conventional comparative methodologies.
+
+<!-- \maketitle\unmarkedfntext{\textsuperscript{1}Specifically, the following two sections are spurred from dialogue with \href{https://paleobiology.si.edu/staff/individuals/hunt.cfm}{Dr Gene Hunt} and \href{https://www.earth.ox.ac.uk/people/erin-saupe/}{Associate Professor Erin E.
+Saupe}} -->
+The study by @Martins_2018 has already been referred to several times in the main text of both papers and for good reason.
+This elegant study was able to do what no one else has; reliably measure extinction rate over meaningful time scales.
+Using fossil ostracods with various levels of sexual dimorphism, @Martins_2018 found that species with higher levels of dimorphism were swept to extinction more quickly.
+From stratographic occurrence data a capture–mark–recapture model was used to determine the rate of extinction of different species.
+Their finding challenges the results from experimental evolution studies [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015] and our meta-analysis that find a positive effect of sexual selection in mitigating extinction.
+Importantly the time-scale used by @Martins_2018 is a key advantage of their study.
+Moving forward, it would be incredibly valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual metamorphism to conduct such a study.
+In the interim I would be eager to see how ``BAMM`` would perform on a ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected among extant taxa.
+For me, this will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction from phylogenetic inference.
+
+One aspect and only foreseeable shortcoming of the ostracod study is that it did not account for environmental variability; a challenging proposal for extinct species nevertheless.
+However the use of paleoclimate models in evolution is now a reality; with recent studies incorporating averaged values across multiple paleoclimate models as predictors of speciation [*see* @quintero_2018].
+While the uncertainty surrounding paleoclimates was beyond the scope of the study future studies should remain vigilant of the everlasting need to incorporate key ecological processes in macroevolutionary research [*reviewed in* @Weber_2017].
+In our study on passerine birds we found that range size is negatively associated with speciation.
+The role of range size in speciation may be further investigated through ancestral geographic range reconstruction [@Quintero_2015], as ancestral ranges may be beneficial in understanding modes of speciation and the causality of range size expansions and contractions [@Losos_2003].
+Within Palaeobiology the role of range size and niche breadth on diversification patterns has been the subject of recent interest [*e.g.* @Saupe_2015; @Saupe_2016].
+Across 92 bivalve species that lived during a mid-Pliocene warming period @Saupe_2015 used environmental niche modelling to assess the impact of ecological processes on macroevolutionary patterns of extinction.
+Harnessing ecological niche modelling in macroevolutionary studies across expansive extant taxa such as passerine birds is a daunting challenge but offers an Eco-evo synthesis that may be particularly pertinent in questions relating to extinction, especially if the taxonomic group under inquiry are of high conservation and cultural value, for which the songbirds certainly are.
+
+ Importantly, macroevolutionary comparative studies can examine the association between sexual selection and speciation across taxa, over evolutionary time-scales and in natural environments.

### 2018-10-07 Sexual Selection and Extinction.Rmd
Manuscript and Supp Headings Changed

@@
 
-###Time-constant
-
-We also ran a time-constant MCC of BAMM
-```{r, eval=FALSE}
-#Read in the tree and MCMC to check for convergence 
-MCC.BAMM.tree  <- read.tree("data/BAMM_MCC/MCC.passerine.tre") #Same as other MCC tree already loaded
-MCC.BAMM.Constant.mcmc <- read.csv( "data/BAMM_MCC/TimeConstant/tree_MCC_Constant_mcmc_out.txt" , stringsAsFactors=F)
-
-#Plot of convergence can be generated by:
-#plot(MCC.BAMM.mcmc$logLik ~ MCC.BAMM.mcmc$generation)
-
-#Looks like it has converged so let's discard burn in: 
-burnstart <- floor(0.1 * nrow(MCC.BAMM.Constant.mcmc))
-postburn <- MCC.BAMM.Constant.mcmc[burnstart:nrow(MCC.BAMM.Constant.mcmc), ]
-
-#We can also check effective population sizes of the log-likelihhod and number of shift events in each sample
-#We want at least 200 (although that's on the low side)
-
-cbind(effectiveSize(postburn$N_shifts), effectiveSize(postburn$logLik)) %>% `colnames<-`(c("N_Shifts", "logLik")) %>% `rownames<-`("Effective Sample Size") %>% pander()
-
-# Read in Event Data
-MCC.BAMM.ED.Constant <- getEventData(MCC.BAMM.tree,  "data/BAMM_MCC/TimeConstant/tree_MCC_Constant_event_data.txt", burnin=0.1, nsamples=1000)
-saveRDS(MCC.BAMM.ED.Constant, 'data/MCC.BAMM.ED.Constant.rds')
-#From the Event Data we can extract
-
-css2 <- credibleShiftSet(MCC.BAMM.ED.Constant, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95)
-
-#Now we obtain the number of distinct shifts: (Out of 1000 samples this is super high, essentially each one is distinct)
-
-css2$number.distinct
-
-plotPrior(MCC.BAMM.Constant.mcmc, expectedNumberOfShifts=100)
-
-MCC.BAMM.df.Constant <- BAMM.extraction.function(MCC.BAMM.ED.Constant)
-
-MCC.BAMM.df.Constant %>% ggplot(aes(x = mean.lambda, y = MCC.BAMM.df$mean.lambda))+
-  geom_point()
-cor(MCC.BAMM.df.Constant$mean.lambda, MCC.BAMM.df$mean.lambda)
-```
+<!-- ###Time-constant -->
+
+<!-- We also ran a time-constant MCC of BAMM -->
+<!-- ```{r, eval=FALSE} -->
+<!-- #Read in the tree and MCMC to check for convergence  -->
+<!-- MCC.BAMM.tree  <- read.tree("data/BAMM_MCC/MCC.passerine.tre") #Same as other MCC tree already loaded -->
+<!-- MCC.BAMM.Constant.mcmc <- read.csv( "data/BAMM_MCC/TimeConstant/tree_MCC_Constant_mcmc_out.txt" , stringsAsFactors=F) -->
+
+<!-- #Plot of convergence can be generated by: -->
+<!-- #plot(MCC.BAMM.mcmc$logLik ~ MCC.BAMM.mcmc$generation) -->
+
+<!-- #Looks like it has converged so let's discard burn in:  -->
+<!-- burnstart <- floor(0.1 * nrow(MCC.BAMM.Constant.mcmc)) -->
+<!-- postburn <- MCC.BAMM.Constant.mcmc[burnstart:nrow(MCC.BAMM.Constant.mcmc), ] -->
+
+<!-- #We can also check effective population sizes of the log-likelihhod and number of shift events in each sample -->
+<!-- #We want at least 200 (although that's on the low side) -->
+
+<!-- cbind(effectiveSize(postburn$N_shifts), effectiveSize(postburn$logLik)) %>% `colnames<-`(c("N_Shifts", "logLik")) %>% `rownames<-`("Effective Sample Size") %>% pander() -->
+
+<!-- # Read in Event Data -->
+<!-- MCC.BAMM.ED.Constant <- getEventData(MCC.BAMM.tree,  "data/BAMM_MCC/TimeConstant/tree_MCC_Constant_event_data.txt", burnin=0.1, nsamples=1000) -->
+<!-- saveRDS(MCC.BAMM.ED.Constant, 'data/MCC.BAMM.ED.Constant.rds') -->
+<!-- #From the Event Data we can extract  -->
+
+<!-- css2 <- credibleShiftSet(MCC.BAMM.ED.Constant, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95) -->
+
+<!-- #Now we obtain the number of distinct shifts: (Out of 1000 samples this is super high, essentially each one is distinct) -->
+
+<!-- css2$number.distinct -->
+
+<!-- plotPrior(MCC.BAMM.Constant.mcmc, expectedNumberOfShifts=100) -->
+
+<!-- MCC.BAMM.df.Constant <- BAMM.extraction.function(MCC.BAMM.ED.Constant) -->
+
+<!-- MCC.BAMM.df.Constant %>% ggplot(aes(x = mean.lambda, y = MCC.BAMM.df$mean.lambda))+ -->
+<!--   geom_point() -->
+<!-- cor(MCC.BAMM.df.Constant$mean.lambda, MCC.BAMM.df$mean.lambda) -->
+<!-- ``` -->
 
@@
 
-##Subsetted analysis with Armenta data
+##Subsetted analysis with spectrophotometry data
 
@@
 
-##Analysis using male-bias measure of sexual selection
+##Analysis using male-biased measure of sexual selection
 

### 2018-10-07 Manuscript.Rmd
Manuscript edits

@@
 The relative contribution of male and female RGB colour values were averaged across body patches and provide 'male-like' and 'female-like' plumage scores.
-Here we use the absolute difference between males and female plumage scores as an estimate of sexual dichromatism.
+Here we use the absolute difference between male and female plumage scores as an estimate of sexual dichromatism.
 Additionally, we used a measure of dichromatism from spectral data, used to estimate colour distance in avian colour space [@Armenta_2008].
@@
 However, this spectrophotometry data is not as extensive as the RGB measures; with only 581 passerine species available for this analysis.
-While there is a correlation between these two measures, there is residual variation ($r = 0.79$; **Figure S9**).
+While there is a correlation between these two measures, there is residual variation ($r = 0.79$; \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S9}).
 
@@
 Therefore, we also use an existing dataset [@Dale_2015] of male-biased sexual selection based on the first component of a phylogenetic principle component analysis (PPCA) of three characteristics associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
-This measure of male-biased sexual selection is available for only 2,465 species and is weakly correlated with the absolute values of sexual dichromatism using RGB measures ($r = 0.34$; **Figure S10**).
+This measure of male-biased sexual selection is available for only 2,465 species and is weakly correlated with the absolute values of sexual dichromatism using RGB measures ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S11}).
 
@@
 To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature (BIO4) and precipitation seasonality (BIO15).
-PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (**Table S1**).
-Given that species range is a potentially informative predictor of speciation and extinction we controlled for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; **Figure S1**).
+PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
+Given that species range is a potentially informative predictor of speciation and extinction we controlled for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}).
 To obtain long-term variation in climates for each species range we take the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
-Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates  (**Table S2**, **Figure S2**).The five predictors of environmental variability are not strongly correlated (**Figure S3**).
-Details and code to generate these predictors can be found within the ESM.
+Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
+Details and code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.
 
@@
 Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification [@Belmaker_2015; @Rabosky_diversification_2018].
-Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events and $\lambda_{ND}$ is a measure of speciation across the root to tip path.
-These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE) models.
-Based on previous simulation studies $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
+Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events and $\lambda_{ND}$ is a measure of speciation across the root-to-tip path.
+These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE).
+But, based on previous simulation studies $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
 
@@
 Importantly, these models provide tip-rate estimates of speciation and extinction rate.
-The parameters of the 100 BAMM runs are detailed in full in the ESM; briefly we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
+The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}; briefly we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
 Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPU's.
-Each run of BAMM reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200 (an arbitrary value, above which phylogeneticists believe posterior tree distributions can be accurately inferred) for the key model parameters of log-likelihood and number of rate shifts (**Table S3**, **Table S4**).
-Further details of BAMM parameters and output are available in the ESM, with tip-rate means and variances provided.
+Each run of BAMM reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200 (an arbitrary value, above which posterior distributions can often be accurately inferred) for the key model parameters of log-likelihood and number of rate shifts (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
+Further details of BAMM parameters and output are available in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}, with tip-rate means and variances provided.
 Additionally, given the variability in BAMM estimates, we also provide some analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on the MCC tree from a genetic-only phylogeny across all birds [@Harvey_2017].
@@
 To test the association between speciation/extinction and sexual selection, environmental variability and their interaction, we used phylogenetic least squares (PGLS) models in the ``nlme`` package [@nlme_2018].
-We began by estimating the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using ``corPagel`` in the ``ape`` package [@ape_2004] on the MCC tree.
+We began by estimating the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using the ``corPagel`` function in the ``ape`` package [@ape_2004] on the MCC tree.
 The estimate of $\lambda$ was then fixed for model selection, which was conducted on the same MCC tree derived from 2,500 draws of the posterior distribution [@Jetz_2012].
@@
 For $\lambda_{BAMM}$ and $\mu_{BAMM}$ we were also able to use model weights sourced from the inverse of the variance of tip-rates for a given species from the posterior distribution (*n* = 1,000).
+Model weights thus reflect the degree of precision to which each species tip-rate is measured in BAMM.
 Using model selection we only compared interaction terms between a measure of sexual selection (sexual dichromatism or male-biased sexual selection) and environmental measures.
@@
 The phylogenetic path analysis used tip-rates and correlation structure from the MCC tree.
-Further details of the path analysis including reasons for path directions can be found within the ESM along with all other analyses and the relevant ``R`` code to reproduce results.
+Further details of the path analysis including reasons for path directions can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
 
@@
 PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in model selection.
-However, no interaction terms were significant or present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; **Table S5**, **Table S6**, **Table S10**, **Table S13**).
+However, no interaction terms were significant or present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
 Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
-Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, **Figure S10**).
+Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
 
@@
 Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-PGLS analyses using subsetted data (n = 581), where sexual dichromatism was measured using spectrophotometry [@Armenta_2008] yielded results concordant to the full data set; that is, no association between sexual dichromatism and speciation or extinction (**Figure S10**).
+PGLS analyses using subsetted data (*n* = 581), where sexual dichromatism was measured using spectrophotometry [@Armenta_2008] yielded results concordant to the full data set; that is, no association between sexual dichromatism and speciation or extinction (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
 Our results from MCC models are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 100 random trees.
-Model estimates are normally distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, see HPD (highest posterior density) intervals in **Table S8**).
-For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (`r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (**Table S12**).
+Model estimates are normally distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, see HPD (highest posterior density) intervals in \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
+For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (`r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
 
@@
 The distribution of estimates from PGLS models on 100 random trees was similar to the estimate from an MCC tree: among the 100 tress there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
-Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within **Table S15**.
+Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}.
 
@@
 Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subsetted for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
-Model using data subsetted for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (**Figure S10, Table S11, Table S2**).
+Model using data subsetted for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
 Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with highest power using near-complete taxon sampling (*n* = 5,812).
@@
 
-Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; **Figure S13**).
+Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; \href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S13}).
 Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from male-biased sexual selection to sexual dichromatism was relatively weak, ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`), despite dichromatism being widely used as an indicator of the strength of sexual selection.
@@
 
-We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (**FigureS10, Table S11, Table S12**) or male-biased sexual selection (\autoref{model_results}b, **Table S14, Table S15**).
+We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
 
@@
 We used a range of methods to overcome uncertainty in the power and precision of various tip-rate estimates as well as the variation between trees.
-We found that that the tip-rate estimates across the 100 trees were quite variable for both $\lambda_{DR}$ and $\lambda_{BAMM}$ (**Figure S14**).
-Despite this, tip-rate estimates of speciation from the two methods were moderately correlated across the 100 trees ($r = 0.75$) and for the MCC tree ($r = 0.68$; **Figure S15**).
+We found that that the tip-rate estimates across the 100 trees were quite variable for both $\lambda_{DR}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S14}).
+Despite this, tip-rate estimates of speciation from the two methods were moderately correlated across the 100 trees ($r = 0.75$) and for the MCC tree ($r = 0.68$; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
 However, in comparison to $\lambda_{DR}$, the value of $\lambda_{BAMM}$ was the _mean_ drawn from a posterior distribution (*n* = 1,000) of BAMM generations, thus estimates of $\lambda_{BAMM}$ (and $\mu_{BAMM}$) have an added level of variation.
 To account for this variation, weights (using the inverse of the variance) were used for the PGLS models.
-From 1000 posterior samples of the MCC BAMM run the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* **Figures S5 and S7**).
-Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (**Table S3 and S4**) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) was high.
-This means that although the number of rate shifts reached convergence ($median = 59$; **Figure S6**), the locations of the rate shifts (*i.e.* the credible shift set) across the tree are highly heterogeneous.
+From 1000 posterior samples of the MCC BAMM run the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figures S5} and \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S7}).
+Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) was high.
+This means that although the number of rate shifts reached convergence ($median = 59$; \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6}), the locations of the rate shifts (*i.e.* the credible shift set) across the tree are highly heterogeneous.
 The variability in the locations for the shift configurations is a likely source of uncertainty in downstream tip-rates used as the response variable in PGLS models.
-Notably, the 95 % HPD interval for model estimates using BAMM rates from 100 trees was about 20-fold the 95 % confidence intervals of the estimate from the MCC tree; whereas for $\lambda_{DR}$ and $\lambda_{ND}$ the HPD 95 % interval width was equal to or less than the MCC 95 % CI (\autoref{model_results}, **Table S9**).
+Notably, the 95 % HPD interval for model estimates using BAMM rates from 100 trees was about 20-fold the 95 % confidence intervals of the estimate from the MCC tree; whereas for $\lambda_{DR}$ and $\lambda_{ND}$ the HPD 95 % interval width was equal to or less than the MCC 95 % CI (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S9}).
 
@@
 Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection, and that variance in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
-There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is enigmatic.
+There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
 Firstly, ecological pressures could drive sexual dichromatism in the absence of mate choice if sex-specific niches provide opportunity for the evolution of sex-limited traits [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989].
@@
 
-In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we find that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ --- but this association was not depenendent on variable or stressful environments.
+In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we find that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ --- but this association was not dependent on variable or stressful environments.
 When evolving under stressful conditions, sexually antagonistic selection is hypothesised to be weakened relative to selection that is concordant between the sexes [@Connallon_SA_2015; @Connallon_Hall_2016].
@@
 Several studies have used morphological measures of wing shape (Hand-wing index/ Kipp's distance) --- a proxy for dispersal ability --- as predictors of speciation and coexistence [*e.g.*, @Kennedy_2016; @Pigot_2016].
-But currently published datasets of these measurements are not available for many passerines.
+But currently, published datasets of these measurements are not available for many passerines.
 Alternatively, reduced range size may be a consequence of speciation.
@@
 Darwin's finches).
-One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
+One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
 
@@
 However, BAMM has been subject to controversy in its ability to model rate shifts and extinction rates [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], and while several tip-rate estimates exist for speciation (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$) extinction estimates are not as easily obtained.
-Here, we found that BAMM often produced homogeneous $\lambda_{BAMM}$ and $\mu_{BAMM}$ for clades with few rate shifts; arguably this could reduce our power to detect extinction rates [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
+Here, we found that BAMM often produced homogeneous speciation and extinction rates for clades with few rate shifts; arguably this could reduce our power to detect samll differences in extinction rates among closely related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
 The ability to detect extinction rates rarely improves when fossil records are used, as many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
-But, there are exceptions, ostracods --- a group of marine invertebratesone --- have extensive fossil records and show identifiable sexual dimorphism.
+But, there are exceptions: ostracods --- a group of marine invertebrates --- have extensive fossil records and show identifiable sexual dimorphism.
 Across 93 species of ostracods, @Martins_2018 found that ostracod taxa with larger sexual dimorphism (representing exaggerated investment by males in reproductive organs) had higher extinction rates; implying that sexual selection increases extinction risk.
@@
 Improvements in estimating tip-rates of speciation has allowed us to revisit the association between sexual selection and speciation.
-With high power and alongside extensive ecological predictors we find that male-biased sexual selection predicts speciation but sexual dichromatism does not.
+With high power and alongside extensive ecological predictors we find support that male-biased sexual selection predicts speciation.
 If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold with sexual selection a potentially valuable utility in conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].

### 2018-10-08 Manuscript.Rmd
Critical Analysis Edits

@@
 
-Perhaps one of the biggest unravelings of this project was the decoupling of male-bias sexual selection from sexual dichromatism.
-Despite the complex statistical models, monsterous sample size and modern sexual conflict theory the most fascinating finding was one that Wallace and Darwin were at odds over 130 years prior [@Kottler_1980].
-The widespread use of sexual dichromatism largely goes unchallenged as our instinctive Darwinian mind reminisces of peacocks and birds of paradise.
-My thoughts were such when I began this scientific inquiry, but am now skeptical and somewhat disheartened.
+Perhaps one of the biggest unravelings of this project was the decoupling of male-biased sexual selection from sexual dichromatism, remarkably this same controversy engrossed Wallace and Darwin 130 years ago [@Kottler_1980].
+The widespread use of sexual dichromatism largely goes unchallenged as our instinctive Darwinian mind reminisces of peacocks and birds of paradise; my thoughts were such when I began this scientific inquiry.
 Of course, I was expecting sexual dichromatism from RGB measures to be a noisy measure of sexual selection, but across 5,831 species I anticipated a signal would permeate through the noise.
-And despite the correlation between sexual dichromatism and male-bias sexual selection is present, the relationship is surprisingly weak (see Figure 3 in main text).
+And despite the correlation between sexual dichromatism and male-biased sexual selection is present, the relationship is surprisingly weak (see \autoref{Path_Analysis} in main text).
 With this in mind, the practical use of sexual dichromatism in macroevolutionary research is severely weakened.
-This finding is particularly pertinent as efficient proxy measurements are incredibly valuable in comparative analyses: that's because they increase the accessibility, available sample size and transfer-ability of a study making it more conducive to research synthesis.
+This finding is particularly pertinent as efficient proxy measurements are incredibly valuable in comparative analyses: that's because they increase the accessibility, available sample size and transferability of a study, making it more conducive to research synthesis.
 But, as Wallace remarks, the matter is inevitably more complex despite the irresistibility of the Darwinian argument\footnote{Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took more interest in a view where natural selection on predator-prone nesting females generated dullness in females but not males.} [@Kottler_1980; @Caro_2017].
-On the bright side, the multivariate measure of sexual selection that is the PPCA of sexual dimorphism, social polygyny and [lack of] paternal care is a dataset that better predicts speciation and would benefit from being expanded across birds.
+On the bright side, the multivariate measure of male-biased sexual selection that is: sexual dimorphism, social polygyny and [lack of] paternal care is a dataset that better predicts speciation and would benefit from being expanded across birds.
 
@@
 
-Not only are the measurement of sexual selection under a cloud of uncertainty, the estimates of speciation and extinction are another point of controversy.
+Not only are the measurements of sexual selection under a cloud of uncertainty, the estimates of speciation and extinction are another point of controversy.
 Macroevolutionary research is a field plagued by methodological uncertainty and conflicts.
-And during the second part of this MSc I jumped head-first into the controversy.
 Originally, I had planned to continue my meta-analysis by investigating extinction rate across a large taxonomic group.
@@
 Extinction rate is arguably the best measure of population fitness.
-And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured over long, evolutionarily meaningful timescales [*but see* @Martins_2018].
+And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured accurately over long, evolutionarily meaningful timescales [*but see* @Martins_2018].
 We included measurements of extinction rate within the paper, however these findings were played down.
-The reason for this movement away from extinction and towards speciation is the controversy surrounding extinction rate measures.
-Measuring extinction rates within the bird phylogeny are problematic because of the nature of phylogenetic construction; for passerine birds a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction rate.
-Furthermore, the very detection of an absence of data --- which is what such methods attempt to do --- is a hard sell both conceptually and methodically come peer review: see @Rabosky_2010 for a cautionary tale on measuring extinction rate with ``BAMM``.
+The reason for this movement away from extinction and towards speciation is the controversy surrounding extinction rate measures as well as phylogentic uncertainty.
+Measuring extinction rates within the bird phylogeny is problematic because of the nature of phylogenetic construction; for passerine birds a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction rate [@Beaulieu_2015].
+Furthermore, the very detection of an absence of data --- which is what such methods attempt to do --- is a hard sell both conceptually and methodically come peer review: see @Rabosky_Extinction_2016 for a cautionary tale on measuring extinction rate with ``BAMM``.
 The mechanisms and rates at which speciation and extinction occur are an active area of research.
 In a recent pre-print @Diaz_2018 used ``BAMM`` across 104 time-calibrated molecular phylogenies and found that despite variation in ecology, biogeography and taxonomy, younger clades consistently had higher rates of speciation and extinction.
-This novel finding of time-dependency in rates of macroevolutionary diversification makes comparative analyses between clades challenging and calls into question countless studies [@Diaz_2018].
+This finding of time-dependency in rates of diversification makes comparative analyses between clades challenging and calls into question countless studies [@Diaz_2018].
 Similar to Wallace's view of sexual dichromatism we are left with a view of diversification rate measures where we cannot yet account for the underlying true variation; which is vastly more complex than we can currently measure.
@@
 In total, for 5,831 species we extracted 338,203,831 point measures\footnote{5,831 species x 1000 samples per range x 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.} of environmental variability, but our measures may not account for variation in biotic components such as nutritional resources or predation.
-However, the methods employed here are arguably much better than those implemented by other studies; where the range mid-point is used to obtain a single measure of temperature, precipitation or seasonality or where only one bioclimatic variable is used.
+However, the methods employed here are arguably much better than those traditionally implemented; where the range mid-point is used to obtain a single measure of temperature, precipitation or seasonality or where only one bioclimatic variable is used.
 I would argue that reducing the multivariate data into several biologically meaningful predictors is a conceptually challenging exercise, but researchers should not feel limited by the amount of raw data that they can access\footnote{This statement applies to species with documented range sizes, which varies across taxa but is comprehensive for birds, mammals, amphibians and many marine groups.
 The \href{http://www.iucnredlist.org/technical-documents/spatial-data}{IUCN} has a large and accessible database for research use.
-Details of how we extracted data from these ranges is provided as a potentially useful resource in the Supplementary Material.}.
+Details of how we extracted data from these ranges is provided as a potentially useful resource in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.}.
 In fact, previous to our study on birds the same methodology was applied to ~ 100 species of bovids with speciation, extinction and environmental variability obtained.
@@
 The use of extensive bioclimatic predictors and spatial data in macroevolutionary research is imperative as we scale up studies.
-At very least, if bioclimatic variables do not encapsulate key ecological parameters (*e.g.*, resources, predation, habitat structure) these predictors will account for some of the noise in the dataset and may help isolate the effect of the key moderator variables of interest.
-
-<!-- \maketitle\unmarkedfntext{\textsuperscript{1}} -->
-<!-- \unmarkedfntext{\textsuperscript{2}} -->
+At very least, if bioclimatic variables do not encapsulate key ecological parameters (*e.g.*, resources, predation, habitat structure) these predictors may account for some of the noise in the dataset and may help isolate the effect of the key moderator variables of interest.
 
@@
 As such, many questions regarding this topic remain unanswered.
-Here I discuss the directions that we may follow in order to consolidate our understanding of sexual selection and its effects on fitness and diversification.
+Here I discuss the broader directions that we may follow in order to consolidate our understanding of sexual selection and its effects on fitness and diversification.
 I will draw heavily on the discussions from a Royal Society meeting on _'Sexual selection: patterns in the history of life'_\footnote{Specifically, the following two sections are spurred from dialogue with \href{https://paleobiology.si.edu/staff/individuals/hunt.cfm}{Dr Gene Hunt} and \href{https://www.earth.ox.ac.uk/people/erin-saupe/}{Associate Professor Erin E.
@@
 
-<!-- \maketitle\unmarkedfntext{\textsuperscript{1}Specifically, the following two sections are spurred from dialogue with \href{https://paleobiology.si.edu/staff/individuals/hunt.cfm}{Dr Gene Hunt} and \href{https://www.earth.ox.ac.uk/people/erin-saupe/}{Associate Professor Erin E.
-Saupe}} -->
 The study by @Martins_2018 has already been referred to several times in the main text of both papers and for good reason.
 This elegant study was able to do what no one else has; reliably measure extinction rate over meaningful time scales.
-Using fossil ostracods with various levels of sexual dimorphism, @Martins_2018 found that species with higher levels of dimorphism were swept to extinction more quickly.
-From stratographic occurrence data a capture–mark–recapture model was used to determine the rate of extinction of different species.
-Their finding challenges the results from experimental evolution studies [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015] and our meta-analysis that find a positive effect of sexual selection in mitigating extinction.
-Importantly the time-scale used by @Martins_2018 is a key advantage of their study.
-Moving forward, it would be incredibly valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual metamorphism to conduct such a study.
-In the interim I would be eager to see how ``BAMM`` would perform on a ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected among extant taxa.
+Using fossil ostracods with various levels of sexual dimorphism, @Martins_2018 found that species with high levels of sexual dimorphism were quickly swept to extinction.
+From stratographic occurrence data a capture–mark–recapture model was used to determine the rate of extinction for 93 species.
+Their finding challenges the results from experimental evolution studies [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015] and our meta-analysis that find sexual selection in mitigates extinction.
+Moving forward, it would be incredibly valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual dimorphism to conduct such a study.
+In the interim I would be eager to see how ``BAMM`` would perform on an ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected among extant taxa.
 For me, this will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction from phylogenetic inference.
@@
 However the use of paleoclimate models in evolution is now a reality; with recent studies incorporating averaged values across multiple paleoclimate models as predictors of speciation [*see* @quintero_2018].
-While the uncertainty surrounding paleoclimates was beyond the scope of the study future studies should remain vigilant of the everlasting need to incorporate key ecological processes in macroevolutionary research [*reviewed in* @Weber_2017].
+While the uncertainty surrounding paleoclimates was beyond the scope of the study, future studies should remain vigilant of the everlasting need to incorporate key ecological processes in macroevolutionary research [*reviewed in* @Weber_2017].
 In our study on passerine birds we found that range size is negatively associated with speciation.
@@
 Within Palaeobiology the role of range size and niche breadth on diversification patterns has been the subject of recent interest [*e.g.* @Saupe_2015; @Saupe_2016].
-Across 92 bivalve species that lived during a mid-Pliocene warming period @Saupe_2015 used environmental niche modelling to assess the impact of ecological processes on macroevolutionary patterns of extinction.
-Harnessing ecological niche modelling in macroevolutionary studies across expansive extant taxa such as passerine birds is a daunting challenge but offers an Eco-evo synthesis that may be particularly pertinent in questions relating to extinction, especially if the taxonomic group under inquiry are of high conservation and cultural value, for which the songbirds certainly are.
-
- Importantly, macroevolutionary comparative studies can examine the association between sexual selection and speciation across taxa, over evolutionary time-scales and in natural environments.
+Across 92 bivalve species that lived during a mid-Pliocene warming period, @Saupe_2015 used environmental niche modelling to assess the impact of ecological processes on extinction.
+Harnessing ecological niche modelling in macroevolutionary studies across extant taxa such as passerine birds is a daunting challenge but offers an Eco-evo synthesis that may be particularly pertinent in questions relating to extinction; especially if the taxonomic group under inquiry is of high conservation value, for which the songbirds certainly are.

### 2018-10-09 Manuscript.Rmd
a few more changes and references to add

@@
 Sexual selection is predicted to increase speciation rate as it can facilitate adaptation to new environments, and work alongside divergent natural selection to swiftly promote reproductive isolation.
-However, sexual selection frequently favours investment in costly or harmful traits that may reduce adaptation affect diversification.
+However, sexual selection frequently favours investment in costly or harmful traits that may reduce adaptation and affect diversification.
 Here, we investigate the association of sexual selection with speciation and extinction in passerine birds.
@@
 
-Selection pressures drive speciation.
+Selection pressures are the main engine of speciation.
 As populations diverge they accumulate genetic and phenotypic changes that ultimately erect barriers against gene flow [@Presgraves_2010; *but see* @Shaw_2011].
-In this view, populations will speciate and adapt more readily to new divergent environments if selection is strong and/or persistant.
+In this view, populations will speciate and adapt more readily to new and different environments if selection is strong and/or persistant.
 Hence, the rate at which speciation occurs is likely dependent on both the populations response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour @Hugall_2012; @Rabosky_morph_2013].
-Sexual selection is an evolutionary force likely to affect fitness components, such as reproductive success or viability, as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition.
+Sexual selection is an evolutionary force likely to affect fitness components of a population, such as reproductive success or viability, as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition (**add couple of refs**).
 Sexual selection is therefore hypothesised to be a key driver of diversification through shifting the rates of adaptation, speciation and extinction.
@@
 However, this positive correlation is not ubiquitous; indeed sexual selection frequently favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006].
-For instance, in species with high levels of sexual selection, mating success becomes more variable and uncertain; thus individuals must invest heavily to increase their chances of successfully reproducing.
-High investment into costly weaponry and sexual signals, may trade off against strategies that would improve offspring fitness (*e.g.* parental care).
+For instance, in species with high levels of sexual selection, mating success becomes more variable and uncertain; thus individuals must invest heavily in costly weaponry and sexual signals to increase their chances of successfully reproducing.
+Such high investment, may trade off against strategies that would improve offspring fitness (*e.g.* parental care).
 This is a costly consequence of sexual selection that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].
@@
 Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].
-Concurrently, sexual selection can promote the extent to which populations diverge by enlarging the phenotypic space diverging species can evolve into.
+Concurrently, sexual selection can increase the extent to which populations diverge by enlarging the phenotypic space diverging species can evolve into.
 These predictions emerge as sexual selection may increase genetic diversity as a result of trade-offs between naturally- and sexually-selected traits or trade-offs under intra-locus sexual conflict between male and female trait expressions [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
@@
 We might therefore expect bird populations experiencing stronger sexual selection to have increased adaptation and speciation rates.
+**I think the link between the increased heritability of of the ornamentation and adaptation and speciation should be explained more clearly**
 
@@
 In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signal efficacy in sticklebacks [@Wong_2007; @Candolin_2007].
+Alternatively, sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signal efficacy in sticklebacks [@Wong_2007; @Candolin_2007].
 In this conflicting view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.
+*** would be good to have in this paragraph a bit of how the unstability of the environment can be viewed as a proxy for harsher conditions, why we think that unstable environments impose larger selective pressures**
 
-The association between sexual selection and speciation has been investigated using many macroevolutionary approaches applied to various taxa.
-A meta-analysis of 64 effect sizes from 20 studies found a small, but significant positive association between sexual selection and speciation [@Kraaijeveld_2011].
-However, there was large variation in effect sizes due to variation in the measures of sexual selection used and other methodological choices (*e.g.* species richness *vs* rate estimates).
-Given that species richness and speciation rate can be inversely correlated [*e.g.* @Rabosky_fish_2018] and the recent refinement of methods to obtain tip-rate diversification estimates [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting.
-Firstly, the strength of sexual selection strength in macroevolutionary studies is often estimated using measures of sexual dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa.
-Secondlly, many studies use incomplete taxon sampling, with family-level analysis restricted to well sampled avian clades [@Seddon_2008; @Huang_2014].
-Thirdly, until recently, studies on the association between sexual selection and speciation have largely ignored the contribution and interaction with environmental variables to explain variation in speciation rates.
-Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague efforts to identify drivers of speciation and extinction.
+**we need a sentence earlier showing that we are doing this in birds, may be at the beginning of this paragraph, i added something**
+Birds are notorious for the wide arrange of ornaments that have presumably evolved in response to sexual selection, such as plumage colour, size dimorphism, complex courtship rituals and male song (**add ref**).
+The association between sexual selection and speciation has been investigated previously in birds using macroevolutionary approaches.
+A meta-analysis of 64 effect sizes from 20 studies found a small, but significant positive association between sexual selection and speciation across a wide range of taxa [@Kraaijeveld_2011].
+Within birds, they found a positive association but much lower than that for insects and fish.
+However, there was large variation in effect sizes due to variation in the measures of sexual selection used and diversification proxies (*e.g.* species/subspecies richness *vs* rate estimates).
+Contrary to this, it was recently shown by Huang and Rabosky and Cooney et al.
+**CITE: Sexual selection, speciation, and constraints on geographical range overlap in birds** that there is no link between sexual dichromatism and speciation in a sample of ~1000 species and ~1300 pairs of species, respectively.
+Given these contrasting results, and given that species richness and speciation rate are not necessarily correlated [*e.g.* @Rabosky_fish_2018] and the recent refinement of methods to obtain tip-rate diversification estimates [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018], the association between sexual selection and speciation rate is worth revisiting.
+Firstly, the strength of sexual selection in macroevolutionary studies is often estimated using measures of sexual dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa.
+Secondly, until recently, studies on the association between sexual selection and speciation have largely ignored the contribution and interaction with environmental variables to explain variation in speciation rates.
+Thirdly, many studies use incomplete taxon sampling, with family-level analysis restricted to well sampled avian clades [@Seddon_2008; @Huang_2014].Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague efforts to identify drivers of speciation and extinction.
 
-Here, we perform a comprehensive analysis of the relationship between sexual selection and speciation or extinction in passerine birds (Order: Passeriformes), arguably the largest group for which there exists comprehensive data.
-Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset [*n* = 918; @Armenta_2008].
+Here, we perform a macroevolutionary analysis of the relationship between sexual selection and speciation or extinction in passerine birds (Order: Passeriformes), arguably the largest group for which there exists comprehensive data.
+Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset [*n* = 918; @Armenta_2008] ** mention here the Cooney et al.
+paper ,or move this bit up where I added it earlier**.
 We build on these previous studies by using published datasets of sexual dichromatism measured from images [*n* = 5,812; @Dale_2015] alongside smaller datasets of sexual dichromatism measured from spectrophotometry [*n* = 581; @Armenta_2008] and a composite measure of male-biased sexual selection [*n* = 2,465; @Dale_2015].
@@
 We found no evidence that sexual dichromatism alters the rate of speciation in passerine birds independently or via interactions with environmental variability.
-This result is consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and two measures of dichromatism (spectral and RGB).
+This result is consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and two measures of dichromatism (spectral and RGB), and is consistent with previous studies performed at a smaller scale (Huang & Rabosky, Cooney et al.).
 However, we did find that when using a measure of male-biased sexual selection --- instead of sexual dichromatism --- there was a positive association with speciation, though the strength of this relationship varied across different rate estimates.
@@
 Firstly, ecological pressures could drive sexual dichromatism in the absence of mate choice if sex-specific niches provide opportunity for the evolution of sex-limited traits [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989].
-In damselflies, differences between the sexes in their exposure to UV radiation drives dichromatism between the sexes [@Cooper_2010; @Punzalan_2010].
+**In damselflies, differences between the sexes in their exposure to UV radiation drives dichromatism between the sexes [@Cooper_2010; @Punzalan_2010] I would erase this example and find stuff on birds** .
+Natural selection on females maybe important in explaining patterns of sexual dichromatism in birds ** cite: Hofmann CM, Cronin TW, Omland KE.
+Evolution of sexual dichromatism.
+1.
+Convergent losses of elaborate female coloration in New World Orioles (Icterus spp.).
+The Auk.
+2008;125:778–89.Price JJ, Eaton MD.
+Reconstructing the evolution of sexual dichromatism: current color diversity does not reflect past rates of male and female change.
+Evolution.
+2014;68(7):2026–37.** In fairy-wrens (Malurus cyaneus), female colouration has evolved in response to environmental pressures, leading to increased sexual dichromatism in regions where females are under stronger selection from predation (**cite: Medina et al.
+https://doi.org/10.1186/s12862-016-0861-3**).
 Ecological impacts on sexual dichromatism are supported by our path analysis, which reveals that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation) albeit relatively weakly.
-Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition.
+Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition **cite: https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/1365-2656.12824**.
 And thirdly, not all plumage colouration honestly reflects mate quality.
@@
 
-In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we find that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ --- but this association was not dependent on variable or stressful environments.
+In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ --- but this association was not dependent on variable or stressful environments.
 When evolving under stressful conditions, sexually antagonistic selection is hypothesised to be weakened relative to selection that is concordant between the sexes [@Connallon_SA_2015; @Connallon_Hall_2016].
@@
 Given that we did not detect an effect of environmental variability (by itself or as an interaction with sexual selection) we are left with several possible explanations.
-Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability the species is evolving under.
+** would be good to at least mention how/why we expected one of the environmenta variables to have an effect on speciation, up to here there are not much details in the causal relationship we were expecting with environment** Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability the species is evolving under.
 Specifically, sexual antagonism may be more easily purged in environments where cyclic pressures (*e.g.* seasonality) are perpetuated; whereas when environmental variability is directional (*e.g.* long-term climatic change) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
@@
 The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates; however reduced range size can be a cause or consequence of speciation.
-Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995].
-However, birds with smaller range sizes are also predicted to have lower dispersal ability.
+** compare the effects??
+looks like association with sexual selection is stronger than that with range size, which is interesting**  Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995].
+However, **why however?
+seems in line with previous sentence** birds with smaller range sizes are also predicted to have lower dispersal ability.
 This ensues that individuals might not be able to overcome geographical barriers and maintain gene flow; which would promote reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
@@
 However, BAMM has been subject to controversy in its ability to model rate shifts and extinction rates [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], and while several tip-rate estimates exist for speciation (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$) extinction estimates are not as easily obtained.
-Here, we found that BAMM often produced homogeneous speciation and extinction rates for clades with few rate shifts; arguably this could reduce our power to detect samll differences in extinction rates among closely related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
+Here, we found that BAMM often produced homogeneous speciation and extinction rates for clades with few rate shifts; arguably this could reduce our power to detect small differences in extinction rates among closely related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
 The ability to detect extinction rates rarely improves when fossil records are used, as many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
@@
 Unfortunately, these methods often rely on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List [*e.g.*, @Morrow_2003; @Bro-Jorgensen_2014; @Morrow_2004] and may be biased to species under greater anthropogenic threat, culturally valued or well researched.
+** if you need to cut words I think that this paragraph could be just summarised **
 

### 2018-10-10 Manuscript.Rmd
Edits and added data objects needed for recompilation + figures

@@
 output:
-  pdf_document: default
-  html_document: default
   word_document:
     pandoc_args: --smart
+  html_document: default
+  pdf_document: default
 header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
@@
 Hence, the rate at which speciation occurs is likely dependent on both the populations response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour @Hugall_2012; @Rabosky_morph_2013].
-Sexual selection is an evolutionary force likely to affect fitness components of a population, such as reproductive success or viability, as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition (**add couple of refs**).
+Sexual selection is an evolutionary force likely to affect fitness components of a population [@Cally_in_review], such as reproductive success or viability, as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition [@McCullough_2016].
 Sexual selection is therefore hypothesised to be a key driver of diversification through shifting the rates of adaptation, speciation and extinction.
@@
 In male homogametic species (ZZ/ZW or ZZ/ZO) --- such as birds --- males are predicted to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003].
-We might therefore expect bird populations experiencing stronger sexual selection to have increased adaptation and speciation rates.
-**I think the link between the increased heritability of of the ornamentation and adaptation and speciation should be explained more clearly**
+If fitness and associated mating traits are more heritable in birds we might expect bird populations experiencing stronger sexual selection to have increased adaptation and speciation rates compared to male heterogametic taxa.
 
@@
 In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-Alternatively, sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signal efficacy in sticklebacks [@Wong_2007; @Candolin_2007].
+Therefore, strong sexual selection and strong natural selection --- such as environmental change over space and time --- may work synergestically to increase speciation rates.
+
+<!-- Alternatively, sexual selection may interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signal efficacy in sticklebacks [@Wong_2007; @Candolin_2007].
 In this conflicting view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.
-*** would be good to have in this paragraph a bit of how the unstability of the environment can be viewed as a proxy for harsher conditions, why we think that unstable environments impose larger selective pressures**
+-->
 
-**we need a sentence earlier showing that we are doing this in birds, may be at the beginning of this paragraph, i added something**
-Birds are notorious for the wide arrange of ornaments that have presumably evolved in response to sexual selection, such as plumage colour, size dimorphism, complex courtship rituals and male song (**add ref**).
+Birds are notorious for the variety of ornaments that have evolved in response to sexual selection, such as plumage colour, size dimorphism, complex courtship rituals and male song [@Gontard_1999].
 The association between sexual selection and speciation has been investigated previously in birds using macroevolutionary approaches.
@@
 However, there was large variation in effect sizes due to variation in the measures of sexual selection used and diversification proxies (*e.g.* species/subspecies richness *vs* rate estimates).
-Contrary to this, it was recently shown by Huang and Rabosky and Cooney et al.
-**CITE: Sexual selection, speciation, and constraints on geographical range overlap in birds** that there is no link between sexual dichromatism and speciation in a sample of ~1000 species and ~1300 pairs of species, respectively.
-Given these contrasting results, and given that species richness and speciation rate are not necessarily correlated [*e.g.* @Rabosky_fish_2018] and the recent refinement of methods to obtain tip-rate diversification estimates [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018], the association between sexual selection and speciation rate is worth revisiting.
+Contrary to this, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset [*n* = 918; @Armenta_2008]; similarly @Cooney_2017 found no effect of sexual dichromatism on diversification rates when using 1,306 pairs of species and human estimates of dichromatism.
+Given methodological advancements in tip-rate estimates [@Rabosky_diversification_2018], the association between sexual selection and speciation is worth revisiting in order to overcome previous limitations.
 Firstly, the strength of sexual selection in macroevolutionary studies is often estimated using measures of sexual dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa.
@@
 Here, we perform a macroevolutionary analysis of the relationship between sexual selection and speciation or extinction in passerine birds (Order: Passeriformes), arguably the largest group for which there exists comprehensive data.
-Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset [*n* = 918; @Armenta_2008] ** mention here the Cooney et al.
-paper ,or move this bit up where I added it earlier**.
 We build on these previous studies by using published datasets of sexual dichromatism measured from images [*n* = 5,812; @Dale_2015] alongside smaller datasets of sexual dichromatism measured from spectrophotometry [*n* = 581; @Armenta_2008] and a composite measure of male-biased sexual selection [*n* = 2,465; @Dale_2015].
@@
 We found no evidence that sexual dichromatism alters the rate of speciation in passerine birds independently or via interactions with environmental variability.
-This result is consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and two measures of dichromatism (spectral and RGB), and is consistent with previous studies performed at a smaller scale (Huang & Rabosky, Cooney et al.).
-However, we did find that when using a measure of male-biased sexual selection --- instead of sexual dichromatism --- there was a positive association with speciation, though the strength of this relationship varied across different rate estimates.
+This result is consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and two measures of dichromatism (spectral and RGB).
+The results are also consistent with previous studies performed at a smaller scale [@Huang_2014; @Cooney_2017].
+However, we found that when using a measure of male-biased sexual selection --- instead of sexual dichromatism --- there was a positive association with speciation, though the strength of this relationship varied across different rate estimates.
 Our findings suggest that the components of the composite measure of male-biased sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care) are better predictors of speciation than sexual dichromatism.
@@
 Firstly, ecological pressures could drive sexual dichromatism in the absence of mate choice if sex-specific niches provide opportunity for the evolution of sex-limited traits [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989].
-**In damselflies, differences between the sexes in their exposure to UV radiation drives dichromatism between the sexes [@Cooper_2010; @Punzalan_2010] I would erase this example and find stuff on birds** .
-Natural selection on females maybe important in explaining patterns of sexual dichromatism in birds ** cite: Hofmann CM, Cronin TW, Omland KE.
-Evolution of sexual dichromatism.
-1.
-Convergent losses of elaborate female coloration in New World Orioles (Icterus spp.).
-The Auk.
-2008;125:778–89.Price JJ, Eaton MD.
-Reconstructing the evolution of sexual dichromatism: current color diversity does not reflect past rates of male and female change.
-Evolution.
-2014;68(7):2026–37.** In fairy-wrens (Malurus cyaneus), female colouration has evolved in response to environmental pressures, leading to increased sexual dichromatism in regions where females are under stronger selection from predation (**cite: Medina et al.
-https://doi.org/10.1186/s12862-016-0861-3**).
+Specifically, natural selection on females may be important in explaining patterns of sexual dichromatism in birds [@Price_2014].
+In fairy-wrens (*Malurus cyaneus*), female colouration has evolved in response to environmental pressures, leading to increased sexual dichromatism in regions where females are under stronger selection from predation [@Medina_2017].
 Ecological impacts on sexual dichromatism are supported by our path analysis, which reveals that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation) albeit relatively weakly.
-Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition **cite: https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/1365-2656.12824**.
+Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
 And thirdly, not all plumage colouration honestly reflects mate quality.
@@
 
-In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ --- but this association was not dependent on variable or stressful environments.
+In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ estimates --- but this association was not dependent on variable or stressful environments.
 When evolving under stressful conditions, sexually antagonistic selection is hypothesised to be weakened relative to selection that is concordant between the sexes [@Connallon_SA_2015; @Connallon_Hall_2016].
 While these predictions are supported in several lab (experimental evolution) studies [@Long_2012; @Berger_2014; @Punzalan_2013], other studies have not found the association [@Delcourt_2009; @Martinossi-Allibert_Savković_2017; @Holman_Jacomb_2017].
+The environmental variability predictors used here often reflected changes in climate across space and time, whereby high environmental variation would likely increase the strength of natural selection on a given species.
 Given that we did not detect an effect of environmental variability (by itself or as an interaction with sexual selection) we are left with several possible explanations.
-** would be good to at least mention how/why we expected one of the environmenta variables to have an effect on speciation, up to here there are not much details in the causal relationship we were expecting with environment** Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability the species is evolving under.
+Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability the species is evolving under.
 Specifically, sexual antagonism may be more easily purged in environments where cyclic pressures (*e.g.* seasonality) are perpetuated; whereas when environmental variability is directional (*e.g.* long-term climatic change) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
@@
 The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates; however reduced range size can be a cause or consequence of speciation.
-** compare the effects??
-looks like association with sexual selection is stronger than that with range size, which is interesting**  Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995].
-However, **why however?
-seems in line with previous sentence** birds with smaller range sizes are also predicted to have lower dispersal ability.
+Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995].
+However, birds with smaller range sizes are predicted to have lower dispersal ability.
 This ensues that individuals might not be able to overcome geographical barriers and maintain gene flow; which would promote reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
@@
 Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives.
-Within Himalayan songbirds, range expansions are prevented by the competition for resources by related species [@Price_2014]; however, this limitation on range expansion is then expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011].
+Within Himalayan songbirds, range expansions are prevented by the competition for resources by related species [@Price_2014]; this supression on range expansion is then expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011].
 Unlike Himalayan songbirds, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
-Across islands we expect to see a correlation between speciation and small range size, not because of any inherent properties of the range size, but because they reflect newly formed islands with empty niches.
+Across islands we expect to see a correlation between speciation and small range size because they reflect newly formed environments with empty niches.
 Thus, our findings may be heavily dependent on species whose range size is restricted to a single island within an archipelago (*e.g.* *Geospiza* sp.; a.k.a.
@@
 One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
+Given the problems of determining causality and mode of speciation the reason for the negative association found here remains obscured.
 
@@
 Unfortunately, these methods often rely on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List [*e.g.*, @Morrow_2003; @Bro-Jorgensen_2014; @Morrow_2004] and may be biased to species under greater anthropogenic threat, culturally valued or well researched.
-** if you need to cut words I think that this paragraph could be just summarised **
 

### 2018-10-10 Sexual Selection and Speciation.Rmd
Renaming files

--- 2018-10-07 Sexual Selection and Extinction.Rmd
@@
 ---
-title: 'Sexual selection and speciation: a macro-evolutionary approach'
+title: 'Sexual selection, environmental variation, range size and speciation in passerine
+  birds'
 author: "Justin G. Cally^1^, Iliana Medina Guzmán^1^ Luke Holman^1^, Devi Stuart-Fox^1^ <br></br> <br></br> ^1^The University of Melbourne"
@@
 
-Here we provide the supplementary material for our study on sexual selection and extinction risk in passerine birds.
-
 #Compiling the data

### 2018-10-10 Manuscript.Rmd
Fix CV for BAMM

@@
 output:
+  pdf_document: default
+  html_document: default
   word_document:
     pandoc_args: --smart
-  html_document: default
-  pdf_document: default
 header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}

### 2018-10-10 Sexual Selection and Speciation.Rmd
Fix CV for BAMM

@@
 
-```{r, eval = F, echo = F}
+```{r, eval = T, echo = F, results='hide', warning=F}
 MCC.BAMM.df$CV.lambda <- sqrt(exp(MCC.BAMM.df$var.lambda) - 1)*100
@@
 
-#saveRDS(MCC.BAMM.df, 'data/MCC.BAMM.df.rds')
-
-MCC.BAMM.df %>% ggplot(aes(y = CV.lambda)) +
-  geom_boxplot(outlier.shape = NA)+
-  scale_y_continuous(trans = "log")
+saveRDS(MCC.BAMM.df, 'data/MCC.BAMM.df.rds')
+# 
+# MCC.BAMM.df %>% ggplot(aes(y = CV.lambda)) +
+#   geom_boxplot(outlier.shape = NA)+
+#   scale_y_continuous(trans = "log")
 ```

### 2018-10-11 Manuscript.Rmd
Luke MS edits 1

@@
 csl: References/AmNat.csl
-abstract: 'Sexual selection may influence patterns of diversification; yet controversy remains on whether sexual selection leads to a net increase or decrease in speciation rate, or even elevates extinction.
-Sexual selection is predicted to increase speciation rate as it can facilitate adaptation to new environments, and work alongside divergent natural selection to swiftly promote reproductive isolation.
-However, sexual selection frequently favours investment in costly or harmful traits that may reduce adaptation and affect diversification.
-Here, we investigate the association of sexual selection with speciation and extinction in passerine birds.
-We use two measures of sexual selection: sexual dichromatism (n = 5,812 species) and a multivariate measure of male-biased sexual selection (sexual dimorphism, polygyny and paternal care; n = 2,465), three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$).
-Importantly, we test whether the effects of sexual selection on speciation are dependent upon ecological pressures such as temperature variability across time and space, primary productivity and range size.
-We find that male-biased sexual selection but not sexual dichromatism predicts speciation.
-Additionally, we find that birds with smaller ranges have higher speciation rates; although the direction of causality is unclear.'
+abstract: 'Many theories related to sexual selection predict that it shapes phylogenetic diversity by altering speciation or extinction rates.
+The predicted effects on diversity are often contradictory; for example, sexual selection might promote speciation by creating reproductive isolation, or hinder it by selecting for traits that facilitate gene flow.
+Similarly, sexual selection might precipitate extinction by favouring competitive phenotypes with poor survival, or stave it off by helping populations shed deleterious mutations.
+Here, we investigate the relationship between sexual selection and rates of speciation and extinction in passerine birds.
+We use two measures of sexual selection -- sexual dichromatism (n = 5,812 species) and a multivariate measure of male-biased sexual selection (sexual dimorphism, polygyny and paternal care; n = 2,465) -- as well as three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$), and one measure of extinction ($\mu_{BAMM}$).
+We also tested whether the effects of sexual selection is contingent on ecological variables such as spatiotemporal temperature variation, primary productivity, and range size.
+We found a positive correlation between speciation rate and male-biased sexual selection, as well as a negative correlation between range size and speciation rate.
+**ONE MORE SENTENCE HERE**'
 ---
@@
 In this view, populations will speciate and adapt more readily to new and different environments if selection is strong and/or persistant.
-Hence, the rate at which speciation occurs is likely dependent on both the populations response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour @Hugall_2012; @Rabosky_morph_2013].
-Sexual selection is an evolutionary force likely to affect fitness components of a population [@Cally_in_review], such as reproductive success or viability, as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition [@McCullough_2016].
-Sexual selection is therefore hypothesised to be a key driver of diversification through shifting the rates of adaptation, speciation and extinction.
+Hence, the rate at which speciation occurs is likely dependent on both the population's response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour @Hugall_2012; @Rabosky_morph_2013].
+Sexual selection affects a diverse range of phenotypic traits important to population growth and persistance [@Cally_in_review], such as reproductive success or viability, as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition [@McCullough_2016].
+Sexual selection is therefore hypothesised to be a key driver of diversification, via its effects on adaptation, speciation and extinction.
 
-Sexual selection is applied by one sex on the other --- more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and the strength of sexual selection can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in* @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
-Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial mutations in a population [@Whitlock_2000].
-Thus, sexual selection is thought to be a key process in mitigating extinction risk and facilitating adaptation.
-Sexual selection can also increase the fitness of a population when the genotypes under sexual selection (*e.g.* those that increase male mating/fertilisation success) are genetically positively correlated with traits affecting fitness (*e.g.* female fecundity or survival).
-However, this positive correlation is not ubiquitous; indeed sexual selection frequently favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006].
+Sexual selection results from competition for matings, fertilisations, or associated resources, and is typically stronger in males than females [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008].
+Sexual selection has both positive and negative consequences for 'population fitness', i.e.
+the average absolute fitness of individuals in the population [e.g.
+@Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
+For example, sexual selection can elevate population fitness by helping to purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial ones [@Whitlock_2000], which should mitigate the risk of extinction due to mutation accumulation or a failure to adapte to changing environments.
+However, sexual selection does not always favour the evolution of traits that benefit the population...
+
+this positive correlation is not ubiquitous; indeed sexual selection frequently favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006].
 For instance, in species with high levels of sexual selection, mating success becomes more variable and uncertain; thus individuals must invest heavily in costly weaponry and sexual signals to increase their chances of successfully reproducing.
@@
 Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
-The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 100 random trees can be found within the ESM.
+The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 100 random trees can be found in the ESM.
 Density curves are based on estimates from 100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC tree.

### 2018-10-12 Manuscript.Rmd
Luke edits 2 - intro finished

--- 2018-10-11 Manuscript.Rmd
@@
 Hence, the rate at which speciation occurs is likely dependent on both the population's response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour @Hugall_2012; @Rabosky_morph_2013].
-Sexual selection affects a diverse range of phenotypic traits important to population growth and persistance [@Cally_in_review], such as reproductive success or viability, as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition [@McCullough_2016].
-Sexual selection is therefore hypothesised to be a key driver of diversification, via its effects on adaptation, speciation and extinction.
+Experiments suggest that sexual selection can influence the evolution of a surprisingly diverse set of traits [@Cally_in_review], and may therefore strongly influence extant patterns of phylogenetic diversity.
+**I removed the bit about weapons as it kinda came out of nowhere**
 
 Sexual selection results from competition for matings, fertilisations, or associated resources, and is typically stronger in males than females [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008].
-Sexual selection has both positive and negative consequences for 'population fitness', i.e.
+This competition can have both positive and negative consequences for 'population fitness', i.e.
 the average absolute fitness of individuals in the population [e.g.
 @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
-For example, sexual selection can elevate population fitness by helping to purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial ones [@Whitlock_2000], which should mitigate the risk of extinction due to mutation accumulation or a failure to adapte to changing environments.
-However, sexual selection does not always favour the evolution of traits that benefit the population...
+For example, sexual selection can elevate population fitness by helping to purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial ones [@Whitlock_2000], which should mitigate the risk of extinction due to mutation accumulation or a failure to adapt quickly to environmental change.
+However, sexual selection might also promote extinction, for example by selecting for male traits that have deterimental pleiotropic effects on female traits [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006], or selecting for male phenotypes with improved mating success but which harm population productivity, e.g.
+by harming females or under-investing in parental care **ref here - Kokko/Rankin maybe**.
 
-this positive correlation is not ubiquitous; indeed sexual selection frequently favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006].
-For instance, in species with high levels of sexual selection, mating success becomes more variable and uncertain; thus individuals must invest heavily in costly weaponry and sexual signals to increase their chances of successfully reproducing.
-Such high investment, may trade off against strategies that would improve offspring fitness (*e.g.* parental care).
-This is a costly consequence of sexual selection that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].
-
-Alongside changes to fitness, sexual selection has been proposed to facilitate speciation through the evolution of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982].
+** I think this next paragraph is quite muddled and needs work!**
+**Also I think you might be using 'per se' wrong, you seem to mean e.g. here**
+As well as affecting population demography, sexual selection has been proposed to facilitate speciation through the evolution of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982].
 For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism, are more species rich [@Seddon_2008].
@@
 However, the genetic architecture (*per se* heritability) of fitness and sexually selected traits may vary due to different sex determination systems across taxa.
-In male homogametic species (ZZ/ZW or ZZ/ZO) --- such as birds --- males are predicted to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003].
+In male homogametic species (ZZ/ZW or ZZ/ZO **Is there such a thing as ZO?**) --- such as birds --- males are predicted to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003].
 If fitness and associated mating traits are more heritable in birds we might expect bird populations experiencing stronger sexual selection to have increased adaptation and speciation rates compared to male heterogametic taxa.
 
+**Not sure if this next para is needed.
+You already talk about SS helping adaptation, so maybe merge your favourite parts of this with that part?
+Also I think the value of including all those environmental variables is to control for extraneous variation in speciation rate, and therefore get a better estimate of the importance of SS as distinct from that background**
 Importantly, the effects of sexual selection on adaptation and speciation may depend on environmental conditions.
@@
 
-Birds are notorious for the variety of ornaments that have evolved in response to sexual selection, such as plumage colour, size dimorphism, complex courtship rituals and male song [@Gontard_1999].
-The association between sexual selection and speciation has been investigated previously in birds using macroevolutionary approaches.
-A meta-analysis of 64 effect sizes from 20 studies found a small, but significant positive association between sexual selection and speciation across a wide range of taxa [@Kraaijeveld_2011].
-Within birds, they found a positive association but much lower than that for insects and fish.
-However, there was large variation in effect sizes due to variation in the measures of sexual selection used and diversification proxies (*e.g.* species/subspecies richness *vs* rate estimates).
-Contrary to this, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset [*n* = 918; @Armenta_2008]; similarly @Cooney_2017 found no effect of sexual dichromatism on diversification rates when using 1,306 pairs of species and human estimates of dichromatism.
-Given methodological advancements in tip-rate estimates [@Rabosky_diversification_2018], the association between sexual selection and speciation is worth revisiting in order to overcome previous limitations.
-Firstly, the strength of sexual selection in macroevolutionary studies is often estimated using measures of sexual dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa.
-Secondly, until recently, studies on the association between sexual selection and speciation have largely ignored the contribution and interaction with environmental variables to explain variation in speciation rates.
-Thirdly, many studies use incomplete taxon sampling, with family-level analysis restricted to well sampled avian clades [@Seddon_2008; @Huang_2014].Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague efforts to identify drivers of speciation and extinction.
+As a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumge, birds are a popular focus for macroevolutionary studies of sexual selection and diversification **refs**.
+A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation overall, though the average effect size was weaker in birds than in insects and fish [@Kraaijeveld_2011].
+The meta-analysis also found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of diversification and the strength of sexual selection.
+More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species), as measured used spectrophotometry **OF REAL BIRDS?** [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012].
+Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores from human observers.
 
-Here, we perform a macroevolutionary analysis of the relationship between sexual selection and speciation or extinction in passerine birds (Order: Passeriformes), arguably the largest group for which there exists comprehensive data.
-We build on these previous studies by using published datasets of sexual dichromatism measured from images [*n* = 5,812; @Dale_2015] alongside smaller datasets of sexual dichromatism measured from spectrophotometry [*n* = 581; @Armenta_2008] and a composite measure of male-biased sexual selection [*n* = 2,465; @Dale_2015].
-We assess the relationship between sexual selection, environmental variability and their interaction with three tip-rate measures of speciation and one measure of extinction ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$).
-Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species’ range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
+Here, we investigate the association between sexual selection and diversification in birds while extending previous work in multiple ways.
+Firstly, many macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual dimorphism or dichromatism, but these are imperfect correlates.
+We used a very large (*n* = 5,812) dataset of sexual dichromatism measures (measured spectrophotometrically from images; @Dale_2015]), as well as a smaller but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection that captures variation in **TRAIT NAMES HERE**.
+These datasets allow us to cover the great majority of bird genera, in contrast to previous work that has often focused on particular clades (*e.g.* [@Seddon_2008; @Huang_2014]).
+Secondly, we use the latest methods for estimating speciation and exinction rates at the tips of phylogenetic trees, including BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but less controversial methods such as **XXX**.
+We also conduct analyses that allow phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] to affect the precision of our estimates, rather than implicitly assuming that the single best-supported phylogenetic tree is correct.
+Thirdly, our analysis includes several potentially important environmental variables as predictors: range size, seasonal variation in temperature, spatial temperature variation (across a species’ range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
+By including these environmental variables in our models, we statistically account for their effects on variation in diversification rates and thereby improve our estimates of the effects of sexual selection.
+We also test whether the effects of sexual selection are contingent on the environment, as predicted by some theories (e.g.
+the theory that sexual selection is more likely to elevate popualtion fitness in changeable relative to static environments; **REFS**)
 

### 2018-10-12 Manuscript.Rmd
Title Change

--- 2018-10-11 Manuscript.Rmd
@@
 ---
-title: Sexual selection, environmental variation, range size and speciation in passerine
-  birds
+title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds
 author:
@@
   \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
-  \rhead[]{Sexual Selection and Speciation} \usepackage[noabbrev,capitalise]{cleveref}
+  \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref}
   \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
@@
   \raggedbottom
-csl: References/AmNat.csl
+csl: References/evolution.csl
 abstract: 'Many theories related to sexual selection predict that it shapes phylogenetic diversity by altering speciation or extinction rates.
@@
 We also tested whether the effects of sexual selection is contingent on ecological variables such as spatiotemporal temperature variation, primary productivity, and range size.
-We found a positive correlation between speciation rate and male-biased sexual selection, as well as a negative correlation between range size and speciation rate.
-**ONE MORE SENTENCE HERE**'
+We find that male-biased sexual selection but not sexual dichromatism predicts speciation.
+Additionally, we find that birds with smaller ranges have higher speciation rates; although the direction of causality is unclear.
+\vspace{5mm} \par\noindent \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage dichromatism, Range size, Extinction rate.'
 ---
@@
 Perhaps one of the biggest unravelings of this project was the decoupling of male-biased sexual selection from sexual dichromatism, remarkably this same controversy engrossed Wallace and Darwin 130 years ago [@Kottler_1980].
-The widespread use of sexual dichromatism largely goes unchallenged as our instinctive Darwinian mind reminisces of peacocks and birds of paradise; my thoughts were such when I began this scientific inquiry.
-Of course, I was expecting sexual dichromatism from RGB measures to be a noisy measure of sexual selection, but across 5,831 species I anticipated a signal would permeate through the noise.
-And despite the correlation between sexual dichromatism and male-biased sexual selection is present, the relationship is surprisingly weak (see \autoref{Path_Analysis} in main text).
+The widespread use of sexual dichromatism as a proxy for sexual selection often goes unchallenged as our instinctive Darwinian minds reminisce of peacocks and birds of paradise; my thoughts were such when I began this scientific inquiry.
+Although, sexual dichromatism measurements from colour drawings were always going to be a less than perfect estimate of sexual selection, across 5,831 species I anticipated a signal would permeate through the noise.
+Alas, this was not so and male-biased sexual selection and sexual dichromatism were very different measures (see \autoref{Path_Analysis} in main text).
 With this in mind, the practical use of sexual dichromatism in macroevolutionary research is severely weakened.
@@
 But, as Wallace remarks, the matter is inevitably more complex despite the irresistibility of the Darwinian argument\footnote{Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took more interest in a view where natural selection on predator-prone nesting females generated dullness in females but not males.} [@Kottler_1980; @Caro_2017].
-On the bright side, the multivariate measure of male-biased sexual selection that is: sexual dimorphism, social polygyny and [lack of] paternal care is a dataset that better predicts speciation and would benefit from being expanded across birds.
-
-<!-- \maketitle\unmarkedfntext{\textsuperscript{1}} -->
+On the bright side, the multivariate measure of male-biased sexual selection (sexual dimorphism, social polygyny and [lack of] paternal care) is a dataset that would benefit from being expanded across birds.
 
@@
 There were signs that extinction rate could be measured from phylogenetic methods with relative confidence [@Pyron_2016].
-However, these hopes soon dissipated, and with them the seeming \emph{raison d'\^{e}tre} of the project.
+However, these hopes soon dissipated, and with them the seeming \emph{raison d'\^{e}tre} of the study.
 Extinction rate is arguably the best measure of population fitness.
-And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured accurately over long, evolutionarily meaningful timescales [*but see* @Martins_2018].
+And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured accurately over evolutionarily meaningful timescales [*but see* @Martins_2018].
 We included measurements of extinction rate within the paper, however these findings were played down.
 The reason for this movement away from extinction and towards speciation is the controversy surrounding extinction rate measures as well as phylogentic uncertainty.
-Measuring extinction rates within the bird phylogeny is problematic because of the nature of phylogenetic construction; for passerine birds a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction rate [@Beaulieu_2015].
-Furthermore, the very detection of an absence of data --- which is what such methods attempt to do --- is a hard sell both conceptually and methodically come peer review: see @Rabosky_Extinction_2016 for a cautionary tale on measuring extinction rate with ``BAMM``.
+Measuring extinction rates within the bird phylogeny is problematic because of the nature of phylogenetic construction: for passerine birds a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction rate [@Beaulieu_2015].
+Furthermore, the very detection of absence of data --- which is what such methods attempt to do --- is a hard sell both conceptually and methodically come peer review: see @Rabosky_Extinction_2016 for a cautionary tale on measuring extinction rate with ``BAMM``.
 The mechanisms and rates at which speciation and extinction occur are an active area of research.
@@
 As with sexual dichromatism and macroevolutionary rates, the most severe limitation for environmental variation was the coarseness of the measures.
-In total, for 5,831 species we extracted 338,203,831 point measures\footnote{5,831 species x 1000 samples per range x 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.} of environmental variability, but our measures may not account for variation in biotic components such as nutritional resources or predation.
-However, the methods employed here are arguably much better than those traditionally implemented; where the range mid-point is used to obtain a single measure of temperature, precipitation or seasonality or where only one bioclimatic variable is used.
-I would argue that reducing the multivariate data into several biologically meaningful predictors is a conceptually challenging exercise, but researchers should not feel limited by the amount of raw data that they can access\footnote{This statement applies to species with documented range sizes, which varies across taxa but is comprehensive for birds, mammals, amphibians and many marine groups.
+In total, for 5,831 species we extracted 338,203,831 point measures\footnote{5,831 species x 1000 samples per range x 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.} of environmental variability, but these measures may not account for important variation in biotic components such as nutritional resources or predation.
+Arguably, the methods used here are much better than those traditionally implemented; where the range mid-point is used to obtain a single measure of temperature, precipitation or seasonality or where only one bioclimatic variable is used.
+I would argue that reducing the multivariate data into several biologically meaningful predictors is a conceptually challenging exercise, but researchers should not feel limited by the amount of raw data that they can access\footnote{This statement applies to species with documented range sizes, which varies across clades but is comprehensive for birds, mammals, amphibians and many marine groups.
 The \href{http://www.iucnredlist.org/technical-documents/spatial-data}{IUCN} has a large and accessible database for research use.
 Details of how we extracted data from these ranges is provided as a potentially useful resource in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.}.
-In fact, previous to our study on birds the same methodology was applied to ~ 100 species of bovids with speciation, extinction and environmental variability obtained.
+In fact, previous to our study on birds the same methodology was applied to ~ 100 species of bovids with speciation, extinction and environmental variability estimates obtained.
 This study was prematurely halted as there was homogeneity in the response variable (speciation and extinction rates) across the smaller bovid phylogeny.
@@
 
-I had initially envisaged that this comparative study would allow for the reliable measure of extinction rate, utilise multiple taxonomic groups and use more robust measures of sexual selection [*see* @Anthes_2016].
+I had initially envisaged that this comparative study would allow for the reliable measure of extinction rate, use multiple taxonomic groups and more robust measures of sexual selection [*see* @Anthes_2016].
 As such, many questions regarding this topic remain unanswered.
@@
 
-The study by @Martins_2018 has already been referred to several times in the main text of both papers and for good reason.
+The study by @Martins_2018 has already been referred to in the main text of both papers and for good reason.
 This elegant study was able to do what no one else has; reliably measure extinction rate over meaningful time scales.
@@
 From stratographic occurrence data a capture–mark–recapture model was used to determine the rate of extinction for 93 species.
-Their finding challenges the results from experimental evolution studies [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015] and our meta-analysis that find sexual selection in mitigates extinction.
-Moving forward, it would be incredibly valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual dimorphism to conduct such a study.
-In the interim I would be eager to see how ``BAMM`` would perform on an ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected among extant taxa.
-For me, this will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction from phylogenetic inference.
+Their finding challenges the results from experimental evolution studies [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015] and our meta-analysis that find sexual selection mitigates extinction.
+In future, it would be incredibly valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual dimorphism to conduct such a study.
+In the interim I would be eager to see how ``BAMM`` performs on an ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected from phylogenies.
+For me, this will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction rates across a wider variety of species.
 
 One aspect and only foreseeable shortcoming of the ostracod study is that it did not account for environmental variability; a challenging proposal for extinct species nevertheless.
-However the use of paleoclimate models in evolution is now a reality; with recent studies incorporating averaged values across multiple paleoclimate models as predictors of speciation [*see* @quintero_2018].
+However, the use of paleoclimate models in evolution is now a reality; with recent studies incorporating averaged values across multiple paleoclimate models as predictors of speciation [*see* @quintero_2018].
 While the uncertainty surrounding paleoclimates was beyond the scope of the study, future studies should remain vigilant of the everlasting need to incorporate key ecological processes in macroevolutionary research [*reviewed in* @Weber_2017].

### 2018-10-12 Sexual Selection and Speciation.Rmd
Title Change

@@
 ---
-title: 'Sexual selection, environmental variation, range size and speciation in passerine
-  birds'
+title: 'Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds'
 author: "Justin G. Cally^1^, Iliana Medina Guzmán^1^ Luke Holman^1^, Devi Stuart-Fox^1^ <br></br> <br></br> ^1^The University of Melbourne"
@@
 
-For this project we aimed to compile data on a large set of species in the Passerine bird order that would allow us to account for many evolutionary and environmental predictors of speciation.
+For this project we aimed to compile data on a large set of species in the passerine bird order that would allow us to account for many evolutionary and environmental predictors of speciation.
 The following details the source and use of the data.
@@
 
-The code for extracting the environmental data from the species is provided in another R-markdown document **Github link**, however no raw data is provided alongside this file due to the large file size of the shapefile and raster information.
+Below we outline the code used for extracting spatial data from bird ranges, however no raw data is provided alongside this file due to the large file size of the shapefile and raster information.
 We have put the extracted environmental variables in a csv filed called ``complete.dataframe.csv`` for convenience.

### 2018-10-14 Sexual Selection and Speciation.Rmd
Devi Edits to Supplementary Material

@@
 ---
-title: 'Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds'
-author: "Justin G. Cally^1^, Iliana Medina Guzmán^1^ Luke Holman^1^, Devi Stuart-Fox^1^ <br></br> <br></br> ^1^The University of Melbourne"
-subtitle: Supplementary Material
-bibliography: References/SS_Extinction.bib
+title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation
+  in birds
+author: Justin G. Cally^1^, Devi Stuart-Fox^1^, Luke Holman^1^, Iliana Medina Guzmán^1^
+  <br></br> <br></br> ^1^The University of Melbourne
 csl: References/AmNat.csl
@@
   html_document:
-    toc: true # table of content true
-    toc_float: true # make 
-    depth: 3  # upto three depths of headings (specified by #, ## and ###)
-    number_sections: false  ## if you want number sections at each table header
+    code_folding: hide
+    depth: 3
+    number_sections: no
     theme: cosmo
-    code_folding: hide # awesome buttons to show/hide the code
-  pdf_document: default
+    toc: yes
+    toc_float: yes
+subtitle: Supplementary Material
+bibliography: References/SS_Extinction.bib
 ---
@@
 
-For this project we aimed to compile data on a large set of species in the passerine bird order that would allow us to account for many evolutionary and environmental predictors of speciation.
+We aimed to compile data on Passerine birds, which represent the largest group for which relatively complete plylogenetic, trait and environmental data are available.
 The following details the source and use of the data.
@@
 To obtain speciation rate estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM).
-Additionally, we compared our BAMM output files (event data) against a pre-existing extinction rate dataset [@Harvey_2017].
-
-Rather than obtaining one estimate of speciation rate from one tree, we ran BAMM 100 times on 100 trees plus an MCC tree to obtain uncertainty estimates of the speciation rate generated from the phylogeny created by @Jetz_2012.
-We also estimated extinction rate from BAMM.
+Additionally, we compared our BAMM output files (event data)  to an existing extinction rate dataset [@Harvey_2017].
+
+Rather than obtaining one estimate of speciation rate from one tree, we ran BAMM 100 times on 100 trees plus an MCC tree to obtain uncertainty estimates of the speciation rate generated from the phylogeny produced by @Jetz_2012.
+We also estimated extinction rate from BAMM and speciation rates using the diversification rate and node density statistic ($\lambda_{DR}$ and ($\lambda_{ND}$).
+See @Rabosky_diversification_2018 for a comparison of methods.
 
@@
 Our proxy for sexual selection was based on measures of sexual dichromatism.
-Notably, we use two existing datasets; a complete set of male and female plumage scores in Passerines from @Dale_2015 and reflectance data from 1000 birds from @Armenta_2008.
-
-Briefly, @Dale_2015 obtained male and female plumage score, which can be divided to obtain a sexual dichromatism index (SDi).
+We used two existing datasets; a complete set of male and female plumage scores in Passerines from @Dale_2015 and reflectance data from 1000 birds from @Armenta_2008.
+
+the ratio of which provides an index of sexual dichromatism (SDi).
 The plumage score was scored using the _Handbook of the Birds of the World_ @delhoyo2011.
-By scanning images of males and females in this book across multiple patches @Dale_2015 were able to obtain mean plumage scores from RGB values.
-Notably concern about the accuracy of these measurements were raised, thus this dataset was previously compared with a more thorough reflectance dataset of Australian birds, the two datasets correlated but was expectedly noisy.
-Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008 using colour discriminability as the parameter.
-
-Alongside sexual dichromatism data there is also partial data for other metrics analysed in @Dale_2015 such as: Body size, tropical life history, Sexual selection, Cooperative breeding and Migration.
-We use a phylogenetic PCA (PC1) that estimates male-bias sexual selection from a species level of sexual size dimorphism, social polygyny and paternal care.
+By scanning images of males and females in this book across multiple patches, @Dale_2015 obtained mean plumage scores from RGB values.
+@Dale_2015 compared their RGB scores to a dataset of reflectance measurements of Australian birds.
+The two datasets were correlated but the relationship was expectedly noisy.
+Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008, which estimated sexual dichromatism as  the mean number of just noticeable differences (JNDs) between male and female plumage colour (discriminability), based on a model of bird colour vision.
+
+Alongside sexual dichromatism, @Dale_2015 compiled available data for traits such as: Body size, tropical life history, Sexual selection, Cooperative breeding and Migration.
+We use a phylogenetic PCA (PC1) that estimates male-biased sexual selection from a species level dataset of sexual size dimorphism, social polygyny and paternal care.
 
@@
 
-We obtained species range distribution maps from Birdlife International (through IUCN) [@birdlife_2017].
+We obtained species range distribution maps from Birdlife International [@birdlife_2017].
 These species range maps cover nearly all species of birds.
-From these spatial information files we were able to obtain estimates of the following (note that not all the data extracted was subsequently used in the analysis):
+From these spatial information files we obtained estimates of the following (note that not all the data extracted was subsequently used in the analysis):
 
@@
 
-#Generating biologically relevent predictors
-
-Here we assess the relationship between sexual selection and extinction risk.
-However, in doing so we attempt to take into account as many other predictors of extinction as possible, primarily through environmental variables.
-Basically, our model structure seeks to contain:
+#Generating biologically relevant predictors
+
+In this study we assess the relationship between sexual selection and extinction risk.
+In doing so we attempt to take into account as many other predictors of extinction as possible, primarily through environmental variables.
+Our model structure seeks to contain:
 
@@
 
-To obtain estimates of spatial variability we can obtain PCAs of the standard errors of the bioclimatic variables (excluding seasonality in temperature and precipitation).
-From there we can run GAM models with the PCA components and range size and extract residuals.
-We do this as we expect variability will increase alongside increased range size, thus taking the residuals allows us to correct for this association.
-
-**Table S1:** Loadings for the first two PCs, when we conduct a PCA on the variation (standard error, where n ~ 1,000) pf each bioclim variable except the standard error of temperature seasonality (se.bioclim4) and the standard error of precipitation seasonality (se.bioclim15), as these are already measures of variability and based on other bioclim variables.
-We use the standard error as, while we attempted to sample each species range 1,000 times, some points returned NA values and would thus affect the standard deviation, this inconvenience should not affect standard error.
+To obtain estimates of spatial variability we performed PCAs of the standard errors of the bioclimatic variables (excluding seasonality in temperature and precipitation).
+We expect variability to increase with increased range size.
+To correct for this association, we took the residuals from GAM models of the relationship between PCA components and range size.
+
+**Table S1:** Loadings for the first two PCs, from a PCA on the variation (standard error, where n ~ 1,000) of each bioclim variable except the standard error of temperature seasonality (se.bioclim4) and the standard error of precipitation seasonality (se.bioclim15), as these are already measures of variability and based on other bioclim variables.
+We use the standard error as, while we attempted to sample each species range 1,000 times, some points returned NA values and would thus affect the standard deviation, but not standard error.
 ```{r}
@@
 **Figure S1:** The relationship between spatial variability in the temperature components (PC1) and log-range size is relatively strong.
-But not as strong for spatial variability in precipitation (PC2).
-In the analysis we only used the residuals from PC1
+The relationship between spatial variability in precipitation (PC2) and log-range size is weaker.
+In the analysis we only used the residuals from PC1.
 
@@
 
-To gain estimates for change in climate over the past ~130,000 years we can use the difference in bioclim variables between the LIG and present values.
-The plots show the two PCs, broadly representing temperature and prescipitation.
-Data has also been provided for the last-glacial maximum (LGM) but was not used in the analysis.
+To gain estimates for change in climate over the past ~130,000 years we used the difference in bioclim variables between the LIG and present values.
+The plots show the two PCs, broadly representing temperature and precipitation.
+Data is also provided for the last-glacial maximum (LGM) but was not used in the analysis.
 
 **Table S2:** Loadings for the first two PCs of each of the PCAs for the difference in bioclimatic variables between today and the LIG.
-Here, PC1 is more heavily loaded for absoluted temperature difference, while PC2 is more heavily loaded for absolute difference in precipitation.
+Here, PC1 is more heavily loaded for absolute temperature difference, while PC2 is more heavily loaded for absolute difference in precipitation.
 
@@
 
-**Figure S2:** The distribution of the two PCA compontents for the absolute difference in bioclimatic variables between today and the LIG.
+**Figure S2:** The distribution of the two PCA components for the absolute difference in bioclimatic variables between today and the LIG.
 
@@
 
-We can check the correlations between environmental predictors that we will use in our model. Specifically we want to know whether the following environmental variables are correlated:
+We checked the correlations between environmental predictors used in subsequent phylogenetic comparative analyses (PGLS models).
+Specifically we tested whether the following environmental variables are correlated:
 
@@
 
-To check some of the correlations between the environmental predictors to be used in the PGLS models we can inspect the following correlation plot with the correlation value plotted:
+To check the correlations between the environmental predictors to be used in the PGLS models we inspected the following correlation plots with the correlation value plotted:
 
@@
 **Figure S3:** The highest correlation is between Long term temperature variation and seasonal variation.
-However this correlation is not excessively high (0.53) so we can be satisfied in including these predictors within our model.
+However, this correlation is moderate (0.53) so we can be confident that collinearity with not be an issue for our PGLS models.
 
@@
 
-The sexual dimorphism dataset looks to have an overdispersed distribution (see phylogenic plot in manuscript).
+The sexual dimorphism dataset looks to have an over-dispersed distribution (see phylogenic plot in manuscript).
 Unfortunately, transformations do not greatly improve the distribution.
-Our model fit is expected to be reduced by this but further transformations are not obvious and would risk problems of interperetation.
+Our model fit is expected to be reduced by this but further transformations are not obvious and would risk problems of interpretation.
 
@@
 
-Diversification rates can be simply estimated from thew 'tippiness of the phylogeny'.
-Here a log equal splits (also called DR) value can be achieved that is weighted to more recent diversification.
-It was used by @Jetz_2012 and recently (in comparison with BAMM) in a study ivestigation bird diversity alongside elevation [@quintero_2018].
-We can look at speciation rate using ES-Sim adapted by @Harvey_essim_2017 and found here https://github.com/mgharvey/ES-sim.
+We obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
+Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root-to-tip, while the diversification rate (DR) [*e.g.*, @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each more basal node having the sum of lengths down-weighted.
 
@@
 Here they are run on 100 trees plus a full MCC tree of passerine birds based on 2500 samples of the posterior [@Jetz_2012].
-This was done using the ``phanghorn`` [@phangorn_2010].
+This was done using the ``phanghorn`` package [@phangorn_2010].
 
@@
 
-To gain an estimate of variation between trees we can take the values and obtain a mean, variance and coeeficient of variation (CV).
+To gain an estimate of variation between trees we can take the values and obtain a mean, variance and coefficient of variation (CV).
 We obtain a CV value to account for unequal variation at higher estimates.
@@
 
-Furthermore, we can account for overdispersion by obtaining a transformed version of CV. In such a case we can calculate CV for a log-normal distribution:
+Furthermore, we can account for over-dispersion by obtaining a transformed version of CV. In such a case we can calculate CV for a log-normal distribution:
 
@@
 ```
-**Figure S4:** $\lambda_{DR}$ and $\lambda_{ND}$ are correlated in their estimation of species rate.
-Furthermore, we see similar variation across each species in the estimates across 100 trees as the inverse of the coefficeint of variation (weight) is approximately equivalent across all values of ND and DR (colours and size).
+**Figure S4:** $\lambda_{DR}$ and $\lambda_{ND}$ estimates of speciation rate are correlated.
+Furthermore, we see similar variation across each species in the estimates across 100 trees as the inverse of the coefficient of variation (weight) is approximately equivalent across all values of ND and DR (colours and size).
 Notably, these estimates speciation rate are relatively uncertain across all species, with 50 % CIs in both axes plotted.
-
-Combine the estimates of DR and ND with the restricted dataframe containing all the other variables. We can inspect the trend in SDi in relation to the speciation rates.
 
@@
 **Figure S5:** Scatter plots showing the raw relationship between sexual dichromatism (SDi) and speciation rates.
-Across all three measures the pattern and spread is similar, with no obvious relationship.
-Note that terminal branch lenght (mean.tb) was not used in the analysis.
+Across all three measures of speciation the pattern and spread is similar, with no obvious relationship.
+Note that terminal branch length (mean.tb) was not used in the analysis.
 
@@
 For the use of BAMM we used the following code to generate the parameters across the 100 trees.
-Each paramaeter value is specified in this code chunk to ensure reproducibility when re-running on any of the trees.
+Each parameter value is specified in this code chunk to ensure reproducibility.
 These same parameters were also used for the MCC tree (with the seed set at 2500).
@@
 
-Those values are well above 200, so we are good to use the MCC data in analysis!
+Those values are well above 200, so we can use the MCC data in analysis!
 We can also check the convergence of BAMM across 100 runs of BAMM.
@@
 To plot the credible shift set, we need the prior distribution on the number of rate shifts (this is generated internally by BAMMtools).
-We can then estimate the credible set of rate shifts using the BAMMtools function credibleShiftSet:
+We can then estimate the credible set of rate shifts using the BAMMtools function ``credibleShiftSet``:
 
@@
 
-css$number.distinct
-```
-
-This value indicates that from a posterior sample of 1000 we have 950 unique rate shifts, which is very high.
-Arguably, across such a large phylogenetic tree, even if the number of shifts reaches convergence the location of those > 50 shifts will be quite variable, with the variation likely impacting tip-rate estimates.
-Even if we increase the sampling of the posterior to 2,500 instead of 1,000 we still see that the CSS is high (close to 2500).
-We can have a look what our top few shifts are:
-
-```{r}
 summary(css)
@@
 Notably, each of the top shifts has low probability, indicating that out of the entire posterior sample we cannot differentiate which shift configuration is more likely.
-Based on the follwing eastimates of the BayesFactor, we are convinced that the number of shifts is non-zero, however it is still quite a wide distribution, sourced from potentially weakly non-informative priors.
-But, this is at odds with the ESS and the following plot:
+Based on the following estimates of the BayesFactor, we are convinced that the number of shifts is non-zero, however it is still quite a wide distribution.
 
@@
 ```
-**Figure S6:** The apparant convergence of the number of shifts in the posterior is at odds with the variability seen in the CSS and although there is greater certainty in the number of shifts being within the range above, the position of those shifts remain highly variable.
-
-We can compare our run of BAMM against @Harvey_2017 who used BAMM on a genetic-only MCC tree with different parameters
-
-```{r}
+**Figure S6:** The apparent convergence of the number of shifts in the posterior is at odds with the variability seen in the CSS and although there is greater certainty in the number of shifts being within the range above, the position of those shifts remains variable.
+
+We can compare our run of BAMM against @Harvey_2017 who used BAMM on a genetic-only MCC tree with different parameters.
+
+```{r, fig.width=10, fig.height=6}
 load('data/Hackett_split_eventsample.rda')
 css3 <- credibleShiftSet(ed, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95)
-css3$number.distinct
-```
-
-Based on Their posterior sample of 1250, 1188 are unique. Like our run on the MCC, we find high uncertainity in the position of shift-configurations.
-
-Let's run a model based on genetic data from the Harvey MCC
-
-```{r, fig.width=10, fig.height=6}
+
 Harvey.BAMM.df <- BAMM.extraction.function(ed)
@@
 The value is fixed for subsequent models as independently estimating it in each case becomes computationally intensive.
-Based on preliminary analysis even slight differences in lambda lead to large reductions in AIC.
-
-**2:** Create a global model of six predictor variables plus the fiver interactions between sexual dichromatism and the other variables.
+
+**2:** Create a global model of six predictor variables plus the five interactions between sexual dichromatism and the other variables.
 
@@
 
-The lambda is high and not too dissimilar than if we assume Brownian Motion $\lambda = 1$.
-However, given the large sample size this difference may have an effect on the results so, to play safe we included it as a fixed value for $\lambda$ in all successive models.
+The lambda is high and similar to if we assume Brownian Motion $\lambda = 1$.However, given the large sample size, this difference may have an effect on the results so we included it as a fixed value for $\lambda$ in all successive models.
 
@@
 **Table S5:** The dredged models both show the top model is one with no interactions, with $\delta AICc > 4$ in both cases.
-We can be resonably confident that interactions are unlikely to affect the pattern of speciation we see in passerine birds.
+We can be reasonably confident that interactions are unlikely to affect the pattern of speciation we see in passerine birds.
 ```{r}
@@
 From the dredged model list we can see that the top model is one with no interactions.
-Run this model on the MCC and 100 trees.
-Noting that each model uses a unique set of values for DR/ND and a unique tree in the correlation structure.
+We ran this model on the MCC tree and 100 trees, noting that each model uses a unique set of values for $\lambda_{DR}$/$\lambda_{ND}$ and a unique tree in the correlation structure.
 
@@
 
-From these models we can get a distribution of the estimates for the models alongside the 95 % CIs of the MCC model, this will show the variation beween trees, and the variation associated with the top MCC model.
+From these models we obtained a distribution of the estimates for the models alongside the 95 % CIs of the MCC model.
+This enabled us to show the variation between trees, and the variation associated with the top MCC model.
 
@@
 
-Using the previously generated BAMM dataframe we can now undertake model selection for speciation and extinction.
-The process is similar to that used on DR and ND, except given the use of a Bayesian approach in BAMM we can make use of varying levels of uncertainty between tips (species) by constructing a weighted model, where the weight is the inverse of the variance, meaning more precise estimates of speciation or extinction at a given tip (species) holds higher weight in the model.
-
-Based on preliminary findings we found that Pagal's lambda was = 1 and running ``corPagel`` lead to problemd of convergence.
-Therefore the following models are run assuming Brownian Motion with ``corBrownian``.
+Using the previously generated BAMM dataframe we undertook model selection for speciation and extinction.
+The process is similar to that used on DR and ND, except given the use of a Bayesian approach in BAMM we can make use of varying levels of uncertainty between tips (species) by constructing a weighted model, where the weight is the inverse of the variance, such that more precise estimates of speciation or extinction at a given tip (species) holds higher weight in the model.
+
+Based on preliminary findings we found that Pagel's lambda was = 1 and running ``corPagel`` lead to problems of convergence.
+Therefore we ran the following models assuming Brownian Motion with ``corBrownian``.
 
@@
 
-**Table S6:** The dredged models for BAMM speciation and BAMM extinction both show the top model is one with no interactions , with $\delta AICc > 4$.
+**Table S6:** The dredged models for BAMM speciation and BAMM extinction both show the top model is one with no interactions, with $\delta AICc > 4$.
 This is the same situation as DR/ND models (see above).
@@
 
-Given that we are running models weighted according to the variance of the response ($\lambda_{BAMM}$ and $\mu_{BAMM}$), we can check to see whether a weighted model isw favourable to an unweighted model using an ``anova`` to compare AIC values.
+Given that we ran models weighted according to the variance of the response ($\lambda_{BAMM}$ and $\mu_{BAMM}$), we checked to see whether a weighted model is favourable to an unweighted model using an ``anova`` to compare AIC values.
 
@@
 **Table S6:** Models with a set of weights for the response based on the inverse of the variance of the posterior distribution in $\lambda_{BAMM}$ and $\mu_{BAMM}$ have much lower AIC values, indicating that the weighting scheme improves the fit of the model.
-In any case, we found that an unweighted model did not change the main conclusions drawn in that they did not generate any statistically significant relationships.
+In any case, we found that an unweighted model did not change qualitative results.
 ```{r}
@@
 
-We can run the top (no interactions) model on the 100 trees, each with unique estimates of speciation and extinction from the BAMM runs.
+We ran the top (no interactions) model on the 100 trees, each with unique estimates of speciation and extinction from the BAMM runs.
 
@@
 ```
-**Figure S8:** This figure is the same basic figure as seen in the manuscript (FIgure 1), it provides model estimates for four response variables across 100 random trees alongside the MCC tree.
+**Figure S8:** This figure is the same basic figure as seen in the manuscript (Figure 1). It provides model estimates for four response variables across 100 random trees alongside the MCC tree.
 
@@
 
-From the estimates of 100 trees we can calculate HPD intervals.
+From the estimates of 100 trees we calculated Highest posterior density (HPD) intervals.
 Note that these do no account for the CI of each estimate.
@@
 
-**Table S8:** Highest posterior density intervals are calculated for the above figure (ie; models using RGB values of sexual dichromatism).
+**Table S8:**  HPD intervals calculated for the above figure (*i.e.* models using RGB values of sexual dichromatism).
 The HPD range is determined using the ``hdi`` function with a 95 % credible interval.
 These intervals do not take into account the variance associated with each interval and thus are not an estimate of model precision.
-Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that in the same direction (+ or -).
+Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that is in the same direction (+ or -).
 ```{r, warning = FALSE}
@@
 
-From the HPD 95 % intervals and the intervals for the 95 % CI we can gain an estimate of how much variability there is within a model in comparison to variability between trees.
-To do this we calculate the width of the HPD interval relativel to the 95 % CI interval from the MCC.
+From the HPD 95 % intervals and the intervals for the 95 % CI we can gain an indication of how much variability there is within a model in comparison to variability between trees.
+To do this we calculated the width of the HPD interval relative to the 95 % CI interval from the MCC tree.
 
 **Table S9:** Model estimates from $\lambda_{BAMM}$ and $\mu_{BAMM}$ tip-rate estimates show high levels of variability between trees relative to estimates from $\lambda_{DR}$ and $\lambda_{ND}$.
-The interval ratio is calculated as the wisth of the HPD 95 % CI relative to the MCC 95 % CI.
+The interval ratio is calculated as the width of the HPD 95 % CI relative to the MCC 95 % CI.
 A value of 1 suggests that model estimates from 95 % of the trees fall within the MCC 95 % CI.
@@
 
-Using the dataset from @Armenta_2008 we can conduct a subsetted analysis where dichromatism values are measured from spectrophotmetry and synthesised into a colour discriminability measure, which is a difference measure between the sexes.
-To make this dataset comprable with ours we use the absolute difference between the sexes.
-Therefore making the scale from monochromatism to dichromatism rather than female colouration to male colouration.
+Using the dataset from @Armenta_2008 we conducted the analysis on a subset of species for which dichromatism values were derived from spectrophotometry data.
+The measure of dichromatism is a difference measure between the sexes, based on a model of bird colour vision.
+To make this dataset comparable with the RGB measure of dichromatism, we use the absolute difference between the sexes; thereby making the scale from monochromatism to dichromatism rather than female colouration to male colouration.
 
@@
 
-**Figure S9:** There is a correlation between the RGB measures and the Spectophotometry measures, The RGB measures seem to be more noisy around the lower values.
+**Figure S9:** There is a correlation between the RGB measures and the Spectophotometry measures. The RGB measures seem to be more noisy around the lower values.
 
@@
 
-**Table S10:** The dredged models for Models using a subset dataset where sexual dichromatism was measured using spectrophotometry, all four show the top model is one with no interactions , with $\delta AICc > 4$.
+**Table S10:** The dredged models using a subset of species for which sexual dichromatism was measured using spectrophotometry.
+All four show the top model is one with no interactions, with $\delta AICc > 4$.
 ```{r}
@@
 
-Using the MCC models as well as the 100 PGLS models we can generate the plots similar to Supp Fig S8:
+Using the MCC models as well as the 100 PGLS models we can generate the plots similar to Figure S8:
 
@@
 ```
-**Figure S10:** Measures of sexual dichromatism using a more precise measure in spectrophotometry do not change the main patterns seen within the PGLS models analysed here.
-This dataset is a subset of the complete dataset (n = 581), thus drawing conclusions for the other predictors (e.g.
-borderline long term temperature variation and spatial temperature variation) potentially risks type I error; additionally these environmental predictors were not measured differently in this analysis compared to the previous above so we have no obvious reason to make conclusions from this subset for the effects of the environmental predictors.
+**Figure S10:** Measures of sexual dichromatism using spectrophotometry do not change the main patterns.
+This dataset is a subset of the complete dataset (n = 581), thus drawing conclusions for the other predictors (*e.g.* borderline long term temperature variation and spatial temperature variation) potentially risks type I error.
+Additionally, these environmental predictors were not measured differently in this analysis compared to the analysis using RGB measures of dichromatism for all species so we refrain from drawing conclusions on the effects of the environmental predictors based on this smaller dataset.
 
@@
 
-**Table S12:** Highest posterior density intervals are calculated for the above figure (ie; models using Spectrophotometry values of sexual dichromatism).
-The HPD range is determined using the ``hdi`` function with a 95 % credible interval.
-These intervals do not take into account the variance associated with each interval and thus are not an estimate of model precision.
-Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that in the same direction (+ or -).
-These intervals are calculated in the same way as in **Table S7**.
+**Table S12:**  HPD intervals calculated for the above figure (*i.e.* models using Spectrophotometry values of sexual dichromatism).
+These intervals do not take into account the variance associated with each estimate and thus are not an estimate of model precision.
+Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that are in the same direction (+ or -).
+These intervals are calculated in the same way as in Table S7.
 ```{r, warning = FALSE}
@@
 Sexual dichromatism is expected to be a good measure of sexual selection strength in birds.
-However the relationship between sexual dichromatism and male-biased measures of sexual selection (social mating system, sexual size dimorphism, and paternal care) is expected to be relatively noisy given the precision of the measurment used [@Dale_2015].
-Here we use a dataset of sexual selection for 2,465 species within the orginal dataset of sexual dichromatism scores from RGB measures.
-This male-bias sexual selection score is based on three components, combined in a phylogenetic pca (ppca) with the following loadings:
+However, the relationship between sexual dichromatism and male-biased measures of sexual selection (social mating system, sexual size dimorphism and paternal care) is expected to be relatively noisy given the precision of the measurement used [@Dale_2015].
+Here we use a dataset of sexual selection for 2,465 species of the 5,812 species with sexual dichromatism scores from RGB measures.
+This male-biased sexual selection score is based on three components, combined in a phylogenetic PCA (ppca) with the following loadings:
 
@@
 
-As such, high values for this score are expected to have high sexual selection (high dimorphism, high polygyny and low paternal care).
-PGLS models using this dataset are calculated in the same process as above.
+As such, species with high values for this score are expected to have high male-biased sexual selection (high dimorphism, high polygyny and low paternal care).
+PGLS models using this dataset were run using the same process as above.
 
@@
 ```
-**Figure S11:** The relationship between absolute sexual dichromatism and the sexual selection.
-Although positively correlated the distribution is very noisy with species having high sexual dichromatism and low sexual selection and visa versa.Colour.discriminability
+**Figure S11:** The relationship between absolute sexual dichromatism and male-baised sexual selection.
+Although positively correlated, the distribution is very noisy with some species having high sexual dichromatism and low male-biased sexual selection and vice versa.
+
 ```{r}
@@
 **Table S13:** All top models ($\delta AICc > 4$) using male-biased sexual selection measures do not contain interactions between sexual selection and the environmental variables.
-Thus interaction terms are not included in further analysis.
+Thus interaction terms were not included in further analysis.
 
@@
 ```
-**Figure S12:** Measures of sexual selection a phylogenetic pca of sexual dimorphism, social polygyny and paternal care provide some evidence to suggest sexual selection predicts extinction rate (DR and ND models).
-This figure is seen in Figure 1 within the manuscript
+**Figure S12:** PGLS analyses using measures of male-biased sexual selection (*n* = 2,465) using three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
+The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 100 random trees can be found in the ESM.
+Density curves are based on estimates from 100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC tree.
+For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to interpret the MCC 95 % CIs.
+This figure is part **b** of Figure 1 within the manuscript.
+n
 
@@
 
-**Table S15:** Highest posterior density intervals are calculated for the above figure (ie; models using Male-bias sexual selection measures).
+**Table S15:** HPD intervals are calculated for the above figure (i.e.
+models using Male-bias sexual selection measures).
 using this measure the models using $\lambda_{DR}$ have model estimates that fall on the positive side 95 % of the time in 100 trees.
@@
 $\lambda_{BAMM}$ also shows a lesser positive skew.
-The HPD range is determined using the ``hdi`` function with a 95 % credible interval.
-These intervals do not take into account the variance associated with each interval and thus are not an estimate of model precision.
-Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that in the same direction (+ or -).
-These intervals are calculated in the same way as in **Table S7** and **Table S9**.
+These intervals do not take into account the variance associated with each estimate and thus are not an estimate of model precision.
+Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter are in the same direction (+ or -).
+These intervals are calculated in the same way as in Table S7 and Table S9.
 ```{r, warning = FALSE}
@@
 We undertook a phylogenetic path analysis using a model of causal pathways *a priori*.
-To do this we use the ``phylopath`` function.
-Our path analysis is undertaken on the dataset subset to species with measures of male bias sexual selection.
-Additionallty, the path analysis is restricted to using one measure of speciation ($\lambda_{DR}$) and the MCC tree.
-Our reasoning for the causal links are described breifly below:
-
-1) **$\lambda_{DR}$** dependent upon **Sexual Dichromatism**: Sexual dichromatism would directly impact speciation if there is an association with the evolvability of chromatism and speciation, this may more rapidly lead to niche divergence or reproductive isolation.
+To do this we used the ``phylopath`` function.
+Our path analysis was undertaken on the dataset subset to species with measures of male-biased sexual selection.
+Additionally, the path analysis was restricted to using one measure of speciation ($\lambda_{DR}$) and the MCC tree.
+Our reasoning for the causal links are described briefly below:
+
+1) **$\lambda_{DR}$** dependent upon **Sexual Dichromatism**: Sexual dichromatism would directly impact speciation if there is an association with the evolvability of plumage colour and speciation.
+This may more rapidly lead to niche divergence or reproductive isolation..
 2) **$\lambda_{DR}$** dependent upon **Sexual Selection**: This was a major hypothesis in our study (see introduction).  
 3) **$\lambda_{DR}$** dependent upon **Temperature Seasonality**: More variable environments are broadly expected to impact the resources, niches and fitness of species; thus there is an expectation that more variable environments may impact speciation rate.
-4) **$\lambda_{DR}$** dependent upon **Range Size**: As previously seen in the PGLS models range size shows a negative correlation with speciation rate; expected reasons for this causal link include lower dispersal ability in smaller ranges correlating with lower gene flow and greater speciation.
-It may also reflect niche specialisation and thus in line with the previous point increase speciation and reduce gene flow.
+4) **$\lambda_{DR}$** dependent upon **Range Size**: As previously seen in the PGLS models, range size shows a negative correlation with speciation rate.
+Expected reasons for this causal link include lower dispersal ability for species with smaller ranges correlating with lower gene flow and greater speciation.
+It may also reflect niche specialisation and thus, in line with the previous point, increase speciation and reduce gene flow.
 5) **Sexual Dichromatism** dependent upon **Sexual Selection**: Higher levels of sexual selection will promote showy males and drab females (sexual dichromatism).  
@@
 7) **Sexual Selection** dependent upon **Temperature Seasonality**: Greater environmental variability/stress may impact the usefulness (fitness effects) of sexual selection (see introduction).  
-8) **Range Size** dependent upon **Sexual Selection**: Sexual selection may often depend on local adaptations and the alignment of sexually selected traits with locally valuable traits, range size may limit this leading to a negative correlation.
-Alternatively if sexual selection has added fitness and evolvability benefits there may be increased range sizes in sexually selected species.
-9) **Range Size** dependent upon **Temperature Seasonality**: Anticipated to be more of a correlation than causal link.
-Yet a speecies range size may be dependent upon how variable their environment is and the availability to local resources.
+8) **Range Size** dependent upon **Sexual Selection**: Sexual selection may often depend on local adaptations and the alignment of sexually selected traits with traits increase local fitness.
+Thus, species with smaller ranges may have higher rates of local adaptation, with sexual selection further promoting the maintenance of local adaptations, restricted migration and thus smaller range sizes.
+Alternatively, if sexual selection has added fitness and evolvability benefits, it may enable species to colonise a greater range of environments leading to larger range size.
+9) **Range Size** dependent upon **Temperature Seasonality**: Anticipated to be more of a correlation than causal link; however a species range size may depend on the extent of environmental variation and availability of local resources.
 
@@
 
-**Figure S13:** Path analysis standardized regression coefficients vary across relationships.
+**Figure S13:** Path analysis standardised regression coefficients vary across relationships.
 Error bars are derived from confidence intervals through 500 bootstrapped iterations.
-Paths with error bars not everlapping zero are presented with an asterix in Figure 3 within the manuscript.
+Paths with error bars not overlapping zero are presented with an astericks in Figure 3 within the manuscript.
 

### 2018-10-15 Sexual Selection and Speciation.Rmd
Just a few changes. check the numbering of the figs after s6.

@@
 
-We aimed to compile data on Passerine birds, which represent the largest group for which relatively complete plylogenetic, trait and environmental data are available.
+We aimed to compile data on Passerine birds, which represent the largest group for which relatively complete phylogenetic, trait and environmental data are available.
 The following details the source and use of the data.
@@
 To obtain speciation rate estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM).
-Additionally, we compared our BAMM output files (event data)  to an existing extinction rate dataset [@Harvey_2017].
+Additionally, we compared our BAMM output files (event data)  to an existing rate dataset [@Harvey_2017].
 
@@
 
-##Sexual Selection
-
-Our proxy for sexual selection was based on measures of sexual dichromatism.
+##Sexual dichromatism and Male-biased Sexual Selection
+
+The first proxy for sexual selection that we used was based on measures of sexual dichromatism.
 We used two existing datasets; a complete set of male and female plumage scores in Passerines from @Dale_2015 and reflectance data from 1000 birds from @Armenta_2008.
 
-the ratio of which provides an index of sexual dichromatism (SDi).
-The plumage score was scored using the _Handbook of the Birds of the World_ @delhoyo2011.
+Briefly, @Dale_2015 obtained male and female plumage score, the ratio of which provides an index of sexual dichromatism (SDi).
+The plumage score was extracted using the _Handbook of the Birds of the World_ @delhoyo2011.
 By scanning images of males and females in this book across multiple patches, @Dale_2015 obtained mean plumage scores from RGB values.
 @Dale_2015 compared their RGB scores to a dataset of reflectance measurements of Australian birds.
-The two datasets were correlated but the relationship was expectedly noisy.
+The two datasets were correlated but the relationship was expectedly noisy **include R square of correlation**.
 Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008, which estimated sexual dichromatism as  the mean number of just noticeable differences (JNDs) between male and female plumage colour (discriminability), based on a model of bird colour vision.
@@
 Alongside sexual dichromatism, @Dale_2015 compiled available data for traits such as: Body size, tropical life history, Sexual selection, Cooperative breeding and Migration.
-We use a phylogenetic PCA (PC1) that estimates male-biased sexual selection from a species level dataset of sexual size dimorphism, social polygyny and paternal care.
+As a second measure of sexual selection, we used a phylogenetic PCA (PC1) that estimates male-biased sexual selection from a species level dataset of sexual size dimorphism, social polygyny and paternal care.
+Higher values indicate higher size dimorphism (larger males), higher polygyny and less paternal care.
 
@@
 ```{r, eval = FALSE}
-#Obtain means
+#Obtain means per variable per species
 bird.frame <- lapply(bird.values, function(x) {as.data.frame(x)})
@@
 
-**4)** This process was repeated when we used gridded data for the LGM, LIG, NPP and human population density, while range size was obtained from the following code:
+**4)** The process above was repeated when we used gridded data for the LGM, LIG, NPP and human population density, while range size was obtained from the following code:
 
@@
 
-##Spatial Variability PCAs
-
-To obtain estimates of spatial variability we performed PCAs of the standard errors of the bioclimatic variables (excluding seasonality in temperature and precipitation).
+##Environmental Spatial Variability PCAs
+
+To obtain estimates of spatial variability in environment we performed PCAs of the standard errors of the bioclimatic variables across the 1000 random points extracted per species (excluding seasonality in temperature and precipitation).
 We expect variability to increase with increased range size.
-To correct for this association, we took the residuals from GAM models of the relationship between PCA components and range size.
-
-**Table S1:** Loadings for the first two PCs, from a PCA on the variation (standard error, where n ~ 1,000) of each bioclim variable except the standard error of temperature seasonality (se.bioclim4) and the standard error of precipitation seasonality (se.bioclim15), as these are already measures of variability and based on other bioclim variables.
-We use the standard error as, while we attempted to sample each species range 1,000 times, some points returned NA values and would thus affect the standard deviation, but not standard error.
+To correct for this association, and to be able to include both variables in the model, we took the residuals from GAM models of the relationship between PCA components and range size.
+
+**Table S1:** Loadings for the first two PCs, from a PCA on the variation (standard error, where n ~ 1,000) of each bioclim variable except the standard error of temperature seasonality (se.bioclim4) and the standard error of precipitation seasonality (se.bioclim15), as these are already measures of variability and based on other bioclimatic variables.
+We use the standard error as, while we attempted to sample each species range 1,000 times, some points returned NA values and would thus affect the standard deviation, but not standard error.** mmm..SD is the sum of the distances of each point to the mean, divided by n-1, so it already takes into account whether some species have more or less points, it's like an average of distances to the mean.
+I think the standard way of measuring spread data is the SD more than the SE...anyway, I don't think that will change much if all Ns are very similar, but let's think about that for the publication**
 ```{r}
@@
 ```{r, fig.width=10, fig.height=5}
-#run GAM
-PC1.gam <- mgcv::gam(PC1*(-1) ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian") #Inverse as the PC1 loads to the negative (COUNTER-NTUITIVE)
+#run GAM of association between range size and spatial climatic variation
+PC1.gam <- mgcv::gam(PC1*(-1) ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian") #Inverse as the PC1 loads to the negative (COUNTER-INTUITIVE)
 
@@
 In the analysis we only used the residuals from PC1.
+**How much % is explained by PC1??**
 
@@
 
+Climate stability through time can potentially affect diversification dynamics.
 To gain estimates for change in climate over the past ~130,000 years we used the difference in bioclim variables between the LIG and present values.
@@
 Data is also provided for the last-glacial maximum (LGM) but was not used in the analysis.
+**last sentence not needed, or otherwise explain why we chose LIG over LGM?**
 
@@
 Our model fit is expected to be reduced by this but further transformations are not obvious and would risk problems of interpretation.
+In any case, the phylogenetically corrected models we employ do not assume normality in the response variable.
 
@@
 The functions that generate these tip rates estimates are: ``calculate.log.es``, ``calculate.nd`` and ``calculate.tb``.
-Here they are run on 100 trees plus a full MCC tree of passerine birds based on 2500 samples of the posterior [@Jetz_2012].
+Here they are run on 100 trees plus a full MCC tree of passerine birds based on 2500 sample trees of the posterior [@Jetz_2012].
 This was done using the ``phanghorn`` package [@phangorn_2010].
@@
 ```
-**Figure S5:** The tip-rate estimate for BAMM are highly variable within each run of BAMM.
+**Figure S6:** **this changes to S6 and next to S7 etc**The tip-rate estimates for BAMM are highly variable within each run of BAMM.
 Across most species there is high variability in the posterior distribution of tip-rate estimates.
@@
 
-Given the high variability in tip-rate estimates from the above plot, below we peform some diagnostics on BAMM to demonstrate that the variability is unlikely an error in sampling or parameters, rather an inherent aspect of BAMM.
+Given the high variability in tip-rate estimates from the above plot, below we peformed some diagnostics on BAMM to demonstrate that the variability is unlikely an error in sampling or parameters, rather an inherent aspect of BAMM.
 The following is not inherently necessary to understand the conclusions drawn from the paper, however they do raise a set of methodological questions about BAMM that warrant further investigation.
@@
 Notably, each of the top shifts has low probability, indicating that out of the entire posterior sample we cannot differentiate which shift configuration is more likely.
-Based on the following estimates of the BayesFactor, we are convinced that the number of shifts is non-zero, however it is still quite a wide distribution.
+Based on the following estimates of the BayesFactor, we are confident that the number of shifts is non-zero, however it is still quite a wide distribution.
 
@@
 ```
-**Figure S6:** The apparent convergence of the number of shifts in the posterior is at odds with the variability seen in the CSS and although there is greater certainty in the number of shifts being within the range above, the position of those shifts remains variable.
+**Figure S7:** The apparent convergence of the number of shifts in the posterior is at odds with the variability seen in the CSS and although there is greater certainty in the number of shifts being within the range above, the position of those shifts remains variable.
 
@@
 ```
-**Figure S7:** Based on the event data from the BAMM run by @Harvey_2017 we find that in both our case and theirs the tip-rate estimates for many species are extrememly variable.
-here we present mean estimates with 50 % CIs.
+**Figure S8:** Based on the event data from the BAMM run by @Harvey_2017 we find that in both our case and theirs the tip-rate estimates for many species are extremely variable across samples of the posterior probability.
+Here we present mean estimates with 50 % CIs.
 

### 2018-10-16 Manuscript.Rmd
Changes to manuscript

@@
 Here, we investigate the relationship between sexual selection and rates of speciation and extinction in passerine birds.
-We use two measures of sexual selection -- sexual dichromatism (n = 5,812 species) and a multivariate measure of male-biased sexual selection (sexual dimorphism, polygyny and paternal care; n = 2,465) -- as well as three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$), and one measure of extinction ($\mu_{BAMM}$).
+We use two measures of sexual selection -- sexual dichromatism (*n* = 5,812 species) and a multivariate measure of male-biased sexual selection (sexual dimorphism, polygyny and paternal care; *n* = 2,465) -- as well as three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$), and one measure of extinction ($\mu_{BAMM}$).
 We also tested whether the effects of sexual selection is contingent on ecological variables such as spatiotemporal temperature variation, primary productivity, and range size.
@@
 In this view, populations will speciate and adapt more readily to new and different environments if selection is strong and/or persistant.
-Hence, the rate at which speciation occurs is likely dependent on both the population's response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour @Hugall_2012; @Rabosky_morph_2013].
+Hence, the rate at which speciation occurs is likely dependent on both the population's response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.*, morphology or colour @Hugall_2012; @Rabosky_morph_2013].
 Experiments suggest that sexual selection can influence the evolution of a surprisingly diverse set of traits [@Cally_in_review], and may therefore strongly influence extant patterns of phylogenetic diversity.
-**I removed the bit about weapons as it kinda came out of nowhere**
 
 Sexual selection results from competition for matings, fertilisations, or associated resources, and is typically stronger in males than females [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008].
-This competition can have both positive and negative consequences for 'population fitness', i.e.
-the average absolute fitness of individuals in the population [e.g.
-@Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
-For example, sexual selection can elevate population fitness by helping to purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial ones [@Whitlock_2000], which should mitigate the risk of extinction due to mutation accumulation or a failure to adapt quickly to environmental change.
-However, sexual selection might also promote extinction, for example by selecting for male traits that have deterimental pleiotropic effects on female traits [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006], or selecting for male phenotypes with improved mating success but which harm population productivity, e.g.
-by harming females or under-investing in parental care **ref here - Kokko/Rankin maybe**.
+This evolutionary force has been proposed to facilitate speciation through the evolution of divergent signals associated with improved mating success [@Lande_1981; @Lande_1982].
+For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism, are more species rich [@Seddon_2008].
+Under this model of speciation, divergent mating signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].
+Alongside facilitating the evolution of divergent mating signals, sexual selection can increase the extent populations diverge by enlarging the available phenotypic space for mating signals.
+These predictions emerge as sexual selection may increase genetic diversity as a result of trade-offs between naturally- and sexually-selected traits or trade-offs under *intra*-locus sexual conflict between male and female trait expressions [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
+Additionally, sexual selection can have both positive and negative consequences for adaptation [*e.g.*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
+For example, sexual selection can elevate average fitness by helping to purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial ones [@Whitlock_2000], which should mitigate extinction risk and facilitate adaptation to environmental change.
+However, sexual selection might also promote extinction, by selecting for male traits that have deterimental pleiotropic effects on female traits [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006], or selecting for male phenotypes with improved mating success but which harm population productivity, such as under-investing in parental care [@RN277].
 
-** I think this next paragraph is quite muddled and needs work!**
-**Also I think you might be using 'per se' wrong, you seem to mean e.g. here**
-As well as affecting population demography, sexual selection has been proposed to facilitate speciation through the evolution of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982].
-For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism, are more species rich [@Seddon_2008].
-Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].
-Concurrently, sexual selection can increase the extent to which populations diverge by enlarging the phenotypic space diverging species can evolve into.
-These predictions emerge as sexual selection may increase genetic diversity as a result of trade-offs between naturally- and sexually-selected traits or trade-offs under intra-locus sexual conflict between male and female trait expressions [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
-However, the genetic architecture (*per se* heritability) of fitness and sexually selected traits may vary due to different sex determination systems across taxa.
-In male homogametic species (ZZ/ZW or ZZ/ZO **Is there such a thing as ZO?**) --- such as birds --- males are predicted to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003].
-If fitness and associated mating traits are more heritable in birds we might expect bird populations experiencing stronger sexual selection to have increased adaptation and speciation rates compared to male heterogametic taxa.
-
-**Not sure if this next para is needed.
-You already talk about SS helping adaptation, so maybe merge your favourite parts of this with that part?
-Also I think the value of including all those environmental variables is to control for extraneous variation in speciation rate, and therefore get a better estimate of the importance of SS as distinct from that background**
 Importantly, the effects of sexual selection on adaptation and speciation may depend on environmental conditions.
@@
 Conversely, benign or stable environments ensure consistent selection, preferentially eroding genetic variation at sexually concordant loci (*i.e.* loci where the optimum genotype is equivelant for both sexes).
-In this situation, genetic variation remains disproportionately at sexually antagonistic loci; causing displacement of male and female traits from their optimum.
-It follows from these predictions that populations with stronger sexual selection would have increased fitness and adaptive benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
-In certain circumstances, sexual selection will reinforce locally adaptated phenotypes through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011].
-Additionally, if mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to speed-up the evolution of divergent mating signals [@Lande_1981].
-In several bird taxa, beak morphology is a trait that arose under natural selection and is now a trait under sexual selection [a.k.a.
-"magic traits"; @Servedio_2011].
+In these stable environments, genetic variation remains disproportionately at sexually antagonistic loci; causing displacement of male and female traits from their optimum.
+It follows from these predictions that populations with stronger sexual selection would have fitness benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
+Additionally, sexual selection can reinforce locally adaptated phenotypes through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011].
+For instance, beak morphology is a trait that arose under natural selection in several taxa and is now a trait under sexual selection [a.k.a.
+*magic traits*; @Servedio_2011].
 In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-Therefore, strong sexual selection and strong natural selection --- such as environmental change over space and time --- may work synergestically to increase speciation rates.
+Through these synergenetic effects sexual selection and natural selection --- such as environmental change over space and time --- may lead to increased adaptation and speciation rates.
 
-<!-- Alternatively, sexual selection may interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signal efficacy in sticklebacks [@Wong_2007; @Candolin_2007].
-In this conflicting view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.
--->
-
-As a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumge, birds are a popular focus for macroevolutionary studies of sexual selection and diversification **refs**.
-A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation overall, though the average effect size was weaker in birds than in insects and fish [@Kraaijeveld_2011].
-The meta-analysis also found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of diversification and the strength of sexual selection.
-More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species), as measured used spectrophotometry **OF REAL BIRDS?** [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012].
+As a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumge, birds are a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g*, @Seddon_2008; @Huang_2014; @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003].
+A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, though the average effect size was weaker in birds than in insects and fish but stronger than in mammals [@Kraaijeveld_2011].
+Birds are often thought to be more conducive in benefiting from sexual selection due to their sex determination system.
+In male homogametic species (ZZ/ZW) --- such as birds --- males are predicted to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003].
+As such, we might expect bird populations experiencing stronger sexual selection to have increased adaptation and speciation rates compared to male heterogametic taxa.
+However, @Kraaijeveld_2011 found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of speciation and the strength of sexual selection.
+More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) when using spectrophotometry measures on taxonomic specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012].
 Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores from human observers.
@@
 Firstly, many macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual dimorphism or dichromatism, but these are imperfect correlates.
-We used a very large (*n* = 5,812) dataset of sexual dichromatism measures (measured spectrophotometrically from images; @Dale_2015]), as well as a smaller but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection that captures variation in **TRAIT NAMES HERE**.
-These datasets allow us to cover the great majority of bird genera, in contrast to previous work that has often focused on particular clades (*e.g.* [@Seddon_2008; @Huang_2014]).
-Secondly, we use the latest methods for estimating speciation and exinction rates at the tips of phylogenetic trees, including BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but less controversial methods such as **XXX**.
+We used a very large (*n* = 5,812) dataset of sexual dichromatism measures [sourced from illustrated drawings; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection, capturing variation in sexual dimorphism, social polygyny and [lack of] paternal care.
+These datasets allow us to cover the great majority of bird genera, in contrast to previous work that has often focused on particular clades [*e.g.* @Seddon_2008].
+Secondly, we use the latest methods for estimating speciation and exinction rates at the tips of phylogenetic trees, including BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but less controversial tip-rate statistics, commonly reffered to as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
 We also conduct analyses that allow phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] to affect the precision of our estimates, rather than implicitly assuming that the single best-supported phylogenetic tree is correct.
@@
 By including these environmental variables in our models, we statistically account for their effects on variation in diversification rates and thereby improve our estimates of the effects of sexual selection.
-We also test whether the effects of sexual selection are contingent on the environment, as predicted by some theories (e.g.
-the theory that sexual selection is more likely to elevate popualtion fitness in changeable relative to static environments; **REFS**)
+We also test whether the effects of sexual selection are contingent on the environment, as predicted by some theories [*e.g.* the theory that sexual selection is more likely to elevate adaptation in changeable relative to static environments; @Long_2012; @Connallon_Hall_2016].
 
@@
 Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014].
-This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate a probability distribution of evolutionary rate-shift configurations with variable speciation and extinction rates.
+This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate probability distributions of evolutionary rate-shift configurations with variable speciation and extinction rates.
 Importantly, these models provide tip-rate estimates of speciation and extinction rate.
@@
 Finally, using data subsetted for species with measurements of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
-The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
+The phylogenetic path analysis was used to assess causal paths *between* variables not able to be modelled within the univariate response of PGLS.
 That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection and range size to have effects on each other and not just on speciation rate.
@@
 Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-PGLS analyses using subsetted data (*n* = 581), where sexual dichromatism was measured using spectrophotometry [@Armenta_2008] yielded results concordant to the full data set; that is, no association between sexual dichromatism and speciation or extinction (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
+PGLS analyses using subsetted data (*n* = 581), where sexual dichromatism was measured using spectrophotometry [@Armenta_2008] yielded results concordant to the full data set; *i.e.* no association between sexual dichromatism and speciation or extinction (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
 Our results from MCC models are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 100 random trees.
-Model estimates are normally distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, see HPD (highest posterior density) intervals in \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
+The HPD (highest posterior density) intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
 For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (`r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
@@
 Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subsetted for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
-Model using data subsetted for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+Models using data subsetted for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
 Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with highest power using near-complete taxon sampling (*n* = 5,812).
@@
 Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; \href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S13}).
-Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from male-biased sexual selection to sexual dichromatism was relatively weak, ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`), despite dichromatism being widely used as an indicator of the strength of sexual selection.
+Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from male-biased sexual selection to sexual dichromatism was relatively weak, ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
 Additionally, the large direct effect of temperature seasonality on range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`) suggests an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}), given the negative we identified between $\lambda_{DR}$ and range size in PGLS models.
@@
 To account for this variation, weights (using the inverse of the variance) were used for the PGLS models.
-From 1000 posterior samples of the MCC BAMM run the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figures S5} and \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S7}).
+From 1,000 posterior samples of the MCC BAMM run the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figures S5} and \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S7}).
 Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) was high.
@@
 In fairy-wrens (*Malurus cyaneus*), female colouration has evolved in response to environmental pressures, leading to increased sexual dichromatism in regions where females are under stronger selection from predation [@Medina_2017].
-Ecological impacts on sexual dichromatism are supported by our path analysis, which reveals that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation) albeit relatively weakly.
+Ecological impacts on sexual dichromatism are supported by our path analysis, which reveals that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit relatively weakly.
 Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
-And thirdly, not all plumage colouration honestly reflects mate quality.
-A recent meta-analysis found that converted carotenoids (but not those sourced from diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
+And lastly, not all plumage colouration honestly reflects mate quality.
+A recent meta-analysis found that converted carotenoids (but *not* those sourced from diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
 Given the importance of honest signaling in sexual selection and the variability in colour production across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
@@
 Secondly, the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
-Specifically, access to dietary resources, and the impacts of predation or parasitism are unaccounted for here: arguably these are key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
+Specifically, access to dietary resources, and the impacts of predation or parasitism are unaccounted for here, these are likely key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
 Thirdly, it is also possible that there is no effect of environmental variability on speciation rates; at least not in the species investigated here.
-
-<!-- **Make it about evolvability** -->
-<!-- The effects of sexual selection are complex and are expected to vary amongst taxa.
-Although birds have a sex determination system conducive to increased father-to-son heritability of male-specific fitness traits on the Z chromosome [@Hastings_1994; @Reeve_2003] there are counteravailing arguments suggesting that deleterious recessive alleles on the Z chromosome in males (ZZ) are less exposed to selection, thus limiting the evolvability of traits [@Charlesworth_1987; @Charlesworth_2018].
-In a macro-evolutionary meta-analysis on sexual selection and speciation rate @Kraaijeveld_2011 found no significant taxon based differences across a diverse sample of vertebrates and invertebrates (fish, insects, birds, spiders, reptiles, mammals).
-Yet sexual selection in male homogametic taxa (birds) produced noticibly faster speciation rates than in male heterogametic taxa **Would be good to get raw data off them and do a pair-wise comparison**.
-The evolvability of the   -->
 
@@
 Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives.
-Within Himalayan songbirds, range expansions are prevented by the competition for resources by related species [@Price_2014]; this supression on range expansion is then expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011].
+Within Himalayan songbirds, range expansions are prevented by the competition for resources by related species [@Price_2014]; this supression on range expansion is consequently expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011].
 Unlike Himalayan songbirds, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
@@
 
-<!-- Despite the strong predictions that sexual selection is more likely to increase speciation than not **cite all those paper**, there are countervailing arguments that sexual selection may dampen speciation rate.
-Studies have suggested that increased selection on males through sexual selection may elevate inter-population mating rates as males seek to overcome local gene flow between populations and impede speciation [@Parker_1998; @Servedio_2014].
-**How?** Sexual selection may also lead to the evolution of traits that constrain further divergence or adaptation (i.e.
-lead to an evolutionary dead end) **CITE**.
--->
-
 \newpage
@@
 The widespread use of sexual dichromatism as a proxy for sexual selection often goes unchallenged as our instinctive Darwinian minds reminisce of peacocks and birds of paradise; my thoughts were such when I began this scientific inquiry.
-Although, sexual dichromatism measurements from colour drawings were always going to be a less than perfect estimate of sexual selection, across 5,831 species I anticipated a signal would permeate through the noise.
-Alas, this was not so and male-biased sexual selection and sexual dichromatism were very different measures (see \autoref{Path_Analysis} in main text).
+Although, sexual dichromatism measurements from illustrations were always going to be an imperfect estimate of sexual selection, across 5,831 species I anticipated a signal would permeate through the noise.
+Alas --- this was not so ---- and male-biased sexual selection and sexual dichromatism had different effects on speciation.
 With this in mind, the practical use of sexual dichromatism in macroevolutionary research is severely weakened.
-This finding is particularly pertinent as efficient proxy measurements are incredibly valuable in comparative analyses: that's because they increase the accessibility, available sample size and transferability of a study, making it more conducive to research synthesis.
-But, as Wallace remarks, the matter is inevitably more complex despite the irresistibility of the Darwinian argument\footnote{Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took more interest in a view where natural selection on predator-prone nesting females generated dullness in females but not males.} [@Kottler_1980; @Caro_2017].
+This finding is pertinent as efficient proxy measurements are invaluable for comparative analyses, proxies of sexual selection increase the accessibility, available sample size and transferability of research, making a study more conducive to research synthesis.
+But, as Wallace remarks, the matter is inevitably more complex, despite the irresistibility of the Darwinian argument\footnote{Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took the position where natural selection on predator-prone nesting females led to females losing their plumage colouration.} [@Kottler_1980; @Caro_2017].
 On the bright side, the multivariate measure of male-biased sexual selection (sexual dimorphism, social polygyny and [lack of] paternal care) is a dataset that would benefit from being expanded across birds.
+Future research may also benefit from expanding this multivariate measure to account for ornamentation, often overlooked by sexual dimorphism and dichromatism measures [*see* Gomes_2016].
 
@@
 However, these hopes soon dissipated, and with them the seeming \emph{raison d'\^{e}tre} of the study.
-Extinction rate is arguably the best measure of population fitness.
+Extinction rate is arguably the best measure of population fitness as it tracks survival over multiple generations.
 And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured accurately over evolutionarily meaningful timescales [*but see* @Martins_2018].
-We included measurements of extinction rate within the paper, however these findings were played down.
-The reason for this movement away from extinction and towards speciation is the controversy surrounding extinction rate measures as well as phylogentic uncertainty.
-Measuring extinction rates within the bird phylogeny is problematic because of the nature of phylogenetic construction: for passerine birds a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction rate [@Beaulieu_2015].
-Furthermore, the very detection of absence of data --- which is what such methods attempt to do --- is a hard sell both conceptually and methodically come peer review: see @Rabosky_Extinction_2016 for a cautionary tale on measuring extinction rate with ``BAMM``.
+We included measurements of extinction rate in this study, however the findings were played-down.
+The movement away from extinction and towards speciation is because of the controversy surrounding extinction rate measures as well as phylogentic uncertainty.
+Measuring extinction rates from the @Jetz_2012 bird phylogeny is problematic because of the nature of phylogenetic construction: a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction rate [@Beaulieu_2015].
+Furthermore, the very detection of absence of data --- which is what such methods attempt to do --- is a hard sell, both conceptually and methodologically come peer review: see @Rabosky_Extinction_2016 for a cautionary tale on measuring extinction rate with ``BAMM``.
 The mechanisms and rates at which speciation and extinction occur are an active area of research.
 In a recent pre-print @Diaz_2018 used ``BAMM`` across 104 time-calibrated molecular phylogenies and found that despite variation in ecology, biogeography and taxonomy, younger clades consistently had higher rates of speciation and extinction.
-This finding of time-dependency in rates of diversification makes comparative analyses between clades challenging and calls into question countless studies [@Diaz_2018].
+This finding of time-dependency in rates of diversification makes comparative analyses between clades challenging and calls into question countless studies.
 Similar to Wallace's view of sexual dichromatism we are left with a view of diversification rate measures where we cannot yet account for the underlying true variation; which is vastly more complex than we can currently measure.
@@
 In total, for 5,831 species we extracted 338,203,831 point measures\footnote{5,831 species x 1000 samples per range x 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.} of environmental variability, but these measures may not account for important variation in biotic components such as nutritional resources or predation.
-Arguably, the methods used here are much better than those traditionally implemented; where the range mid-point is used to obtain a single measure of temperature, precipitation or seasonality or where only one bioclimatic variable is used.
+Arguably, the methods used here are much better than simply using the range mid-point to obtain a measure of temperature, precipitation or seasonality or where only one bioclimatic variable is used.
 I would argue that reducing the multivariate data into several biologically meaningful predictors is a conceptually challenging exercise, but researchers should not feel limited by the amount of raw data that they can access\footnote{This statement applies to species with documented range sizes, which varies across clades but is comprehensive for birds, mammals, amphibians and many marine groups.
@@
 Here I discuss the broader directions that we may follow in order to consolidate our understanding of sexual selection and its effects on fitness and diversification.
-I will draw heavily on the discussions from a Royal Society meeting on _'Sexual selection: patterns in the history of life'_\footnote{Specifically, the following two sections are spurred from dialogue with \href{https://paleobiology.si.edu/staff/individuals/hunt.cfm}{Dr Gene Hunt} and \href{https://www.earth.ox.ac.uk/people/erin-saupe/}{Associate Professor Erin E.
+I will draw heavily on the discussions from a Royal Society meeting on '*Sexual selection: patterns in the history of life*'\footnote{Specifically, the following two sections are spurred from dialogue with \href{https://paleobiology.si.edu/staff/individuals/hunt.cfm}{Dr Gene Hunt} and \href{https://www.earth.ox.ac.uk/people/erin-saupe/}{Associate Professor Erin E.
 Saupe}} in the UK this year.
@@
 From stratographic occurrence data a capture–mark–recapture model was used to determine the rate of extinction for 93 species.
-Their finding challenges the results from experimental evolution studies [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015] and our meta-analysis that find sexual selection mitigates extinction.
+Their finding challenges the results from experimental evolution studies [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015] and our meta-analysis that suggest sexual selection mitigates extinction.
 In future, it would be incredibly valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual dimorphism to conduct such a study.
-In the interim I would be eager to see how ``BAMM`` performs on an ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected from phylogenies.
+In the interim I would be eager to see how ``BAMM`` performs on an ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected from the phylogeny.
 For me, this will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction rates across a wider variety of species.
@@
 However, the use of paleoclimate models in evolution is now a reality; with recent studies incorporating averaged values across multiple paleoclimate models as predictors of speciation [*see* @quintero_2018].
-While the uncertainty surrounding paleoclimates was beyond the scope of the study, future studies should remain vigilant of the everlasting need to incorporate key ecological processes in macroevolutionary research [*reviewed in* @Weber_2017].
-In our study on passerine birds we found that range size is negatively associated with speciation.
+While the uncertainty surrounding paleoclimates was beyond the scope of the study, future studies should remain vigilant of the everlasting need to use key ecological processes in macroevolutionary research [*reviewed in* @Weber_2017].
+In our study on passerine birds we found that range size is negatively associated with speciation, yet range size encapsulates multiple aspects of ecology and evolution so it is unclear what the finding truly means.
 The role of range size in speciation may be further investigated through ancestral geographic range reconstruction [@Quintero_2015], as ancestral ranges may be beneficial in understanding modes of speciation and the causality of range size expansions and contractions [@Losos_2003].
@@
 Across 92 bivalve species that lived during a mid-Pliocene warming period, @Saupe_2015 used environmental niche modelling to assess the impact of ecological processes on extinction.
-Harnessing ecological niche modelling in macroevolutionary studies across extant taxa such as passerine birds is a daunting challenge but offers an Eco-evo synthesis that may be particularly pertinent in questions relating to extinction; especially if the taxonomic group under inquiry is of high conservation value, for which the songbirds certainly are.
+Harnessing ecological niche modelling in macroevolutionary studies across extant taxa such as passerine birds is a daunting challenge but offers an *eco-evo* synthesis that may be pertinent in questions relating to extinction; especially if the taxonomic group under inquiry is of high conservation value, for which the songbirds certainly are.

### 2018-10-16 Manuscript.Rmd
Supp Material Changes

@@
 Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions.
-Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{electronic supplementary material (ESM)}.
+Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Electronic Supplementary Material (ESM)}.
 
@@
 However, this spectrophotometry data is not as extensive as the RGB measures; with only 581 passerine species available for this analysis.
-While there is a correlation between these two measures, there is residual variation ($r = 0.79$; \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S9}).
+While there is a correlation between these two measures, there is residual variation ($r = 0.79$; \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
 
@@
 Therefore, we also use an existing dataset [@Dale_2015] of male-biased sexual selection based on the first component of a phylogenetic principle component analysis (PPCA) of three characteristics associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
-This measure of male-biased sexual selection is available for only 2,465 species and is weakly correlated with the absolute values of sexual dichromatism using RGB measures ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S11}).
+This measure of male-biased sexual selection is available for only 2,465 species and is weakly correlated with the absolute values of sexual dichromatism using RGB measures ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
 
@@
 Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
-Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
+Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
 
@@
 Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-PGLS analyses using subsetted data (*n* = 581), where sexual dichromatism was measured using spectrophotometry [@Armenta_2008] yielded results concordant to the full data set; *i.e.* no association between sexual dichromatism and speciation or extinction (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
+PGLS analyses using subsetted data (*n* = 581), where sexual dichromatism was measured using spectrophotometry [@Armenta_2008] yielded results concordant to the full data set; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
 Our results from MCC models are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 100 random trees.
@@
 Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subsetted for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
-Models using data subsetted for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+Models using data subsetted for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
 Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with highest power using near-complete taxon sampling (*n* = 5,812).
@@
 
-Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; \href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S13}).
+Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; \href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S14}).
 Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from male-biased sexual selection to sexual dichromatism was relatively weak, ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
@@
 
-We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
 
@@
 We used a range of methods to overcome uncertainty in the power and precision of various tip-rate estimates as well as the variation between trees.
-We found that that the tip-rate estimates across the 100 trees were quite variable for both $\lambda_{DR}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S14}).
-Despite this, tip-rate estimates of speciation from the two methods were moderately correlated across the 100 trees ($r = 0.75$) and for the MCC tree ($r = 0.68$; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
+We found that that the tip-rate estimates across the 100 trees were quite variable for both $\lambda_{DR}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
+Despite this, tip-rate estimates of speciation from the two methods were moderately correlated across the 100 trees ($r = 0.75$) and for the MCC tree ($r = 0.68$; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S16}).
 However, in comparison to $\lambda_{DR}$, the value of $\lambda_{BAMM}$ was the _mean_ drawn from a posterior distribution (*n* = 1,000) of BAMM generations, thus estimates of $\lambda_{BAMM}$ (and $\mu_{BAMM}$) have an added level of variation.
 To account for this variation, weights (using the inverse of the variance) were used for the PGLS models.
-From 1,000 posterior samples of the MCC BAMM run the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figures S5} and \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S7}).
+From 1,000 posterior samples of the MCC BAMM run the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figures S6} and \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
 Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) was high.
-This means that although the number of rate shifts reached convergence ($median = 59$; \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6}), the locations of the rate shifts (*i.e.* the credible shift set) across the tree are highly heterogeneous.
+This means that although the number of rate shifts reached convergence ($median = 59$; \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S7}), the locations of the rate shifts (*i.e.* the credible shift set) across the tree are highly heterogeneous.
 The variability in the locations for the shift configurations is a likely source of uncertainty in downstream tip-rates used as the response variable in PGLS models.

### 2018-10-16 Sexual Selection and Speciation.Rmd
Supp Material Changes

@@
 @Dale_2015 compared their RGB scores to a dataset of reflectance measurements of Australian birds.
-The two datasets were correlated but the relationship was expectedly noisy **include R square of correlation**.
+The two datasets were correlated but the relationship was expectedly noisy (*R^2^* = 0.67).
 Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008, which estimated sexual dichromatism as the mean number of just noticeable differences (JNDs) between male and female plumage colour (discriminability), based on a model of bird colour vision.
@@
 
-**Table S1:** Loadings for the first two PCs, from a PCA on the variation (standard error, where n ~ 1,000) of each bioclim variable except the standard error of temperature seasonality (se.bioclim4) and the standard error of precipitation seasonality (se.bioclim15), as these are already measures of variability and based on other bioclimatic variables.
-We use the standard error as, while we attempted to sample each species range 1,000 times, some points returned NA values and would thus affect the standard deviation, but not standard error.** mmm..SD is the sum of the distances of each point to the mean, divided by n-1, so it already takes into account whether some species have more or less points, it's like an average of distances to the mean.
-I think the standard way of measuring spread data is the SD more than the SE...anyway, I don't think that will change much if all Ns are very similar, but let's think about that for the publication**
+**Table S1:** Loadings for the first two PCs, from a PCA on the variation (standard error, where n ~ 1,000) of each bioclim variable except the standard deviation of temperature seasonality (sd.bioclim4) and the standard error of precipitation seasonality (sd.bioclim15), as these are already measures of variability and based on other bioclimatic variables.
 ```{r}
@@
 In the analysis we only used the residuals from PC1.
-**How much % is explained by PC1??**
+PC1 accounts for 48.1 % of the variation.
 
@@
 The plots show the two PCs, broadly representing temperature and precipitation.
-Data is also provided for the last-glacial maximum (LGM) but was not used in the analysis.
-**last sentence not needed, or otherwise explain why we chose LIG over LGM?**
+We used the difference between the present and LIG climates as these represent a longer (evolutionarily meaningful) time-scale than the difference between the LGM and present climates.
 
@@
 ```
-**Figure S6:** **this changes to S6 and next to S7 etc**The tip-rate estimates for BAMM are highly variable within each run of BAMM.
+**Figure S6:** The tip-rate estimates for BAMM are highly variable within each run of BAMM.
 Across most species there is high variability in the posterior distribution of tip-rate estimates.
@@
 ```
-**Figure S8:** This figure is the same basic figure as seen in the manuscript (Figure 1). It provides model estimates for four response variables across 100 random trees alongside the MCC tree.
+**Figure S9:** This figure is the same basic figure as seen in the manuscript (Figure 1). It provides model estimates for four response variables across 100 random trees alongside the MCC tree.
 
@@
 
-**Figure S9:** There is a correlation between the RGB measures and the Spectophotometry measures. The RGB measures seem to be more noisy around the lower values.
+**Figure S10:** There is a correlation between the RGB measures and the Spectophotometry measures. The RGB measures seem to be more noisy around the lower values.
 
@@
 ```
-**Figure S10:** Measures of sexual dichromatism using spectrophotometry do not change the main patterns.
+**Figure S11:** Measures of sexual dichromatism using spectrophotometry do not change the main patterns.
 This dataset is a subset of the complete dataset (n = 581), thus drawing conclusions for the other predictors (*e.g.* borderline long term temperature variation and spatial temperature variation) potentially risks type I error.
@@
 ```
-**Figure S11:** The relationship between absolute sexual dichromatism and male-baised sexual selection.
+**Figure S12:** The relationship between absolute sexual dichromatism and male-baised sexual selection.
 Although positively correlated, the distribution is very noisy with some species having high sexual dichromatism and low male-biased sexual selection and vice versa.
@@
 ```
-**Figure S12:** PGLS analyses using measures of male-biased sexual selection (*n* = 2,465) using three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
+**Figure S13:** PGLS analyses using measures of male-biased sexual selection (*n* = 2,465) using three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
 The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 100 random trees can be found in the ESM.
@@
 This figure is part **b** of Figure 1 within the manuscript.
-n
 
@@
 
-**Figure S13:** Path analysis  standardised regression coefficients vary across relationships.
+**Figure S14:** Path analysis  standardised regression coefficients vary across relationships.
 Error bars are derived from confidence intervals through 500 bootstrapped iterations.
@@
 ```
-**Figure S14:** Speciation Rate means from 100 trees using either the DR statistic (x axis) or BAMM (y axis).
+**Figure S15:** Speciation Rate means from 100 trees using either the DR statistic (x axis) or BAMM (y axis).
 While there is a clear correlation (*r* = `r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`) there is variability with BAMM showing less heterogeneity.
@@
 ```
-**Figure S15:** Similar to Figure S14, there is a correlation (*r* = `r round(cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda), digits = 2)`) between $\lambda_{DR}$ and $\lambda_{BAMM}$ with BAMM results showing less heterogeneity.
+**Figure S16:** Similar to Figure S14, there is a correlation (*r* = `r round(cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda), digits = 2)`) between $\lambda_{DR}$ and $\lambda_{BAMM}$ with BAMM results showing less heterogeneity.
 Each point represents a species (*n* = 5,965).

### 2018-10-17 Manuscript.Rmd
Updates from Devi

@@
 ---
-title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds
+title: \huge{Male-biased sexual selection, but not sexual dichromatism, predicts speciation
+  in birds}
 author:
-- Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§} and Iliana Medina Guzmán\textsuperscript{§}
+- Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§}
+  and Iliana Medina Guzmán\textsuperscript{§}
 - ~
@@
 output:
-  word_document:
-    pandoc_args: --smart
-  html_document: default
   pdf_document: default
@@
   \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape}
-  \usepackage{hyperref}\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, urlcolor=blue}
-  \raggedbottom
+  \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{}
+  \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{}
+  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage{hyperref}\hypersetup{colorlinks=true,
+  linkcolor=blue, filecolor=blue, urlcolor=blue} \raggedbottom
 csl: References/evolution.csl
-abstract: 'Many theories related to sexual selection predict that it shapes phylogenetic diversity by altering speciation or extinction rates.
-The predicted effects on diversity are often contradictory; for example, sexual selection might promote speciation by creating reproductive isolation, or hinder it by selecting for traits that facilitate gene flow.
-Similarly, sexual selection might precipitate extinction by favouring competitive phenotypes with poor survival, or stave it off by helping populations shed deleterious mutations.
-Here, we investigate the relationship between sexual selection and rates of speciation and extinction in passerine birds.
-We use two measures of sexual selection -- sexual dichromatism (*n* = 5,812 species) and a multivariate measure of male-biased sexual selection (sexual dimorphism, polygyny and paternal care; *n* = 2,465) -- as well as three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$), and one measure of extinction ($\mu_{BAMM}$).
-We also tested whether the effects of sexual selection is contingent on ecological variables such as spatiotemporal temperature variation, primary productivity, and range size.
-We find that male-biased sexual selection but not sexual dichromatism predicts speciation.
-Additionally, we find that birds with smaller ranges have higher speciation rates; although the direction of causality is unclear.
-\vspace{5mm} \par\noindent \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage dichromatism, Range size, Extinction rate.'
+subparagraph: yes
+abstract: Sexual selection can shape phylogenetic diversity by altering speciation
+  or extinction rates. The predicted effects on diversity are often contradictory;
+  for example, sexual selection might promote speciation by creating reproductive
+  isolation, or hinder it by selecting for traits that facilitate gene flow. Similarly,
+  sexual selection might precipitate extinction by favouring competitive phenotypes
+  with poor survival, or stave it off by helping populations shed deleterious mutations.
+  Here, we investigate the relationship between sexual selection and rates of speciation
+  and extinction in passerine birds. We use two measures of sexual selection --- sexual
+  dichromatism (*n* = 5,812 species) and a multivariate measure of male-biased sexual
+  selection (sexual dimorphism, polygyny and paternal care; *n* = 2,465) --- as well
+  as three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$),
+  and one measure of extinction ($\mu_{BAMM}$). We also tested whether the effects
+  of sexual selection is contingent on ecological variables such as spatiotemporal
+  temperature variation, primary productivity, and range size. We find that birds
+  with smaller ranges have higher speciation rates; although the direction of causality
+  is unclear. Moreover, our results show that male-biased sexual selection, but not
+  sexual dichromatism, is an agent of adaptation and speciation. This finding suggests
+  that sexual selection is an evolutionary force shaping patterns of diversity amongst
+  songbirds, which has broad implications for ecology and conservation biology. \vspace{5mm}
+  dichromatism, Range size, Extinction rate.
+  \par\noindent \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage
 ---
@@
 
-\newpage
+\newpage\phantomsection
 
@@
 As populations diverge they accumulate genetic and phenotypic changes that ultimately erect barriers against gene flow [@Presgraves_2010; *but see* @Shaw_2011].
-In this view, populations will speciate and adapt more readily to new and different environments if selection is strong and/or persistant.
-Hence, the rate at which speciation occurs is likely dependent on both the population's response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.*, morphology or colour @Hugall_2012; @Rabosky_morph_2013].
+In this view, populations will speciate and adapt more readily to new and different environments if selection is strong and/or persistent.
+Hence, the rate at which speciation occurs is likely dependent on both the population's response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.*, morphology or colour: @Hugall_2012; @Rabosky_morph_2013].
 Experiments suggest that sexual selection can influence the evolution of a surprisingly diverse set of traits [@Cally_in_review], and may therefore strongly influence extant patterns of phylogenetic diversity.
@@
 This evolutionary force has been proposed to facilitate speciation through the evolution of divergent signals associated with improved mating success [@Lande_1981; @Lande_1982].
-For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism, are more species rich [@Seddon_2008].
+For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism are more species rich [@Seddon_2008].
 Under this model of speciation, divergent mating signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].
@@
 For example, sexual selection can elevate average fitness by helping to purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial ones [@Whitlock_2000], which should mitigate extinction risk and facilitate adaptation to environmental change.
-However, sexual selection might also promote extinction, by selecting for male traits that have deterimental pleiotropic effects on female traits [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006], or selecting for male phenotypes with improved mating success but which harm population productivity, such as under-investing in parental care [@RN277].
+However, sexual selection might also promote extinction, by selecting for male traits that have detrimental pleiotropic effects on female traits [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006], or selecting for male phenotypes with improved mating success but which harm population productivity, such as under-investing in parental care [@RN277].
 
@@
 Stressful environments strengthen the positive correlation between male mating success and female fitness, reducing the burden of negative pleiotropic effects [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017].
-Conversely, benign or stable environments ensure consistent selection, preferentially eroding genetic variation at sexually concordant loci (*i.e.* loci where the optimum genotype is equivelant for both sexes).
+Conversely, benign or stable environments ensure consistent selection, preferentially eroding genetic variation at sexually concordant loci (*i.e.* loci where the optimum genotype is equivalent for both sexes).
 In these stable environments, genetic variation remains disproportionately at sexually antagonistic loci; causing displacement of male and female traits from their optimum.
 It follows from these predictions that populations with stronger sexual selection would have fitness benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
-Additionally, sexual selection can reinforce locally adaptated phenotypes through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011].
+Additionally, sexual selection can reinforce locally adapted phenotypes through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011].
 For instance, beak morphology is a trait that arose under natural selection in several taxa and is now a trait under sexual selection [a.k.a.
@@
 In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-Through these synergenetic effects sexual selection and natural selection --- such as environmental change over space and time --- may lead to increased adaptation and speciation rates.
+Through these synergetic effects sexual selection and natural selection --- such as environmental change over space and time --- may lead to increased adaptation and speciation rates.
 
-As a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumge, birds are a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g*, @Seddon_2008; @Huang_2014; @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003].
-A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, though the average effect size was weaker in birds than in insects and fish but stronger than in mammals [@Kraaijeveld_2011].
-Birds are often thought to be more conducive in benefiting from sexual selection due to their sex determination system.
-In male homogametic species (ZZ/ZW) --- such as birds --- males are predicted to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003].
-As such, we might expect bird populations experiencing stronger sexual selection to have increased adaptation and speciation rates compared to male heterogametic taxa.
+As a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumage, birds are a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003].
+A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects and fish [@Kraaijeveld_2011].
 However, @Kraaijeveld_2011 found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of speciation and the strength of sexual selection.
@@
 Firstly, many macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual dimorphism or dichromatism, but these are imperfect correlates.
-We used a very large (*n* = 5,812) dataset of sexual dichromatism measures [sourced from illustrated drawings; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection, capturing variation in sexual dimorphism, social polygyny and [lack of] paternal care.
+We used a very large (*n* = 5,812) dataset of sexual dichromatism [estimated from illustrated drawings; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection, capturing variation in sexual dimorphism, social polygyny and [lack of] paternal care.
 These datasets allow us to cover the great majority of bird genera, in contrast to previous work that has often focused on particular clades [*e.g.* @Seddon_2008].
-Secondly, we use the latest methods for estimating speciation and exinction rates at the tips of phylogenetic trees, including BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but less controversial tip-rate statistics, commonly reffered to as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
-We also conduct analyses that allow phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] to affect the precision of our estimates, rather than implicitly assuming that the single best-supported phylogenetic tree is correct.
+Secondly, we use the latest methods for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but less controversial tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
+We also conduct analyses that allow phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] to affect the precision of our estimates, rather than implicitly assuming that the single best-supported phylogenetic tree, or consensus tree, is correct.
 Thirdly, our analysis includes several potentially important environmental variables as predictors: range size, seasonal variation in temperature, spatial temperature variation (across a species’ range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
@@
 We also test whether the effects of sexual selection are contingent on the environment, as predicted by some theories [*e.g.* the theory that sexual selection is more likely to elevate adaptation in changeable relative to static environments; @Long_2012; @Connallon_Hall_2016].
+\phantomsection
 
@@
 Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Electronic Supplementary Material (ESM)}.
+\phantomsection
 
 ##Compiling data for sexual selection and environmental stress
+\phantomsection
 
@@
 
-We used a previously published measure of sexual selection for 5,983 species of passerines [@Dale_2015].
+We used a previously published measure of sexual dichromatism for 5,983 species of passerines [@Dale_2015].
 Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forhead, throat, upper breast and lower breast) from _Handbook of the Birds of the World_ (volumes 8–16) [@delhoyo2011].
@@
 Here we use the absolute difference between male and female plumage scores as an estimate of sexual dichromatism.
-Additionally, we used a measure of dichromatism from spectral data, used to estimate colour distance in avian colour space [@Armenta_2008].
+Additionally, we used a more accurate measure of dichromatism corresponding to colour distance in avian colour space derived from spectral data [@Armenta_2008].
 These measurements include variation in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from taxonomic samples (as opposed to colour drawings).
@@
 While there is a correlation between these two measures, there is residual variation ($r = 0.79$; \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
+\phantomsection
 
@@
 This measure of male-biased sexual selection is available for only 2,465 species and is weakly correlated with the absolute values of sexual dichromatism using RGB measures ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
+\phantomsection
 
@@
 We obtained estimates of species range size using expert range maps [@birdlife_2017].
-Because of taxanomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
+Because of taxonomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
 From these distributions we obtained estimates of climatic conditions that each species is exposed to.
@@
 From these values we obtained means and standard deviations for each species.
-Using the same spatial sampling we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006].
-Furthermore, to estimate variability in the energy available to species we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution.
+Using the same spatial sampling, we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006].
+Furthermore, to estimate variability in the energy available to species, we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution.
 Estimates of NPP had 30 second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005].
-Using this data, which we provide as a potentially useful data resource (see ESM), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
+Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
+\phantomsection
 
-###Generating biologically relevent predictors for environmental variation
+###Generating biologically relevant predictors for environmental variation
 
@@
 Details and code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.
+\phantomsection
 
-##Estimating extinction and speciation.
+##Estimating extinction and speciation
 
@@
 These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE).
-But, based on previous simulation studies $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
+But, based on previous simulation studies, $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
 
@@
 The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}; briefly we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
-Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPU's.
+Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
 Each run of BAMM reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200 (an arbitrary value, above which posterior distributions can often be accurately inferred) for the key model parameters of log-likelihood and number of rate shifts (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
 Further details of BAMM parameters and output are available in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}, with tip-rate means and variances provided.
-Additionally, given the variability in BAMM estimates, we also provide some analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on the MCC tree from a genetic-only phylogeny across all birds [@Harvey_2017].
+Additionally, given the variability in BAMM estimates, we also provide analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on the MCC tree from a genetic-only phylogeny across all birds [@Harvey_2017].
 All analyses were conducted on log-rates.
+\phantomsection
 
-##Statistical analysis
-
-###Phylogenetic camparative analysis
+##Phylogenetic comparative analysis
 
@@
 
-Finally, using data subsetted for species with measurements of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
+Finally, using data subset for species with measurements of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
 The phylogenetic path analysis was used to assess causal paths *between* variables not able to be modelled within the univariate response of PGLS.
@@
 The phylogenetic path analysis used tip-rates and correlation structure from the MCC tree.
-Further details of the path analysis including reasons for path directions can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
+Further details of the path analysis, including reasons for path directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
 
@@
 ```
+\phantomsection
+
 #Results
+\phantomsection
 
@@
 Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
+\phantomsection
 
@@
 Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-PGLS analyses using subsetted data (*n* = 581), where sexual dichromatism was measured using spectrophotometry [@Armenta_2008] yielded results concordant to the full data set; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
-Our results from MCC models are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 100 random trees.
+PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
+Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 100 random trees.
 The HPD (highest posterior density) intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
-For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (`r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+\phantomsection
 
@@
 We found a significant positive association between male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-However, this associate was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-The distribution of estimates from PGLS models on 100 random trees was similar to the estimate from an MCC tree: among the 100 tress there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
+However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+The distribution of estimates from PGLS models on 100 random trees was similar to the estimate from the MCC tree: among the 100 tress there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
 Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}.
+\phantomsection
 
@@
 
-Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation we found a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$.
+Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation, we found a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$.
 This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}).
 This association was also evident across the estimates from models using the 100 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a).
-Subsetted models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed variable evidence that range size is negatively associated with speciation.
-Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subsetted for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
+Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed variable evidence that range size is negatively associated with speciation.
+Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subset for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
 Models using data subsetted for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
 Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with highest power using near-complete taxon sampling (*n* = 5,812).
+\phantomsection
 
@@
 Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from male-biased sexual selection to sexual dichromatism was relatively weak, ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
-Additionally, the large direct effect of temperature seasonality on range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`) suggests an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}), given the negative we identified between $\lambda_{DR}$ and range size in PGLS models.
+Additionally, the large direct effect of temperature seasonality on range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`) suggests an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta_{indirect}$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}), given the negative association we identified between $\lambda_{DR}$ and range size in PGLS models.
+\phantomsection
 
@@
 We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+\phantomsection
 
@@
 To account for this variation, weights (using the inverse of the variance) were used for the PGLS models.
-From 1,000 posterior samples of the MCC BAMM run the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figures S6} and \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
+From 1,000 posterior samples of the MCC BAMM run, the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figures S6} and \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
 Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) was high.
@@
 This suggests that species with smaller range sizes have increased speciation rate, however, small range size may be a cause or effect of elevated speciation rate and potentially even an artifact of taxonomic classification.
-Interestingly we found a discrepancy in the model results obtained from tip-rate statistics ($\lambda_{DR}$, $\lambda_{ND}$) and those using tip-rate estimates from a model-based approach ($\lambda_{BAMM}$); whereby BAMM produced precisely null results for most associations between predictors and speciation as well as extinction.
+Interestingly, we found a discrepancy in the model results obtained from tip-rate statistics ($\lambda_{DR}$, $\lambda_{ND}$) and those using tip-rate estimates from a model-based approach ($\lambda_{BAMM}$), whereby BAMM produced precisely null results for most associations between predictors and speciation as well as extinction.
 
@@
 Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015] we find that sexual dichromatism does not predict speciation rate, whereas male-biased sexual selection does.
-Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection, and that variance in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
+Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection, and that variation in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
 There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
@@
 Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
-And lastly, not all plumage colouration honestly reflects mate quality.
+Lastly, not all plumage colouration honestly reflects mate quality.
 A recent meta-analysis found that converted carotenoids (but *not* those sourced from diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
-Given the importance of honest signaling in sexual selection and the variability in colour production across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
+Given the importance of honest signaling in sexual selection and the variability in colour production mechanisms across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
 
-In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ estimates --- but this association was not dependent on variable or stressful environments.
-When evolving under stressful conditions, sexually antagonistic selection is hypothesised to be weakened relative to selection that is concordant between the sexes [@Connallon_SA_2015; @Connallon_Hall_2016].
-While these predictions are supported in several lab (experimental evolution) studies [@Long_2012; @Berger_2014; @Punzalan_2013], other studies have not found the association [@Delcourt_2009; @Martinossi-Allibert_Savković_2017; @Holman_Jacomb_2017].
+In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ estimates --- but this association was not dependent on environmental variability.
 The environmental variability predictors used here often reflected changes in climate across space and time, whereby high environmental variation would likely increase the strength of natural selection on a given species.
@@
 Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability the species is evolving under.
-Specifically, sexual antagonism may be more easily purged in environments where cyclic pressures (*e.g.* seasonality) are perpetuated; whereas when environmental variability is directional (*e.g.* long-term climatic change) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
+Specifically, sexual antagonism may be more easily purged in environments where pressures are cyclic (*e.g.* seasonality); whereas when environmental variability is directional (*e.g.* long-term climate change) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
 Secondly, the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
-Specifically, access to dietary resources, and the impacts of predation or parasitism are unaccounted for here, these are likely key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
+Specifically, access to dietary resources, and the impacts of predation or parasitism are unaccounted for here.
+These are likely key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
 Thirdly, it is also possible that there is no effect of environmental variability on speciation rates; at least not in the species investigated here.
@@
 Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995].
-However, birds with smaller range sizes are predicted to have lower dispersal ability.
-This ensues that individuals might not be able to overcome geographical barriers and maintain gene flow; which would promote reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
-Several studies have used morphological measures of wing shape (Hand-wing index/ Kipp's distance) --- a proxy for dispersal ability --- as predictors of speciation and coexistence [*e.g.*, @Kennedy_2016; @Pigot_2016].
-But currently, published datasets of these measurements are not available for many passerines.
+However, birds with smaller range sizes are predicted to have lower dispersal ability, which would promote reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
 Alternatively, reduced range size may be a consequence of speciation.
-Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives.
-Within Himalayan songbirds, range expansions are prevented by the competition for resources by related species [@Price_2014]; this supression on range expansion is consequently expected to prevent future speciation events [@Rosenzweig_1995; @Weir_2011].
-Unlike Himalayan songbirds, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
+Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives [@Rosenzweig_1995; @Weir_2011; @Price_2014].
+However, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
 Across islands we expect to see a correlation between speciation and small range size because they reflect newly formed environments with empty niches.
-Thus, our findings may be heavily dependent on species whose range size is restricted to a single island within an archipelago (*e.g.* *Geospiza* sp.; a.k.a.
+Thus, our findings may be heavily dependent on island radiations (*e.g.* *Geospiza* sp.; a.k.a.
 Darwin's finches).
 One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
-Given the problems of determining causality and mode of speciation the reason for the negative association found here remains obscured.
+Given the problems of determining causality and mode of speciation, the reason for the negative association found here remains unclear.
 
@@
 Extinction is notoriously hard to estimate accurately [*see* @Rabosky_Extinction_2016].
-It can be estimated from phylogenetic inference, fossil record, records of recent extinctions and IUCN extinction threat status.
+It can be estimated from phylogenies, the fossil record, documented recent extinctions and IUCN extinction threat status, but each approach has limitations.
 Phylogenetic methods --- like BAMM --- allow for speciation and extinction rates to be measured using moderately sized phylogenies.
-However, BAMM has been subject to controversy in its ability to model rate shifts and extinction rates [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], and while several tip-rate estimates exist for speciation (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$) extinction estimates are not as easily obtained.
-Here, we found that BAMM often produced homogeneous speciation and extinction rates for clades with few rate shifts; arguably this could reduce our power to detect small differences in extinction rates among closely related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
-The ability to detect extinction rates rarely improves when fossil records are used, as many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
-But, there are exceptions: ostracods --- a group of marine invertebrates --- have extensive fossil records and show identifiable sexual dimorphism.
-Across 93 species of ostracods, @Martins_2018 found that ostracod taxa with larger sexual dimorphism (representing exaggerated investment by males in reproductive organs) had higher extinction rates; implying that sexual selection increases extinction risk.
-Alternatively, when comparative studies have assessed the relationship between sexual selection and extinction in extant birds [@Doherty_2003; @Mitra_1996; @Morrow_2003] and mammals [@Bro-Jorgensen_2014; @Morrow_2004] there has been no consensus.
-Unfortunately, these methods often rely on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List [*e.g.*, @Morrow_2003; @Bro-Jorgensen_2014; @Morrow_2004] and may be biased to species under greater anthropogenic threat, culturally valued or well researched.
+However, BAMM has been the subject of controversy over its ability to model evolutionary rate shifts and extinction rates [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], and while several tip-rate estimates exist for speciation (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$) extinction estimates are not as easily obtained.
+Here, we found that BAMM often produced homogeneous speciation and extinction rates for clades with few rate shifts which could arguably reduce our power to detect small differences in extinction rates among closely related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
 
-Improvements in estimating tip-rates of speciation has allowed us to revisit the association between sexual selection and speciation.
-With high power and alongside extensive ecological predictors we find support that male-biased sexual selection predicts speciation.
-If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold with sexual selection a potentially valuable utility in conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].
-Furthermore, our finding that high speciation rate is associated with smaller range size has pronounced significance to the future survival of these rapidly speciating species in a world with increased habitat loss and anthropogenic stress.
-Range size is the best predictor of extinction risk [@Harris_2008] and the association found here ensues that many newly speciated clades may be at greater risk of extinction due to their reduced range being impacted by anthropogenic effects.
+We have shown that in passerines, male-biased sexual selection, but not sexual dichromatism, predicts speciation, independent of several measures of environmental variability.
+If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold, including for conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].
+Furthermore, our finding that high speciation rate is associated with smaller range size highlights the threat to the persistence of rapidly speciating lineages in a world with increased habitat loss and anthropogenic stress.
+Range size is the best predictor of extinction risk [@Harris_2008] and the association found here implies that many newly speciated clades may be at greater risk of extinction due to their small range being impacted by anthropogenic effects.
 
-\newpage
+\newpage\phantomsection
+
 # References
@@
 \newpage
-\nolinenumbers
+\nolinenumbers\phantomsection
 
@@
 But it by no means follows that slight differences in the shape, pattern, or colours of the ornamental plumes are what lead a female to give the preference to one male over another; still less that all the females of a species, or the great majority of them, over a wide area of country, and for many successive generations, prefer exactly the same modification of the colour or ornament.\\
---- \large\textsc{Alfred Russel Wallace}}
+--- \large\textsc{Alfred Russel Wallace}}\phantomsection
 
@@
 With this in mind, the practical use of sexual dichromatism in macroevolutionary research is severely weakened.
-This finding is pertinent as efficient proxy measurements are invaluable for comparative analyses, proxies of sexual selection increase the accessibility, available sample size and transferability of research, making a study more conducive to research synthesis.
+This finding is pertinent, as efficient proxy measurements are invaluable for comparative analyses.
+Proxies of sexual selection increase the accessibility, available sample size and transferability of research, making a study more conducive to research synthesis.
 But, as Wallace remarks, the matter is inevitably more complex, despite the irresistibility of the Darwinian argument\footnote{Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took the position where natural selection on predator-prone nesting females led to females losing their plumage colouration.} [@Kottler_1980; @Caro_2017].
@@
 At very least, if bioclimatic variables do not encapsulate key ecological parameters (*e.g.*, resources, predation, habitat structure) these predictors may account for some of the noise in the dataset and may help isolate the effect of the key moderator variables of interest.
+\phantomsection
 
-##Future Directions
+##Future directions
 
@@
 In the interim I would be eager to see how ``BAMM`` performs on an ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected from the phylogeny.
-For me, this will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction rates across a wider variety of species.
+This will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction rates across a wider variety of species.
 
@@
 Harnessing ecological niche modelling in macroevolutionary studies across extant taxa such as passerine birds is a daunting challenge but offers an *eco-evo* synthesis that may be pertinent in questions relating to extinction; especially if the taxonomic group under inquiry is of high conservation value, for which the songbirds certainly are.
+
+<!-- The ability to detect extinction rates rarely improves when fossil records are used, as many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
+But, there are exceptions: ostracods --- a group of marine invertebrates --- have extensive fossil records and show identifiable sexual dimorphism.
+Across 93 species of ostracods, @Martins_2018 found that ostracod taxa with larger sexual dimorphism (representing exaggerated investment by males in reproductive organs) had higher extinction rates; implying that sexual selection increases extinction risk.
+Alternatively, when comparative studies have assessed the relationship between sexual selection and extinction in extant birds [@Doherty_2003; @Mitra_1996; @Morrow_2003] and mammals [@Bro-Jorgensen_2014; @Morrow_2004] there has been no consensus.
+Unfortunately, these methods often rely on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List [*e.g.*, @Morrow_2003; @Bro-Jorgensen_2014; @Morrow_2004] and may be biased to species under greater anthropogenic threat, culturally valued or well researched.
+-->
+
+<!-- Note that Devi changed this para anyway -->

### 2018-10-17 Sexual Selection and Speciation.Rmd
Updates from Devi

@@
 
-**2)** Extracted the bioclim or other (e.g. NPP) data from each of the points:
+**2)** Extracted the bioclim or other (*e.g.* NPP) data from each of the points:
 ```{r, eval = FALSE}
@@
 
-**3)** Summarised the extracted data across the 1,000 points into a summary value of interest (means and se) for each species and exported that data.
+**3)** Summarised the extracted data across the 1,000 points into a summary value of interest (means and sd) for each species and exported that data.
 
@@
 bird.summary <- lapply(bird.frame, function(x) {
-(as.data.frame(apply(x,2,mean, na.rm =T )))})
+(as.data.frame(apply(x,2,mean, na.rm =T)))})
 
@@
 Error bars are derived from confidence intervals through 500 bootstrapped iterations.
-Paths with error bars not overlapping zero are presented with an astericks in Figure 3 within the manuscript.
+Paths with error bars not overlapping zero are presented with an asterisks in Figure 3 within the manuscript.
+
+_______________
 

### 2018-10-18 Manuscript.Rmd
SE to SD Change

@@
   pdf_document: default
+  word_document: default
 header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
@@
 subparagraph: yes
-abstract: Sexual selection can shape phylogenetic diversity by altering speciation
+abstract: 'Sexual selection can shape phylogenetic diversity by altering speciation
   or extinction rates. The predicted effects on diversity are often contradictory;
@@
   that sexual selection is an evolutionary force shaping patterns of diversity amongst
-  songbirds, which has broad implications for ecology and conservation biology. \vspace{5mm}
-  dichromatism, Range size, Extinction rate.
-  \par\noindent \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage
+  songbirds, which has broad implications for ecology and conservation biology.
+\vspace{5mm}  \par\noindent \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage dichromatism, Range size, Extinction rate.'
 ---
@@
 Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
+For instance, sexual selection may promote investments in other mating signals (*e.g.* bird song) that trade-off against plumage colour, leading to a negative correlation between different sexual selection proxies [@Cooney_2018].
 Lastly, not all plumage colouration honestly reflects mate quality.
@@
 
-Perhaps one of the biggest unravelings of this project was the decoupling of male-biased sexual selection from sexual dichromatism, remarkably this same controversy engrossed Wallace and Darwin 130 years ago [@Kottler_1980].
+One of the biggest unravelings of this project was the decoupling of male-biased sexual selection from sexual dichromatism, remarkably this same controversy engrossed Wallace and Darwin 130 years ago [@Kottler_1980].
 The widespread use of sexual dichromatism as a proxy for sexual selection often goes unchallenged as our instinctive Darwinian minds reminisce of peacocks and birds of paradise; my thoughts were such when I began this scientific inquiry.
@@
 
-The study by @Martins_2018 has already been referred to in the main text of both papers and for good reason.
-This elegant study was able to do what no one else has; reliably measure extinction rate over meaningful time scales.
-Using fossil ostracods with various levels of sexual dimorphism, @Martins_2018 found that species with high levels of sexual dimorphism were quickly swept to extinction.
+Estimating extinction risk from the fossil record is equally fraught because many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
+Using fossil ostracods (a group of marine invertebrates), @Martins_2018 found that species with high levels of sexual dimorphism were quickly swept to extinction.
 From stratographic occurrence data a capture–mark–recapture model was used to determine the rate of extinction for 93 species.
-Their finding challenges the results from experimental evolution studies [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015] and our meta-analysis that suggest sexual selection mitigates extinction.
+Their finding challenges the results from experimental evolution studies that suggest sexual selection mitigates extinction [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015].
+By contrast, evidence is mixed regarding the relationship between sexual selection and extinction in extant birds [@Doherty_2003; @Mitra_1996; @Morrow_2003] and mammals [@Bro-Jorgensen_2014; @Morrow_2004].
+Unfortunately, these studies have relied on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List and may be biased to species under greater anthropogenic threat, culturally valued or well researched.
 In future, it would be incredibly valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual dimorphism to conduct such a study.
@@
 Harnessing ecological niche modelling in macroevolutionary studies across extant taxa such as passerine birds is a daunting challenge but offers an *eco-evo* synthesis that may be pertinent in questions relating to extinction; especially if the taxonomic group under inquiry is of high conservation value, for which the songbirds certainly are.
-
-<!-- The ability to detect extinction rates rarely improves when fossil records are used, as many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
-But, there are exceptions: ostracods --- a group of marine invertebrates --- have extensive fossil records and show identifiable sexual dimorphism.
-Across 93 species of ostracods, @Martins_2018 found that ostracod taxa with larger sexual dimorphism (representing exaggerated investment by males in reproductive organs) had higher extinction rates; implying that sexual selection increases extinction risk.
-Alternatively, when comparative studies have assessed the relationship between sexual selection and extinction in extant birds [@Doherty_2003; @Mitra_1996; @Morrow_2003] and mammals [@Bro-Jorgensen_2014; @Morrow_2004] there has been no consensus.
-Unfortunately, these methods often rely on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List [*e.g.*, @Morrow_2003; @Bro-Jorgensen_2014; @Morrow_2004] and may be biased to species under greater anthropogenic threat, culturally valued or well researched.
--->
-
-<!-- Note that Devi changed this para anyway -->

### 2018-10-18 Sexual Selection and Speciation.Rmd
SE to SD Change

@@
 restricted.data <- complete.dataframe %>% drop_na(bioclim1) #Drop species without environmental data
-PCA.bioclim <- prcomp(restricted.data[c('se.bioclim1', 'se.bioclim2', 'se.bioclim3', 'se.bioclim5', 'se.bioclim6', 'se.bioclim7', 'se.bioclim8', 'se.bioclim9', 'se.bioclim10', 'se.bioclim11', 'se.bioclim12', 'se.bioclim13', 'se.bioclim14', 'se.bioclim16', 'se.bioclim17', 'se.bioclim18', 'se.bioclim19')], #Select se.bioclim
+PCA.bioclim <- prcomp(restricted.data[c('sd.bioclim1', 'sd.bioclim2', 'sd.bioclim3', 'sd.bioclim5', 'sd.bioclim6', 'sd.bioclim7', 'sd.bioclim8', 'sd.bioclim9', 'sd.bioclim10', 'sd.bioclim11', 'sd.bioclim12', 'sd.bioclim13', 'sd.bioclim14', 'sd.bioclim16', 'sd.bioclim17', 'sd.bioclim18', 'sd.bioclim19')], #Select sd.bioclim
                       scale = TRUE, center = TRUE)
@@
 #Obtain DR (log(es)) estimates, by calling the calculate.dr function
-passerine.trees <- read.nexus('bamm_extinction/data/trees/Passerine_Trees_Full.nex')
+passerine.trees <- read.nexus('data/trees/Passerine_Trees_Full.nex')
 

### 2018-10-20 Manuscript.Rmd
ReadMe Changes

@@
   \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{}
-  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage{hyperref}\hypersetup{colorlinks=true,
-  linkcolor=blue, filecolor=blue, urlcolor=blue} \raggedbottom
+  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} 
+  \usepackage[hang, flushmargin]{footmisc} 
+  \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false, linkcolor=blue, filecolor=blue, urlcolor=blue}
+  \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
+  \raggedbottom
 csl: References/evolution.csl
@@
   temperature variation, primary productivity, and range size. We find that birds
-  with smaller ranges have higher speciation rates; although the direction of causality
+  with smaller ranges have higher speciation rates, although the direction of causality
   is unclear. Moreover, our results show that male-biased sexual selection, but not
@@
 For example, sexual selection can elevate average fitness by helping to purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial ones [@Whitlock_2000], which should mitigate extinction risk and facilitate adaptation to environmental change.
-However, sexual selection might also promote extinction, by selecting for male traits that have detrimental pleiotropic effects on female traits [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006], or selecting for male phenotypes with improved mating success but which harm population productivity, such as under-investing in parental care [@RN277].
+However, sexual selection might also promote extinction by selecting for male traits that have detrimental pleiotropic effects on female traits [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006], or selecting for male phenotypes with improved mating success but which harm population productivity, such as under-investing in parental care [@RN277].
 
@@
 It follows from these predictions that populations with stronger sexual selection would have fitness benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
-Additionally, sexual selection can reinforce locally adapted phenotypes through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011].
+Additionally, sexual selection can reinforce local adaptations through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011].
 For instance, beak morphology is a trait that arose under natural selection in several taxa and is now a trait under sexual selection [a.k.a.
@@
 
-As a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumage, birds are a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003].
+As a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumage, birds are a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g.*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003].
 A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects and fish [@Kraaijeveld_2011].
@@
 Firstly, many macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual dimorphism or dichromatism, but these are imperfect correlates.
-We used a very large (*n* = 5,812) dataset of sexual dichromatism [estimated from illustrated drawings; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection, capturing variation in sexual dimorphism, social polygyny and [lack of] paternal care.
+We used a very large (*n* = 5,812) dataset of sexual dichromatism [estimated from illustrated drawings; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection, capturing variation in sexual dimorphism, social polygyny and [lack of] paternal care [@Dale_2015].
 These datasets allow us to cover the great majority of bird genera, in contrast to previous work that has often focused on particular clades [*e.g.* @Seddon_2008].
-Secondly, we use the latest methods for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but less controversial tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
+Secondly, we use the latest methods for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but less controversial tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
 We also conduct analyses that allow phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] to affect the precision of our estimates, rather than implicitly assuming that the single best-supported phylogenetic tree, or consensus tree, is correct.
@@
 We used a previously published measure of sexual dichromatism for 5,983 species of passerines [@Dale_2015].
-Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forhead, throat, upper breast and lower breast) from _Handbook of the Birds of the World_ (volumes 8–16) [@delhoyo2011].
+Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forhead, throat, upper breast and lower breast) from _Handbook of the Birds of the World_ [@delhoyo2011].
 The relative contribution of male and female RGB colour values were averaged across body patches and provide 'male-like' and 'female-like' plumage scores.
@@
 From these values we obtained means and standard deviations for each species.
-Using the same spatial sampling, we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006].
+Using the same spatial sampling, we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial [LIG; 120,000 - 140,000 years ago; @Otto-Bliesner_2006].
 Furthermore, to estimate variability in the energy available to species, we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution.
-Estimates of NPP had 30 second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005].
+Estimates of NPP had 30 second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 [MOD17A3; @Zhao_2005].
 Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
@@
 PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
-Given that species range is a potentially informative predictor of speciation and extinction we controlled for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}).
+Given that species range is a potentially informative predictor of speciation and extinction, we controlled for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}).
 To obtain long-term variation in climates for each species range we take the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
 Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
-Details and code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.
+Details and ``R`` code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.
 \phantomsection
@@
 For $\lambda_{BAMM}$ and $\mu_{BAMM}$ we were also able to use model weights sourced from the inverse of the variance of tip-rates for a given species from the posterior distribution (*n* = 1,000).
-Model weights thus reflect the degree of precision to which each species tip-rate is measured in BAMM.
+Model weights thus reflect the degree of precision to which each species' tip-rate is measured in BAMM.
 Using model selection we only compared interaction terms between a measure of sexual selection (sexual dichromatism or male-biased sexual selection) and environmental measures.
@@
 That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection and range size to have effects on each other and not just on speciation rate.
-To minimise path complexity we use temperature seasonality (BIO4) as the single measure for environmental variability and use $\lambda_{DR}$ as the single measure of speciation.
-The phylogenetic path analysis used tip-rates and correlation structure from the MCC tree.
+To minimise path complexity we used temperature seasonality (BIO4) as the single measure for environmental variability, $\lambda_{DR}$ as the single measure of speciation and the tip-rates from the MCC tree.
 Further details of the path analysis, including reasons for path directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
@@
 Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subset for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
-Models using data subsetted for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
 Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with highest power using near-complete taxon sampling (*n* = 5,812).
@@
 $\lambda_{DR}$ are those from the MCC tree and images of birds are from the _Handbook of the Birds of the World_.
-Clockwise the species are: _Sporophila bouvronides_, _Geospiza magnirostris_, _Atlapetes leucopterus_, _Euphonia rufiventris_, _Euplectes franciscanus_,_Promerops gurneyi_, _Phainopepla nitens_, _Zosterops erythropleurus_, _Acrocephalus australis_, _Progne cryptoleuca_, _Mayrornis schistaceus_, _Paradisaea rubra_, _Pachycephala pectoralis_, _Malurus pulcherrimus_, _Cranioleuca curtata_, _Cercomacra manu_, _Muscisaxicola capistratus_, _Lepidothrix coeruleocapilla_.
+Clockwise the species are: _Sporophila bouvronides_, _Geospiza magnirostris_, _Atlapetes leucopterus_, _Euphonia rufiventris_, _Euplectes franciscanus_, _Promerops gurneyi_, _Phainopepla nitens_, _Zosterops erythropleurus_, _Acrocephalus australis_, _Progne cryptoleuca_, _Mayrornis schistaceus_, _Paradisaea rubra_, _Pachycephala pectoralis_, _Malurus pulcherrimus_, _Cranioleuca curtata_, _Cercomacra manu_, _Muscisaxicola capistratus_, _Lepidothrix coeruleocapilla_.
 Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation and should not be interpreted.
@@
 ![Path analysis of evolutionary and ecological variables.
-Arrows represent direct effects with numeric values corresponding to colours (blue = positive, red = negative).
+Arrows represent direct effects with the direction of effect corresponding to colours (blue = positive, red = negative).
 The numeric values are standardised regression slopes and the asterisks indicates that the 95 % confidence intervals of this estimate do not overlap with zero.
@@
 
-We found no evidence that sexual dichromatism alters the rate of speciation in passerine birds independently or via interactions with environmental variability.
-This result is consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and two measures of dichromatism (spectral and RGB).
-The results are also consistent with previous studies performed at a smaller scale [@Huang_2014; @Cooney_2017].
-However, we found that when using a measure of male-biased sexual selection --- instead of sexual dichromatism --- there was a positive association with speciation, though the strength of this relationship varied across different rate estimates.
-Our findings suggest that the components of the composite measure of male-biased sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care) are better predictors of speciation than sexual dichromatism.
-No bioclimatic measures of environmental variability (temperature seasonality, long-term temperature variation and spatial temperature variation) predicted speciation, but range size was negatively associated with two measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$).
-This suggests that species with smaller range sizes have increased speciation rate, however, small range size may be a cause or effect of elevated speciation rate and potentially even an artifact of taxonomic classification.
-Interestingly, we found a discrepancy in the model results obtained from tip-rate statistics ($\lambda_{DR}$, $\lambda_{ND}$) and those using tip-rate estimates from a model-based approach ($\lambda_{BAMM}$), whereby BAMM produced precisely null results for most associations between predictors and speciation as well as extinction.
+We found evidence that the composite measure of male-biased sexual selection, but not the measure of sexual dichromatism, explains variation in the rate of speciation in passerine birds.
+The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB).
+These findings substantiate previous studies performed at smaller scales in birds, where sexual dichromatism measures were obtained using spectrophotometry [@Huang_2014] and human observers [@Cooney_2017].
+The correlation between speciation rate and the index of male-biased sexual selection (which reflects sexual size dimorphism, social polygyny, and [lack of] paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$, although the estimated sign and magnitude of the correlation was broadly similar across all speciation measures.
+Interestingly, we found a negative relationship between range size and speciation rates, at least when this rate was quantified by $\lambda_{DR}$ and $\lambda_{ND}$.
+This suggests that species with smaller range sizes have increased speciation rate.
+None of the bioclimatic measures of environmental variability that we investigated (*i.e.*, temperature seasonality, long-term temperature variation, and spatial temperature variation) predicted speciation rate.
 
@@
 There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
-Firstly, ecological pressures could drive sexual dichromatism in the absence of mate choice if sex-specific niches provide opportunity for the evolution of sex-limited traits [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989].
-Specifically, natural selection on females may be important in explaining patterns of sexual dichromatism in birds [@Price_2014].
-In fairy-wrens (*Malurus cyaneus*), female colouration has evolved in response to environmental pressures, leading to increased sexual dichromatism in regions where females are under stronger selection from predation [@Medina_2017].
-Ecological impacts on sexual dichromatism are supported by our path analysis, which reveals that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit relatively weakly.
+Firstly, sexual dichromatism can evolve for reasons other than sexual selection, such as when males and females occupy different ecological niches [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989] or experience different selective pressures in contexts other than competition for mates [@Price_2014].
+For example, in superb fairy-wrens (*Malurus cyaneus*) female colouration has evolved in response to spatial variation in predation pressure [@Medina_2017].
+Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit relatively weakly.
 Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
-For instance, sexual selection may promote investments in other mating signals (*e.g.* bird song) that trade-off against plumage colour, leading to a negative correlation between different sexual selection proxies [@Cooney_2018].
+For instance, sexual selection may promote investments in other mating signals (*e.g.* bird song) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
 Lastly, not all plumage colouration honestly reflects mate quality.
-A recent meta-analysis found that converted carotenoids (but *not* those sourced from diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
+A recent meta-analysis found that converted carotenoids (but *not* carotenoids sourced from the diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
 Given the importance of honest signaling in sexual selection and the variability in colour production mechanisms across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
 
-In-line with predictions and a previous meta-analysis [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate --- at least for $\lambda_{DR}$ and $\lambda_{ND}$ estimates --- but this association was not dependent on environmental variability.
-The environmental variability predictors used here often reflected changes in climate across space and time, whereby high environmental variation would likely increase the strength of natural selection on a given species.
-Given that we did not detect an effect of environmental variability (by itself or as an interaction with sexual selection) we are left with several possible explanations.
-Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability the species is evolving under.
-Specifically, sexual antagonism may be more easily purged in environments where pressures are cyclic (*e.g.* seasonality); whereas when environmental variability is directional (*e.g.* long-term climate change) sexual antagonism can remain indefinitely as a fitness burden [@Connallon_Hall_2016].
+In-line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least for $\lambda_{DR}$ and $\lambda_{ND}$ estimates.
+Additionally, we found that this association appears to be independent of net primary productivity and spatiotemporal variation in the environment.
+The lack of an effect of these environmental variables on speciation rate (either alone or in combination with sexual selection) has several possible interpretations.
+Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability under which the species is evolving.
+Specifically, speciation rates may be impacted by genetic constraints on adaptation, that vary across environments.
+Theory suggests that sexual antagonism (often arising from increased sexual selection) may be easily purged in environments where pressures are cyclic (*e.g.* seasonality), whereas when environmental variability is directional (*e.g.* long-term climate change) sexual antagonism can indefinitely limit adaptation rates [@Connallon_Hall_2016].
 Secondly, the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
@@
 
-The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates; however reduced range size can be a cause or consequence of speciation.
+The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates.
 Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995].
-However, birds with smaller range sizes are predicted to have lower dispersal ability, which would promote reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
-Alternatively, reduced range size may be a consequence of speciation.
+However, birds with smaller range sizes might often have lower dispersal ability or more specialised niches, which would promote reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
+Alternatively, reduced range size may be a consequence (rather than a cause) of speciation.
 Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives [@Rosenzweig_1995; @Weir_2011; @Price_2014].
 However, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
-Across islands we expect to see a correlation between speciation and small range size because they reflect newly formed environments with empty niches.
-Thus, our findings may be heavily dependent on island radiations (*e.g.* *Geospiza* sp.; a.k.a.
-Darwin's finches).
+Across islands we expect to see a correlation between speciation and small range size because islands often reflect newly formed environments with empty niches.
+Thus, our findings may be heavily dependent on island radiations (*e.g.* "Darwin's finches", *Geospiza* spp.).
 One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
@@
 Sexual selection may increase extinction risk as it often leads individuals (usually males) to invest in exaggerated mating signals as opposed to traits --- such as parental care --- that elevate offspring fitness [*reviewed in* @Kokko_Brooks_2003].
-Using BAMM, we found no association between extinction rate and sexual dichromatism, male-biased sexual selection or measures of environmental variability.
-Extinction is notoriously hard to estimate accurately [*see* @Rabosky_Extinction_2016].
-It can be estimated from phylogenies, the fossil record, documented recent extinctions and IUCN extinction threat status, but each approach has limitations.
-Phylogenetic methods --- like BAMM --- allow for speciation and extinction rates to be measured using moderately sized phylogenies.
-However, BAMM has been the subject of controversy over its ability to model evolutionary rate shifts and extinction rates [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], and while several tip-rate estimates exist for speciation (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$) extinction estimates are not as easily obtained.
-Here, we found that BAMM often produced homogeneous speciation and extinction rates for clades with few rate shifts which could arguably reduce our power to detect small differences in extinction rates among closely related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
+Using the model-based approach of BAMM, we found no association between extinction rate and sexual dichromatism, male-biased sexual selection or measures of environmental variability.
+Extinction is notoriously difficult to estimate accurately from phylogenies, principally because many different combinations of speciation and extinction rates can give rise to similar patterns of diversity [*see* @Rabosky_Extinction_2016].
+Phylogenetic methods such as BAMM allow for speciation and extinction rates to be measured using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates  remains a subject of debate [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
+Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models and are frought with issues of sampling bias [@Davis_2013].
+Although extinction rates can be estimated from alternative sources, such as the fossil record, documented recent extinctions and IUCN extinction threat status, each approach has limitations.
+Across the passerine bird phylogeny, we found that BAMM often produced homogeneous speciation and extinction rates for smaller clades showing few rate shifts, which might reduce our power to detect small differences in extinction rates among closely-related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
+Thus, this methodological constraint likely decreases our ability to accurately measure the correlation between metrics of sexual selection and the probability of extinction.
 
-We have shown that in passerines, male-biased sexual selection, but not sexual dichromatism, predicts speciation, independent of several measures of environmental variability.
+To summarise, we have shown that in passerines, male-biased sexual selection, but not sexual dichromatism, predicts speciation, independent of several measures of environmental variability.
 If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold, including for conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].
 Furthermore, our finding that high speciation rate is associated with smaller range size highlights the threat to the persistence of rapidly speciating lineages in a world with increased habitat loss and anthropogenic stress.
-Range size is the best predictor of extinction risk [@Harris_2008] and the association found here implies that many newly speciated clades may be at greater risk of extinction due to their small range being impacted by anthropogenic effects.
+This concern arises as the best predictor of extinction risk is range size [@Harris_2008] and the association found here implies that many newly-speciated clades have small range sizes and could thus be at greater risk of extinction.
 
@@
 
-One of the biggest unravelings of this project was the decoupling of male-biased sexual selection from sexual dichromatism, remarkably this same controversy engrossed Wallace and Darwin 130 years ago [@Kottler_1980].
-The widespread use of sexual dichromatism as a proxy for sexual selection often goes unchallenged as our instinctive Darwinian minds reminisce of peacocks and birds of paradise; my thoughts were such when I began this scientific inquiry.
+One of the biggest unravelings of this project was the decoupling of male-biased sexual selection from sexual dichromatism in predicting speciation, remarkably this same controversy engrossed Wallace and Darwin 130 years ago [@Kottler_1980].
+The widespread use of sexual dichromatism as a proxy for sexual selection often goes unchallenged as our instinctive Darwinian minds reminisce of peacocks and birds of paradise.
 Although, sexual dichromatism measurements from illustrations were always going to be an imperfect estimate of sexual selection, across 5,831 species I anticipated a signal would permeate through the noise.
-Alas --- this was not so ---- and male-biased sexual selection and sexual dichromatism had different effects on speciation.
+Alas --- this was not so --- and male-biased sexual selection and sexual dichromatism had different effects on speciation.
 With this in mind, the practical use of sexual dichromatism in macroevolutionary research is severely weakened.
@@
 Proxies of sexual selection increase the accessibility, available sample size and transferability of research, making a study more conducive to research synthesis.
-But, as Wallace remarks, the matter is inevitably more complex, despite the irresistibility of the Darwinian argument\footnote{Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took the position where natural selection on predator-prone nesting females led to females losing their plumage colouration.} [@Kottler_1980; @Caro_2017].
-On the bright side, the multivariate measure of male-biased sexual selection (sexual dimorphism, social polygyny and [lack of] paternal care) is a dataset that would benefit from being expanded across birds.
-Future research may also benefit from expanding this multivariate measure to account for ornamentation, often overlooked by sexual dimorphism and dichromatism measures [*see* Gomes_2016].
+But, as Wallace remarks, the matter is inevitably more complex, despite the irresistibility of the Darwinian argument\footnote[1]{Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took the position where natural selection on predator-prone nesting females led to females losing their plumage colouration.} [@Kottler_1980; @Caro_2017].
+On the bright side, the multivariate measure of male-biased sexual selection (sexual dimorphism, social polygyny and [lack of] paternal care) is a dataset that would benefit from being embiggened across birds.
+Future research may also benefit from expanding this multivariate measure to account for ornamentation, often overlooked by sexual dimorphism and dichromatism measures [@Gomes_2016].
 
-Not only are the measurements of sexual selection under a cloud of uncertainty, the estimates of speciation and extinction are another point of controversy.
-Macroevolutionary research is a field plagued by methodological uncertainty and conflicts.
+Like sexual selection, estimates of speciation and extinction are a point of controversy, with macroevolutionary research often plagued by methodological uncertainty.
 Originally, I had planned to continue my meta-analysis by investigating extinction rate across a large taxonomic group.
-There were signs that extinction rate could be measured from phylogenetic methods with relative confidence [@Pyron_2016].
+There were signs that extinction rate could be measured from phylogenies with confidence [*see* @Pyron_2016].
 However, these hopes soon dissipated, and with them the seeming \emph{raison d'\^{e}tre} of the study.
 Extinction rate is arguably the best measure of population fitness as it tracks survival over multiple generations.
-And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured accurately over evolutionarily meaningful timescales [*but see* @Martins_2018].
+And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured alongside sexual selection over evolutionarily meaningful timescales [*but see* @Martins_2018].
 We included measurements of extinction rate in this study, however the findings were played-down.
 The movement away from extinction and towards speciation is because of the controversy surrounding extinction rate measures as well as phylogentic uncertainty.
-Measuring extinction rates from the @Jetz_2012 bird phylogeny is problematic because of the nature of phylogenetic construction: a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction rate [@Beaulieu_2015].
-Furthermore, the very detection of absence of data --- which is what such methods attempt to do --- is a hard sell, both conceptually and methodologically come peer review: see @Rabosky_Extinction_2016 for a cautionary tale on measuring extinction rate with ``BAMM``.
+Measuring extinction rates from the @Jetz_2012 bird phylogeny is problematic because of the nature of phylogenetic construction: a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction [@Beaulieu_2015].
+Specifically, a birth-only tree will likely lead to underestimates of extinction, while phylogenies relying on taxon placement can cause imbalance in branching patterns and bias in extinction rate estimates [@Rabosky_Extinction_2016].
+Furthermore, the very detection of absent data --- which is what such methods attempt to do --- is a conceptually hard sell come peer review.
 The mechanisms and rates at which speciation and extinction occur are an active area of research.
 In a recent pre-print @Diaz_2018 used ``BAMM`` across 104 time-calibrated molecular phylogenies and found that despite variation in ecology, biogeography and taxonomy, younger clades consistently had higher rates of speciation and extinction.
-This finding of time-dependency in rates of diversification makes comparative analyses between clades challenging and calls into question countless studies.
-Similar to Wallace's view of sexual dichromatism we are left with a view of diversification rate measures where we cannot yet account for the underlying true variation; which is vastly more complex than we can currently measure.
+This finding of time-dependency in rates of diversification make comparative analyses between clades challenging and calls into question countless studies.
+Similar to Wallace's view of sexual dichromatism we are left with a view of diversification rate measures where we cannot yet account for the underlying true variation; which is vastly more complex than we currently measure.
 
-As with sexual dichromatism and macroevolutionary rates, the most severe limitation for environmental variation was the coarseness of the measures.
-In total, for 5,831 species we extracted 338,203,831 point measures\footnote{5,831 species x 1000 samples per range x 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.} of environmental variability, but these measures may not account for important variation in biotic components such as nutritional resources or predation.
-Arguably, the methods used here are much better than simply using the range mid-point to obtain a measure of temperature, precipitation or seasonality or where only one bioclimatic variable is used.
-I would argue that reducing the multivariate data into several biologically meaningful predictors is a conceptually challenging exercise, but researchers should not feel limited by the amount of raw data that they can access\footnote{This statement applies to species with documented range sizes, which varies across clades but is comprehensive for birds, mammals, amphibians and many marine groups.
-The \href{http://www.iucnredlist.org/technical-documents/spatial-data}{IUCN} has a large and accessible database for research use.
-Details of how we extracted data from these ranges is provided as a potentially useful resource in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.}.
-In fact, previous to our study on birds the same methodology was applied to ~ 100 species of bovids with speciation, extinction and environmental variability estimates obtained.
-This study was prematurely halted as there was homogeneity in the response variable (speciation and extinction rates) across the smaller bovid phylogeny.
-However, this workflow was easily transferred over to passerine birds with similar methods used for extracting environmental predictors.
-The use of extensive bioclimatic predictors and spatial data in macroevolutionary research is imperative as we scale up studies.
+The most severe limitation for environmental variation was the coarseness of the measures.
+In total, for 5,831 species, we extracted 338,203,831 point measures of environmental variability\footnote[2]{5,831 species $\times$ 1,000 samples per range $\times$ 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.}.
+But these measures may not be as valuable as their volume implies, since they do not account for variation in biotic components such as nutritional resources or predation.
+However, our methods are still improvements on using the range mid-point to obtain a measure of climate, or using a single bioclimatic variable.
+While reducing the multivariate data into several biologically meaningful predictors is a conceptually challenging exercise, researchers should not feel limited by the amount of raw data that they can access\footnote[3]{This statement applies to species with documented range sizes, which varies across clades but is comprehensive for birds, mammals, amphibians and many marine groups.
+The \href{http://www.iucnredlist.org/technical-documents/spatial-data}{IUCN} has a large and accessible database for research use.}.
+In fact, previous to our study on birds the same methodology was applied to ~100 species of bovids with speciation, extinction and environmental variability estimates obtained.
+This study was prematurely halted as there was homogeneity in speciation and extinction rates across the smaller bovid phylogeny.
+However, this workflow was easily transferred over to passerine birds, with similar methods used for extracting environmental predictors.
+The use of extensive bioclimatic predictors and spatial data in macroevolutionary research is imperative as we scale-up studies.
 At very least, if bioclimatic variables do not encapsulate key ecological parameters (*e.g.*, resources, predation, habitat structure) these predictors may account for some of the noise in the dataset and may help isolate the effect of the key moderator variables of interest.
@@
 
-I had initially envisaged that this comparative study would allow for the reliable measure of extinction rate, use multiple taxonomic groups and more robust measures of sexual selection [*see* @Anthes_2016].
+I had initially envisaged that this comparative study would allow for the reliable measure of extinction rate, use multiple taxonomic groups and robustly measure sexual selection [*see* @Anthes_2016].
 As such, many questions regarding this topic remain unanswered.
-Here I discuss the broader directions that we may follow in order to consolidate our understanding of sexual selection and its effects on fitness and diversification.
-I will draw heavily on the discussions from a Royal Society meeting on '*Sexual selection: patterns in the history of life*'\footnote{Specifically, the following two sections are spurred from dialogue with \href{https://paleobiology.si.edu/staff/individuals/hunt.cfm}{Dr Gene Hunt} and \href{https://www.earth.ox.ac.uk/people/erin-saupe/}{Associate Professor Erin E.
+Here, I discuss the broader directions that we may follow in order to consolidate our understanding of sexual selection and its effects on fitness and diversification.
+I will draw heavily on the discussions from a Royal Society meeting on '*Sexual selection: patterns in the history of life*'\footnote[4]{Specifically, the following two sections are spurred from dialogue with \href{https://scholar.google.com.au/citations?user=SVV7KH4AAAAJ&hl=en&oi=ao}{Dr Gene Hunt} and \href{https://scholar.google.com.au/citations?user=zG45LlAAAAAJ&hl=en}{Associate Professor Erin E.
 Saupe}} in the UK this year.
-This meeting spurred further inquiry into the subject and how interdisciplinary approaches can be harnessed to liberate ourselves from conventional comparative methodologies.
+This meeting spurred further inquiry into the topic and how interdisciplinary approaches can be harnessed to liberate ourselves from conventional comparative methodologies.
 
-Estimating extinction risk from the fossil record is equally fraught because many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
+Estimating extinction risk from the fossil record is difficult because many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
+But there are exceptions.
 Using fossil ostracods (a group of marine invertebrates), @Martins_2018 found that species with high levels of sexual dimorphism were quickly swept to extinction.
-From stratographic occurrence data a capture–mark–recapture model was used to determine the rate of extinction for 93 species.
-Their finding challenges the results from experimental evolution studies that suggest sexual selection mitigates extinction [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015].
-By contrast, evidence is mixed regarding the relationship between sexual selection and extinction in extant birds [@Doherty_2003; @Mitra_1996; @Morrow_2003] and mammals [@Bro-Jorgensen_2014; @Morrow_2004].
-Unfortunately, these studies have relied on extinction threat status from the International Union for Conservation of Nature (IUCN) Red List and may be biased to species under greater anthropogenic threat, culturally valued or well researched.
-In future, it would be incredibly valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual dimorphism to conduct such a study.
-In the interim I would be eager to see how ``BAMM`` performs on an ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected from the phylogeny.
+From stratographic occurrence data, a capture–mark–recapture model was used to determine the extinction rate of 93 species.
+This finding challenges the results from experimental evolution studies that suggest sexual selection mitigates extinction [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015].
+The use of fossil data overcomes limitations of other comparative analyses, which by contrast, show mixed evidence in the relationship between sexual selection and extinction in birds [@Doherty_2003; @Mitra_1996; @Morrow_2003] and mammals [@Bro-Jorgensen_2014; @Morrow_2004].
+In these cases, studies have often relied on extinction threat status from the IUCN Red List and are likely biased to species under greater anthropogenic threat, culturally valued or well researched.
+In future, it would be valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual dimorphism to conduct such a study.
+In the interim I would be eager to see how ``BAMM`` performs on a robust ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected from the phylogeny.
 This will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction rates across a wider variety of species.
@@
 One aspect and only foreseeable shortcoming of the ostracod study is that it did not account for environmental variability; a challenging proposal for extinct species nevertheless.
-However, the use of paleoclimate models in evolution is now a reality; with recent studies incorporating averaged values across multiple paleoclimate models as predictors of speciation [*see* @quintero_2018].
-While the uncertainty surrounding paleoclimates was beyond the scope of the study, future studies should remain vigilant of the everlasting need to use key ecological processes in macroevolutionary research [*reviewed in* @Weber_2017].
-In our study on passerine birds we found that range size is negatively associated with speciation, yet range size encapsulates multiple aspects of ecology and evolution so it is unclear what the finding truly means.
-The role of range size in speciation may be further investigated through ancestral geographic range reconstruction [@Quintero_2015], as ancestral ranges may be beneficial in understanding modes of speciation and the causality of range size expansions and contractions [@Losos_2003].
+However, the use of paleoclimate models in evolutionary biology is now a reality; with recent studies incorporating averaged values across multiple paleoclimate models as predictors of speciation [@quintero_2018].
+While the uncertainty surrounding paleoclimates was beyond the scope of the study, future studies should remain vigilant of the need to integrate key ecological processes in macroevolutionary research [*reviewed in* @Weber_2017].
+Here, we found that range size is negatively associated with speciation, yet range size encapsulates multiple aspects of ecology and evolution so it is unclear what the finding truly means.
+The role of range size in speciation may be further investigated through ancestral geographic range reconstruction [@Quintero_2015], which may be beneficial in understanding modes of speciation and the causality of range size expansions and contractions [@Losos_2003].
 Within Palaeobiology the role of range size and niche breadth on diversification patterns has been the subject of recent interest [*e.g.* @Saupe_2015; @Saupe_2016].
-Across 92 bivalve species that lived during a mid-Pliocene warming period, @Saupe_2015 used environmental niche modelling to assess the impact of ecological processes on extinction.
-Harnessing ecological niche modelling in macroevolutionary studies across extant taxa such as passerine birds is a daunting challenge but offers an *eco-evo* synthesis that may be pertinent in questions relating to extinction; especially if the taxonomic group under inquiry is of high conservation value, for which the songbirds certainly are.
+Across 92 bivalve species that lived during a mid-Pliocene warming period, @Saupe_2015 used environmental niche modelling to assess the impact of the realized *vs* fundamental niche on extinction.
+Harnessing ecological niche modelling in macroevolutionary studies across extant taxa such as passerine birds is a daunting challenge but offers an *eco-evo* synthesis that may be pertinent in questions relating to extinction; especially if the taxonomic group under inquiry is of high conservation value; for which the songbirds certainly are.

### 2018-10-20 Manuscript.Rmd
Manuscript Changes

@@
   sexual selection might precipitate extinction by favouring competitive phenotypes
-  with poor survival, or stave it off by helping populations shed deleterious mutations.
+  with poor survival or stave it off by helping populations shed deleterious mutations.
   Here, we investigate the relationship between sexual selection and rates of speciation
@@
 This evolutionary force has been proposed to facilitate speciation through the evolution of divergent signals associated with improved mating success [@Lande_1981; @Lande_1982].
-For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism are more species rich [@Seddon_2008].
+For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism are more species-rich [@Seddon_2008].
 Under this model of speciation, divergent mating signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].
@@
 In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-Through these synergetic effects sexual selection and natural selection --- such as environmental change over space and time --- may lead to increased adaptation and speciation rates.
+Through these synergetic effects sexual selection and natural selection (*e.g.* environmental change over space and time) may lead to increased adaptation and speciation rates.
 
@@
 We also conduct analyses that allow phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] to affect the precision of our estimates, rather than implicitly assuming that the single best-supported phylogenetic tree, or consensus tree, is correct.
-Thirdly, our analysis includes several potentially important environmental variables as predictors: range size, seasonal variation in temperature, spatial temperature variation (across a species’ range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).
+Thirdly, our analysis includes several potentially important environmental variables as predictors: range size, seasonal variation in temperature, spatial temperature variation (across a species’ range), long-term variation in temperature between now and the last interglacial (LIG) as well net primary productivity (NPP).
 By including these environmental variables in our models, we statistically account for their effects on variation in diversification rates and thereby improve our estimates of the effects of sexual selection.
@@
 We used a previously published measure of sexual dichromatism for 5,983 species of passerines [@Dale_2015].
-Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forhead, throat, upper breast and lower breast) from _Handbook of the Birds of the World_ [@delhoyo2011].
+Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forehead, throat, upper breast, and lower breast) from _Handbook of the Birds of the World_ [@delhoyo2011].
 The relative contribution of male and female RGB colour values were averaged across body patches and provide 'male-like' and 'female-like' plumage scores.
@@
 Because of taxonomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
-From these distributions we obtained estimates of climatic conditions that each species is exposed to.
-That is, we took 1,000 random point samples in each range and extracted 19 present-day bioclimatic variables (representing a variety of biologically relevant annual trends in temperature and precipitation) with 30 seconds (~1 km^2^) spatial resolution [@Fick_2017].
-From these values we obtained means and standard deviations for each species.
+From these distributions, we obtained estimates of climatic conditions that each species is exposed to.
+That is, we took 1,000 random point samples in each range and extracted 19 present-day bioclimatic variables (representing a variety of biologically relevant annual trends in temperature and precipitation) with 30-second (~1 km^2^) spatial resolution [@Fick_2017].
+From these values, we obtained means and standard deviations for each species.
 Using the same spatial sampling, we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial [LIG; 120,000 - 140,000 years ago; @Otto-Bliesner_2006].
 Furthermore, to estimate variability in the energy available to species, we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution.
-Estimates of NPP had 30 second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 [MOD17A3; @Zhao_2005].
+Estimates of NPP had 30-second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 [MOD17A3; @Zhao_2005].
 Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
@@
 To obtain long-term variation in climates for each species range we take the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
-Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
+Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
+The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
 Details and ``R`` code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.
@@
 Phylogenetic information was obtained from www.birdtree.org [@Jetz_2012].
-We used a maximum clade credibility (MCC) tree from 2,500 samples of the posterior distribution that were subset to the passerine bird order (*n* = 5,965) as the main phylogentic tree in our comparative analysis.
+We used a maximum clade credibility (MCC) tree from 2,500 samples of the posterior distribution that were subset to the passerine bird order (*n* = 5,965) as the main phylogenetic tree in our comparative analysis.
 Additionally, a random draw of 100 full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures and BAMM [@Rabosky_2014].
-These trees used a 'Hackett backbone' [@Hackett_2008] and were constructed using the a pure birth (Yule) model.
+These trees used a 'Hackett backbone' [@Hackett_2008] and were constructed using a pure birth (Yule) model.
 We calculated three different tip-rate metrics of speciation and one of extinction across all trees.
@@
 Importantly, these models provide tip-rate estimates of speciation and extinction rate.
-The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}; briefly we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
+The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}; briefly, we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
 Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
@@
 Using model selection we only compared interaction terms between a measure of sexual selection (sexual dichromatism or male-biased sexual selection) and environmental measures.
-That is, we fixed the individual predictors of: dichromatism/male-biased sexual selection measures, log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation and NPP while comparing 32 models with different combinations of interactions (including none).
+That is, we fixed the individual predictors of dichromatism/male-biased sexual selection measures, log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP while comparing 32 models with different combinations of interactions (including none).
 Model selection was done in ``MuMIn`` using the ``dredge`` function [@MuMIn_2017].
@@
 The phylogenetic path analysis was used to assess causal paths *between* variables not able to be modelled within the univariate response of PGLS.
-That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection and range size to have effects on each other and not just on speciation rate.
+That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection, and range size to have effects on each other and not just on speciation rate.
 To minimise path complexity we used temperature seasonality (BIO4) as the single measure for environmental variability, $\lambda_{DR}$ as the single measure of speciation and the tip-rates from the MCC tree.
@@
 
-PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in model selection.
+PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) were included in model selection.
 However, no interaction terms were significant or present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
 Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
-Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
+Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
 \phantomsection
@@
 Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
-Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with highest power using near-complete taxon sampling (*n* = 5,812).
+Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with the highest power using near-complete taxon sampling (*n* = 5,812).
 \phantomsection
@@
 
-We tested our hypothesis that sexual selection is associated with speciation using several methods and across 100 trees drawn from the post burn-in posterior [@Jetz_2012].
+We tested our hypothesis that sexual selection is associated with speciation using several methods and across 100 trees drawn from the post-burn-in posterior [@Jetz_2012].
 We used a range of methods to overcome uncertainty in the power and precision of various tip-rate estimates as well as the variation between trees.
@@
 From 1,000 posterior samples of the MCC BAMM run, the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figures S6} and \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
-Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) was high.
+Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) were high.
 This means that although the number of rate shifts reached convergence ($median = 59$; \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S7}), the locations of the rate shifts (*i.e.* the credible shift set) across the tree are highly heterogeneous.
 The variability in the locations for the shift configurations is a likely source of uncertainty in downstream tip-rates used as the response variable in PGLS models.
-Notably, the 95 % HPD interval for model estimates using BAMM rates from 100 trees was about 20-fold the 95 % confidence intervals of the estimate from the MCC tree; whereas for $\lambda_{DR}$ and $\lambda_{ND}$ the HPD 95 % interval width was equal to or less than the MCC 95 % CI (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S9}).
+Notably, the 95 % HPD interval for model estimates using BAMM rates from 100 trees was about 20-fold the 95 % confidence intervals of the estimate from the MCC tree; whereas for $\lambda_{DR}$ and $\lambda_{ND}$, the HPD 95 % interval width was equal to or less than the MCC 95 % CI (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S9}).
 
@@
 The association between sexual selection and sexual dichromatism in birds is a Darwinian trope [@Darwin_1871] that allows it to be commonly used as a proxy for the strength of sexual selection in comparative studies [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
-Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015] we find that sexual dichromatism does not predict speciation rate, whereas male-biased sexual selection does.
+Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015], we find that sexual dichromatism does not predict speciation rate, whereas male-biased sexual selection does.
 Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection, and that variation in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
@@
 Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
-For instance, sexual selection may promote investments in other mating signals (*e.g.* bird song) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
+For instance, sexual selection may promote investments in other mating signals (*e.g.* birdsong) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
 Lastly, not all plumage colouration honestly reflects mate quality.
@@
 Secondly, the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
-Specifically, access to dietary resources, and the impacts of predation or parasitism are unaccounted for here.
+Specifically, access to dietary resources and the impacts of predation or parasitism are unaccounted for here.
 These are likely key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
@@
 However, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
-Across islands we expect to see a correlation between speciation and small range size because islands often reflect newly formed environments with empty niches.
+Across islands, we expect to see a correlation between speciation and small range size because small islands often reflect newly formed environments with empty niches.
 Thus, our findings may be heavily dependent on island radiations (*e.g.* "Darwin's finches", *Geospiza* spp.).
@@
 Phylogenetic methods such as BAMM allow for speciation and extinction rates to be measured using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates remains a subject of debate [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
-Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models and are frought with issues of sampling bias [@Davis_2013].
+Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models and are fraught with issues of sampling bias [@Davis_2013].
 Although extinction rates can be estimated from alternative sources, such as the fossil record, documented recent extinctions and IUCN extinction threat status, each approach has limitations.
@@
 The widespread use of sexual dichromatism as a proxy for sexual selection often goes unchallenged as our instinctive Darwinian minds reminisce of peacocks and birds of paradise.
-Although, sexual dichromatism measurements from illustrations were always going to be an imperfect estimate of sexual selection, across 5,831 species I anticipated a signal would permeate through the noise.
+Although sexual dichromatism measurements from illustrations were always going to be an imperfect estimate of sexual selection, across 5,831 species I anticipated a signal would permeate through the noise.
 Alas --- this was not so --- and male-biased sexual selection and sexual dichromatism had different effects on speciation.
@@
 And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured alongside sexual selection over evolutionarily meaningful timescales [*but see* @Martins_2018].
-We included measurements of extinction rate in this study, however the findings were played-down.
-The movement away from extinction and towards speciation is because of the controversy surrounding extinction rate measures as well as phylogentic uncertainty.
+We included measurements of extinction rate in this study, but the findings were played-down.
+The movement away from extinction and towards speciation is because of the controversy surrounding extinction rate measures as well as phylogenetic uncertainty.
 Measuring extinction rates from the @Jetz_2012 bird phylogeny is problematic because of the nature of phylogenetic construction: a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction [@Beaulieu_2015].
-Specifically, a birth-only tree will likely lead to underestimates of extinction, while phylogenies relying on taxon placement can cause imbalance in branching patterns and bias in extinction rate estimates [@Rabosky_Extinction_2016].
+Specifically, a birth-only tree will likely lead to underestimates of extinction, while phylogenies relying on taxon placement can cause an imbalance in branching patterns and bias in extinction rate estimates [@Rabosky_Extinction_2016].
 Furthermore, the very detection of absent data --- which is what such methods attempt to do --- is a conceptually hard sell come peer review.
-The mechanisms and rates at which speciation and extinction occur are an active area of research.
-In a recent pre-print @Diaz_2018 used ``BAMM`` across 104 time-calibrated molecular phylogenies and found that despite variation in ecology, biogeography and taxonomy, younger clades consistently had higher rates of speciation and extinction.
+The mechanisms and rates at which speciation and extinction occur is an active area of research.
+In a recent pre-print @Diaz_2018 used ``BAMM`` across 104 time-calibrated molecular phylogenies and found that despite variation in ecology, biogeography, and taxonomy, younger clades consistently had higher rates of speciation and extinction.
 This finding of time-dependency in rates of diversification make comparative analyses between clades challenging and calls into question countless studies.
-Similar to Wallace's view of sexual dichromatism we are left with a view of diversification rate measures where we cannot yet account for the underlying true variation; which is vastly more complex than we currently measure.
+Similar to Wallace's view of sexual dichromatism, we are left with a view of diversification rate measures where we cannot yet account for the underlying true variation; which is vastly more complex than we currently measure.
 
@@
 In total, for 5,831 species, we extracted 338,203,831 point measures of environmental variability\footnote[2]{5,831 species $\times$ 1,000 samples per range $\times$ 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.}.
-But these measures may not be as valuable as their volume implies, since they do not account for variation in biotic components such as nutritional resources or predation.
-However, our methods are still improvements on using the range mid-point to obtain a measure of climate, or using a single bioclimatic variable.
+But these measures may not be as valuable as their volume implies since they do not account for variation in biotic components such as nutritional resources or predation.
+However, our methods are still improvements on using the range mid-point to obtain a measure of climate or using a single bioclimatic variable.
 While reducing the multivariate data into several biologically meaningful predictors is a conceptually challenging exercise, researchers should not feel limited by the amount of raw data that they can access\footnote[3]{This statement applies to species with documented range sizes, which varies across clades but is comprehensive for birds, mammals, amphibians and many marine groups.
@@
 
-Estimating extinction risk from the fossil record is difficult because many clades (*e.g.* birds) lack the stratographic sampling across time to statistically calculate extinction rates.
+Estimating extinction risk from the fossil record is difficult because many clades (*e.g.* birds) lack the stratigraphic sampling across time to statistically calculate extinction rates.
 But there are exceptions.
 Using fossil ostracods (a group of marine invertebrates), @Martins_2018 found that species with high levels of sexual dimorphism were quickly swept to extinction.
-From stratographic occurrence data, a capture–mark–recapture model was used to determine the extinction rate of 93 species.
+From stratigraphic occurrence data, a capture–mark–recapture model was used to determine the extinction rate of 93 species.
 This finding challenges the results from experimental evolution studies that suggest sexual selection mitigates extinction [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015].
@@
 In future, it would be valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual dimorphism to conduct such a study.
-In the interim I would be eager to see how ``BAMM`` performs on a robust ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected from the phylogeny.
+In the interim, I would be eager to see how ``BAMM`` performs on a robust ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected from the phylogeny.
 This will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction rates across a wider variety of species.

### 2019-05-22 Sexual Selection and Speciation.Rmd
Updated DR and ND: Now with 1000 models

@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r setup, message=FALSE, warning=FALSE}
 library(ggplot2)
@@
 library(ggraph)
+library(stringr)
 #devtools::install_github("Ax3man/phylopath")
@@
 source('functions/gheatmap2.R') #Gets rid of spaces between tiles
-```
-
-```{r setup, include=FALSE}
 knitr::opts_chunk$set(cache=TRUE)
@@
 
-#Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
-DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
-  left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
-            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ES")), 
-            by = "TipLabel")
-})
-
-#PGLS needs tiplabel as rowname
-DR.model.data <- lapply(DR.model.data, function(x) {
-  tibble::column_to_rownames(x, "TipLabel")})
-
-#Prune the trees
-pruned.trees<-lapply(passerine.trees, function(x) {
-  drop.tip(x,x$tip.label[-match(restricted.data$TipLabel, x$tip.label)])
-})
-
-#Use mapply to create a list of PGLS global models
-DR.pgls.models <- mcmapply(function(x,y) {
-  gls(ES ~ SDi 
-         + log(range.size.m2)
-         + bioclim4 #Seasonal variation
-         + residuals.PC1 #Spatial variation
-         + PC1.LIG #Long-term climate variation
-         + NPP,
-    corPagel(0.985, phy = y, fixed = TRUE), 
-    data = x, 
-    method = "REML")
-}, x = DR.model.data, y = pruned.trees,
-SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
-mc.cores = 8) #Specify core number
-
-saveRDS(DR.pgls.models, "data/DR.pgls.models.rds")
-
-#Now for Node Density:
-ND.model.data <- lapply(nd.list, function(x) {
-  left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
-            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
-            by = "TipLabel")
-})
-
-#PGLS needs tiplabel as rowname
-ND.model.data <- lapply(ND.model.data, function(x) {
-  tibble::column_to_rownames(x, "TipLabel")})
-
-#Use mapply to create a list of PGLS global models
-ND.pgls.models <- mcmapply(function(x,y) {
-gls(ND ~ SDi 
-         + log(range.size.m2)
-         + bioclim4 #Seasonal variation
-         + residuals.PC1 #Spatial variation
-         + PC1.LIG #Long-term climate variation
-         + NPP,
-    corPagel(0.9996, phy = y, fixed = TRUE), 
-    data = x, 
-    method = "REML")
-}, x = ND.model.data, y = pruned.trees, 
-SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
-mc.cores = 8) #Specify core number
-
-saveRDS(ND.pgls.models, "data/ND.pgls.models.rds")
+# No longer used now that we use 1000 trees
+
+# #Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
+# DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
+#   left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+#             x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ES")), 
+#             by = "TipLabel")
+# })
+# 
+# #PGLS needs tiplabel as rowname
+# DR.model.data <- lapply(DR.model.data, function(x) {
+#   tibble::column_to_rownames(x, "TipLabel")})
+# 
+# #Prune the trees
+# pruned.trees<-lapply(passerine.trees, function(x) {
+#   drop.tip(x,x$tip.label[-match(restricted.data$TipLabel, x$tip.label)])
+# })
+# 
+# #Use mapply to create a list of PGLS global models
+# DR.pgls.models <- mcmapply(function(x,y) {
+#   gls(ES ~ SDi 
+#          + log(range.size.m2)
+#          + bioclim4 #Seasonal variation
+#          + residuals.PC1 #Spatial variation
+#          + PC1.LIG #Long-term climate variation
+#          + NPP,
+#     corPagel(0.985, phy = y, fixed = TRUE), 
+#     data = x, 
+#     method = "REML")
+# }, x = DR.model.data, y = pruned.trees,
+# SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+# mc.cores = 8) #Specify core number 
+# 
+# saveRDS(DR.pgls.models, "data/DR.pgls.models.rds")
+# 
+# #Now for Node Density:
+# ND.model.data <- lapply(nd.list, function(x) {
+#   left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
+#             x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
+#             by = "TipLabel")
+# })
+# 
+# #PGLS needs tiplabel as rowname
+# ND.model.data <- lapply(ND.model.data, function(x) {
+#   tibble::column_to_rownames(x, "TipLabel")})
+# 
+# #Use mapply to create a list of PGLS global models
+# ND.pgls.models <- mcmapply(function(x,y) {
+# gls(ND ~ SDi 
+#          + log(range.size.m2)
+#          + bioclim4 #Seasonal variation
+#          + residuals.PC1 #Spatial variation
+#          + PC1.LIG #Long-term climate variation
+#          + NPP,
+#     corPagel(0.9996, phy = y, fixed = TRUE), 
+#     data = x, 
+#     method = "REML")
+# }, x = ND.model.data, y = pruned.trees, 
+# SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+# mc.cores = 8) #Specify core number 
+# 
+# saveRDS(ND.pgls.models, "data/ND.pgls.models.rds")
 ```
@@
 This enabled us to show the variation between trees, and the variation associated with the top MCC model.
+Given the variability across trees we ran 1000 ND and DR models, each allowed to estimate a phylogenetic signal ($\lambda$).
+The models were saved as an `.rds` object and read back into R with key coefficients extracted.
+Below we provide the code for this:
+
+```{r, eval = F}
+##########################################################
+####### Read in DR models and extract estimates: #########
+##########################################################
+
+#We ran DR and ND models on 1000 trees
+files.DR.SD <- list.files(path = "/Users/justincally/Dropbox/Runs Spartan/DR_SD/", pattern = "\\.rds$", full.names = TRUE) #1000 models
+df <- list()
+lapply(files.DR.SD, function(x) {
+  tryCatch({
+  model.x <- readRDS(x)
+  name.x <- str_sub(x, -19, -5)
+  df[[name.x]] <<- data.frame(model.x$coefficients,
+                              confint(model.x),
+                              coef(summary(model.x))[,2], #Std.Error
+                              coef(summary(model.x))[,3], #t-val
+                              coef(summary(model.x))[,4], #pval
+                              model.x[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = name.x) %>% tibble::rownames_to_column()
+  rm(model.x)
+  gc(verbose = FALSE)
+  },
+  error = function(e) NULL
+  )
+})
+DR.pgls.summary <- bind_rows(df) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
+#saveRDS(DR.pgls.summary, "data/DR.pgls.summary.rds") #Save a simple df with the model and coeff
+
+##########################################################
+####### Read in ND models and extract estimates: #########
+##########################################################
+
+files.ND.SD <- list.files(path = "/Users/justincally/Dropbox/Runs Spartan/ND_SD/", pattern = "\\.rds$", full.names = TRUE) #1000 models
+df <- list()
+lapply(files.ND.SD, function(x) {
+  tryCatch({
+  model.x <- readRDS(x) 
+  name.x <- str_sub(x, -19, -5)
+  df[[name.x]] <<- data.frame(model.x$coefficients, 
+                              confint(model.x), 
+                              coef(summary(model.x))[,2], #Std.Error
+                              coef(summary(model.x))[,3], #t-val
+                              coef(summary(model.x))[,4], #pval
+                              model.x[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = name.x) %>% tibble::rownames_to_column()
+  rm(model.x)
+  gc(verbose = FALSE)
+  },
+  error = function(e) NULL
+  )
+})
+ND.pgls.summary <- bind_rows(df) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
+#saveRDS(ND.pgls.summary, "data/ND.pgls.summary.rds") #Save a simple df with the model and coeff
+```
 
 ```{r, fig.height = 8, fig.width = 4, message = FALSE, warning = FALSE}
-DR.pgls.models <- readRDS("data/DR.pgls.models.rds")
+#Read in the 1000 tree summary df and the MCC tree model
+DR.pgls.summary <- readRDS("data/DR.pgls.summary.rds")
 MCC.DR.top <- readRDS('data/MCC.DR.top.rds')
-
-DR.pgls.summary <- lapply(DR.pgls.models, function(x) {
-  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
-})
-
-MCC.DR.summary <- data.frame(MCC.DR.top$coefficients, confint(MCC.DR.top)) %>% tibble::rownames_to_column()
-
-DR.pgls.summary <- bind_rows(DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
-
-colnames(MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+MCC.DR.summary <-  data.frame(MCC.DR.top$coefficients,
+                              confint(MCC.DR.top),
+                              coef(summary(MCC.DR.top))[,2], #Std.Error
+                              coef(summary(MCC.DR.top))[,3], #t-val
+                              coef(summary(MCC.DR.top))[,4], #pval
+                              MCC.DR.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% 
+  tibble::rownames_to_column() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
+
+#Now plot this for Node Density (ND):
+
+ND.pgls.summary <- readRDS("data/ND.pgls.summary.rds")
+MCC.ND.top <- readRDS('data/MCC.ND.top.rds')
+MCC.ND.summary <- data.frame(MCC.ND.top$coefficients,
+                              confint(MCC.ND.top),
+                              coef(summary(MCC.ND.top))[,2], #Std.Error
+                              coef(summary(MCC.ND.top))[,3], #t-val
+                              coef(summary(MCC.ND.top))[,4], #pval
+                              MCC.ND.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% 
+  tibble::rownames_to_column() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
 
@@
                position = position_nudge(y= 0.6))+
-  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
+  #geom_errorbarh(aes(xmin = LCI, xmax = UCI, colour = Parameter), position = position_jitter(seed = 1), height = 0)+
+  geom_point(shape = 21, alpha = 0.5, size = 0.75, position = position_jitter(seed = 1))+
   geom_vline(xintercept = 0)+
@@
 
-#Now plot this for Node Density (ND):
-
-ND.pgls.models <- readRDS("data/ND.pgls.models.rds")
-MCC.ND.top <- readRDS('data/MCC.ND.top.rds')
-
-ND.pgls.summary <- lapply(ND.pgls.models, function(x) {
-  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
-})
-
-MCC.ND.summary <- data.frame(MCC.ND.top$coefficients, confint(MCC.ND.top)) %>% tibble::rownames_to_column()
-
-ND.pgls.summary <- bind_rows(ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
-
-colnames(MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
-
 ND.plot <-ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
@@
 BAMM.lambda.pgls.models <- readRDS("data/BAMM.lambda.pgls.models.rds")
-MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
 
@@
 })
+
+#Now plot this for Lambda
+
+MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
+MCC.lambda.summary <- data.frame(MCC.Lambda.top$coefficients,
+                              confint(MCC.Lambda.top),
+                              coef(summary(MCC.Lambda.top))[,2], #Std.Error
+                              coef(summary(MCC.Lambda.top))[,3], #t-val
+                              coef(summary(MCC.Lambda.top))[,4], #pval
+                              MCC.Lambda.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% 
+  tibble::rownames_to_column() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
 
@@
 
-MCC.lambda.summary <- data.frame(MCC.Lambda.top$coefficients, confint(MCC.Lambda.top)) %>% tibble::rownames_to_column()
-
 BAMM.lambda.pgls.summary <- bind_rows(BAMM.lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
-
-colnames(MCC.lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
 
@@
 BAMM.mu.pgls.models <- readRDS("data/BAMM.mu.pgls.models.rds")
-MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')
 
@@
 
-MCC.mu.summary <- data.frame(MCC.Mu.top$coefficients, confint(MCC.Mu.top)) %>% tibble::rownames_to_column()
-
 BAMM.mu.pgls.summary <- bind_rows(BAMM.mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
 
-colnames(MCC.mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+#Now plot this for Lambda
+
+MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')
+MCC.mu.summary <- data.frame(MCC.Mu.top$coefficients,
+                              confint(MCC.Mu.top),
+                              coef(summary(MCC.Mu.top))[,2], #Std.Error
+                              coef(summary(MCC.Mu.top))[,3], #t-val
+                              coef(summary(MCC.Mu.top))[,4], #pval
+                              MCC.Mu.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% 
+  tibble::rownames_to_column() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
 
@@
 ```{r}
-MCC.DR.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates")
-MCC.ND.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates")
-MCC.lambda.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates")
-MCC.mu.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates")
+MCC.DR.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates", split.table = Inf)
+MCC.ND.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates", split.table = Inf)
+MCC.lambda.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates", split.table = Inf)
+MCC.mu.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates", split.table = Inf)
 ```
@@
 
-#Run the 100 models for DR and ND using the best model:
-Subset.data.noMCC <- SS.subset %>% dplyr::select(TipLabel, Sexual_selection_ppca, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP)
-
-#Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
-Subset.DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
-  left_join(Subset.data.noMCC, 
-            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "DR")), 
-            by = "TipLabel")
-})
-
-#PGLS needs tiplabel as rowname
-Subset.DR.model.data <- lapply(Subset.DR.model.data, function(x) {
-  tibble::column_to_rownames(x, "TipLabel")})
-
-#Prune the trees
-Subset.pruned.trees<-lapply(passerine.trees, function(x) {
-  drop.tip(x,x$tip.label[-match(SS.subset$TipLabel, x$tip.label)])
-})
-
-#Use mapply to create a list of PGLS global models
-Subset.DR.pgls.models <- mcmapply(function(x,y) {
-  gls(DR ~ Sexual_selection_ppca 
-         + log(range.size.m2)
-         + bioclim4 #Seasonal variation
-         + residuals.PC1 #Spatial variation
-         + PC1.LIG #Long-term climate variation
-         + NPP,
-    correlation = corPagel(0.9711, phy = y, fixed = TRUE), 
-    data = x, 
-    method = "REML")
-}, x = Subset.DR.model.data, y = Subset.pruned.trees,
-SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
-mc.cores = 8) #Specify core number
-
-saveRDS(Subset.DR.pgls.models, "data/Subset.DR.pgls.models.rds")
-
-#Now for Node Density:
-Subset.ND.model.data <- lapply(nd.list, function(x) { #es.list is a list of ES values calculated earlier
-  left_join(Subset.data.noMCC, 
-            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
-            by = "TipLabel")
-})
-
-#PGLS needs tiplabel as rowname
-Subset.ND.model.data <- lapply(Subset.ND.model.data, function(x) {
-  tibble::column_to_rownames(x, "TipLabel")})
-
-#Use mapply to create a list of PGLS global models
-Subset.ND.pgls.models <- mcmapply(function(x,y) {
-gls(ND ~ Sexual_selection_ppca 
-         + log(range.size.m2)
-         + bioclim4 #Seasonal variation
-         + residuals.PC1 #Spatial variation
-         + PC1.LIG #Long-term climate variation
-         + NPP,
-    corPagel(1, phy = y, fixed = TRUE), 
-    data = x, 
-    method = "REML")
-}, x = Subset.ND.model.data, y = Subset.pruned.trees, 
-SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
-mc.cores = 8) #Specify core number
-
-saveRDS(Subset.ND.pgls.models, "data/Subset.ND.pgls.models.rds")
+## No longer needed
+
+# #Run the 100 models for DR and ND using the best model:
+# Subset.data.noMCC <- SS.subset %>% dplyr::select(TipLabel, Sexual_selection_ppca, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP)
+# 
+# #Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
+# Subset.DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
+#   left_join(Subset.data.noMCC, 
+#             x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "DR")), 
+#             by = "TipLabel")
+# })
+# 
+# #PGLS needs tiplabel as rowname
+# Subset.DR.model.data <- lapply(Subset.DR.model.data, function(x) {
+#   tibble::column_to_rownames(x, "TipLabel")})
+# 
+# #Prune the trees
+# Subset.pruned.trees<-lapply(passerine.trees, function(x) {
+#   drop.tip(x,x$tip.label[-match(SS.subset$TipLabel, x$tip.label)])
+# })
+# 
+# #Use mapply to create a list of PGLS global models
+# Subset.DR.pgls.models <- mcmapply(function(x,y) {
+#   gls(DR ~ Sexual_selection_ppca 
+#          + log(range.size.m2)
+#          + bioclim4 #Seasonal variation
+#          + residuals.PC1 #Spatial variation
+#          + PC1.LIG #Long-term climate variation
+#          + NPP,
+#     correlation = corPagel(0.9711, phy = y, fixed = TRUE), 
+#     data = x, 
+#     method = "REML")
+# }, x = Subset.DR.model.data, y = Subset.pruned.trees,
+# SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+# mc.cores = 8) #Specify core number 
+# 
+# saveRDS(Subset.DR.pgls.models, "data/Subset.DR.pgls.models.rds")
+# 
+# #Now for Node Density:
+# Subset.ND.model.data <- lapply(nd.list, function(x) { #es.list is a list of ES values calculated earlier
+#   left_join(Subset.data.noMCC, 
+#             x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
+#             by = "TipLabel")
+# })
+# 
+# #PGLS needs tiplabel as rowname
+# Subset.ND.model.data <- lapply(Subset.ND.model.data, function(x) {
+#   tibble::column_to_rownames(x, "TipLabel")})
+# 
+# #Use mapply to create a list of PGLS global models
+# Subset.ND.pgls.models <- mcmapply(function(x,y) {
+# gls(ND ~ Sexual_selection_ppca 
+#          + log(range.size.m2)
+#          + bioclim4 #Seasonal variation
+#          + residuals.PC1 #Spatial variation
+#          + PC1.LIG #Long-term climate variation
+#          + NPP,
+#     corPagel(1, phy = y, fixed = TRUE), 
+#     data = x, 
+#     method = "REML")
+# }, x = Subset.ND.model.data, y = Subset.pruned.trees, 
+# SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
+# mc.cores = 8) #Specify core number 
+# 
+# saveRDS(Subset.ND.pgls.models, "data/Subset.ND.pgls.models.rds")
 
@@
 
+```{r, eval = F}
+
+##########################################################
+####### Read in DR models and extract estimates: #########
+##########################################################
+
+#We ran DR and ND models on 1000 trees
+files.DR.SS <- list.files(path = "/Users/justincally/Dropbox/Runs Spartan/DR_SS/", pattern = "\\.rds$", full.names = TRUE) #1000 models
+df <- list()
+rds.list <- list()
+models <- list()
+c.list <- NULL
+
+lapply(files.DR.SS, function(x) {
+models <- readRDS(x)
+rds.list <- unlist(models, recursive = F, use.names = T)
+c.list <<- c(c.list, rds.list[-grep(".log", names(rds.list))])
+rm(models)
+rm(rds.list)
+gc()
+})
+
+lapply(names(c.list), function(x) {
+  tryCatch({
+  model <- c.list[[x]]
+  df[[x]] <<- data.frame(model$coefficients,
+                              confint(model),
+                              coef(summary(model))[,2], #Std.Error
+                              coef(summary(model))[,3], #t-val
+                              coef(summary(model))[,4], #pval
+                              model[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = x) %>% tibble::rownames_to_column()
+  gc(verbose = FALSE)
+  },
+  error = function(e) NULL
+  )
+})
+
+Subset.DR.pgls.summary <- bind_rows(df) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
+#saveRDS(Subset.DR.pgls.summary, "data/Subset.DR.pgls.summary.rds") #Save a simple df with the model and coeff
+
+#read in 1000 trees:
+
+#We ran DR and ND models on 1000 trees
+files.ND.SS <- list.files(path = "/Users/justincally/Dropbox/Runs Spartan/ND_SS/", pattern = "\\.rds$", full.names = TRUE) #1000 models
+df <- list()
+rds.list <- list()
+models <- list()
+c.list <- NULL
+
+lapply(files.ND.SS, function(x) {
+models <- readRDS(x)
+rds.list <- unlist(models, recursive = F, use.names = T)
+c.list <<- c(c.list, rds.list[-grep(".log", names(rds.list))])
+rm(models)
+rm(rds.list)
+gc()
+})
+
+lapply(names(c.list), function(x) {
+  tryCatch({
+  model <- c.list[[x]]
+  df[[x]] <<- data.frame(model$coefficients,
+                              confint(model),
+                              coef(summary(model))[,2], #Std.Error
+                              coef(summary(model))[,3], #t-val
+                              coef(summary(model))[,4], #pval
+                              model[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = x) %>% tibble::rownames_to_column()
+  gc(verbose = FALSE)
+  },
+  error = function(e) NULL
+  )
+})
+
+Subset.ND.pgls.summary <- bind_rows(df) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
+#saveRDS(Subset.ND.pgls.summary, "data/Subset.ND.pgls.summary.rds") #Save a simple df with the model and coeff
+rm(c.list)
+############################
+```
+
 ```{r}
-Subset.DR.pgls.models <- readRDS("data/Subset.DR.pgls.models.rds")
+Subset.DR.pgls.summary <- readRDS("data/Subset.DR.pgls.summary.rds")
+Subset.ND.pgls.summary <- readRDS("data/Subset.ND.pgls.summary.rds")
+
+#Cols for plot: 
+Subset.cols <- brewer.pal(n = 7, name = "Dark2")[-6]
+
+### MCC
 Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')
 
-Subset.cols <- brewer.pal(n = 7, name = "Dark2")[-6]
-
-Subset.DR.pgls.summary <- lapply(Subset.DR.pgls.models, function(x) {
-  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
-})
-
-Subset.MCC.DR.summary <- data.frame(Subset.MCC.DR.top$coefficients, confint(Subset.MCC.DR.top)) %>% tibble::rownames_to_column()
-
-Subset.DR.pgls.summary <- bind_rows(Subset.DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
-
-colnames(Subset.MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+Subset.MCC.DR.summary <- data.frame(Subset.MCC.DR.top$coefficients,
+                              confint(Subset.MCC.DR.top),
+                              coef(summary(Subset.MCC.DR.top))[,2], #Std.Error
+                              coef(summary(Subset.MCC.DR.top))[,3], #t-val
+                              coef(summary(Subset.MCC.DR.top))[,4], #pval
+                              Subset.MCC.DR.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% tibble::rownames_to_column()
+
+colnames(Subset.MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name")
 
@@
 
-Subset.ND.pgls.models <- readRDS("data/Subset.ND.pgls.models.rds")
 Subset.MCC.ND.top <- readRDS('data/Subset.MCC.ND.top.rds')
 
-Subset.ND.pgls.summary <- lapply(Subset.ND.pgls.models, function(x) {
-  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
-})
-
-Subset.MCC.ND.summary <- data.frame(Subset.MCC.ND.top$coefficients, confint(Subset.MCC.ND.top)) %>% tibble::rownames_to_column()
-
-Subset.ND.pgls.summary <- bind_rows(Subset.ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
-
-colnames(Subset.MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+Subset.MCC.ND.summary <- data.frame(Subset.MCC.ND.top$coefficients,
+                              confint(Subset.MCC.ND.top),
+                              coef(summary(Subset.MCC.ND.top))[,2], #Std.Error
+                              coef(summary(Subset.MCC.ND.top))[,3], #t-val
+                              coef(summary(Subset.MCC.ND.top))[,4], #pval
+                              Subset.MCC.ND.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% tibble::rownames_to_column()
+
+colnames(Subset.MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name")
 
@@
 
-Subset.MCC.Lambda.summary <- data.frame(Subset.MCC.Lambda.top$coefficients, confint(Subset.MCC.Lambda.top)) %>% tibble::rownames_to_column()
-
 Subset.Lambda.pgls.summary <- bind_rows(Subset.Lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
 
-colnames(Subset.MCC.Lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+#MCC
+
+Subset.MCC.Lambda.summary <- data.frame(Subset.MCC.Lambda.top$coefficients,
+                              confint(Subset.MCC.Lambda.top),
+                              coef(summary(Subset.MCC.Lambda.top))[,2], #Std.Error
+                              coef(summary(Subset.MCC.Lambda.top))[,3], #t-val
+                              coef(summary(Subset.MCC.Lambda.top))[,4], #pval
+                              Subset.MCC.Lambda.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% tibble::rownames_to_column()
+
+colnames(Subset.MCC.Lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name")
 
@@
 })
-
-Subset.MCC.Mu.summary <- data.frame(Subset.MCC.Mu.top$coefficients, confint(Subset.MCC.Mu.top)) %>% tibble::rownames_to_column()
-
 Subset.Mu.pgls.summary <- bind_rows(Subset.Mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
 
-colnames(Subset.MCC.Mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI")
+#MCC
+
+Subset.MCC.Mu.summary <- data.frame(Subset.MCC.Mu.top$coefficients,
+                              confint(Subset.MCC.Mu.top),
+                              coef(summary(Subset.MCC.Mu.top))[,2], #Std.Error
+                              coef(summary(Subset.MCC.Mu.top))[,3], #t-val
+                              coef(summary(Subset.MCC.Mu.top))[,4], #pval
+                              Subset.MCC.Mu.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% tibble::rownames_to_column()
+
+colnames(Subset.MCC.Mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name")
 
@@
 ```{r}
-Subset.MCC.DR.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates from Male-bias SS")
-Subset.MCC.ND.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates from Male-bias SS")
-Subset.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates from Male-bias SS")
-Subset.MCC.Mu.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates from Male-bias SS")
+Subset.MCC.DR.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates from Male-bias SS", split.table = Inf)
+Subset.MCC.ND.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates from Male-bias SS", split.table = Inf)
+Subset.MCC.Lambda.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates from Male-bias SS", split.table = Inf)
+Subset.MCC.Mu.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates from Male-bias SS", split.table = Inf)
 ```

### 2019-05-22 thousand_trees.R

@@ -1,0 +1,427 @@
+###############################################################################
+### Testing the Association Between Sexual Dichromatism/Sexual Selection 
+###                   and speciation using 1,000 trees       ##################
+###############################################################################
+
+##############################
+#####       Setup      #######
+##############################
+library(foreach)
+library(doParallel)
+library(nlme)
+library(phytools)
+library(ape)
+library(dplyr)
+source("functions/diversification_rate_calculation.R")
+
+trees <- list()
+
+do_one_tree <- function(tree){
+  xfjvbzdfsg
+  write.rds(zxfv)
+}
+
+slurm.apply(trees, do_one_tree)
+
+
+###Generate 1000 trees for the analysis
+
+#thousand_trees <- passerine.trees
+#t5.trees <- read.nexus('data/trees/passerine_2500.nex')
+# '%ni%' <- Negate('%in%')
+# hundred.tree.names <- names(passerine.trees)
+# 
+# for(x in names(t5.trees)){
+#       ifelse(length(thousand_trees) < 1000, 
+#              thousand_trees[[x]] <- t5.trees[[x]],
+#              break)
+#     }
+
+##Save DFs
+
+# full.data.for.model <- restricted.data %>% select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP)
+# 
+# subset.data.for.model <- SS.subset %>% select(binomial = binomial.x, TipLabel, Sexual_selection_ppca, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP)
+
+#saveRDS(thousand_trees, file = "data/trees/thousand_trees.rds")
+# saveRDS(full.data.for.model, file = "data/full.data.for.model.rds")
+# saveRDS(subset.data.for.model, file = "data/subset.data.for.model.rds")
+
+###############################################################
+###Running the loop: 
+
+##Read in the trees and data for model
+
+thousand_trees <- readRDS(file = "data/trees/thousand_trees.rds")
+
+full.data.for.model <- readRDS(file = "data/full.data.for.model.rds")
+subset.data.for.model <- readRDS(file = "data/subset.data.for.model.rds")
+
+#Set up cluster and register it with R
+
+########################################
+######### SPECIFY CORES HERE!!!! #######
+########################################
+
+
+#setup parallel backend to use many processors
+cl <- makeCluster(4) #not to overload your computer
[…]

### 2019-07-12 ComparativePlots.Rmd
changes to figures

@@
         axis.ticks.x.bottom = element_line())
+
+SS.subset %>% ggplot(aes(x = Sexual_selection_ppca, y = 0, fill = ..x..))+
+  geom_density_ridges_gradient(scale = 3, size = 0.05, gradient_lwd = 0.75)+
+  scale_fill_distiller(palette = "Greys", direction = 1, guide = F)+
+  theme_minimal()+
+  scale_y_discrete(expand = c(0,0))+
+  theme(axis.text.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 14),
+        axis.line.y = element_blank(), 
+        panel.grid = element_blank(),
+        axis.title.x = element_blank(),
+        axis.ticks.x.bottom = element_line())
 ```
@@
           plot.margin = unit(c(.5,.5,.5,.5), "cm"))
-
-#Get more leasonable intercept for gls slope
-SDi.intercept <- coef(gls(MCC.DR ~ SDi, data = restricted.data))[1]
-
-DR.scatter <- restricted.data %>% ggplot(aes(y = exp(MCC.DR), x = SDi, fill = ..x..))+
-  geom_point(shape = 21, stroke = 0.35)+
-  scale_fill_distiller(palette = "OrRd", direction = 1, guide = F)+
-  theme_bw()+
-  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
-  scale_x_continuous(expand = c(0,0.3))+
-  xlab("Sexual Dichromatism")+
-  ylab("Speciation Rate")+
-  geom_smooth(method = "glm", se = F, color = "black", linetype = 5)+ 
-    theme(panel.grid = element_blank(),
-          axis.line = element_line(),
-          panel.border = element_blank(),
-          plot.margin = unit(c(.5,.5,.5,.5), "cm"))
-
-#Get more leasonable intercept for gls slope
-range.model <- coef(gls(MCC.DR ~ log(range.size.m2), 
-                            data = restricted.data))
-
-newRange <- data.frame(`range.size.m2` = restricted.data$range.size,
-                         SDi = mean(restricted.data$SDi),
-                         NPP = mean(restricted.data$NPP),
-                         bioclim4 = mean(restricted.data$bioclim4),
-                         residuals.PC1 = mean(restricted.data$residuals.PC1),
-                         PC1.LIG = mean(restricted.data$PC1.LIG))
-restricted.data$Range.Predict <- predict(MCC.DR.top, newRange)
-
-Range.scatter <- restricted.data %>% ggplot(aes(y = exp(MCC.DR), x = range.size.m2, fill = ..x..))+
-  geom_point(shape = 21, stroke = 0.35)+
-  scale_fill_distiller(palette = "BuPu", direction = 1, guide = F, trans = "log")+
-  theme_bw()+
-  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
-  scale_x_continuous(expand = c(0,0.3), trans = "log", breaks = c(1, 1*(10^7), 1*(10^10), 1*(10^13)))+
-  xlab("Range Size")+
-  ylab("Speciation Rate")+
-    geom_smooth(method = "glm", se = F, color = "black", linetype = 5)+
-    theme(panel.grid = element_blank(),
-          axis.line = element_line(),
-          panel.border = element_blank(),
-          plot.margin = unit(c(.5,.5,.5,.5), "cm"))
-
-BAMM.scatter
-DR.scatter
-Range.scatter
 ```
@@
 
-# ggtree.plot <-gheatmap2(ggtree.plot, data = MCC.BAMM.model.data %>% select(SDi), width=0.05, colnames = FALSE, low = "orange", high = "red")
+ ggtree.plot <-gheatmap2(ggtree.plot, data = MCC.BAMM.model.data %>% left_join(SS.subset %>% select(TipLabel, Sexual_selection_ppca), by = "TipLabel") %>% select(Sexual_selection_ppca), width=0.05, colnames = FALSE, low = "grey90", high = "black")
 
@@
   xlim(-75, NA)+
-  scale_colour_gradientn(colours = cols, 
-                         #breaks = c(-4, -2, 1.75, 1.2),
-                         guide = F) +
+  scale_color_distiller(palette = "YlGn", direction = 1, guide = F, trans = "log", breaks=c(0.05, 0.15, 0.5, 1.5))+
   theme(legend.position = "bottom",
@@
 
-# ggtree.plot.dr <- gheatmap2(ggtree.plot.dr, data = MCC.BAMM.model.data %>% select(SDi), width=0.05, colnames = FALSE) + 
-#   scale_fill_distiller(palette = "OrRd", direction = 1, na.value = "white", guide = guide_colorbar(direction = "horizontal", label.position = "bottom", barheight = 1.5, barwidth = 25, label.vjust = -0.5))+
+# ggtree.plot.dr.1 <- gheatmap2(ggtree.plot.dr, data = MCC.BAMM.model.data %>% `rownames<-`(MCC.BAMM.model.data$TipLabel) %>% select(SDi), width=0.05, colnames = FALSE)  +
+# scale_fill_distiller(palette = "OrRd", direction = 1, na.value = "white", guide = F)
+# 
+# ggtree.plot.dr.2 <- ggtree.plot.dr.1 + new_scale_fill() 
+# ggtree.plot.dr.2 <- gheatmap2(ggtree.plot.dr.2, data = MCC.BAMM.model.data %>% `rownames<-`(MCC.BAMM.model.data$TipLabel) %>% select(SDi), width=0.05, colnames = FALSE)  +
+# scale_fill_distiller(palette = "BuPu", direction = 1, na.value = "white", guide = F)
+# 
+# ggtree.plot.dr.3 <- ggtree.plot.dr.2 + new_scale_fill() 
+# ggtree.plot.dr.3 <- gheatmap2(ggtree.plot.dr.3, data = MCC.BAMM.model.data %>% `rownames<-`(MCC.BAMM.model.data$TipLabel) %>% left_join(SS.subset %>% select(TipLabel, Sexual_selection_ppca), by = "TipLabel") %>% select(Sexual_selection_ppca), width=0.05, colnames = FALSE)  +
+# scale_fill_distiller(palette = "Greys", direction = 1, na.value = "white", guide = F)
+
+ggtree.plot.dr.3 <- gheatmap2(ggtree.plot.dr, data = MCC.BAMM.model.data %>% left_join(SS.subset %>% select(TipLabel, Sexual_selection_ppca), by = "TipLabel")%>% `rownames<-`(MCC.BAMM.model.data$TipLabel) %>% select(Sexual_selection_ppca), width=0.05, colnames = FALSE, low = "Grey90", high = "Black")
+
 #     theme(legend.position = "bottom",
 #         legend.text = element_text(size = 18))
+# 
+# ggtree.plot.dr.h <- gheatmap2(ggtree.plot.dr, data = MCC.BAMM.model.data %>% select(SDi), width=0.5, colnames = FALSE, low = "grey90", high = "black")
+
 ```
@@
 ```{r}
+
 Lepidothrix.coeruleocapilla.data <- restricted.data %>% filter(TipLabel == "Lepidothrix_coeruleocapilla")
@@
 ```
+
+#Figure 1 in manuscript (filtered model results)
+
+```{r, fig.height = 8, fig.width = 4, message = FALSE, warning = FALSE}
+library(dplyr)
+library(ggplot2)
+library(Cairo)
+library(data.table)
+
+#Cols for plot: 
+cols1 <- c("#8856a7", "#c6dbef")
+cols2 <- c("#8856a7", "#fc8d59")
+
+#Read in the 1000 tree summary df and the MCC tree model
+DR.pgls.summary <- readRDS("data/DR.pgls.summary.rds")
+MCC.DR.top <- readRDS('data/MCC.DR.top.rds')
+MCC.DR.summary <-  data.frame(MCC.DR.top$coefficients,
+                              confint(MCC.DR.top),
+                              coef(summary(MCC.DR.top))[,2], #Std.Error
+                              coef(summary(MCC.DR.top))[,3], #t-val
+                              coef(summary(MCC.DR.top))[,4], #pval
+                              MCC.DR.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% 
+  tibble::rownames_to_column() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
+
+#Now plot this for Node Density (ND):
+
+ND.pgls.summary <- readRDS("data/ND.pgls.summary.rds")
+MCC.ND.top <- readRDS('data/MCC.ND.top.rds')
+MCC.ND.summary <- data.frame(MCC.ND.top$coefficients,
+                              confint(MCC.ND.top),
+                              coef(summary(MCC.ND.top))[,2], #Std.Error
+                              coef(summary(MCC.ND.top))[,3], #t-val
+                              coef(summary(MCC.ND.top))[,4], #pval
+                              MCC.ND.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% 
+  tibble::rownames_to_column() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
+
+parameter_names <- c(
+                    `bioclim4` = "Temperature Seasonality",
+                    `log(range.size.m2)` = "Range Size (log-transformed)",
+                    `NPP` = "NPP",
+                    `PC1.LIG` = "Long-term Temperature Variation",
+                    `residuals.PC1` = "Spatial Temperature Variation",
+                    `SDi` = "Sexual Dichromatism"
+                    )
+
+DR.plot.sub <-DR.pgls.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")) %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  #geom_errorbarh(aes(xmin = LCI, xmax = UCI, colour = Parameter), position = position_jitter(seed = 1), height = 0)+
+  geom_point(shape = 21, alpha = 0.25, size = 0.75, position = position_jitter(seed = 1))+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+DR.plot.sub <- DR.plot.sub + geom_errorbarh(data = MCC.DR.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = MCC.DR.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = cols1)+
+  scale_color_manual(values = cols1)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.title = element_text(hjust = 0.5),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle(expression(paste(lambda['DR'], sep = "")))
+
+ND.plot.sub <-ND.pgls.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")) %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.25, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+ND.plot.sub <- ND.plot.sub + geom_errorbarh(data = MCC.ND.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = MCC.ND.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,0.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = cols1)+
+  scale_color_manual(values = cols1)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.title = element_text(hjust = 0.5),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle(expression(paste(lambda['ND'], sep = "")))
+
+BAMM.lambda.pgls.models <- readRDS("data/BAMM.lambda.pgls.models.rds")
+
+BAMM.lambda.pgls.summary <- lapply(BAMM.lambda.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+#Now plot this for Lambda
+
+MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
+MCC.lambda.summary <- data.frame(MCC.Lambda.top$coefficients,
+                              confint(MCC.Lambda.top),
+                              coef(summary(MCC.Lambda.top))[,2], #Std.Error
+                              coef(summary(MCC.Lambda.top))[,3], #t-val
+                              coef(summary(MCC.Lambda.top))[,4], #pval
+                              MCC.Lambda.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% 
+  tibble::rownames_to_column() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
+
+BAMM.lambda.pgls.summary <- bind_rows(BAMM.lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+  # for (x in unique(BAMM.lambda.pgls.summary$Parameter)[2:7]){
+  #   filter(Parameter == x & between(Estimate, left = as.numeric(hpd.Lambda.top[1,x]), right = as.numeric(hpd.Lambda.top[2,x])))
+  #   }
+
+remove_outliers <- function(x, na.rm = TRUE, ...) {
+  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
+  H <- 1.5 * IQR(x, na.rm = na.rm)
+  y <- x
+  y[x < (qnt[1] - H)] <- NA
+  y[x > (qnt[2] + H)] <- NA
+  y
+}
+
+BAMM.lambda.pgls.summary.RO <- data.table::dcast(BAMM.lambda.pgls.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), Estimate ~ Parameter, value.var = "Estimate")
+BAMM.lambda.pgls.summary.RO$Estimate <- NULL
+BAMM.lambda.pgls.summary.RO <- sapply(BAMM.lambda.pgls.summary.RO, function(x) {
+  remove_outliers(x, na.rm = T)})
+BAMM.lambda.pgls.summary.RO <-melt(BAMM.lambda.pgls.summary.RO) %>% na.omit()
+BAMM.lambda.pgls.summary.RO$Var1 <- NULL
+colnames(BAMM.lambda.pgls.summary.RO) <- c("Parameter", "Estimate")
+
+BAMM.lambda.plot.sub <- BAMM.lambda.pgls.summary.RO %>%
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.25, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+BAMM.lambda.plot.sub <- BAMM.lambda.plot.sub + geom_errorbarh(data = MCC.lambda.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = MCC.lambda.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,0.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 7, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = cols1)+
+  scale_color_manual(values = cols1)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.title = element_text(hjust = 0.5),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle(expression(paste(lambda['BAMM'], sep = "")))
+
+#Figure for extinction
+
+BAMM.mu.pgls.models <- readRDS("data/BAMM.mu.pgls.models.rds")
+
+BAMM.mu.pgls.summary <- lapply(BAMM.mu.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+BAMM.mu.pgls.summary <- bind_rows(BAMM.mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+#Now plot this for Lambda
+
+MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')
+MCC.mu.summary <- data.frame(MCC.Mu.top$coefficients,
+                              confint(MCC.Mu.top),
+                              coef(summary(MCC.Mu.top))[,2], #Std.Error
+                              coef(summary(MCC.Mu.top))[,3], #t-val
+                              coef(summary(MCC.Mu.top))[,4], #pval
+                              MCC.Mu.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% 
+  tibble::rownames_to_column() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name"))
+
+#Get rid of outliers
+BAMM.mu.pgls.summary.RO <- data.table::dcast(BAMM.mu.pgls.summary %>% dplyr::select(Parameter,Estimate) %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), Estimate ~ Parameter, value.var = "Estimate")
+BAMM.mu.pgls.summary.RO$Estimate <- NULL
+BAMM.mu.pgls.summary.RO <- sapply(BAMM.mu.pgls.summary.RO, function(x) {
+  remove_outliers(x, na.rm = T)})
+BAMM.mu.pgls.summary.RO <-melt(BAMM.mu.pgls.summary.RO) %>% na.omit()
+BAMM.mu.pgls.summary.RO$Var1 <- NULL
+colnames(BAMM.mu.pgls.summary.RO) <- c("Parameter", "Estimate")
+
+BAMM.Mu.plot.sub <-BAMM.mu.pgls.summary.RO %>%
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.25, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+BAMM.Mu.plot.sub <- BAMM.Mu.plot.sub + geom_errorbarh(data = MCC.mu.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = MCC.mu.summary %>% filter(Parameter %in% c("log(range.size.m2)", "SDi")), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,0.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 7, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = cols1)+
+  scale_color_manual(values = cols1)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.title = element_text(hjust = 0.5),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle(expression(paste(mu['BAMM'], sep = "")))
+
+Subset.DR.pgls.summary <- readRDS("data/Subset.DR.pgls.summary.rds")
+Subset.ND.pgls.summary <- readRDS("data/Subset.ND.pgls.summary.rds")
+
+### MCC
+Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')
+
+Subset.MCC.DR.summary <- data.frame(Subset.MCC.DR.top$coefficients,
+                              confint(Subset.MCC.DR.top),
+                              coef(summary(Subset.MCC.DR.top))[,2], #Std.Error
+                              coef(summary(Subset.MCC.DR.top))[,3], #t-val
+                              coef(summary(Subset.MCC.DR.top))[,4], #pval
+                              Subset.MCC.DR.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% tibble::rownames_to_column()
+
+colnames(Subset.MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name")
+
+parameter_names <- c(
+                    `bioclim4` = "Temperature Seasonality",
+                    `log(range.size.m2)` = "Range Size (log-transformed)",
+                    `NPP` = "NPP",
+                    `PC1.LIG` = "Long-term Temperature Variation",
+                    `residuals.PC1` = "Spatial Temperature Variation",
+                    `Sexual_selection_ppca` = "Sexual Selection"
+                    )
+
+Subset.DR.pgls.summary$Parameter = factor(Subset.DR.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.MCC.DR.summary$Parameter = factor(Subset.MCC.DR.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.DR.plot.sub <-Subset.DR.pgls.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")) %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.25, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Subset.DR.plot.sub <- Subset.DR.plot.sub + geom_errorbarh(data = Subset.MCC.DR.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Subset.MCC.DR.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = cols2)+
+  scale_color_manual(values = cols2)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.title = element_text(hjust = 0.5),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle(expression(paste(lambda['DR'], sep = "")))
+
+####################################
+#For ND
+
+Subset.MCC.ND.top <- readRDS('data/Subset.MCC.ND.top.rds')
+
+Subset.MCC.ND.summary <- data.frame(Subset.MCC.ND.top$coefficients,
+                              confint(Subset.MCC.ND.top),
+                              coef(summary(Subset.MCC.ND.top))[,2], #Std.Error
+                              coef(summary(Subset.MCC.ND.top))[,3], #t-val
+                              coef(summary(Subset.MCC.ND.top))[,4], #pval
+                              Subset.MCC.ND.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% tibble::rownames_to_column()
+
+colnames(Subset.MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name")
+
+Subset.ND.pgls.summary$Parameter = factor(Subset.ND.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.MCC.ND.summary$Parameter = factor(Subset.MCC.ND.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.ND.plot.sub <-Subset.ND.pgls.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")) %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.25, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Subset.ND.plot.sub <- Subset.ND.plot.sub + geom_errorbarh(data = Subset.MCC.ND.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Subset.MCC.ND.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = cols2)+
+  scale_color_manual(values = cols2)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.title = element_text(hjust = 0.5),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle(expression(paste(lambda['ND'], sep = "")))
+
+######################
+
+#For Lambda
+Subset.BAMM.lambda.pgls.models <- readRDS("data/Subset.BAMM.lambda.pgls.models.rds")
+Subset.MCC.Lambda.top <- readRDS('data/Subset.MCC.Lambda.top.rds')
+
+Subset.Lambda.pgls.summary <- lapply(Subset.BAMM.lambda.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+
+Subset.Lambda.pgls.summary <- bind_rows(Subset.Lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+#MCC
+
+Subset.MCC.Lambda.summary <- data.frame(Subset.MCC.Lambda.top$coefficients,
+                              confint(Subset.MCC.Lambda.top),
+                              coef(summary(Subset.MCC.Lambda.top))[,2], #Std.Error
+                              coef(summary(Subset.MCC.Lambda.top))[,3], #t-val
+                              coef(summary(Subset.MCC.Lambda.top))[,4], #pval
+                              Subset.MCC.Lambda.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% tibble::rownames_to_column()
+
+colnames(Subset.MCC.Lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name")
+
+Subset.Lambda.pgls.summary$Parameter = factor(Subset.Lambda.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.MCC.Lambda.summary$Parameter = factor(Subset.MCC.Lambda.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.Lambda.pgls.summary.RO <- data.table::dcast(Subset.Lambda.pgls.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), Estimate ~ Parameter, value.var = "Estimate")
+Subset.Lambda.pgls.summary.RO$Estimate <- NULL
+Subset.Lambda.pgls.summary.RO <- sapply(Subset.Lambda.pgls.summary.RO, function(x) {
+  remove_outliers(x, na.rm = T)})
+Subset.Lambda.pgls.summary.RO <-melt(Subset.Lambda.pgls.summary.RO) %>% na.omit()
+Subset.Lambda.pgls.summary.RO$Var1 <- NULL
+colnames(Subset.Lambda.pgls.summary.RO) <- c("Parameter", "Estimate")
+
+Subset.Lambda.plot.sub <-Subset.Lambda.pgls.summary.RO %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.25, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Subset.Lambda.plot.sub <- Subset.Lambda.plot.sub + geom_errorbarh(data = Subset.MCC.Lambda.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Subset.MCC.Lambda.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = cols2)+
+  scale_color_manual(values = cols2)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.title = element_text(hjust = 0.5),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle(expression(paste(lambda['BAMM'], sep = "")))
+
+######################
+
+#For Mu
+
+Subset.BAMM.Mu.pgls.models <- readRDS("data/Subset.BAMM.mu.pgls.models.rds")
+Subset.MCC.Mu.top <- readRDS('data/Subset.MCC.Mu.top.rds')
+
+Subset.Mu.pgls.summary <- lapply(Subset.BAMM.Mu.pgls.models, function(x) {
+  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
+})
+Subset.Mu.pgls.summary <- bind_rows(Subset.Mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+
+#MCC
+
+Subset.MCC.Mu.summary <- data.frame(Subset.MCC.Mu.top$coefficients,
+                              confint(Subset.MCC.Mu.top),
+                              coef(summary(Subset.MCC.Mu.top))[,2], #Std.Error
+                              coef(summary(Subset.MCC.Mu.top))[,3], #t-val
+                              coef(summary(Subset.MCC.Mu.top))[,4], #pval
+                              Subset.MCC.Mu.top[["modelStruct"]][["corStruct"]][1], #lambda value
+                              model = "MCC_model") %>% tibble::rownames_to_column()
+
+colnames(Subset.MCC.Mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI", "SE", "tval", "pval" , "lambda" ,"model_name")
+
+Subset.Mu.pgls.summary$Parameter = factor(Subset.Mu.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.MCC.Mu.summary$Parameter = factor(Subset.MCC.Mu.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))
+
+Subset.Mu.pgls.summary.RO <- data.table::dcast(Subset.Mu.pgls.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), Estimate ~ Parameter, value.var = "Estimate")
+Subset.Mu.pgls.summary.RO$Estimate <- NULL
+Subset.Mu.pgls.summary.RO <- sapply(Subset.Mu.pgls.summary.RO, function(x) {
+  remove_outliers(x, na.rm = T)})
+Subset.Mu.pgls.summary.RO <-melt(Subset.Mu.pgls.summary.RO) %>% na.omit()
+Subset.Mu.pgls.summary.RO$Var1 <- NULL
+colnames(Subset.Mu.pgls.summary.RO) <- c("Parameter", "Estimate")
+
+Subset.Mu.plot.sub <-Subset.Mu.pgls.summary.RO %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6))+
+  geom_jitter(shape = 21, alpha = 0.25, size = 0.75)+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")
+
+Subset.Mu.plot.sub <- Subset.Mu.plot.sub + geom_errorbarh(data = Subset.MCC.Mu.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), 
+                 aes(xmin = LCI,
+                     xmax = UCI, y = Parameter,
+                     color = Parameter), 
+                 height = 0, show.legend = F, 
+                 position = position_nudge(y= -0.75))+
+
+  geom_point(data = Subset.MCC.Mu.summary %>% filter(Parameter %in% c("log(range.size.m2)", "Sexual_selection_ppca")), 
+             aes(x = Estimate, y = Parameter, fill = Parameter), 
+             shape=21, color = "grey20",
+             size = 3,
+             position = position_nudge(y= -0.75)) +
+  scale_y_discrete(expand = c(0.5,.5))+
+  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
+  scale_fill_manual(values = cols2)+
+  scale_color_manual(values = cols2)+
+  theme(panel.grid.major.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.x = element_text(size = 7),
+        plot.title = element_text(hjust = 0.5),
+        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
+  ggtitle(expression(paste(mu['BAMM'], sep = "")))
+
+#########################################
+```
+
+Add the scatter plots
+
+```{r}
+library(nlme)
+# #Get more leasonable intercept for gls slope
+# SDi.intercept <- coef(gls(MCC.DR ~ SDi, data = restricted.data))[1]
+# 
+# DR.scatter <- restricted.data %>% ggplot(aes(y = exp(MCC.DR), x = SDi, fill = ..x..))+
+#   geom_point(shape = 21, stroke = 0.35)+
+#   scale_fill_distiller(palette = "OrRd", direction = 1, guide = F)+
+#   theme_bw()+
+#   scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
+#   scale_x_continuous(expand = c(0,0.3))+
+#   xlab("Sexual Dichromatism")+
+#   ylab("Speciation Rate")+
+#   geom_smooth(method = "glm", se = F, color = "black", linetype = 5)+ 
+#     theme(panel.grid = element_blank(),
+#           axis.line = element_line(),
+#           panel.border = element_blank(),
+#           plot.margin = unit(c(.5,.5,.5,.5), "cm"))
+
+###################
+SS.intercept <- coef(gls(MCC.DR ~ Sexual_selection_ppca, data = SS.subset))[1]
+
+DR.SS.scatter <- SS.subset %>% ggplot(aes(y = exp(MCC.DR), x = Sexual_selection_ppca, fill = ..x..))+
+  geom_point(shape = 21, stroke = 0.35, alpha = 0.6)+
+  scale_fill_gradient(low = "#fdae6b", high = "#e6550d", guide = F)+
+  geom_text(aes(x = max(SS.subset$Sexual_selection_ppca, na.rm = T), y = max(c(exp(SS.subset$MCC.DR)), exp(restricted.data$MCC.DR))), label = paste0("n = ", prettyNum(nrow(SS.subset), big.mark = ",")), fontface = 3, vjust = "inward", hjust = "inward")+
+  theme_bw()+
+  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
+  scale_x_continuous(expand = c(0,0.3))+
+  xlab("Sexual Selection")+
+  ylab(expression(paste("Speciation Rate (",lambda['DR'], ') ', sep = "")))+
+  geom_smooth(method = "glm", se = F, color = "black", linetype = 5)+ 
+    theme(panel.grid = element_blank(),
+          axis.line = element_line(),
+          panel.border = element_blank(),
+          plot.margin = unit(c(.5,2,.5,2), "cm"))
+
+#Get more leasonable intercept for gls slope
+range.model <- coef(gls(MCC.DR ~ log(range.size.m2), 
+                            data = restricted.data))
+
+newRange <- data.frame(`range.size.m2` = restricted.data$range.size,
+                         SDi = mean(restricted.data$SDi),
+                         NPP = mean(restricted.data$NPP),
+                         bioclim4 = mean(restricted.data$bioclim4),
+                         residuals.PC1 = mean(restricted.data$residuals.PC1),
+                         PC1.LIG = mean(restricted.data$PC1.LIG))
+restricted.data$Range.Predict <- predict(MCC.DR.top, newRange)
+
+Range.scatter <- restricted.data %>% ggplot(aes(y = exp(MCC.DR), x = range.size.m2, fill = ..x..))+
+  geom_point(shape = 21, stroke = 0.35, alpha = 0.6)+
+  scale_fill_gradient(low = "#bcbddc", high = "#54278f", guide = F)+
+  geom_text(aes(x = max(restricted.data$range.size.m2, na.rm = T), y = max(c(exp(restricted.data$MCC.DR)), exp(SS.subset$MCC.DR))), label = paste0("n = ", prettyNum(nrow(restricted.data), big.mark = ",")), fontface = 3, vjust = "inward", hjust = "inward")+
+  theme_bw()+
+  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
+  scale_x_continuous(expand = c(0,0.3), trans = "log", breaks = c(1, 1*(10^7), 1*(10^10), 1*(10^13)))+
+  xlab("Range Size")+
+  ylab(expression(paste("Speciation Rate (",lambda['DR'], ') ', sep = "")))+
+    geom_smooth(method = "glm", se = F, color = "black", linetype = 5)+
+    theme(panel.grid = element_blank(),
+          axis.line = element_line(),
+          panel.border = element_blank(),
+          plot.margin = unit(c(.5,2,.5,2), "cm"))
+
+```
+
+```{r}
+library(gridExtra)
+ggsave(file = "Figures/Estimates_and_scatter.pdf", width = 12, height = 10, device = cairo_pdf,
+plot = grid.arrange(
+  symmetrise_scale(DR.plot.sub, "x"),
+  symmetrise_scale(ND.plot.sub, "x"),
+  symmetrise_scale(BAMM.lambda.plot.sub, "x"),
+  symmetrise_scale(BAMM.Mu.plot.sub, "x"),
+  Range.scatter,
+  symmetrise_scale(Subset.DR.plot.sub, "x"),
+  symmetrise_scale(Subset.ND.plot.sub, "x"),
+  symmetrise_scale(Subset.Lambda.plot.sub, "x"),
+  symmetrise_scale(Subset.Mu.plot.sub, "x"),
+  DR.SS.scatter,
+  layout_matrix = rbind(c(1,2,3,4),
+                        c(5,5,10,10),
+                        c(6,7,8,9))
+))
+
+ggsave(file = "Figures/Estimates_and_scatter.pdf", width = 12, height = 10, device = cairo_pdf,
+plot = symmetrise_scale(DR.plot.sub, "x"),
+)
+```

### 2019-07-12 Manuscript_Publication.Rmd
changes to figures

--- 2018-10-20 Manuscript.Rmd
@@
   and Iliana Medina Guzmán\textsuperscript{§}
-- ~
 - '*\textit{justin.g.cally@gmail.com}'
@@
   pdf_document: default
-  word_document: default
+  html_document:
+    df_print: paged
 header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
@@
   \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{}
-  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} 
-  \usepackage[hang, flushmargin]{footmisc} 
-  \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false, linkcolor=blue, filecolor=blue, urlcolor=blue}
-  \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
+  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
+  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
+  linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
   \raggedbottom
@@
 subparagraph: yes
-abstract: 'Sexual selection can shape phylogenetic diversity by altering speciation
-  or extinction rates. The predicted effects on diversity are often contradictory;
-  for example, sexual selection might promote speciation by creating reproductive
-  isolation, or hinder it by selecting for traits that facilitate gene flow. Similarly,
-  sexual selection might precipitate extinction by favouring competitive phenotypes
-  with poor survival or stave it off by helping populations shed deleterious mutations.
-  Here, we investigate the relationship between sexual selection and rates of speciation
-  and extinction in passerine birds. We use two measures of sexual selection --- sexual
-  dichromatism (*n* = 5,812 species) and a multivariate measure of male-biased sexual
-  selection (sexual dimorphism, polygyny and paternal care; *n* = 2,465) --- as well
-  as three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$),
-  and one measure of extinction ($\mu_{BAMM}$). We also tested whether the effects
-  of sexual selection is contingent on ecological variables such as spatiotemporal
-  temperature variation, primary productivity, and range size. We find that birds
-  with smaller ranges have higher speciation rates, although the direction of causality
-  is unclear. Moreover, our results show that male-biased sexual selection, but not
-  sexual dichromatism, is an agent of adaptation and speciation. This finding suggests
-  that sexual selection is an evolutionary force shaping patterns of diversity amongst
-  songbirds, which has broad implications for ecology and conservation biology.
-\vspace{5mm}  \par\noindent \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage dichromatism, Range size, Extinction rate.'
+abstract: Sexual selection can shape phylogenetic diversity by affecting speciation or extinction rates.
+The predicted effects on diversity are often contradictory; for example, sexual selection might promote speciation by creating reproductive isolation, or hinder it by selecting for traits that facilitate gene flow.
+Here, we investigate the relationship between sexual selection and diversification in passerine birds.
+Multiple studies use sexual dichromatism as a proxy for sexual selection, however,  this relationship can be weak in many clades.
+Thus, we use two measures of sexual selection --- sexual dichromatism and a multivariate measure of male-biased sexual selection --- to test the link between sexual selection and diversification.
+We also test whether the effect of sexual selection on diversification is contingent on ecological variables.
+Our results show that male-biased sexual selection, but not sexual dichromatism, can explain speciation rates ($\lambda_{DR}$) in passerines.
+We also find that birds with smaller ranges have higher speciation rates, but there is no interaction between environmental variables and sexual selection.
+Our findings show that sexual dichromatism is a poor proxy for sexual selection at least a a broad-scale, and support the view that sexual selection is an evolutionary force shaping patterns of diversity amongst songbirds.
+\vspace{5mm}  \par\noindent
+  \textbf{Keywords:} Speciation rate, Comparative analysis,Plumage dichromatism, Range size, Extinction rate.
 ---
@@
 
-#Introduction
-
-Selection pressures are the main engine of speciation.
-As populations diverge they accumulate genetic and phenotypic changes that ultimately erect barriers against gene flow [@Presgraves_2010; *but see* @Shaw_2011].
-In this view, populations will speciate and adapt more readily to new and different environments if selection is strong and/or persistent.
-Hence, the rate at which speciation occurs is likely dependent on both the population's response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.*, morphology or colour: @Hugall_2012; @Rabosky_morph_2013].
-Experiments suggest that sexual selection can influence the evolution of a surprisingly diverse set of traits [@Cally_in_review], and may therefore strongly influence extant patterns of phylogenetic diversity.
+# Introduction
 
@@
 This evolutionary force has been proposed to facilitate speciation through the evolution of divergent signals associated with improved mating success [@Lande_1981; @Lande_1982].
+Experiments suggest that sexual selection can influence the evolution of a surprisingly diverse set of traits [@Cally_2019], and may therefore strongly influence extant patterns of species diversity.
 For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism are more species-rich [@Seddon_2008].
-Under this model of speciation, divergent mating signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].
+Divergent mating signals in one sex (usually males) co-evolves with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].
+
 Alongside facilitating the evolution of divergent mating signals, sexual selection can increase the extent populations diverge by enlarging the available phenotypic space for mating signals.
@@
 In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-Through these synergetic effects sexual selection and natural selection (*e.g.* environmental change over space and time) may lead to increased adaptation and speciation rates.
+Through these synergetic effects sexual selection and natural selection (*e.g.* environmental variation over space and time) may lead to increased speciation rates.
 
-As a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumage, birds are a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g.*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003].
+Birds are a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumage, thus are a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g.*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003].
 A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects and fish [@Kraaijeveld_2011].
@@
 Here, we investigate the association between sexual selection and diversification in birds while extending previous work in multiple ways.
-Firstly, many macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual dimorphism or dichromatism, but these are imperfect correlates.
-We used a very large (*n* = 5,812) dataset of sexual dichromatism [estimated from illustrated drawings; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection, capturing variation in sexual dimorphism, social polygyny and [lack of] paternal care [@Dale_2015].
-These datasets allow us to cover the great majority of bird genera, in contrast to previous work that has often focused on particular clades [*e.g.* @Seddon_2008].
-Secondly, we use the latest methods for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but less controversial tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
-We also conduct analyses that allow phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] to affect the precision of our estimates, rather than implicitly assuming that the single best-supported phylogenetic tree, or consensus tree, is correct.
-Thirdly, our analysis includes several potentially important environmental variables as predictors: range size, seasonal variation in temperature, spatial temperature variation (across a species’ range), long-term variation in temperature between now and the last interglacial (LIG) as well net primary productivity (NPP).
-By including these environmental variables in our models, we statistically account for their effects on variation in diversification rates and thereby improve our estimates of the effects of sexual selection.
-We also test whether the effects of sexual selection are contingent on the environment, as predicted by some theories [*e.g.* the theory that sexual selection is more likely to elevate adaptation in changeable relative to static environments; @Long_2012; @Connallon_Hall_2016].
+Firstly, multiple macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual dimorphism or dichromatism, but sexual selection in birds can involve many other traits, such as song, or the use of ornaments like long tail feathers, bowers, crests and displays [*e.g.*, @Uy_2000; @Pryke_2001].
+In fact, it was recently highlighted that song divergence is inversely correlated to sexual dichromatism in passerines, suggesting that there is a trade-off between signalling modalities [@Cooney_2018].
+Our study compares the role of sexual dichromatism and other sexual selection proxies in the generation of species.
+Additionally, we test the relative contribution of environmental variables to the generation of diversity, and how environment and sexual selection could interact to drive speciation processes [@Long_2012; @Connallon_Hall_2016].
+
+To test the link between different measures of sexual selection and diversification processes we use a large (*n* = 5,812) dataset of sexual dichromatism [estimated from illustrated drawings; @Dale_2015], as well as a reduced but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection, capturing variation in sexual dimorphism, social polygyny and [lack of] paternal care [@Dale_2015].
+These datasets allow us to cover the majority of bird genera.
+Additionally, we use multiple metrics for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
 \phantomsection
 
-#Materials and methods
+# Materials and methods
 
@@
 
-##Compiling data for sexual selection and environmental stress
+## Compiling data for sexual selection and environmental stress
 \phantomsection
 
-###Sexual dichromatism
+### Sexual dichromatism
 
@@
 Here we use the absolute difference between male and female plumage scores as an estimate of sexual dichromatism.
-Additionally, we used a more accurate measure of dichromatism corresponding to colour distance in avian colour space derived from spectral data [@Armenta_2008].
-These measurements include variation in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from taxonomic samples (as opposed to colour drawings).
+Additionally, we used another measure of dichromatism corresponding to colour distance in avian colour space derived from spectral data [@Armenta_2008].
+These measurements include variation in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from museum samples (as opposed to colour drawings).
 However, this spectrophotometry data is not as extensive as the RGB measures; with only 581 passerine species available for this analysis.
@@
 
-###Male-biased sexual selection
+### Male-biased sexual selection
 
 Sexual dichromatism is a widely used indicator of sexual selection in birds [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014], but only partially captures variation in the strength of sexual selection across taxa.
-Therefore, we also use an existing dataset [@Dale_2015] of male-biased sexual selection based on the first component of a phylogenetic principle component analysis (PPCA) of three characteristics associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
+Therefore, we also use an existing dataset [@Dale_2015] of male-biased sexual selection based on the first component of a phylogenetic principle component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
 This measure of male-biased sexual selection is available for only 2,465 species and is weakly correlated with the absolute values of sexual dichromatism using RGB measures ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
@@
 
-###Environmental variables
+### Environmental variables
 
@@
 
-###Generating biologically relevant predictors for environmental variation
+### Generating biologically relevant predictors for environmental variation
 
-Using the extracted environmental variables from each species range size we developed biologically meaningful predictors of environmental variation/stress relating to (*i*) seasonal climate variation, (*ii*) spatial climate variation and (*iii*) long-term climate variation.
-To obtain seasonal climate variation we used mean values of temperature seasonality (BIO4) for each range.
-To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature (BIO4) and precipitation seasonality (BIO15).
+Given that stressful environments are expected to interact with sexual selection and have a positive effect on adaptation [@Cally_2019], we used the extracted environmental variables from each species range size to develop biologically meaningful predictors of environmental variation/stress.
+Firstly we used (*i*) the average NPP in each species' range and (*v*) the log-transformed range size as potentially informative predictors of speciation rates.
+We also used three environmental predictors dervived from bioclimatic data.
+These predictors relate to seasonal climate variation, spatial climate variation and long-term climate variation.
+To obtain seasonal climate variation we used (*iii*) mean values of temperature seasonality (BIO4) for each range.
+(*iV*) To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO 4 and BIO15).
 PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
-Given that species range is a potentially informative predictor of speciation and extinction, we controlled for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}).
-To obtain long-term variation in climates for each species range we take the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
+Species range is a potentially informative predictor of speciation and extinction, so we controlled for the correlation between environmental spatial variation and range size --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
+To obtain long-term variation in climates for each species range we take (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
 Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
@@
 
-##Estimating extinction and speciation
+## Estimating extinction and speciation
 
 Phylogenetic information was obtained from www.birdtree.org [@Jetz_2012].
-We used a maximum clade credibility (MCC) tree from 2,500 samples of the posterior distribution that were subset to the passerine bird order (*n* = 5,965) as the main phylogenetic tree in our comparative analysis.
-Additionally, a random draw of 100 full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures and BAMM [@Rabosky_2014].
+We used a maximum clade credibility (MCC) tree from 2,500 samples of the posterior distribution (*n* = 5,965) as the main phylogenetic hypothesis in our comparative analysis.
+Additionally, a random draw of full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures (1000 trees) and BAMM (100 trees)[@Rabosky_2014].
 These trees used a 'Hackett backbone' [@Hackett_2008] and were constructed using a pure birth (Yule) model.
@@
 Firstly, we obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
-Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [*e.g.*, @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each more basal node having the sum of lengths down-weighted.
+Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [*e.g.*, @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each edge towards the root having the length down-weighted.
 Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification [@Belmaker_2015; @Rabosky_diversification_2018].
@@
 Importantly, these models provide tip-rate estimates of speciation and extinction rate.
-The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}; briefly, we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
-Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
+The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{ESM}; briefly, we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
+Each BAMM model (one MCC tree and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
 Each run of BAMM reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200 (an arbitrary value, above which posterior distributions can often be accurately inferred) for the key model parameters of log-likelihood and number of rate shifts (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
@@
 
-##Phylogenetic comparative analysis
+## Phylogenetic comparative analysis
 
 To test the association between speciation/extinction and sexual selection, environmental variability and their interaction, we used phylogenetic least squares (PGLS) models in the ``nlme`` package [@nlme_2018].
-We began by estimating the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using the ``corPagel`` function in the ``ape`` package [@ape_2004] on the MCC tree.
-The estimate of $\lambda$ was then fixed for model selection, which was conducted on the same MCC tree derived from 2,500 draws of the posterior distribution [@Jetz_2012].
-Model selection used either $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ tip-rate estimates from the MCC tree as the response variable.
+Firstly we conducted model selection using $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ tip-rate estimates from the same MCC tree, which was derived from 2,500 draws of the posterior distribution [@Jetz_2012].
 For $\lambda_{BAMM}$ and $\mu_{BAMM}$ we were also able to use model weights sourced from the inverse of the variance of tip-rates for a given species from the posterior distribution (*n* = 1,000).
@@
 Model selection was done in ``MuMIn`` using the ``dredge`` function [@MuMIn_2017].
-Using the terms from the top-ranking model (lowest AICc), we ran the equivalent model for each of the 100 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$, using the unique response variables and phylogenetic tree correlation structure in each model.
-This method enabled us to present model estimates for an MCC tree alongside 100 trees from the posterior distribution of trees to account for phylogenetic uncertainty.
+Using the terms from the top-ranking model (lowest AICc), we ran the equivalent model for each of the 1,000 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and each of the 100 used to derive $\lambda_{BAMM}$ and $\mu_{BAMM}$.
+In each model we used the unique response variables and phylogenetic tree correlation structure.
+Specifically, for models using tip-rate metrics ($\lambda_{DR}$, $\lambda_{ND}$) we estimated the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using the ``corPagel`` function in the ``ape`` package [@ape_2004] independently for each of the 1,000 trees/models.
+Alternitively, for models using speciation and extinction estimates derived using BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$) we found $\lambda$ was consistently estimated at 1 and hence assumed Brownian motion (using the ``corBrownian`` function) to estimate the correlation structure.
+This method enabled us to present model estimates for an MCC tree alongside 1,000/100 trees from the posterior distribution of trees to account for phylogenetic uncertainty.
 This approach was repeated on three datasets corresponding to each measure of sexual selection: dichromatism derived from RGB values of images (*n* = 5,812); dichromatism from spectrophotometry (*n* = 581) and the multivariate measure of male-biased sexual selection (*n* = 2,465).
@@
 
+#Correlations
+all.trees.join <- readRDS("data/all.trees.correlations.rds")
+Harvey.vs.us <- readRDS("data/Harvey.vs.us.rds")
+
 #Male bias SS
@@
 
-#Results
+# Results
 \phantomsection
 
-##Effects of sexual dichromatism/selection on speciation are not dependent on environmental variability
+## Variability across phylogenetic hypotheses and measures of speciation and sexual selection
+
+Estimates of speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % highest posterior density (HPD) interval across 100 trees was more than 20 times larger than the 95 % confidence interval calculated for the MCC tree.
+This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval across 1,000 trees was near-equivelant to the 95 % confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
+Mean measures of speciation across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
+Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
+Details of the BAMM results are presented in the supplementary materials.
+
+There was a strong correlation between spectral measures of colour dichromatism and RGB values ($r=0.79$; Figure S10) and a weaker correlation between sexual dichromatism and the sexual selection index ($r= 0.34$, Figure S12).
+
+\phantomsection
+## Effects of sexual dichromatism/selection on speciation are not dependent on environmental variability
 
@@
 
-##No evidence that sexual dichromatism affects speciation
+## No evidence that sexual dichromatism affects speciation
 
@@
 PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
-Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 100 random trees.
+Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 random trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 random trees for $\lambda_{BAMM}$.
 The HPD (highest posterior density) intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
@@
 
-##Male-biased sexual selection increases speciation
+## Male-biased sexual selection increases speciation
 
@@
 However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-The distribution of estimates from PGLS models on 100 random trees was similar to the estimate from the MCC tree: among the 100 tress there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
+The distribution of estimates from PGLS models on 1,000 random trees was similar to the estimate from the MCC tree: among the 1,000 tress there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
 Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}.
@@
 
-##Species with smaller ranges have increased rates of speciation
+## Species with smaller ranges have increased rates of speciation
 
@@
 This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}).
-This association was also evident across the estimates from models using the 100 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a).
+This association was also evident across the estimates from models using the 1,000 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a).
 Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed variable evidence that range size is negatively associated with speciation.
@@
 
-##Phylogenetic path analysis
+## Phylogenetic path analysis
 
@@
 
-##Extinction rate
+## Extinction rate
 
 We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
-\phantomsection
 
-##Variability across trees and methods
+![Model estimates (**a**) for the effect of log-range size and sexual dichromatism on speciation and extinction rates using PGLS analyses with the sexual dichromatism dataset (*n* = 5,812).
+(**b**) depicts the scatter plot of speciation rate ($\lambda_{DR}$) and log-range size with the model estimate presented as a dashed line.
+(**c**) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection PPCA for a subsetted dataset (*n* = 2,465).
+Similar to (**a**), (**d**) presents Model estimates for PGLS analyses using a restricted dataset with measures of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
+Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
+The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 1,000 random trees (for $\lambda_{DR}$ and $\lambda_{ND}$) or 100 random trees for $\lambda_{BAMM}$ and $\mu_{BAMM}$ can be found in the ESM.
+Density curves are based on model estimates from 1,000/100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC tree.
+For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to be able to visualise the MCC 95 % CIs.
+\label{model_results}](Figures/Estimates_and_scatter_edit.pdf)
 
-We tested our hypothesis that sexual selection is associated with speciation using several methods and across 100 trees drawn from the post-burn-in posterior [@Jetz_2012].
-We used a range of methods to overcome uncertainty in the power and precision of various tip-rate estimates as well as the variation between trees.
-We found that that the tip-rate estimates across the 100 trees were quite variable for both $\lambda_{DR}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
-Despite this, tip-rate estimates of speciation from the two methods were moderately correlated across the 100 trees ($r = 0.75$) and for the MCC tree ($r = 0.68$; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S16}).
-However, in comparison to $\lambda_{DR}$, the value of $\lambda_{BAMM}$ was the _mean_ drawn from a posterior distribution (*n* = 1,000) of BAMM generations, thus estimates of $\lambda_{BAMM}$ (and $\mu_{BAMM}$) have an added level of variation.
-To account for this variation, weights (using the inverse of the variance) were used for the PGLS models.
-From 1,000 posterior samples of the MCC BAMM run, the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figures S6} and \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
-Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) were high.
-This means that although the number of rate shifts reached convergence ($median = 59$; \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S7}), the locations of the rate shifts (*i.e.* the credible shift set) across the tree are highly heterogeneous.
-The variability in the locations for the shift configurations is a likely source of uncertainty in downstream tip-rates used as the response variable in PGLS models.
-Notably, the 95 % HPD interval for model estimates using BAMM rates from 100 trees was about 20-fold the 95 % confidence intervals of the estimate from the MCC tree; whereas for $\lambda_{DR}$ and $\lambda_{ND}$, the HPD 95 % interval width was equal to or less than the MCC 95 % CI (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S9}).
-
-![Model estimates for (**a**) PGLS analyses using sexual dichromatism (*n* = 5,812) and (**b**) PGLS analyses using measures of male-biased sexual selection (*n* = 2,465).
-Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
-The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 100 random trees can be found in the ESM.
-Density curves are based on estimates from 100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC tree.
-For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to interpret the MCC 95 % CIs.
-\label{model_results}](Figures/ModelEstimates.pdf)
-
-![Speciation rate ($\lambda_{DR}$) across all passerine birds (*n* = 5,965) with estimates of sexual dichromatism and range size available for 5,812 of them.
-Across these species there was a small but significant negative association between $\lambda_{DR}$ and log-range size but no significant association between $\lambda_{DR}$ and sexual dichromatism based on RGB.
+![Speciation rate ($\lambda_{DR}$) across all passerine birds (*n* = 5,965) with estimates of sexual dichromatism, range size available for 5,812 species and male-biased sexual selection available for 2,465 species.
+Across these species there was a small but significant negative association between $\lambda_{DR}$ and log-range size as well as a significant positive association between $\lambda_{DR}$ and male-biased sexual selection but no significant association between $\lambda_{DR}$ and sexual dichromatism based on RGB measures.
 $\lambda_{DR}$ are those from the MCC tree and images of birds are from the _Handbook of the Birds of the World_.
@@
 
-#Discussion
+# Discussion
 
@@
 The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB).
-These findings substantiate previous studies performed at smaller scales in birds, where sexual dichromatism measures were obtained using spectrophotometry [@Huang_2014] and human observers [@Cooney_2017].
-The correlation between speciation rate and the index of male-biased sexual selection (which reflects sexual size dimorphism, social polygyny, and [lack of] paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$, although the estimated sign and magnitude of the correlation was broadly similar across all speciation measures.
-Interestingly, we found a negative relationship between range size and speciation rates, at least when this rate was quantified by $\lambda_{DR}$ and $\lambda_{ND}$.
-This suggests that species with smaller range sizes have increased speciation rate.
+These findings substantiate previous studies performed at smaller scales in birds, where sexual dichromatism measures were obtained using spectrophotometry [@Huang_2014] and human observers [@Cooney_2017] and no association was found.
+The correlation between speciation rate and the index of male-biased sexual selection (which reflects sexual size dimorphism, social polygyny, and [lack of] paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$, although the estimated direction and magnitude of the correlation was broadly similar across all speciation measures.
+Interestingly, we found a consistent negative relationship between range size and speciation rates, at least when this rate was quantified by $\lambda_{DR}$ and $\lambda_{ND}$.
+This suggests that species with smaller range sizes have increased speciation rates.
 None of the bioclimatic measures of environmental variability that we investigated (*i.e.*, temperature seasonality, long-term temperature variation, and spatial temperature variation) predicted speciation rate.
@@
 The association between sexual selection and sexual dichromatism in birds is a Darwinian trope [@Darwin_1871] that allows it to be commonly used as a proxy for the strength of sexual selection in comparative studies [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
-Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015], we find that sexual dichromatism does not predict speciation rate, whereas male-biased sexual selection does.
-Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection, and that variation in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
+Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015]--$r= 0.34$-- we find that sexual dichromatism does not predict speciation rate, whereas male-biased sexual selection does.
+Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection at least at this broad scale, and that variation in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
 There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
 Firstly, sexual dichromatism can evolve for reasons other than sexual selection, such as when males and females occupy different ecological niches [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989] or experience different selective pressures in contexts other than competition for mates [@Price_2014].
-For example, in superb fairy-wrens (*Malurus cyaneus*) female colouration has evolved in response to spatial variation in predation pressure [@Medina_2017].
+For example, in superb fairy-wrens (*Malurus cyaneus*) female colouration has probably evolved in response to spatial variation in predation pressure, increasing dichromatism [@Medina_2017].
 Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit relatively weakly.
@@
 Theory suggests that sexual antagonism (often arising from increased sexual selection) may be easily purged in environments where pressures are cyclic (*e.g.* seasonality), whereas when environmental variability is directional (*e.g.* long-term climate change) sexual antagonism can indefinitely limit adaptation rates [@Connallon_Hall_2016].
-Secondly, the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
+Another possibility is that the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
 Specifically, access to dietary resources and the impacts of predation or parasitism are unaccounted for here.
 These are likely key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
-Thirdly, it is also possible that there is no effect of environmental variability on speciation rates; at least not in the species investigated here.
+Finally, it is also possible that there is no effect of environmental variability on speciation rates; at least not in the species investigated here.
 
 The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates.
-Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995].
-However, birds with smaller range sizes might often have lower dispersal ability or more specialised niches, which would promote reproductive isolation and speciation [@Birand_2012; @Claramunt_2012].
-Alternatively, reduced range size may be a consequence (rather than a cause) of speciation.
+Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995; @Castiglione_2017].
+However, the opposite pattern has also been suggested and birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
+Alternatively, reduced range size may be correlated but not causally related to speciation.
 Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives [@Rosenzweig_1995; @Weir_2011; @Price_2014].
@@
 Across islands, we expect to see a correlation between speciation and small range size because small islands often reflect newly formed environments with empty niches.
-Thus, our findings may be heavily dependent on island radiations (*e.g.* "Darwin's finches", *Geospiza* spp.).
+Thus, our findings could be heavily dependent on island radiations (*e.g.* "Darwin's finches", *Geospiza* spp.).
 One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
-Given the problems of determining causality and mode of speciation, the reason for the negative association found here remains unclear.
+
+One outcome of our analyses was the fact that different measures of speciation rates presented different results.
+This is not completely unexpected, given that each of the rates is calculated in a different way (discussed in detail in @Rabosky_diversification_2018).
+For instance, $\lambda_{DR}$ is weighted higher towards speciation events close to the tips, and allows more rate heterogeneity compared to BAMM estimates (although increasing error rate).
+The higher variation in $\lambda_{DR}$ could explain why we detected a significant pattern using this metric and not the other two.
+Moreover, it has been shown that the power of PGLS approaches is reduced compared to other alternatives such as simulations [@Harvey_essim_2017], however given the complexity of our model this was the most straight forward approach.
+This opens the possibility that the patterns detected with $\lambda_{DR}$ are also present in the other speciation metrics but there is not enough statistical power to detect them (at least for $\lambda_{BAMM}$).
+We think it is unlikely that the significant correlation between $\lambda_{DR}$ and sexual selection is spurious given the limited power of the PGLS approach, the large heterogeneity in $\lambda_{DR}$ values and the consistency in the calculated estimates across the 1,000 trees.
 
@@
 Extinction is notoriously difficult to estimate accurately from phylogenies, principally because many different combinations of speciation and extinction rates can give rise to similar patterns of diversity [*see* @Rabosky_Extinction_2016].
-Phylogenetic methods such as BAMM allow for speciation and extinction rates to be measured using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates remains a subject of debate [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
+Phylogenetic methods such as BAMM allow for speciation and extinction rates to be estimated using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates remains a subject of debate [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
 Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models and are fraught with issues of sampling bias [@Davis_2013].
@@
 To summarise, we have shown that in passerines, male-biased sexual selection, but not sexual dichromatism, predicts speciation, independent of several measures of environmental variability.
+These findings support the idea that sexual dechromatism is not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
 If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold, including for conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].
@@
 <div id="refs"></div>
-
-\newpage
-\nolinenumbers\phantomsection
-
-#Critical analysis
-
-\epigraph{[On Darwinism]...
-But it by no means follows that slight differences in the shape, pattern, or colours of the ornamental plumes are what lead a female to give the preference to one male over another; still less that all the females of a species, or the great majority of them, over a wide area of country, and for many successive generations, prefer exactly the same modification of the colour or ornament.\\
---- \large\textsc{Alfred Russel Wallace}}\phantomsection
-
-##Strengths and limitations
-
-One of the biggest unravelings of this project was the decoupling of male-biased sexual selection from sexual dichromatism in predicting speciation, remarkably this same controversy engrossed Wallace and Darwin 130 years ago [@Kottler_1980].
-The widespread use of sexual dichromatism as a proxy for sexual selection often goes unchallenged as our instinctive Darwinian minds reminisce of peacocks and birds of paradise.
-Although sexual dichromatism measurements from illustrations were always going to be an imperfect estimate of sexual selection, across 5,831 species I anticipated a signal would permeate through the noise.
-Alas --- this was not so --- and male-biased sexual selection and sexual dichromatism had different effects on speciation.
-With this in mind, the practical use of sexual dichromatism in macroevolutionary research is severely weakened.
-This finding is pertinent, as efficient proxy measurements are invaluable for comparative analyses.
-Proxies of sexual selection increase the accessibility, available sample size and transferability of research, making a study more conducive to research synthesis.
-But, as Wallace remarks, the matter is inevitably more complex, despite the irresistibility of the Darwinian argument\footnote[1]{Wallace was initially supportive of the evolution of sexual dichromatism as a result of mate choice but later took the position where natural selection on predator-prone nesting females led to females losing their plumage colouration.} [@Kottler_1980; @Caro_2017].
-On the bright side, the multivariate measure of male-biased sexual selection (sexual dimorphism, social polygyny and [lack of] paternal care) is a dataset that would benefit from being embiggened across birds.
-Future research may also benefit from expanding this multivariate measure to account for ornamentation, often overlooked by sexual dimorphism and dichromatism measures [@Gomes_2016].
-
-Like sexual selection, estimates of speciation and extinction are a point of controversy, with macroevolutionary research often plagued by methodological uncertainty.
-Originally, I had planned to continue my meta-analysis by investigating extinction rate across a large taxonomic group.
-There were signs that extinction rate could be measured from phylogenies with confidence [*see* @Pyron_2016].
-However, these hopes soon dissipated, and with them the seeming \emph{raison d'\^{e}tre} of the study.
-Extinction rate is arguably the best measure of population fitness as it tracks survival over multiple generations.
-And while a high profile study has measured extinction rate in experimental evolution [@Lumley_2015], it has been rarely measured alongside sexual selection over evolutionarily meaningful timescales [*but see* @Martins_2018].
-We included measurements of extinction rate in this study, but the findings were played-down.
-The movement away from extinction and towards speciation is because of the controversy surrounding extinction rate measures as well as phylogenetic uncertainty.
-Measuring extinction rates from the @Jetz_2012 bird phylogeny is problematic because of the nature of phylogenetic construction: a birth-only model of tree construction and incomplete genetic taxon sampling are large barriers that limit the ability to use model-based methods to detect extinction [@Beaulieu_2015].
-Specifically, a birth-only tree will likely lead to underestimates of extinction, while phylogenies relying on taxon placement can cause an imbalance in branching patterns and bias in extinction rate estimates [@Rabosky_Extinction_2016].
-Furthermore, the very detection of absent data --- which is what such methods attempt to do --- is a conceptually hard sell come peer review.
-The mechanisms and rates at which speciation and extinction occur is an active area of research.
-In a recent pre-print @Diaz_2018 used ``BAMM`` across 104 time-calibrated molecular phylogenies and found that despite variation in ecology, biogeography, and taxonomy, younger clades consistently had higher rates of speciation and extinction.
-This finding of time-dependency in rates of diversification make comparative analyses between clades challenging and calls into question countless studies.
-Similar to Wallace's view of sexual dichromatism, we are left with a view of diversification rate measures where we cannot yet account for the underlying true variation; which is vastly more complex than we currently measure.
-
-The most severe limitation for environmental variation was the coarseness of the measures.
-In total, for 5,831 species, we extracted 338,203,831 point measures of environmental variability\footnote[2]{5,831 species $\times$ 1,000 samples per range $\times$ 59 environmental variability measures including bioclimatic predictors now, in the last inter-glacial and the last glacial maximum as well as NPP.}.
-But these measures may not be as valuable as their volume implies since they do not account for variation in biotic components such as nutritional resources or predation.
-However, our methods are still improvements on using the range mid-point to obtain a measure of climate or using a single bioclimatic variable.
-While reducing the multivariate data into several biologically meaningful predictors is a conceptually challenging exercise, researchers should not feel limited by the amount of raw data that they can access\footnote[3]{This statement applies to species with documented range sizes, which varies across clades but is comprehensive for birds, mammals, amphibians and many marine groups.
-The \href{http://www.iucnredlist.org/technical-documents/spatial-data}{IUCN} has a large and accessible database for research use.}.
-In fact, previous to our study on birds the same methodology was applied to ~100 species of bovids with speciation, extinction and environmental variability estimates obtained.
-This study was prematurely halted as there was homogeneity in speciation and extinction rates across the smaller bovid phylogeny.
-However, this workflow was easily transferred over to passerine birds, with similar methods used for extracting environmental predictors.
-The use of extensive bioclimatic predictors and spatial data in macroevolutionary research is imperative as we scale-up studies.
-At very least, if bioclimatic variables do not encapsulate key ecological parameters (*e.g.*, resources, predation, habitat structure) these predictors may account for some of the noise in the dataset and may help isolate the effect of the key moderator variables of interest.
-\phantomsection
-
-##Future directions
-
-I had initially envisaged that this comparative study would allow for the reliable measure of extinction rate, use multiple taxonomic groups and robustly measure sexual selection [*see* @Anthes_2016].
-As such, many questions regarding this topic remain unanswered.
-Here, I discuss the broader directions that we may follow in order to consolidate our understanding of sexual selection and its effects on fitness and diversification.
-I will draw heavily on the discussions from a Royal Society meeting on '*Sexual selection: patterns in the history of life*'\footnote[4]{Specifically, the following two sections are spurred from dialogue with \href{https://scholar.google.com.au/citations?user=SVV7KH4AAAAJ&hl=en&oi=ao}{Dr Gene Hunt} and \href{https://scholar.google.com.au/citations?user=zG45LlAAAAAJ&hl=en}{Associate Professor Erin E.
-Saupe}} in the UK this year.
-This meeting spurred further inquiry into the topic and how interdisciplinary approaches can be harnessed to liberate ourselves from conventional comparative methodologies.
-
-Estimating extinction risk from the fossil record is difficult because many clades (*e.g.* birds) lack the stratigraphic sampling across time to statistically calculate extinction rates.
-But there are exceptions.
-Using fossil ostracods (a group of marine invertebrates), @Martins_2018 found that species with high levels of sexual dimorphism were quickly swept to extinction.
-From stratigraphic occurrence data, a capture–mark–recapture model was used to determine the extinction rate of 93 species.
-This finding challenges the results from experimental evolution studies that suggest sexual selection mitigates extinction [@Jarzebowska_2010; @Plesnar_2012; @Lumley_2015].
-The use of fossil data overcomes limitations of other comparative analyses, which by contrast, show mixed evidence in the relationship between sexual selection and extinction in birds [@Doherty_2003; @Mitra_1996; @Morrow_2003] and mammals [@Bro-Jorgensen_2014; @Morrow_2004].
-In these cases, studies have often relied on extinction threat status from the IUCN Red List and are likely biased to species under greater anthropogenic threat, culturally valued or well researched.
-In future, it would be valuable to see whether other groups of marine invertebrates have the fossil data and recognisable sexual dimorphism to conduct such a study.
-In the interim, I would be eager to see how ``BAMM`` performs on a robust ostracod phylogeny and whether the same positive association between extinction rate and sexual dimorphism could be detected from the phylogeny.
-This will be a true test of the applicability of ``BAMM`` and a key step in the effort to reliably measure extinction rates across a wider variety of species.
-
-One aspect and only foreseeable shortcoming of the ostracod study is that it did not account for environmental variability; a challenging proposal for extinct species nevertheless.
-However, the use of paleoclimate models in evolutionary biology is now a reality; with recent studies incorporating averaged values across multiple paleoclimate models as predictors of speciation [@quintero_2018].
-While the uncertainty surrounding paleoclimates was beyond the scope of the study, future studies should remain vigilant of the need to integrate key ecological processes in macroevolutionary research [*reviewed in* @Weber_2017].
-Here, we found that range size is negatively associated with speciation, yet range size encapsulates multiple aspects of ecology and evolution so it is unclear what the finding truly means.
-The role of range size in speciation may be further investigated through ancestral geographic range reconstruction [@Quintero_2015], which may be beneficial in understanding modes of speciation and the causality of range size expansions and contractions [@Losos_2003].
-Within Palaeobiology the role of range size and niche breadth on diversification patterns has been the subject of recent interest [*e.g.* @Saupe_2015; @Saupe_2016].
-Across 92 bivalve species that lived during a mid-Pliocene warming period, @Saupe_2015 used environmental niche modelling to assess the impact of the realized *vs* fundamental niche on extinction.
-Harnessing ecological niche modelling in macroevolutionary studies across extant taxa such as passerine birds is a daunting challenge but offers an *eco-evo* synthesis that may be pertinent in questions relating to extinction; especially if the taxonomic group under inquiry is of high conservation value; for which the songbirds certainly are.

### 2019-07-12 Sexual Selection and Speciation.Rmd
changes to figures

@@
   in birds
-author: Justin G. Cally^1^, Devi Stuart-Fox^1^, Luke Holman^1^, Iliana Medina Guzmán^1^
-  <br></br> <br></br> ^1^The University of Melbourne
-csl: References/AmNat.csl
+author: "Justin G. Cally, Devi Stuart-Fox, Luke Holman, Iliana Medina Guzmán"
+csl: References/evolution.csl
 output:
   html_document:
-    code_folding: hide
-    depth: 3
-    number_sections: no
-    theme: cosmo
-    toc: yes
-    toc_float: yes
+    toc: true # table of content true
+    toc_float: true # make 
+    depth: 3  # upto three depths of headings (specified by #, ## and ###)
+    number_sections: false  ## if you want number sections at each table header
+    theme: yeti # lovely fonts and colours
+    code_folding: hide # awesome buttons to show/hide the code
 subtitle: Supplementary Material
@@
 library(grid)
+library(svglite)
 library(gridExtra)
@@
 library(ggraph)
+library(ggiraphExtra)
 library(stringr)
+library(Cairo)
+library(gdtools)
+library(data.table)
+library(bindrcpp)
 #devtools::install_github("Ax3man/phylopath")
+#devtools::install_github("thomasp85/patchwork")
+#devtools::install_github('VPetukhov/ggrastr')
+library(ggrastr)
+library(patchwork)
 library(phylopath)
+if(!requireNamespace("BiocManager", quietly = TRUE)){
+    install.packages("BiocManager")
+  }
+#BiocManager::install(c("EBImage", "ggtree"))
 library("EBImage") # for images
 library("ggtree")
+library("ggnewscale")
 source("bamm_extinction/functions/check_and_fix_ultrametric.R")
@@
 
-#Compiling the data
+# Compiling the data
 
@@
 
-##Speciation Rate
+## Speciation Rate
 
@@
 
-##Sexual dichromatism and Male-biased Sexual Selection
+## Sexual dichromatism and Male-biased Sexual Selection
 
@@
 
-##Environmental predictors
+## Environmental predictors
 
@@
 
-```{r, eval = FALSE}
+```{r spatial example one, eval = FALSE}
 #Increse the iterations (defaukt is 4) so we can obtain complete samples of each range
@@
 **2)** Extracted the bioclim or other (*e.g.* NPP) data from each of the points:
-```{r, eval = FALSE}
+```{r spatial example two, eval = FALSE}
 bird.values <- lapply(bird.points, function(x) {raster::extract(bioclim_data, x)})
@@
 
-```{r, eval = FALSE}
+```{r spatial example three, eval = FALSE}
 #Obtain means per variable per species
@@
 #Now add column of species name: Same order carried through
-bird.means <- cbind.data.frame(shps.jetz, bird.means)
+bird.means <- cbind.data.frame(bird.names, bird.means)
 
@@
 
-```{r, eval = FALSE}
+```{r spatial example four, eval = FALSE}
 bird.range.size <- sapply(bird.ranges,
@@
 
-#Generating biologically relevant predictors
+# Generating biologically relevant predictors
 
@@
 
-These predictors can be obtained from the compiled datasets seen here:
-
-```{r, warning=FALSE}
+These predictors can be obtained from the compiled datasets read in here:
+
+```{r read data, warning=FALSE}
 plumage.scores <- read.csv('data/plumage_scores.csv')
@@
 complete.dataframe <- left_join(plumage.scores %>% dplyr::select(binomial, Male_plumage_score, Female_plumage_score, SDi), complete.dataframe, by = "binomial")
-```
-
-##Environmental Spatial Variability PCAs
+
+head(complete.dataframe) %>%
+  kable("html") %>% 
+  kable_styling() %>%
+  scroll_box(width = "100%", height = "500px")
+```
+
+## Environmental Spatial Variability PCAs
 
@@
 **Table S1:** Loadings for the first two PCs, from a PCA on the variation (standard error, where n ~ 1,000) of each bioclim variable except the standard deviation of temperature seasonality (sd.bioclim4) and the standard error of precipitation seasonality (sd.bioclim15), as these are already measures of variability and based on other bioclimatic variables.
-```{r}
+```{r PCA bioclim}
 restricted.data <- complete.dataframe %>% drop_na(bioclim1) #Drop species without environmental data
@@
 
-```{r, fig.width=10, fig.height=5}
+```{r PCA GAM, fig.width=10, fig.height=5}
 #run GAM of association between range size and spatial climatic variation
@@
 
-##Long term climate variability (LIG anomaly)
+## Long term climate variability (LIG anomaly)
 
@@
 
-```{r}
+```{r historical data}
 historical.variation.data <- as.data.frame(restricted.data[1])
@@
 
-```{r, fig.width=8, fig.height=5}
+```{r PCA historical, fig.width=8, fig.height=5}
 #Now we can predict the PCA results: 
@@
 
-##Correlations between environmental predictors
+## Correlations between environmental predictors
 
@@
 
-```{r, fig.width=10, fig.height=10}
+```{r model predictor covariance, fig.width=10, fig.height=10}
 #Draw plot
@@
 
-##Sexual Dimorphism
+## Sexual Dimorphism
 
@@
 
-#Analysis
-
-##Simple speciation measures: Diversification rate (DR) and Node Density (ND)
+# Analysis
+
+## Simple speciation measures: Diversification rate (DR) and Node Density (ND)
 
@@
 The functions that generate these tip rates estimates are: ``calculate.log.es``, ``calculate.nd`` and ``calculate.tb``.
-Here they are run on 100 trees plus a full MCC tree of passerine birds based on 2500 sample trees of the posterior [@Jetz_2012].
+Here they are run on 1000 trees plus a full MCC tree of passerine birds based on 2500 sample trees of the posterior [@Jetz_2012].
 This was done using the ``phanghorn`` package [@phangorn_2010].
 
-```{r}
+```{r speciation statistics}
 #Obtain DR (log(es)) estimates, by calling the calculate.dr function
@@
 
-To gain an estimate of variation between trees we can take the values and obtain a mean, variance and coefficient of variation (CV).
+<!-- To gain an estimate of variation between trees we can take the values and obtain a mean, variance and coefficient of variation (CV).
 We obtain a CV value to account for unequal variation at higher estimates.
-The CV is expressed as the ratio of the standard deviation ($\sigma$) to the mean ($\mu$) (converting to a percentage not needed):
-
-$C_V = \frac{\sigma}{\mu} \times 100$
-
-Furthermore, we can account for over-dispersion by obtaining a transformed version of CV. In such a case we can calculate CV for a log-normal distribution:
-
-${\widehat {c_{\rm {v}}}}_{\rm {ln}}={\sqrt {\mathrm {e} ^{s_{\rm {ln}}^{2}}-1}}$
-
-Given that our _ES_ value is log-transformed, the calculation of ${\widehat {c_{\rm {v}}}}_{\rm {ln}}$ is perhaps more ideal.
+The CV is expressed as the ratio of the standard deviation ($\sigma$) to the mean ($\mu$) (converting to a percentage not needed): -->
+
+<!-- $C_V = \frac{\sigma}{\mu} \times 100$ -->
+
+<!-- Furthermore, we can account for over-dispersion by obtaining a transformed version of CV. In such a case we can calculate CV for a log-normal distribution:  -->
+
+<!-- ${\widehat {c_{\rm {v}}}}_{\rm {ln}}={\sqrt {\mathrm {e} ^{s_{\rm {ln}}^{2}}-1}}$ -->
+
+<!-- Given that our _ES_ value is log-transformed, the calculation of ${\widehat {c_{\rm {v}}}}_{\rm {ln}}$ is perhaps more ideal.
 As for _ND_ we can see from the a ``qqnorm`` plot that value of ND is normally distributed, therefore the regular CV is more appropriate for this distribution.
-
-```{r, eval = FALSE}
+-->
+
+```{r speciation statistics wrangling, eval = FALSE}
 #For ND
@@
 
-```{r}
+```{r code for plot speciation rates RAW}
 dr.summary <- readRDS('data/dr.summary.rds')
 #Plot a summary of logES and TB with the DRs with their weightings
-dr.plot <- dr.summary %>% ggplot(aes(x = mean.loges, y = mean.nd, 
-                          fill = (1/(CV.nd))/sum((1/(CV.nd)))*100, 
-                      size = (1/(CV.loges))/sum((1/(CV.loges)))*100))+
+dr.plot <- dr.summary %>% ggplot(aes(x = MCC.DR, y = MCC.ND, 
+                          fill = MCC.DR, 
+                      size = MCC.ND))+
   geom_point(shape = 21, alpha = 0.5)+
   theme_minimal()+
-  geom_errorbarh(aes(xmin = mean.loges - 0.674*sqrt(var.loges), xmax = mean.loges + 0.674*sqrt(var.loges)), 
-                 size = 0.015)+
-  geom_errorbar(aes(ymin = mean.nd - 0.674*sqrt(var.nd), ymax = mean.nd + 0.674*sqrt(var.nd)), 
-                size = 0.015)+
   theme(legend.position="bottom")+
-  labs(size = 'Weight (%) [using DR]', y='Node Density [ND]', x= 'Log Diversification Rate [DR]', fill = 'Weight (%) [using ND] \n')+
-  scale_fill_distiller(guide = "colorbar", palette = "RdPu", direction = 1, breaks=c(0,0.015,.03),
-                           limits=c(0,0.03))
+  labs(size = 'Node Density [ND]', y='Node Density [ND]', x= 'Diversification Rate [DR]\n (log-equal splits)', fill = 'Diversification Rate [DR]')+
+  scale_fill_distiller(guide = "colorbar", palette = "RdPu", direction = 1)
 marginal.dr <-ggMarginal(
@@
 
-```{r, fig.width=8, fig.height=6}
+```{r plot speciation rates RAW, fig.width=8, fig.height=6}
 grid.newpage()
@@
 ```
-**Figure S4:** $\lambda_{DR}$ and $\lambda_{ND}$ estimates of speciation rate are correlated.
-Furthermore, we see similar variation across each species in the estimates across 100 trees as the inverse of the coefficient of variation (weight) is approximately equivalent across all values of ND and DR (colours and size).
-Notably, these estimates speciation rate are relatively uncertain across all species, with 50 % CIs in both axes plotted.
-
-```{r, warning=FALSE, fig.width=8, fig.height=10}
+**Figure S4:** $\lambda_{DR}$ and $\lambda_{ND}$ estimates of speciation rate are correlated for the MCC tree used in this study. Both of these response variables are normally distributed.
+
+```{r speciation rates plot two, warning=FALSE, fig.width=8, fig.height=10}
 restricted.data <- left_join(restricted.data, dr.summary, by = 'TipLabel')
@@
 
-##BAMM measures of speciation and extinction
-
-###Set up BAMM parameters
+## BAMM measures of speciation and extinction
+
+### Set up BAMM parameters
 
@@
 
-```{r, eval = FALSE, results = 'hide'}
+```{r BAMM settings, eval = FALSE, results = 'hide'}
 name.passerine.tree <- names(passerine.trees)
@@
 
-###Run analysis
+### Run analysis
 
@@
 
-```{r, eval = FALSE}
+```{r BAmm export, eval = FALSE}
 bamm.commands <- list()
@@
 
-###Read in the event data and extract tip data
+### Read in the event data and extract tip data
 
@@
 
-```{r}
+```{r BAMM inspection}
 #Read in the tree and MCMC to check for convergence 
@@
 **Table S4:** For the 100 trees that BAMM was run on effective sample size (ESS) for the two key BAMM parameters (number of evolutionary shifts and log-Likelihood) also indicates that BAMM converges with the minimum for each parameter across the 100 trees being over 200.
-```{r}
+```{r ESS}
 ESS <- readRDS('data/ESS.rds')
@@
 
-```{r}
+```{r BAMM wrangling}
 # Read in Event Data
@@
 
-```{r}
+```{r BAMM plot}
 MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')
+MCC.BAMM.df$CV.lambda <- sqrt(exp(MCC.BAMM.df$var.lambda) - 1)*100
 #Plot a summary of logES and TB with the DRs with their weightings
@@
 
-```{r, fig.width=10, fig.height=6}
+```{r BAMM plot print, fig.width=10, fig.height=6}
 grid.newpage()
@@
 
-###Analysis of BAMM results
+### Analysis of BAMM results
 
@@
 
-####Credible number of shifts
+**Credible number of shifts**
 
@@
 
-```{r}
+```{r CSS}
 css <- credibleShiftSet(MCC.BAMM.ED, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95)
@@
 
-```{r} 
+```{r CSS two} 
 round(computeBayesFactors(MCC.BAMM.mcmc, expectedNumberOfShifts=100, burnin=0.1)[,1], digits = 2)
@@
 
-```{r, fig.width = 8, fig.height = 5}
+```{r plot prior BAMM, fig.width = 8, fig.height = 5}
 plotPrior(MCC.BAMM.mcmc, expectedNumberOfShifts=100)
@@
 
-```{r, fig.width=10, fig.height=6}
+```{r Hackett Data, fig.width=10, fig.height=6}
 load('data/Hackett_split_eventsample.rda')
@@
 
+Harvey.vs.us <- full_join(MCC.BAMM.df %>% select(TipLabel, mean.lambda), Harvey.BAMM.df %>% select(TipLabel, Harvey.lambda = mean.lambda), by = "TipLabel")
+saveRDS(Harvey.vs.us, "data/Harvey.vs.us.rds")
 # BAMM.variance <- ggExtra::ggMarginal(
@@
 
+In comparison to $\lambda_{DR}$, the value of $\lambda_{BAMM}$ was the _mean_ drawn from a posterior distribution (*n* = 1,000) of BAMM generations, thus estimates of $\lambda_{BAMM}$ (and $\mu_{BAMM}$) have an added level of variation.
+To account for this variation, weights (using the inverse of the variance) were used for the PGLS models.
+From 1,000 posterior samples of the MCC BAMM run, the coefficient of variation (*CV*) for all log-rates of $\lambda_{BAMM}$ was relatively low (*mean CV* = `r round(mean(MCC.BAMM.df$CV.lambda), digits = 2)`, *median CV* = `r round(median(MCC.BAMM.df$CV.lambda), digits = 2)` ; *see* Figures S6 and Figure S8).
+Despite the convergence of the BAMM model in all runs --- where effective sample sizes of the number of shifts and log-likelihood were all greater than 200 (Table S3,Table S4) --- we found that the unique combinations of rate shifts across the large phylogenetic tree (*n* = 5,966 species) were high.
+This means that although the number of rate shifts reached convergence ($median = 59$; Figure S7), the locations of the rate shifts (*i.e.* the credible shift set) across the tree are highly heterogeneous.
+The variability in the locations for the shift configurations is a likely source of uncertainty in downstream tip-rates used as the response variable in PGLS models.
+Notably, the 95 % HPD interval for model estimates using BAMM rates from 100 trees was about 20-fold the 95 % confidence intervals of the estimate from the MCC tree; whereas for $\lambda_{DR}$ and $\lambda_{ND}$, the HPD 95 % interval width was equal to or less than the MCC 95 % CI (Table S9).
+
 <!-- ###Time-constant -->
@@
 
-##PGLS Models
+## PGLS Models
 
@@
 
-###PGLS Models on DR and ND
+### PGLS Models on DR and ND
 
@@
 
-```{r, eval = FALSE}
+```{r MCC DR and ND, eval = FALSE}
 #Prune tree
@@
 Inspect the $\lambda$ value, which can then be fixed for successive models.
-```{r, warning=FALSE}
+```{r MCC DR and ND lambda, warning=FALSE}
 MCC.DR.corPagel <- readRDS('data/MCC.DR.corPagel.rds')
@@
 
-```{r, eval = FALSE}
+```{r MCC DR ND models, eval = FALSE}
 #Run model for DR
@@
 We can be reasonably confident that interactions are unlikely to affect the pattern of speciation we see in passerine birds.
-```{r}
+```{r MCC DR ND model tables}
 dredged.DR.model <- readRDS("data/dredged.DR.model.rds")
@@
 
-```{r, eval=FALSE}
+```{r MCC ND DR top, eval=FALSE}
 #In both cases the top model is 1/2/3/4/5/6 no interaction terms. With no models within delta < 4:
@@
 
-```{r, eval = F}
-##########################################################
-####### Read in DR models and extract estimates: #########
-##########################################################
+```{r trycatch all models, eval = F}
+
+#Read in DR models and extract estimates:
 
@@
 
-##########################################################
-####### Read in ND models and extract estimates: #########
-##########################################################
+# Read in ND models and extract estimates
 
@@
 
-```{r, fig.height = 8, fig.width = 4, message = FALSE, warning = FALSE}
+```{r  DR and ND plots for model, fig.height = 8, fig.width = 4, message = FALSE, warning = FALSE}
 #Read in the 1000 tree summary df and the MCC tree model
@@
 
-##PGLS Models on BAMM estimates
+## PGLS Models on BAMM estimates
 
@@
 
-```{r, eval = FALSE}
+```{r BAMM mcc dredge, eval = FALSE}
 MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')
@@
 This is the same situation as DR/ND models (see above).
-```{r}
+```{r MCC BAMM table}
 dredged.MCC.lambda <- readRDS("data/dredged.MCC.lambda.rds")
@@
 
-```{r, eval=FALSE}
+```{r MCC BAMM top, eval=FALSE}
 #Run the top model for the MCC
@@
 In any case, we found that an unweighted model did not change qualitative results.
-```{r}
+```{r MCC BAMM AICs}
 MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
@@
 
-```{r, eval=FALSE}
+```{r MCC BAMM wrangling, eval=FALSE}
 #Read in the BAMM data for the 100 trees
@@
 
-```{r, warning=F, error = F, fig.height = 8, fig.width = 4}
+```{r BAMM plots, warning=F, error = F, fig.height = 8, fig.width = 4}
 BAMM.lambda.pgls.models <- readRDS("data/BAMM.lambda.pgls.models.rds")
@@
 
-###PGLS Model Summary
-
-```{r, fig.width=12, fig.height=9}
+### PGLS Model Summary
+
+```{r Model Summary, fig.width=12, fig.height=9}
 grid.arrange(
@@
 **Table S7:** The estimates of the MCC models plotted above are based on the following data tables.
-```{r}
+```{r Values For Models}
 MCC.DR.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates", split.table = Inf)
@@
 Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that is in the same direction (+ or -).
-```{r, warning = FALSE}
+```{r HPD, warning = FALSE}
 hpd.DR.top <- list()
@@
 Here, the large values for the estimates from BAMM tip-rates suggest that there is high variation in tip-rate estimates across trees that is not see within the DR and ND models.
-```{r}
+```{r HPD CI interval ratio}
 hpd.DR.top2 <- hpd.DR.top %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Parameter")
@@
 
-##Subsetted analysis with spectrophotometry data
+## Subsetted analysis with spectrophotometry data
 
@@
 
-```{r, warning = FALSE, fig.width=10, fig.height=6}
+```{r Armenta inspection data, warning = FALSE, fig.width=10, fig.height=6}
 #Read in Armenta data
@@
 R-squared and correlation for relationship between RGB and Spectrophotometry Data:
-```{r}
+```{r Armenta r val}
 data_frame(
@@
 Run PGLS model for the data
-```{r, eval = FALSE}
+```{r Armenta models, eval = FALSE}
 #Prune tree
@@
 
-```{r, eval = FALSE}
+```{r Armenta cluster, eval = FALSE}
 #Set up cluster
@@
 All four show the top model is one with no interactions, with $\delta AICc > 4$.
-```{r}
+```{r Armenta dredged}
 Armenta.dredged.ND.model <- readRDS("data/Armenta.dredged.ND.model.rds")
@@
 
-```{r, eval = FALSE}
+```{r Armenta top, eval = FALSE}
 #In both cases the top model is 1/2/3/4/5/6 no interaction terms. With no models within delta < 4:
@@
 
-```{r, fig.height = 8, fig.width = 4, warning=FALSE}
+```{r Armenta plots, fig.height = 8, fig.width = 4, warning=FALSE}
 
@@
 
-####################################
 #For ND
@@
 
-######################
-
 #For Lambda
@@
 
-######################
-
 #For Mu
@@
 
-```{r, fig.width=12, fig.height=9}
+```{r Armenta plot plots, fig.width=12, fig.height=9}
 grid.arrange(symmetrise_scale(Armenta.DR.plot, "x"),
@@
 **Table S11:** MCC estimates from the above plot are presented as numerical values below. 
-```{r}
+```{r Armenta moel summary}
 Armenta.MCC.DR.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates from Spec measures")
@@
 These intervals are calculated in the same way as in Table S7.
-```{r, warning = FALSE}
+```{r Armenta HPD, warning = FALSE}
 Armenta.hpd.DR.top <- list()
@@
 
-###############
 Armenta.hpd.Lambda.top <- list()
@@
 
-################
 Armenta.hpd.Mu.top <- list()
@@
 
-##Analysis using male-biased measure of sexual selection
+## Analysis using male-biased measure of sexual selection
 
@@
 
-```{r, warning = FALSE, fig.width=10, fig.height=6}
+```{r SS subset inspection, warning = FALSE, fig.width=10, fig.height=6}
 SS.subset <- plumage.scores %>% filter(Sexual_selection_ppca != "NA")
@@
 
-```{r}
+```{r SS R val}
 data_frame(
@@
 
-```{r, eval = FALSE}
+```{r SS MCC, eval = FALSE}
 #Prune tree
@@
 
-```{r, eval = FALSE}
+```{r Dredge SS, eval = FALSE}
 #Set up cluster
@@
 
-```{r}
+```{r Table for SS}
 Subset.dredged.ND.model <- readRDS("data/Subset.dredged.ND.model.rds")
@@
 
-```{r, eval = FALSE}
+```{r SS MCC Top, eval = FALSE}
 #Run model for DR
@@
 
-```{r, warning = FALSE}
+```{r SS MCC lambda, warning = FALSE}
 Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')
@@
 
-```{r, eval = F}
-
-##########################################################
-####### Read in DR models and extract estimates: #########
-##########################################################
+```{r Read in Models, eval = F}
+
+#Read in DR models and extract estimates:
 
@@
 rm(c.list)
-############################
-```
-
-```{r}
+```
+
+```{r SS plots}
 Subset.DR.pgls.summary <- readRDS("data/Subset.DR.pgls.summary.rds")
@@
 
-#########################################
-```
-
-```{r, fig.width=12, fig.height=9}
+```
+
+```{r SS plot combined, fig.width=12, fig.height=9}
 grid.arrange(symmetrise_scale(Subset.DR.plot, "x"),
@@
 **Table S14:** MCC model estimates from the above model are presented here as numberical values with 95 % CIs.
-```{r}
+```{r Table for SS plots}
 Subset.MCC.DR.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates from Male-bias SS", split.table = Inf)
@@
 These intervals are calculated in the same way as in Table S7 and Table S9.
-```{r, warning = FALSE}
+```{r SS HPD, warning = FALSE}
 Subset.hpd.DR.top <- list()
@@
 
-###############
 Subset.hpd.Lambda.top <- list()
@@
 
-################
 Subset.hpd.Mu.top <- list()
@@
 
-##Phylogenetic Path Analysis
+## Phylogenetic Path Analysis
 
@@
 
-```{r, eval = F}
+```{r Phylopath, eval = F}
 #Set rownames to match tree
@@
 
-```{r, fig.width = 6, fig.height = 4, }
+```{r phylopath model, fig.width = 6, fig.height = 4, }
 best_model <- readRDS('data/path_model.rds')
@@
 
-#Additional Figures and Tables
-
-```{r, fig.width = 8, fig.height = 6}
+# Additional Figures and Tables
+
+```{r hundy trees, fig.width = 8, fig.height = 6}
 BAMM.df <- readRDS('data/BAMM.df.rds')
@@
 
-all.trees.join <- left_join(DR.Lambda.100Trees %>% select(TipLabel, DR), BAMM.Lambda.100Trees %>% select(mean.lambda, TipLabel), by = "TipLabel")
+ND.list <- plyr::llply(nd.list, function(x) {
+  x %>% as.data.frame %>% rownames_to_column()
+  })
+ND.list <- do.call(rbind, ND.list)
+ND.Lambda.100Trees <- ggiraphExtra::summarySE(ND.list, measurevar = ".", groupvars = "rowname")
+ND.Lambda.100Trees <- ND.Lambda.100Trees %>% rename(ND.Lambda.100Trees, ND = ., TipLabel = rowname)
+
+all.trees.join <- left_join(DR.Lambda.100Trees %>% select(TipLabel, DR) %>% 
+                              left_join(ND.Lambda.100Trees %>% select(TipLabel, ND), by = "TipLabel"), 
+                            BAMM.Lambda.100Trees %>% select(mean.lambda, TipLabel), by = "TipLabel")
 
@@
 
+saveRDS(all.trees.join, "data/all.trees.correlations.rds")
 # mean(DR.Lambda.100Trees$sd)
@@
 
-```{r, echo = F}
+```{r MCC compare one, echo = F}
 MCC.BAMM.model.data <- readRDS('data/MCC.BAMM.model.data.rds')
@@
 
-```{r, fig.width = 8, fig.height = 6}
+```{r MCC compare two, fig.width = 8, fig.height = 6}
 restricted.data %>% ggplot(aes(x = MCC.DR, y = MCC.BAMM.model.data$mean.lambda, fill = MCC.DR))+
@@
 
-```{r, eval = T, echo = F, results='hide', warning=F}
+```{r some random calcs, eval = T, echo = F, results='hide', warning=F}
 MCC.BAMM.df$CV.lambda <- sqrt(exp(MCC.BAMM.df$var.lambda) - 1)*100
@@
 
-```{r, eval = FALSE, echo = F}
+```{r random calcs two, eval = FALSE, echo = F}
 #Run subset model with SDi
@@
 
-#R Session information
+# R Session information
 

### 2019-07-12 thousand_trees.R

@@ -1,427 +1,0 @@
-###############################################################################
-### Testing the Association Between Sexual Dichromatism/Sexual Selection 
-###                   and speciation using 1,000 trees       ##################
-###############################################################################
-
-##############################
-#####       Setup      #######
-##############################
-library(foreach)
-library(doParallel)
-library(nlme)
-library(phytools)
-library(ape)
-library(dplyr)
-source("functions/diversification_rate_calculation.R")
-
-trees <- list()
-
-do_one_tree <- function(tree){
-  xfjvbzdfsg
-  write.rds(zxfv)
-}
-
-slurm.apply(trees, do_one_tree)
-
-
-###Generate 1000 trees for the analysis
-
-#thousand_trees <- passerine.trees
-#t5.trees <- read.nexus('data/trees/passerine_2500.nex')
-# '%ni%' <- Negate('%in%')
-# hundred.tree.names <- names(passerine.trees)
-# 
-# for(x in names(t5.trees)){
-#       ifelse(length(thousand_trees) < 1000, 
-#              thousand_trees[[x]] <- t5.trees[[x]],
-#              break)
-#     }
-
-##Save DFs
-
-# full.data.for.model <- restricted.data %>% select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP)
-# 
-# subset.data.for.model <- SS.subset %>% select(binomial = binomial.x, TipLabel, Sexual_selection_ppca, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP)
-
-#saveRDS(thousand_trees, file = "data/trees/thousand_trees.rds")
-# saveRDS(full.data.for.model, file = "data/full.data.for.model.rds")
-# saveRDS(subset.data.for.model, file = "data/subset.data.for.model.rds")
-
-###############################################################
-###Running the loop: 
-
-##Read in the trees and data for model
-
-thousand_trees <- readRDS(file = "data/trees/thousand_trees.rds")
-
-full.data.for.model <- readRDS(file = "data/full.data.for.model.rds")
-subset.data.for.model <- readRDS(file = "data/subset.data.for.model.rds")
-
-#Set up cluster and register it with R
-
-########################################
-######### SPECIFY CORES HERE!!!! #######
-########################################
-
-
-#setup parallel backend to use many processors
-cl <- makeCluster(4) #not to overload your computer
[…]

### 2019-07-19 ComparativePlots.Rmd
Changes after Devis comments

@@
 cols <- c("#543005", "#8c510a", "#bf812d", "#dfc27d", "#c0d884",
-           brewer.pal(n = 9, name = "BuGn")[4:9])
+           brewer.pal(n = 9, name = "YlGn")[4:9])
 cols <- c('#B76643','#B76643','#B76643','#B76643','#B47339','#AC8032','#A08E33','#909B3C','#7DA64D', brewer.pal(n = 9, name = "Greens")[5:9], "#00441B", "#00441B", "#00441B")
@@
            brewer.pal(n = 9, name = "YlGn")[4:9])
-
-cols <- c(
-'#543005',
-'#8c510a',
-'#fd8d3c',
-'#feb24c',
-'#fed976',
-'#ffeda0',
-#'#f7fcb9',
-'#d9f0a3',
-'#addd8e',
-'#78c679',
-'#41ab5d',
-'#238443',
-'#006837',
-'#006837',
-'#004529',
-'#004529',
-'#004529')
-
-cols <- c(
-'#211F29',
-'#262B37',
-'#2A3945',
-'#2C4751',
-'#2D565C',
-'#2F6565',
-'#34746C',
-'#3E8470',
-'#4D9371',
-'#60A171',
-'#78B06F',
-'#92BD6D',
-'#B0C96B',
-'#D0D46B'
-)
+# 
+# cols <- c(
+# '#543005',
+# '#8c510a',
+# '#fd8d3c',
+# '#feb24c',
+# '#fed976',
+# '#ffeda0',
+# #'#f7fcb9',
+# '#d9f0a3',
+# '#addd8e',
+# '#78c679',
+# '#41ab5d',
+# '#238443',
+# '#006837',
+# '#006837',
+# '#004529',
+# '#004529',
+# '#004529')
+# 
+# cols <- c(
+# '#211F29',
+# '#262B37',
+# '#2A3945',
+# '#2C4751',
+# '#2D565C',
+# '#2F6565',
+# '#34746C',
+# '#3E8470',
+# '#4D9371',
+# '#60A171',
+# '#78B06F',
+# '#92BD6D',
+# '#B0C96B',
+# '#D0D46B'
+# )
 
@@
   geom_density_ridges_gradient(scale = 3, size = 0.05, gradient_lwd = 0.75)+
-  scale_fill_gradientn(colours = cols, guide = F, breaks = c(0.02, 0.1, 0.5, 2.5))+
-  theme_minimal()+
-  scale_x_continuous(trans = "log", breaks = c(0.02, 0.1, 0.5, 2.5))+
+  scale_fill_distiller(palette = "YlGn", direction = 1, labels = )+
+  theme_minimal()+
+  scale_x_continuous(breaks = c(0.02, 0.1, 0.5, 2.5), trans = "log")+
   scale_y_discrete(expand = c(0,0))+
   theme(axis.text.y = element_blank(),
+        legend.direction = "horizontal",
+        legend.position = "bottom",
         axis.title.y = element_blank(),
@@
   xlim(-75, NA)+
-  scale_color_distiller(palette = "YlGn", direction = 1, guide = F, trans = "log", breaks=c(0.05, 0.15, 0.5, 1.5))+
+  scale_color_distiller(palette = "YlGn", direction = 1, trans = "log", breaks=c(0.05, 0.15, 0.5, 1.5))+
   theme(legend.position = "bottom",

### 2019-07-19 Manuscript_Publication.Rmd
Changes after Devis comments

@@
   \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
-  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
+  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}}
   \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref}
@@
 Using the same spatial sampling, we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial [LIG; 120,000 - 140,000 years ago; @Otto-Bliesner_2006].
+We include estimates of climate variability during recent evolutionary history as they may be a better indicators of environmental effects on speciation than present-day  environmental variability.
 Furthermore, to estimate variability in the energy available to species, we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution.
@@
 
-### Generating biologically relevant predictors for environmental variation
+### Generating biologically relevant predictors for environmental stress
 
@@
 Firstly we used (*i*) the average NPP in each species' range and (*v*) the log-transformed range size as potentially informative predictors of speciation rates.
-We also used three environmental predictors dervived from bioclimatic data.
+We also used three environmental predictors derived from bioclimatic data.
 These predictors relate to seasonal climate variation, spatial climate variation and long-term climate variation.
@@
 To obtain long-term variation in climates for each species range we take (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
-Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
+Similar to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
 The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
@@
 Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [*e.g.*, @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each edge towards the root having the length down-weighted.
-Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification [@Belmaker_2015; @Rabosky_diversification_2018].
+Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
 Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events and $\lambda_{ND}$ is a measure of speciation across the root-to-tip path.
-These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE).
-But, based on previous simulation studies, $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
+These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE); but, based on previous simulation studies, $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
 
@@
 Specifically, for models using tip-rate metrics ($\lambda_{DR}$, $\lambda_{ND}$) we estimated the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using the ``corPagel`` function in the ``ape`` package [@ape_2004] independently for each of the 1,000 trees/models.
-Alternitively, for models using speciation and extinction estimates derived using BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$) we found $\lambda$ was consistently estimated at 1 and hence assumed Brownian motion (using the ``corBrownian`` function) to estimate the correlation structure.
+Alternatively, for models using speciation and extinction estimates derived using BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$) we found $\lambda$ was consistently estimated at 1 and hence assumed Brownian motion (using the ``corBrownian`` function) to estimate the correlation structure.
 This method enabled us to present model estimates for an MCC tree alongside 1,000/100 trees from the posterior distribution of trees to account for phylogenetic uncertainty.
@@
 
-## Variability across phylogenetic hypotheses and measures of speciation and sexual selection
+## Variability across phylogenetic hypotheses and measures of speciation
 
@@
 This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval across 1,000 trees was near-equivelant to the 95 % confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
+Given the computational requirements of BAMM, the great majority of studies only derive estimates for a single consensus tree; so the high variability in BAMM estimates among trees is notable.
 Mean measures of speciation across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
@@
 
-There was a strong correlation between spectral measures of colour dichromatism and RGB values ($r=0.79$; Figure S10) and a weaker correlation between sexual dichromatism and the sexual selection index ($r= 0.34$, Figure S12).
-
 \phantomsection
-## Effects of sexual dichromatism/selection on speciation are not dependent on environmental variability
+## Effects of sexual dichromatism/selection on speciation do not depend on environmental variability
 
@@
 
-We found a significant positive association between male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+We found a significant positive association between male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ for the MCC tree ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
 However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
@@
 Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; \href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S14}).
-Notably, environmental variability (temperature seasonality) directly affected sexual dichromatism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and the path from male-biased sexual selection to sexual dichromatism was relatively weak, ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
-Additionally, the large direct effect of temperature seasonality on range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`) suggests an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta_{indirect}$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}), given the negative association we identified between $\lambda_{DR}$ and range size in PGLS models.
+There was a modest effect of male-biased sexual selection on sexual dimorphism ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
+Additionally, temperature seasonality weakly affected sexual dimorphism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and strongly affected range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`).
+This suggests an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta_{indirect}$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}), given the negative association we identified between $\lambda_{DR}$ and range size in PGLS models.
 \phantomsection
@@
 
-![Model estimates (**a**) for the effect of log-range size and sexual dichromatism on speciation and extinction rates using PGLS analyses with the sexual dichromatism dataset (*n* = 5,812).
+![Model estimates (**a**) for the effect of log-range size and sexual dichromatism on speciation and extinction rates using PGLS analyses with the sexual dichromatism dataset (RGB values, *n* = 5,812).
 (**b**) depicts the scatter plot of speciation rate ($\lambda_{DR}$) and log-range size with the model estimate presented as a dashed line.
-(**c**) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection PPCA for a subsetted dataset (*n* = 2,465).
-Similar to (**a**), (**d**) presents Model estimates for PGLS analyses using a restricted dataset with measures of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
+(**c**) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection (*n* = 2,465).
+Similar to (**a**), (**d**) presents model estimates for PGLS analyses using a restricted dataset with measures of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
 Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
@@
 The association between sexual selection and sexual dichromatism in birds is a Darwinian trope [@Darwin_1871] that allows it to be commonly used as a proxy for the strength of sexual selection in comparative studies [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
-Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015]--$r= 0.34$-- we find that sexual dichromatism does not predict speciation rate, whereas male-biased sexual selection does.
+Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015]--$r= 0.34$-- we find that sexual dichromatism does not predict any measure of speciation rate, whereas male-biased sexual selection predicts $\lambda_{DR}$.
 Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection at least at this broad scale, and that variation in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
@@
 For example, in superb fairy-wrens (*Malurus cyaneus*) female colouration has probably evolved in response to spatial variation in predation pressure, increasing dichromatism [@Medina_2017].
-Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit relatively weakly.
+Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit weakly.
 Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
-For instance, sexual selection may promote investments in other mating signals (*e.g.* birdsong) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
+For instance, sexual selection may promote investments in other mating signals (*e.g.* song) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
 Lastly, not all plumage colouration honestly reflects mate quality.
@@
 
-In-line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least for $\lambda_{DR}$ and $\lambda_{ND}$ estimates.
+In-line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least for $\lambda_{DR}$ estimates.
 Additionally, we found that this association appears to be independent of net primary productivity and spatiotemporal variation in the environment.
@@
 Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995; @Castiglione_2017].
-However, the opposite pattern has also been suggested and birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
-Alternatively, reduced range size may be correlated but not causally related to speciation.
+However, the opposite pattern has also been suggested because birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
+In amphibians (329 genera) diversification was shown to be higher in taxa with smaller range size [@Greenberg_2017].
+Alternatively, smaller range size may be correlated but not causally related to speciation.
 Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives [@Rosenzweig_1995; @Weir_2011; @Price_2014].
@@
 
-One outcome of our analyses was the fact that different measures of speciation rates presented different results.
+One outcome of our analyses was that different measures of speciation rates presented different results.
 This is not completely unexpected, given that each of the rates is calculated in a different way (discussed in detail in @Rabosky_diversification_2018).
-For instance, $\lambda_{DR}$ is weighted higher towards speciation events close to the tips, and allows more rate heterogeneity compared to BAMM estimates (although increasing error rate).
+For instance, $\lambda_{DR}$ is weighted more towards speciation events close to the tips and allows more rate heterogeneity compared to BAMM estimates (although increasing error rate).
 The higher variation in $\lambda_{DR}$ could explain why we detected a significant pattern using this metric and not the other two.
-Moreover, it has been shown that the power of PGLS approaches is reduced compared to other alternatives such as simulations [@Harvey_essim_2017], however given the complexity of our model this was the most straight forward approach.
-This opens the possibility that the patterns detected with $\lambda_{DR}$ are also present in the other speciation metrics but there is not enough statistical power to detect them (at least for $\lambda_{BAMM}$).
-We think it is unlikely that the significant correlation between $\lambda_{DR}$ and sexual selection is spurious given the limited power of the PGLS approach, the large heterogeneity in $\lambda_{DR}$ values and the consistency in the calculated estimates across the 1,000 trees.
+We think it is unlikely that the significant correlation between $\lambda_{DR}$ and sexual selection is spurious given the variation in $\lambda_{DR}$ values and the consistency in the calculated estimates across the 1,000 trees.
 
@@
 Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models and are fraught with issues of sampling bias [@Davis_2013].
-Although extinction rates can be estimated from alternative sources, such as the fossil record, documented recent extinctions and IUCN extinction threat status, each approach has limitations.
+Although extinction rates can be estimated from alternative sources, such as the fossil record, documented recent extinctions and IUCN extinction threat status, each approach has significant limitations.
 Across the passerine bird phylogeny, we found that BAMM often produced homogeneous speciation and extinction rates for smaller clades showing few rate shifts, which might reduce our power to detect small differences in extinction rates among closely-related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
@@
 
+# Acknowledgements
+
+We would like to acknowledge Adnan Moussalli for his assistance in conducting BAMM runs across 100 phylogenetic trees. **Any others?**
+
+\newpage\phantomsection
+
 # References

### 2019-07-24 Manuscript_Publication.Rmd
Luke edits. Still working on it

@@
 
+```{r, echo=FALSE}
+MCC.DR.top <- readRDS('data/MCC.DR.top.rds')
+MCC.ND.top <- readRDS('data/MCC.ND.top.rds')
+MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
+MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')
+
+#hpds
+hpd.DR.top <- readRDS('data/hpd.DR.top.rds')
+hpd.ND.top <- readRDS('data/hpd.ND.top.rds')
+Armenta.hpd.DR.top <- readRDS('data/Armenta.hpd.DR.top.rds')
+
+#Correlations
+all.trees.join <- readRDS("data/all.trees.correlations.rds")
+Harvey.vs.us <- readRDS("data/Harvey.vs.us.rds")
+
+#Male bias SS
+Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')
+Subset.MCC.ND.top <- readRDS('data/Subset.MCC.ND.top.rds')
+Subset.MCC.Lambda.top <- readRDS('data/Subset.MCC.Lambda.top.rds')
+Subset.hpd.DR.top <- readRDS('data/Subset.hpd.DR.top.rds')
+Subset.hpd.ND.top <- readRDS('data/Subset.hpd.ND.top.rds')
+Subset.hpd.Lambda.top <- readRDS('data/Subset.hpd.Lambda.top.rds')
+MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')
+
+#path analysis
+best_model <- readRDS('data/path_model.rds')
+```
+
 \newpage\phantomsection
@@
 
-Sexual selection results from competition for matings, fertilisations, or associated resources, and is typically stronger in males than females [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008].
-This evolutionary force has been proposed to facilitate speciation through the evolution of divergent signals associated with improved mating success [@Lande_1981; @Lande_1982].
-Experiments suggest that sexual selection can influence the evolution of a surprisingly diverse set of traits [@Cally_2019], and may therefore strongly influence extant patterns of species diversity.
-For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism are more species-rich [@Seddon_2008].
-Divergent mating signals in one sex (usually males) co-evolves with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].
+Sexual selection results from competition between same-sex individuals for mates, their gametes, or associated resources, and is typically stronger in males than females [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008].
+There is long-standing interest in the relationship between sexual selection and the processes of speciation, extinction, and diversification [e.g.
+@Lande_1981; @Lande_1982].
+Experimental work suggests that sexual selection shapes the evolution of many different traits --- not only the obvious ones such as courthship songs or antlers [@Cally_2019] --- while some evolutionary comparative analyses have found that variation in sexual selection between clades predicts phylogenetic diversity.
+For instance, in antbirds (Thamnophilidae), genera with complex songs and pronounced sexual dichromatism are more species-rich [@Seddon_2008].
+ADD ONE MORE EXAMPLE IF POSSIBLE
 
-Alongside facilitating the evolution of divergent mating signals, sexual selection can increase the extent populations diverge by enlarging the available phenotypic space for mating signals.
-These predictions emerge as sexual selection may increase genetic diversity as a result of trade-offs between naturally- and sexually-selected traits or trade-offs under *intra*-locus sexual conflict between male and female trait expressions [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016].
-Additionally, sexual selection can have both positive and negative consequences for adaptation [*e.g.*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013].
-For example, sexual selection can elevate average fitness by helping to purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial ones [@Whitlock_2000], which should mitigate extinction risk and facilitate adaptation to environmental change.
-However, sexual selection might also promote extinction by selecting for male traits that have detrimental pleiotropic effects on female traits [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006], or selecting for male phenotypes with improved mating success but which harm population productivity, such as under-investing in parental care [@RN277].
+Several hypotheses predict a positive relationship across clades between the strength of sexual selection and the diversification rate (or the amount of extant phylogenetic diversity).
+For example, sexual selection is the mechanism underlying reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation (**REFERENCES**).
+Additionally, sexual selection often operates on traits that can create reproductive isolation when they differ between lineages, such as sexual signals and preferences, sperm-egg interactions, or genitalia (**Safran_2013?, MORE - @Lande_1981; @Lande_1982 ?**).
+Furthermore, sexual selection is hypothesised to increase diversification by maintaining trait combinations that would be disadvantageous under pure natural selection (**REFS - Bonduriansky_2011 & Radwan_2016?
+NB I think the journal name might be wrong for Radwan_2016, it's probably JEB**), and by helping to prevent extinction by purging deleterious mutations [@Agrawal_2001; @Siller_2001], fixing beneficial ones [@Whitlock_2000], and accelerating the rate at which populations can adapt to novel environments [@Lorch_2003; @Candolin_2008; @Cally_2019].
+Also, sexual selection is sometimes diversifying (as opposed to directional or stabilising), for example when individuals with rare phenotypes have elevated mating success (**REFS?
+- Holman et al BMC Evol Biol 2009 about Crozier's paradox is one, and likely has some refs in it**): diversifying selection promotes speciaiton by increasing the rate at which isolated populations become genetically distinct.
 
-Importantly, the effects of sexual selection on adaptation and speciation may depend on environmental conditions.
-Both theoretical and empirical studies suggest that under stressful/changing environments, sexual selection may have greater fitness benefits than under benign environments.
-Stressful environments strengthen the positive correlation between male mating success and female fitness, reducing the burden of negative pleiotropic effects [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017].
-Conversely, benign or stable environments ensure consistent selection, preferentially eroding genetic variation at sexually concordant loci (*i.e.* loci where the optimum genotype is equivalent for both sexes).
-In these stable environments, genetic variation remains disproportionately at sexually antagonistic loci; causing displacement of male and female traits from their optimum.
+Conversely, there are multiple ways in which sexual selection is hypothesised to reduce the diversification rate.
+Firstly, sexual selection can make extinction more likely by favouring traits that improve mating success but reduce population fitness.
+For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the mating success of individuals while reducing population-wide productivity [e.g.
+infanticide, harassment, and reduced parental care in mate-seeking males; @RN37; @Holman_Kokko_2013].
+Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013] **Add paper with Mikatake, Hosken and other.
+Current Biology, about selection on a male trait in beetles constraining female trait evolution even tho the male trait is male-only**.
+Secondly, sexual selection might decrease the speciation rate by ensuring that individuals with novel trait values have low reproductive success.
+
+Importantly, the relationship between sexual selection and diversification may depend on the environment.
+Some theoretical and empirical results suggest that sexual selection has a more positive effect on adaptation and population fitness in variable relative to stable environments [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017; @Cally_2019].
+The hypothesised mechanism is that stable environments lead to the erosion of genetic variation at sexually concordant loci (*i.e.* loci where the fittest allele is the same in both sexes), while sexually antagonistic loci remain comparatively genetically variable, reducing the genetic correlation between male and female fitness.
+In turn, this means that the most attractive members of the more competitive sex (typically males) produce less fit offspring of the demographically important sex (usually female), such that sexual selection is less likely to favour female-detrimental, extinction-promoting alleles in variable environments.
+Furthermore, sexual selection might be especially helpful in preventing extinction in variable environments, because (in the words of Charles Darwin), "sexual selection will give its aid to ordinary selection, by assuring to the most vigorous and best adapted males the greatest number of offspring" (DARWIN REF; see also @Agrawal_2001; @Siller_2001; @Boughman_2002; @Maan_2011; ALSO KOKKO + HEUBEL 2008 model about GxE).
+
+<!-- .; causing displacement of male and female traits from their optimum.
 It follows from these predictions that populations with stronger sexual selection would have fitness benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
@@
 Through these synergetic effects sexual selection and natural selection (*e.g.* environmental variation over space and time) may lead to increased speciation rates.
+-->
 
-Birds are a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumage, thus are a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g.*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003].
+Birds have been a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g.*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003].
+Birds are well-suited for this questions as their phylogenetic relationships are comparateively well-known, and they have diverse mating systems and conspicuous sexually-selected traits such as song and colourful plumage.
 A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects and fish [@Kraaijeveld_2011].
-However, @Kraaijeveld_2011 found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of speciation and the strength of sexual selection.
-More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) when using spectrophotometry measures on taxonomic specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012].
+However, the meta-analysis found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of speciation and the strength of sexual selection.
+More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012].
 Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores from human observers.
@@
 Here, we investigate the association between sexual selection and diversification in birds while extending previous work in multiple ways.
-Firstly, multiple macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual dimorphism or dichromatism, but sexual selection in birds can involve many other traits, such as song, or the use of ornaments like long tail feathers, bowers, crests and displays [*e.g.*, @Uy_2000; @Pryke_2001].
-In fact, it was recently highlighted that song divergence is inversely correlated to sexual dichromatism in passerines, suggesting that there is a trade-off between signalling modalities [@Cooney_2018].
-Our study compares the role of sexual dichromatism and other sexual selection proxies in the generation of species.
-Additionally, we test the relative contribution of environmental variables to the generation of diversity, and how environment and sexual selection could interact to drive speciation processes [@Long_2012; @Connallon_Hall_2016].
+Firstly, multiple macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual size dimorphism or dichromatism, but sexual selection in birds can involve many other traits, such as song, or the use of ornaments like long tail feathers, bowers, crests and displays [*e.g.*, @Uy_2000; @Pryke_2001].
+In fact, it was recently highlighted that song divergence is inversely correlated with sexual dichromatism in passerines, suggesting that a focus on one signalling modality may cause the mismeasurement sexual selection [@Cooney_2018].
+Our study compares the role of sexual dichromatism and other sexual selection proxies in the **generation of species (Really?
+Isn't the composite of speciation and extinction?)** .
+Additionally, we test the relative effect of some ecological and environmental variables **(I say "ecological" because of range size)** on diversification, and how environment and sexual selection could interact to drive speciation processes [@Long_2012; @Connallon_Hall_2016].
 
-To test the link between different measures of sexual selection and diversification processes we use a large (*n* = 5,812) dataset of sexual dichromatism [estimated from illustrated drawings; @Dale_2015], as well as a reduced but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection, capturing variation in sexual dimorphism, social polygyny and [lack of] paternal care [@Dale_2015].
+To investigate the link between different measures of sexual selection and diversification, we draw upon a large (*n* = 5,812 species) dataset of sexual dichromatism [estimated from illustrations; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that provides an index of male-biased sexual selection, which uses principal components analysis to capture (co)variation in sexual dimorphism, social polygyny and [lack of] paternal care [@Dale_2015].
 These datasets allow us to cover the majority of bird genera.
 Additionally, we use multiple metrics for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
+We also check the robustness of the results to phylogenetic uncertainty, by considering the posterior distribution of possible phylogenies rather than restricting our analysis to the single most likely phylogeny **(REF?)**.
 \phantomsection
 
+# Results
+\phantomsection
+
+## Variability across phylogenetic trees and speciation rate measures
+
+**Is it possible to move this later in the Results? It's maybe a bad idea to start with confusing, worrying results that don't pertain to our main non-null finding regarding the SS index**
+
+We first estimated speciation rates across the phylogeny, using the maximum clade credibility (MCC) tree, as well as 100 or 1000 trees from the posterior distribution of trees [@Jetz_2012].
+Estimates of speciation rates varied between phylogenetic trees, especially for rates calcualted by BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95% highest posterior density (HPD) interval across 100 trees was >20-fold larger than the 95% confidence interval calculated for the MCC tree.
+Conversely, for other **DIVERSIFICATION??** rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), the 95% HPD interval across 1,000 trees was near-identical to the 95 % confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
+Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar likelihoods is notable.
+Mean measures of speciation rate across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
+Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
+Details of the BAMM results are presented in the supplementary materials.
+
+\phantomsection
+## Effects of sexual dichromatism/selection on speciation do not depend on environmental variability
+
+**I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Intro.**
+
+We next fit PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) and compared them using AICc model selection.
+However, no interaction terms were present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
+Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
+Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
+\phantomsection
+
+## No evidence that sexual dichromatism affects speciation
+
+We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation.
+Using three different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) as the response variable, the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree.
+Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
+Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
+PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
+Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 random trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 random trees for $\lambda_{BAMM}$.
+The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
+For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+\phantomsection
+
+## Male-biased sexual selection increases speciation rate
+
+We found a significant positive association between male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ for the MCC tree ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+The distribution of estimates from PGLS models on 1,000 random trees was similar to the estimate from the MCC tree: among the 1,000 tress there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
+Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}.
+\phantomsection
+
+## Species with smaller ranges have increased rates of speciation
+
+Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation, we found a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$.
+This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}).
+This association was also evident across the estimates from models using the 1,000 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a).
+Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed equivocal evidence that range size is negatively associated with speciation.
+Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subset for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
+Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with the highest power using near-complete taxon sampling (*n* = 5,812).
+\phantomsection
+
+## Phylogenetic path analysis
+
+Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; \href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S14}).
+There was a modest effect of male-biased sexual selection on sexual dimorphism ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
+Additionally, temperature seasonality weakly affected sexual dimorphism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and strongly affected range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`).
+This suggests an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta_{indirect}$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}), given the negative association we identified between $\lambda_{DR}$ and range size in PGLS models.
+\phantomsection
+
+## Extinction rate
+
+We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+
+![Model estimates (**a**) for the effect of log-range size and sexual dichromatism on speciation and extinction rates using PGLS analyses with the sexual dichromatism dataset (RGB values, *n* = 5,812).
+(**b**) depicts the scatter plot of speciation rate ($\lambda_{DR}$) and log-range size with the model estimate presented as a dashed line.
+(**c**) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection (*n* = 2,465).
+Similar to (**a**), (**d**) presents model estimates for PGLS analyses using a restricted dataset with measures of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
+Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
+The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 1,000 random trees (for $\lambda_{DR}$ and $\lambda_{ND}$) or 100 random trees for $\lambda_{BAMM}$ and $\mu_{BAMM}$ can be found in the ESM.
+Density curves are based on model estimates from 1,000/100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC tree.
+For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to be able to visualise the MCC 95 % CIs.
+\label{model_results}](Figures/Estimates_and_scatter_edit.pdf)
+
+![Speciation rate ($\lambda_{DR}$) across all passerine birds (*n* = 5,965) with estimates of sexual dichromatism, range size available for 5,812 species and male-biased sexual selection available for 2,465 species.
+Across these species there was a small but significant negative association between $\lambda_{DR}$ and log-range size as well as a significant positive association between $\lambda_{DR}$ and male-biased sexual selection but no significant association between $\lambda_{DR}$ and sexual dichromatism based on RGB measures.
+$\lambda_{DR}$ are those from the MCC tree and images of birds are from the _Handbook of the Birds of the World_.
+Clockwise the species are: _Sporophila bouvronides_, _Geospiza magnirostris_, _Atlapetes leucopterus_, _Euphonia rufiventris_, _Euplectes franciscanus_, _Promerops gurneyi_, _Phainopepla nitens_, _Zosterops erythropleurus_, _Acrocephalus australis_, _Progne cryptoleuca_, _Mayrornis schistaceus_, _Paradisaea rubra_, _Pachycephala pectoralis_, _Malurus pulcherrimus_, _Cranioleuca curtata_, _Cercomacra manu_, _Muscisaxicola capistratus_, _Lepidothrix coeruleocapilla_.
+Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation and should not be interpreted.
+\label{Speciation_Figure}](Figures/DRSpeciationFigure.pdf)
+
+![Path analysis of evolutionary and ecological variables.
+Arrows represent direct effects with the direction of effect corresponding to colours (blue = positive, red = negative).
+The numeric values are standardised regression slopes and the asterisks indicates that the 95 % confidence intervals of this estimate do not overlap with zero.
+The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and male-biased sexual selection measures (*n* = 2,465).
+\label{Path_Analysis}](Figures/Path_Analysis.pdf)
+
+\newpage\phantomsection
+
+# Discussion
+
+We found evidence that the composite measure of male-biased sexual selection, but not the measure of sexual dichromatism, explains variation in the rate of speciation in passerine birds.
+The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB).
+These findings substantiate previous studies performed at smaller scales in birds, where sexual dichromatism measures were obtained using spectrophotometry [@Huang_2014] and human observers [@Cooney_2017] and no association was found.
+The correlation between speciation rate and the index of male-biased sexual selection (which encapsulates variation in sexual size dimorphism, social polygyny, and [lack of] paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$, although the estimated direction and magnitude of the correlation was broadly similar across all speciation measures.
+Interestingly, we found a consistent negative relationship between range size and speciation rate, at least when this rate was quantified using $\lambda_{DR}$ and $\lambda_{ND}$.
+This suggests that species with smaller range sizes have increased speciation rates.
+None of the bioclimatic measures of environmental variability that we investigated (*i.e.*, temperature seasonality, long-term temperature variation, and spatial temperature variation) predicted speciation rate.
+
+**Find a paper about Eclectus parrots to add to the paragraph below, maybe.
+Males and green and females are bright red, and I think females compete for male parental care, so SS is reversed.
+Another reason why male-biased SS might not be well-correlated with dichromatism**
+
+Although many comparative studies of birds have used sexual dichromatism as a proxy for the strength of sexual selection [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014], we found that sexual dichromatism did not predict any measure of speciation rate, while the index of the male-biased sexual selection predicted $\lambda_{DR}$.
+Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection at least at this broad scale, and that variation in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
+There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
+Firstly, sexual dichromatism can evolve for reasons other than sexual selection, such as when males and females occupy different ecological niches [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989] or experience different selective pressures in contexts other than competition for mates [@Price_2014].
+For example, in superb fairy-wrens (*Malurus cyaneus*) female colouration has probably evolved in response to spatial variation in predation pressure, increasing dichromatism [@Medina_2017].
+Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit weakly.
+Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
+For instance, sexual selection may promote investments in other mating signals (*e.g.* song) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
+Lastly, not all plumage colouration honestly reflects mate quality.
+A recent meta-analysis found that converted carotenoids (but *not* carotenoids sourced from the diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
+Given the importance of honest signaling in sexual selection and the variability in colour production mechanisms across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
+
+In-line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least for $\lambda_{DR}$ estimates.
+Additionally, we found that this association appears to be independent of net primary productivity and spatiotemporal variation in the environment.
+The lack of an effect of these environmental variables on speciation rate (either alone or in combination with sexual selection) has several possible interpretations.
+Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability under which the species is evolving.
+Specifically, speciation rates may be impacted by genetic constraints on adaptation, that vary across environments.
+Theory suggests that sexual antagonism (often arising from increased sexual selection) may be easily purged in environments where pressures are cyclic (*e.g.* seasonality), whereas when environmental variability is directional (*e.g.* long-term climate change) sexual antagonism can indefinitely limit adaptation rates [@Connallon_Hall_2016].
+Another possibility is that the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
+Specifically, access to dietary resources and the impacts of predation or parasitism are unaccounted for here.
+These are likely key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
+Finally, it is also possible that there is no effect of environmental variability on speciation rates; at least not in the species investigated here.
+
+The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates.
+Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995; @Castiglione_2017].
+However, the opposite pattern has also been suggested because birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
+In amphibians (329 genera) diversification was shown to be higher in taxa with smaller range size [@Greenberg_2017].
+Alternatively, smaller range size may be correlated but not causally related to speciation.
+Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives [@Rosenzweig_1995; @Weir_2011; @Price_2014].
+However, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
+Across islands, we expect to see a correlation between speciation and small range size because small islands often reflect newly formed environments with empty niches.
+Thus, our findings could be heavily dependent on island radiations (*e.g.* "Darwin's finches", *Geospiza* spp.).
+One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
+
+One outcome of our analyses was that different measures of speciation rates presented different results.
+This is not completely unexpected, given that each of the rates is calculated in a different way (discussed in detail in @Rabosky_diversification_2018).
+For instance, $\lambda_{DR}$ is weighted more towards speciation events close to the tips and allows more rate heterogeneity compared to BAMM estimates (although increasing error rate).
+The higher variation in $\lambda_{DR}$ could explain why we detected a significant pattern using this metric and not the other two.
+We think it is unlikely that the significant correlation between $\lambda_{DR}$ and sexual selection is spurious given the variation in $\lambda_{DR}$ values and the consistency in the calculated estimates across the 1,000 trees.
+
+In addition to speciation, sexual selection has frequently been invoked as a driver of extinction.
+Sexual selection may increase extinction risk as it often leads individuals (usually males) to invest in exaggerated mating signals as opposed to traits --- such as parental care --- that elevate offspring fitness [*reviewed in* @Kokko_Brooks_2003].
+Using the model-based approach of BAMM, we found no association between extinction rate and sexual dichromatism, male-biased sexual selection or measures of environmental variability.
+Extinction is notoriously difficult to estimate accurately from phylogenies, principally because many different combinations of speciation and extinction rates can give rise to similar patterns of diversity [*see* @Rabosky_Extinction_2016].
+Phylogenetic methods such as BAMM allow for speciation and extinction rates to be estimated using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates remains a subject of debate [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
+Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models and are fraught with issues of sampling bias [@Davis_2013].
+Although extinction rates can be estimated from alternative sources, such as the fossil record, documented recent extinctions and IUCN extinction threat status, each approach has significant limitations.
+Across the passerine bird phylogeny, we found that BAMM often produced homogeneous speciation and extinction rates for smaller clades showing few rate shifts, which might reduce our power to detect small differences in extinction rates among closely-related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
+Thus, this methodological constraint likely decreases our ability to accurately measure the correlation between metrics of sexual selection and the probability of extinction.
+
+To summarise, we have shown that in passerines, male-biased sexual selection, but not sexual dichromatism, predicts speciation, independent of several measures of environmental variability.
+These findings support the idea that sexual dechromatism is not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
+If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold, including for conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].
+Furthermore, our finding that high speciation rate is associated with smaller range size highlights the threat to the persistence of rapidly speciating lineages in a world with increased habitat loss and anthropogenic stress.
+This concern arises as the best predictor of extinction risk is range size [@Harris_2008] and the association found here implies that many newly-speciated clades have small range sizes and could thus be at greater risk of extinction.
+
+\newpage\phantomsection
 # Materials and methods
 
-We examined the effect of sexual selection on speciation and extinction rate in 97 % of passerines (*n* = 5,812 species; 58 % of all birds).
+We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds).
 Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions.
@@
 
-```{r, echo=FALSE}
-MCC.DR.top <- readRDS('data/MCC.DR.top.rds')
-MCC.ND.top <- readRDS('data/MCC.ND.top.rds')
-MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
-MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')
-
-#hpds
-hpd.DR.top <- readRDS('data/hpd.DR.top.rds')
-hpd.ND.top <- readRDS('data/hpd.ND.top.rds')
-Armenta.hpd.DR.top <- readRDS('data/Armenta.hpd.DR.top.rds')
-
-#Correlations
-all.trees.join <- readRDS("data/all.trees.correlations.rds")
-Harvey.vs.us <- readRDS("data/Harvey.vs.us.rds")
-
-#Male bias SS
-Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')
-Subset.MCC.ND.top <- readRDS('data/Subset.MCC.ND.top.rds')
-Subset.MCC.Lambda.top <- readRDS('data/Subset.MCC.Lambda.top.rds')
-Subset.hpd.DR.top <- readRDS('data/Subset.hpd.DR.top.rds')
-Subset.hpd.ND.top <- readRDS('data/Subset.hpd.ND.top.rds')
-Subset.hpd.Lambda.top <- readRDS('data/Subset.hpd.Lambda.top.rds')
-MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')
-
-#path analysis
-best_model <- readRDS('data/path_model.rds')
-```
-\phantomsection
-
-# Results
-\phantomsection
-
-## Variability across phylogenetic hypotheses and measures of speciation
-
-Estimates of speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % highest posterior density (HPD) interval across 100 trees was more than 20 times larger than the 95 % confidence interval calculated for the MCC tree.
-This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval across 1,000 trees was near-equivelant to the 95 % confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
-Given the computational requirements of BAMM, the great majority of studies only derive estimates for a single consensus tree; so the high variability in BAMM estimates among trees is notable.
-Mean measures of speciation across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
-Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
-Details of the BAMM results are presented in the supplementary materials.
-
-\phantomsection
-## Effects of sexual dichromatism/selection on speciation do not depend on environmental variability
-
-PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) were included in model selection.
-However, no interaction terms were significant or present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
-Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
-Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
-\phantomsection
-
-## No evidence that sexual dichromatism affects speciation
-
-We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation.
-Using three response measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$), the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree.
-Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
-Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 random trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 random trees for $\lambda_{BAMM}$.
-The HPD (highest posterior density) intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
-For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
-\phantomsection
-
-## Male-biased sexual selection increases speciation
-
-We found a significant positive association between male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ for the MCC tree ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-The distribution of estimates from PGLS models on 1,000 random trees was similar to the estimate from the MCC tree: among the 1,000 tress there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
-Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}.
-\phantomsection
-
-## Species with smaller ranges have increased rates of speciation
-
-Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation, we found a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$.
-This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}).
-This association was also evident across the estimates from models using the 1,000 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a).
-Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed variable evidence that range size is negatively associated with speciation.
-Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subset for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
-Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
-Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with the highest power using near-complete taxon sampling (*n* = 5,812).
-\phantomsection
-
-## Phylogenetic path analysis
-
-Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; \href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S14}).
-There was a modest effect of male-biased sexual selection on sexual dimorphism ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
-Additionally, temperature seasonality weakly affected sexual dimorphism ($\beta$ = `r round(best_model[["coef"]][1,4], digits = 2)`) and strongly affected range size ($\beta$ = `r round(best_model[["coef"]][1,3], digits = 2)`).
-This suggests an indirect effect of temperature seasonality on $\lambda_{DR}$ ($\beta_{indirect}$ = `r round(best_model[["coef"]][1,3]*best_model[["coef"]][3,5], digits = 2)`; \autoref{Path_Analysis}), given the negative association we identified between $\lambda_{DR}$ and range size in PGLS models.
-\phantomsection
-
-## Extinction rate
-
-We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
-
-![Model estimates (**a**) for the effect of log-range size and sexual dichromatism on speciation and extinction rates using PGLS analyses with the sexual dichromatism dataset (RGB values, *n* = 5,812).
-(**b**) depicts the scatter plot of speciation rate ($\lambda_{DR}$) and log-range size with the model estimate presented as a dashed line.
-(**c**) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection (*n* = 2,465).
-Similar to (**a**), (**d**) presents model estimates for PGLS analyses using a restricted dataset with measures of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
-Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
-The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 1,000 random trees (for $\lambda_{DR}$ and $\lambda_{ND}$) or 100 random trees for $\lambda_{BAMM}$ and $\mu_{BAMM}$ can be found in the ESM.
-Density curves are based on model estimates from 1,000/100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC tree.
-For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to be able to visualise the MCC 95 % CIs.
-\label{model_results}](Figures/Estimates_and_scatter_edit.pdf)
-
-![Speciation rate ($\lambda_{DR}$) across all passerine birds (*n* = 5,965) with estimates of sexual dichromatism, range size available for 5,812 species and male-biased sexual selection available for 2,465 species.
-Across these species there was a small but significant negative association between $\lambda_{DR}$ and log-range size as well as a significant positive association between $\lambda_{DR}$ and male-biased sexual selection but no significant association between $\lambda_{DR}$ and sexual dichromatism based on RGB measures.
-$\lambda_{DR}$ are those from the MCC tree and images of birds are from the _Handbook of the Birds of the World_.
-Clockwise the species are: _Sporophila bouvronides_, _Geospiza magnirostris_, _Atlapetes leucopterus_, _Euphonia rufiventris_, _Euplectes franciscanus_, _Promerops gurneyi_, _Phainopepla nitens_, _Zosterops erythropleurus_, _Acrocephalus australis_, _Progne cryptoleuca_, _Mayrornis schistaceus_, _Paradisaea rubra_, _Pachycephala pectoralis_, _Malurus pulcherrimus_, _Cranioleuca curtata_, _Cercomacra manu_, _Muscisaxicola capistratus_, _Lepidothrix coeruleocapilla_.
-Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation and should not be interpreted.
-\label{Speciation_Figure}](Figures/DRSpeciationFigure.pdf)
-
-![Path analysis of evolutionary and ecological variables.
-Arrows represent direct effects with the direction of effect corresponding to colours (blue = positive, red = negative).
-The numeric values are standardised regression slopes and the asterisks indicates that the 95 % confidence intervals of this estimate do not overlap with zero.
-The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and male-biased sexual selection measures (*n* = 2,465).
-\label{Path_Analysis}](Figures/Path_Analysis.pdf)
-
-\newpage\phantomsection
-
-# Discussion
-
-We found evidence that the composite measure of male-biased sexual selection, but not the measure of sexual dichromatism, explains variation in the rate of speciation in passerine birds.
-The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB).
-These findings substantiate previous studies performed at smaller scales in birds, where sexual dichromatism measures were obtained using spectrophotometry [@Huang_2014] and human observers [@Cooney_2017] and no association was found.
-The correlation between speciation rate and the index of male-biased sexual selection (which reflects sexual size dimorphism, social polygyny, and [lack of] paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$, although the estimated direction and magnitude of the correlation was broadly similar across all speciation measures.
-Interestingly, we found a consistent negative relationship between range size and speciation rates, at least when this rate was quantified by $\lambda_{DR}$ and $\lambda_{ND}$.
-This suggests that species with smaller range sizes have increased speciation rates.
-None of the bioclimatic measures of environmental variability that we investigated (*i.e.*, temperature seasonality, long-term temperature variation, and spatial temperature variation) predicted speciation rate.
-
-The association between sexual selection and sexual dichromatism in birds is a Darwinian trope [@Darwin_1871] that allows it to be commonly used as a proxy for the strength of sexual selection in comparative studies [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
-Despite the association between sexual selection and sexual dichromatism in passerine birds [@Dale_2015]--$r= 0.34$-- we find that sexual dichromatism does not predict any measure of speciation rate, whereas male-biased sexual selection predicts $\lambda_{DR}$.
-Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection at least at this broad scale, and that variation in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
-There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
-Firstly, sexual dichromatism can evolve for reasons other than sexual selection, such as when males and females occupy different ecological niches [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989] or experience different selective pressures in contexts other than competition for mates [@Price_2014].
-For example, in superb fairy-wrens (*Malurus cyaneus*) female colouration has probably evolved in response to spatial variation in predation pressure, increasing dichromatism [@Medina_2017].
-Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit weakly.
-Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
-For instance, sexual selection may promote investments in other mating signals (*e.g.* song) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
-Lastly, not all plumage colouration honestly reflects mate quality.
-A recent meta-analysis found that converted carotenoids (but *not* carotenoids sourced from the diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
-Given the importance of honest signaling in sexual selection and the variability in colour production mechanisms across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
-
-In-line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least for $\lambda_{DR}$ estimates.
-Additionally, we found that this association appears to be independent of net primary productivity and spatiotemporal variation in the environment.
-The lack of an effect of these environmental variables on speciation rate (either alone or in combination with sexual selection) has several possible interpretations.
-Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability under which the species is evolving.
-Specifically, speciation rates may be impacted by genetic constraints on adaptation, that vary across environments.
-Theory suggests that sexual antagonism (often arising from increased sexual selection) may be easily purged in environments where pressures are cyclic (*e.g.* seasonality), whereas when environmental variability is directional (*e.g.* long-term climate change) sexual antagonism can indefinitely limit adaptation rates [@Connallon_Hall_2016].
-Another possibility is that the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
-Specifically, access to dietary resources and the impacts of predation or parasitism are unaccounted for here.
-These are likely key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
-Finally, it is also possible that there is no effect of environmental variability on speciation rates; at least not in the species investigated here.
-
-The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates.
-Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995; @Castiglione_2017].
-However, the opposite pattern has also been suggested because birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
-In amphibians (329 genera) diversification was shown to be higher in taxa with smaller range size [@Greenberg_2017].
-Alternatively, smaller range size may be correlated but not causally related to speciation.
-Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives [@Rosenzweig_1995; @Weir_2011; @Price_2014].
-However, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
-Across islands, we expect to see a correlation between speciation and small range size because small islands often reflect newly formed environments with empty niches.
-Thus, our findings could be heavily dependent on island radiations (*e.g.* "Darwin's finches", *Geospiza* spp.).
-One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
-
-One outcome of our analyses was that different measures of speciation rates presented different results.
-This is not completely unexpected, given that each of the rates is calculated in a different way (discussed in detail in @Rabosky_diversification_2018).
-For instance, $\lambda_{DR}$ is weighted more towards speciation events close to the tips and allows more rate heterogeneity compared to BAMM estimates (although increasing error rate).
-The higher variation in $\lambda_{DR}$ could explain why we detected a significant pattern using this metric and not the other two.
-We think it is unlikely that the significant correlation between $\lambda_{DR}$ and sexual selection is spurious given the variation in $\lambda_{DR}$ values and the consistency in the calculated estimates across the 1,000 trees.
-
-In addition to speciation, sexual selection has frequently been invoked as a driver of extinction.
-Sexual selection may increase extinction risk as it often leads individuals (usually males) to invest in exaggerated mating signals as opposed to traits --- such as parental care --- that elevate offspring fitness [*reviewed in* @Kokko_Brooks_2003].
-Using the model-based approach of BAMM, we found no association between extinction rate and sexual dichromatism, male-biased sexual selection or measures of environmental variability.
-Extinction is notoriously difficult to estimate accurately from phylogenies, principally because many different combinations of speciation and extinction rates can give rise to similar patterns of diversity [*see* @Rabosky_Extinction_2016].
-Phylogenetic methods such as BAMM allow for speciation and extinction rates to be estimated using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates remains a subject of debate [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
-Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models and are fraught with issues of sampling bias [@Davis_2013].
-Although extinction rates can be estimated from alternative sources, such as the fossil record, documented recent extinctions and IUCN extinction threat status, each approach has significant limitations.
-Across the passerine bird phylogeny, we found that BAMM often produced homogeneous speciation and extinction rates for smaller clades showing few rate shifts, which might reduce our power to detect small differences in extinction rates among closely-related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
-Thus, this methodological constraint likely decreases our ability to accurately measure the correlation between metrics of sexual selection and the probability of extinction.
-
-To summarise, we have shown that in passerines, male-biased sexual selection, but not sexual dichromatism, predicts speciation, independent of several measures of environmental variability.
-These findings support the idea that sexual dechromatism is not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
-If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold, including for conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].
-Furthermore, our finding that high speciation rate is associated with smaller range size highlights the threat to the persistence of rapidly speciating lineages in a world with increased habitat loss and anthropogenic stress.
-This concern arises as the best predictor of extinction risk is range size [@Harris_2008] and the association found here implies that many newly-speciated clades have small range sizes and could thus be at greater risk of extinction.
-
 \newpage\phantomsection

### 2019-07-25 Manuscript_Publication.Rmd
abstract

--- 2019-07-19 Manuscript_Publication.Rmd
@@
 subparagraph: yes
-abstract: Sexual selection can shape phylogenetic diversity by affecting speciation or extinction rates.
-The predicted effects on diversity are often contradictory; for example, sexual selection might promote speciation by creating reproductive isolation, or hinder it by selecting for traits that facilitate gene flow.
-Here, we investigate the relationship between sexual selection and diversification in passerine birds.
-Multiple studies use sexual dichromatism as a proxy for sexual selection, however,  this relationship can be weak in many clades.
-Thus, we use two measures of sexual selection --- sexual dichromatism and a multivariate measure of male-biased sexual selection --- to test the link between sexual selection and diversification.
-We also test whether the effect of sexual selection on diversification is contingent on ecological variables.
-Our results show that male-biased sexual selection, but not sexual dichromatism, can explain speciation rates ($\lambda_{DR}$) in passerines.
+abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates.
+However, many of the hypothesised effects of sexual selection are contradictory, and the net effect of sexual selection on diversification is unclear.
+Here, we investigate the relationship between sexual selection and diversification in passerine birds, using two measures of sexual selection --- sexual dichromatism, plus an index that captures variance in sexual size dimorphism, polygyny and male parental care.
+We also test whether the relationship between sexual selection on diversification is contingent on ecological variables, as some theories predict.
+Our results show that male-biased sexual selection, but not sexual dichromatism, predicts speciation rates ($\lambda_{DR}$) in passerines.
 We also find that birds with smaller ranges have higher speciation rates, but there is no interaction between environmental variables and sexual selection.
-Our findings show that sexual dichromatism is a poor proxy for sexual selection at least a a broad-scale, and support the view that sexual selection is an evolutionary force shaping patterns of diversity amongst songbirds.
+Our findings support the view that sexual selection is a potent evolutionary force that has shaped diversification in songbirds.
 \vspace{5mm}  \par\noindent
-  \textbf{Keywords:} Speciation rate, Comparative analysis,Plumage dichromatism, Range size, Extinction rate.
+  \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage dichromatism, Range size, Extinction rate.
 ---

### 2019-07-26 Manuscript_Publication.Rmd
More writing

@@
 abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates.
-However, many of the hypothesised effects of sexual selection are contradictory, and the net effect of sexual selection on diversification is unclear.
-Here, we investigate the relationship between sexual selection and diversification in passerine birds, using two measures of sexual selection --- sexual dichromatism, plus an index that captures variance in sexual size dimorphism, polygyny and male parental care.
-We also test whether the relationship between sexual selection on diversification is contingent on ecological variables, as some theories predict.
-Our results show that male-biased sexual selection, but not sexual dichromatism, predicts speciation rates ($\lambda_{DR}$) in passerines.
-We also find that birds with smaller ranges have higher speciation rates, but there is no interaction between environmental variables and sexual selection.
+However, the net effect of sexual selection on diversification is hard to predict _a priori_, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes.
+Here, we test whether the strength of sexual selection predicts variation in speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera).
+We tested for associations between two measures of sexual selection (sexual dichromatism, plus an index that captures variance in sexual size dimorphism, polygyny, and male parental care), several environmental variables, and multiple measures of speciation and extinction rates.
+Our results show that male-biased sexual selection, but not sexual dichromatism, predicts speciation rates ($\lambda_{DR}$) in passerines, and found no evidence that this relationship varies with the environment.
+We also found a strong negative relationship between range size and speciation rate.
+There was no correlation between sexual selection and extinction rate **XXXXXX**.
 Our findings support the view that sexual selection is a potent evolutionary force that has shaped diversification in songbirds.
@@
 
+<!-- Luke general comments:
+In a few places you say things like "1000 random trees".
+To me this means trees with a random topology (which are sometimes used as some sort of null model), but I think you actually mean 1000 trees randomly sampled (without replacement) from the posterior of trees in Jetz et al.
+Is that right?
+If so, probably best to search for the word "random" and remove/reword it.
+-->
+
 # Introduction
 
-Sexual selection results from competition between same-sex individuals for mates, their gametes, or associated resources, and is typically stronger in males than females [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008].
-There is long-standing interest in the relationship between sexual selection and the processes of speciation, extinction, and diversification [e.g.
-@Lande_1981; @Lande_1982].
-Experimental work suggests that sexual selection shapes the evolution of many different traits --- not only the obvious ones such as courthship songs or antlers [@Cally_2019] --- while some evolutionary comparative analyses have found that variation in sexual selection between clades predicts phylogenetic diversity.
-For instance, in antbirds (Thamnophilidae), genera with complex songs and pronounced sexual dichromatism are more species-rich [@Seddon_2008].
-ADD ONE MORE EXAMPLE IF POSSIBLE
+Sexual selection results from competition between same-sex individuals for mates, their gametes, or associated resources, and in most species it predominently operates on males [@Kokko_Jennions2008; @Fromhage_2016; @Janicke_2016].
+There is long-standing interest in the relationship between sexual selection and the processes of speciation, extinction, and phylogenetic diversification [e.g.
+@Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018].
+Recent empirical and methodological advances, such as growing evidence that sexual selection can profoundly affect many traits [e.g.
+life history, body mass, and immunocompetence; @Cally_2019], larger and more accurate phylogenies [@Jetz_2012], and new phylogenetic methods (**REFERENCES**), present new opportunities to test whether and how sexual selection drives diversification.
 
-Several hypotheses predict a positive relationship across clades between the strength of sexual selection and the diversification rate (or the amount of extant phylogenetic diversity).
-For example, sexual selection is the mechanism underlying reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation (**REFERENCES**).
-Additionally, sexual selection often operates on traits that can create reproductive isolation when they differ between lineages, such as sexual signals and preferences, sperm-egg interactions, or genitalia (**Safran_2013?, MORE - @Lande_1981; @Lande_1982 ?**).
-Furthermore, sexual selection is hypothesised to increase diversification by maintaining trait combinations that would be disadvantageous under pure natural selection (**REFS - Bonduriansky_2011 & Radwan_2016?
-NB I think the journal name might be wrong for Radwan_2016, it's probably JEB**), and by helping to prevent extinction by purging deleterious mutations [@Agrawal_2001; @Siller_2001], fixing beneficial ones [@Whitlock_2000], and accelerating the rate at which populations can adapt to novel environments [@Lorch_2003; @Candolin_2008; @Cally_2019].
-Also, sexual selection is sometimes diversifying (as opposed to directional or stabilising), for example when individuals with rare phenotypes have elevated mating success (**REFS?
-- Holman et al BMC Evol Biol 2009 about Crozier's paradox is one, and likely has some refs in it**): diversifying selection promotes speciaiton by increasing the rate at which isolated populations become genetically distinct.
+Several hypotheses predict a positive relationship across species between the strength of sexual selection and the rate of speciation and/or extinction.
+For example, sexual selection is involved in reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation (**REF**).
+Additionally, sexual selection often operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia].
+Furthermore, sexual selection has been hypothesised to increase diversification by maintaining trait combinations that would be selected out under pure natural selection [@Bonduriansky_2011; @Radwan_2016], and by helping to prevent extinction by purging deleterious mutations [@Whitlock_Agrawal_2009], fixing beneficial ones [@Whitlock_2000], and accelerating adapation in environments [@Lorch_2003; @Candolin_2008; @Cally_2019].
+Also, sexual selection sometimes favours individuals with rare phenotypes [@holman2013crozier], providing a source of diversifying selection that might increase the rate at which isolated populations diverge genetically.
 
-Conversely, there are multiple ways in which sexual selection is hypothesised to reduce the diversification rate.
-Firstly, sexual selection can make extinction more likely by favouring traits that improve mating success but reduce population fitness.
-For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the mating success of individuals while reducing population-wide productivity [e.g.
-infanticide, harassment, and reduced parental care in mate-seeking males; @RN37; @Holman_Kokko_2013].
-Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013] **Add paper with Mikatake, Hosken and other.
-Current Biology, about selection on a male trait in beetles constraining female trait evolution even tho the male trait is male-only**.
-Secondly, sexual selection might decrease the speciation rate by ensuring that individuals with novel trait values have low reproductive success.
+Conversely, there are multiple ways in which sexual selection might hinder speciation or make extinction more likely.
+Firstly, sexual selection might promote extinction by favouring traits that improve mating success but reduce population fitness.
+For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the reproductive success of individuals while reducing population-wide productivity [e.g.
+infanticide, harassment, and reduced parental care by mate-seeking males; @RN37; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016].
+Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013].
+Secondly, sexual selection might promote evolutionary stasis by ensuring that individuals with novel trait values have low mating success.
 
-Importantly, the relationship between sexual selection and diversification may depend on the environment.
-Some theoretical and empirical results suggest that sexual selection has a more positive effect on adaptation and population fitness in variable relative to stable environments [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017; @Cally_2019].
-The hypothesised mechanism is that stable environments lead to the erosion of genetic variation at sexually concordant loci (*i.e.* loci where the fittest allele is the same in both sexes), while sexually antagonistic loci remain comparatively genetically variable, reducing the genetic correlation between male and female fitness.
-In turn, this means that the most attractive members of the more competitive sex (typically males) produce less fit offspring of the demographically important sex (usually female), such that sexual selection is less likely to favour female-detrimental, extinction-promoting alleles in variable environments.
-Furthermore, sexual selection might be especially helpful in preventing extinction in variable environments, because (in the words of Charles Darwin), "sexual selection will give its aid to ordinary selection, by assuring to the most vigorous and best adapted males the greatest number of offspring" (DARWIN REF; see also @Agrawal_2001; @Siller_2001; @Boughman_2002; @Maan_2011; ALSO KOKKO + HEUBEL 2008 model about GxE).
+Importantly, the relationship between sexual selection and diversification might be contingent on the environment.
+Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [e.g.
+@Long_2012; @Connallon_Hall_2016], though the empirical evidence for this theory is mixed and largely limited to insects [@Cally_2019].
+The mechanism underpinning this prediction is that the environment and the selective history of the population should influence the relative amounts of genetic variation at sexually concordant loci (*i.e.* loci where the same allele is fittest for both sexes) and sexually antagonistic loci (where different alleles are favoured in each sex).
+In stable environments, variation is thought to be preferentially depleted at sexually concordant loci, leading to stronger gender load and reduced net benefits of sexual selection [e.g.
+@Connallon_Hall_2016], and potentially to a more negative relationship between sexual selection and diversification rates across species.
+Conversely, sexual selection might be especially important at promoting local adaptation when the environment displays variability at the relevant spatial and temporal scales [e.g.
+@Boughman_2002; @Lorch_2003; @Maan_2011; @kokko2008condition]
 
@@
 
-Birds have been a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g.*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003].
-Birds are well-suited for this questions as their phylogenetic relationships are comparateively well-known, and they have diverse mating systems and conspicuous sexually-selected traits such as song and colourful plumage.
-A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects and fish [@Kraaijeveld_2011].
-However, the meta-analysis found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of speciation and the strength of sexual selection.
+Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [*e.g.* @Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014].
+Birds are well-suited for this purpose due to their diverse and well-characterised mating systems and sexually-selected traits, and because their phylogenetic relationships are comparatively well-known.
+A 2011 meta-analysis covering 20 primary studies of birds and other taxa found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects or fish [@Kraaijeveld_2011].
+However, there was large variation in effect size estimates across the 20 studies, likely reflecting differences in methodology, such as metrics used to characterise speciation and sexual selection, in addition to true biological differences.
 More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012].
-Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores from human observers.
+Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.
 
-Here, we investigate the association between sexual selection and diversification in birds while extending previous work in multiple ways.
-Firstly, multiple macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual size dimorphism or dichromatism, but sexual selection in birds can involve many other traits, such as song, or the use of ornaments like long tail feathers, bowers, crests and displays [*e.g.*, @Uy_2000; @Pryke_2001].
-In fact, it was recently highlighted that song divergence is inversely correlated with sexual dichromatism in passerines, suggesting that a focus on one signalling modality may cause the mismeasurement sexual selection [@Cooney_2018].
-Our study compares the role of sexual dichromatism and other sexual selection proxies in the **generation of species (Really?
-Isn't the composite of speciation and extinction?)** .
-Additionally, we test the relative effect of some ecological and environmental variables **(I say "ecological" because of range size)** on diversification, and how environment and sexual selection could interact to drive speciation processes [@Long_2012; @Connallon_Hall_2016].
-
-To investigate the link between different measures of sexual selection and diversification, we draw upon a large (*n* = 5,812 species) dataset of sexual dichromatism [estimated from illustrations; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that provides an index of male-biased sexual selection, which uses principal components analysis to capture (co)variation in sexual dimorphism, social polygyny and [lack of] paternal care [@Dale_2015].
-These datasets allow us to cover the majority of bird genera.
-Additionally, we use multiple metrics for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
-We also check the robustness of the results to phylogenetic uncertainty, by considering the posterior distribution of possible phylogenies rather than restricting our analysis to the single most likely phylogeny **(REF?)**.
+Here, we investigate the association between sexual selection and diversification in birds while building upon previous approaches in multiple ways.
+We use two measures of the strength of sexual selection: sexual dichromatism [*n* = 5,812 species; estimated from illustrations by @Dale_2015], as well as an index of male-biased sexual selection [*n* = 2,465 species; derived via principal components analysis by @Dale_2015], which captures (co)variation in sexual size dimorphism, social polygyny and paternal care.
+Most **OR ALL??** earlier macroevolutionary studies of this topic only quantified sexual selection using dichromatism, which is problematic since dichromatism does not always signal the presence of strong sexual selection and _vice versa_ [@Dale_2015].
+For example, male and female dunnocks (_Prunella modularis_) are similarly coloured yet sexual selection appears to be strong [@davies1986re].
+Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018].
+Additionally, our analysis includes multiple ecological and environmental variables on diversification, allowing us to adjust for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016].
+We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
+Finally, we incorporate phylogenetic uncertainty into our analyses, improving over earlier work which assumed the statistically best-fitting phylogeny is correct.
 \phantomsection
@@
 
-**Is it possible to move this later in the Results? It's maybe a bad idea to start with confusing, worrying results that don't pertain to our main non-null finding regarding the SS index**
+**[Is it possible to move this later in the Results?
+It's possibly not ideal to start by highlighting the dependence of our conclusions on the methods we choose, and not mentioning our main non-null finding regarding the SS index.
+Having said that I can see a case for leaving it here too.
+- Luke]**
 
-We first estimated speciation rates across the phylogeny, using the maximum clade credibility (MCC) tree, as well as 100 or 1000 trees from the posterior distribution of trees [@Jetz_2012].
-Estimates of speciation rates varied between phylogenetic trees, especially for rates calcualted by BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95% highest posterior density (HPD) interval across 100 trees was >20-fold larger than the 95% confidence interval calculated for the MCC tree.
-Conversely, for other **DIVERSIFICATION??** rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), the 95% HPD interval across 1,000 trees was near-identical to the 95 % confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
-Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar likelihoods is notable.
+We first estimated speciation rates across the phylogeny, using the maximum clade credibility (MCC) tree, as well as 100 or 1000 trees sampled from the posterior distribution of trees in @Jetz_2012.
+Estimates of speciation rates varied between phylogenetic trees, especially for rates calculated by BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95% highest posterior density (HPD) interval across 100 trees was >20-fold larger than the **95% confidence interval [SHOULD THIS BE HPD TOO?]** calculated for the MCC tree.
+Conversely, for our estimates of diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$), the 95% HPD interval across 1,000 trees was near-identical to the 95% confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
+Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar statistical support is notable.
 Mean measures of speciation rate across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
 Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
-Details of the BAMM results are presented in the supplementary materials.
+Full details of the BAMM results are presented as supplementary materials.
 
 \phantomsection
-## Effects of sexual dichromatism/selection on speciation do not depend on environmental variability
+## No evidence that the relationship between sexual selection and speciation depends on environmental variability
 
-**I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Intro.**
+**[I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Abstract and Intro. - Luke]**
 
-We next fit PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) and compared them using AICc model selection.
+We next fit Phylogenetic Generalized Least Squares (PGLS) models with **2-way [CORRECT?]** interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) and compared them using AICc model selection.
 However, no interaction terms were present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
@@
 PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
-Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 random trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 random trees for $\lambda_{BAMM}$.
+Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 trees for $\lambda_{BAMM}$.
 The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
@@
 However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-The distribution of estimates from PGLS models on 1,000 random trees was similar to the estimate from the MCC tree: among the 1,000 tress there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
+The distribution of estimates from PGLS models on 1,000 trees was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
 Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}.
@@
 The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 1,000 random trees (for $\lambda_{DR}$ and $\lambda_{ND}$) or 100 random trees for $\lambda_{BAMM}$ and $\mu_{BAMM}$ can be found in the ESM.
-Density curves are based on model estimates from 1,000/100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC tree.
-For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to be able to visualise the MCC 95 % CIs.
+Density curves are based on model estimates from 1,000/100 trees and the circle below with error bars is the estimate and 95% CIs from the MCC tree.
+For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to be able to visualise the MCC 95% CIs.
 \label{model_results}](Figures/Estimates_and_scatter_edit.pdf)
@@
 Arrows represent direct effects with the direction of effect corresponding to colours (blue = positive, red = negative).
-The numeric values are standardised regression slopes and the asterisks indicates that the 95 % confidence intervals of this estimate do not overlap with zero.
+The numeric values are standardised regression slopes and the asterisks indicates that the 95% confidence intervals of this estimate do not overlap with zero.
 The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and male-biased sexual selection measures (*n* = 2,465).
@@
 
-We found evidence that the composite measure of male-biased sexual selection, but not the measure of sexual dichromatism, explains variation in the rate of speciation in passerine birds.
-The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB).
-These findings substantiate previous studies performed at smaller scales in birds, where sexual dichromatism measures were obtained using spectrophotometry [@Huang_2014] and human observers [@Cooney_2017] and no association was found.
-The correlation between speciation rate and the index of male-biased sexual selection (which encapsulates variation in sexual size dimorphism, social polygyny, and [lack of] paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$, although the estimated direction and magnitude of the correlation was broadly similar across all speciation measures.
-Interestingly, we found a consistent negative relationship between range size and speciation rate, at least when this rate was quantified using $\lambda_{DR}$ and $\lambda_{ND}$.
-This suggests that species with smaller range sizes have increased speciation rates.
-None of the bioclimatic measures of environmental variability that we investigated (*i.e.*, temperature seasonality, long-term temperature variation, and spatial temperature variation) predicted speciation rate.
+We found evidence that the composite index of male-biased sexual selection, but not measures of sexual dichromatism, is correlated with the rate of speciation in passerine birds.
+The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB), and it cannot be explained by a difference in statistical power since the sample size for the dichromatism analyses was much larger.
+These findings reaffirm the conclusions of previous, smaller studies in which sexual dichromatism was measured using spectrophotometry [@Huang_2014] or human observers [@Cooney_2017].
+The correlation between speciation rate and the index of male-biased sexual selection (which encapsulates variation in sexual size dimorphism, social polygyny, and paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$, although the estimated direction and magnitude of the correlation was broadly similar across all three measures.
+Interestingly, we also found a consistent negative relationship between range size and speciation rate, at least when this rate was quantified using $\lambda_{DR}$ and $\lambda_{ND}$.
+None of the bioclimatic measures of environmental variability that we investigated (*i.e.*, temperature seasonality, long-term temperature variation, and spatial temperature variation) significantly predicted speciation rate.
 
-**Find a paper about Eclectus parrots to add to the paragraph below, maybe.
-Males and green and females are bright red, and I think females compete for male parental care, so SS is reversed.
-Another reason why male-biased SS might not be well-correlated with dichromatism**
-
-Although many comparative studies of birds have used sexual dichromatism as a proxy for the strength of sexual selection [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014], we found that sexual dichromatism did not predict any measure of speciation rate, while the index of the male-biased sexual selection predicted $\lambda_{DR}$.
-Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection at least at this broad scale, and that variation in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study.
+The difference in findings between the analyses of sexual dichromatism _versus_ the index of sexual selection is noteworthy, because the majority of earlier studies used dichromatism alone as their proxy for sexual selection [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
+Given our findings, and the modest correlation between dichromatism and the sexual selection index [$r = 0.34$; @Dale_2015], we hypothesise that sexual dichromatism may not be a robust proxy for sexual selection.
+Although dichromatism clearly reflects sexual selection to some extent, it is possibly that it is too indirect a measure to detectany association with speciation rate, even with a high sample size.
 There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
@@
 Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit weakly.
-Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018].
+Secondly, colour is only one of the traits subject to sexual selection [@Miles_2018; @Cally_2019].
 For instance, sexual selection may promote investments in other mating signals (*e.g.* song) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
 Lastly, not all plumage colouration honestly reflects mate quality.
-A recent meta-analysis found that converted carotenoids (but *not* carotenoids sourced from the diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
-Given the importance of honest signaling in sexual selection and the variability in colour production mechanisms across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
+A recent meta-analysis found that converted carotenoids (but not carotenoids sourced from the diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
+Thus, bright colours may signal direct benefits or 'good genes' in some species but not others, and some of the colours used to calculate dichromatism might be more functionally relevant than others.
 
-In-line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least for $\lambda_{DR}$ estimates.
+In-line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least when speciation is measured by $\lambda_{DR}$.
 Additionally, we found that this association appears to be independent of net primary productivity and spatiotemporal variation in the environment.
-The lack of an effect of these environmental variables on speciation rate (either alone or in combination with sexual selection) has several possible interpretations.
-Firstly, the effects of sexual selection on adaptation and thus speciation may depend on the type of environmental variability under which the species is evolving.
-Specifically, speciation rates may be impacted by genetic constraints on adaptation, that vary across environments.
-Theory suggests that sexual antagonism (often arising from increased sexual selection) may be easily purged in environments where pressures are cyclic (*e.g.* seasonality), whereas when environmental variability is directional (*e.g.* long-term climate change) sexual antagonism can indefinitely limit adaptation rates [@Connallon_Hall_2016].
-Another possibility is that the environmental predictors used here may not account for the key ecological forces/natural selection pressures that interact with sexual selection to drive speciation.
-Specifically, access to dietary resources and the impacts of predation or parasitism are unaccounted for here.
-These are likely key processes affecting sexual selection and speciation [*reviewed in* @Maan_2011].
-Finally, it is also possible that there is no effect of environmental variability on speciation rates; at least not in the species investigated here.
+The lack of an effect of these environmental variables on speciation rate has several possible interpretations.
+Firstly, the effects of sexual selection on adaptation and speciation may depend on the type of environmental variability under which the species is evolving.
+Specifically, speciation rates might be impacted by genetic constraints on adaptation, that vary across environments.
+Theory suggests that sexual antagonism (which is often exacerbated in species with strong sexual selection) may be lower in habitats experiencing cyclical environmental variation (*e.g.* seasonality), relative to those experiencing directional change in the environment [@Connallon_Hall_2016].
+Another possibility is that the environmental predictors we chose may not account for the key ecological sources of selection that interact with sexual selection to drive speciation.
+For example, our study does not include direct measure of food availability or the severity of predation and parasitism, which are both hypothesised to affect sexual selection and speciation [*reviewed in* @Maan_2011].
+Finally, it is possible that environmental variability genuinely has little effect on speciation rates, at least in the taxa investigated here.
 
-The most consistent finding uncovered in this comparative analysis is that smaller range sizes are associated with increased speciation rates.
-Intuitively, large range size should promote speciation by creating greater opportunities for geographic barriers to form [@Rosenzweig_1995; @Castiglione_2017].
-However, the opposite pattern has also been suggested because birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
-In amphibians (329 genera) diversification was shown to be higher in taxa with smaller range size [@Greenberg_2017].
-Alternatively, smaller range size may be correlated but not causally related to speciation.
-Under this view, high speciation rate may lead to smaller range sizes as niche filling by recently diverged species will suppress the expansion of newly speciated relatives [@Rosenzweig_1995; @Weir_2011; @Price_2014].
-However, species undergoing adaptive radiation in new niches are unlikely to be limited by competition for resources from existing taxa.
+The most robust finding in our study is the finding that species with smaller ranges have elevated speciation rates.
+This result is similar to a study of 329 amphibian genera, which found higher diversification rates in taxa with smaller range size [@Greenberg_2017].
+Intuitively, large range size should promote speciation by creating more opportunities for geographic barriers to form [@Rosenzweig_1995; @Castiglione_2017].
+However, the opposite pattern is also plausible because birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
+**Please check this sentence, I didn't understand the original but I tried to re-draft for clarity** It is also possible that high speciation rates cause smaller range sizes, rather than the other way around, for example because repeatedly-speciating lineages tend to fill niches in ways that hinder the geographical expansion of new species [@Rosenzweig_1995; @Weir_2011; @Price_2014].
+However, species undergoing adaptive radiation in new habitats are unlikely to be limited by competition for resources from existing taxa.
 Across islands, we expect to see a correlation between speciation and small range size because small islands often reflect newly formed environments with empty niches.
-Thus, our findings could be heavily dependent on island radiations (*e.g.* "Darwin's finches", *Geospiza* spp.).
+**As a jerk reviewer, I would say "can't you re-run it without the island radiations??"** Thus, our findings might be driven by island radiations (*e.g.* "Darwin's finches", *Geospiza* spp.).
 One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
 
-One outcome of our analyses was that different measures of speciation rates presented different results.
-This is not completely unexpected, given that each of the rates is calculated in a different way (discussed in detail in @Rabosky_diversification_2018).
-For instance, $\lambda_{DR}$ is weighted more towards speciation events close to the tips and allows more rate heterogeneity compared to BAMM estimates (although increasing error rate).
-The higher variation in $\lambda_{DR}$ could explain why we detected a significant pattern using this metric and not the other two.
-We think it is unlikely that the significant correlation between $\lambda_{DR}$ and sexual selection is spurious given the variation in $\lambda_{DR}$ values and the consistency in the calculated estimates across the 1,000 trees.
-
-In addition to speciation, sexual selection has frequently been invoked as a driver of extinction.
-Sexual selection may increase extinction risk as it often leads individuals (usually males) to invest in exaggerated mating signals as opposed to traits --- such as parental care --- that elevate offspring fitness [*reviewed in* @Kokko_Brooks_2003].
-Using the model-based approach of BAMM, we found no association between extinction rate and sexual dichromatism, male-biased sexual selection or measures of environmental variability.
-Extinction is notoriously difficult to estimate accurately from phylogenies, principally because many different combinations of speciation and extinction rates can give rise to similar patterns of diversity [*see* @Rabosky_Extinction_2016].
-Phylogenetic methods such as BAMM allow for speciation and extinction rates to be estimated using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates remains a subject of debate [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
-Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models and are fraught with issues of sampling bias [@Davis_2013].
-Although extinction rates can be estimated from alternative sources, such as the fossil record, documented recent extinctions and IUCN extinction threat status, each approach has significant limitations.
+In addition to speciation, sexual selection is hypothesised to affect extinction.
+Using the model-based approach of BAMM, we found no association between the estiamted extinction rate and sexual dichromatism, male-biased sexual selection, or our measures of environmental variability.
+However, these extinction results should not be regarded as definitive because extinction is notoriously difficult to estimate accurately from phylogenies, principally because different combinations of speciation and extinction rates can give rise to similar patterns of diversity [see @Rabosky_Extinction_2016].
+Phylogenetic methods such as BAMM allow for speciation and extinction rates to be estimated using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates is debated [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
+Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models which are sensitive to sampling bias [@Davis_2013].
+Although extinction rates can be inferred from alternative sources, such as the fossil record [@Martins_2018], direct observation extinction, or IUCN red list status [@Greenberg_2017], each approach has its limitations.
 Across the passerine bird phylogeny, we found that BAMM often produced homogeneous speciation and extinction rates for smaller clades showing few rate shifts, which might reduce our power to detect small differences in extinction rates among closely-related taxa [@Rabosky_BAMM_flawed_2017; @Rabosky_diversification_2018].
@@
 
-To summarise, we have shown that in passerines, male-biased sexual selection, but not sexual dichromatism, predicts speciation, independent of several measures of environmental variability.
-These findings support the idea that sexual dechromatism is not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
-If sexual selection promotes speciation through improvements in fitness and adaptation the implications are manifold, including for conservation [*reviewed in*, @Holman_Kokko_2013] and captive breeding programs for threatened species [*reviewed in*, @Charge_2014].
-Furthermore, our finding that high speciation rate is associated with smaller range size highlights the threat to the persistence of rapidly speciating lineages in a world with increased habitat loss and anthropogenic stress.
+One outcome of our analyses was that different measures of speciation rates presented different results.
+This is not completely surprising, because each of the rates is calculated differently (see @Rabosky_diversification_2018).
+For instance, $\lambda_{DR}$ is weighted more towards speciation events close to the tips and allows more rate heterogeneity compared to BAMM estimates.
+This leads to greater variation in $\lambda_{DR}$ relative to the BAMM estimates, potentially explaining the difference in results.
+The BAMM estimates were also more sensitive to phylogenetic uncertainty.
+
+To summarise, we have shown that male-biased sexual selection, but not sexual dichromatism, predicts speciation in passerines, and that this relationship is not markedly affected by environmental variability.
+These findings imply that sexual dechromatism is not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
+Our results also add indirect suppport to the hypothesis that sexual selection promotes adaptation, which has implications for conservation [@Holman_Kokko_2013] and captive breeding programs for threatened species [@Charge_2014].
+Furthermore, our finding that high speciation rate is associated with smaller range size highlights the threat to the persistence of rapidly-speciating lineages in a world with increased habitat loss and anthropogenic stress.
 This concern arises as the best predictor of extinction risk is range size [@Harris_2008] and the association found here implies that many newly-speciated clades have small range sizes and could thus be at greater risk of extinction.

### 2019-07-26 Manuscript_Publication.Rmd
More writing

@@
 If so, probably best to search for the word "random" and remove/reword it.
+
+I think it might make sense to call it "the index of male-biased sexual selection" throughout, including in the figures.
+Otherwise it's not clear when you mean the general phenomenon vs our measure of it.
+See what you think after reading my edits.
 -->
@@
 
-We used a previously published measure of sexual dichromatism for 5,983 species of passerines [@Dale_2015].
-Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forehead, throat, upper breast, and lower breast) from _Handbook of the Birds of the World_ [@delhoyo2011].
+We used a previously-published measure of sexual dichromatism for 5,983 species of passerines [@Dale_2015].
+Briefly, @Dale_2015 obtained sex-specific RGB (red-green-blue) values across six body patches (nape, crown, forehead, throat, upper breast, and lower breast) from _Handbook of the Birds of the World_ [@delhoyo2011].
 The relative contribution of male and female RGB colour values were averaged across body patches and provide 'male-like' and 'female-like' plumage scores.
@@
 Additionally, we used another measure of dichromatism corresponding to colour distance in avian colour space derived from spectral data [@Armenta_2008].
-These measurements include variation in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from museum samples (as opposed to colour drawings).
-However, this spectrophotometry data is not as extensive as the RGB measures; with only 581 passerine species available for this analysis.
-While there is a correlation between these two measures, there is residual variation ($r = 0.79$; \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
+These measurements include variation in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from museum specimens as opposed to illustrations.
+The spectrophotometry data covers only 581 passerine species (10-fold fewer than the RGB data), although there was a substantial correlation between the two dichromatism measures ($r = 0.79$; \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
 \phantomsection
@@
 
-Sexual dichromatism is a widely used indicator of sexual selection in birds [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014], but only partially captures variation in the strength of sexual selection across taxa.
-Therefore, we also use an existing dataset [@Dale_2015] of male-biased sexual selection based on the first component of a phylogenetic principle component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
-This measure of male-biased sexual selection is available for only 2,465 species and is weakly correlated with the absolute values of sexual dichromatism using RGB measures ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
+Because sexual dichromatism is presumably imperfectly correlated with variation in the strength of sexual selection across taxa, we sourced an additional measure of sexual selection, also from @Dale_2015, referred to here as the 'index of male-biased sexual selection'.
+This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
+This measure of male-biased sexual selection is available for only 2,465 species, and shows a moderate correlation with the RGB measure of sexual dichromatism ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
 \phantomsection
@@
 
-Given that stressful environments are expected to interact with sexual selection and have a positive effect on adaptation [@Cally_2019], we used the extracted environmental variables from each species range size to develop biologically meaningful predictors of environmental variation/stress.
+Given that stressful environments are expected to interact with sexual selection and have a positive effect on adaptation [@Cally_2019], we used the extracted environmental variables from each species range size to create predictors of environmental variation/stress.
 Firstly we used (*i*) the average NPP in each species' range and (*v*) the log-transformed range size as potentially informative predictors of speciation rates.
@@
 Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
-Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events and $\lambda_{ND}$ is a measure of speciation across the root-to-tip path.
+Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path.
 These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE); but, based on previous simulation studies, $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
@@
 To test the association between speciation/extinction and sexual selection, environmental variability and their interaction, we used phylogenetic least squares (PGLS) models in the ``nlme`` package [@nlme_2018].
-Firstly we conducted model selection using $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ tip-rate estimates from the same MCC tree, which was derived from 2,500 draws of the posterior distribution [@Jetz_2012].
-For $\lambda_{BAMM}$ and $\mu_{BAMM}$ we were also able to use model weights sourced from the inverse of the variance of tip-rates for a given species from the posterior distribution (*n* = 1,000).
-Model weights thus reflect the degree of precision to which each species' tip-rate is measured in BAMM.
-Using model selection we only compared interaction terms between a measure of sexual selection (sexual dichromatism or male-biased sexual selection) and environmental measures.
-That is, we fixed the individual predictors of dichromatism/male-biased sexual selection measures, log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP while comparing 32 models with different combinations of interactions (including none).
+Firstly, we conducted model selection to compare models in which $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ were the response variable: these tip-rate estimates all came from the same MCC tree (derived from 2,500 draws of the posterior distribution in @Jetz_2012).
+For models of $\lambda_{BAMM}$ and $\mu_{BAMM}$ we used the inverse of the variance associated with each tip rate estimate as weights, to account for the variable precision of the estimates provided by BAMM.
+The most complicated model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (*i.e.* log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP), and all of the 2-way interactions between sexual selection and each of the environmental measures.
+The simpler models contained all of the same main effects, but had fewer 2-way interaction terms (potentially none).
 Model selection was done in ``MuMIn`` using the ``dredge`` function [@MuMIn_2017].
-Using the terms from the top-ranking model (lowest AICc), we ran the equivalent model for each of the 1,000 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and each of the 100 used to derive $\lambda_{BAMM}$ and $\mu_{BAMM}$.
+Using the terms from the top-ranked model (ranked by AICc), we ran the equivalent model for each of the 1,000 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and each of the 100 trees used to derive $\lambda_{BAMM}$ and $\mu_{BAMM}$.
 In each model we used the unique response variables and phylogenetic tree correlation structure.
@@
 
-Finally, using data subset for species with measurements of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
+Finally, using the subset of species with measurements of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
 The phylogenetic path analysis was used to assess causal paths *between* variables not able to be modelled within the univariate response of PGLS.
@@
 To minimise path complexity we used temperature seasonality (BIO4) as the single measure for environmental variability, $\lambda_{DR}$ as the single measure of speciation and the tip-rates from the MCC tree.
-Further details of the path analysis, including reasons for path directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
+Further details of the path analysis, including our rationale for each path's directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
 

### 2019-08-07 Manuscript_Publication.Rmd

@@
 - Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§}
-  and Iliana Medina Guzmán\textsuperscript{§}
+  and Iliana Medina\textsuperscript{§}
 - '*\textit{justin.g.cally@gmail.com}'
@@
 output:
+  word_document: default
   pdf_document: default
@@
   \raggedbottom
-csl: References/evolution.csl
+csl: References/NComms.csl
 subparagraph: yes
-abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates.
-However, the net effect of sexual selection on diversification is hard to predict _a priori_, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes.
-Here, we test whether the strength of sexual selection predicts variation in speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera).
-We tested for associations between two measures of sexual selection (sexual dichromatism, plus an index that captures variance in sexual size dimorphism, polygyny, and male parental care), several environmental variables, and multiple measures of speciation and extinction rates.
-Our results show that male-biased sexual selection, but not sexual dichromatism, predicts speciation rates ($\lambda_{DR}$) in passerines, and found no evidence that this relationship varies with the environment.
-We also found a strong negative relationship between range size and speciation rate.
-There was no correlation between sexual selection and extinction rate **XXXXXX**.
-Our findings support the view that sexual selection is a potent evolutionary force that has shaped diversification in songbirds.
-\vspace{5mm}  \par\noindent
-  \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage dichromatism, Range size, Extinction rate.
+abstract: Sexual selection is thought to shape phylogenetic diversity by affecting
+  speciation or extinction rates. However, the net effect of sexual selection on diversification
+  is hard to predict _a priori_, because many of the hypothesised effects on speciation
+  or extinction have opposing signs and uncertain magnitudes. Here, we test whether
+  the strength of sexual selection predicts variation in speciation and extinction
+  rates across passerine birds (up to 5,812 species, covering most genera). We tested
+  for associations between two measures of sexual selection (sexual dichromatism and
+  a multivariate measure of male-biased sexual selection), several environmental variables,
+  and multiple measures of speciation and extinction rates. Our results show that
+  male-biased sexual selection, but not sexual dichromatism, predicts speciation rates
+  ($\lambda_{DR}$) in passerines, and found no evidence that this relationship varies
+  with the environment. We also found a strong negative relationship between range
+  size and speciation rate. There was no correlation between sexual selection and
+  extinction rate ($\mu_{BAMM}$). Our findings show that sexual dichromatism is a
+  poor proxy for sexual selection at least a broad-scale and support the view that
+  sexual selection is a potent evolutionary force that has shaped diversification
+  analysis, Plumage dichromatism, Range size, Extinction rate.
+  in songbirds. \vspace{5mm}  \par\noindent \textbf{Keywords:} Speciation rate, Comparative
 ---
-\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}
+\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia}
 
@@
 <!-- Luke general comments:
-In a few places you say things like "1000 random trees".
+In a few places you say things like "1000 rrandom trees".
 To me this means trees with a random topology (which are sometimes used as some sort of null model), but I think you actually mean 1000 trees randomly sampled (without replacement) from the posterior of trees in Jetz et al.
@@
 Recent empirical and methodological advances, such as growing evidence that sexual selection can profoundly affect many traits [e.g.
-life history, body mass, and immunocompetence; @Cally_2019], larger and more accurate phylogenies [@Jetz_2012], and new phylogenetic methods (**REFERENCES**), present new opportunities to test whether and how sexual selection drives diversification.
+life history, body mass, and immunocompetence; @Cally_2019], larger and more accurate phylogenies [@Jetz_2012], and new phylogenetic methods [@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.
 
 Several hypotheses predict a positive relationship across species between the strength of sexual selection and the rate of speciation and/or extinction.
-For example, sexual selection is involved in reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation (**REF**).
+For example, sexual selection is involved in reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation [see @Ritchie_2007].
 Additionally, sexual selection often operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia].
 Furthermore, sexual selection has been hypothesised to increase diversification by maintaining trait combinations that would be selected out under pure natural selection [@Bonduriansky_2011; @Radwan_2016], and by helping to prevent extinction by purging deleterious mutations [@Whitlock_Agrawal_2009], fixing beneficial ones [@Whitlock_2000], and accelerating adapation in environments [@Lorch_2003; @Candolin_2008; @Cally_2019].
-Also, sexual selection sometimes favours individuals with rare phenotypes [@holman2013crozier], providing a source of diversifying selection that might increase the rate at which isolated populations diverge genetically.
 
-Conversely, there are multiple ways in which sexual selection might hinder speciation or make extinction more likely.
-Firstly, sexual selection might promote extinction by favouring traits that improve mating success but reduce population fitness.
-For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the reproductive success of individuals while reducing population-wide productivity [e.g.
-infanticide, harassment, and reduced parental care by mate-seeking males; @RN37; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016].
+Conversely, sexual selection might hinder speciation or make extinction more likely by favouring traits that improve mating success but reduce population fitness.
+For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the reproductive success of individuals while reducing population-wide productivity [e.g., infanticide, harassment, and reduced parental care by mate-seeking males; @RN37; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016].
 Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013].
-Secondly, sexual selection might promote evolutionary stasis by ensuring that individuals with novel trait values have low mating success.
 
@@
 @Long_2012; @Connallon_Hall_2016], though the empirical evidence for this theory is mixed and largely limited to insects [@Cally_2019].
-The mechanism underpinning this prediction is that the environment and the selective history of the population should influence the relative amounts of genetic variation at sexually concordant loci (*i.e.* loci where the same allele is fittest for both sexes) and sexually antagonistic loci (where different alleles are favoured in each sex).
+The mechanism underpinning this prediction is that the environment and the selective history of the population should influence the relative amounts of genetic variation at sexually concordant loci (i.e.
+loci where the same allele is fittest for both sexes) and sexually antagonistic loci (where different alleles are favoured in each sex).
 In stable environments, variation is thought to be preferentially depleted at sexually concordant loci, leading to stronger gender load and reduced net benefits of sexual selection [e.g.
@@
 
-<!-- .; causing displacement of male and female traits from their optimum.
-It follows from these predictions that populations with stronger sexual selection would have fitness benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008].
-Additionally, sexual selection can reinforce local adaptations through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011].
-For instance, beak morphology is a trait that arose under natural selection in several taxa and is now a trait under sexual selection [a.k.a.
-*magic traits*; @Servedio_2011].
-In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-Through these synergetic effects sexual selection and natural selection (*e.g.* environmental variation over space and time) may lead to increased speciation rates.
--->
-
-Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [*e.g.* @Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014].
+Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [e.g.
+@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014].
 Birds are well-suited for this purpose due to their diverse and well-characterised mating systems and sexually-selected traits, and because their phylogenetic relationships are comparatively well-known.
@@
 We use two measures of the strength of sexual selection: sexual dichromatism [*n* = 5,812 species; estimated from illustrations by @Dale_2015], as well as an index of male-biased sexual selection [*n* = 2,465 species; derived via principal components analysis by @Dale_2015], which captures (co)variation in sexual size dimorphism, social polygyny and paternal care.
-Most **OR ALL??** earlier macroevolutionary studies of this topic only quantified sexual selection using dichromatism, which is problematic since dichromatism does not always signal the presence of strong sexual selection and _vice versa_ [@Dale_2015].
+These two measures are used because sexual dichromatism does not always signal the presence of strong sexual selection and _vice versa_ [@Dale_2015].
 For example, male and female dunnocks (_Prunella modularis_) are similarly coloured yet sexual selection appears to be strong [@davies1986re].
@@
 Additionally, our analysis includes multiple ecological and environmental variables on diversification, allowing us to adjust for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016].
-We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
-Finally, we incorporate phylogenetic uncertainty into our analyses, improving over earlier work which assumed the statistically best-fitting phylogeny is correct.
+We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; see, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
 \phantomsection
@@
 Having said that I can see a case for leaving it here too.
-- Luke]**
+- Luke] [We initially had this at the end but moved it up to the start.
+What do you think Devi?
+- Justin]**
 
-We first estimated speciation rates across the phylogeny, using the maximum clade credibility (MCC) tree, as well as 100 or 1000 trees sampled from the posterior distribution of trees in @Jetz_2012.
-Estimates of speciation rates varied between phylogenetic trees, especially for rates calculated by BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95% highest posterior density (HPD) interval across 100 trees was >20-fold larger than the **95% confidence interval [SHOULD THIS BE HPD TOO?]** calculated for the MCC tree.
-Conversely, for our estimates of diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$), the 95% HPD interval across 1,000 trees was near-identical to the 95% confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
+Estimates of the effect of predictor variables on speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % highest posterior density (HPD) interval across phylogenetic least squares (PGLS) model estimates from 100 trees was often > 20 times larger than the 95 % confidence interval for estimates from a single PGLS model using the MCC (maximum clade credibility) tree.
+This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for pgls model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
 Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar statistical support is notable.
@@
 
-**[I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Abstract and Intro. - Luke]**
+**[I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Abstract and Intro.
+- Luke] [Statistically, I think it is best to rule out the interactions first but Yes I see your point - Justin]**
 
-We next fit Phylogenetic Generalized Least Squares (PGLS) models with **2-way [CORRECT?]** interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) and compared them using AICc model selection.
-However, no interaction terms were present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
+We next fit PGLS models with 2-way interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) were included in the full model for posterior model selection.
+However, no interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\Delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
 Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
@@
 Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
+PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; i.e.
+no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
 Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 trees for $\lambda_{BAMM}$.
-The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
+The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top$SDi, format = "e", digits = 3)`, $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top$SDi, format = "e", digits = 3)`,\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
 For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
@@
 However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-The distribution of estimates from PGLS models on 1,000 trees was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and a smaller positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
+The distribution of estimates from PGLS models on 1,000 trees was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
 Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}.
@@
 Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
-The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 1,000 random trees (for $\lambda_{DR}$ and $\lambda_{ND}$) or 100 random trees for $\lambda_{BAMM}$ and $\mu_{BAMM}$ can be found in the ESM.
+The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 1,000 randomly sampled trees (for $\lambda_{DR}$ and $\lambda_{ND}$) or 100 randomly sampled trees for $\lambda_{BAMM}$ and $\mu_{BAMM}$ can be found in the ESM.
 Density curves are based on model estimates from 1,000/100 trees and the circle below with error bars is the estimate and 95% CIs from the MCC tree.
-For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to be able to visualise the MCC 95% CIs.
+For this figure we removed outliers from estimates coming from the 100 randomly sampled trees for BAMM models in order to be able to visualise the MCC 95% CIs.
 \label{model_results}](Figures/Estimates_and_scatter_edit.pdf)
@@
 $\lambda_{DR}$ are those from the MCC tree and images of birds are from the _Handbook of the Birds of the World_.
-Clockwise the species are: _Sporophila bouvronides_, _Geospiza magnirostris_, _Atlapetes leucopterus_, _Euphonia rufiventris_, _Euplectes franciscanus_, _Promerops gurneyi_, _Phainopepla nitens_, _Zosterops erythropleurus_, _Acrocephalus australis_, _Progne cryptoleuca_, _Mayrornis schistaceus_, _Paradisaea rubra_, _Pachycephala pectoralis_, _Malurus pulcherrimus_, _Cranioleuca curtata_, _Cercomacra manu_, _Muscisaxicola capistratus_, _Lepidothrix coeruleocapilla_.
+Clockwise the six species are: _Sporophila bouvronides_, _Euplectes franciscanus_, _Phainopepla nitens_, _Paradisaea rubra_, _Malurus pulcherrimus_, _Lepidothrix coeruleocapilla_.
 Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation and should not be interpreted.
@@
 Interestingly, we also found a consistent negative relationship between range size and speciation rate, at least when this rate was quantified using $\lambda_{DR}$ and $\lambda_{ND}$.
-None of the bioclimatic measures of environmental variability that we investigated (*i.e.*, temperature seasonality, long-term temperature variation, and spatial temperature variation) significantly predicted speciation rate.
+None of the bioclimatic measures of environmental variability that we investigated (i.e., temperature seasonality, long-term temperature variation, and spatial temperature variation) significantly predicted speciation rate.
 
-The difference in findings between the analyses of sexual dichromatism _versus_ the index of sexual selection is noteworthy, because the majority of earlier studies used dichromatism alone as their proxy for sexual selection [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
+The difference in findings between the analyses of sexual dichromatism _versus_ the index of sexual selection is noteworthy, because the majority of earlier studies used dichromatism alone as their proxy for sexual selection [e.g., @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
 Given our findings, and the modest correlation between dichromatism and the sexual selection index [$r = 0.34$; @Dale_2015], we hypothesise that sexual dichromatism may not be a robust proxy for sexual selection.
@@
 Secondly, colour is only one of the traits subject to sexual selection [@Miles_2018; @Cally_2019].
-For instance, sexual selection may promote investments in other mating signals (*e.g.* song) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
+For instance, sexual selection may promote investments in other mating signals (e.g.
+song) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018].
 Lastly, not all plumage colouration honestly reflects mate quality.
 A recent meta-analysis found that converted carotenoids (but not carotenoids sourced from the diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018].
-Thus, bright colours may signal direct benefits or 'good genes' in some species but not others, and some of the colours used to calculate dichromatism might be more functionally relevant than others.
+Given the importance of honest signaling in sexual selection and the variability in colour production mechanisms across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
 
@@
 Specifically, speciation rates might be impacted by genetic constraints on adaptation, that vary across environments.
-Theory suggests that sexual antagonism (which is often exacerbated in species with strong sexual selection) may be lower in habitats experiencing cyclical environmental variation (*e.g.* seasonality), relative to those experiencing directional change in the environment [@Connallon_Hall_2016].
+Theory suggests that sexual antagonism (which is often exacerbated in species with strong sexual selection) may be lower in habitats experiencing cyclical environmental variation (e.g.
+seasonality), relative to those experiencing directional change in the environment [@Connallon_Hall_2016].
 Another possibility is that the environmental predictors we chose may not account for the key ecological sources of selection that interact with sexual selection to drive speciation.
-For example, our study does not include direct measure of food availability or the severity of predation and parasitism, which are both hypothesised to affect sexual selection and speciation [*reviewed in* @Maan_2011].
+For example, our study does not include direct measure of food availability or the severity of predation and parasitism, which are both hypothesised to affect sexual selection and speciation [reviewed in @Maan_2011].
 Finally, it is possible that environmental variability genuinely has little effect on speciation rates, at least in the taxa investigated here.
@@
 However, the opposite pattern is also plausible because birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
-**Please check this sentence, I didn't understand the original but I tried to re-draft for clarity** It is also possible that high speciation rates cause smaller range sizes, rather than the other way around, for example because repeatedly-speciating lineages tend to fill niches in ways that hinder the geographical expansion of new species [@Rosenzweig_1995; @Weir_2011; @Price_2014].
+It is also possible that high speciation rates cause smaller range sizes, rather than the other way around, for example because repeatedly-speciating lineages tend to fill niches in ways that hinder the geographical expansion of new species [@Rosenzweig_1995; @Weir_2011; @Price_2014].
 However, species undergoing adaptive radiation in new habitats are unlikely to be limited by competition for resources from existing taxa.
-Across islands, we expect to see a correlation between speciation and small range size because small islands often reflect newly formed environments with empty niches.
-**As a jerk reviewer, I would say "can't you re-run it without the island radiations??"** Thus, our findings might be driven by island radiations (*e.g.* "Darwin's finches", *Geospiza* spp.).
 One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
@@
 However, these extinction results should not be regarded as definitive because extinction is notoriously difficult to estimate accurately from phylogenies, principally because different combinations of speciation and extinction rates can give rise to similar patterns of diversity [see @Rabosky_Extinction_2016].
-Phylogenetic methods such as BAMM allow for speciation and extinction rates to be estimated using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates is debated [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
-Additionally, while several tip-rate estimates exist for speciation rate (*e.g.*, $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models which are sensitive to sampling bias [@Davis_2013].
+Phylogenetic methods such as BAMM allow for speciation and extinction rates to be estimated using moderately-sized phylogenies, although the ability of BAMM to model evolutionary rate shifts and extinction rates is debated [see, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017].
+Additionally, while several tip-rate estimates exist for speciation rate (e.g., $\lambda_{DR}$ and $\lambda_{ND}$), tip-rate estimates of extinction rate are difficult to obtain without complex Bayesian models which are sensitive to sampling bias [@Davis_2013].
 Although extinction rates can be inferred from alternative sources, such as the fossil record [@Martins_2018], direct observation extinction, or IUCN red list status [@Greenberg_2017], each approach has its limitations.
@@
 We obtained estimates of species range size using expert range maps [@birdlife_2017].
-Because of taxonomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
+Because of taxonomic changes to 1,230 species in the Birdlife database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
 From these distributions, we obtained estimates of climatic conditions that each species is exposed to.
@@
 To obtain seasonal climate variation we used (*iii*) mean values of temperature seasonality (BIO4) for each range.
-(*iV*) To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO 4 and BIO15).
+(*iv*) To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO 4 and BIO15).
 PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
@@
 To obtain long-term variation in climates for each species range we take (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
-Similar to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
+Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
 The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
@@
 Firstly, we obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
-Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [*e.g.*, @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each edge towards the root having the length down-weighted.
+Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [e.g., @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each edge towards the root having the length down-weighted.
 Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
@@
 For models of $\lambda_{BAMM}$ and $\mu_{BAMM}$ we used the inverse of the variance associated with each tip rate estimate as weights, to account for the variable precision of the estimates provided by BAMM.
-The most complicated model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (*i.e.* log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP), and all of the 2-way interactions between sexual selection and each of the environmental measures.
+The most complicated model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (i.e.
+log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP), and all of the 2-way interactions between sexual selection and each of the environmental measures.
 The simpler models contained all of the same main effects, but had fewer 2-way interaction terms (potentially none).
@@
 Finally, using the subset of species with measurements of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
-The phylogenetic path analysis was used to assess causal paths *between* variables not able to be modelled within the univariate response of PGLS.
+The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
 That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection, and range size to have effects on each other and not just on speciation rate.
@@
 
-We would like to acknowledge Adnan Moussalli for his assistance in conducting BAMM runs across 100 phylogenetic trees. **Any others?**
+We would like to acknowledge Adnan Moussalli for his assistance in conducting BAMM runs across 100 phylogenetic trees.
 

### 2019-08-09 Manuscript_Publication.Rmd
Added Devis changes

@@
 ---
-title: \huge{Male-biased sexual selection, but not sexual dichromatism, predicts speciation
-  in birds}
+title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation
+  in birds
 author:
@@
   pdf_document: default
-  html_document:
-    df_print: paged
 header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
@@
   rates across passerine birds (up to 5,812 species, covering most genera). We tested
-  for associations between two measures of sexual selection (sexual dichromatism and
-  a multivariate measure of male-biased sexual selection), several environmental variables,
-  and multiple measures of speciation and extinction rates. Our results show that
-  male-biased sexual selection, but not sexual dichromatism, predicts speciation rates
-  ($\lambda_{DR}$) in passerines, and found no evidence that this relationship varies
-  with the environment. We also found a strong negative relationship between range
-  size and speciation rate. There was no correlation between sexual selection and
-  extinction rate ($\mu_{BAMM}$). Our findings show that sexual dichromatism is a
-  poor proxy for sexual selection at least a broad-scale and support the view that
-  sexual selection is a potent evolutionary force that has shaped diversification
-  analysis, Plumage dichromatism, Range size, Extinction rate.
-  in songbirds. \vspace{5mm}  \par\noindent \textbf{Keywords:} Speciation rate, Comparative
+  for associations between speciation or extinction and two measures of sexual selection
+  (sexual dichromatism and a multivariate measure of male-biased sexual selection),
+  accounting for range size and measures of environmental variability. Male-biased
+  sexual selection, but not sexual dichromatism, predicted speciation rates ($\lambda_{DR}$)
+  in passerines. This relationship was independent of range size or environmental
+  variability, though species with smaller ranges had higher speciation rates. There
+  was no association between sexual selection and extinction rate ($\mu_{BAMM}$).
+  Our findings show that sexual dichromatism is a poor proxy for sexual selection
+  at least a broad-scale and support the view that sexual selection has shaped diversification
+  in songbirds.
 ---
@@
 
-Sexual selection results from competition between same-sex individuals for mates, their gametes, or associated resources, and in most species it predominently operates on males [@Kokko_Jennions2008; @Fromhage_2016; @Janicke_2016].
-There is long-standing interest in the relationship between sexual selection and the processes of speciation, extinction, and phylogenetic diversification [e.g.
-@Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018].
-Recent empirical and methodological advances, such as growing evidence that sexual selection can profoundly affect many traits [e.g.
-life history, body mass, and immunocompetence; @Cally_2019], larger and more accurate phylogenies [@Jetz_2012], and new phylogenetic methods [@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.
+Sexual selection is a fundamental evolutionary process; yet there is long-standing debate about how it shapes patterns of species diversity[@Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018].
+Sexual selection can promote speciation because it operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection[@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia].
+Sexual selection could also promote speciation or prevent extinction by purging deleterious mutations[@Whitlock_Agrawal_2009], fixing beneficial ones[@Whitlock_2000], and accelerating adaptation in different environments[@Lorch_2003; @Candolin_2008; @Cally_2019].
+Conversely, sexual selection might hinder speciation or make extinction more likely by favouring traits that improve mating success but reduce population fitness[@Rankin_2011; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016].
+For example, species with costly sexual signals may be less resilient to environmental change[@Kokko_Brooks_2003].
+Sexual selection might also promote extinction by causing maladaptation (‘gender load’) in female traits that are genetically correlated with sexually-selected male traits[@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013].
+Although numerous studies have examined the relationship between sexual selection and speciation or extinction rates[@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Kraaijeveld_2011; @Seddon_2013; @Huang_2014], the availability of more complete phenotypic, ecological and phylogenetic data[@Jetz_2012], together with significant advances in phylogenetic methods[@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.
 
-Several hypotheses predict a positive relationship across species between the strength of sexual selection and the rate of speciation and/or extinction.
-For example, sexual selection is involved in reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation [see @Ritchie_2007].
-Additionally, sexual selection often operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia].
-Furthermore, sexual selection has been hypothesised to increase diversification by maintaining trait combinations that would be selected out under pure natural selection [@Bonduriansky_2011; @Radwan_2016], and by helping to prevent extinction by purging deleterious mutations [@Whitlock_Agrawal_2009], fixing beneficial ones [@Whitlock_2000], and accelerating adapation in environments [@Lorch_2003; @Candolin_2008; @Cally_2019].
+The relationship between sexual selection and diversification may depend on the environment.
+Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones[@Long_2012; @Connallon_Hall_2016].
+In stable environments, consistent selection depletes genetic variation at sexually concordant loci (i.e.
+loci where the same allele is fittest for both sexes).
+In these environments, genetic variation remains disproportionately at sexually antagonistic loci, leading to stronger gender load and reduced net benefits of sexual selection[@Connallon_Hall_2016].
+By contrast, in spatially or temporally variable environments, sexual selection can enhance local adaptation.
+For example, in Darwin’s finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating[@Huber_2007].
+Despite the potential interaction between sexual selection and environmental variability in diversification, comparative tests are currently lacking.
 
-Conversely, sexual selection might hinder speciation or make extinction more likely by favouring traits that improve mating success but reduce population fitness.
-For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the reproductive success of individuals while reducing population-wide productivity [e.g., infanticide, harassment, and reduced parental care by mate-seeking males; @RN37; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016].
-Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013].
-
-Importantly, the relationship between sexual selection and diversification might be contingent on the environment.
-Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [e.g.
-@Long_2012; @Connallon_Hall_2016], though the empirical evidence for this theory is mixed and largely limited to insects [@Cally_2019].
-The mechanism underpinning this prediction is that the environment and the selective history of the population should influence the relative amounts of genetic variation at sexually concordant loci (i.e.
-loci where the same allele is fittest for both sexes) and sexually antagonistic loci (where different alleles are favoured in each sex).
-In stable environments, variation is thought to be preferentially depleted at sexually concordant loci, leading to stronger gender load and reduced net benefits of sexual selection [e.g.
-@Connallon_Hall_2016], and potentially to a more negative relationship between sexual selection and diversification rates across species.
-Conversely, sexual selection might be especially important at promoting local adaptation when the environment displays variability at the relevant spatial and temporal scales [e.g.
-@Boughman_2002; @Lorch_2003; @Maan_2011; @kokko2008condition]
-
-Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [e.g.
-@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014].
-Birds are well-suited for this purpose due to their diverse and well-characterised mating systems and sexually-selected traits, and because their phylogenetic relationships are comparatively well-known.
+Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014] because their biology and phylogenetic relationships are comparatively well-known.
 A 2011 meta-analysis covering 20 primary studies of birds and other taxa found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects or fish [@Kraaijeveld_2011].
 However, there was large variation in effect size estimates across the 20 studies, likely reflecting differences in methodology, such as metrics used to characterise speciation and sexual selection, in addition to true biological differences.
-More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012].
-Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.
+More recently, Huang & Rabosky [@Huang_2014] found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny[@Jetz_2012].
+Similarly, Cooney et al.
+[@Cooney_2017] found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.
 
 Here, we investigate the association between sexual selection and diversification in birds while building upon previous approaches in multiple ways.
-We use two measures of the strength of sexual selection: sexual dichromatism [*n* = 5,812 species; estimated from illustrations by @Dale_2015], as well as an index of male-biased sexual selection [*n* = 2,465 species; derived via principal components analysis by @Dale_2015], which captures (co)variation in sexual size dimorphism, social polygyny and paternal care.
-These two measures are used because sexual dichromatism does not always signal the presence of strong sexual selection and _vice versa_ [@Dale_2015].
-For example, male and female dunnocks (_Prunella modularis_) are similarly coloured yet sexual selection appears to be strong [@davies1986re].
+We use two measures of the strength of sexual selection: sexual dichromatism [@Dale_2015], as well as an index of male-biased sexual selection [@Dale_2015], which captures (co)variation in sexual size dimorphism, social polygyny and paternal care.
+We use these two measures because sexual dichromatism does not always signal the presence of strong sexual selection and *vice versa* [@Dale_2015].
+For example, male and female dunnocks (*Prunella modularis*) are similarly coloured yet sexual selection appears to be strong [@davies1986re].
 Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018].
-Additionally, our analysis includes multiple ecological and environmental variables on diversification, allowing us to adjust for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016].
-We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; see, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).
+Additionally, our analysis includes multiple ecological and environmental variables, allowing us to adjust for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016].
+We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [@Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, namely diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$) [@Jetz_2012].
 \phantomsection
@@
 # Results
+
+We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds).
+We calculated three different tip-rate metrics of speciation and one of extinction for a maximum clade credibility (MCC) tree and, to account for phylogenetic uncertainty, 1,000 trees for $\lambda_{DR}$ and $\lambda_{ND}$ and 100 trees for $\lambda_{BAMM}$ [@Rabosky_2014].
+$\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path.
+Alternatively, $\lambda_{BAMM}$ uses a Bayesian approach to assess the probability of evolutionary rate-shift configurations, from which it generates tip-rate speciation and extinction estimates.
+
+To test the association between speciation/extinction and sexual selection, environmental variability and their interaction, we fitted phylogenetic least squares (PGLS) models with $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ as the response variable.
+Predictors included one measure of sexual selection (one of two measures of sexual dichromatism or the index of male-biased sexual selection), five environmental measures, and 2-way interactions between the measure of sexual selection and each of the environmental measures, with subsequent model simplification using AIC model selection.
+
+The two measures of sexual dichromatism were from previously published data, one based on RGB (red-green-blue) values from images in *Handbook of the Birds of the World* [@delhoyo2011] (*n* = 5,983), and the other based on spectrophotometry and avian colour space (*n* = 581) [@Armenta_2008].
+The index of male-biased sexual selection is associated with greater sexual size dimorphism, social polygyny and lack of paternal care (*n* = 2,465).
+
+The five environmental variables were (*i*) the log-transformed range size; (*ii*) the average Net Primary Productivity (NPP) in each species’ range; (*iii*) mean temperature seasonality (BIO4) for each range; (*iv*) variation in temperature across a species’ range and (*v*) long-term climate variation, which primarily reflects temperature differences between the last interglacial and current climates.
+We included an estimate of climate variability during recent evolutionary history as it may be a better indicator of environmental effects on speciation than present-day environmental variability.
 \phantomsection
 
-## Variability across phylogenetic trees and speciation rate measures
+## Male-biased sexual selection, but not sexual dichromatism, affects speciation
 
-**[Is it possible to move this later in the Results?
-It's possibly not ideal to start by highlighting the dependence of our conclusions on the methods we choose, and not mentioning our main non-null finding regarding the SS index.
-Having said that I can see a case for leaving it here too.
-- Luke] [We initially had this at the end but moved it up to the start.
-What do you think Devi?
-- Justin]**
+We found a significant positive association between the index of male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ for the MCC tree ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
+The distribution of estimates from PGLS models on 1,000 trees was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
 
-Estimates of the effect of predictor variables on speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % highest posterior density (HPD) interval across phylogenetic least squares (PGLS) model estimates from 100 trees was often > 20 times larger than the 95 % confidence interval for estimates from a single PGLS model using the MCC (maximum clade credibility) tree.
-This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for pgls model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
-Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar statistical support is notable.
-Mean measures of speciation rate across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
-Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
-Full details of the BAMM results are presented as supplementary materials.
-
-\phantomsection
-## No evidence that the relationship between sexual selection and speciation depends on environmental variability
-
-**[I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Abstract and Intro.
-- Luke] [Statistically, I think it is best to rule out the interactions first but Yes I see your point - Justin]**
-
-We next fit PGLS models with 2-way interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) were included in the full model for posterior model selection.
-However, no interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\Delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
-Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
-Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
-\phantomsection
-
-## No evidence that sexual dichromatism affects speciation
-
-We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation.
-Using three different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) as the response variable, the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree.
-Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
-Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
+In contrast to male-biased sexual selection, we found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation.
+Sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}), $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a) or $\lambda_{BAMM}$ ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
 PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; i.e.
@@
 For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
-\phantomsection
 
-## Male-biased sexual selection increases speciation rate
-
-We found a significant positive association between male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ for the MCC tree ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-The distribution of estimates from PGLS models on 1,000 trees was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (HPD Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`).
-Complete HPD intervals for models using male-bias sexual selection PPCA as a predictor can be found within \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}.
+No interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and the index of male-biased sexual selection; $\Delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
+Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
+Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
 \phantomsection
@@
 Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed equivocal evidence that range size is negatively associated with speciation.
-Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subset for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
+Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subset for species with an index of male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
 Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
@@
 We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+\phantomsection
+
+## Variability across phylogenetic trees and speciation rate measures
+
+Estimates of the effect of predictor variables on speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % HPD interval across PGLS model estimates from 100 trees was often > 20 times larger than the 95 % confidence interval for estimates from a single PGLS model using the MCC tree.
+This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for pgls model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
+Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar statistical support is notable.
+Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
+Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
+Full details of the BAMM results are presented as supplementary materials.
 
@@
 (**c**) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection (*n* = 2,465).
-Similar to (**a**), (**d**) presents model estimates for PGLS analyses using a restricted dataset with measures of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
+Similar to (**a**), (**d**) presents model estimates for PGLS analyses using a restricted dataset with measures of an index of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
 Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
@@
 
-![Speciation rate ($\lambda_{DR}$) across all passerine birds (*n* = 5,965) with estimates of sexual dichromatism, range size available for 5,812 species and male-biased sexual selection available for 2,465 species.
+![Speciation rate ($\lambda_{DR}$) across all passerine birds (*n* = 5,965) with estimates of sexual dichromatism, range size available for 5,812 species and an index of male-biased sexual selection available for 2,465 species.
 Across these species there was a small but significant negative association between $\lambda_{DR}$ and log-range size as well as a significant positive association between $\lambda_{DR}$ and male-biased sexual selection but no significant association between $\lambda_{DR}$ and sexual dichromatism based on RGB measures.
@@
 The numeric values are standardised regression slopes and the asterisks indicates that the 95% confidence intervals of this estimate do not overlap with zero.
-The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and male-biased sexual selection measures (*n* = 2,465).
+The confidence intervals were obtained from 500 bootstrapped iterations and the data used in this analysis was subset to species with both sexual dichromatism and an index of male-biased sexual selection measures (*n* = 2,465).
 \label{Path_Analysis}](Figures/Path_Analysis.pdf)
@@
 These findings reaffirm the conclusions of previous, smaller studies in which sexual dichromatism was measured using spectrophotometry [@Huang_2014] or human observers [@Cooney_2017].
-The correlation between speciation rate and the index of male-biased sexual selection (which encapsulates variation in sexual size dimorphism, social polygyny, and paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$, although the estimated direction and magnitude of the correlation was broadly similar across all three measures.
+The correlation between speciation rate and the index of male-biased sexual selection (which encapsulates variation in sexual size dimorphism, social polygyny, and paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$.
 Interestingly, we also found a consistent negative relationship between range size and speciation rate, at least when this rate was quantified using $\lambda_{DR}$ and $\lambda_{ND}$.
@@
 The difference in findings between the analyses of sexual dichromatism _versus_ the index of sexual selection is noteworthy, because the majority of earlier studies used dichromatism alone as their proxy for sexual selection [e.g., @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
-Given our findings, and the modest correlation between dichromatism and the sexual selection index [$r = 0.34$; @Dale_2015], we hypothesise that sexual dichromatism may not be a robust proxy for sexual selection.
-Although dichromatism clearly reflects sexual selection to some extent, it is possibly that it is too indirect a measure to detectany association with speciation rate, even with a high sample size.
+Given our findings, and the modest correlation between dichromatism and the sexual selection index [$r = 0.34$; @Dale_2015], we suggest that sexual dichromatism may not be a robust proxy for sexual selection.
+Although dichromatism clearly reflects sexual selection to some extent, it may be too indirect a measure to detect any association with speciation rate, even with large sample size.
 There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
@@
 One outcome of our analyses was that different measures of speciation rates presented different results.
-This is not completely surprising, because each of the rates is calculated differently (see @Rabosky_diversification_2018).
+This is not completely surprising, because each of the rates is calculated differently [@Rabosky_diversification_2018].
 For instance, $\lambda_{DR}$ is weighted more towards speciation events close to the tips and allows more rate heterogeneity compared to BAMM estimates.
@@
 
-To summarise, we have shown that male-biased sexual selection, but not sexual dichromatism, predicts speciation in passerines, and that this relationship is not markedly affected by environmental variability.
+To summarise, we have shown that male-biased sexual selection, but not sexual dichromatism, predicts speciation in passerines, that the magnitude of this effect is modest, and that this relationship is not markedly affected by environmental variability.
 These findings imply that sexual dechromatism is not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
@@
 We used a previously-published measure of sexual dichromatism for 5,983 species of passerines [@Dale_2015].
-Briefly, @Dale_2015 obtained sex-specific RGB (red-green-blue) values across six body patches (nape, crown, forehead, throat, upper breast, and lower breast) from _Handbook of the Birds of the World_ [@delhoyo2011].
+Briefly, Dale et al.
+[@Dale_2015] obtained sex-specific RGB (red-green-blue) values across six body patches (nape, crown, forehead, throat, upper breast, and lower breast) from _Handbook of the Birds of the World_ [@delhoyo2011].
 The relative contribution of male and female RGB colour values were averaged across body patches and provide 'male-like' and 'female-like' plumage scores.
@@
 
-Because sexual dichromatism is presumably imperfectly correlated with variation in the strength of sexual selection across taxa, we sourced an additional measure of sexual selection, also from @Dale_2015, referred to here as the 'index of male-biased sexual selection'.
+Because sexual dichromatism is presumably imperfectly correlated with variation in the strength of sexual selection across taxa, we sourced an additional measure of sexual selection [@Dale_2015], referred to here as the 'index of male-biased sexual selection'.
 This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
@@
 From these values, we obtained means and standard deviations for each species.
-Using the same spatial sampling, we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial [LIG; 120,000 - 140,000 years ago; @Otto-Bliesner_2006].
-We include estimates of climate variability during recent evolutionary history as they may be a better indicators of environmental effects on speciation than present-day environmental variability.
-Furthermore, to estimate variability in the energy available to species, we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution.
-Estimates of NPP had 30-second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 [MOD17A3; @Zhao_2005].
+Using the same spatial sampling, we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006].
+To estimate variability in the energy available to species, we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution.
+Estimates of NPP had 30-second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005].
 Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
@@
 To obtain seasonal climate variation we used (*iii*) mean values of temperature seasonality (BIO4) for each range.
-(*iv*) To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO 4 and BIO15).
+(*iv*) To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO4 and BIO15).
 PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
@@
 We used a maximum clade credibility (MCC) tree from 2,500 samples of the posterior distribution (*n* = 5,965) as the main phylogenetic hypothesis in our comparative analysis.
-Additionally, a random draw of full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures (1000 trees) and BAMM (100 trees)[@Rabosky_2014].
+Additionally, a random draw of full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures (1,000 trees) and BAMM (100 trees)[@Rabosky_2014].
 These trees used a 'Hackett backbone' [@Hackett_2008] and were constructed using a pure birth (Yule) model.
@@
 Firstly, we obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
-Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [e.g., @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each edge towards the root having the length down-weighted.
+Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each edge towards the root having the length down-weighted.
 Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
@@
 To test the association between speciation/extinction and sexual selection, environmental variability and their interaction, we used phylogenetic least squares (PGLS) models in the ``nlme`` package [@nlme_2018].
-Firstly, we conducted model selection to compare models in which $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ were the response variable: these tip-rate estimates all came from the same MCC tree (derived from 2,500 draws of the posterior distribution in @Jetz_2012).
+Firstly, we conducted model selection to compare models in which $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ were the response variable: these tip-rate estimates all came from the same MCC tree (derived from 2,500 draws of the posterior distribution [@Jetz_2012]).
 For models of $\lambda_{BAMM}$ and $\mu_{BAMM}$ we used the inverse of the variance associated with each tip rate estimate as weights, to account for the variable precision of the estimates provided by BAMM.
@@
 This method enabled us to present model estimates for an MCC tree alongside 1,000/100 trees from the posterior distribution of trees to account for phylogenetic uncertainty.
-This approach was repeated on three datasets corresponding to each measure of sexual selection: dichromatism derived from RGB values of images (*n* = 5,812); dichromatism from spectrophotometry (*n* = 581) and the multivariate measure of male-biased sexual selection (*n* = 2,465).
+This approach was repeated on three datasets corresponding to each measure of sexual selection: dichromatism derived from RGB values of images (*n* = 5,812); dichromatism from spectrophotometry (*n* = 581) and the index of male-biased sexual selection (*n* = 2,465).
 
-Finally, using the subset of species with measurements of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
+Finally, using the subset of species with an index of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
 The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
@@
 Further details of the path analysis, including our rationale for each path's directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
-
-\newpage\phantomsection
+\phantomsection
 

### 2019-08-11 Manuscript_Publication.Rmd
addition of polygyny - speciation rate reference

@@
 output:
+  pdf_document: default
   word_document: default
-  pdf_document: default
 header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
@@
 For example, species with costly sexual signals may be less resilient to environmental change[@Kokko_Brooks_2003].
-Sexual selection might also promote extinction by causing maladaptation (‘gender load’) in female traits that are genetically correlated with sexually-selected male traits[@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013].
-Although numerous studies have examined the relationship between sexual selection and speciation or extinction rates[@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Kraaijeveld_2011; @Seddon_2013; @Huang_2014], the availability of more complete phenotypic, ecological and phylogenetic data[@Jetz_2012], together with significant advances in phylogenetic methods[@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.
+Extinction risk may also be exacerbated by sexual selection causing maladaptation (‘gender load’) in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013].
+Although numerous studies have examined the relationship between sexual selection and speciation or extinction rates [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Kraaijeveld_2011; @Seddon_2013; @Huang_2014], the availability of more complete phenotypic, ecological and phylogenetic data[@Jetz_2012], together with significant advances in phylogenetic methods[@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.
 
@@
 [@Cooney_2017] found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.
+However, another study using the prevelance of polygyny (as oppossed to sexual dichromatism) as a proxy of sexual selection found a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].
 
@@
 We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds).
-We calculated three different tip-rate metrics of speciation and one of extinction for a maximum clade credibility (MCC) tree and, to account for phylogenetic uncertainty, 1,000 trees for $\lambda_{DR}$ and $\lambda_{ND}$ and 100 trees for $\lambda_{BAMM}$ [@Rabosky_2014].
+We calculated three different tip-rate metrics of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one of extinction ($\mu_{BAMM}$) for a maximum clade credibility (MCC) tree and, to account for phylogenetic uncertainty, 1,000 trees for $\lambda_{DR}$ and $\lambda_{ND}$ and 100 trees for $\lambda_{BAMM}$ [@Rabosky_2014].
 $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path.

### 2019-10-02 Manuscript_Publication.Rmd
Manuscript ammendments

@@
 author:
-- Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§}
-  and Iliana Medina\textsuperscript{§}
-- '*\textit{justin.g.cally@gmail.com}'
+# - Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§}
+#   and Iliana Medina\textsuperscript{§}
+# - '*\textit{justin.g.cally@gmail.com}'
 bibliography: References/SS_Extinction.bib
@@
   \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
-  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
+  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{draft,colorlinks=false, hyperfootnotes=false,
   linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
@@
 subparagraph: yes
-abstract: Sexual selection is thought to shape phylogenetic diversity by affecting
-  speciation or extinction rates. However, the net effect of sexual selection on diversification
-  is hard to predict _a priori_, because many of the hypothesised effects on speciation
-  or extinction have opposing signs and uncertain magnitudes. Here, we test whether
-  the strength of sexual selection predicts variation in speciation and extinction
-  rates across passerine birds (up to 5,812 species, covering most genera). We tested
-  for associations between speciation or extinction and two measures of sexual selection
-  (sexual dichromatism and a multivariate measure of male-biased sexual selection),
-  accounting for range size and measures of environmental variability. Male-biased
-  sexual selection, but not sexual dichromatism, predicted speciation rates ($\lambda_{DR}$)
-  in passerines. This relationship was independent of range size or environmental
-  variability, though species with smaller ranges had higher speciation rates. There
-  was no association between sexual selection and extinction rate ($\mu_{BAMM}$).
-  Our findings show that sexual dichromatism is a poor proxy for sexual selection
-  at least a broad-scale and support the view that sexual selection has shaped diversification
-  in songbirds.
+abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates.
+However, the net effect of sexual selection on diversification is hard to predict _a priori_, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes.
+Here, we test whether the strength of sexual selection predicts variation in speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera).
+We use two measures of sexual selection (sexual dichromatism and a multivariate measure of male-biased sexual selection) and adjust for range size and environmental variability.
+Male-biased sexual selection, but not sexual dichromatism, predicted speciation rates ($\lambda_{DR}$) in passerines.
+This relationship was independent of range size or environmental variability, though species with smaller ranges had higher speciation rates.
+There was no association between sexual selection and extinction rate ($\mu_{BAMM}$).
+Our findings suggest that sexual dichromatism is a poor proxy for sexual selection at least a broad-scale and support the view that sexual selection has shaped diversification in songbirds.
 ---
-\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia}
+<!-- \maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia} -->
 
 ```{r, echo=FALSE}
+#devtools::install_github('jpshanno/ecoflux')
+library(xtable)
+
 MCC.DR.top <- readRDS('data/MCC.DR.top.rds')
@@
 
-\newpage\phantomsection
+\newpage
 
-<!-- Luke general comments:
-In a few places you say things like "1000 rrandom trees".
-To me this means trees with a random topology (which are sometimes used as some sort of null model), but I think you actually mean 1000 trees randomly sampled (without replacement) from the posterior of trees in Jetz et al.
-Is that right?
-If so, probably best to search for the word "random" and remove/reword it.
-
-I think it might make sense to call it "the index of male-biased sexual selection" throughout, including in the figures.
-Otherwise it's not clear when you mean the general phenomenon vs our measure of it.
-See what you think after reading my edits.
--->
-
-# Introduction
-
-Sexual selection is a fundamental evolutionary process; yet there is long-standing debate about how it shapes patterns of species diversity[@Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018].
-Sexual selection can promote speciation because it operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection[@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia].
-Sexual selection could also promote speciation or prevent extinction by purging deleterious mutations[@Whitlock_Agrawal_2009], fixing beneficial ones[@Whitlock_2000], and accelerating adaptation in different environments[@Lorch_2003; @Candolin_2008; @Cally_2019].
-Conversely, sexual selection might hinder speciation or make extinction more likely by favouring traits that improve mating success but reduce population fitness[@Rankin_2011; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016].
-For example, species with costly sexual signals may be less resilient to environmental change[@Kokko_Brooks_2003].
+Sexual selection is a ubiquitous evolutionary process whose effect on diversification is much debated [@Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018].
+Sexual selection can promote speciation because it operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia].
+Sexual selection could also promote speciation or prevent extinction by purging deleterious mutations[@Whitlock_Agrawal_2009], fixing beneficial ones [@Whitlock_2000], and accelerating adaptation in different environments [@Lorch_2003; @Candolin_2008; @Cally_2019].
+Conversely, sexual selection might hinder speciation or make extinction more likely by favouring traits that improve mating success but reduce population fitness [@Rankin_2011; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016].
+For example, species with costly sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003].
 Extinction risk may also be exacerbated by sexual selection causing maladaptation (‘gender load’) in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013].
-Although numerous studies have examined the relationship between sexual selection and speciation or extinction rates [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Kraaijeveld_2011; @Seddon_2013; @Huang_2014], the availability of more complete phenotypic, ecological and phylogenetic data[@Jetz_2012], together with significant advances in phylogenetic methods[@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.
+Although numerous studies have examined the relationship between sexual selection and speciation or extinction rates [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Kraaijeveld_2011; @Seddon_2013; @Huang_2014], the availability of more complete phenotypic, ecological and phylogenetic data [@Jetz_2012], together with significant advances in phylogenetic methods [@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.
 
@@
 For example, in Darwin’s finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating[@Huber_2007].
-Despite the potential interaction between sexual selection and environmental variability in diversification, comparative tests are currently lacking.
+Despite the potential interaction between sexual selection and environmental variability in diversification, phylogenetic tests are currently lacking.
 
@@
 Similarly, Cooney et al.
-[@Cooney_2017] found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.
-However, another study using the prevelance of polygyny (as oppossed to sexual dichromatism) as a proxy of sexual selection found a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].
+[@Cooney_2017] found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.More recently, social polygyny (a proxy for sexual selection) was found to have a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].
 
@@
 We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [@Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, namely diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$) [@Jetz_2012].
+Our results show that (*i*) a composite measure of sexual selection, but not sexual dichromatism, significantly predicts speciation rates, (*ii*) species with smaller ranges have higher speciation rates and (*iii*) there is no evidence that environmental variables or their interaction with sexual selection have an impact on diversification rates.
+Therefore, we provide evidence at a very large scale that sexual selection can have positive effects on diversification in the largest radiation of birds.
+Furthermore, we suggest that the use of sexual dichromatism as the sole proxy for sexual selection should be reconsidered, since it can misguide conclusions about the role of sexual selection in different scenarios.
 \phantomsection
@@
 # Results
-
-We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds).
-We calculated three different tip-rate metrics of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one of extinction ($\mu_{BAMM}$) for a maximum clade credibility (MCC) tree and, to account for phylogenetic uncertainty, 1,000 trees for $\lambda_{DR}$ and $\lambda_{ND}$ and 100 trees for $\lambda_{BAMM}$ [@Rabosky_2014].
-$\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path.
-Alternatively, $\lambda_{BAMM}$ uses a Bayesian approach to assess the probability of evolutionary rate-shift configurations, from which it generates tip-rate speciation and extinction estimates.
-
-To test the association between speciation/extinction and sexual selection, environmental variability and their interaction, we fitted phylogenetic least squares (PGLS) models with $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ as the response variable.
-Predictors included one measure of sexual selection (one of two measures of sexual dichromatism or the index of male-biased sexual selection), five environmental measures, and 2-way interactions between the measure of sexual selection and each of the environmental measures, with subsequent model simplification using AIC model selection.
-
-The two measures of sexual dichromatism were from previously published data, one based on RGB (red-green-blue) values from images in *Handbook of the Birds of the World* [@delhoyo2011] (*n* = 5,983), and the other based on spectrophotometry and avian colour space (*n* = 581) [@Armenta_2008].
-The index of male-biased sexual selection is associated with greater sexual size dimorphism, social polygyny and lack of paternal care (*n* = 2,465).
-
-The five environmental variables were (*i*) the log-transformed range size; (*ii*) the average Net Primary Productivity (NPP) in each species’ range; (*iii*) mean temperature seasonality (BIO4) for each range; (*iv*) variation in temperature across a species’ range and (*v*) long-term climate variation, which primarily reflects temperature differences between the last interglacial and current climates.
-We included an estimate of climate variability during recent evolutionary history as it may be a better indicator of environmental effects on speciation than present-day environmental variability.
 \phantomsection
@@
 
-We found a significant positive association between the index of male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ for the MCC tree ($\beta$ = `r formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; $\lambda_{BAMM}$: $\beta$ = `r formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 3)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 3)`; \autoref{model_results}b).
-The distribution of estimates from PGLS models on 1,000 trees was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 3)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 3)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 3)`; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds).
+To do this, we tested the association between speciation/extinction and sexual selection, environmental variability and their interaction using phylogenetic least squares (PGLS) models with $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ as the response variable.
+These are three different tip-rate metrics of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one of extinction ($\mu_{BAMM}$, see Methods for details).
+Predictors included one measure of sexual selection (one of two measures of sexual dichromatism or the index of male-biased sexual selection), four measures of environmental variation, range size and 2-way interactions between the measure of sexual selection and each of the environmental measures including range size, with subsequent model simplification using AIC model selection.
+
+We found a significant positive association between the index of male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ maximum credibility (MCC) tree ($\beta$ = `r sanitize.numbers(formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b).
+However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; $\lambda_{BAMM}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b).
+When we took into account phylogenetic uncertainty by running the models using 1,000 trees, the distribution of estimates from PGLS models was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r sanitize.numbers(formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
 
 In contrast to male-biased sexual selection, we found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation.
-Sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}), $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a) or $\lambda_{BAMM}$ ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a).
+Sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r sanitize.numbers(formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 2)`; \autoref{model_results}a, \autoref{Speciation_Figure}), $\lambda_{ND}$ ($\beta$ = `r sanitize.numbers(formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 2)`; \autoref{model_results}a) or $\lambda_{BAMM}$ ($\beta$ = `r sanitize.numbers(formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 2)`; \autoref{model_results}a).
 PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; i.e.
@@
 Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 trees for $\lambda_{BAMM}$.
-The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top$SDi, format = "e", digits = 3)`, $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top$SDi, format = "e", digits = 3)`,\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
-For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism ($\lambda_{DR}$: HPD Interval = `r sanitize.numbers(formatC(hpd.DR.top$SDi, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, $\lambda_{ND}$: HPD Interval = `r sanitize.numbers(formatC(hpd.ND.top$SDi, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, \autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
+For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r sanitize.numbers(formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
 
 No interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and the index of male-biased sexual selection; $\Delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
-Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation.
-Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
+Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation or range size.
+Furthermore, we found no evidence that these environmental factors --- seasonal temperature variation, long-term temperature variation, spatial temperature variation, and Net Primary Productivity (NPP) --- predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
 \phantomsection
@@
 Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation, we found a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$.
-This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}).
-This association was also evident across the estimates from models using the 1,000 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a).
+This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r sanitize.numbers(formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 2)`; $\lambda_{ND}$: $\beta$ = `r sanitize.numbers(formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 2)`; \autoref{model_results}a, \autoref{Speciation_Figure}).
+This association was also evident across the estimates from models using the 1,000 trees ($\lambda_{DR}$: HPD Interval = `r sanitize.numbers(formatC(hpd.DR.top[[2]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; $\lambda_{ND}$: HPD Interval = `r sanitize.numbers(formatC(hpd.ND.top[[2]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \autoref{model_results}a).
 Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed equivocal evidence that range size is negatively associated with speciation.
@@
 
-We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r sanitize.numbers(formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 2)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
 \phantomsection
@@
 Estimates of the effect of predictor variables on speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % HPD interval across PGLS model estimates from 100 trees was often > 20 times larger than the 95 % confidence interval for estimates from a single PGLS model using the MCC tree.
-This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for pgls model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
+This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
 Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar statistical support is notable.
@@
 Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation and should not be interpreted.
+Illustrations reproduced by permission of Lynx Editions.
 \label{Speciation_Figure}](Figures/DRSpeciationFigure.pdf)
@@
 We found evidence that the composite index of male-biased sexual selection, but not measures of sexual dichromatism, is correlated with the rate of speciation in passerine birds.
-The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB), and it cannot be explained by a difference in statistical power since the sample size for the dichromatism analyses was much larger.
+The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB), and it cannot be explained by a difference in statistical power since the sample size for the dichromatism analyses was larger.
 These findings reaffirm the conclusions of previous, smaller studies in which sexual dichromatism was measured using spectrophotometry [@Huang_2014] or human observers [@Cooney_2017].
@@
 Given our findings, and the modest correlation between dichromatism and the sexual selection index [$r = 0.34$; @Dale_2015], we suggest that sexual dichromatism may not be a robust proxy for sexual selection.
-Although dichromatism clearly reflects sexual selection to some extent, it may be too indirect a measure to detect any association with speciation rate, even with large sample size.
+Although dichromatism almost certainly provides some insight into the operation of sexual selection, it may be too indirect a measure to detect any association with speciation rate, even with large sample size.
 There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
@@
 
-In-line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least when speciation is measured by $\lambda_{DR}$.
+In line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least when speciation is measured by $\lambda_{DR}$.
 Additionally, we found that this association appears to be independent of net primary productivity and spatiotemporal variation in the environment.
@@
 To summarise, we have shown that male-biased sexual selection, but not sexual dichromatism, predicts speciation in passerines, that the magnitude of this effect is modest, and that this relationship is not markedly affected by environmental variability.
-These findings imply that sexual dechromatism is not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
+These findings imply that sexual dichromatism is not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
 Our results also add indirect suppport to the hypothesis that sexual selection promotes adaptation, which has implications for conservation [@Holman_Kokko_2013] and captive breeding programs for threatened species [@Charge_2014].
@@
 
-\newpage\phantomsection
+\phantomsection
 # Materials and methods
@@
 Here we use the absolute difference between male and female plumage scores as an estimate of sexual dichromatism.
+Technically, this measures differences in the ‘degree of male-ness’ between males and females, rather than sex differences in colour _per se_ (i.e.
+dichromatism in the strict sense).
+For example, the metric would fail to capture dichromatism when both the male and female possess a single, but differently coloured 'male-like' patch.
+However, the metric is highly correlated with dichromatism measured from spectral data (see below).
+
 Additionally, we used another measure of dichromatism corresponding to colour distance in avian colour space derived from spectral data [@Armenta_2008].
@@
 
-# Acknowledgements
+<!-- # Acknowledgements -->
 
-We would like to acknowledge Adnan Moussalli for his assistance in conducting BAMM runs across 100 phylogenetic trees.
+<!-- We thank Adnan Moussalli for his assistance in conducting BAMM runs across 100 phylogenetic trees, Ignacio Quintero for providing BAMM results for comparison and for discussion of the results as well as Chris Cooney for comments on the manuscript.
+-->
 

### 2019-10-10 Manuscript_Publication.Rmd
Evolution formatting

@@
 author:
-# - Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§}
-#   and Iliana Medina\textsuperscript{§}
-# - '*\textit{justin.g.cally@gmail.com}'
 bibliography: References/SS_Extinction.bib
 output:
-  pdf_document: default
   word_document: default
+  pdf_document:
+    keep_tex: yes
 header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
@@
   \raggedbottom
-csl: References/NComms.csl
+csl: References/evolution.csl
 subparagraph: yes
@@
 \newpage
+\phantomsection  
+# Introduction
 
@@
 Sexual selection can promote speciation because it operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia].
-Sexual selection could also promote speciation or prevent extinction by purging deleterious mutations[@Whitlock_Agrawal_2009], fixing beneficial ones [@Whitlock_2000], and accelerating adaptation in different environments [@Lorch_2003; @Candolin_2008; @Cally_2019].
+Sexual selection could also promote speciation or prevent extinction by purging deleterious mutations [@Whitlock_Agrawal_2009], fixing beneficial ones [@Whitlock_2000], and accelerating adaptation in different environments [@Lorch_2003; @Candolin_2008; @Cally_2019].
 Conversely, sexual selection might hinder speciation or make extinction more likely by favouring traits that improve mating success but reduce population fitness [@Rankin_2011; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016].
@@
 The relationship between sexual selection and diversification may depend on the environment.
-Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones[@Long_2012; @Connallon_Hall_2016].
+Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [@Long_2012; @Connallon_Hall_2016].
 In stable environments, consistent selection depletes genetic variation at sexually concordant loci (i.e.
 loci where the same allele is fittest for both sexes).
-In these environments, genetic variation remains disproportionately at sexually antagonistic loci, leading to stronger gender load and reduced net benefits of sexual selection[@Connallon_Hall_2016].
+In these environments, genetic variation remains disproportionately at sexually antagonistic loci, leading to stronger gender load and reduced net benefits of sexual selection [@Connallon_Hall_2016].
 By contrast, in spatially or temporally variable environments, sexual selection can enhance local adaptation.
-For example, in Darwin’s finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating[@Huber_2007].
+For example, in Darwin’s finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
 Despite the potential interaction between sexual selection and environmental variability in diversification, phylogenetic tests are currently lacking.
@@
 However, there was large variation in effect size estimates across the 20 studies, likely reflecting differences in methodology, such as metrics used to characterise speciation and sexual selection, in addition to true biological differences.
-More recently, Huang & Rabosky [@Huang_2014] found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny[@Jetz_2012].
-Similarly, Cooney et al.
-[@Cooney_2017] found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.More recently, social polygyny (a proxy for sexual selection) was found to have a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].
+More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012].
+Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.More recently, social polygyny (a proxy for sexual selection) was found to have a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].
 
@@
 Furthermore, we suggest that the use of sexual dichromatism as the sole proxy for sexual selection should be reconsidered, since it can misguide conclusions about the role of sexual selection in different scenarios.
+
+\phantomsection
+# Materials and methods
+
+We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds).
+Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions.
+Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Electronic Supplementary Material (ESM)}.
+\phantomsection
+
+## Compiling data for sexual selection and environmental stress
+\phantomsection
+
+### Sexual dichromatism
+
+We used a previously-published measure of sexual dichromatism for 5,983 species of passerines [@Dale_2015].
+Briefly, Dale et al.
+[@Dale_2015] obtained sex-specific RGB (red-green-blue) values across six body patches (nape, crown, forehead, throat, upper breast, and lower breast) from _Handbook of the Birds of the World_ [@delhoyo2011].
+The relative contribution of male and female RGB colour values were averaged across body patches and provide 'male-like' and 'female-like' plumage scores.
+Here we use the absolute difference between male and female plumage scores as an estimate of sexual dichromatism.
+Technically, this measures differences in the ‘degree of male-ness’ between males and females, rather than sex differences in colour _per se_ (i.e.
+dichromatism in the strict sense).
+For example, the metric would fail to capture dichromatism when both the male and female possess a single, but differently coloured 'male-like' patch.
+However, the metric is highly correlated with dichromatism measured from spectral data (see below).
+
+Additionally, we used another measure of dichromatism corresponding to colour distance in avian colour space derived from spectral data [@Armenta_2008].
+These measurements include variation in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from museum specimens as opposed to illustrations.
+The spectrophotometry data covers only 581 passerine species (10-fold fewer than the RGB data), although there was a substantial correlation between the two dichromatism measures ($r = 0.79$; \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
+\phantomsection
+
+### Male-biased sexual selection
+
+Because sexual dichromatism is presumably imperfectly correlated with variation in the strength of sexual selection across taxa, we sourced an additional measure of sexual selection [@Dale_2015], referred to here as the 'index of male-biased sexual selection'.
+This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
+This measure of male-biased sexual selection is available for only 2,465 species, and shows a moderate correlation with the RGB measure of sexual dichromatism ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
+\phantomsection
+
+### Environmental variables
+
+We obtained estimates of species range size using expert range maps [@birdlife_2017].
+Because of taxonomic changes to 1,230 species in the Birdlife database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
+From these distributions, we obtained estimates of climatic conditions that each species is exposed to.
+That is, we took 1,000 random point samples in each range and extracted 19 present-day bioclimatic variables (representing a variety of biologically relevant annual trends in temperature and precipitation) with 30-second (~1 km^2^) spatial resolution [@Fick_2017].
+From these values, we obtained means and standard deviations for each species.
+Using the same spatial sampling, we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006].
+To estimate variability in the energy available to species, we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution.
+Estimates of NPP had 30-second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005].
+Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
+\phantomsection
+
+### Generating biologically relevant predictors for environmental stress
+
+Given that stressful environments are expected to interact with sexual selection and have a positive effect on adaptation [@Cally_2019], we used the extracted environmental variables from each species range size to create predictors of environmental variation/stress.
+Firstly we used (*i*) the average NPP in each species' range and (*v*) the log-transformed range size as potentially informative predictors of speciation rates.
+We also used three environmental predictors derived from bioclimatic data.
+These predictors relate to seasonal climate variation, spatial climate variation and long-term climate variation.
+To obtain seasonal climate variation we used (*iii*) mean values of temperature seasonality (BIO4) for each range.
+(*iv*) To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO4 and BIO15).
+PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
+Species range is a potentially informative predictor of speciation and extinction, so we controlled for the correlation between environmental spatial variation and range size --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
+To obtain long-term variation in climates for each species range we take (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
+Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
+The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
+Details and ``R`` code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.
+\phantomsection
+
+## Estimating extinction and speciation
+
+Phylogenetic information was obtained from www.birdtree.org [@Jetz_2012].
+We used a maximum clade credibility (MCC) tree from 2,500 samples of the posterior distribution (*n* = 5,965) as the main phylogenetic hypothesis in our comparative analysis.
+Additionally, a random draw of full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures (1,000 trees) and BAMM (100 trees) [@Rabosky_2014].
+These trees used a 'Hackett backbone' [@Hackett_2008] and were constructed using a pure birth (Yule) model.
+We calculated three different tip-rate metrics of speciation and one of extinction across all trees.
+
+Firstly, we obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
+Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each edge towards the root having the length down-weighted.
+Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
+Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path.
+These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE); but, based on previous simulation studies, $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
+
+Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014].
+This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate probability distributions of evolutionary rate-shift configurations with variable speciation and extinction rates.
+Importantly, these models provide tip-rate estimates of speciation and extinction rate.
+The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{ESM}; briefly, we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
+Each BAMM model (one MCC tree and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
+Each run of BAMM reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200 (an arbitrary value, above which posterior distributions can often be accurately inferred) for the key model parameters of log-likelihood and number of rate shifts (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
+Further details of BAMM parameters and output are available in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}, with tip-rate means and variances provided.
+Additionally, given the variability in BAMM estimates, we also provide analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on the MCC tree from a genetic-only phylogeny across all birds [@Harvey_2017].
+All analyses were conducted on log-rates.
+\phantomsection
+
+## Phylogenetic comparative analysis
+
+To test the association between speciation/extinction and sexual selection, environmental variability and their interaction, we used phylogenetic least squares (PGLS) models in the ``nlme`` package [@nlme_2018].
+Firstly, we conducted model selection to compare models in which $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ were the response variable: these tip-rate estimates all came from the same MCC tree (derived from 2,500 draws of the posterior distribution [@Jetz_2012]).
+For models of $\lambda_{BAMM}$ and $\mu_{BAMM}$ we used the inverse of the variance associated with each tip rate estimate as weights, to account for the variable precision of the estimates provided by BAMM.
+The most complicated model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (i.e.
+log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP), and all of the 2-way interactions between sexual selection and each of the environmental measures.
+The simpler models contained all of the same main effects, but had fewer 2-way interaction terms (potentially none).
+Model selection was done in ``MuMIn`` using the ``dredge`` function [@MuMIn_2017].
+Using the terms from the top-ranked model (ranked by AICc), we ran the equivalent model for each of the 1,000 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and each of the 100 trees used to derive $\lambda_{BAMM}$ and $\mu_{BAMM}$.
+In each model we used the unique response variables and phylogenetic tree correlation structure.
+Specifically, for models using tip-rate metrics ($\lambda_{DR}$, $\lambda_{ND}$) we estimated the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using the ``corPagel`` function in the ``ape`` package [@ape_2004] independently for each of the 1,000 trees/models.
+Alternatively, for models using speciation and extinction estimates derived using BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$) we found $\lambda$ was consistently estimated at 1 and hence assumed Brownian motion (using the ``corBrownian`` function) to estimate the correlation structure.
+This method enabled us to present model estimates for an MCC tree alongside 1,000/100 trees from the posterior distribution of trees to account for phylogenetic uncertainty.
+This approach was repeated on three datasets corresponding to each measure of sexual selection: dichromatism derived from RGB values of images (*n* = 5,812); dichromatism from spectrophotometry (*n* = 581) and the index of male-biased sexual selection (*n* = 2,465).
+
+Finally, using the subset of species with an index of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
+The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
+That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection, and range size to have effects on each other and not just on speciation rate.
+To minimise path complexity we used temperature seasonality (BIO4) as the single measure for environmental variability, $\lambda_{DR}$ as the single measure of speciation and the tip-rates from the MCC tree.
+Further details of the path analysis, including our rationale for each path's directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
 \phantomsection
@@
 
-\phantomsection
-# Materials and methods
-
-We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds).
-Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions.
-Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Electronic Supplementary Material (ESM)}.
-\phantomsection
-
-## Compiling data for sexual selection and environmental stress
-\phantomsection
-
-### Sexual dichromatism
-
-We used a previously-published measure of sexual dichromatism for 5,983 species of passerines [@Dale_2015].
-Briefly, Dale et al.
-[@Dale_2015] obtained sex-specific RGB (red-green-blue) values across six body patches (nape, crown, forehead, throat, upper breast, and lower breast) from _Handbook of the Birds of the World_ [@delhoyo2011].
-The relative contribution of male and female RGB colour values were averaged across body patches and provide 'male-like' and 'female-like' plumage scores.
-Here we use the absolute difference between male and female plumage scores as an estimate of sexual dichromatism.
-Technically, this measures differences in the ‘degree of male-ness’ between males and females, rather than sex differences in colour _per se_ (i.e.
-dichromatism in the strict sense).
-For example, the metric would fail to capture dichromatism when both the male and female possess a single, but differently coloured 'male-like' patch.
-However, the metric is highly correlated with dichromatism measured from spectral data (see below).
-
-Additionally, we used another measure of dichromatism corresponding to colour distance in avian colour space derived from spectral data [@Armenta_2008].
-These measurements include variation in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from museum specimens as opposed to illustrations.
-The spectrophotometry data covers only 581 passerine species (10-fold fewer than the RGB data), although there was a substantial correlation between the two dichromatism measures ($r = 0.79$; \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
-\phantomsection
-
-### Male-biased sexual selection
-
-Because sexual dichromatism is presumably imperfectly correlated with variation in the strength of sexual selection across taxa, we sourced an additional measure of sexual selection [@Dale_2015], referred to here as the 'index of male-biased sexual selection'.
-This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
-This measure of male-biased sexual selection is available for only 2,465 species, and shows a moderate correlation with the RGB measure of sexual dichromatism ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
-\phantomsection
-
-### Environmental variables
-
-We obtained estimates of species range size using expert range maps [@birdlife_2017].
-Because of taxonomic changes to 1,230 species in the Birdlife database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
-From these distributions, we obtained estimates of climatic conditions that each species is exposed to.
-That is, we took 1,000 random point samples in each range and extracted 19 present-day bioclimatic variables (representing a variety of biologically relevant annual trends in temperature and precipitation) with 30-second (~1 km^2^) spatial resolution [@Fick_2017].
-From these values, we obtained means and standard deviations for each species.
-Using the same spatial sampling, we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006].
-To estimate variability in the energy available to species, we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution.
-Estimates of NPP had 30-second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005].
-Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
-\phantomsection
-
-### Generating biologically relevant predictors for environmental stress
-
-Given that stressful environments are expected to interact with sexual selection and have a positive effect on adaptation [@Cally_2019], we used the extracted environmental variables from each species range size to create predictors of environmental variation/stress.
-Firstly we used (*i*) the average NPP in each species' range and (*v*) the log-transformed range size as potentially informative predictors of speciation rates.
-We also used three environmental predictors derived from bioclimatic data.
-These predictors relate to seasonal climate variation, spatial climate variation and long-term climate variation.
-To obtain seasonal climate variation we used (*iii*) mean values of temperature seasonality (BIO4) for each range.
-(*iv*) To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO4 and BIO15).
-PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
-Species range is a potentially informative predictor of speciation and extinction, so we controlled for the correlation between environmental spatial variation and range size --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
-To obtain long-term variation in climates for each species range we take (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
-Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
-The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
-Details and ``R`` code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.
-\phantomsection
-
-## Estimating extinction and speciation
-
-Phylogenetic information was obtained from www.birdtree.org [@Jetz_2012].
-We used a maximum clade credibility (MCC) tree from 2,500 samples of the posterior distribution (*n* = 5,965) as the main phylogenetic hypothesis in our comparative analysis.
-Additionally, a random draw of full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures (1,000 trees) and BAMM (100 trees)[@Rabosky_2014].
-These trees used a 'Hackett backbone' [@Hackett_2008] and were constructed using a pure birth (Yule) model.
-We calculated three different tip-rate metrics of speciation and one of extinction across all trees.
-
-Firstly, we obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
-Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each edge towards the root having the length down-weighted.
-Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
-Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path.
-These tip-rate measures are alternatives to state-dependent diversification models such as Quantitative State Speciation-Extinction (QuaSSE); but, based on previous simulation studies, $\lambda_{DR}$ and $\lambda_{ND}$ are robust and intuitive measures that provide high power and low false discovery rate with large phylogenies when incorporated into Phylogenetic Generalized Least Squares (PGLS) models [@Harvey_essim_2017].
-
-Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014].
-This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate probability distributions of evolutionary rate-shift configurations with variable speciation and extinction rates.
-Importantly, these models provide tip-rate estimates of speciation and extinction rate.
-The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{ESM}; briefly, we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
-Each BAMM model (one MCC tree and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
-Each run of BAMM reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200 (an arbitrary value, above which posterior distributions can often be accurately inferred) for the key model parameters of log-likelihood and number of rate shifts (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
-Further details of BAMM parameters and output are available in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}, with tip-rate means and variances provided.
-Additionally, given the variability in BAMM estimates, we also provide analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on the MCC tree from a genetic-only phylogeny across all birds [@Harvey_2017].
-All analyses were conducted on log-rates.
-\phantomsection
-
-## Phylogenetic comparative analysis
-
-To test the association between speciation/extinction and sexual selection, environmental variability and their interaction, we used phylogenetic least squares (PGLS) models in the ``nlme`` package [@nlme_2018].
-Firstly, we conducted model selection to compare models in which $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ were the response variable: these tip-rate estimates all came from the same MCC tree (derived from 2,500 draws of the posterior distribution [@Jetz_2012]).
-For models of $\lambda_{BAMM}$ and $\mu_{BAMM}$ we used the inverse of the variance associated with each tip rate estimate as weights, to account for the variable precision of the estimates provided by BAMM.
-The most complicated model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (i.e.
-log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP), and all of the 2-way interactions between sexual selection and each of the environmental measures.
-The simpler models contained all of the same main effects, but had fewer 2-way interaction terms (potentially none).
-Model selection was done in ``MuMIn`` using the ``dredge`` function [@MuMIn_2017].
-Using the terms from the top-ranked model (ranked by AICc), we ran the equivalent model for each of the 1,000 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and each of the 100 trees used to derive $\lambda_{BAMM}$ and $\mu_{BAMM}$.
-In each model we used the unique response variables and phylogenetic tree correlation structure.
-Specifically, for models using tip-rate metrics ($\lambda_{DR}$, $\lambda_{ND}$) we estimated the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using the ``corPagel`` function in the ``ape`` package [@ape_2004] independently for each of the 1,000 trees/models.
-Alternatively, for models using speciation and extinction estimates derived using BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$) we found $\lambda$ was consistently estimated at 1 and hence assumed Brownian motion (using the ``corBrownian`` function) to estimate the correlation structure.
-This method enabled us to present model estimates for an MCC tree alongside 1,000/100 trees from the posterior distribution of trees to account for phylogenetic uncertainty.
-This approach was repeated on three datasets corresponding to each measure of sexual selection: dichromatism derived from RGB values of images (*n* = 5,812); dichromatism from spectrophotometry (*n* = 581) and the index of male-biased sexual selection (*n* = 2,465).
-
-Finally, using the subset of species with an index of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
-The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
-That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection, and range size to have effects on each other and not just on speciation rate.
-To minimise path complexity we used temperature seasonality (BIO4) as the single measure for environmental variability, $\lambda_{DR}$ as the single measure of speciation and the tip-rates from the MCC tree.
-Further details of the path analysis, including our rationale for each path's directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
-\phantomsection
-
+<!-- \phantomsection -->
 <!-- # Acknowledgements -->

### 2019-10-10 Sexual Selection and Speciation.Rmd
Evolution formatting

@@
 ---
-title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation
-  in birds
-author: "Justin G. Cally, Devi Stuart-Fox, Luke Holman, Iliana Medina Guzmán"
+title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds
+author: 
+# "Justin G. Cally, Devi Stuart-Fox, Luke Holman, Iliana Medina Guzmán"
 csl: References/evolution.csl
@@
     code_folding: hide # awesome buttons to show/hide the code
-subtitle: Supplementary Material
+subtitle: Electronic Supplementary Material
 bibliography: References/SS_Extinction.bib

### 2020-05-30 Manuscript_Publication.Rmd
Updated supp material

@@
 output:
-  word_document: default
   pdf_document:

### 2020-05-30 Sexual Selection and Speciation.Rmd
Updated supp material

@@
 library(phylopath)
+library(sjPlot)
+library(sjmisc)
 if(!requireNamespace("BiocManager", quietly = TRUE)){
@@
 ```{r trycatch all models, eval = F}
-
 #Read in DR models and extract estimates:
@@
   theme_minimal()
+
+#Set rownames to match tree
+rownames(SS.subset) <- SS.subset$TipLabel
 ```
@@
 saveRDS(pruned.MCC.Subset.tree, 'data/pruned.MCC.Subset.tree.rds')
-
-#Set rownames to match tree
-rownames(SS.subset) <- SS.subset$TipLabel
 
@@
 
-```{r random calcs two, eval = FALSE, echo = F}
+## Testing the sexual dichromatism dataset using the same data points as the sexual selection dataset
+
+In our analysis we conducted tested for the effect of sexual dichromatism on speciation using a dataset with a sample size of 5,812 passerine bird species.
+However, when testing for the effects of sexual selection (using a principal component score) we only utilised 2,465 species.
+This discrepancy is because data for the principal components analysis is incomplete for many species.
+Here we conduct an analysis on the effects of sexual dichromatism on speciation using the same model structure as in our primary analysis, however using the same data/species as we used in the analysis of the effect of sexual selection (*n = 2,465*).
+We conduct the analyses on MCC trees for $\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$ and compare the results to the analyses conducted on 5,812 passerine bird species.
+
+```{r subset SDi, eval = TRUE, warning=FALSE, message=FALSE, fig.height=12, fig.width=8}
+if(!file.exists("data/sample_size_comparison.rds")){
+# Read in tree 
+pruned.MCC.Subset.tree <- readRDS('data/pruned.MCC.Subset.tree.rds')
+
 #Run subset model with SDi
 #Run model for DR
-Subset.MCC.DR.SDi <- gls(MCC.DR ~ SDi
+Subset_SDi_Models <- list()
+
+Subset_SDi_Models[["DR"]] <- gls(MCC.DR ~ SDi
                                + log(range.size.m2)
@@
                                + NPP,
-                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = FALSE), 
+                      correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = F), #Subset Tree 
                       data = SS.subset, 
@@
 
+#Run model for ND
+Subset_SDi_Models[["ND"]] <- gls(MCC.ND ~ SDi
+                               + log(range.size.m2)
+                               + bioclim4 #Seasonal variation
+                               + residuals.PC1 #Spatial variation
+                               + PC1.LIG #Long-term climate variation
+                               + NPP,
+                         correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = F), #Subset Tree 
+                         data = SS.subset, 
+                         method = "REML")
+
+#Run model for ND
+Subset_SDi_Models[["BAMM"]] <- gls(mean.lambda ~ SDi
+                                 + log(range.size.m2)
+                                 + bioclim4 #Seasonal variation
+                                 + residuals.PC1 #Spatial variation
+                                 + PC1.LIG #Long-term climate variation
+                                 + NPP,
+                           weights = ~ sqrt(var.lambda),
+                           correlation = corBrownian(phy = pruned.MCC.Subset.tree), #Subset Tree
+                           data = SS.subset, 
+                           method = "REML")
+
+# Main analysis models 
+Main_SDi_Models <- list()
+
+Main_SDi_Models[["DR"]] <- readRDS('data/MCC.DR.top.rds')
+Main_SDi_Models[["ND"]] <- readRDS('data/MCC.ND.top.rds')
+Main_SDi_Models[["BAMM"]] <- readRDS('data/MCC.Lambda.top.rds')
+
+sample_size_comparison <- list(Subset_SDi_Models = Subset_SDi_Models, Main_SDi_Models = Main_SDi_Models)
+saveRDS(sample_size_comparison, "data/sample_size_comparison.rds")
+} else {
+  sample_size_comparison  <- readRDS("data/sample_size_comparison.rds")
+}
+
+plot_mods <- function(model1, model2) {
+
+}
+sjplot_custom <- function(sub, full, type, min, max) {
+  theme_set(theme_sjplot())
+plot <-  sjPlot::plot_models(sub, 
+                      full, 
+                      show.p = FALSE, 
+                      transform = NULL, 
+                      std.est = "std2", 
+                      m.labels = c("Subset data (n = 2,465)", "Full data (n = 5,812)"), 
+                      legend.title = "Data Points Used",
+                      title = paste("Comparing", type, "models with subset data")) + 
+    ylim(min, max) +
+    theme_sjplot2() + 
+  scale_color_sjplot("simply")
+  return(plot)
+}
+
+grid.arrange(
+sjplot_custom(sample_size_comparison[["Subset_SDi_Models"]][["DR"]], 
+              sample_size_comparison[["Main_SDi_Models"]][["DR"]], 
+              type = "DR", min = -0.025, max = 0.025),
+sjplot_custom(sample_size_comparison[["Subset_SDi_Models"]][["ND"]], 
+              sample_size_comparison[["Main_SDi_Models"]][["ND"]], 
+              type = "ND", min = -0.00075, max = 0.00075),
+sjplot_custom(sample_size_comparison[["Subset_SDi_Models"]][["BAMM"]], 
+              sample_size_comparison[["Main_SDi_Models"]][["BAMM"]], 
+              type = "Lambda-BAMM", min = -0.002, max = 0.002))
+
 #stargazer(Subset.MCC.DR.SDi, type = "html")
 ```
+
+**Figure S17:** A comparison of full and subset data was used in models testing for the effect of sexual dichromatism on speciation rate ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$).
+While there is variation between model estimates due to the different data points used, these differences do not lead us to make alternive conclusions.
+In all three cases the partial regression coefficient for sexual dichromatism (SDi) has 95 % confidence intervals that overlap zero, thus we have no reason to suspect that the effect that this study found of sexual selection on speciation is due to sampling bias.
 

### 2020-08-18 Manuscript_Publication.Rmd
addition of manuscript changes after review, new figure and extra analyses

@@
 ---
-title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation
-  in birds
+title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds
 author:
@@
   \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
-  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{draft,colorlinks=false, hyperfootnotes=false,
+  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{draft, colorlinks=false, hyperfootnotes=false,
   linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
@@
 However, the net effect of sexual selection on diversification is hard to predict _a priori_, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes.
-Here, we test whether the strength of sexual selection predicts variation in speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera).
-We use two measures of sexual selection (sexual dichromatism and a multivariate measure of male-biased sexual selection) and adjust for range size and environmental variability.
-Male-biased sexual selection, but not sexual dichromatism, predicted speciation rates ($\lambda_{DR}$) in passerines.
-This relationship was independent of range size or environmental variability, though species with smaller ranges had higher speciation rates.
-There was no association between sexual selection and extinction rate ($\mu_{BAMM}$).
-Our findings suggest that sexual dichromatism is a poor proxy for sexual selection at least a broad-scale and support the view that sexual selection has shaped diversification in songbirds.
+**Theoretical work also suggests that the net effect of sexual selection on diversification should depend strongly on ecological factors, though this prediction has seldom been tested.
+Here, we test whether variation in sexual selection can predict speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera) and whether this relationship is mediated by environmental factors.
+Male-biased sexual selection, and specifically sexual size dimorphism, predicted two of the three measures of speciation rates that we examined**.
+The link we observed between sexual selection and speciation was independent of environmental variability, though species with smaller ranges had higher speciation rates.
+There was no association between sexual selection and extinction rate.
+**Our findings support the view that male-biased sexual selection, as measured by predictors of male-male competition, has shaped diversification in the largest radiation of birds.**
 ---
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, message = FALSE, warning = FALSE}
 #devtools::install_github('jpshanno/ecoflux')
 library(xtable)
+library(dplyr)
 
@@
 
+# Additional analyses on indidual PPCA components: HPD Intervals
+nd.wing <- readRDS('data/Subset.ND.pgls.summaryA.rds')[1:300] %>% 
+  dplyr::bind_rows() %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>%
+  dplyr::filter(Parameter == 'abs(SSD_wing)') %>%
+  dplyr::pull(Estimate) %>%
+  coda::as.mcmc(.) %>%
+  coda::HPDinterval(., prob = 0.95)
+
+nd.mating <- readRDS('data/Subset.ND.pgls.summary_mating.rds')[1:300] %>% 
+  dplyr::bind_rows() %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>%
+  dplyr::filter(Parameter == 'as.factor(mating_system)1') %>%
+  dplyr::pull(Estimate) %>%
+  coda::as.mcmc(.) %>%
+  coda::HPDinterval(., prob = 0.95)
+
+nd.paternal <- readRDS('data/Subset.ND.pgls.summary_paternal.rds') %>% 
+  dplyr::bind_rows() %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>%
+  dplyr::filter(Parameter == 'as.factor(paternal_care)1') %>%
+  dplyr::pull(Estimate) %>%
+  coda::as.mcmc(.) %>%
+  coda::HPDinterval(., prob = 0.95)
+
+dr.wing <- readRDS('data/Subset.DR.pgls.summary_sswing.rds') %>% 
+  dplyr::bind_rows() %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>%
+  dplyr::filter(Parameter == 'abs(SSD_wing)') %>%
+  dplyr::pull(Estimate) %>%
+  coda::as.mcmc(.) %>%
+  coda::HPDinterval(., prob = 0.95)
+
+dr.mating <- readRDS('data/Subset.DR.pgls.summary_mating.rds') %>% 
+  dplyr::bind_rows() %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>%
+  dplyr::filter(Parameter == 'as.factor(mating_system)1') %>%
+  dplyr::pull(Estimate) %>%
+  coda::as.mcmc(.) %>%
+  coda::HPDinterval(., prob = 0.95)
+
+dr.paternal <- readRDS('data/Subset.DR.pgls.summary_paternal.rds') %>% 
+  dplyr::bind_rows() %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>%
+  dplyr::filter(Parameter == 'as.factor(paternal_care)1') %>%
+  dplyr::pull(Estimate) %>%
+  coda::as.mcmc(.) %>%
+  coda::HPDinterval(., prob = 0.95)
+
+b.wing <- readRDS('data/Subset.Lambda.pgls.summaryB.rds') %>% 
+  dplyr::bind_rows() %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>%
+  dplyr::filter(Parameter == 'abs(SSD_wing)') %>%
+  dplyr::pull(Estimate) %>%
+  coda::as.mcmc(.) %>%
+  coda::HPDinterval(., prob = 0.95)
+
+b.mating <- readRDS('data/Subset.Lambda.pgls.summary_mating.rds') %>% 
+  dplyr::bind_rows() %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>% 
+  dplyr::filter(Parameter == 'as.factor(mating_system)1') %>%
+  dplyr::pull(Estimate) %>%
+  coda::as.mcmc(.) %>%
+  coda::HPDinterval(., prob = 0.95)
+
+b.paternal <- readRDS('data/Subset.Lambda.pgls.summary_paternal.rds') %>% 
+  dplyr::bind_rows() %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>%
+  dplyr::filter(Parameter == 'as.factor(paternal_care)1') %>%
+  dplyr::pull(Estimate) %>%
+  coda::as.mcmc(.) %>%
+  coda::HPDinterval(., prob = 0.95)
+
 #path analysis
@@
 
-Sexual selection is a ubiquitous evolutionary process whose effect on diversification is much debated [@Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018].
+Sexual selection is a ubiquitous evolutionary process whose effect on phylogenic diversification is much debated **[@Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018; @tsuji2020sexual]**.
 Sexual selection can promote speciation because it operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia].
@@
 Extinction risk may also be exacerbated by sexual selection causing maladaptation (‘gender load’) in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013].
+
 Although numerous studies have examined the relationship between sexual selection and speciation or extinction rates [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Kraaijeveld_2011; @Seddon_2013; @Huang_2014], the availability of more complete phenotypic, ecological and phylogenetic data [@Jetz_2012], together with significant advances in phylogenetic methods [@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.
+**Furthermore, the diversity of outcomes and approaches in previous studies suggests that the association between species diversity and sexual selection is far from clear (reviewed in @tsuji2020sexual).**
 
-The relationship between sexual selection and diversification may depend on the environment.
-Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [@Long_2012; @Connallon_Hall_2016].
+**A possible reason for the above uncertainty regarding the relationship between sexual selection and diversification is that this relationship may strongly depend on the environment.** Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [@Long_2012; @Connallon_Hall_2016].
 In stable environments, consistent selection depletes genetic variation at sexually concordant loci (i.e.
@@
 For example, in Darwin’s finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
+**Under these circumstances we predict that the effect of sexual selection on phylogenetic rates of divergence may depend on the variability of the species’ environment**.
 Despite the potential interaction between sexual selection and environmental variability in diversification, phylogenetic tests are currently lacking.
@@
 Additionally, our analysis includes multiple ecological and environmental variables, allowing us to adjust for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016].
-We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [@Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, namely diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$) [@Jetz_2012].
+We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including Bayesian Analysis of Macroevolutionary Mixtures [BAMM; @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, namely diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$) [@Jetz_2012].
 Our results show that (*i*) a composite measure of sexual selection, but not sexual dichromatism, significantly predicts speciation rates, (*ii*) species with smaller ranges have higher speciation rates and (*iii*) there is no evidence that environmental variables or their interaction with sexual selection have an impact on diversification rates.
 Therefore, we provide evidence at a very large scale that sexual selection can have positive effects on diversification in the largest radiation of birds.
-Furthermore, we suggest that the use of sexual dichromatism as the sole proxy for sexual selection should be reconsidered, since it can misguide conclusions about the role of sexual selection in different scenarios.
+Furthermore, we suggest that the use of sexual dichromatism as the sole proxy for sexual selection should be reconsidered, since it appears to be inconsistently associated with the operation of sexual selection.
 
@@
 Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions.
-Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Electronic Supplementary Material (ESM)}.
+Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}.
 \phantomsection
@@
 This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
+**The variables included in this index have all been positively linked to the intensity of sexual selection, and are usually correlated [@bjorklund_1990; @owens_1998; @dunn_2001], which is why they were combined into a single metric in previous studies [@Dale_2015]**.
 This measure of male-biased sexual selection is available for only 2,465 species, and shows a moderate correlation with the RGB measure of sexual dichromatism ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
@@
 Estimates of NPP had 30-second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005].
-Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
+Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
 \phantomsection
@@
 Given that stressful environments are expected to interact with sexual selection and have a positive effect on adaptation [@Cally_2019], we used the extracted environmental variables from each species range size to create predictors of environmental variation/stress.
-Firstly we used (*i*) the average NPP in each species' range and (*v*) the log-transformed range size as potentially informative predictors of speciation rates.
+Firstly we used (*i*) the average NPP in each species' range and (*ii*) the log-transformed range size as potentially informative predictors of speciation rates.
 We also used three environmental predictors derived from bioclimatic data.
@@
 To obtain seasonal climate variation we used (*iii*) mean values of temperature seasonality (BIO4) for each range.
-(*iv*) To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO4 and BIO15).
-PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
-Species range is a potentially informative predictor of speciation and extinction, so we controlled for the correlation between environmental spatial variation and range size --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
+(*iv*) To estimate levels of spatial environmental variation a species may endure, we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO4 and BIO15).
+PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 **largely** reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
+**A taxon’s range size often correlates with speciation and extinction rates [@Rosenzweig_1995; @Castiglione_2017]**, so we controlled for the correlation between environmental spatial variation and species' range sizes --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
 To obtain long-term variation in climates for each species range we take (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
@@
 The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
-Details and ``R`` code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}.
+Details and ``R`` code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}.
 \phantomsection
@@
 Additionally, a random draw of full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures (1,000 trees) and BAMM (100 trees) [@Rabosky_2014].
-These trees used a 'Hackett backbone' [@Hackett_2008] and were constructed using a pure birth (Yule) model.
-We calculated three different tip-rate metrics of speciation and one of extinction across all trees.
+**These trees had crown clades with a topology that was heavily constrained on the basis of a previously published study [‘Hacket backbone’; @Hackett_2008] and were constructed using a pure birth (Yule) model.** We calculated three different tip-rate metrics of speciation and one of extinction across all trees.
 
-Firstly, we obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
-Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES), also known as diversification rate (DR) [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each edge towards the root having the length down-weighted.
-Studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
+**Diversification is the result of two processes, speciation and extinction through time.
+To estimate speciation rates, we first obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.** Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES; also referred to as diversification rate/DR) **is derived from the sum of edge lengths from each tip towards the root, with each edge towards the root having the length down-weighted [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018].** Crucially, studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
 Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path.
@@
 
-Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014].
-This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate probability distributions of evolutionary rate-shift configurations with variable speciation and extinction rates.
-Importantly, these models provide tip-rate estimates of speciation and extinction rate.
-The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{ESM}; briefly, we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
-Each BAMM model (one MCC tree and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
-Each run of BAMM reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200 (an arbitrary value, above which posterior distributions can often be accurately inferred) for the key model parameters of log-likelihood and number of rate shifts (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
-Further details of BAMM parameters and output are available in the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM}, with tip-rate means and variances provided.
+Secondly, we used BAMM to model the dynamics of speciation and extinction across the **101 phylogenetic trees (one MCC tree and 100 random draws of the posterior)**.
+This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate probability distributions of evolutionary rate-shift configurations with variable speciation and extinction rates [@Rabosky_2014].
+**These models provide tip-rate estimates of speciation and extinction rate that can be easily used in comparative analyses**.
+The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{Supplementary Information}; **briefly, we used a time-variable model with the prior expected number of evolutionary rate shifts set at 100 and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
+BAMM models were run independently for the 101 phylogenetic trees for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
+Important BAMM parameters (log-likelihood and number of rate shifts) reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200; an arbitrary value, above which posterior distributions can often be accurately inferred** (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
+Further details of BAMM parameters and output are available in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}, with tip-rate means and variances provided.
 Additionally, given the variability in BAMM estimates, we also provide analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on the MCC tree from a genetic-only phylogeny across all birds [@Harvey_2017].
@@
 For models of $\lambda_{BAMM}$ and $\mu_{BAMM}$ we used the inverse of the variance associated with each tip rate estimate as weights, to account for the variable precision of the estimates provided by BAMM.
-The most complicated model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (i.e.
+**For each response variable, we conducted model selection to compare models with different combinations of predictor variables.
+** The most complicated model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (i.e.
 log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP), and all of the 2-way interactions between sexual selection and each of the environmental measures.
@@
 Using the terms from the top-ranked model (ranked by AICc), we ran the equivalent model for each of the 1,000 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and each of the 100 trees used to derive $\lambda_{BAMM}$ and $\mu_{BAMM}$.
-In each model we used the unique response variables and phylogenetic tree correlation structure.
+**Additionally, we investigated the effect of the individual principal component scores for the index of male-biased sexual selection on speciation rate.
+For these pgls models we replaced the composite index score with the individual biological variable (sexual size dimorphism, social polygyny and [lack of] paternal care) and ran the equivalent model for 300 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and 100 trees used to derive $\lambda_{BAMM}$**.
+
+**Across all our analyses we corrected for the phylogenetic signal.
+Our models used the unique response variables and correlation structure for a given phylogenetic tree**.
 Specifically, for models using tip-rate metrics ($\lambda_{DR}$, $\lambda_{ND}$) we estimated the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using the ``corPagel`` function in the ``ape`` package [@ape_2004] independently for each of the 1,000 trees/models.
@@
 
-Finally, using the subset of species with an index of male-biased sexual selection, we conducted a phylogenetic path analysis using the phylopath ``R`` package [@Wouter_2018].
+Finally, using the subset of species with an index of male-biased sexual selection, we conducted a phylogenetic path analysis using the ``phylopath R`` package [@Wouter_2018].
 The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
@@
 To minimise path complexity we used temperature seasonality (BIO4) as the single measure for environmental variability, $\lambda_{DR}$ as the single measure of speciation and the tip-rates from the MCC tree.
-Further details of the path analysis, including our rationale for each path's directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{ESM} along with all other analyses and the relevant ``R`` code to reproduce results.
+Further details of the path analysis, including our rationale for each path's directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information} along with all other analyses and the relevant ``R`` code to reproduce results.
 \phantomsection
@@
 
-We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds).
+We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds; \autoref{Speciation_Figure}).
 To do this, we tested the association between speciation/extinction and sexual selection, environmental variability and their interaction using phylogenetic least squares (PGLS) models with $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ as the response variable.
@@
 Predictors included one measure of sexual selection (one of two measures of sexual dichromatism or the index of male-biased sexual selection), four measures of environmental variation, range size and 2-way interactions between the measure of sexual selection and each of the environmental measures including range size, with subsequent model simplification using AIC model selection.
+**In all cases the best model had a $\Delta$ AICc above 20 against the second model, and an evidence ratio above 100.**
 
@@
 When we took into account phylogenetic uncertainty by running the models using 1,000 trees, the distribution of estimates from PGLS models was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r sanitize.numbers(formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+
+**Given that this male-biased sexual selection index comprises three different (but correlated) variables, we investigated which of these variables was driving the association observed with $\lambda_{DR}$.
+Our results over 300 trees showed that this pattern is mainly driven by the sexual size dimorphism component (HPD Interval = `r sanitize.numbers(formatC(dr.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), with the effects of other components overlapping zero; paternal care (HPD Interval = `r sanitize.numbers(formatC(dr.paternal, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`) and mating system (HPD Interval= `r sanitize.numbers(formatC(dr.mating, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`).
+Importantly, the association between sexual size dimorphism and speciation rates is also present when using $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(nd.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), but not when using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(b.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`, \autoref{SS_components}).**
 
@@
 
+**Our analyses also show that the differences in results between sexual dichromatism and male-biased sexual selection (i.e.
+association with speciation rates only for the latter) were not due to differences in the size of the datasets used (5,812 species vs.
+2,465,** \href{https://justincally.github.io/SexualSelection_Speciation/#testing_the_sexual_dichromatism_dataset_using_the_same_data_points_as_the_sexual_selection_dataset}{Figure S17}).
 No interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and the index of male-biased sexual selection; $\Delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
@@
 Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed equivocal evidence that range size is negatively associated with speciation.
-Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subset for species with an index of male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
+Range size is **significantly associated with** $\lambda_{DR}$ (\autoref{model_results}b) using data subset for species with an index of male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
 Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
@@
 This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
-Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar statistical support is notable.
+Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, **and so it is concerning that there are such large differences in the BAMM estimates between alternative, similarly plausible phylogenies**.
 Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
@@
 Full details of the BAMM results are presented as supplementary materials.
-
-![Model estimates (**a**) for the effect of log-range size and sexual dichromatism on speciation and extinction rates using PGLS analyses with the sexual dichromatism dataset (RGB values, *n* = 5,812).
-(**b**) depicts the scatter plot of speciation rate ($\lambda_{DR}$) and log-range size with the model estimate presented as a dashed line.
-(**c**) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection (*n* = 2,465).
-Similar to (**a**), (**d**) presents model estimates for PGLS analyses using a restricted dataset with measures of an index of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
-Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
-The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 1,000 randomly sampled trees (for $\lambda_{DR}$ and $\lambda_{ND}$) or 100 randomly sampled trees for $\lambda_{BAMM}$ and $\mu_{BAMM}$ can be found in the ESM.
-Density curves are based on model estimates from 1,000/100 trees and the circle below with error bars is the estimate and 95% CIs from the MCC tree.
-For this figure we removed outliers from estimates coming from the 100 randomly sampled trees for BAMM models in order to be able to visualise the MCC 95% CIs.
-\label{model_results}](Figures/Estimates_and_scatter_edit.pdf)
 
@@
 \label{Speciation_Figure}](Figures/DRSpeciationFigure.pdf)
+
+![Model estimates (**a**) **showing the effect size (i.e.
+slope) of log-range size** and sexual dichromatism on speciation and extinction rates using PGLS analyses with the sexual dichromatism dataset (RGB values, *n* = 5,812).
+(**b**) depicts the scatter plot of speciation rate ($\lambda_{DR}$) and log-range size with the model estimate presented as a dashed line.
+(**c**) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection (*n* = 2,465).
+Similar to (**a**), (**d**) presents model estimates for PGLS analyses using a restricted dataset with measures of an index of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
+Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
+The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 1,000 randomly sampled trees (for $\lambda_{DR}$ and $\lambda_{ND}$) or 100 randomly sampled trees for $\lambda_{BAMM}$ and $\mu_{BAMM}$ can be found in the Supplementary Information.
+Density curves are based on model estimates from 1,000/100 trees and the circle below with error bars is the estimate and 95% CIs from the MCC tree.
+For this figure we removed outliers from estimates coming from the 100 randomly sampled trees for BAMM models in order to be able to visualise the MCC 95% CIs.
+\label{model_results}](Figures/Estimates_and_scatter_edit.pdf)
+
+![Estimates of the effect of individual sexual selection components included in the PPCA (paternal care and wing size dimorphism) on three measures of speciation rate ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$).
+Estimates are presented as density intervals from pgls models on 300 phylogentic trees that used species with available data for these sexual selection measures (*n* = 2,465).
+The bar under each density ridge is the 95 % Highest Posterior Density Interval.
+\label{SS_components}](Figures/Individual_SS_Estimates.pdf)
 
@@
 We found evidence that the composite index of male-biased sexual selection, but not measures of sexual dichromatism, is correlated with the rate of speciation in passerine birds.
-The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB), and it cannot be explained by a difference in statistical power since the sample size for the dichromatism analyses was larger.
+The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB), **and it cannot be explained by a difference in statistical power or sampling**.
 These findings reaffirm the conclusions of previous, smaller studies in which sexual dichromatism was measured using spectrophotometry [@Huang_2014] or human observers [@Cooney_2017].
 The correlation between speciation rate and the index of male-biased sexual selection (which encapsulates variation in sexual size dimorphism, social polygyny, and paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$.
+**This pattern seems to be mainly driven by an association between sexual size dimorphism and speciation**.
 Interestingly, we also found a consistent negative relationship between range size and speciation rate, at least when this rate was quantified using $\lambda_{DR}$ and $\lambda_{ND}$.
-None of the bioclimatic measures of environmental variability that we investigated (i.e., temperature seasonality, long-term temperature variation, and spatial temperature variation) significantly predicted speciation rate.
+None of the bioclimatic measures of environmental variability that we investigated (i.e., temperature seasonality, long-term temperature variation, and spatial temperature variation) **were significantly associated speciation rate, nor mediated the relationship between sexual selection and diversification**.
 
-The difference in findings between the analyses of sexual dichromatism _versus_ the index of sexual selection is noteworthy, because the majority of earlier studies used dichromatism alone as their proxy for sexual selection [e.g., @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
-Given our findings, and the modest correlation between dichromatism and the sexual selection index [$r = 0.34$; @Dale_2015], we suggest that sexual dichromatism may not be a robust proxy for sexual selection.
+The difference in findings between the analyses of sexual dichromatism _versus_ the index of male-biased sexual selection is noteworthy, because the majority of earlier studies used dichromatism alone as their proxy for sexual selection [e.g., @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014].
+Given our findings, and the modest correlation between dichromatism and the sexual selection index [$r = 0.34$; @Dale_2015], we suggest that sexual dichromatism may not be a robust proxy for sexual selection [@Cooney_2018].
 Although dichromatism almost certainly provides some insight into the operation of sexual selection, it may be too indirect a measure to detect any association with speciation rate, even with large sample size.
 There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic.
-Firstly, sexual dichromatism can evolve for reasons other than sexual selection, such as when males and females occupy different ecological niches [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989] or experience different selective pressures in contexts other than competition for mates [@Price_2014].
+Sexual dichromatism can evolve for reasons other than sexual selection, such as when males and females occupy different ecological niches [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989] or experience different selective pressures in contexts other than competition for mates [@Price_2014].
 For example, in superb fairy-wrens (*Malurus cyaneus*) female colouration has probably evolved in response to spatial variation in predation pressure, increasing dichromatism [@Medina_2017].
-Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit weakly.
+**In fact, our path analysis detected a weak relationship between environment and sexual dichromatism, where sexual dichromatism was positively predicted by temperature seasonality (a measure of environmental variation).**
+
+<!-- Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit weakly.
 Secondly, colour is only one of the traits subject to sexual selection [@Miles_2018; @Cally_2019].
@@
 Given the importance of honest signaling in sexual selection and the variability in colour production mechanisms across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.
+-->
 
 In line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least when speciation is measured by $\lambda_{DR}$.
-Additionally, we found that this association appears to be independent of net primary productivity and spatiotemporal variation in the environment.
+**Many of the species that have both high scores of male-biased sexual selection and high diversification rates belong to the genera _Ploceus_, _Euplectes_ (Ploceidae) and _Paradisaea_ (Paradiaseidae).
+Multiple weaver species (Ploceidae) are polygynous and lack paternal care, and both weavers and birds of paradise have strong size dimorphism.
+The association between speciation rates and principal component scores that we report seems to be mainly driven by sexual size dimorphism and, to a lesser extent, paternal care.
+Speciation rates (both $\lambda_{DR}$ and $\lambda_{ND}$) are higher in species with larger sexual dimorphism and $\lambda_{DR}$ also has a tendency to be higher in species with no paternal care.
+Size dimorphism is often thought to arise as a consequence of intrasexual competition, where one of the sexes (males in most birds) has to compete for access to the other sex, leading to selection for larger body sizes and thus greater dimorphism [@bjorklund_1990; @owens_1998].
+Therefore, competition between males could be the underlying driver of the high speciation rates that we detect in some clades.**
+
+**Sexual selection due to competition within sexes contrasts with the drivers of sexual dichromatism.
+Plumage dichromatism can evolve as a consequence of female cryptic choice and be related to extra-bond fertilizations, but not necessarily paternal care or mating system [@owens_1998].
+It can also arise as a result of selection on the level of crypsis of the sex that cares for offspring [@Dale_2015].
+The fact that traits linked with competition (such as size dimorphism) are the ones associated with higher $\lambda_{DR}$ values – rather than sexual dichromatism – supports the general view that antagonistic interactions and sexual conflict can lead to increased diversity [@Bonduriansky_2011; @Hoskin_2010; @tsuji2020sexual].
+Moreover, body size is a trait that influences multiple aspects of the physiology and ecology of a species.
+Differences in body size (as a result of sexual selection) could be linked to changes in diet, vulnerability to predators or environmental tolerance [@Damuth_1993; @Liow_2008; @Bonduriansky_2011], and such differences could ultimately increase the likelihood of divergence between young lineages.
+In mammals, sexual selection has driven the evolution of large body size which in turn has allowed diversification of ecological strategies in the clade, and higher speciation rates.**
+
+**We also found that the association between sexual selection and speciation appears to be independent of net primary productivity and spatiotemporal variation in the environment**.
 The lack of an effect of these environmental variables on speciation rate has several possible interpretations.
@@
 
-The most robust finding in our study is the finding that species with smaller ranges have elevated speciation rates.
+**We found that species with smaller ranges have elevated speciation rates**.
 This result is similar to a study of 329 amphibian genera, which found higher diversification rates in taxa with smaller range size [@Greenberg_2017].
@@
 However, the opposite pattern is also plausible because birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
+Moreover, species that have recently split as a consequence of vicariant divergence might have smaller ranges as a result of the split of the ancestral lineage, leading to link between smaller ranges and shorter divergence times.
 It is also possible that high speciation rates cause smaller range sizes, rather than the other way around, for example because repeatedly-speciating lineages tend to fill niches in ways that hinder the geographical expansion of new species [@Rosenzweig_1995; @Weir_2011; @Price_2014].
 However, species undergoing adaptive radiation in new habitats are unlikely to be limited by competition for resources from existing taxa.
-One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
+**One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.**
 
@@
 This is not completely surprising, because each of the rates is calculated differently [@Rabosky_diversification_2018].
-For instance, $\lambda_{DR}$ is weighted more towards speciation events close to the tips and allows more rate heterogeneity compared to BAMM estimates.
-This leads to greater variation in $\lambda_{DR}$ relative to the BAMM estimates, potentially explaining the difference in results.
-The BAMM estimates were also more sensitive to phylogenetic uncertainty.
+For instance, $\lambda_{DR}$ is weighted more towards speciation events close to the tips and allows more rate heterogeneity compared to $\lambda_{BAMM}$ estimates.
+**Rate shifts are unlikely to be detected in smaller clades in BAMM, meaning that it is not uncommon for whole genera to have the same rate.
+Using the $\lambda_{DR}$ metric, only sister species are guaranteed to have the same rate.
+This leads to greater variation in $\lambda_{DR}$ relative to the $\lambda_{BAMM}$ estimates, which is suggested to be an advantage when studying diversification patterns [@Quintero_2015].
+Additionally, $\lambda_{BAMM}$ estimates were also more sensitive to phylogenetic uncertainty when compared to $\lambda_{DR}$, and estimates are 20 times more variable across trees in $\lambda_{BAMM}$ compared to $\lambda_{DR}$.
+We cannot completely reject the idea that the lack of association between $\lambda_{BAMM}$ and sexual selection could be the result of low statistical power, due to the combination of both low variation across species in the speciation rates and high levels of variation in the estimates across trees.**
 
-To summarise, we have shown that male-biased sexual selection, but not sexual dichromatism, predicts speciation in passerines, that the magnitude of this effect is modest, and that this relationship is not markedly affected by environmental variability.
-These findings imply that sexual dichromatism is not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
-Our results also add indirect suppport to the hypothesis that sexual selection promotes adaptation, which has implications for conservation [@Holman_Kokko_2013] and captive breeding programs for threatened species [@Charge_2014].
-Furthermore, our finding that high speciation rate is associated with smaller range size highlights the threat to the persistence of rapidly-speciating lineages in a world with increased habitat loss and anthropogenic stress.
-This concern arises as the best predictor of extinction risk is range size [@Harris_2008] and the association found here implies that many newly-speciated clades have small range sizes and could thus be at greater risk of extinction.
+To summarise, we have shown that sexual size dimorphism (a putative proxy for male-biased sexual selection), but not sexual dichromatism, predicts speciation in passerines, that the magnitude of this effect is modest, and that this relationship is not markedly affected by environmental variability.
+We have also shown that there is no evidence of an association between sexual selection and extinction rates.
+Overall, our findings imply that male-male competition could be the mechanism driving increased speciation rates in birds, that sexual dichromatism may not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
 

### 2020-08-18 Sexual Selection and Speciation.Rmd
addition of manuscript changes after review, new figure and extra analyses

@@
 library(sjmisc)
+library(egg)
 if(!requireNamespace("BiocManager", quietly = TRUE)){
@@
 
-**Table S6:** Models with a set of weights for the response based on the inverse of the variance of the posterior distribution in $\lambda_{BAMM}$ and $\mu_{BAMM}$ have much lower AIC values, indicating that the weighting scheme improves the fit of the model.
+**Table S7:** Models with a set of weights for the response based on the inverse of the variance of the posterior distribution in $\lambda_{BAMM}$ and $\mu_{BAMM}$ have much lower AIC values, indicating that the weighting scheme improves the fit of the model.
 In any case, we found that an unweighted model did not change qualitative results.
@@
 
+<!-- ## Sexual selection components effects on speciation   -->
+
+```{r individual_components, fig.height = 8, fig.width = 7, warning = F, message = F, include=FALSE, echo = FALSE, eval = FALSE}
+#### Read files with summary results: Additional analyses were conducted seperate from this RMarkdown; Contact authors for information ----
+
+nd.wing <- readRDS('data/Subset.ND.pgls.summaryA.rds')[1:300]
+nd.mating <- readRDS('data/Subset.ND.pgls.summary_mating.rds')[1:300]
+nd.paternal <- readRDS('data/Subset.ND.pgls.summary_paternal.rds')
+dr.wing <- readRDS('data/Subset.DR.pgls.summary_sswing.rds')
+dr.mating <- readRDS('data/Subset.DR.pgls.summary_mating.rds')
+dr.paternal <- readRDS('data/Subset.DR.pgls.summary_paternal.rds')
+b.wing <- readRDS('data/Subset.Lambda.pgls.summaryB.rds')
+b.mating <- readRDS('data/Subset.Lambda.pgls.summary_mating.rds')
+b.paternal <- readRDS('data/Subset.Lambda.pgls.summary_paternal.rds')
+
+# Subset.ND.pgls.summary <- dplyr::bind_rows(b.wing, b.paternal) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+# 
+# Subset.ND.pgls.summary <- dplyr::filter(Subset.ND.pgls.summary, Parameter %in% c('abs(SSD_wing)', 'as.factor(paternal_care)1'))
+
+#HPDinterval(as.mcmc(Subset.ND.pgls.summary$Estimate), prob=0.95)
+
+#### Graph for paternal care ####
+
+Subset.ND.pgls.summary=nd.paternal
+Subset.ND.pgls.summary <- bind_rows(Subset.ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.ND.pgls.summary = filter(Subset.ND.pgls.summary,Parameter == 'as.factor(paternal_care)1')
+HPD_Interval_ND_PC <- HPDinterval(as.mcmc(Subset.ND.pgls.summary$Estimate), prob=0.95) %>% as.data.frame %>% dplyr::mutate(Parameter = 'as.factor(paternal_care)1')
+
+Subset.ND.plot <-Subset.ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6), fill="orange", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1, colour="black", fill="orange")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")+ scale_x_continuous(limits = c(-0.25, 0.25)) +
+  annotate("text", x=-0.22, y=1.5, label= expression(lambda[ND])) + xlab(" ") + ylab(" ") +
+  geom_errorbarh(data = HPD_Interval_ND_PC, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "orange", size = 1)
+
+Subset.DR.pgls.summary=dr.paternal
+Subset.DR.pgls.summary <- bind_rows(Subset.DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.DR.pgls.summary = filter(Subset.DR.pgls.summary,Parameter == 'as.factor(paternal_care)1')
+HPD_Interval_DR_PC <- HPDinterval(as.mcmc(Subset.DR.pgls.summary$Estimate), prob=0.95) %>% 
+  as.data.frame %>% 
+  dplyr::mutate(Parameter = 'as.factor(paternal_care)1')
+
+Subset.DR.plot <-Subset.DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6),fill="orange", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="orange")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")+ scale_x_continuous(limits = c(-0.25, 0.25))+
+  annotate("text", x=-0.22, y=1.5, label= expression(lambda[DR])) + xlab(" ") + ylab(" ") + 
+    geom_errorbarh(data = HPD_Interval_DR_PC, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "orange", size = 1)
+
+Subset.B.pgls.summary=b.paternal
+Subset.B.pgls.summary <- bind_rows(Subset.B.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.B.pgls.summary = filter(Subset.B.pgls.summary,Parameter == 'as.factor(paternal_care)1')
+HPD_Interval_B_PC <- HPDinterval(as.mcmc(Subset.B.pgls.summary$Estimate), prob=0.95) %>% 
+  as.data.frame %>% 
+  dplyr::mutate(Parameter = 'as.factor(paternal_care)1')
+
+Subset.B.plot <-Subset.B.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6),fill="orange", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="orange")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none") + scale_x_continuous(limits = c(-0.25, 0.25)) +
+  annotate("text", x=-0.20, y=1.5, label= expression(lambda[BAMM])) +
+  geom_errorbarh(data = HPD_Interval_B_PC, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "orange", size = 1) +
+  xlab("Estimate (paternal care)") + ylab("")
+
+#Paternal_graph=ggarrange(Subset.ND.plot,Subset.DR.plot,Subset.B.plot, ncol=1)
+
+#### Graph for sexual dimorphism ####
+
+Subset.ND.pgls.summary=nd.wing
+Subset.ND.pgls.summary <- bind_rows(Subset.ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.ND.pgls.summary = filter(Subset.ND.pgls.summary,Parameter == 'abs(SSD_wing)')
+HPD_Interval_ND_W <- HPDinterval(as.mcmc(Subset.ND.pgls.summary$Estimate), prob=0.95) %>% as.data.frame %>% dplyr::mutate(Parameter = 'abs(SSD_wing)')
+
+Subset.ND.plot2 <-Subset.ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6), fill="darkturquoise", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1, colour="black", fill="darkturquoise")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")+ scale_x_continuous(limits = c(-2.5, 4)) +
+  xlab(" ") + ylab(" ") +
+    geom_errorbarh(data = HPD_Interval_ND_W, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "darkturquoise", size = 1)
+
+Subset.DR.pgls.summary=dr.wing
+Subset.DR.pgls.summary <- bind_rows(Subset.DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.DR.pgls.summary = filter(Subset.DR.pgls.summary,Parameter == 'abs(SSD_wing)')
+HPD_Interval_DR_W <- HPDinterval(as.mcmc(Subset.DR.pgls.summary$Estimate), prob=0.95) %>% as.data.frame %>% dplyr::mutate(Parameter = 'abs(SSD_wing)')
+
+Subset.DR.plot2 <-Subset.DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6),fill="darkturquoise", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="darkturquoise")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")+ scale_x_continuous(limits = c(-2.5, 4)) +
+  xlab(" ") + ylab(" ") +
+  geom_errorbarh(data = HPD_Interval_DR_W, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "darkturquoise", size = 1)
+
+Subset.B.pgls.summary=b.wing
+Subset.B.pgls.summary <- bind_rows(Subset.B.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.B.pgls.summary = filter(Subset.B.pgls.summary,Parameter == 'abs(SSD_wing)')
+HPD_Interval_B_W <- HPDinterval(as.mcmc(Subset.B.pgls.summary$Estimate), prob=0.95) %>% as.data.frame %>% dplyr::mutate(Parameter = 'abs(SSD_wing)')
+
+Subset.B.plot2 <-Subset.B.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6),fill="darkturquoise", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="darkturquoise")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none") + scale_x_continuous(limits = c(-2.5, 4)) +
+  xlab("Estimate (sexual size dimorphism)") + ylab("") +
+  geom_errorbarh(data = HPD_Interval_B_W, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "darkturquoise", size = 1)
+
+ggarrange(Subset.ND.plot,Subset.ND.plot2,Subset.DR.plot,Subset.DR.plot2,Subset.B.plot,Subset.B.plot2, ncol=2)
+#saved as 8x7
+
+```
+
+**Figure S18:** Estimates of the effect of individual sexual selection components included in the PPCA (paternal care and wing size dimorphism) on three measures of speciation rate ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$).
+Estimates are presented as density intervals from pgls models on 300 phylogentic trees.
+
 # R Session information

### 2020-08-22 Sexual Selection and Speciation.Rmd
new figure with mating system

--- 2020-05-30 Sexual Selection and Speciation.Rmd
@@
 library(bindrcpp)
+library(captioner)
 #devtools::install_github("Ax3man/phylopath")
@@
 knitr::opts_chunk$set(cache=TRUE)
+
+supp_figs <- captioner::captioner(prefix = "Figure S", auto_space = F)
+
+counter_fig <- function() {
+  if(!exists('counter_iteration')) {
+    counter_iteration <<- 1L
+  } else {
+    counter_iteration <<- counter_iteration + 1L
+  }
+  print(counter_iteration)
+}
 ```
@@
 
+<!-- ## Sexual selection components effects on speciation   -->
+
+```{r individual_components, fig.height = 8, fig.width = 7, warning = F, message = F, include=FALSE, echo = FALSE, eval = FALSE}
+#### Read files with summary results: Additional analyses were conducted seperate from this RMarkdown; Contact authors for information ----
+
+nd.wing <- readRDS('data/Subset.ND.pgls.summaryA.rds')[1:300]
+nd.mating <- readRDS('data/Subset.ND.pgls.summary_mating.rds')[1:300]
+nd.paternal <- readRDS('data/Subset.ND.pgls.summary_paternal.rds')
+dr.wing <- readRDS('data/Subset.DR.pgls.summary_sswing.rds')
+dr.mating <- readRDS('data/Subset.DR.pgls.summary_mating.rds')
+dr.paternal <- readRDS('data/Subset.DR.pgls.summary_paternal.rds')
+b.wing <- readRDS('data/Subset.Lambda.pgls.summaryB.rds')
+b.mating <- readRDS('data/Subset.Lambda.pgls.summary_mating.rds')
+b.paternal <- readRDS('data/Subset.Lambda.pgls.summary_paternal.rds')
+
+# Subset.ND.pgls.summary <- dplyr::bind_rows(b.wing, b.paternal) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+# 
+# Subset.ND.pgls.summary <- dplyr::filter(Subset.ND.pgls.summary, Parameter %in% c('abs(SSD_wing)', 'as.factor(paternal_care)1'))
+
+#HPDinterval(as.mcmc(Subset.ND.pgls.summary$Estimate), prob=0.95)
+
+#### Graph for paternal care ####
+
+Subset.ND.pgls.summary=nd.paternal
+Subset.ND.pgls.summary <- bind_rows(Subset.ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.ND.pgls.summary = filter(Subset.ND.pgls.summary,Parameter == 'as.factor(paternal_care)1')
+HPD_Interval_ND_PC <- HPDinterval(as.mcmc(Subset.ND.pgls.summary$Estimate), prob=0.95) %>% as.data.frame %>% dplyr::mutate(Parameter = 'as.factor(paternal_care)1')
+
+Subset.ND.plot <-Subset.ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6), fill="orange", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1, colour="black", fill="orange")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")+ scale_x_continuous(limits = c(-0.25, 0.25)) +
+  annotate("text", x=-0.22, y=1.5, label= expression(lambda[ND])) + xlab(" ") + ylab(" ") +
+  geom_errorbarh(data = HPD_Interval_ND_PC, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "orange", size = 1) +
+  theme(panel.grid.major.y = element_blank())
+  # geom_text(aes(y=Parameter, x = Estimate, label=""), position = position_nudge(y = -0.8), inherit.aes = F)
+
+Subset.DR.pgls.summary=dr.paternal
+Subset.DR.pgls.summary <- bind_rows(Subset.DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.DR.pgls.summary = filter(Subset.DR.pgls.summary,Parameter == 'as.factor(paternal_care)1')
+HPD_Interval_DR_PC <- HPDinterval(as.mcmc(Subset.DR.pgls.summary$Estimate), prob=0.95) %>% 
+  as.data.frame %>% 
+  dplyr::mutate(Parameter = 'as.factor(paternal_care)1')
+
+Subset.DR.plot <-Subset.DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6),fill="orange", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="orange")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")+ scale_x_continuous(limits = c(-0.25, 0.25))+
+  annotate("text", x=-0.22, y=1.5, label= expression(lambda[DR])) + xlab(" ") + ylab(" ") + 
+    geom_errorbarh(data = HPD_Interval_DR_PC, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "orange", size = 1) +
+  theme(panel.grid.major.y = element_blank())
+
+Subset.B.pgls.summary=b.paternal
+Subset.B.pgls.summary <- bind_rows(Subset.B.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.B.pgls.summary = filter(Subset.B.pgls.summary,Parameter == 'as.factor(paternal_care)1')
+HPD_Interval_B_PC <- HPDinterval(as.mcmc(Subset.B.pgls.summary$Estimate), prob=0.95) %>% 
+  as.data.frame %>% 
+  dplyr::mutate(Parameter = 'as.factor(paternal_care)1')
+
+Subset.B.plot <-Subset.B.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6),fill="orange", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="orange")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none") + scale_x_continuous(limits = c(-0.25, 0.25)) +
+  annotate("text", x=-0.20, y=1.5, label= expression(lambda[BAMM])) +
+  geom_errorbarh(data = HPD_Interval_B_PC, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "orange", size = 1) +
+  xlab("Estimate (paternal care)") + ylab("") +
+  theme(panel.grid.major.y = element_blank())
+
+#Paternal_graph=ggarrange(Subset.ND.plot,Subset.DR.plot,Subset.B.plot, ncol=1)
+
+#### Graph for sexual dimorphism ####
+
+Subset.ND.pgls.summary=nd.wing
+Subset.ND.pgls.summary <- bind_rows(Subset.ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.ND.pgls.summary = filter(Subset.ND.pgls.summary,Parameter == 'abs(SSD_wing)')
+HPD_Interval_ND_W <- HPDinterval(as.mcmc(Subset.ND.pgls.summary$Estimate), prob=0.95) %>% as.data.frame %>% dplyr::mutate(Parameter = 'abs(SSD_wing)')
+
+Subset.ND.plot2 <-Subset.ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6), fill="darkturquoise", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1, colour="black", fill="darkturquoise")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")+ scale_x_continuous(limits = c(-2.5, 4)) +
+  xlab(" ") + ylab(" ") +
+    geom_errorbarh(data = HPD_Interval_ND_W, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "darkturquoise", size = 1) +
+  theme(panel.grid.major.y = element_blank())
+
+Subset.DR.pgls.summary=dr.wing
+Subset.DR.pgls.summary <- bind_rows(Subset.DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.DR.pgls.summary = filter(Subset.DR.pgls.summary,Parameter == 'abs(SSD_wing)')
+HPD_Interval_DR_W <- HPDinterval(as.mcmc(Subset.DR.pgls.summary$Estimate), prob=0.95) %>% as.data.frame %>% dplyr::mutate(Parameter = 'abs(SSD_wing)')
+
+Subset.DR.plot2 <-Subset.DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6),fill="darkturquoise", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="darkturquoise")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none")+ scale_x_continuous(limits = c(-2.5, 4)) +
+  xlab(" ") + ylab(" ") +
+  geom_errorbarh(data = HPD_Interval_DR_W, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "darkturquoise", size = 1) +
+  theme(panel.grid.major.y = element_blank())
+
+Subset.B.pgls.summary=b.wing
+Subset.B.pgls.summary <- bind_rows(Subset.B.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))
+Subset.B.pgls.summary = filter(Subset.B.pgls.summary,Parameter == 'abs(SSD_wing)')
+HPD_Interval_B_W <- HPDinterval(as.mcmc(Subset.B.pgls.summary$Estimate), prob=0.95) %>% as.data.frame %>% dplyr::mutate(Parameter = 'abs(SSD_wing)')
+
+Subset.B.plot2 <-Subset.B.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
+  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               position = position_nudge(y= 0.6),fill="darkturquoise", colour="black")+
+  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="darkturquoise")+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none") + scale_x_continuous(limits = c(-2.5, 4)) +
+  xlab("Estimate (sexual size dimorphism)") + ylab("") +
+  geom_errorbarh(data = HPD_Interval_B_W, aes(y = Parameter, xmin = lower, xmax = upper), 
+                 inherit.aes = F, height = 0, position = position_nudge(y = -0.5), colour = "darkturquoise", size = 1) +
+  theme(panel.grid.major.y = element_blank())
+
+# Subrtact the intercept (monogamy) from the partial estimates to get their estimated effect on speciation rate in comparison to the reference level (monogamy)
+Subset.DR.pgls.summary <- dr.mating %>% 
+  dplyr::bind_rows() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>% #estimate now in 5th col
+   dplyr::filter(str_detect(string = Parameter, "mating_system|Intercept")) %>% 
+  dplyr::mutate(Parameter = case_when(Parameter == "as.factor(mating_system)1" ~ "<5% Polygyny",
+                                      Parameter == "as.factor(mating_system)2" ~ "5-20% Polygny", 
+                                      Parameter == "as.factor(mating_system)3" ~ ">20% Polygny", 
+                                      TRUE ~ Parameter))
+
+#Factor for plot display
+Subset.DR.pgls.summary$Parameter <- factor(Subset.DR.pgls.summary$Parameter, 
+                                           levels = c("(Intercept)", "<5% Polygyny", "5-20% Polygny", ">20% Polygny"))
+
+#Get HPD estimates
+HPD_DR_Mating <- Subset.DR.pgls.summary %>%
+  dplyr::group_by(Parameter) %>%
+  summarise(lower = HPDinterval(as.mcmc(Estimate), prob=0.95)[1],
+            upper = HPDinterval(as.mcmc(Estimate), prob=0.95)[2])
+
+Subset.DR.Mating.Plot <-Subset.DR.pgls.summary %>% 
+  filter(Parameter != "(Intercept)") %>% 
+  mutate(Test = "Mating System (vs Monogamy)") %>%
+  ggplot(aes(x = Estimate, y = Test, fill = Parameter, colour = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               alpha = 0.4,
+               position = position_nudge(y= 1), colour="black")+
+  #geom_jitter(shape = 21, alpha = 0.5, size = 1, fill="black", position = position_nudge(y= 0.4))+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none") + scale_x_continuous(limits = c(-0.3, 0.3)) +
+  xlab("") + ylab("") +
+  geom_errorbarh(data = HPD_DR_Mating %>% 
+  filter(Parameter != "(Intercept)") %>% 
+  mutate(Test = "Mating System (vs Monogamy)"), 
+  aes(y = Test, 
+      xmin = lower, 
+      xmax = upper, 
+      colour = Parameter), 
+  inherit.aes = F, height = 0, position = position_nudge(y = c(-0.1,-0.25, -0.4)), size = 1) +
+  geom_text(data = HPD_DR_Mating %>% 
+  filter(Parameter != "(Intercept)") %>% 
+  mutate(Test = "Mating System (vs Monogamy)"), 
+  aes(y = Test, 
+      x = lower, 
+      colour = Parameter, 
+      label = Parameter), 
+  inherit.aes = F, height = 0,
+  position = position_nudge(y = c(-0.1,-0.25, -0.4), x = -0.095), size = 3.25) +
+  scale_fill_manual(values = brewer.pal(8, "Greens")[6:8]) +
+  scale_color_manual(values = brewer.pal(8, "Greens")[6:8])
+
+#### ND ####
+
+# Subrtact the intercept (monogamy) from the partial estimates to get their estimated effect on speciation rate in comparison to the reference level (monogamy)
+Subset.ND.pgls.summary <- nd.mating %>% 
+  dplyr::bind_rows() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>% #estimate now in 5th col
+   dplyr::filter(str_detect(string = Parameter, "mating_system|Intercept")) %>% 
+  dplyr::mutate(Parameter = case_when(Parameter == "as.factor(mating_system)1" ~ "<5% Polygyny",
+                                      Parameter == "as.factor(mating_system)2" ~ "5-20% Polygny", 
+                                      Parameter == "as.factor(mating_system)3" ~ ">20% Polygny",
+                                      TRUE ~ Parameter))
+
+#Factor for plot display
+Subset.ND.pgls.summary$Parameter <- factor(Subset.ND.pgls.summary$Parameter, 
+                                           levels = c("(Intercept)", "<5% Polygyny", "5-20% Polygny", ">20% Polygny"))
+
+#Get HPD estimates
+HPD_ND_Mating <- Subset.ND.pgls.summary %>%
+  dplyr::group_by(Parameter) %>%
+  summarise(lower = HPDinterval(as.mcmc(Estimate), prob=0.95)[1],
+            upper = HPDinterval(as.mcmc(Estimate), prob=0.95)[2])
+
+Subset.ND.Mating.Plot <-Subset.ND.pgls.summary %>% 
+  filter(Parameter != "(Intercept)") %>% 
+  mutate(Test = "Mating System (vs Monogamy)") %>%
+  ggplot(aes(x = Estimate, y = Test, fill = Parameter, colour = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               alpha = 0.4,
+               position = position_nudge(y= 1), colour="black")+
+  #geom_jitter(shape = 21, alpha = 0.5, size = 1, fill="black", position = position_nudge(y= 0.4))+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none") + scale_x_continuous(limits = c(-0.3, 0.3)) +
+  xlab("") + ylab("") +
+  geom_errorbarh(data = HPD_ND_Mating %>% 
+  filter(Parameter != "(Intercept)") %>% 
+  mutate(Test = "Mating System (vs Monogamy)"), 
+  aes(y = Test, 
+      xmin = lower, 
+      xmax = upper, 
+      colour = Parameter), 
+  inherit.aes = F, height = 0, position = position_nudge(y = c(-0.1,-0.25, -0.4)), size = 1) +
+  geom_text(data = HPD_ND_Mating %>% 
+  filter(Parameter != "(Intercept)") %>% 
+  mutate(Test = "Mating System (vs Monogamy)"), 
+  aes(y = Test, 
+      x = lower, 
+      colour = Parameter, 
+      label = Parameter), 
+  inherit.aes = F, height = 0,
+  position = position_nudge(y = c(-0.1,-0.25, -0.4), x = -0.095), size = 3.25) +
+  scale_fill_manual(values = brewer.pal(8, "Greens")[6:8]) +
+  scale_color_manual(values = brewer.pal(8, "Greens")[6:8])
+
+# Subrtact the intercept (monogamy) from the partial estimates to get their estimated effect on speciation rate in comparison to the reference level (monogamy)
+Subset.B.pgls.summary <- b.mating %>% 
+  dplyr::bind_rows() %>% 
+  `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI")) %>% #estimate now in 5th col
+   dplyr::filter(str_detect(string = Parameter, "mating_system|(Intercept)")) %>% 
+  dplyr::mutate(Parameter = case_when(Parameter == "as.factor(mating_system)1" ~ "<5% Polygyny",
+                                      Parameter == "as.factor(mating_system)2" ~ "5-20% Polygny", 
+                                      Parameter == "as.factor(mating_system)3" ~ ">20% Polygny", 
+                                      TRUE ~ Parameter))
+
+#Factor for plot display
+Subset.B.pgls.summary$Parameter <- factor(Subset.B.pgls.summary$Parameter, 
+                                           levels = c("(Intercept)", "<5% Polygyny", "5-20% Polygny", ">20% Polygny"))
+
+#Get HPD estimates
+HPD_B_Mating <- Subset.B.pgls.summary %>%
+  dplyr::group_by(Parameter) %>%
+  summarise(lower = HPDinterval(as.mcmc(Estimate), prob=0.95)[1],
+            upper = HPDinterval(as.mcmc(Estimate), prob=0.95)[2])
+
+Subset.b.mating.Plot <-Subset.B.pgls.summary %>% 
+  filter(Parameter != "(Intercept)") %>% 
+  mutate(Test = "Mating System (vs Monogamy)") %>%
+  ggplot(aes(x = Estimate, y = Test, fill = Parameter, colour = Parameter)) +
+  geom_density(aes(y = ..scaled..),
+               alpha = 0.4,
+               position = position_nudge(y= 1), colour="black")+
+  #geom_jitter(shape = 21, alpha = 0.5, size = 1, fill="black", position = position_nudge(y= 0.4))+
+  geom_vline(xintercept = 0)+
+  theme_minimal()+
+  theme(axis.text.y = element_blank(),
+        legend.position = "none") + scale_x_continuous(limits = c(-0.3, 0.3)) +
+  xlab("Estimate (Mating System vs Monogamy)") + ylab("") +
+  geom_errorbarh(data = HPD_B_Mating %>% 
+  filter(Parameter != "(Intercept)") %>% 
+  mutate(Test = "Mating System (vs Monogamy)"), 
+  aes(y = Test, 
+      xmin = lower, 
+      xmax = upper, 
+      colour = Parameter), 
+  inherit.aes = F, height = 0, position = position_nudge(y = c(-0.1,-0.25, -0.4)), size = 1) +
+  geom_text(data = HPD_B_Mating %>% 
+  filter(Parameter != "(Intercept)") %>% 
+  mutate(Test = "Mating System (vs Monogamy)"), 
+  aes(y = Test, 
+      x = lower, 
+      colour = Parameter, 
+      label = Parameter), 
+  inherit.aes = F, height = 0,
+  position = position_nudge(y = c(-0.1,-0.25, -0.4), x = -0.095), size = 3.25) +
+  scale_fill_manual(values = brewer.pal(8, "Greens")[6:8]) +
+  scale_color_manual(values = brewer.pal(8, "Greens")[6:8])
+
+ggarrange(Subset.ND.plot,Subset.ND.plot2,Subset.ND.Mating.Plot,
+          Subset.DR.plot,Subset.DR.plot2,Subset.DR.Mating.Plot,
+          Subset.B.plot,Subset.B.plot2,Subset.b.mating.Plot, ncol=3)
+```
+
+<!-- **Figure S18:** Estimates of the effect of individual sexual selection components included in the PPCA (paternal care and wing size dimorphism) on three measures of speciation rate ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$).
+Estimates are presented as density intervals from pgls models on 300 phylogentic trees.
+-->
+
 # R Session information

### 2020-08-23 Manuscript_Publication.Rmd
slight changes to manuscript and figure

@@
 The link we observed between sexual selection and speciation was independent of environmental variability, though species with smaller ranges had higher speciation rates.
-There was no association between sexual selection and extinction rate.
+There was no association between any proxies of sexual selection and extinction rate.
 **Our findings support the view that male-biased sexual selection, as measured by predictors of male-male competition, has shaped diversification in the largest radiation of birds.**
@@
 More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012].
-Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.More recently, social polygyny (a proxy for sexual selection) was found to have a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].
+Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.
+More recently, social polygyny (a proxy for sexual selection) was found to have a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].
 
@@
 Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018].
-Additionally, our analysis includes multiple ecological and environmental variables, allowing us to adjust for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016].
+Additionally, our analysis includes multiple ecological and environmental variables, allowing us to control for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016].
+
 We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including Bayesian Analysis of Macroevolutionary Mixtures [BAMM; @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, namely diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$) [@Jetz_2012].
-Our results show that (*i*) a composite measure of sexual selection, but not sexual dichromatism, significantly predicts speciation rates, (*ii*) species with smaller ranges have higher speciation rates and (*iii*) there is no evidence that environmental variables or their interaction with sexual selection have an impact on diversification rates.
+Our results show that (*i*) a composite measure of sexual selection, but not sexual dichromatism, significantly predicts speciation rates, (*ii*) **the significant association between the composite measure of sexual selection and speciation rate is largely driven by sexual size dimorphism**, (*iii*) species with smaller ranges have higher speciation rates and (*iv*) there is no evidence that environmental variables or their interaction with sexual selection have an impact on diversification rates.
 Therefore, we provide evidence at a very large scale that sexual selection can have positive effects on diversification in the largest radiation of birds.
@@
 
-Secondly, we used BAMM to model the dynamics of speciation and extinction across the **101 phylogenetic trees (one MCC tree and 100 random draws of the posterior)**.
+We used BAMM to model the dynamics of speciation and extinction across the **101 phylogenetic trees (one MCC tree and 100 random draws of the posterior)**.
 This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate probability distributions of evolutionary rate-shift configurations with variable speciation and extinction rates [@Rabosky_2014].
@@
 **For each response variable, we conducted model selection to compare models with different combinations of predictor variables.
-** The most complicated model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (i.e.
+** The most complex model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (i.e.
 log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP), and all of the 2-way interactions between sexual selection and each of the environmental measures.
@@
 Using the terms from the top-ranked model (ranked by AICc), we ran the equivalent model for each of the 1,000 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and each of the 100 trees used to derive $\lambda_{BAMM}$ and $\mu_{BAMM}$.
-**Additionally, we investigated the effect of the individual principal component scores for the index of male-biased sexual selection on speciation rate.
+**Additionally, we investigated the effect of the individual variables used to derive the index of male-biased sexual selection on speciation rate.
 For these pgls models we replaced the composite index score with the individual biological variable (sexual size dimorphism, social polygyny and [lack of] paternal care) and ran the equivalent model for 300 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and 100 trees used to derive $\lambda_{BAMM}$**.
@@
 Finally, using the subset of species with an index of male-biased sexual selection, we conducted a phylogenetic path analysis using the ``phylopath R`` package [@Wouter_2018].
-The phylogenetic path analysis was used to assess causal paths between variables not able to be modelled within the univariate response of PGLS.
+The phylogenetic path analysis was used to assess causal paths between variables unable to be modelled within the univariate response of PGLS.
 That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection, and range size to have effects on each other and not just on speciation rate.
@@
 
-**Given that this male-biased sexual selection index comprises three different (but correlated) variables, we investigated which of these variables was driving the association observed with $\lambda_{DR}$.
+**Given that this male-biased sexual selection index comprises of three different (but correlated) variables, we investigated which of these variables was driving the association observed with $\lambda_{DR}$.
 Our results over 300 trees showed that this pattern is mainly driven by the sexual size dimorphism component (HPD Interval = `r sanitize.numbers(formatC(dr.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), with the effects of other components overlapping zero; paternal care (HPD Interval = `r sanitize.numbers(formatC(dr.paternal, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`) and mating system (HPD Interval= `r sanitize.numbers(formatC(dr.mating, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`).
@@
 Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
-Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}).
+Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}, @Harvey_2017).
 Full details of the BAMM results are presented as supplementary materials.
@@
 
-![Estimates of the effect of individual sexual selection components included in the PPCA (paternal care and wing size dimorphism) on three measures of speciation rate ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$).
+![Estimates of the effect of individual sexual selection components included in the PPCA (paternal care, sexual size dimorphism and mating system) on three measures of speciation rate ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$).
 Estimates are presented as density intervals from pgls models on 300 phylogentic trees that used species with available data for these sexual selection measures (*n* = 2,465).
 The bar under each density ridge is the 95 % Highest Posterior Density Interval.
+Given that the mating system variable is categorical, model estimates for three polygynous mating system levels are in reference to a striclty monogamous mating system (0% polygyny).
 \label{SS_components}](Figures/Individual_SS_Estimates.pdf)
@@
 
-**Sexual selection due to competition within sexes contrasts with the drivers of sexual dichromatism.
+**Sexual dimorphism due to competition within sexes contrasts with the drivers of sexual dichromatism.
 Plumage dichromatism can evolve as a consequence of female cryptic choice and be related to extra-bond fertilizations, but not necessarily paternal care or mating system [@owens_1998].
@@
 Differences in body size (as a result of sexual selection) could be linked to changes in diet, vulnerability to predators or environmental tolerance [@Damuth_1993; @Liow_2008; @Bonduriansky_2011], and such differences could ultimately increase the likelihood of divergence between young lineages.
-In mammals, sexual selection has driven the evolution of large body size which in turn has allowed diversification of ecological strategies in the clade, and higher speciation rates.**
+In mammals, sexual selection is suggested to have driven the evolution of large body size which in turn has allowed diversification of ecological strategies in the clade, and higher speciation rates [@Mclain_1993; @Bonduriansky_2011].**
 
@@
 In addition to speciation, sexual selection is hypothesised to affect extinction.
-Using the model-based approach of BAMM, we found no association between the estiamted extinction rate and sexual dichromatism, male-biased sexual selection, or our measures of environmental variability.
+Using the model-based approach of BAMM, we found no association between the estimated extinction rate and sexual dichromatism, male-biased sexual selection, or our measures of environmental variability.
 However, these extinction results should not be regarded as definitive because extinction is notoriously difficult to estimate accurately from phylogenies, principally because different combinations of speciation and extinction rates can give rise to similar patterns of diversity [see @Rabosky_Extinction_2016].

### 2020-08-23 Sexual Selection and Speciation.Rmd
slight changes to manuscript and figure

@@
 
-**Table S7:** The estimates of the MCC models plotted above are based on the following data tables.
+**Table S8:** The estimates of the MCC models plotted above are based on the following data tables.
 ```{r Values For Models}
@@
 
-**Table S8:**  HPD intervals calculated for the above figure (*i.e.* models using RGB values of sexual dichromatism).
+**Table S9:**  HPD intervals calculated for the above figure (*i.e.* models using RGB values of sexual dichromatism).
 The HPD range is determined using the ``hdi`` function with a 95 % credible interval.
@@
 
-**Table S9:** Model estimates from $\lambda_{BAMM}$ and $\mu_{BAMM}$ tip-rate estimates show high levels of variability between trees relative to estimates from $\lambda_{DR}$ and $\lambda_{ND}$.
+**Table S10:** Model estimates from $\lambda_{BAMM}$ and $\mu_{BAMM}$ tip-rate estimates show high levels of variability between trees relative to estimates from $\lambda_{DR}$ and $\lambda_{ND}$.
 The interval ratio is calculated as the width of the HPD 95 % CI relative to the MCC 95 % CI.
@@
 
-**Table S10:** The dredged models using a subset of species for which sexual dichromatism was measured using spectrophotometry.
+**Table S11:** The dredged models using a subset of species for which sexual dichromatism was measured using spectrophotometry.
 All four show the top model is one with no interactions, with $\delta AICc > 4$.
@@
 
-**Table S11:** MCC estimates from the above plot are presented as numerical values below. 
+**Table S12:** MCC estimates from the above plot are presented as numerical values below. 
 ```{r Armenta moel summary}
@@
 
-**Table S12:**  HPD intervals calculated for the above figure (*i.e.* models using Spectrophotometry values of sexual dichromatism).
+**Table S13:**  HPD intervals calculated for the above figure (*i.e.* models using Spectrophotometry values of sexual dichromatism).
 These intervals do not take into account the variance associated with each estimate and thus are not an estimate of model precision.
 Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that are in the same direction (+ or -).
-These intervals are calculated in the same way as in Table S7.
+These intervals are calculated in the same way as in Table S9.
 ```{r Armenta HPD, warning = FALSE}
@@
 
-**Table S13:** All top models ($\delta AICc > 4$) using male-biased sexual selection measures do not contain interactions between sexual selection and the environmental variables.
+**Table S14:** All top models ($\delta AICc > 4$) using male-biased sexual selection measures do not contain interactions between sexual selection and the environmental variables.
 Thus interaction terms were not included in further analysis.
@@
 
-**Table S14:** MCC model estimates from the above model are presented here as numberical values with 95 % CIs.
+**Table S15:** MCC model estimates from the above model are presented here as numberical values with 95 % CIs.
 ```{r Table for SS plots}
@@
 
-**Table S15:** HPD intervals are calculated for the above figure (i.e.
+**Table S16:** HPD intervals are calculated for the above figure (i.e.
 models using Male-bias sexual selection measures).
@@
 Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter are in the same direction (+ or -).
-These intervals are calculated in the same way as in Table S7 and Table S9.
+These intervals are calculated in the same way as in Table S9.
 ```{r SS HPD, warning = FALSE}
@@
                position = position_nudge(y= 0.6), fill="orange", colour="black")+
-  geom_jitter(shape = 21, alpha = 0.5, size = 1, colour="black", fill="orange")+
+  geom_jitter(shape = 21, alpha = 0, size = 1, colour="black", fill="orange")+
   geom_vline(xintercept = 0)+
@@
                position = position_nudge(y= 0.6),fill="orange", colour="black")+
-  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="orange")+
+  geom_jitter(shape = 21, alpha = 0, size = 1,fill="black", colour="orange")+
   geom_vline(xintercept = 0)+
@@
                position = position_nudge(y= 0.6), fill="darkturquoise", colour="black")+
-  geom_jitter(shape = 21, alpha = 0.5, size = 1, colour="black", fill="darkturquoise")+
+  geom_jitter(shape = 21, alpha = 0, size = 1, colour="black", fill="darkturquoise")+
   geom_vline(xintercept = 0)+
@@
                position = position_nudge(y= 0.6),fill="darkturquoise", colour="black")+
-  geom_jitter(shape = 21, alpha = 0.5, size = 1,fill="black", colour="darkturquoise")+
+  geom_jitter(shape = 21, alpha = 0, size = 1,fill="black", colour="darkturquoise")+
   geom_vline(xintercept = 0)+
@@
    dplyr::filter(str_detect(string = Parameter, "mating_system|Intercept")) %>% 
-  dplyr::mutate(Parameter = case_when(Parameter == "as.factor(mating_system)1" ~ "<5% Polygyny",
-                                      Parameter == "as.factor(mating_system)2" ~ "5-20% Polygny", 
-                                      Parameter == "as.factor(mating_system)3" ~ ">20% Polygny", 
+  dplyr::mutate(Parameter = case_when(Parameter == "as.factor(mating_system)1" ~ "<5% polygyny",
+                                      Parameter == "as.factor(mating_system)2" ~ "5-20% polygyny", 
+                                      Parameter == "as.factor(mating_system)3" ~ ">20% polygyny", 
                                       TRUE ~ Parameter))
@@
 Subset.DR.pgls.summary$Parameter <- factor(Subset.DR.pgls.summary$Parameter, 
-                                           levels = c("(Intercept)", "<5% Polygyny", "5-20% Polygny", ">20% Polygny"))
+                                           levels = c("(Intercept)", "<5% polygyny", "5-20% polygyny", ">20% polygyny"))
 
@@
    dplyr::filter(str_detect(string = Parameter, "mating_system|Intercept")) %>% 
-  dplyr::mutate(Parameter = case_when(Parameter == "as.factor(mating_system)1" ~ "<5% Polygyny",
-                                      Parameter == "as.factor(mating_system)2" ~ "5-20% Polygny", 
-                                      Parameter == "as.factor(mating_system)3" ~ ">20% Polygny",
+  dplyr::mutate(Parameter = case_when(Parameter == "as.factor(mating_system)1" ~ "<5% polygyny",
+                                      Parameter == "as.factor(mating_system)2" ~ "5-20% polygyny", 
+                                      Parameter == "as.factor(mating_system)3" ~ ">20% polygyny",
                                       TRUE ~ Parameter))
@@
 Subset.ND.pgls.summary$Parameter <- factor(Subset.ND.pgls.summary$Parameter, 
-                                           levels = c("(Intercept)", "<5% Polygyny", "5-20% Polygny", ">20% Polygny"))
+                                           levels = c("(Intercept)", "<5% polygyny", "5-20% polygyny", ">20% polygyny"))
 
@@
    dplyr::filter(str_detect(string = Parameter, "mating_system|(Intercept)")) %>% 
-  dplyr::mutate(Parameter = case_when(Parameter == "as.factor(mating_system)1" ~ "<5% Polygyny",
-                                      Parameter == "as.factor(mating_system)2" ~ "5-20% Polygny", 
-                                      Parameter == "as.factor(mating_system)3" ~ ">20% Polygny", 
+  dplyr::mutate(Parameter = case_when(Parameter == "as.factor(mating_system)1" ~ "<5% polygyny",
+                                      Parameter == "as.factor(mating_system)2" ~ "5-20% polygyny", 
+                                      Parameter == "as.factor(mating_system)3" ~ ">20% polygyny", 
                                       TRUE ~ Parameter))
@@
 Subset.B.pgls.summary$Parameter <- factor(Subset.B.pgls.summary$Parameter, 
-                                           levels = c("(Intercept)", "<5% Polygyny", "5-20% Polygny", ">20% Polygny"))
+                                           levels = c("(Intercept)", "<5% polygyny", "5-20% polygyny", ">20% polygyny"))
 
@@
         legend.position = "none") + scale_x_continuous(limits = c(-0.3, 0.3)) +
-  xlab("Estimate (Mating System vs Monogamy)") + ylab("") +
+  xlab("Estimate (mating system vs monogamy)") + ylab("") +
   geom_errorbarh(data = HPD_B_Mating %>% 

### 2020-08-29 Manuscript_Publication.Rmd
updated supp info

@@
   \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
-  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{draft, colorlinks=false, hyperfootnotes=false,
+  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
   linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
@@
 Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions.
-Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}.
+Analyses are documented with reproducible code in the \href{https://ssspeciation.github.io/SexualSelection_Speciation/}{Supplementary Information}.
 \phantomsection
@@
 These measurements include variation in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from museum specimens as opposed to illustrations.
-The spectrophotometry data covers only 581 passerine species (10-fold fewer than the RGB data), although there was a substantial correlation between the two dichromatism measures ($r = 0.79$; \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
+The spectrophotometry data covers only 581 passerine species (10-fold fewer than the RGB data), although there was a substantial correlation between the two dichromatism measures ($r = 0.79$; \href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
 \phantomsection
@@
 **The variables included in this index have all been positively linked to the intensity of sexual selection, and are usually correlated [@bjorklund_1990; @owens_1998; @dunn_2001], which is why they were combined into a single metric in previous studies [@Dale_2015]**.
-This measure of male-biased sexual selection is available for only 2,465 species, and shows a moderate correlation with the RGB measure of sexual dichromatism ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
+This measure of male-biased sexual selection is available for only 2,465 species, and shows a moderate correlation with the RGB measure of sexual dichromatism ($r = 0.34$; \href{https://ssspeciation.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
 \phantomsection
@@
 Estimates of NPP had 30-second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005].
-Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
+Using these data, which we provide as a potentially useful data resource (see \href{https://ssspeciation.github.io/SexualSelection_Speciation/}{Supplementary Information}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
 \phantomsection
@@
 (*iv*) To estimate levels of spatial environmental variation a species may endure, we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO4 and BIO15).
-PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 **largely** reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
-**A taxon’s range size often correlates with speciation and extinction rates [@Rosenzweig_1995; @Castiglione_2017]**, so we controlled for the correlation between environmental spatial variation and species' range sizes --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
+PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 **largely** reflects the variation in temperature across a species' range (\href{https://ssspeciation.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
+**A taxon’s range size often correlates with speciation and extinction rates [@Rosenzweig_1995; @Castiglione_2017]**, so we controlled for the correlation between environmental spatial variation and species' range sizes --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://ssspeciation.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
 To obtain long-term variation in climates for each species range we take (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
-Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
-The five predictors of environmental variability are not strongly correlated (\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
-Details and ``R`` code to generate these predictors can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}.
+Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://ssspeciation.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
+The five predictors of environmental variability are not strongly correlated (\href{https://ssspeciation.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
+Details and ``R`` code to generate these predictors can be found within the \href{https://ssspeciation.github.io/SexualSelection_Speciation/}{Supplementary Information}.
 \phantomsection
@@
 **These models provide tip-rate estimates of speciation and extinction rate that can be easily used in comparative analyses**.
-The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{Supplementary Information}; **briefly, we used a time-variable model with the prior expected number of evolutionary rate shifts set at 100 and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
+The parameters of the 100 BAMM runs are detailed in full in the \href{https://ssspeciation.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{Supplementary Information}; **briefly, we used a time-variable model with the prior expected number of evolutionary rate shifts set at 100 and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
 BAMM models were run independently for the 101 phylogenetic trees for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
-Important BAMM parameters (log-likelihood and number of rate shifts) reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200; an arbitrary value, above which posterior distributions can often be accurately inferred** (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
-Further details of BAMM parameters and output are available in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}, with tip-rate means and variances provided.
+Important BAMM parameters (log-likelihood and number of rate shifts) reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200; an arbitrary value, above which posterior distributions can often be accurately inferred** (\href{https://ssspeciation.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
+Further details of BAMM parameters and output are available in the \href{https://ssspeciation.github.io/SexualSelection_Speciation/}{Supplementary Information}, with tip-rate means and variances provided.
 Additionally, given the variability in BAMM estimates, we also provide analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on the MCC tree from a genetic-only phylogeny across all birds [@Harvey_2017].
@@
 To minimise path complexity we used temperature seasonality (BIO4) as the single measure for environmental variability, $\lambda_{DR}$ as the single measure of speciation and the tip-rates from the MCC tree.
-Further details of the path analysis, including our rationale for each path's directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information} along with all other analyses and the relevant ``R`` code to reproduce results.
+Further details of the path analysis, including our rationale for each path's directions, can be found within the \href{https://ssspeciation.github.io/SexualSelection_Speciation/}{Supplementary Information} along with all other analyses and the relevant ``R`` code to reproduce results.
 \phantomsection
@@
 However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; $\lambda_{BAMM}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b).
-When we took into account phylogenetic uncertainty by running the models using 1,000 trees, the distribution of estimates from PGLS models was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r sanitize.numbers(formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+When we took into account phylogenetic uncertainty by running the models using 1,000 trees, the distribution of estimates from PGLS models was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r sanitize.numbers(formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \href{https://ssspeciation.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
 
@@
 PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; i.e.
-no association between sexual dichromatism and speciation  (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
+no association between sexual dichromatism and speciation  (\href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
 Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 trees for $\lambda_{BAMM}$.
-The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism ($\lambda_{DR}$: HPD Interval = `r sanitize.numbers(formatC(hpd.DR.top$SDi, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, $\lambda_{ND}$: HPD Interval = `r sanitize.numbers(formatC(hpd.ND.top$SDi, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, \autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
-For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r sanitize.numbers(formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism ($\lambda_{DR}$: HPD Interval = `r sanitize.numbers(formatC(hpd.DR.top$SDi, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, $\lambda_{ND}$: HPD Interval = `r sanitize.numbers(formatC(hpd.ND.top$SDi, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, \autoref{model_results}a, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
+For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r sanitize.numbers(formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
 
@@
 association with speciation rates only for the latter) were not due to differences in the size of the datasets used (5,812 species vs.
-2,465,** \href{https://justincally.github.io/SexualSelection_Speciation/#testing_the_sexual_dichromatism_dataset_using_the_same_data_points_as_the_sexual_selection_dataset}{Figure S17}).
-No interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and the index of male-biased sexual selection; $\Delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
+2,465,** \href{https://ssspeciation.github.io/SexualSelection_Speciation/#testing_the_sexual_dichromatism_dataset_using_the_same_data_points_as_the_sexual_selection_dataset}{Figure S17}).
+No interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and the index of male-biased sexual selection; $\Delta$ AICc > 4; \href{https://ssspeciation.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}).
 Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation or range size.
-Furthermore, we found no evidence that these environmental factors --- seasonal temperature variation, long-term temperature variation, spatial temperature variation, and Net Primary Productivity (NPP) --- predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
+Furthermore, we found no evidence that these environmental factors --- seasonal temperature variation, long-term temperature variation, spatial temperature variation, and Net Primary Productivity (NPP) --- predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
 \phantomsection
@@
 Range size is **significantly associated with** $\lambda_{DR}$ (\autoref{model_results}b) using data subset for species with an index of male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
-Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
+Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S13}).
 Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with the highest power using near-complete taxon sampling (*n* = 5,812).
@@
 
-Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; \href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S14}).
+Using a phylogenetic path analysis, we found multiple significant paths between variables used in the PGLS (\autoref{Path_Analysis}; \href{https://ssspeciation.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S14}).
 There was a modest effect of male-biased sexual selection on sexual dimorphism ($\beta$ = `r round(best_model[["coef"]][2,4], digits = 2)`).
@@
 
-We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r sanitize.numbers(formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 2)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r sanitize.numbers(formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 2)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S13}) or male-biased sexual selection (\autoref{model_results}b, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S16}).
 \phantomsection
@@
 Estimates of the effect of predictor variables on speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % HPD interval across PGLS model estimates from 100 trees was often > 20 times larger than the 95 % confidence interval for estimates from a single PGLS model using the MCC tree.
-This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
+This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://ssspeciation.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S9}).
 Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, **and so it is concerning that there are such large differences in the BAMM estimates between alternative, similarly plausible phylogenies**.
-Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
-Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}, @Harvey_2017).
+Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://ssspeciation.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
+Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}, @Harvey_2017).
 Full details of the BAMM results are presented as supplementary materials.

### 2020-08-29 Sexual Selection and Speciation.Rmd
updated supp info

@@
 source('functions/gheatmap2.R') #Gets rid of spaces between tiles
-knitr::opts_chunk$set(cache=TRUE)
-
-supp_figs <- captioner::captioner(prefix = "Figure S", auto_space = F)
-
-counter_fig <- function() {
-  if(!exists('counter_iteration')) {
-    counter_iteration <<- 1L
-  } else {
-    counter_iteration <<- counter_iteration + 1L
-  }
-  print(counter_iteration)
-}
+knitr::opts_chunk$set(cache=TRUE, warning =F)
+
+# supp_figs <- captioner::captioner(prefix = "Figure S", auto_space = F)
+# 
+# 
+# counter_fig <- function() {
+#   if(!exists('counter_iteration')) {
+#     counter_iteration <<- 1L
+#   } else {
+#     counter_iteration <<- counter_iteration + 1L
+#   }
+#   print(counter_iteration)
+# }
 ```

### 2020-09-05 Manuscript_Publication.Rmd
slight changes

@@
 For example, in Darwin’s finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-**Under these circumstances we predict that the effect of sexual selection on phylogenetic rates of divergence may depend on the variability of the species’ environment**.
+**Under these circumstances we predict that the effect of sexual selection on rates of divergence may depend on the variability of the species’ environment**.
 Despite the potential interaction between sexual selection and environmental variability in diversification, phylogenetic tests are currently lacking.
@@
 Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://ssspeciation.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
-Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}, @Harvey_2017).
+Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://ssspeciation.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8}, @Harvey_2017).
 Full details of the BAMM results are presented as supplementary materials.
@@
 Clockwise the six species are: _Sporophila bouvronides_, _Euplectes franciscanus_, _Phainopepla nitens_, _Paradisaea rubra_, _Malurus pulcherrimus_, _Lepidothrix coeruleocapilla_.
-Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation and should not be interpreted.
+Edge colours for the terminal branch correspond to $\lambda_{DR}$ but all precluding branches have been generated for graphical purposes using ancestral character state estimation [@phytools_2012] and should not be interpreted.
 Illustrations reproduced by permission of Lynx Editions.

### 2020-12-03 Manuscript_Publication.Rmd
third round of revisions

--- 2020-08-23 Manuscript_Publication.Rmd
@@
   \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
-  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{draft, colorlinks=false, hyperfootnotes=false,
-  linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
+  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
+  linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
   \raggedbottom
 csl: References/evolution.csl
+nocite: | 
+  @Kraaijeveld_2011, @Maia_2013, @Huang_2014, @Gomes_2016, @Cooney_2017, @Janicke_2018, @Mason_2017, @Iglesias-Carrasco_2019, @Hosner_2020, @Price-Waldman_2020
 subparagraph: yes
@@
 However, the net effect of sexual selection on diversification is hard to predict _a priori_, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes.
-**Theoretical work also suggests that the net effect of sexual selection on diversification should depend strongly on ecological factors, though this prediction has seldom been tested.
+Theoretical work also suggests that the net effect of sexual selection on diversification should depend strongly on ecological factors, though this prediction has seldom been tested.
 Here, we test whether variation in sexual selection can predict speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera) and whether this relationship is mediated by environmental factors.
-Male-biased sexual selection, and specifically sexual size dimorphism, predicted two of the three measures of speciation rates that we examined**.
+Male-biased sexual selection, and specifically sexual size dimorphism, predicted two of the three measures of speciation rates that we examined.
 The link we observed between sexual selection and speciation was independent of environmental variability, though species with smaller ranges had higher speciation rates.
 There was no association between any proxies of sexual selection and extinction rate.
-**Our findings support the view that male-biased sexual selection, as measured by predictors of male-male competition, has shaped diversification in the largest radiation of birds.**
+Our findings support the view that male-biased sexual selection, as measured by frequent predictors of male-male competition, has shaped diversification in the largest radiation of birds.
 ---
@@
 library(dplyr)
+library(readxl)
+library(pander)
+
+#add literature table  
+literature_table <- readxl::read_excel("literature_table.xlsx") %>% filter(!is.na(Outcome))
 
@@
 
-Sexual selection is a ubiquitous evolutionary process whose effect on phylogenic diversification is much debated **[@Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018; @tsuji2020sexual]**.
+Sexual selection is a ubiquitous evolutionary process whose effect on phylogenetic diversification is much debated [@Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018; @tsuji2020sexual].
 Sexual selection can promote speciation because it operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia].
@@
 Although numerous studies have examined the relationship between sexual selection and speciation or extinction rates [@Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Kraaijeveld_2011; @Seddon_2013; @Huang_2014], the availability of more complete phenotypic, ecological and phylogenetic data [@Jetz_2012], together with significant advances in phylogenetic methods [@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.
-**Furthermore, the diversity of outcomes and approaches in previous studies suggests that the association between species diversity and sexual selection is far from clear (reviewed in @tsuji2020sexual).**
+Furthermore, the diversity of outcomes and approaches in previous studies suggests that the association between species diversity and sexual selection is far from clear (reviewed in @tsuji2020sexual).
+**See Table 1 for a review of the major findings from previous studies testing the association between sexual selection and speciation**.
 
-**A possible reason for the above uncertainty regarding the relationship between sexual selection and diversification is that this relationship may strongly depend on the environment.** Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [@Long_2012; @Connallon_Hall_2016].
+A possible reason for the above uncertainty regarding the relationship between sexual selection and diversification is that this relationship may strongly depend on the environment.
+Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [@Long_2012; @Connallon_Hall_2016].
 In stable environments, consistent selection depletes genetic variation at sexually concordant loci (i.e.
@@
 By contrast, in spatially or temporally variable environments, sexual selection can enhance local adaptation.
-For example, in Darwin’s finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
-**Under these circumstances we predict that the effect of sexual selection on phylogenetic rates of divergence may depend on the variability of the species’ environment**.
+For example, in Darwin’s finches, divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007].
+Under these circumstances we predict that the effect of sexual selection on rates of divergence may depend on the variability of the species’ environment.
 Despite the potential interaction between sexual selection and environmental variability in diversification, phylogenetic tests are currently lacking.
@@
 We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including Bayesian Analysis of Macroevolutionary Mixtures [BAMM; @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, namely diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$) [@Jetz_2012].
-Our results show that (*i*) a composite measure of sexual selection, but not sexual dichromatism, significantly predicts speciation rates, (*ii*) **the significant association between the composite measure of sexual selection and speciation rate is largely driven by sexual size dimorphism**, (*iii*) species with smaller ranges have higher speciation rates and (*iv*) there is no evidence that environmental variables or their interaction with sexual selection have an impact on diversification rates.
+Our results show that (*i*) a composite measure of sexual selection, but not sexual dichromatism, significantly predicts speciation rates, (*ii*) the significant association between the composite measure of sexual selection and speciation rate is largely driven by sexual size dimorphism, (*iii*) species with smaller ranges have higher speciation rates and (*iv*) there is no evidence that environmental variables or their interaction with sexual selection have an impact on diversification rates.
 Therefore, we provide evidence at a very large scale that sexual selection can have positive effects on diversification in the largest radiation of birds.
 Furthermore, we suggest that the use of sexual dichromatism as the sole proxy for sexual selection should be reconsidered, since it appears to be inconsistently associated with the operation of sexual selection.
+
+```{r lit, echo = FALSE, warning=FALSE}
+options(knitr.kable.NA = ' ')
+literature_table %>%
+  select(-Citation) %>% 
+  # pandoc.table(style = "rmarkdown", 
+  #              caption = "Review of studies testing the association between sexual selection and speciation")
+  knitr::kable(format = "latex",
+        booktabs = T,
+        escape = T,
+        caption = "Review of studies testing the association between sexual selection and speciation") %>%
+  kableExtra::kable_styling(latex_options = c("HOLD_position", "scale_down"),
+                            full_width = FALSE, 
+                            font_size = 8) %>%
+  kableExtra::column_spec(5, "27em") %>%
+  kableExtra::column_spec(3, "10em") %>%
+  kableExtra::column_spec(4, "3em") %>%
+  kableExtra::collapse_rows(1:2, row_group_label_position = "identity") 
+```
 
@@
 This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
-**The variables included in this index have all been positively linked to the intensity of sexual selection, and are usually correlated [@bjorklund_1990; @owens_1998; @dunn_2001], which is why they were combined into a single metric in previous studies [@Dale_2015]**.
+The variables included in this index have all been positively linked to the intensity of sexual selection, and are usually correlated [@bjorklund_1990; @owens_1998; @dunn_2001], which is why they were combined into a single metric in previous studies [@Dale_2015].
 This measure of male-biased sexual selection is available for only 2,465 species, and shows a moderate correlation with the RGB measure of sexual dichromatism ($r = 0.34$; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
@@
 
-We obtained estimates of species range size using expert range maps [@birdlife_2017].
-Because of taxonomic changes to 1,230 species in the Birdlife database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015].
-From these distributions, we obtained estimates of climatic conditions that each species is exposed to.
-That is, we took 1,000 random point samples in each range and extracted 19 present-day bioclimatic variables (representing a variety of biologically relevant annual trends in temperature and precipitation) with 30-second (~1 km^2^) spatial resolution [@Fick_2017].
-From these values, we obtained means and standard deviations for each species.
+**We obtained estimates of species range size using expert range maps [@birdlife_2017].
+The names of 1,230 species in the Birdlife database [@delhoyo_2016] have been recently changed, so we manually matched these taxa with the names used in the sexual dichromatism  dataset [@delhoyo_2016].
+For each species’ range, we obtained estimates of climatic conditions by extracting 1,000 random point samples of each bioclimatic variable.
+We extracted 19 present-day bioclimatic variables (representing a variety of biologically relevant annual trends in temperature  and precipitation) with 30-second (~1 km^2^) spatial resolution [@Fick_2017].
+From the 1000 values of each bioclimatic variable, we obtained means and standard deviations for each species.
 Using the same spatial sampling, we extracted  means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006].
@@
 Estimates of NPP had 30-second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005].
-Using these data, which we provide as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}), we generated five predictors of speciation associated with different patterns in environmental variability (see below).
-\phantomsection
+We provide these data as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}).**  \phantomsection
 
@@
 Given that stressful environments are expected to interact with sexual selection and have a positive effect on adaptation [@Cally_2019], we used the extracted environmental variables from each species range size to create predictors of environmental variation/stress.
-Firstly we used (*i*) the average NPP in each species' range and (*ii*) the log-transformed range size as potentially informative predictors of speciation rates.
+We used (*i*) the average NPP in each species' range and (*ii*) the log-transformed range size as potentially informative predictors of speciation rates.
 We also used three environmental predictors derived from bioclimatic data.
@@
 (*iv*) To estimate levels of spatial environmental variation a species may endure, we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature and precipitation seasonality (BIO4 and BIO15).
-PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 **largely** reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
-**A taxon’s range size often correlates with speciation and extinction rates [@Rosenzweig_1995; @Castiglione_2017]**, so we controlled for the correlation between environmental spatial variation and species' range sizes --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
-To obtain long-term variation in climates for each species range we take (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
+PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 largely reflects the variation in temperature across a species' range (\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
+A taxon’s range size often correlates with speciation and extinction rates [@Rosenzweig_1995; @Castiglione_2017], so we controlled for the correlation between environmental spatial variation and species' range sizes --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
+To obtain long-term variation in climates for each species range we **took** (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
 Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
@@
 Additionally, a random draw of full trees (including species without genetic data) from the posterior distribution of phylogenetic trees was used for diversification analyses using tip-rate measures (1,000 trees) and BAMM (100 trees) [@Rabosky_2014].
-**These trees had crown clades with a topology that was heavily constrained on the basis of a previously published study [‘Hacket backbone’; @Hackett_2008] and were constructed using a pure birth (Yule) model.** We calculated three different tip-rate metrics of speciation and one of extinction across all trees.
+These trees had crown clades with a topology that was heavily constrained on the basis of a previously published study [‘Hacket backbone’; @Hackett_2008] and were constructed using a pure birth (Yule) model.
+We calculated three different tip-rate metrics of speciation and one of extinction across all trees.
 
-**Diversification is the result of two processes, speciation and extinction through time.
-To estimate speciation rates, we first obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.** Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES; also referred to as diversification rate/DR) **is derived from the sum of edge lengths from each tip towards the root, with each edge towards the root having the length down-weighted [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018].** Crucially, studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
+Diversification is the result of two processes, speciation and extinction through time.
+To estimate speciation rates, we first obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny.
+Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES; also referred to as diversification rate/DR) is derived from the sum of edge lengths from each tip towards the root, with each edge towards the root having the length down-weighted [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018].
+Crucially, studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
 Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path.
@@
 
-We used BAMM to model the dynamics of speciation and extinction across the **101 phylogenetic trees (one MCC tree and 100 random draws of the posterior)**.
+We used BAMM to model the dynamics of speciation and extinction across the 101 phylogenetic trees (one MCC tree and 100 random draws of the posterior).
 This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate probability distributions of evolutionary rate-shift configurations with variable speciation and extinction rates [@Rabosky_2014].
-**These models provide tip-rate estimates of speciation and extinction rate that can be easily used in comparative analyses**.
-The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{Supplementary Information}; **briefly, we used a time-variable model with the prior expected number of evolutionary rate shifts set at 100 and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014].
-BAMM models were run independently for the 101 phylogenetic trees for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
-Important BAMM parameters (log-likelihood and number of rate shifts) reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200; an arbitrary value, above which posterior distributions can often be accurately inferred** (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
+These models provide tip-rate estimates of speciation and extinction rate that can be easily used in comparative analyses.
+The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{Supplementary Information}; briefly, we used a time-variable model with the prior expected number of evolutionary rate shifts set at 100 and prior rates set from the **initial tip-level estimates of speciation and extinction** using the ``BAMMtools`` R package [@bammtools_2014].
+**BAMM models were run independently for the 101 phylogenetic trees for 100 million generations.
+Given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.** Important BAMM parameters (log-likelihood and number of rate shifts) reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200; an arbitrary value, above which posterior distributions can often be accurately inferred (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
 Further details of BAMM parameters and output are available in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}, with tip-rate means and variances provided.
@@
 Firstly, we conducted model selection to compare models in which $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ were the response variable: these tip-rate estimates all came from the same MCC tree (derived from 2,500 draws of the posterior distribution [@Jetz_2012]).
-For models of $\lambda_{BAMM}$ and $\mu_{BAMM}$ we used the inverse of the variance associated with each tip rate estimate as weights, to account for the variable precision of the estimates provided by BAMM.
-**For each response variable, we conducted model selection to compare models with different combinations of predictor variables.
-** The most complex model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (i.e.
+For models of $\lambda_{BAMM}$ and $\mu_{BAMM}$, we used the inverse of the variance associated with each tip rate estimate as weights, to account for the variable precision of the estimates provided by BAMM.
+For each response variable, we conducted model selection to compare models with different combinations of predictor variables.
+The most complex model in each set under comparison contained one of the measures of sexual selection (sexual dichromatism or the index of male-biased sexual selection), all of the environmental measures (i.e.
 log-transformed range size, seasonal temperature variation, spatial temperature variation, long-term temperature variation, and NPP), and all of the 2-way interactions between sexual selection and each of the environmental measures.
@@
 Using the terms from the top-ranked model (ranked by AICc), we ran the equivalent model for each of the 1,000 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and each of the 100 trees used to derive $\lambda_{BAMM}$ and $\mu_{BAMM}$.
-**Additionally, we investigated the effect of the individual variables used to derive the index of male-biased sexual selection on speciation rate.
-For these pgls models we replaced the composite index score with the individual biological variable (sexual size dimorphism, social polygyny and [lack of] paternal care) and ran the equivalent model for 300 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and 100 trees used to derive $\lambda_{BAMM}$**.
+Additionally, we investigated the effect of the individual variables used to derive the index of male-biased sexual selection on speciation rate.
+For these pgls models we replaced the composite index score with the individual biological variable (sexual size dimorphism, social polygyny and [lack of] paternal care) and ran the equivalent model for 300 phylogenetic trees used to derive $\lambda_{DR}$, $\lambda_{ND}$ and 100 trees used to derive $\lambda_{BAMM}$.
 
-**Across all our analyses we corrected for the phylogenetic signal.
-Our models used the unique response variables and correlation structure for a given phylogenetic tree**.
-Specifically, for models using tip-rate metrics ($\lambda_{DR}$, $\lambda_{ND}$) we estimated the phylogenetic signal --- Pagel's $\lambda$ [@Pagel_1999] --- using the ``corPagel`` function in the ``ape`` package [@ape_2004] independently for each of the 1,000 trees/models.
-Alternatively, for models using speciation and extinction estimates derived using BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$) we found $\lambda$ was consistently estimated at 1 and hence assumed Brownian motion (using the ``corBrownian`` function) to estimate the correlation structure.
+Across all our analyses we corrected for the phylogenetic signal.
+**Our models used the unique response variables and correlation structure for a given phylogenetic tree.
+Specifically, for models using tip-rate metrics ($\lambda_{DR}$, $\lambda_{ND}$) we estimated the phylogenetic signal independently for each of the 1,000 trees/models.
+Phylogenetic signal was estimated as Pagel's $\lambda$ [@Pagel_1999] using the ``corPagel`` function in the ``ape`` package [@ape_2004]**.
+Alternatively, for models using speciation and extinction estimates derived using BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), we found that $\lambda$ was consistently estimated at 1 and hence assumed Brownian motion (using the ``corBrownian`` function) to estimate the correlation structure.
 This method enabled us to present model estimates for an MCC tree alongside 1,000/100 trees from the posterior distribution of trees to account for phylogenetic uncertainty.
@@
 Predictors included one measure of sexual selection (one of two measures of sexual dichromatism or the index of male-biased sexual selection), four measures of environmental variation, range size and 2-way interactions between the measure of sexual selection and each of the environmental measures including range size, with subsequent model simplification using AIC model selection.
-**In all cases the best model had a $\Delta$ AICc above 20 against the second model, and an evidence ratio above 100.**
+In all cases the best model had a $\Delta$ AICc above 20 against the second model, and an evidence ratio above 100.
 
@@
 However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; $\lambda_{BAMM}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b).
-When we took into account phylogenetic uncertainty by running the models using 1,000 trees, the distribution of estimates from PGLS models was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r sanitize.numbers(formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; ) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+When we took into account phylogenetic uncertainty by running the models using 1,000 trees, the distribution of estimates from PGLS models was similar to the estimate from the MCC tree: among the 1,000 trees there was a positive association between sexual selection and $\lambda_{DR}$ (highest posterior density (HPD) Interval = `r sanitize.numbers(formatC(Subset.hpd.DR.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`), and the distribution skewed towards a positive association between sexual selection and $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.ND.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`) as well as the 100 models using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(Subset.hpd.Lambda.top$Sexual_selection_ppca, format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
 
-**Given that this male-biased sexual selection index comprises of three different (but correlated) variables, we investigated which of these variables was driving the association observed with $\lambda_{DR}$.
-Our results over 300 trees showed that this pattern is mainly driven by the sexual size dimorphism component (HPD Interval = `r sanitize.numbers(formatC(dr.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), with the effects of other components overlapping zero; paternal care (HPD Interval = `r sanitize.numbers(formatC(dr.paternal, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`) and mating system (HPD Interval= `r sanitize.numbers(formatC(dr.mating, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`).
-Importantly, the association between sexual size dimorphism and speciation rates is also present when using $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(nd.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), but not when using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(b.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`, \autoref{SS_components}).**
+**We investigated which of the three variables comprising the index of male-biased sexual selection was driving the association observed with $\lambda_{DR}$.** Our results over 300 trees showed that this pattern is mainly driven by the sexual size dimorphism component (HPD Interval = `r sanitize.numbers(formatC(dr.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), with the effects of other components overlapping zero; paternal care (HPD Interval = `r sanitize.numbers(formatC(dr.paternal, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`) and mating system (HPD Interval= `r sanitize.numbers(formatC(dr.mating, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`).
+Importantly, the association between sexual size dimorphism and speciation rates is also present when using $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(nd.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), but not when using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(b.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`, \autoref{SS_components}).
 
@@
 
-**Our analyses also show that the differences in results between sexual dichromatism and male-biased sexual selection (i.e.
+Our analyses also show that the differences in results between sexual dichromatism and male-biased sexual selection (i.e.
 association with speciation rates only for the latter) were not due to differences in the size of the datasets used (5,812 species vs.
-2,465,** \href{https://justincally.github.io/SexualSelection_Speciation/#testing_the_sexual_dichromatism_dataset_using_the_same_data_points_as_the_sexual_selection_dataset}{Figure S17}).
-No interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and the index of male-biased sexual selection; $\Delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}).
+2,465, \href{https://justincally.github.io/SexualSelection_Speciation/#testing_the_sexual_dichromatism_dataset_using_the_same_data_points_as_the_sexual_selection_dataset}{Figure S17}).
+No interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and the index of male-biased sexual selection; $\Delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}).
 Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation or range size.
@@
 Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation, we found a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$.
-This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r sanitize.numbers(formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 2)`; $\lambda_{ND}$: $\beta$ = `r sanitize.numbers(formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 2)`; \autoref{model_results}a, \autoref{Speciation_Figure}).
+This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r sanitize.numbers(formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r sanitize.numbers(formatC(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; $\lambda_{ND}$: $\beta$ = `r sanitize.numbers(formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 2)`; \autoref{model_results}a, \autoref{Speciation_Figure}).
 This association was also evident across the estimates from models using the 1,000 trees ($\lambda_{DR}$: HPD Interval = `r sanitize.numbers(formatC(hpd.DR.top[[2]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; $\lambda_{ND}$: HPD Interval = `r sanitize.numbers(formatC(hpd.ND.top[[2]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`; \autoref{model_results}a).
 Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed equivocal evidence that range size is negatively associated with speciation.
-Range size is **significantly associated with** $\lambda_{DR}$ (\autoref{model_results}b) using data subset for species with an index of male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
-Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
-Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with the highest power using near-complete taxon sampling (*n* = 5,812).
+Range size is significantly associated with $\lambda_{DR}$ (\autoref{model_results}b) using data subset for species with an index of male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$.
+Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S13}).
+Because the range size dataset is the same across the three data subsets, we draw our conclusions from the models with the highest power using near-complete taxon sampling (*n* = 5,812).
 \phantomsection
@@
 
-We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r sanitize.numbers(formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 2)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
+We found no evidence that extinction ($\mu_{BAMM}$) was impacted by the extent of sexual dichromatism for full-taxon sampling ($\beta$ = `r sanitize.numbers(formatC(MCC.Mu.top[["coefficients"]][["SDi"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(MCC.Mu.top)[["SDi","p-value"]], digits = 2)`; \autoref{model_results}a), nor spectrophotometry-based measures of sexual dichromatism (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S13}) or male-biased sexual selection (\autoref{model_results}b, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S16}).
 \phantomsection
@@
 Estimates of the effect of predictor variables on speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % HPD interval across PGLS model estimates from 100 trees was often > 20 times larger than the 95 % confidence interval for estimates from a single PGLS model using the MCC tree.
-This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}).
-Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, **and so it is concerning that there are such large differences in the BAMM estimates between alternative, similarly plausible phylogenies**.
-Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
-Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}, @Harvey_2017).
+This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S9}).
+**Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree.
+However, our results suggest that BAMM estimates between alternative, similarly plausible phylogenies vary substantially.** Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
+Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8}, @Harvey_2017).
 Full details of the BAMM results are presented as supplementary materials.
@@
 
-![Model estimates (**a**) **showing the effect size (i.e.
-slope) of log-range size** and sexual dichromatism on speciation and extinction rates using PGLS analyses with the sexual dichromatism dataset (RGB values, *n* = 5,812).
-(**b**) depicts the scatter plot of speciation rate ($\lambda_{DR}$) and log-range size with the model estimate presented as a dashed line.
-(**c**) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection (*n* = 2,465).
-Similar to (**a**), (**d**) presents model estimates for PGLS analyses using a restricted dataset with measures of an index of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
+![Model estimates (a) showing the effect size (i.e.
+slope) of log-range size and sexual dichromatism on speciation and extinction rates using PGLS analyses with the sexual dichromatism dataset (RGB values, *n* = 5,812).
+(b) depicts the scatter plot of speciation rate ($\lambda_{DR}$) and log-range size with the model estimate presented as a dashed line.
+(c) shows the scatter plot of speciation rate ($\lambda_{DR}$) and male-biased sexual selection (*n* = 2,465).
+Similar to (a), (d) presents model estimates for PGLS analyses using a restricted dataset with measures of an index of male-biased sexual selection (*n* = 2,465) and shows the effect of log-range size and male-biased sexual selection on speciation and extinction rates.
 Both datasets were used for analyses with three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables.
@@
 We found evidence that the composite index of male-biased sexual selection, but not measures of sexual dichromatism, is correlated with the rate of speciation in passerine birds.
-The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB), **and it cannot be explained by a difference in statistical power or sampling**.
+The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB), and it cannot be explained by a difference in statistical power or sampling.
 These findings reaffirm the conclusions of previous, smaller studies in which sexual dichromatism was measured using spectrophotometry [@Huang_2014] or human observers [@Cooney_2017].
 The correlation between speciation rate and the index of male-biased sexual selection (which encapsulates variation in sexual size dimorphism, social polygyny, and paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$.
-**This pattern seems to be mainly driven by an association between sexual size dimorphism and speciation**.
+This pattern seems to be mainly driven by an association between sexual size dimorphism and speciation.
 Interestingly, we also found a consistent negative relationship between range size and speciation rate, at least when this rate was quantified using $\lambda_{DR}$ and $\lambda_{ND}$.
-None of the bioclimatic measures of environmental variability that we investigated (i.e., temperature seasonality, long-term temperature variation, and spatial temperature variation) **were significantly associated speciation rate, nor mediated the relationship between sexual selection and diversification**.
+None of the bioclimatic measures of environmental variability that we investigated (i.e., temperature seasonality, long-term temperature variation, and spatial temperature variation) were significantly associated speciation rate, nor mediated the relationship between sexual selection and diversification.
 
@@
 For example, in superb fairy-wrens (*Malurus cyaneus*) female colouration has probably evolved in response to spatial variation in predation pressure, increasing dichromatism [@Medina_2017].
-**In fact, our path analysis detected a weak relationship between environment and sexual dichromatism, where sexual dichromatism was positively predicted by temperature seasonality (a measure of environmental variation).**
+In fact, our path analysis detected a weak relationship between environment and sexual dichromatism, where sexual dichromatism was positively predicted by temperature seasonality (a measure of environmental variation).
 
@@
 In line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least when speciation is measured by $\lambda_{DR}$.
-**Many of the species that have both high scores of male-biased sexual selection and high diversification rates belong to the genera _Ploceus_, _Euplectes_ (Ploceidae) and _Paradisaea_ (Paradiaseidae).
+Many of the species that have both high scores of male-biased sexual selection and high diversification rates belong to the genera _Ploceus_, _Euplectes_ (Ploceidae) and _Paradisaea_ (Paradiaseidae).
 Multiple weaver species (Ploceidae) are polygynous and lack paternal care, and both weavers and birds of paradise have strong size dimorphism.
@@
 Size dimorphism is often thought to arise as a consequence of intrasexual competition, where one of the sexes (males in most birds) has to compete for access to the other sex, leading to selection for larger body sizes and thus greater dimorphism [@bjorklund_1990; @owens_1998].
-Therefore, competition between males could be the underlying driver of the high speciation rates that we detect in some clades.**
+Therefore, competition between males could be the underlying driver of the high speciation rates that we detect in some clades.
 
-**Sexual dimorphism due to competition within sexes contrasts with the drivers of sexual dichromatism.
+Sexual dimorphism due to competition within sexes contrasts with the drivers of sexual dichromatism.
 Plumage dichromatism can evolve as a consequence of female cryptic choice and be related to extra-bond fertilizations, but not necessarily paternal care or mating system [@owens_1998].
@@
 Differences in body size (as a result of sexual selection) could be linked to changes in diet, vulnerability to predators or environmental tolerance [@Damuth_1993; @Liow_2008; @Bonduriansky_2011], and such differences could ultimately increase the likelihood of divergence between young lineages.
-In mammals, sexual selection is suggested to have driven the evolution of large body size which in turn has allowed diversification of ecological strategies in the clade, and higher speciation rates [@Mclain_1993; @Bonduriansky_2011].**
+In mammals, sexual selection is suggested to have driven the evolution of large body size which in turn has allowed diversification of ecological strategies in the clade, and higher speciation rates [@Mclain_1993; @Bonduriansky_2011].
 
-**We also found that the association between sexual selection and speciation appears to be independent of net primary productivity and spatiotemporal variation in the environment**.
+We also found that the association between sexual selection and speciation appears to be independent of net primary productivity and spatiotemporal variation in the environment.
 The lack of an effect of these environmental variables on speciation rate has several possible interpretations.
@@
 
-**We found that species with smaller ranges have elevated speciation rates**.
+We found that species with smaller ranges have elevated speciation rates.
 This result is similar to a study of 329 amphibian genera, which found higher diversification rates in taxa with smaller range size [@Greenberg_2017].
@@
 However, the opposite pattern is also plausible because birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
-Moreover, species that have recently split as a consequence of vicariant divergence might have smaller ranges as a result of the split of the ancestral lineage, leading to link between smaller ranges and shorter divergence times.
+Moreover, species that have recently split as a consequence of vicariant divergence might have smaller ranges as a result of the split of the ancestral lineage, leading to **a** link between smaller ranges and shorter divergence times.
 It is also possible that high speciation rates cause smaller range sizes, rather than the other way around, for example because repeatedly-speciating lineages tend to fill niches in ways that hinder the geographical expansion of new species [@Rosenzweig_1995; @Weir_2011; @Price_2014].
 However, species undergoing adaptive radiation in new habitats are unlikely to be limited by competition for resources from existing taxa.
-**One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.**
+One further explanation for the negative association between range size and sexual dichromatism/sexual selection is the potential bias of taxonomic classification, whereby over-splitting of species in clades with large ranges leads to increased recent phylogenetic branching as well as smaller ranges.
 
@@
 For instance, $\lambda_{DR}$ is weighted more towards speciation events close to the tips and allows more rate heterogeneity compared to $\lambda_{BAMM}$ estimates.
-**Rate shifts are unlikely to be detected in smaller clades in BAMM, meaning that it is not uncommon for whole genera to have the same rate.
+Rate shifts are unlikely to be detected in smaller clades in BAMM, meaning that it is not uncommon for whole genera to have the same rate.
 Using the $\lambda_{DR}$ metric, only sister species are guaranteed to have the same rate.
 This leads to greater variation in $\lambda_{DR}$ relative to the $\lambda_{BAMM}$ estimates, which is suggested to be an advantage when studying diversification patterns [@Quintero_2015].
-Additionally, $\lambda_{BAMM}$ estimates were also more sensitive to phylogenetic uncertainty when compared to $\lambda_{DR}$, and estimates are 20 times more variable across trees in $\lambda_{BAMM}$ compared to $\lambda_{DR}$.
-We cannot completely reject the idea that the lack of association between $\lambda_{BAMM}$ and sexual selection could be the result of low statistical power, due to the combination of both low variation across species in the speciation rates and high levels of variation in the estimates across trees.**
+**Additionally, $\lambda_{BAMM}$ estimates were more sensitive to phylogenetic uncertainty and were 20 times more variable across trees in compared to $\lambda_{DR}$ estimates.** We cannot completely reject the idea that the lack of association between $\lambda_{BAMM}$ and sexual selection could be the result of low statistical power, due to the combination of both low variation across species in the speciation rates and high levels of variation in the estimates across trees.
 
@@
 We have also shown that there is no evidence of an association between sexual selection and extinction rates.
-Overall, our findings imply that male-male competition could be the mechanism driving increased speciation rates in birds, that sexual dichromatism may not a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
+Overall, our findings imply that male-male competition could be the mechanism driving increased speciation rates in birds, that sexual dichromatism may not **be** a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
 

### 2020-12-03 Sexual Selection and Speciation.Rmd
submitted 2nd round
third round of revisions

@@
 
-To obtain speciation rate estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM).
+To obtain speciation rate estimates, we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM).
 Additionally, we compared our BAMM output files (event data)  to an existing rate dataset [@Harvey_2017].
@@
 
-In this study we assess the relationship between sexual selection and extinction risk.
+In this study, we assess the relationship between sexual selection and extinction risk.
 In doing so we attempt to take into account as many other predictors of extinction as possible, primarily through environmental variables.
@@
 
-To obtain estimates of spatial variability in environment we performed PCAs of the standard errors of the bioclimatic variables across the 1000 random points extracted per species (excluding seasonality in temperature and precipitation).
+To obtain estimates of spatial variability in environment, we performed PCAs of the standard errors of the bioclimatic variables across the 1000 random points extracted per species (excluding seasonality in temperature and precipitation).
 We expect variability to increase with increased range size.
@@
 
-To check the correlations between the environmental predictors to be used in the PGLS models we inspected the following correlation plots with the correlation value plotted:
+To check the correlations between the environmental predictors to be used in the PGLS models, we inspected the following correlation plots with the correlation value plotted:
 
@@
 
-For the use of BAMM we used the following code to generate the parameters across the 100 trees.
+For the use of BAMM, we used the following code to generate the parameters across the 100 trees.
 Each parameter value is specified in this code chunk to ensure reproducibility.
@@
 
-Now we have a series of trees and event data we can read in. Firstly and most importantly let's check for convergence in the MCC tree through the following:
-
-**Table S3:** Effective sample size (ESS) for the two key BAMM parameters (number of evolutionary shifts and log-Likelihood) for the run on the MCC indicate that BAMM converges (ESS > 200).
+We checked the convergence of BAMM results by assessing effective sample size (ESS).
+
+**Table S3:** ESS for the two key BAMM parameters (number of evolutionary shifts and log-Likelihood) for the run on the MCC indicate that BAMM converges (ESS > 200).
 
@@
 
-Those values are well above 200, so we can use the MCC data in analysis!
 We can also check the convergence of BAMM across 100 runs of BAMM.
@@
 This enabled us to show the variation between trees, and the variation associated with the top MCC model.
-Given the variability across trees we ran 1000 ND and DR models, each allowed to estimate a phylogenetic signal ($\lambda$).
+Given the variability across phylogenetic trees we ran 1000 ND and DR models.
+Each model had an independently derived phylogenetic signal ($\lambda$) based on the unique phylogenetic tree used.
 The models were saved as an `.rds` object and read back into R with key coefficients extracted.
@@
 
-**Table S15:** MCC model estimates from the above model are presented here as numberical values with 95 % CIs.
+**Table S15:** MCC model estimates from the above model are presented here as numerical values with 95 % CIs.
 ```{r Table for SS plots}
@@
 Subset.MCC.ND.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates from Male-bias SS", split.table = Inf)
-Subset.MCC.Lambda.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates from Male-bias SS", split.table = Inf)
-Subset.MCC.Mu.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates from Male-bias SS", split.table = Inf)
-```
-
-**Table S16:** HPD intervals are calculated for the above figure (i.e.
-models using Male-bias sexual selection measures).
-using this measure the models using $\lambda_{DR}$ have model estimates that fall on the positive side 95 % of the time in 100 trees.
+Subset.MCC.Lambda.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates from Male-biased SS", split.table = Inf)
+Subset.MCC.Mu.summary %>% select(-model_name) %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates from Male-biased SS", split.table = Inf)
+```
+
+**Table S16:** HPD intervals for models using male-biased sexual selection measures.
+Across 100 trees, estimates from models using $\lambda_{DR}$ are positive >95% of the time.
 For $\lambda_{ND}$, there is a positive skew, however the 95 % HPD interval overlaps zero.

### 2020-12-07 Manuscript_Publication.Rmd
update with Ilianas comments

@@
 abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates.
-However, the net effect of sexual selection on diversification is hard to predict _a priori_, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes.
+However, the net effect of sexual selection on diversification is hard to predict, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes.
 Theoretical work also suggests that the net effect of sexual selection on diversification should depend strongly on ecological factors, though this prediction has seldom been tested.
@@
 options(knitr.kable.NA = ' ')
+
 literature_table %>%
@@
         escape = T,
-        caption = "Review of studies testing the association between sexual selection and speciation") %>%
+        caption = "Previous studies testing the association between sexual selection and speciation") %>%
   kableExtra::kable_styling(latex_options = c("HOLD_position", "scale_down"),
@@
   kableExtra::column_spec(4, "3em") %>%
-  kableExtra::collapse_rows(1:2, row_group_label_position = "identity") 
+  kableExtra::collapse_rows(1:2, row_group_label_position = "identity") %>% 
+  kableExtra::footnote(general_title = "", 
+                       general = "Studies were obtained by searching 'Web of Science' for articles published from 2011 for terms containing ‘speciation’, ‘diversification’ and ‘sexual selection’.
+We summarised all the studies we found relevant and comparable to our study.", threeparttable = TRUE)
 ```
@@
 The bar under each density ridge is the 95 % Highest Posterior Density Interval.
-Given that the mating system variable is categorical, model estimates for three polygynous mating system levels are in reference to a striclty monogamous mating system (0% polygyny).
+Given that the mating system is a categorical variable, model estimates for three polygynous mating system levels are in reference to a striclty monogamous mating system (0% polygyny).
 \label{SS_components}](Figures/Individual_SS_Estimates.pdf)
@@
 It can also arise as a result of selection on the level of crypsis of the sex that cares for offspring [@Dale_2015].
-The fact that traits linked with competition (such as size dimorphism) are the ones associated with higher $\lambda_{DR}$ values – rather than sexual dichromatism – supports the general view that antagonistic interactions and sexual conflict can lead to increased diversity [@Bonduriansky_2011; @Hoskin_2010; @tsuji2020sexual].
+The fact that traits linked with competition (such as size dimorphism) are the ones associated with higher $\lambda_{DR}$ values – rather than sexual dichromatism – supports the general view that antagonistic interactions and sexual conflict can lead to increased diversity [@Bonduriansky_2011; @Tinghitella_2018; @Qvarnstrom_2012; @tsuji2020sexual].
 Moreover, body size is a trait that influences multiple aspects of the physiology and ecology of a species.
@@
 This leads to greater variation in $\lambda_{DR}$ relative to the $\lambda_{BAMM}$ estimates, which is suggested to be an advantage when studying diversification patterns [@Quintero_2015].
-**Additionally, $\lambda_{BAMM}$ estimates were more sensitive to phylogenetic uncertainty and were 20 times more variable across trees in compared to $\lambda_{DR}$ estimates.** We cannot completely reject the idea that the lack of association between $\lambda_{BAMM}$ and sexual selection could be the result of low statistical power, due to the combination of both low variation across species in the speciation rates and high levels of variation in the estimates across trees.
+**Additionally, $\lambda_{BAMM}$ estimates were more sensitive to phylogenetic uncertainty and were 20 times more variable across trees compared to $\lambda_{DR}$ estimates.** We cannot completely reject the idea that the lack of association between $\lambda_{BAMM}$ and sexual selection could be the result of low statistical power, due to the combination of both low variation across species in the speciation rates and high levels of variation in the estimates across trees.
 

### 2021-01-17 Manuscript_Publication.Rmd
manuscript changes

@@
 ---
-<!-- \maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia} -->
+<!-- \maketitle\unmarkedfntext{\textsuperscript{†}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia \\ -->
+<!-- \textsuperscript{‡}School of Natural and Computational Sciences, Massey University (Albany Campus), Auckland, New Zealand \\ -->
+<!-- \textsuperscript{§}\textit{justin.g.cally@gmail.com}} -->
 
@@
 Furthermore, the diversity of outcomes and approaches in previous studies suggests that the association between species diversity and sexual selection is far from clear (reviewed in @tsuji2020sexual).
-**See Table 1 for a review of the major findings from previous studies testing the association between sexual selection and speciation**.
 
@@
 More recently, social polygyny (a proxy for sexual selection) was found to have a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].
+**We summarize the major findings from previous studies testing the association between sexual selection and speciation in birds and other taxa since @Kraaijeveld_2011 meta-analysis (Table 1).**
 
@@
 Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018].
-Additionally, our analysis includes multiple ecological and environmental variables, allowing us to control for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016].
+Additionally, our analysis includes multiple ecological and environmental variables, allowing us to control for potential confounds, **to identify environmental factors, including spatial and temporal environmental variability, interact with sexual selection as theory predicts** [@Connallon_Hall_2016].
 
@@
 
-Because sexual dichromatism is presumably imperfectly correlated with variation in the strength of sexual selection across taxa, we sourced an additional measure of sexual selection [@Dale_2015], referred to here as the 'index of male-biased sexual selection'.
-This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
+**Sexual dichromatism is likely to be imperfectly correlated with variation in the strength of sexual selection across taxa.
+For this reason, we sourced an additional measure of sexual selection [@Dale_2015], referred to here as the ‘index of male-biased sexual selection'.** This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
 The variables included in this index have all been positively linked to the intensity of sexual selection, and are usually correlated [@bjorklund_1990; @owens_1998; @dunn_2001], which is why they were combined into a single metric in previous studies [@Dale_2015].
@@
 Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root to the tip, while the log-transformed equal splits (logES; also referred to as diversification rate/DR) is derived from the sum of edge lengths from each tip towards the root, with each edge towards the root having the length down-weighted [@Jetz_2012; @quintero_2018; @Rabosky_fish_2018].
-Crucially, studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification; this is because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
+Crucially, studies have suggested that DR and ND (henceforth referred to as $\lambda_{DR}$ and $\lambda_{ND}$) are more reflective estimates of speciation than diversification.
+Because $\lambda_{DR}$ and $\lambda_{ND}$ cannot account for whole-clade extinctions and thus under-estimate extinction rate, which makes the composite measure of diversification more dependent on speciation [@Belmaker_2015; @Rabosky_diversification_2018].
 Therefore, $\lambda_{DR}$ is a measure of speciation rate more heavily weighted to recent speciation events while $\lambda_{ND}$ measures speciation across the root-to-tip path.
@@
 
-We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds; \autoref{Speciation_Figure}).
-To do this, we tested the association between speciation/extinction and sexual selection, environmental variability and their interaction using phylogenetic least squares (PGLS) models with $\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ or $\mu_{BAMM}$ as the response variable.
-These are three different tip-rate metrics of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one of extinction ($\mu_{BAMM}$, see Methods for details).
-Predictors included one measure of sexual selection (one of two measures of sexual dichromatism or the index of male-biased sexual selection), four measures of environmental variation, range size and 2-way interactions between the measure of sexual selection and each of the environmental measures including range size, with subsequent model simplification using AIC model selection.
-In all cases the best model had a $\Delta$ AICc above 20 against the second model, and an evidence ratio above 100.
-
-We found a significant positive association between the index of male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ maximum credibility (MCC) tree ($\beta$ = `r sanitize.numbers(formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b).
+**We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds; \autoref{Speciation_Figure})**.
+We found a significant positive association between the index of male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ from the maximum clade credibility (MCC) tree ($\beta$ = `r sanitize.numbers(formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b).
 However, this association was not significant for the other two measures of speciation rate ($\lambda_{ND}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.ND.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.ND.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; $\lambda_{BAMM}$: $\beta$ = `r sanitize.numbers(formatC(Subset.MCC.Lambda.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.Lambda.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b).
@@
 This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S9}).
-**Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree.
+**The great majority of earlier studies have based their estimates on a single consensus tree due to the computational requirements of BAMM.
 However, our results suggest that BAMM estimates between alternative, similarly plausible phylogenies vary substantially.** Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
-Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8}, @Harvey_2017).
+The calculation of BAMM rates can be affected by the settings of the run and the use of different priors.
+We therefore compared the estimate of our MCC tree with that of previous published analyses on birds and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8}, @Harvey_2017).
 Full details of the BAMM results are presented as supplementary materials.
@@
 
+<!-- \phantomsection -->
+<!-- # Author Contributions   -->
+
+<!-- All authors contributed to conception, design and approach to analyses; J.G.C.
+and J.D.
+compiled the data; J.G.C.
+and I.M.
+conducted analyses; J.G.C.
+wrote the first draft of the manuscript; all authors contributed substantially to further manuscript revisions.
+-->
+
+<!-- \phantomsection -->
+<!-- # Data Accessibility -->
+<!-- We have sought to make the data and Supplementary Material (including `R` code) freely accessible to readers.
+The code and Supplementary Material for this manuscript are available on \href{https://justincally.github.io/SexualSelection_Speciation/}{GitHub} and as an attached HTML document (Supplementary Material) for reviewers.
+Data is archived in the adjoined GitHub repository and will be made publicly available on acceptance.
+-->
+
 \newpage\phantomsection

### 2021-01-17 Manuscript_Publication.Rmd
commas and title page

@@
 author: 
+- Justin G. Cally\textsuperscript{†§}, Devi Stuart-Fox\textsuperscript{†}, Luke Holman\textsuperscript{†}, James Dale\textsuperscript{‡} and Iliana Medina\textsuperscript{†}
 bibliography: References/SS_Extinction.bib
@@
 ---
-<!-- \maketitle\unmarkedfntext{\textsuperscript{†}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia \\ -->
-<!-- \textsuperscript{‡}School of Natural and Computational Sciences, Massey University (Albany Campus), Auckland, New Zealand \\ -->
-<!-- \textsuperscript{§}\textit{justin.g.cally@gmail.com}} -->
+\maketitle\unmarkedfntext{\textsuperscript{†}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia \\
+\textsuperscript{‡}School of Natural and Computational Sciences, Massey University (Albany Campus), Auckland, New Zealand \\
+\textsuperscript{§}\textit{justin.g.cally@gmail.com}}
 
@@
 More recently, social polygyny (a proxy for sexual selection) was found to have a positive association with speciation rate across 954 species of birds [@Iglesias-Carrasco_2019].
-**We summarize the major findings from previous studies testing the association between sexual selection and speciation in birds and other taxa since @Kraaijeveld_2011 meta-analysis (Table 1).**
+We summarize the major findings from previous studies testing the association between sexual selection and speciation in birds and other taxa since @Kraaijeveld_2011 meta-analysis (Table 1).
 
@@
 Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018].
-Additionally, our analysis includes multiple ecological and environmental variables, allowing us to control for potential confounds, **to identify environmental factors, including spatial and temporal environmental variability, interact with sexual selection as theory predicts** [@Connallon_Hall_2016].
+Additionally, our analysis includes multiple ecological and environmental variables, allowing us to control for potential confounds, to identify environmental factors, including spatial and temporal environmental variability, interact with sexual selection as theory predicts [@Connallon_Hall_2016].
 
@@
 
-**Sexual dichromatism is likely to be imperfectly correlated with variation in the strength of sexual selection across taxa.
-For this reason, we sourced an additional measure of sexual selection [@Dale_2015], referred to here as the ‘index of male-biased sexual selection'.** This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
+Sexual dichromatism is likely to be imperfectly correlated with variation in the strength of sexual selection across taxa.
+For this reason, we sourced an additional measure of sexual selection [@Dale_2015], referred to here as the ‘index of male-biased sexual selection'.
+This index is the first principal component from a phylogenetic principal component analysis (PPCA) of three characteristics possitively associated with sexual selection (sexual size dimorphism, social polygyny and [lack of] paternal care).
 The variables included in this index have all been positively linked to the intensity of sexual selection, and are usually correlated [@bjorklund_1990; @owens_1998; @dunn_2001], which is why they were combined into a single metric in previous studies [@Dale_2015].
@@
 
-**We obtained estimates of species range size using expert range maps [@birdlife_2017].
+We obtained estimates of species range size using expert range maps [@birdlife_2017].
 The names of 1,230 species in the Birdlife database [@delhoyo_2016] have been recently changed, so we manually matched these taxa with the names used in the sexual dichromatism  dataset [@delhoyo_2016].
@@
 Estimates of NPP had 30-second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005].
-We provide these data as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}).**  \phantomsection
+We provide these data as a potentially useful data resource (see \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}).
+\phantomsection
 
@@
 A taxon’s range size often correlates with speciation and extinction rates [@Rosenzweig_1995; @Castiglione_2017], so we controlled for the correlation between environmental spatial variation and species' range sizes --- where larger ranges have larger variation in PC1 --- by using the residuals of a fitted general additive model (GAM; \href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1}) as a predictor.
-To obtain long-term variation in climates for each species range we **took** (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
+To obtain long-term variation in climates for each species range, we took (*v*) the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values.
 Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates (\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
@@
 These models provide tip-rate estimates of speciation and extinction rate that can be easily used in comparative analyses.
-The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{Supplementary Information}; briefly, we used a time-variable model with the prior expected number of evolutionary rate shifts set at 100 and prior rates set from the **initial tip-level estimates of speciation and extinction** using the ``BAMMtools`` R package [@bammtools_2014].
-**BAMM models were run independently for the 101 phylogenetic trees for 100 million generations.
-Given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.** Important BAMM parameters (log-likelihood and number of rate shifts) reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200; an arbitrary value, above which posterior distributions can often be accurately inferred (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
+The parameters of the 100 BAMM runs are detailed in full in the \href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{Supplementary Information}; briefly, we used a time-variable model with the prior expected number of evolutionary rate shifts set at 100 and prior rates set from the initial tip-level estimates of speciation and extinction using the ``BAMMtools`` R package [@bammtools_2014].
+BAMM models were run independently for the 101 phylogenetic trees for 100 million generations.
+Given the computationally intensive nature of BAMM, runs were conducted across multiple CPUs.
+Important BAMM parameters (log-likelihood and number of rate shifts) reached convergence with effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples surpassing 200; an arbitrary value, above which posterior distributions can often be accurately inferred (\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3}, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
 Further details of BAMM parameters and output are available in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}, with tip-rate means and variances provided.
@@
 Across all our analyses we corrected for the phylogenetic signal.
-**Our models used the unique response variables and correlation structure for a given phylogenetic tree.
-Specifically, for models using tip-rate metrics ($\lambda_{DR}$, $\lambda_{ND}$) we estimated the phylogenetic signal independently for each of the 1,000 trees/models.
-Phylogenetic signal was estimated as Pagel's $\lambda$ [@Pagel_1999] using the ``corPagel`` function in the ``ape`` package [@ape_2004]**.
+Our models used the unique response variables and correlation structure for a given phylogenetic tree.
+Specifically, for models using tip-rate metrics ($\lambda_{DR}$, $\lambda_{ND}$), we estimated the phylogenetic signal independently for each of the 1,000 trees/models.
+Phylogenetic signal was estimated as Pagel's $\lambda$ [@Pagel_1999] using the ``corPagel`` function in the ``ape`` package [@ape_2004].
 Alternatively, for models using speciation and extinction estimates derived using BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), we found that $\lambda$ was consistently estimated at 1 and hence assumed Brownian motion (using the ``corBrownian`` function) to estimate the correlation structure.
@@
 That is, a phylogenetic path analysis allowed us to model relationships between the predictor variables used in our PGLS analysis as we anticipate environmental variability, sexual dichromatism/selection, and range size to have effects on each other and not just on speciation rate.
-To minimise path complexity we used temperature seasonality (BIO4) as the single measure for environmental variability, $\lambda_{DR}$ as the single measure of speciation and the tip-rates from the MCC tree.
+To minimise path complexity, we used temperature seasonality (BIO4) as the single measure for environmental variability, $\lambda_{DR}$ as the single measure of speciation, and the tip-rates from the MCC tree.
 Further details of the path analysis, including our rationale for each path's directions, can be found within the \href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information} along with all other analyses and the relevant ``R`` code to reproduce results.
@@
 
-**We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds; \autoref{Speciation_Figure})**.
+We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds; \autoref{Speciation_Figure}).
 We found a significant positive association between the index of male-biased sexual selection (*n* = 2,465) and $\lambda_{DR}$ from the maximum clade credibility (MCC) tree ($\beta$ = `r sanitize.numbers(formatC(Subset.MCC.DR.top[["coefficients"]][["Sexual_selection_ppca"]], format = "e", digits = 2), type = "latex", math.style.exponents = TRUE)`, p = `r round(anova(Subset.MCC.DR.top)[["Sexual_selection_ppca","p-value"]], digits = 2)`; \autoref{model_results}b).
@@
 
-**We investigated which of the three variables comprising the index of male-biased sexual selection was driving the association observed with $\lambda_{DR}$.** Our results over 300 trees showed that this pattern is mainly driven by the sexual size dimorphism component (HPD Interval = `r sanitize.numbers(formatC(dr.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), with the effects of other components overlapping zero; paternal care (HPD Interval = `r sanitize.numbers(formatC(dr.paternal, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`) and mating system (HPD Interval= `r sanitize.numbers(formatC(dr.mating, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`).
+We investigated which of the three variables comprising the index of male-biased sexual selection was driving the association observed with $\lambda_{DR}$.
+Our results over 300 trees showed that this pattern is mainly driven by the sexual size dimorphism component (HPD Interval = `r sanitize.numbers(formatC(dr.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), with the effects of other components overlapping zero; paternal care (HPD Interval = `r sanitize.numbers(formatC(dr.paternal, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`) and mating system (HPD Interval= `r sanitize.numbers(formatC(dr.mating, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`).
 Importantly, the association between sexual size dimorphism and speciation rates is also present when using $\lambda_{ND}$ (HPD Interval = `r sanitize.numbers(formatC(nd.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`), but not when using $\lambda_{BAMM}$ (HPD Interval = `r sanitize.numbers(formatC(b.wing, format = "e", digits = 2, drop0trailing = T), type = "latex", math.style.exponents = T)`, \autoref{SS_components}).
@@
 This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for PGLS model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S9}).
-**The great majority of earlier studies have based their estimates on a single consensus tree due to the computational requirements of BAMM.
-However, our results suggest that BAMM estimates between alternative, similarly plausible phylogenies vary substantially.** Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
+The great majority of earlier studies have based their estimates on a single consensus tree due to the computational requirements of BAMM.
+However, our results suggest that BAMM estimates between alternative, similarly plausible phylogenies vary substantially.
+Mean measures of speciation rate across 100 trees were positively correlated between measures ($\lambda_{DR}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, $\lambda_{DR}$ - $\lambda_{ND}$: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, $\lambda_{ND}$ - $\lambda_{BAMM}$: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
 The calculation of BAMM rates can be affected by the settings of the run and the use of different priors.
-We therefore compared the estimate of our MCC tree with that of previous published analyses on birds and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8}, @Harvey_2017).
+We therefore compared the estimate of our MCC tree with that of previously published analyses on birds and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8}, @Harvey_2017).
 Full details of the BAMM results are presented as supplementary materials.
@@
 
-In line with some theoretical predictions and previous studies [@Kraaijeveld_2011] we found that male-biased sexual selection increases speciation rate, at least when speciation is measured by $\lambda_{DR}$.
+In line with some theoretical predictions and previous studies [@Kraaijeveld_2011], we found that male-biased sexual selection increases speciation rate, at least when speciation is measured by $\lambda_{DR}$.
 Many of the species that have both high scores of male-biased sexual selection and high diversification rates belong to the genera _Ploceus_, _Euplectes_ (Ploceidae) and _Paradisaea_ (Paradiaseidae).
@@
 However, the opposite pattern is also plausible because birds with limited dispersal or more specialised niches can have more fragmented populations, which would promote vicariant divergence and higher speciation rates [@Birand_2012; @Claramunt_2012; @Jablonski_2003].
-Moreover, species that have recently split as a consequence of vicariant divergence might have smaller ranges as a result of the split of the ancestral lineage, leading to **a** link between smaller ranges and shorter divergence times.
+Moreover, species that have recently split as a consequence of vicariant divergence might have smaller ranges as a result of the split of the ancestral lineage, leading to a link between smaller ranges and shorter divergence times.
 It is also possible that high speciation rates cause smaller range sizes, rather than the other way around, for example because repeatedly-speciating lineages tend to fill niches in ways that hinder the geographical expansion of new species [@Rosenzweig_1995; @Weir_2011; @Price_2014].
@@
 This leads to greater variation in $\lambda_{DR}$ relative to the $\lambda_{BAMM}$ estimates, which is suggested to be an advantage when studying diversification patterns [@Quintero_2015].
-**Additionally, $\lambda_{BAMM}$ estimates were more sensitive to phylogenetic uncertainty and were 20 times more variable across trees compared to $\lambda_{DR}$ estimates.** We cannot completely reject the idea that the lack of association between $\lambda_{BAMM}$ and sexual selection could be the result of low statistical power, due to the combination of both low variation across species in the speciation rates and high levels of variation in the estimates across trees.
+Additionally, $\lambda_{BAMM}$ estimates were more sensitive to phylogenetic uncertainty and were 20 times more variable across trees compared to $\lambda_{DR}$ estimates.
+We cannot completely reject the idea that the lack of association between $\lambda_{BAMM}$ and sexual selection could be the result of low statistical power, due to the combination of both low variation across species in the speciation rates and high levels of variation in the estimates across trees.
 
@@
 We have also shown that there is no evidence of an association between sexual selection and extinction rates.
-Overall, our findings imply that male-male competition could be the mechanism driving increased speciation rates in birds, that sexual dichromatism may not **be** a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
+Overall, our findings imply that male-male competition could be the mechanism driving increased speciation rates in birds, that sexual dichromatism may not be a reliable proxy for sexual selection, and that alternative measures of sexual selection are more directly related to diversification.
 
-<!-- \phantomsection -->
-<!-- # Acknowledgements   -->
+\phantomsection
+# Acknowledgements
 
-<!-- We thank Adnan Moussalli for his assistance in conducting BAMM runs across 100 phylogenetic trees, Ignacio Quintero for providing BAMM results for comparison and for discussion of the results as well as Chris Cooney for comments on the manuscript.
--->
+We thank Adnan Moussalli for his assistance in conducting BAMM runs across 100 phylogenetic trees, Ignacio Quintero for providing BAMM results for comparison and for discussion of the results as well as Chris Cooney for comments on the manuscript.
 
-<!-- \phantomsection -->
-<!-- # Author Contributions   -->
+\phantomsection
+# Author Contributions
 
-<!-- All authors contributed to conception, design and approach to analyses; J.G.C.
+All authors contributed to conception, design and approach to analyses; J.G.C.
 and J.D.
@@
 wrote the first draft of the manuscript; all authors contributed substantially to further manuscript revisions.
--->
 
-<!-- \phantomsection -->
-<!-- # Data Accessibility -->
-<!-- We have sought to make the data and Supplementary Material (including `R` code) freely accessible to readers.
+\phantomsection
+# Data Accessibility
+We have sought to make the data and Supplementary Material (including `R` code) freely accessible to readers.
 The code and Supplementary Material for this manuscript are available on \href{https://justincally.github.io/SexualSelection_Speciation/}{GitHub} and as an attached HTML document (Supplementary Material) for reviewers.
 Data is archived in the adjoined GitHub repository and will be made publicly available on acceptance.
--->
 

### 2021-01-18 Manuscript_Publication.Rmd
add tex

@@
 Sexual dimorphism due to competition within sexes contrasts with the drivers of sexual dichromatism.
-Plumage dichromatism can evolve as a consequence of female cryptic choice and be related to extra-bond fertilizations, but not necessarily paternal care or mating system [@owens_1998].
+Plumage dichromatism can evolve as a consequence of female cryptic choice and be related to extra-pair fertilizations, but not necessarily paternal care or mating system [@owens_1998].
 It can also arise as a result of selection on the level of crypsis of the sex that cares for offspring [@Dale_2015].
@@
 The code and Supplementary Material for this manuscript are available on \href{https://justincally.github.io/SexualSelection_Speciation/}{GitHub} and as an attached HTML document (Supplementary Material) for reviewers.
-Data is archived in the adjoined GitHub repository and will be made publicly available on acceptance.
+Data is archived in the adjoined GitHub repository and the environmental dataset collated for this manuscript is available on Dryad: https://doi.org/10.5061/dryad.573n5tb6n.
 

### 2021-01-18 Manuscript_Publication.tex
add tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds},
            pdfauthor={Justin G. Cally, Devi Stuart-Fox, Luke Holman, James Dale and Iliana Medina},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Male-biased sexual selection, but not sexual dichromatism, predicts
speciation in birds}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Justin G. Cally\textsuperscript{†§}, Devi Stuart-Fox\textsuperscript{†},
Luke Holman\textsuperscript{†}, James Dale\textsuperscript{‡} and Iliana
Medina\textsuperscript{†}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath} \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling} \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large} \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}} \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption} \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt} \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip} \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape} \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{} \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{} \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang, flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false, linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}} \raggedbottom

\begin{document}
\maketitle
\begin{abstract}
Sexual selection is thought to shape phylogenetic diversity by affecting
speciation or extinction rates. However, the net effect of sexual
selection on diversification is hard to predict, because many of the
hypothesised effects on speciation or extinction have opposing signs and
uncertain magnitudes. Theoretical work also suggests that the net effect
of sexual selection on diversification should depend strongly on
ecological factors, though this prediction has seldom been tested. Here,
we test whether variation in sexual selection can predict speciation and
extinction rates across passerine birds (up to 5,812 species, covering
most genera) and whether this relationship is mediated by environmental
factors. Male-biased sexual selection, and specifically sexual size
dimorphism, predicted two of the three measures of speciation rates that
we examined. The link we observed between sexual selection and
speciation was independent of environmental variability, though species
with smaller ranges had higher speciation rates. There was no
association between any proxies of sexual selection and extinction rate.
Our findings support the view that male-biased sexual selection, as
measured by frequent predictors of male-male competition, has shaped
diversification in the largest radiation of birds.
\end{abstract}

\maketitle\unmarkedfntext{\textsuperscript{†}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia \\
\textsuperscript{‡}School of Natural and Computational Sciences, Massey University (Albany Campus), Auckland, New Zealand \\
\textsuperscript{§}\textit{justin.g.cally@gmail.com}}

\newpage
\phantomsection

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Sexual selection is a ubiquitous evolutionary process whose effect on
phylogenetic diversification is much debated (Lande 1981, 1982;
West-Eberhard 1983; Seddon et al. 2008; Cooney et al. 2018; Tsuji and
Fukami 2020). Sexual selection can promote speciation because it
operates on traits that can create reproductive isolation when they
diverge between lineages, such as signals and preferences involved in
mate selection (Lande 1981, 1982; Safran et al. 2013), sperm-egg
interactions (Swanson and Vacquier 1998), or genital morphology (Sloan
and Simmons 2019). Sexual selection could also promote speciation or
prevent extinction by purging deleterious mutations (Whitlock and
Agrawal 2009), fixing beneficial ones (Whitlock 2000), and accelerating
adaptation in different environments (Lorch et al. 2003; Candolin and
Heuschele 2008; Cally et al. 2019). Conversely, sexual selection might
hinder speciation or make extinction more likely by favouring traits
that improve mating success but reduce population fitness (Kokko and
Jennions 2008; Rankin et al. 2011; Holman and Kokko 2013; Fromhage and
Jennions 2016). For example, species with costly sexual signals may be
less resilient to environmental change (Kokko and Brooks 2003).
Extinction risk may also be exacerbated by sexual selection causing
maladaptation (`gender load') in female traits that are genetically
correlated with sexually-selected male traits (Pischedda and Chippindale
2006; Bonduriansky and Chenoweth 2009; Harano et al. 2010; Pennell and
Morrow 2013; Berger et al. 2014).

Although numerous studies have examined the relationship between sexual
selection and speciation or extinction rates (Barraclough et al. 1995;
Morrow et al. 2003; Seddon et al. 2008, 2013; Kraaijeveld et al. 2011;
Huang and Rabosky 2014), the availability of more complete phenotypic,
ecological and phylogenetic data (Jetz et al. 2012), together with
significant advances in phylogenetic methods (Rabosky 2014; Harvey
Michael et al. 2017), present new opportunities to test whether and how
sexual selection drives diversification. Furthermore, the diversity of
outcomes and approaches in previous studies suggests that the
association between species diversity and sexual selection is far from
clear (reviewed in Tsuji and Fukami (2020)).

A possible reason for the above uncertainty regarding the relationship
between sexual selection and diversification is that this relationship
may strongly depend on the environment. Theoretical work predicts that
sexual selection should have a more positive effect on adaptation and
population fitness in variable environments relative to stable ones
(Long et al. 2012; Connallon and Hall 2016). In stable environments,
consistent selection depletes genetic variation at sexually concordant
loci (i.e.~loci where the same allele is fittest for both sexes). In
these environments, genetic variation remains disproportionately at
sexually antagonistic loci, leading to stronger gender load and reduced
net benefits of sexual selection (Connallon and Hall 2016). By contrast,
in spatially or temporally variable environments, sexual selection can
enhance local adaptation. For example, in Darwin's finches, divergent
beak morphology is an adaptation to local food availability that has
been maintained through assortative mating (Huber et al. 2007). Under
these circumstances we predict that the effect of sexual selection on
rates of divergence may depend on the variability of the species'
environment. Despite the potential interaction between sexual selection
and environmental variability in diversification, phylogenetic tests are
currently lacking.

Birds have been a popular focus of macroevolutionary studies of sexual
selection and diversification (Barraclough et al. 1995; Morrow et al.
2003; Seddon et al. 2008, 2013; Huang and Rabosky 2014) because their
biology and phylogenetic relationships are comparatively well-known. A
2011 meta-analysis covering 20 primary studies of birds and other taxa
found a small but significant positive association between sexual
selection and speciation, with the average effect size in birds stronger
than in mammals but weaker than in insects or fish (Kraaijeveld et al.
2011). However, there was large variation in effect size estimates
across the 20 studies, likely reflecting differences in methodology,
such as metrics used to characterise speciation and sexual selection, in
addition to true biological differences. More recently, Huang and
Rabosky (2014) found no association between sexual dichromatism and
speciation (\emph{n} = 918 species) in a study using spectrophotometric
measurements of museum specimens (Armenta et al. 2008) and tip-rate
estimates from a molecular-only phylogeny (Jetz et al. 2012). Similarly,
Cooney et al. (2017) found no effect of sexual dichromatism on
diversification across 1,306 pairs of species, using dichromatism scores
provided by human observers. More recently, social polygyny (a proxy for
sexual selection) was found to have a positive association with
speciation rate across 954 species of birds (Iglesias-Carrasco et al.
2019). We summarize the major findings from previous studies testing the
association between sexual selection and speciation in birds and other
taxa since Kraaijeveld et al. (2011) meta-analysis (Table 1).

Here, we investigate the association between sexual selection and
diversification in birds while building upon previous approaches in
multiple ways. We use two measures of the strength of sexual selection:
sexual dichromatism (Dale et al. 2015), as well as an index of
male-biased sexual selection (Dale et al. 2015), which captures
(co)variation in sexual size dimorphism, social polygyny and paternal
care. We use these two measures because sexual dichromatism does not
always signal the presence of strong sexual selection and \emph{vice
versa} (Dale et al. 2015). For example, male and female dunnocks
(\emph{Prunella modularis}) are similarly coloured yet sexual selection
appears to be strong (Davies and Houston 1986). Furthermore, a recent
comparative study found a negative relationship between dichromatism and
another sexually-selected trait (song) across species, suggesting that a
multi-trait focus would improve estimates of sexual selection intensity
(Cooney et al. 2018). Additionally, our analysis includes multiple
ecological and environmental variables, allowing us to control for
potential confounds, to identify environmental factors, including
spatial and temporal environmental variability, interact with sexual
selection as theory predicts (Connallon and Hall 2016).

We use multiple approaches for quantifying speciation and extinction
rates at the tips of phylogenetic trees, including Bayesian Analysis of
Macroevolutionary Mixtures (BAMM; Beaulieu and O'Meara 2015; Rabosky
2016; Moore et al. 2016; Rabosky et al. 2017), as well as older but
reliable tip-rate statistics, namely diversification rate
(\(\lambda_{DR}\)) and node density (\(\lambda_{ND}\)) (Jetz et al.
2012). Our results show that (\emph{i}) a composite measure of sexual
selection, but not sexual dichromatism, significantly predicts
speciation rates, (\emph{ii}) the significant association between the
composite measure of sexual selection and speciation rate is largely
driven by sexual size dimorphism, (\emph{iii}) species with smaller
ranges have higher speciation rates and (\emph{iv}) there is no evidence
that environmental variables or their interaction with sexual selection
have an impact on diversification rates. Therefore, we provide evidence
at a very large scale that sexual selection can have positive effects on
diversification in the largest radiation of birds. Furthermore, we
suggest that the use of sexual dichromatism as the sole proxy for sexual
selection should be reconsidered, since it appears to be inconsistently
associated with the operation of sexual selection.

\begin{table}[H]

\caption{\label{tab:lit}Previous studies testing the association between sexual selection and speciation}
\centering
\resizebox{\linewidth}{!}{
\fontsize{8}{10}\selectfont
\begin{threeparttable}
\begin{tabular}{ll>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{3em}>{\raggedright\arraybackslash}p{27em}}
\toprule
Study & Taxa studied & Proxy for sexual selection & Support? & Outcome\\
\midrule
 &  & Plumage dichromatism & Yes & Across all birds, evidence in 4/6 studies\\
\cmidrule{3-5}
 &  & Mating system & Yes & Across all birds, evidence in 4/4 studies\\
\cmidrule{3-5}
\multirow{-3}{*}{\raggedright\arraybackslash Kraaijeveld et al. (2011)} & \multirow{-3}{*}{\raggedright\arraybackslash Meta-analysis across all animals} & Size dimorphism & Mixed & Across all birds, evidence in 1/2 studies\\
\cmidrule{1-5}
Maia et al. (2013) & Starlings (Sturnidae), 113 species & Ornamental innovations & Yes & Lineages with derived melanosomes (an ornamental innovation) diversify faster\\
\cmidrule{1-5}
Huang \& Rabosky (2014) & Across birds, \textasciitilde{}1000 species & Plumage dichromatism & No & No association between different measures of dichromatism and diversification\\
\cmidrule{1-5}
Gomes et al. (2016) & Estrildid finches, 134 species & Colour ornamentation & No & More ornamented lineages do not speciate more (but ornaments do evolve faster)\\
\cmidrule{1-5}
Cooney et al. (2017) & Across birds, 1306 pairs of species & Plumage dichromatism & No & Plumage dichromatism does not predict diversification rates, but might reduce the rate of fusion of lineages after secondary contact\\
\cmidrule{1-5}
Janicke et al. (2018) & Meta-analysis across all animals & Bateman gradient & Yes & Steepness of Bateman gradient in males predicts species richness\\
\cmidrule{1-5}
Mason et al. (2017) & Thraupids and Furnariids, 581 species & Vocal evolution & Yes & Bursts of speciation and song evolution are coincident\\
\cmidrule{1-5}
Iglesias-Carrasco et al. (2019) & Across birds, 954 species & Degree of polygyny & Yes & A higher degree of polygyny and rapid molecular evolution are linked with rate of diversification\\
\cmidrule{1-5}
Hosner et al. (2020) & Gallopheasants, 22 species & Sexual dimorphisn (range of traits) & No & No role of sexual selection in relation to diversification\\
\cmidrule{1-5}
Price-Waldman et al. (2020) & Thraupidae, 355 species & Plumage complexity & Yes & Elevated rates of plumage complexity evolution are associated with higher speciation rates\\
\cmidrule{1-5}
 &  & Size dimorphism & Yes & Sexual size dimorphism predicts two out of three measures of speciation rates\\
\cmidrule{3-5}
\multirow{-2}{*}{\raggedright\arraybackslash This study} & \multirow{-2}{*}{\raggedright\arraybackslash Across passerines, 5812 species} & Plumage dichromatism & No & There was no link between plumage dichromatism (measured from spectral info or RGB values) and any speciation rate\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item Studies were obtained by searching 'Web of Science' for articles published from 2011 for terms containing ‘speciation’, ‘diversification’ and ‘sexual selection’. We summarised all the studies we found relevant and comparable to our study.
\end{tablenotes}
\end{threeparttable}}
\end{table}

\phantomsection

\hypertarget{materials-and-methods}{%
\section{Materials and methods}\label{materials-and-methods}}

We examined the effect of sexual selection on speciation and extinction
rate in 97\% of passerines (\emph{n} = 5,812 species; 58\% of all
birds). Specifically, we (\emph{i}) compiled datasets for sexual
dichromatism/selection strength and environmental variability,
(\emph{ii}) obtained estimates of speciation and extinction rates across
passerines, and (\emph{iii}) conducted phylogenetic generalized
least-squares (PGLS) regressions. Analyses are documented with
reproducible code in the
\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}.
\phantomsection

\hypertarget{compiling-data-for-sexual-selection-and-environmental-stress}{%
\subsection{Compiling data for sexual selection and environmental
stress}\label{compiling-data-for-sexual-selection-and-environmental-stress}}

\phantomsection

\hypertarget{sexual-dichromatism}{%
\subsubsection{Sexual dichromatism}\label{sexual-dichromatism}}

We used a previously-published measure of sexual dichromatism for 5,983
species of passerines (Dale et al. 2015). Briefly, Dale et al. (Dale et
al. 2015) obtained sex-specific RGB (red-green-blue) values across six
body patches (nape, crown, forehead, throat, upper breast, and lower
breast) from \emph{Handbook of the Birds of the World} (Del Hoyo et al.
2011). The relative contribution of male and female RGB colour values
were averaged across body patches and provide `male-like' and
`female-like' plumage scores. Here we use the absolute difference
between male and female plumage scores as an estimate of sexual
dichromatism. Technically, this measures differences in the `degree of
male-ness' between males and females, rather than sex differences in
colour \emph{per se} (i.e.~dichromatism in the strict sense). For
example, the metric would fail to capture dichromatism when both the
male and female possess a single, but differently coloured `male-like'
patch. However, the metric is highly correlated with dichromatism
measured from spectral data (see below).

Additionally, we used another measure of dichromatism corresponding to
colour distance in avian colour space derived from spectral data
(Armenta et al. 2008). These measurements include variation in the
ultraviolet and bird-visible range and --- unlike the RGB measures ---
are sourced from museum specimens as opposed to illustrations. The
spectrophotometry data covers only 581 passerine species (10-fold fewer
than the RGB data), although there was a substantial correlation between
the two dichromatism measures (\(r = 0.79\);
\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
\phantomsection

\hypertarget{male-biased-sexual-selection}{%
\subsubsection{Male-biased sexual
selection}\label{male-biased-sexual-selection}}

Sexual dichromatism is likely to be imperfectly correlated with
variation in the strength of sexual selection across taxa. For this
reason, we sourced an additional measure of sexual selection (Dale et
al. 2015), referred to here as the `index of male-biased sexual
selection'. This index is the first principal component from a
phylogenetic principal component analysis (PPCA) of three
characteristics possitively associated with sexual selection (sexual
size dimorphism, social polygyny and {[}lack of{]} paternal care). The
variables included in this index have all been positively linked to the
intensity of sexual selection, and are usually correlated (Björklund
1990; Owens and Hartley 1998; Dunn et al. 2001), which is why they were
combined into a single metric in previous studies (Dale et al. 2015).
This measure of male-biased sexual selection is available for only 2,465
species, and shows a moderate correlation with the RGB measure of sexual
dichromatism (\(r = 0.34\);
\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
\phantomsection

\hypertarget{environmental-variables}{%
\subsubsection{Environmental variables}\label{environmental-variables}}

We obtained estimates of species range size using expert range maps
(BirdLife International and Handbook of the Birds of the World 2017).
The names of 1,230 species in the Birdlife database (Hoyo and Collar
2016) have been recently changed, so we manually matched these taxa with
the names used in the sexual dichromatism dataset (Hoyo and Collar
2016). For each species' range, we obtained estimates of climatic
conditions by extracting 1,000 random point samples of each bioclimatic
variable. We extracted 19 present-day bioclimatic variables
(representing a variety of biologically relevant annual trends in
temperature and precipitation) with 30-second (\textasciitilde{}1
km\textsuperscript{2}) spatial resolution (Fick and Hijmans 2017). From
the 1000 values of each bioclimatic variable, we obtained means and
standard deviations for each species. Using the same spatial sampling,
we extracted means and standard deviations of bioclimatic variables from
the paleoclimate during the last interglacial (LIG; 120,000 - 140,000
years ago) (Otto-Bliesner et al. 2006). To estimate variability in the
energy available to species, we obtained the mean and standard deviation
of net primary productivity (NPP) values between 2000 - 2015 across each
species distribution. Estimates of NPP had 30-second resolution and were
obtained through MODIS (Moderate Resolution Imaging Spectroradiometer)
primary production products stage 3 (MOD17A3) (Zhao et al. 2005). We
provide these data as a potentially useful data resource (see
\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}).
\phantomsection

\hypertarget{generating-biologically-relevant-predictors-for-environmental-stress}{%
\subsubsection{Generating biologically relevant predictors for
environmental
stress}\label{generating-biologically-relevant-predictors-for-environmental-stress}}

Given that stressful environments are expected to interact with sexual
selection and have a positive effect on adaptation (Cally et al. 2019),
we used the extracted environmental variables from each species range
size to create predictors of environmental variation/stress. We used
(\emph{i}) the average NPP in each species' range and (\emph{ii}) the
log-transformed range size as potentially informative predictors of
speciation rates. We also used three environmental predictors derived
from bioclimatic data. These predictors relate to seasonal climate
variation, spatial climate variation and long-term climate variation. To
obtain seasonal climate variation we used (\emph{iii}) mean values of
temperature seasonality (BIO4) for each range. (\emph{iv}) To estimate
levels of spatial environmental variation a species may endure, we used
the first principle component (PC1) from a PCA on standard deviations
from all bioclimatic variables, excluding temperature and precipitation
seasonality (BIO4 and BIO15). PC1 was heavily loaded towards bioclimatic
variables relating to temperature, thus PC1 largely reflects the
variation in temperature across a species' range
(\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
A taxon's range size often correlates with speciation and extinction
rates (Rosenzweig 1995; Castiglione et al. 2017), so we controlled for
the correlation between environmental spatial variation and species'
range sizes --- where larger ranges have larger variation in PC1 --- by
using the residuals of a fitted general additive model (GAM;
\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1})
as a predictor. To obtain long-term variation in climates for each
species range, we took (\emph{v}) the first principal component of the
absolute difference in the bioclimatic variables between the LIG and
current values. Similarly to spatial variation, the long-term climate
variation is primarily loaded to temperature differences between the LIG
and current climates
(\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
The five predictors of environmental variability are not strongly
correlated
(\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
Details and \texttt{R} code to generate these predictors can be found
within the
\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}.
\phantomsection

\hypertarget{estimating-extinction-and-speciation}{%
\subsection{Estimating extinction and
speciation}\label{estimating-extinction-and-speciation}}

Phylogenetic information was obtained from www.birdtree.org (Jetz et al.
2012). We used a maximum clade credibility (MCC) tree from 2,500 samples
of the posterior distribution (\emph{n} = 5,965) as the main
phylogenetic hypothesis in our comparative analysis. Additionally, a
random draw of full trees (including species without genetic data) from
the posterior distribution of phylogenetic trees was used for
diversification analyses using tip-rate measures (1,000 trees) and BAMM
(100 trees) (Rabosky 2014). These trees had crown clades with a topology
that was heavily constrained on the basis of a previously published
study (``Hacket backbone''; Hackett et al. 2008) and were constructed
using a pure birth (Yule) model. We calculated three different tip-rate
metrics of speciation and one of extinction across all trees.

Diversification is the result of two processes, speciation and
extinction through time. To estimate speciation rates, we first obtained
two tip-rate metrics of speciation using statistics derived from the
properties of the nodes and branches along root-to-tip paths of the
phylogeny. Node density (ND) is a simple statistic calculating the
density of nodes from the phylogenetic root to the tip, while the
log-transformed equal splits (logES; also referred to as diversification
rate/DR) is derived from the sum of edge lengths from each tip towards
the root, with each edge towards the root having the length
down-weighted (Jetz et al. 2012; Quintero and Jetz 2018; Rabosky et al.
2018). Crucially, studies have suggested that DR and ND (henceforth
referred to as \(\lambda_{DR}\) and \(\lambda_{ND}\)) are more
reflective estimates of speciation than diversification. Because
\(\lambda_{DR}\) and \(\lambda_{ND}\) cannot account for whole-clade
extinctions and thus under-estimate extinction rate, which makes the
composite measure of diversification more dependent on speciation
(Belmaker and Jetz 2015; Title and Rabosky 2018). Therefore,
\(\lambda_{DR}\) is a measure of speciation rate more heavily weighted
to recent speciation events while \(\lambda_{ND}\) measures speciation
across the root-to-tip path. These tip-rate measures are alternatives to
state-dependent diversification models such as Quantitative State
Speciation-Extinction (QuaSSE); but, based on previous simulation
studies, \(\lambda_{DR}\) and \(\lambda_{ND}\) are robust and intuitive
measures that provide high power and low false discovery rate with large
phylogenies when incorporated into Phylogenetic Generalized Least
Squares (PGLS) models (Harvey Michael et al. 2017).

We used BAMM to model the dynamics of speciation and extinction across
the 101 phylogenetic trees (one MCC tree and 100 random draws of the
posterior). This software uses a Bayesian approach (reversible-jump
Markov Chain Monte Carlo) to generate probability distributions of
evolutionary rate-shift configurations with variable speciation and
extinction rates (Rabosky 2014). These models provide tip-rate estimates
of speciation and extinction rate that can be easily used in comparative
analyses. The parameters of the 100 BAMM runs are detailed in full in
the
\href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{Supplementary Information};
briefly, we used a time-variable model with the prior expected number of
evolutionary rate shifts set at 100 and prior rates set from the initial
tip-level estimates of speciation and extinction using the
\texttt{BAMMtools} R package (Rabosky et al. 2014). BAMM models were run
independently for the 101 phylogenetic trees for 100 million
generations. Given the computationally intensive nature of BAMM, runs
were conducted across multiple CPUs. Important BAMM parameters
(log-likelihood and number of rate shifts) reached convergence with
effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples
surpassing 200; an arbitrary value, above which posterior distributions
can often be accurately inferred
(\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3},
\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
Further details of BAMM parameters and output are available in the
\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information},
with tip-rate means and variances provided. Additionally, given the
variability in BAMM estimates, we also provide analysis of BAMM shift
configurations and tip-rate estimates from our run on the MCC tree and
within a BAMM run on the MCC tree from a genetic-only phylogeny across
all birds (Harvey et al. 2017). All analyses were conducted on
log-rates. \phantomsection

\hypertarget{phylogenetic-comparative-analysis}{%
\subsection{Phylogenetic comparative
analysis}\label{phylogenetic-comparative-analysis}}

To test the association between speciation/extinction and sexual
selection, environmental variability and their interaction, we used
phylogenetic least squares (PGLS) models in the \texttt{nlme} package
(Pinheiro et al. 2018). Firstly, we conducted model selection to compare
models in which \(\lambda_{DR}\), \(\lambda_{ND}\), \(\lambda_{BAMM}\)
or \(\mu_{BAMM}\) were the response variable: these tip-rate estimates
all came from the same MCC tree (derived from 2,500 draws of the
posterior distribution (Jetz et al. 2012)). For models of
\(\lambda_{BAMM}\) and \(\mu_{BAMM}\), we used the inverse of the
variance associated with each tip rate estimate as weights, to account
for the variable precision of the estimates provided by BAMM. For each
response variable, we conducted model selection to compare models with
different combinations of predictor variables. The most complex model in
each set under comparison contained one of the measures of sexual
selection (sexual dichromatism or the index of male-biased sexual
selection), all of the environmental measures (i.e.~log-transformed
range size, seasonal temperature variation, spatial temperature
variation, long-term temperature variation, and NPP), and all of the
2-way interactions between sexual selection and each of the
environmental measures. The simpler models contained all of the same
main effects, but had fewer 2-way interaction terms (potentially none).
Model selection was done in \texttt{MuMIn} using the \texttt{dredge}
function (Bartoń 2017). Using the terms from the top-ranked model
(ranked by AICc), we ran the equivalent model for each of the 1,000
phylogenetic trees used to derive \(\lambda_{DR}\), \(\lambda_{ND}\) and
each of the 100 trees used to derive \(\lambda_{BAMM}\) and
\(\mu_{BAMM}\). Additionally, we investigated the effect of the
individual variables used to derive the index of male-biased sexual
selection on speciation rate. For these pgls models we replaced the
composite index score with the individual biological variable (sexual
size dimorphism, social polygyny and {[}lack of{]} paternal care) and
ran the equivalent model for 300 phylogenetic trees used to derive
\(\lambda_{DR}\), \(\lambda_{ND}\) and 100 trees used to derive
\(\lambda_{BAMM}\).

Across all our analyses we corrected for the phylogenetic signal. Our
models used the unique response variables and correlation structure for
a given phylogenetic tree. Specifically, for models using tip-rate
metrics (\(\lambda_{DR}\), \(\lambda_{ND}\)), we estimated the
phylogenetic signal independently for each of the 1,000 trees/models.
Phylogenetic signal was estimated as Pagel's \(\lambda\) (Pagel 1999)
using the \texttt{corPagel} function in the \texttt{ape} package
(Paradis et al. 2004). Alternatively, for models using speciation and
extinction estimates derived using BAMM (\(\lambda_{BAMM}\) and
\(\mu_{BAMM}\)), we found that \(\lambda\) was consistently estimated at
1 and hence assumed Brownian motion (using the \texttt{corBrownian}
function) to estimate the correlation structure. This method enabled us
to present model estimates for an MCC tree alongside 1,000/100 trees
from the posterior distribution of trees to account for phylogenetic
uncertainty. This approach was repeated on three datasets corresponding
to each measure of sexual selection: dichromatism derived from RGB
values of images (\emph{n} = 5,812); dichromatism from spectrophotometry
(\emph{n} = 581) and the index of male-biased sexual selection (\emph{n}
= 2,465).

Finally, using the subset of species with an index of male-biased sexual
selection, we conducted a phylogenetic path analysis using the
\texttt{phylopath\ R} package (Bijl 2018). The phylogenetic path
analysis was used to assess causal paths between variables unable to be
modelled within the univariate response of PGLS. That is, a phylogenetic
path analysis allowed us to model relationships between the predictor
variables used in our PGLS analysis as we anticipate environmental
variability, sexual dichromatism/selection, and range size to have
effects on each other and not just on speciation rate. To minimise path
complexity, we used temperature seasonality (BIO4) as the single measure
for environmental variability, \(\lambda_{DR}\) as the single measure of
speciation, and the tip-rates from the MCC tree. Further details of the
path analysis, including our rationale for each path's directions, can
be found within the
\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}
along with all other analyses and the relevant \texttt{R} code to
reproduce results. \phantomsection

\hypertarget{results}{%
\section{Results}\label{results}}

\phantomsection

\hypertarget{male-biased-sexual-selection-but-not-sexual-dichromatism-affects-speciation}{%
\subsection{Male-biased sexual selection, but not sexual dichromatism,
affects
speciation}\label{male-biased-sexual-selection-but-not-sexual-dichromatism-affects-speciation}}

We examined the effect of sexual selection on speciation and extinction
rate in 97\% of passerines (\emph{n} = 5,812 species; 58\% of all birds;
\autoref{Speciation_Figure}). We found a significant positive
association between the index of male-biased sexual selection (\emph{n}
= 2,465) and \(\lambda_{DR}\) from the maximum clade credibility (MCC)
tree (\(\beta\) = \(3.89 \times 10^{-2}\), p = 0.01;
\autoref{model_results}b). However, this association was not significant
for the other two measures of speciation rate (\(\lambda_{ND}\):
\(\beta\) = \(4.38 \times 10^{-4}\), p = 0.35; \(\lambda_{BAMM}\):
\(\beta\) = \(9.42 \times 10^{-4}\), p = 0.76;
\autoref{model_results}b). When we took into account phylogenetic
uncertainty by running the models using 1,000 trees, the distribution of
estimates from PGLS models was similar to the estimate from the MCC
tree: among the 1,000 trees there was a positive association between
sexual selection and \(\lambda_{DR}\) (highest posterior density (HPD)
Interval = \(4.51 \times 10^{-3}\), \(5.72 \times 10^{-2}\)), and the
distribution skewed towards a positive association between sexual
selection and \(\lambda_{ND}\) (HPD Interval = \(-5.04 \times 10^{-4}\),
\(1.58 \times 10^{-3}\)) as well as the 100 models using
\(\lambda_{BAMM}\) (HPD Interval = \(-1.30 \times 10^{-2}\),
\(3.09 \times 10^{-2}\);
\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).

We investigated which of the three variables comprising the index of
male-biased sexual selection was driving the association observed with
\(\lambda_{DR}\). Our results over 300 trees showed that this pattern is
mainly driven by the sexual size dimorphism component (HPD Interval =
\(8.53 \times 10^{-1}\), 3.11), with the effects of other components
overlapping zero; paternal care (HPD Interval =
\(-1.78 \times 10^{-1}\), \(7.90 \times 10^{-3}\)) and mating system
(HPD Interval= \(-7.35 \times 10^{-2}\), \(4.32 \times 10^{-2}\)).
Importantly, the association between sexual size dimorphism and
speciation rates is also present when using \(\lambda_{ND}\) (HPD
Interval = \(1.80 \times 10^{-1}\), \(6.38 \times 10^{-1}\)), but not
when using \(\lambda_{BAMM}\) (HPD Interval = -1.49,
\(7.45 \times 10^{-1}\), \autoref{SS_components}).

In contrast to male-biased sexual selection, we found no evidence that
species with increased sexual dichromatism have higher or lower rates of
speciation. Sexual dichromatism showed no association with
\(\lambda_{DR}\) (\(\beta\) = \(-1.28 \times 10^{-3}\), p = 0.15;
\autoref{model_results}a, \autoref{Speciation_Figure}), \(\lambda_{ND}\)
(\(\beta\) = \(-5.75 \times 10^{-5}\), p = 0.08;
\autoref{model_results}a) or \(\lambda_{BAMM}\) (\(\beta\) =
\(-1.43 \times 10^{-5}\), p = 0.87; \autoref{model_results}a). PGLS
analyses using sexual dichromatism (\emph{n} = 581) measured by
spectrophotometry (Armenta et al. 2008) yielded results concordant with
the full dataset; i.e.~no association between sexual dichromatism and
speciation
(\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
Our results from models based on the MCC tree are largely corroborated
by model estimates from PGLS analyses of the rates and correlation
structures from 1,000 trees (for \(\lambda_{DR}\), \(\lambda_{ND}\)) and
100 trees for \(\lambda_{BAMM}\). The HPD intervals show model estimates
are distributed around zero when using complete taxon sampling models
and RGB measures of sexual dichromatism (\(\lambda_{DR}\): HPD Interval
= \(-1.63 \times 10^{-3}\), \(1.66 \times 10^{-3}\), \(\lambda_{ND}\):
HPD Interval = \(-4.26 \times 10^{-5}\), \(5.50 \times 10^{-5}\),
\autoref{model_results}a,
\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
For PGLS models using spectrophotometry-based measures of sexual
dichromatism, the estimates from the 100 trees in the \(\lambda_{DR}\)
models are positively skewed (HPD Interval = \(-1.78 \times 10^{-2}\),
\(3.49 \times 10^{-2}\)) but normally distributed around zero for
\(\lambda_{ND}\) and \(\lambda_{BAMM}\)
(\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).

Our analyses also show that the differences in results between sexual
dichromatism and male-biased sexual selection (i.e.~association with
speciation rates only for the latter) were not due to differences in the
size of the datasets used (5,812 species vs.~2,465,
\href{https://justincally.github.io/SexualSelection_Speciation/#testing_the_sexual_dichromatism_dataset_using_the_same_data_points_as_the_sexual_selection_dataset}{Figure S17}).
No interaction terms were present in the top models (\(\Delta\) AICc
\textgreater{} 4) for any measure of speciation (\(\lambda_{DR}\),
\(\lambda_{ND}\), \(\lambda_{BAMM}\)) or sexual selection (RGB values,
spectrophotometry and the index of male-biased sexual selection;
\(\Delta\) AICc \textgreater{} 4;
\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5},
\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6},
\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11},
\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}).
Thus we found no evidence that the effect of sexual selection on
speciation is dependent on our measures of environmental variation or
range size. Furthermore, we found no evidence that these environmental
factors --- seasonal temperature variation, long-term temperature
variation, spatial temperature variation, and Net Primary Productivity
(NPP) --- predict speciation independently from sexual
dichromatism/selection (\autoref{model_results},
\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
\phantomsection

\hypertarget{species-with-smaller-ranges-have-increased-rates-of-speciation}{%
\subsection{Species with smaller ranges have increased rates of
speciation}\label{species-with-smaller-ranges-have-increased-rates-of-speciation}}

Based on \(\lambda_{DR}\) and \(\lambda_{ND}\) tip-rate metrics of
speciation, we found a negative association between range size and
speciation; that is, species with smaller ranges show marginally higher
values for \(\lambda_{DR}\) and \(\lambda_{ND}\). This negative
association was small but significant for models using the MCC tree
(\(\lambda_{DR}\): \(\beta\) = \(-6.58 \times 10^{-3}\), p =
\(1.48 \times 10^{-3}\); \(\lambda_{ND}\): \(\beta\) =
\(-1.46 \times 10^{-4}\), p = 0.03; \autoref{model_results}a,
\autoref{Speciation_Figure}). This association was also evident across
the estimates from models using the 1,000 trees (\(\lambda_{DR}\): HPD
Interval = \(-8.87 \times 10^{-3}\), \(-6.61 \times 10^{-4}\);
\(\lambda_{ND}\): HPD Interval = \(-1.51 \times 10^{-4}\),
\(1.72 \times 10^{-5}\); \autoref{model_results}a). Subset models with
reduced sample size and different measures of sexual selection --- but
the same measure of range size --- showed equivocal evidence that range
size is negatively associated with speciation. Range size is
significantly associated with \(\lambda_{DR}\)
(\autoref{model_results}b) using data subset for species with an index
of male-biased sexual selection (\emph{n} = 2,465) but not
\(\lambda_{ND}\) or \(\lambda_{BAMM}\). Models using data subset for
spectrophotometry-based dichromatism (\emph{n} = 581) gave
non-significant estimates for the effect of range size on all measures
of speciation
(\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11},
\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12},
\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S13}).
Because the range size dataset is the same across the three data
subsets, we draw our conclusions from the models with the highest power
using near-complete taxon sampling (\emph{n} = 5,812). \phantomsection

\hypertarget{phylogenetic-path-analysis}{%
\subsection{Phylogenetic path
analysis}\label{phylogenetic-path-analysis}}

Using a phylogenetic path analysis, we found multiple significant paths
between variables used in the PGLS (\autoref{Path_Analysis};
\href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S14}).
There was a modest effect of male-biased sexual selection on sexual
dimorphism (\(\beta\) = 0.22). Additionally, temperature seasonality
weakly affected sexual dimorphism (\(\beta\) = 0.07) and strongly
affected range size (\(\beta\) = 0.52). This suggests an indirect effect
of temperature seasonality on \(\lambda_{DR}\) (\(\beta_{indirect}\) =
-0.02; \autoref{Path_Analysis}), given the negative association we
identified between \(\lambda_{DR}\) and range size in PGLS models.
\phantomsection

\hypertarget{extinction-rate}{%
\subsection{Extinction rate}\label{extinction-rate}}

We found no evidence that extinction (\(\mu_{BAMM}\)) was impacted by
the extent of sexual dichromatism for full-taxon sampling (\(\beta\) =
\(2.38 \times 10^{-5}\), p = 0.93; \autoref{model_results}a), nor
spectrophotometry-based measures of sexual dichromatism
(\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11},
\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12},
\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S13})
or male-biased sexual selection (\autoref{model_results}b,
\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15},
\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S16}).
\phantomsection

\hypertarget{variability-across-phylogenetic-trees-and-speciation-rate-measures}{%
\subsection{Variability across phylogenetic trees and speciation rate
measures}\label{variability-across-phylogenetic-trees-and-speciation-rate-measures}}

Estimates of the effect of predictor variables on speciation rates
varied across phylogenetic trees, especially in the BAMM rates
(\(\lambda_{BAMM}\) and \(\mu_{BAMM}\)), where the 95 \% HPD interval
across PGLS model estimates from 100 trees was often \textgreater{} 20
times larger than the 95 \% confidence interval for estimates from a
single PGLS model using the MCC tree. This contrasts with variation
across trees for the other rate estimates (\(\lambda_{DR}\) and
\(\lambda_{ND}\)), where the 95 \% HPD interval of model estimates for
pgls models using 1,000 trees was near-equivelant to the 95 \%
confidence interval calculated for PGLS model estimates of the MCC tree
(\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S9}).
The great majority of earlier studies have based their estimates on a
single consensus tree due to the computational requirements of BAMM.
However, our results suggest that BAMM estimates between alternative,
similarly plausible phylogenies vary substantially. Mean measures of
speciation rate across 100 trees were positively correlated between
measures (\(\lambda_{DR}\) - \(\lambda_{BAMM}\): \emph{r}=0.75,
\(\lambda_{DR}\) - \(\lambda_{ND}\): \emph{r}=0.65, \(\lambda_{ND}\) -
\(\lambda_{BAMM}\): \emph{r}=0.51;
\href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
The calculation of BAMM rates can be affected by the settings of the run
and the use of different priors. We therefore compared the estimate of
our MCC tree with that of previously published analyses on birds and
found a high correlation (\emph{r}=0.81,
\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8},
Harvey et al. (2017)). Full details of the BAMM results are presented as
supplementary materials.

\begin{figure}
\centering
\includegraphics{Figures/DRSpeciationFigure.pdf}
\caption{Speciation rate (\(\lambda_{DR}\)) across all passerine birds
(\emph{n} = 5,965) with estimates of sexual dichromatism, range size
available for 5,812 species and an index of male-biased sexual selection
available for 2,465 species. Across these species there was a small but
significant negative association between \(\lambda_{DR}\) and log-range
size as well as a significant positive association between
\(\lambda_{DR}\) and male-biased sexual selection but no significant
association between \(\lambda_{DR}\) and sexual dichromatism based on
RGB measures. \(\lambda_{DR}\) are those from the MCC tree and images of
birds are from the \emph{Handbook of the Birds of the World}. Clockwise
the six species are: \emph{Sporophila bouvronides}, \emph{Euplectes
franciscanus}, \emph{Phainopepla nitens}, \emph{Paradisaea rubra},
\emph{Malurus pulcherrimus}, \emph{Lepidothrix coeruleocapilla}. Edge
colours for the terminal branch correspond to \(\lambda_{DR}\) but all
precluding branches have been generated for graphical purposes using
ancestral character state estimation (Revell 2012) and should not be
interpreted. Illustrations reproduced by permission of Lynx Editions.
\label{Speciation_Figure}}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Estimates_and_scatter_edit.pdf}
\caption{Model estimates (a) showing the effect size (i.e.~slope) of
log-range size and sexual dichromatism on speciation and extinction
rates using PGLS analyses with the sexual dichromatism dataset (RGB
values, \emph{n} = 5,812). (b) depicts the scatter plot of speciation
rate (\(\lambda_{DR}\)) and log-range size with the model estimate
presented as a dashed line. (c) shows the scatter plot of speciation
rate (\(\lambda_{DR}\)) and male-biased sexual selection (\emph{n} =
2,465). Similar to (a), (d) presents model estimates for PGLS analyses
using a restricted dataset with measures of an index of male-biased
sexual selection (\emph{n} = 2,465) and shows the effect of log-range
size and male-biased sexual selection on speciation and extinction
rates. Both datasets were used for analyses with three measures of
speciation (\(\lambda_{DR}\), \(\lambda_{ND}\), \(\lambda_{BAMM}\)) and
one measure of extinction (\(\mu_{BAMM}\)) as response variables. The
numerical values for the model estimates using the MCC tree and HPD
intervals of estimates from 1,000 randomly sampled trees (for
\(\lambda_{DR}\) and \(\lambda_{ND}\)) or 100 randomly sampled trees for
\(\lambda_{BAMM}\) and \(\mu_{BAMM}\) can be found in the Supplementary
Information. Density curves are based on model estimates from 1,000/100
trees and the circle below with error bars is the estimate and 95\% CIs
from the MCC tree. For this figure we removed outliers from estimates
coming from the 100 randomly sampled trees for BAMM models in order to
be able to visualise the MCC 95\% CIs. \label{model_results}}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Individual_SS_Estimates.pdf}
\caption{Estimates of the effect of individual sexual selection
components included in the PPCA (paternal care, sexual size dimorphism
and mating system) on three measures of speciation rate
(\(\lambda_{DR}\), \(\lambda_{ND}\) and \(\lambda_{BAMM}\)). Estimates
are presented as density intervals from pgls models on 300 phylogentic
trees that used species with available data for these sexual selection
measures (\emph{n} = 2,465). The bar under each density ridge is the 95
\% Highest Posterior Density Interval. Given that the mating system is a
categorical variable, model estimates for three polygynous mating system
levels are in reference to a striclty monogamous mating system (0\%
polygyny). \label{SS_components}}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figures/Path_Analysis.pdf}
\caption{Path analysis of evolutionary and ecological variables. Arrows
represent direct effects with the direction of effect corresponding to
colours (blue = positive, red = negative). The numeric values are
standardised regression slopes and the asterisks indicates that the 95\%
confidence intervals of this estimate do not overlap with zero. The
confidence intervals were obtained from 500 bootstrapped iterations and
the data used in this analysis was subset to species with both sexual
dichromatism and an index of male-biased sexual selection measures
(\emph{n} = 2,465). \label{Path_Analysis}}
\end{figure}

\newpage\phantomsection

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

We found evidence that the composite index of male-biased sexual
selection, but not measures of sexual dichromatism, is correlated with
the rate of speciation in passerine birds. The absence of a detectable
correlation between sexual dichromatism and speciation rate was
consistent across different measures of speciation (\(\lambda_{DR}\),
\(\lambda_{ND}\) and \(\lambda_{BAMM}\)) and both measures of
dichromatism (spectral and RGB), and it cannot be explained by a
difference in statistical power or sampling. These findings reaffirm the
conclusions of previous, smaller studies in which sexual dichromatism
was measured using spectrophotometry (Huang and Rabosky 2014) or human
observers (Cooney et al. 2017). The correlation between speciation rate
and the index of male-biased sexual selection (which encapsulates
variation in sexual size dimorphism, social polygyny, and paternal care)
was statistically significant for \(\lambda_{DR}\), but not for
\(\lambda_{ND}\) and \(\lambda_{BAMM}\). This pattern seems to be mainly
driven by an association between sexual size dimorphism and speciation.
Interestingly, we also found a consistent negative relationship between
range size and speciation rate, at least when this rate was quantified
using \(\lambda_{DR}\) and \(\lambda_{ND}\). None of the bioclimatic
measures of environmental variability that we investigated (i.e.,
temperature seasonality, long-term temperature variation, and spatial
temperature variation) were significantly associated speciation rate,
nor mediated the relationship between sexual selection and
diversification.

The difference in findings between the analyses of sexual dichromatism
\emph{versus} the index of male-biased sexual selection is noteworthy,
because the majority of earlier studies used dichromatism alone as their
proxy for sexual selection (e.g., Barraclough et al. 1995; Owens et al.
1999; Morrow et al. 2003; Seddon et al. 2013; Huang and Rabosky 2014).
Given our findings, and the modest correlation between dichromatism and
the sexual selection index (\(r = 0.34\); Dale et al. 2015), we suggest
that sexual dichromatism may not be a robust proxy for sexual selection
(Cooney et al. 2018). Although dichromatism almost certainly provides
some insight into the operation of sexual selection, it may be too
indirect a measure to detect any association with speciation rate, even
with large sample size. There are several reasons why the use of sexual
dichromatism as a proxy for sexual selection is problematic. Sexual
dichromatism can evolve for reasons other than sexual selection, such as
when males and females occupy different ecological niches (Wallace 1889;
Kottler 1980; Slatkin 1984; Shine 1989) or experience different
selective pressures in contexts other than competition for mates (Price
and Eaton 2014). For example, in superb fairy-wrens (\emph{Malurus
cyaneus}) female colouration has probably evolved in response to spatial
variation in predation pressure, increasing dichromatism (Medina et al.
2017). In fact, our path analysis detected a weak relationship between
environment and sexual dichromatism, where sexual dichromatism was
positively predicted by temperature seasonality (a measure of
environmental variation).

In line with some theoretical predictions and previous studies
(Kraaijeveld et al. 2011), we found that male-biased sexual selection
increases speciation rate, at least when speciation is measured by
\(\lambda_{DR}\). Many of the species that have both high scores of
male-biased sexual selection and high diversification rates belong to
the genera \emph{Ploceus}, \emph{Euplectes} (Ploceidae) and
\emph{Paradisaea} (Paradiaseidae). Multiple weaver species (Ploceidae)
are polygynous and lack paternal care, and both weavers and birds of
paradise have strong size dimorphism. The association between speciation
rates and principal component scores that we report seems to be mainly
driven by sexual size dimorphism and, to a lesser extent, paternal care.
Speciation rates (both \(\lambda_{DR}\) and \(\lambda_{ND}\)) are higher
in species with larger sexual dimorphism and \(\lambda_{DR}\) also has a
tendency to be higher in species with no paternal care. Size dimorphism
is often thought to arise as a consequence of intrasexual competition,
where one of the sexes (males in most birds) has to compete for access
to the other sex, leading to selection for larger body sizes and thus
greater dimorphism (Björklund 1990; Owens and Hartley 1998). Therefore,
competition between males could be the underlying driver of the high
speciation rates that we detect in some clades.

Sexual dimorphism due to competition within sexes contrasts with the
drivers of sexual dichromatism. Plumage dichromatism can evolve as a
consequence of female cryptic choice and be related to extra-pair
fertilizations, but not necessarily paternal care or mating system
(Owens and Hartley 1998). It can also arise as a result of selection on
the level of crypsis of the sex that cares for offspring (Dale et al.
2015). The fact that traits linked with competition (such as size
dimorphism) are the ones associated with higher \(\lambda_{DR}\) values
-- rather than sexual dichromatism -- supports the general view that
antagonistic interactions and sexual conflict can lead to increased
diversity (Bonduriansky 2011; Qvarnström et al. 2012; Tinghitella et al.
2018; Tsuji and Fukami 2020). Moreover, body size is a trait that
influences multiple aspects of the physiology and ecology of a species.
Differences in body size (as a result of sexual selection) could be
linked to changes in diet, vulnerability to predators or environmental
tolerance (Damuth 1993; Liow et al. 2008; Bonduriansky 2011), and such
differences could ultimately increase the likelihood of divergence
between young lineages. In mammals, sexual selection is suggested to
have driven the evolution of large body size which in turn has allowed
diversification of ecological strategies in the clade, and higher
speciation rates (McLain 1993; Bonduriansky 2011).

We also found that the association between sexual selection and
speciation appears to be independent of net primary productivity and
spatiotemporal variation in the environment. The lack of an effect of
these environmental variables on speciation rate has several possible
interpretations. Firstly, the effects of sexual selection on adaptation
and speciation may depend on the type of environmental variability under
which the species is evolving. Specifically, speciation rates might be
impacted by genetic constraints on adaptation, that vary across
environments. Theory suggests that sexual antagonism (which is often
exacerbated in species with strong sexual selection) may be lower in
habitats experiencing cyclical environmental variation
(e.g.~seasonality), relative to those experiencing directional change in
the environment (Connallon and Hall 2016). Another possibility is that
the environmental predictors we chose may not account for the key
ecological sources of selection that interact with sexual selection to
drive speciation. For example, our study does not include direct measure
of food availability or the severity of predation and parasitism, which
are both hypothesised to affect sexual selection and speciation
(reviewed in Maan and Seehausen 2011). Finally, it is possible that
environmental variability genuinely has little effect on speciation
rates, at least in the taxa investigated here.

We found that species with smaller ranges have elevated speciation
rates. This result is similar to a study of 329 amphibian genera, which
found higher diversification rates in taxa with smaller range size
(Greenberg and Mooers 2017). Intuitively, large range size should
promote speciation by creating more opportunities for geographic
barriers to form (Rosenzweig 1995; Castiglione et al. 2017). However,
the opposite pattern is also plausible because birds with limited
dispersal or more specialised niches can have more fragmented
populations, which would promote vicariant divergence and higher
speciation rates (Jablonski and Roy 2003; Birand et al. 2012; Claramunt
et al. 2012). Moreover, species that have recently split as a
consequence of vicariant divergence might have smaller ranges as a
result of the split of the ancestral lineage, leading to a link between
smaller ranges and shorter divergence times. It is also possible that
high speciation rates cause smaller range sizes, rather than the other
way around, for example because repeatedly-speciating lineages tend to
fill niches in ways that hinder the geographical expansion of new
species (Rosenzweig 1995; Weir and Price 2011; Price and Eaton 2014).
However, species undergoing adaptive radiation in new habitats are
unlikely to be limited by competition for resources from existing taxa.
One further explanation for the negative association between range size
and sexual dichromatism/sexual selection is the potential bias of
taxonomic classification, whereby over-splitting of species in clades
with large ranges leads to increased recent phylogenetic branching as
well as smaller ranges.

In addition to speciation, sexual selection is hypothesised to affect
extinction. Using the model-based approach of BAMM, we found no
association between the estimated extinction rate and sexual
dichromatism, male-biased sexual selection, or our measures of
environmental variability. However, these extinction results should not
be regarded as definitive because extinction is notoriously difficult to
estimate accurately from phylogenies, principally because different
combinations of speciation and extinction rates can give rise to similar
patterns of diversity (see Rabosky 2016). Phylogenetic methods such as
BAMM allow for speciation and extinction rates to be estimated using
moderately-sized phylogenies, although the ability of BAMM to model
evolutionary rate shifts and extinction rates is debated (see, Beaulieu
and O'Meara 2015; Rabosky 2016; Moore et al. 2016; Rabosky et al. 2017).
Additionally, while several tip-rate estimates exist for speciation rate
(e.g., \(\lambda_{DR}\) and \(\lambda_{ND}\)), tip-rate estimates of
extinction rate are difficult to obtain without complex Bayesian models
which are sensitive to sampling bias (Davis et al. 2013). Although
extinction rates can be inferred from alternative sources, such as the
fossil record (Martins et al. 2018), direct observation extinction, or
IUCN red list status (Greenberg and Mooers 2017), each approach has its
limitations. Across the passerine bird phylogeny, we found that BAMM
often produced homogeneous speciation and extinction rates for smaller
clades showing few rate shifts, which might reduce our power to detect
small differences in extinction rates among closely-related taxa
(Rabosky et al. 2017; Title and Rabosky 2018). Thus, this methodological
constraint likely decreases our ability to accurately measure the
correlation between metrics of sexual selection and the probability of
extinction.

One outcome of our analyses was that different measures of speciation
rates presented different results. This is not completely surprising,
because each of the rates is calculated differently (Title and Rabosky
2018). For instance, \(\lambda_{DR}\) is weighted more towards
speciation events close to the tips and allows more rate heterogeneity
compared to \(\lambda_{BAMM}\) estimates. Rate shifts are unlikely to be
detected in smaller clades in BAMM, meaning that it is not uncommon for
whole genera to have the same rate. Using the \(\lambda_{DR}\) metric,
only sister species are guaranteed to have the same rate. This leads to
greater variation in \(\lambda_{DR}\) relative to the \(\lambda_{BAMM}\)
estimates, which is suggested to be an advantage when studying
diversification patterns (Quintero et al. 2015). Additionally,
\(\lambda_{BAMM}\) estimates were more sensitive to phylogenetic
uncertainty and were 20 times more variable across trees compared to
\(\lambda_{DR}\) estimates. We cannot completely reject the idea that
the lack of association between \(\lambda_{BAMM}\) and sexual selection
could be the result of low statistical power, due to the combination of
both low variation across species in the speciation rates and high
levels of variation in the estimates across trees.

To summarise, we have shown that sexual size dimorphism (a putative
proxy for male-biased sexual selection), but not sexual dichromatism,
predicts speciation in passerines, that the magnitude of this effect is
modest, and that this relationship is not markedly affected by
environmental variability. We have also shown that there is no evidence
of an association between sexual selection and extinction rates.
Overall, our findings imply that male-male competition could be the
mechanism driving increased speciation rates in birds, that sexual
dichromatism may not be a reliable proxy for sexual selection, and that
alternative measures of sexual selection are more directly related to
diversification.

\phantomsection

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We thank Adnan Moussalli for his assistance in conducting BAMM runs
across 100 phylogenetic trees, Ignacio Quintero for providing BAMM
results for comparison and for discussion of the results as well as
Chris Cooney for comments on the manuscript.

\phantomsection

\hypertarget{author-contributions}{%
\section{Author Contributions}\label{author-contributions}}

All authors contributed to conception, design and approach to analyses;
J.G.C. and J.D. compiled the data; J.G.C. and I.M. conducted analyses;
J.G.C. wrote the first draft of the manuscript; all authors contributed
substantially to further manuscript revisions.

\phantomsection

\hypertarget{data-accessibility}{%
\section{Data Accessibility}\label{data-accessibility}}

We have sought to make the data and Supplementary Material (including
\texttt{R} code) freely accessible to readers. The code and
Supplementary Material for this manuscript are available on
\href{https://justincally.github.io/SexualSelection_Speciation/}{GitHub}
and as an attached HTML document (Supplementary Material) for reviewers.
Data is archived in the adjoined GitHub repository and the environmental
dataset collated for this manuscript is available on Dryad:
\url{https://doi.org/10.5061/dryad.573n5tb6n}.

\newpage\phantomsection

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Armenta_2008}{}%
Armenta, J. K., P. O. Dunn, and L. A. Whittingham. 2008. Quantifying
avian sexual dichromatism: A comparison of methods. Journal of
Experimental Biology 211:2423.

\leavevmode\hypertarget{ref-Barraclough_1995}{}%
Barraclough, T. G., P. H. Harvey, and S. Nee. 1995. Sexual selection and
taxonomic diversity in passerine birds. Proceedings of the Royal Society
B-Biological Sciences 259:211--215.

\leavevmode\hypertarget{ref-MuMIn_2017}{}%
Bartoń, K. 2017. MuMIn: Multi-model inference.

\leavevmode\hypertarget{ref-Beaulieu_2015}{}%
Beaulieu, J. M., and B. C. O'Meara. 2015. Extinction can be estimated
from moderately sized molecular phylogenies. Evolution 69:1036--1043.

\leavevmode\hypertarget{ref-Belmaker_2015}{}%
Belmaker, J., and W. Jetz. 2015. Relative roles of ecological and
energetic constraints, diversification rates and region history on
global species richness gradients. Ecology Letters 18:563--571.

\leavevmode\hypertarget{ref-Berger_2014}{}%
Berger, D., K. Grieshop, M. I. Lind, J. Goenaga, A. A. Maklakov, and G.
Arnqvist. 2014. Intralocus sexual conflict and environmental stress.
Evolution 68:2184--2196.

\leavevmode\hypertarget{ref-Wouter_2018}{}%
Bijl, W. van der. 2018. \emph{Phylopath}: Easy phylogenetic path
analysis in R. PeerJ 6:e4718.

\leavevmode\hypertarget{ref-Birand_2012}{}%
Birand, A., A. Vose, and S. Gavrilets. 2012. Patterns of species ranges,
speciation, and extinction. American Naturalist 179:1--21.

\leavevmode\hypertarget{ref-birdlife_2017}{}%
BirdLife International and Handbook of the Birds of the World. 2017.
Bird species distribution maps of the world.
\url{http://datazone.birdlife.org/species/requestdis}.

\leavevmode\hypertarget{ref-bjorklund_1990}{}%
Björklund, M. 1990. A phylogenetic interpretation of sexual dimorphism
in body size and ornament in relation to mating system in birds. Journal
of Evolutionary Biology 3:171--183. Wiley Online Library.

\leavevmode\hypertarget{ref-Bonduriansky_2011}{}%
Bonduriansky, R. 2011. Sexual selection and conflict as engines of
ecological diversification. American Naturalist 178:729--745.

\leavevmode\hypertarget{ref-Bonduriansky_Chenoweth_2009}{}%
Bonduriansky, R., and S. F. Chenoweth. 2009. Intralocus sexual conflict.
Trends in Ecology \& Evolution 24:280--8.

\leavevmode\hypertarget{ref-Cally_2019}{}%
Cally, J. G., D. Stuart-Fox, and L. Holman. 2019. Meta-analytic evidence
that sexual selection improves population fitness. Nature communications
10:2017. Nature Publishing Group.

\leavevmode\hypertarget{ref-Candolin_2008}{}%
Candolin, U., and J. Heuschele. 2008. Is sexual selection beneficial
during adaptation to environmental change? Trends in Ecology \&
Evolution 23:446--452.

\leavevmode\hypertarget{ref-Castiglione_2017}{}%
Castiglione, S., A. Mondanaro, M. Melchionna, C. Serio, M. Di Febbraro,
F. Carotenuto, and P. Raia. 2017. Diversification rates and the
evolution of species range size frequency distribution. Frontiers in
Ecology and Evolution 5:147. Frontiers.

\leavevmode\hypertarget{ref-Claramunt_2012}{}%
Claramunt, S., E. P. Derryberry, J. V. Remsen, and R. T. Brumfield.
2012. High dispersal ability inhibits speciation in a continental
radiation of passerine birds. Proceedings of the Royal Society B:
Biological Sciences 279:1567.

\leavevmode\hypertarget{ref-Connallon_Hall_2016}{}%
Connallon, T., and M. D. Hall. 2016. Genetic correlations and
sex-specific adaptation in changing environments. 70:2198.

\leavevmode\hypertarget{ref-Cooney_2018}{}%
Cooney, C. R., H. E. A. MacGregor, N. Seddon, and J. A. Tobias. 2018.
Multi-modal signal evolution in birds: Re-examining a standard proxy for
sexual selection. Proceedings of the Royal Society of London B:
Biological Sciences 285.

\leavevmode\hypertarget{ref-Cooney_2017}{}%
Cooney, C. R., J. A. Tobias, J. T. Weir, C. A. Botero, and N. Seddon.
2017. Sexual selection, speciation and constraints on geographical range
overlap in birds. Ecology Letters 20:863--871.

\leavevmode\hypertarget{ref-Dale_2015}{}%
Dale, J., C. J. Dey, K. Delhey, B. Kempenaers, and M. Valcu. 2015. The
effects of life history and sexual selection on male and female plumage
colouration. Nature 527:367--370.

\leavevmode\hypertarget{ref-Damuth_1993}{}%
Damuth, J. 1993. Cope's rule, the island rule and the scaling of
mammalian population density. Nature 365:748--750. Nature Publishing
Group.

\leavevmode\hypertarget{ref-davies1986re}{}%
Davies, N., and A. Houston. 1986. Reproductive success of dunnocks,
\emph{Prunella modularis}, in a variable mating system. II. Conflicts of
interest among breeding adults. Journal of Animal Ecology 55:139--154.

\leavevmode\hypertarget{ref-Davis_2013}{}%
Davis, M. P., P. E. Midford, and W. Maddison. 2013. Exploring power and
parameter estimation of the BiSSE method for analyzing species
diversification. BMC Evolutionary Biology 13:38.

\leavevmode\hypertarget{ref-delhoyo2011}{}%
Del Hoyo, J., A. Elliott, and D. Christie. 2011. Handbook of the birds
of the world. Lynx Edicions 2003-2011.

\leavevmode\hypertarget{ref-dunn_2001}{}%
Dunn, P. O., L. A. Whittingham, and T. E. Pitcher. 2001. Mating systems,
sperm competition, and the evolution of sexual dimorphism in birds.
Evolution 55:161--175. Wiley Online Library.

\leavevmode\hypertarget{ref-Fick_2017}{}%
Fick, S. E., and R. J. Hijmans. 2017. WorldClim 2: New 1-km spatial
resolution climate surfaces for global land areas. International Journal
of Climatology 37:4302--4315.

\leavevmode\hypertarget{ref-Fromhage_2016}{}%
Fromhage, L., and M. D. Jennions. 2016. Coevolution of parental
investment and sexually selected traits drives sex-role divergence.
Nature Communications 7:12517.

\leavevmode\hypertarget{ref-Gomes_2016}{}%
Gomes, A. C. R., M. D. Sorenson, and G. C. Cardoso. 2016. Speciation is
associated with changing ornamentation rather than stronger sexual
selection. Evolution 70:2823--2838.

\leavevmode\hypertarget{ref-Greenberg_2017}{}%
Greenberg, D. A., and A. Ø. Mooers. 2017. Linking speciation to
extinction: Diversification raises contemporary extinction risk in
amphibians. Evolution Letters 1:40--48.

\leavevmode\hypertarget{ref-Hackett_2008}{}%
Hackett, S. J., R. T. Kimball, S. Reddy, R. C. K. Bowie, E. L. Braun, M.
J. Braun, J. L. Chojnowski, W. A. Cox, K.-L. Han, J. Harshman, C. J.
Huddleston, B. D. Marks, K. J. Miglia, W. S. Moore, F. H. Sheldon, D. W.
Steadman, C. C. Witt, and T. Yuri. 2008. A phylogenomic study of birds
reveals their evolutionary history. Science 320:1763--1768.

\leavevmode\hypertarget{ref-harano2010intralocus}{}%
Harano, T., K. Okada, S. Nakayama, T. Miyatake, and D. J. Hosken. 2010.
Intralocus sexual conflict unresolved by sex-limited trait expression.
Current Biology 20:2036--2039.

\leavevmode\hypertarget{ref-Harvey_2017}{}%
Harvey, M. G., G. F. Seeholzer, B. T. Smith, D. L. Rabosky, A. M.
Cuervo, and R. T. Brumfield. 2017. Positive association between
population genetic differentiation and speciation rates in new world
birds. Proceedings of the National Academy of Sciences 114:6328--6333.

\leavevmode\hypertarget{ref-Harvey_essim_2017}{}%
Harvey Michael, G., L. Rabosky Daniel, and N. Cooper. 2017. Continuous
traits and speciation rates: Alternatives to state‐dependent
diversification models. Methods in Ecology and Evolution 9:984--993.

\leavevmode\hypertarget{ref-Holman_Kokko_2013}{}%
Holman, L., and H. Kokko. 2013. The consequences of polyandry for
population viability, extinction risk and conservation. Philosophical
Transactions of the Royal Society B-Biological Sciences 368.

\leavevmode\hypertarget{ref-Hosner_2020}{}%
Hosner, P. A., H. L. Owens, E. L. Braun, and R. T. Kimball. 2020.
Phylogeny and diversification of the gallopheasants (aves: Galliformes):
Testing roles of sexual selection and environmental niche divergence.
Zoologica Scripta 49:549--562.

\leavevmode\hypertarget{ref-delhoyo_2016}{}%
Hoyo, J. del, and N. J. Collar. 2016. HBW and birdlife international
illustrated checklist of the birds of the world. Lynx Edicions; BirdLife
International.

\leavevmode\hypertarget{ref-Huang_2014}{}%
Huang, H. T., and D. L. Rabosky. 2014. Sexual selection and
diversification: Reexamining the correlation between dichromatism and
speciation rate in birds. American Naturalist 184:E101--E114.

\leavevmode\hypertarget{ref-Huber_2007}{}%
Huber, S. K., L. F. De Leon, A. P. Hendry, E. Bermingham, and J. Podos.
2007. Reproductive isolation of sympatric morphs in a population of
darwin's finches. Proceedings of the Royal Society of London B:
Biological Sciences 274:1709--1714.

\leavevmode\hypertarget{ref-Iglesias-Carrasco_2019}{}%
Iglesias-Carrasco, M., M. D. Jennions, S. Y. W. Ho, and D. A. Duchêne.
2019. Sexual selection, body mass and molecular evolution interact to
predict diversification in birds. Proceedings of the Royal Society B:
Biological Sciences 286:20190172.

\leavevmode\hypertarget{ref-Jablonski_2003}{}%
Jablonski, D., and K. Roy. 2003. Geographical range and speciation in
fossil and living molluscs. Proceedings of the Royal Society of London.
Series B: Biological Sciences 270:401--406. The Royal Society.

\leavevmode\hypertarget{ref-Janicke_2018}{}%
Janicke, T., M. G. Ritchie, E. H. Morrow, and L. Marie-Orleach. 2018.
Sexual selection predicts species richness across the animal kingdom.
Proceedings of the Royal Society B: Biological Sciences 285:20180173.
The Royal Society.

\leavevmode\hypertarget{ref-Jetz_2012}{}%
Jetz, W., G. H. Thomas, J. B. Joy, K. Hartmann, and A. O. Mooers. 2012.
The global diversity of birds in space and time. Nature 491:444--448.

\leavevmode\hypertarget{ref-Kokko_Brooks_2003}{}%
Kokko, H., and R. Brooks. 2003. Sexy to die for? Sexual selection and
the risk of extinction. Annales Zoologici Fennici 40:207--219.

\leavevmode\hypertarget{ref-Kokko_Jennions2008}{}%
Kokko, H., and M. D. Jennions. 2008. Parental investment, sexual
selection and sex ratios. Journal of Evolutionary Biology 21:919--948.

\leavevmode\hypertarget{ref-Kottler_1980}{}%
Kottler, M. J. 1980. Darwin, Wallace, and the origin of sexual
dimorphism. Proceedings of the American Philosophical Society
124:203--226.

\leavevmode\hypertarget{ref-Kraaijeveld_2011}{}%
Kraaijeveld, K., F. J. L. Kraaijeveld-Smit, and M. E. Maan. 2011. Sexual
selection and speciation: The comparative evidence revisited. Biological
Reviews 86:367--377.

\leavevmode\hypertarget{ref-Lande_1981}{}%
Lande, R. 1981. Models of speciation by sexual selection on polygenic
traits. Proceedings of the National Academy of Sciences 78:3721--3725.

\leavevmode\hypertarget{ref-Lande_1982}{}%
Lande, R. 1982. Rapid origin of sexual isolation and character
divergence in a cline. Evolution 36:213--223.

\leavevmode\hypertarget{ref-Liow_2008}{}%
Liow, L. H., M. Fortelius, E. Bingham, K. Lintulaakso, H. Mannila, L.
Flynn, and N. C. Stenseth. 2008. Higher origination and extinction rates
in larger mammals. Proceedings of the National Academy of Sciences
105:6097--6102. National Acad Sciences.

\leavevmode\hypertarget{ref-Long_2012}{}%
Long, T. A. F., A. F. Agrawal, and L. Rowe. 2012. The effect of sexual
selection on offspring fitness depends on the nature of genetic
variation. Current Biology 22:204--208.

\leavevmode\hypertarget{ref-Lorch_2003}{}%
Lorch, P. D., S. Proulx, L. Rowe, and T. Day. 2003. Condition-dependent
sexual selection can accelerate adaptation. Evolutionary Ecology
Research 5:867--881.

\leavevmode\hypertarget{ref-Maan_2011}{}%
Maan, M. E., and O. Seehausen. 2011. Ecology, sexual selection and
speciation. Ecology Letters 14:591--602.

\leavevmode\hypertarget{ref-Maia_2013}{}%
Maia, R., D. R. Rubenstein, and M. D. Shawkey. 2013. Key ornamental
innovations facilitate diversification in an avian radiation.
Proceedings of the National Academy of Sciences 110:10687--10692.
National Acad Sciences.

\leavevmode\hypertarget{ref-Martins_2018}{}%
Martins, M. J. F., T. M. Puckett, R. Lockwood, J. P. Swaddle, and G.
Hunt. 2018. High male sexual investment as a driver of extinction in
fossil ostracods. Nature 556:366.

\leavevmode\hypertarget{ref-Mason_2017}{}%
Mason, N. A., K. J. Burns, J. A. Tobias, S. Claramunt, N. Seddon, and E.
P. Derryberry. 2017. Song evolution, speciation, and vocal learning in
passerine birds. Evolution 71:786--796. Wiley Online Library.

\leavevmode\hypertarget{ref-Mclain_1993}{}%
McLain, D. K. 1993. Cope's rules, sexual selection, and the loss of
ecological plasticity. Oikos 490--500. JSTOR.

\leavevmode\hypertarget{ref-Medina_2017}{}%
Medina, I., K. Delhey, A. Peters, K. E. Cain, M. L. Hall, R. A. Mulder,
and N. E. Langmore. 2017. Habitat structure is linked to the evolution
of plumage colour in female, but not male, fairy-wrens. BMC evolutionary
biology 17:35.

\leavevmode\hypertarget{ref-Moore_2016}{}%
Moore, B. R., S. Hohna, M. R. May, B. Rannala, and J. P. Huelsenbeck.
2016. Critically evaluating the theory and performance of bayesian
analysis of macroevolutionary mixtures. Proceedings of the National
Academy of Sciences 113:9569--9574.

\leavevmode\hypertarget{ref-Morrow_Speciation_2003}{}%
Morrow, E. H., T. E. Pitcher, and G. Arnqvist. 2003. No evidence that
sexual selection is an 'engine of speciation' in birds. Ecology Letters
6:228--234.

\leavevmode\hypertarget{ref-Otto-Bliesner_2006}{}%
Otto-Bliesner, B. L., S. J. Marshall, J. T. Overpeck, G. H. Miller, A.
Hu, and. 2006. Simulating arctic climate warmth and icefield retreat in
the last interglaciation. Science 311:1751--1753.

\leavevmode\hypertarget{ref-Owens_1999}{}%
Owens, I. P. F., P. M. Bennett, and P. H. Harvey. 1999. Species richness
among birds: Body size, life history, sexual selection or ecology?
Proceedings of the Royal Society B-Biological Sciences 266:933--939.

\leavevmode\hypertarget{ref-owens_1998}{}%
Owens, I. P., and I. R. Hartley. 1998. Sexual dimorphism in birds: Why
are there so many different forms of dimorphism? Proceedings of the
Royal Society of London. Series B: Biological Sciences 265:397--407. The
Royal Society.

\leavevmode\hypertarget{ref-Pagel_1999}{}%
Pagel, M. 1999. Inferring the historical patterns of biological
evolution. Nature 401:877.

\leavevmode\hypertarget{ref-ape_2004}{}%
Paradis, E., J. Claude, and K. Strimmer. 2004. APE: Analyses of
phylogenetics and evolution in R language. Bioinformatics 20:289--290.

\leavevmode\hypertarget{ref-Pennell_2013}{}%
Pennell, T. M., and E. H. Morrow. 2013. Two sexes, one genome: The
evolutionary dynamics of intralocus sexual conflict. Ecology and
Evolution 3:1819--1834.

\leavevmode\hypertarget{ref-nlme_2018}{}%
Pinheiro, J., D. Bates, S. DebRoy, D. Sarkar, and R Core Team. 2018.
nlme: Linear and nonlinear mixed effects models.

\leavevmode\hypertarget{ref-Pischedda_2006}{}%
Pischedda, A., and A. K. Chippindale. 2006. Intralocus sexual conflict
diminishes the benefits of sexual selection. PLOS Biology 4:e356.

\leavevmode\hypertarget{ref-Price_2014}{}%
Price, J. J., and M. D. Eaton. 2014. Reconstructing the evolution of
sexual dichromatism: Current color diversity does not reflect past rates
of male and female change. Evolution 68:2026--2037.

\leavevmode\hypertarget{ref-Price-Waldman_2020}{}%
Price-Waldman, R. M., A. J. Shultz, and K. J. Burns. 2020. Speciation
rates are correlated with changes in plumage color complexity in the
largest family of songbirds. Evolution 74:1155--1169.

\leavevmode\hypertarget{ref-quintero_2018}{}%
Quintero, I., and W. Jetz. 2018. Global elevational diversity and
diversification of birds. Nature 555:246.

\leavevmode\hypertarget{ref-Quintero_2015}{}%
Quintero, I., P. Keil, W. Jetz, and F. W. Crawford. 2015. Historical
biogeography using species geographical ranges. Systematic Biology
64:1059--1073.

\leavevmode\hypertarget{ref-Qvarnstrom_2012}{}%
Qvarnström, A., N. Vallin, and A. Rudh. 2012. The role of male contest
competition over mates in speciation. Current Zoology 58:493--509.
Oxford University Press Oxford, Uk.

\leavevmode\hypertarget{ref-bammtools_2014}{}%
Rabosky, D., M. Grundler, C. Anderson, P. Title, J. Shi, J. Brown, H.
Huang, and J. Larson. 2014. BAMMtools: An R package for the analysis of
evolutionary dynamics on phylogenetic trees. Methods in Ecology and
Evolution 5:701--707.

\leavevmode\hypertarget{ref-Rabosky_2014}{}%
Rabosky, D. L. 2014. Automatic detection of key innovations, rate
shifts, and diversity-dependence on phylogenetic trees. PLOS ONE
9:e89543.

\leavevmode\hypertarget{ref-Rabosky_Extinction_2016}{}%
Rabosky, D. L. 2016. Challenges in the estimation of extinction from
molecular phylogenies: A response to Beaulieu and O'Meara. Evolution
70:218--228.

\leavevmode\hypertarget{ref-Rabosky_fish_2018}{}%
Rabosky, D. L., J. Chang, P. O. Title, P. F. Cowman, L. Sallan, M.
Friedman, K. Kaschner, C. Garilao, T. J. Near, M. Coll, and others.
2018. An inverse latitudinal gradient in speciation rate for marine
fishes. Nature 559:392.

\leavevmode\hypertarget{ref-Rabosky_BAMM_flawed_2017}{}%
Rabosky, D. L., J. S. Mitchell, and J. Chang. 2017. Is BAMM flawed?
Theoretical and practical concerns in the analysis of multi-rate
diversification models. Systematic biology 66:477--498.

\leavevmode\hypertarget{ref-Rankin_2011}{}%
Rankin, D. J., U. Dieckmann, and H. Kokko. 2011. Sexual conflict and the
tragedy of the commons. American Naturalist 177:780--791.

\leavevmode\hypertarget{ref-phytools_2012}{}%
Revell, L. J. 2012. Phytools: An r package for phylogenetic comparative
biology (and other things). Methods in Ecology and Evolution 3:217--223.

\leavevmode\hypertarget{ref-Rosenzweig_1995}{}%
Rosenzweig, M. L. 1995. Species diversity in space and time. Cambridge
University Press.

\leavevmode\hypertarget{ref-Safran_2013}{}%
Safran, R. J., E. S. Scordato, L. B. Symes, R. L. Rodrı'guez, and T. C.
Mendelson. 2013. Contributions of natural and sexual selection to the
evolution of premating reproductive isolation: A research agenda. Trends
in Ecology \& Evolution 28:643--650.

\leavevmode\hypertarget{ref-Seddon_2013}{}%
Seddon, N., C. A. Botero, J. A. Tobias, P. O. Dunn, H. E. A. MacGregor,
D. R. Rubenstein, J. A. C. Uy, J. T. Weir, L. A. Whittingham, and R. J.
Safran. 2013. Sexual selection accelerates signal evolution during
speciation in birds. Proceedings of the Royal Society B: Biological
Sciences 280.

\leavevmode\hypertarget{ref-Seddon_2008}{}%
Seddon, N., R. M. Merrill, and J. A. Tobias. 2008. Sexually selected
traits predict patterns of species richness in a diverse clade of
suboscine birds. American Naturalist 171:620--631.

\leavevmode\hypertarget{ref-Shine_1989}{}%
Shine, R. 1989. Ecological causes for the evolution of sexual
dimorphism: A review of the evidence. The Quarterly Review of Biology
64:419--461.

\leavevmode\hypertarget{ref-Slatkin_1984}{}%
Slatkin, M. 1984. Ecological causes of sexual dimorphism. Evolution
38:622--630.

\leavevmode\hypertarget{ref-sloan2019genitalia}{}%
Sloan, N. S., and L. W. Simmons. 2019. The evolution of female
genitalia. Journal of Evolutionary Biology in press.

\leavevmode\hypertarget{ref-swanson1998abalone}{}%
Swanson, W. J., and V. D. Vacquier. 1998. Concerted evolution in an egg
receptor for a rapidly evolving abalone sperm protein. Science
281:710--712.

\leavevmode\hypertarget{ref-Tinghitella_2018}{}%
Tinghitella, R. M., A. C. Lackey, M. Martin, P. D. Dijkstra, J. P.
Drury, R. Heathcote, J. Keagy, E. S. Scordato, and A. M. Tyers. 2018. On
the role of male competition in speciation: A review and research
agenda. Behavioral Ecology 29:783--797. Oxford University Press UK.

\leavevmode\hypertarget{ref-Rabosky_diversification_2018}{}%
Title, P. O., and D. L. Rabosky. 2018. Diversification rates and
phylogenies: What are we estimating, and how good are the estimates?
bioRxiv 369124.

\leavevmode\hypertarget{ref-tsuji2020sexual}{}%
Tsuji, K., and T. Fukami. 2020. Sexual dimorphism and species diversity:
From clades to sites. Trends in Ecology \& Evolution 35:105--114.
Elsevier.

\leavevmode\hypertarget{ref-Wallace_1889}{}%
Wallace, A. R. 1889. Colours and ornaments characteristic of sex.
\emph{in} Darwinism, an exponent of the theory of natural selection,
with some of its applications. Macmillan; Company.

\leavevmode\hypertarget{ref-Weir_2011}{}%
Weir, J. T., and T. D. Price. 2011. Limits to speciation inferred from
times to secondary sympatry and ages of hybridizing species along a
latitudinal gradient. American Naturalist 177:462--469.

\leavevmode\hypertarget{ref-west1983sexual}{}%
West-Eberhard, M. J. 1983. Sexual selection, social competition, and
speciation. Quarterly Review of Biology 58:155--183.

\leavevmode\hypertarget{ref-Whitlock_2000}{}%
Whitlock, M. C. 2000. Fixation of new alleles and the extinction of
small populations: Drift load, beneficial alleles, and sexual selection.
Evolution 54:1855--1861.

\leavevmode\hypertarget{ref-Whitlock_Agrawal_2009}{}%
Whitlock, M. C., and A. F. Agrawal. 2009. Purging the genome with sexual
selection: Reducing mutation load through selection on males. Evolution
63:569--582.

\leavevmode\hypertarget{ref-Zhao_2005}{}%
Zhao, M., F. A. Heinsch, R. R. Nemani, and S. W. Running. 2005.
Improvements of the modis terrestrial gross and net primary production
global data set. Remote Sensing of Environment 95:164--176.

\end{document}

### 2021-01-18 title_page.tex
add tex

--- 2021-01-18 Manuscript_Publication.tex
@@
 
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
+\usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath} \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
 
-\usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath} \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling} \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large} \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}} \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption} \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt} \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip} \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape} \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{} \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{} \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang, flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false, linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}} \raggedbottom
+\pretitle{\begin{flushleft}\LARGE}
+\posttitle{\par\end{flushleft}\vskip 0.5em}
+\preauthor{\begin{flushleft}\large \lineskip 0.5em}
+\postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
+\postdate{\par\end{flushleft}}
+
+\usepackage{fancyhdr}
+
+\pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}}
+\rhead[]{Sexual selection predicts speciation}
+
+\usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{setspace}
+
+\onehalfspacing
+
+\usepackage[font={small}]{caption} \usepackage{epigraph}
+
+\setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
+\setlength\beforeepigraphskip{1\baselineskip}
+\setlength\afterepigraphskip{2\baselineskip}
+
+\renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape} \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{} \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{} \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang, flushmargin]{footmisc} \usepackage{hyperref}
+
+\hypersetup{colorlinks=true, hyperfootnotes=false, linkcolor=blue, filecolor=blue, urlcolor=blue}
+
+\usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
+
+\raggedbottom
 
@@
 \maketitle
-\begin{abstract}
-Sexual selection is thought to shape phylogenetic diversity by affecting
-speciation or extinction rates. However, the net effect of sexual
-selection on diversification is hard to predict, because many of the
-hypothesised effects on speciation or extinction have opposing signs and
-uncertain magnitudes. Theoretical work also suggests that the net effect
-of sexual selection on diversification should depend strongly on
-ecological factors, though this prediction has seldom been tested. Here,
-we test whether variation in sexual selection can predict speciation and
-extinction rates across passerine birds (up to 5,812 species, covering
-most genera) and whether this relationship is mediated by environmental
-factors. Male-biased sexual selection, and specifically sexual size
-dimorphism, predicted two of the three measures of speciation rates that
-we examined. The link we observed between sexual selection and
-speciation was independent of environmental variability, though species
-with smaller ranges had higher speciation rates. There was no
-association between any proxies of sexual selection and extinction rate.
-Our findings support the view that male-biased sexual selection, as
-measured by frequent predictors of male-male competition, has shaped
-diversification in the largest radiation of birds.
-\end{abstract}
 
@@
 \textsuperscript{§}\textit{justin.g.cally@gmail.com}}
-
-\newpage
-\phantomsection
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Sexual selection is a ubiquitous evolutionary process whose effect on
-phylogenetic diversification is much debated (Lande 1981, 1982;
-West-Eberhard 1983; Seddon et al. 2008; Cooney et al. 2018; Tsuji and
-Fukami 2020). Sexual selection can promote speciation because it
-operates on traits that can create reproductive isolation when they
-diverge between lineages, such as signals and preferences involved in
-mate selection (Lande 1981, 1982; Safran et al. 2013), sperm-egg
-interactions (Swanson and Vacquier 1998), or genital morphology (Sloan
-and Simmons 2019). Sexual selection could also promote speciation or
-prevent extinction by purging deleterious mutations (Whitlock and
-Agrawal 2009), fixing beneficial ones (Whitlock 2000), and accelerating
-adaptation in different environments (Lorch et al. 2003; Candolin and
-Heuschele 2008; Cally et al. 2019). Conversely, sexual selection might
-hinder speciation or make extinction more likely by favouring traits
-that improve mating success but reduce population fitness (Kokko and
-Jennions 2008; Rankin et al. 2011; Holman and Kokko 2013; Fromhage and
-Jennions 2016). For example, species with costly sexual signals may be
-less resilient to environmental change (Kokko and Brooks 2003).
-Extinction risk may also be exacerbated by sexual selection causing
-maladaptation (`gender load') in female traits that are genetically
-correlated with sexually-selected male traits (Pischedda and Chippindale
-2006; Bonduriansky and Chenoweth 2009; Harano et al. 2010; Pennell and
-Morrow 2013; Berger et al. 2014).
-
-Although numerous studies have examined the relationship between sexual
-selection and speciation or extinction rates (Barraclough et al. 1995;
-Morrow et al. 2003; Seddon et al. 2008, 2013; Kraaijeveld et al. 2011;
-Huang and Rabosky 2014), the availability of more complete phenotypic,
-ecological and phylogenetic data (Jetz et al. 2012), together with
-significant advances in phylogenetic methods (Rabosky 2014; Harvey
-Michael et al. 2017), present new opportunities to test whether and how
-sexual selection drives diversification. Furthermore, the diversity of
-outcomes and approaches in previous studies suggests that the
-association between species diversity and sexual selection is far from
-clear (reviewed in Tsuji and Fukami (2020)).
-
-A possible reason for the above uncertainty regarding the relationship
-between sexual selection and diversification is that this relationship
-may strongly depend on the environment. Theoretical work predicts that
-sexual selection should have a more positive effect on adaptation and
-population fitness in variable environments relative to stable ones
-(Long et al. 2012; Connallon and Hall 2016). In stable environments,
-consistent selection depletes genetic variation at sexually concordant
-loci (i.e.~loci where the same allele is fittest for both sexes). In
-these environments, genetic variation remains disproportionately at
-sexually antagonistic loci, leading to stronger gender load and reduced
-net benefits of sexual selection (Connallon and Hall 2016). By contrast,
-in spatially or temporally variable environments, sexual selection can
-enhance local adaptation. For example, in Darwin's finches, divergent
-beak morphology is an adaptation to local food availability that has
-been maintained through assortative mating (Huber et al. 2007). Under
-these circumstances we predict that the effect of sexual selection on
-rates of divergence may depend on the variability of the species'
-environment. Despite the potential interaction between sexual selection
-and environmental variability in diversification, phylogenetic tests are
-currently lacking.
-
-Birds have been a popular focus of macroevolutionary studies of sexual
-selection and diversification (Barraclough et al. 1995; Morrow et al.
-2003; Seddon et al. 2008, 2013; Huang and Rabosky 2014) because their
-biology and phylogenetic relationships are comparatively well-known. A
-2011 meta-analysis covering 20 primary studies of birds and other taxa
-found a small but significant positive association between sexual
-selection and speciation, with the average effect size in birds stronger
-than in mammals but weaker than in insects or fish (Kraaijeveld et al.
-2011). However, there was large variation in effect size estimates
-across the 20 studies, likely reflecting differences in methodology,
-such as metrics used to characterise speciation and sexual selection, in
-addition to true biological differences. More recently, Huang and
-Rabosky (2014) found no association between sexual dichromatism and
-speciation (\emph{n} = 918 species) in a study using spectrophotometric
-measurements of museum specimens (Armenta et al. 2008) and tip-rate
-estimates from a molecular-only phylogeny (Jetz et al. 2012). Similarly,
-Cooney et al. (2017) found no effect of sexual dichromatism on
-diversification across 1,306 pairs of species, using dichromatism scores
-provided by human observers. More recently, social polygyny (a proxy for
-sexual selection) was found to have a positive association with
-speciation rate across 954 species of birds (Iglesias-Carrasco et al.
-2019). We summarize the major findings from previous studies testing the
-association between sexual selection and speciation in birds and other
-taxa since Kraaijeveld et al. (2011) meta-analysis (Table 1).
-
-Here, we investigate the association between sexual selection and
-diversification in birds while building upon previous approaches in
-multiple ways. We use two measures of the strength of sexual selection:
-sexual dichromatism (Dale et al. 2015), as well as an index of
-male-biased sexual selection (Dale et al. 2015), which captures
-(co)variation in sexual size dimorphism, social polygyny and paternal
-care. We use these two measures because sexual dichromatism does not
-always signal the presence of strong sexual selection and \emph{vice
-versa} (Dale et al. 2015). For example, male and female dunnocks
-(\emph{Prunella modularis}) are similarly coloured yet sexual selection
-appears to be strong (Davies and Houston 1986). Furthermore, a recent
-comparative study found a negative relationship between dichromatism and
-another sexually-selected trait (song) across species, suggesting that a
-multi-trait focus would improve estimates of sexual selection intensity
-(Cooney et al. 2018). Additionally, our analysis includes multiple
-ecological and environmental variables, allowing us to control for
-potential confounds, to identify environmental factors, including
-spatial and temporal environmental variability, interact with sexual
-selection as theory predicts (Connallon and Hall 2016).
-
-We use multiple approaches for quantifying speciation and extinction
-rates at the tips of phylogenetic trees, including Bayesian Analysis of
-Macroevolutionary Mixtures (BAMM; Beaulieu and O'Meara 2015; Rabosky
-2016; Moore et al. 2016; Rabosky et al. 2017), as well as older but
-reliable tip-rate statistics, namely diversification rate
-(\(\lambda_{DR}\)) and node density (\(\lambda_{ND}\)) (Jetz et al.
-2012). Our results show that (\emph{i}) a composite measure of sexual
-selection, but not sexual dichromatism, significantly predicts
-speciation rates, (\emph{ii}) the significant association between the
-composite measure of sexual selection and speciation rate is largely
-driven by sexual size dimorphism, (\emph{iii}) species with smaller
-ranges have higher speciation rates and (\emph{iv}) there is no evidence
-that environmental variables or their interaction with sexual selection
-have an impact on diversification rates. Therefore, we provide evidence
-at a very large scale that sexual selection can have positive effects on
-diversification in the largest radiation of birds. Furthermore, we
-suggest that the use of sexual dichromatism as the sole proxy for sexual
-selection should be reconsidered, since it appears to be inconsistently
-associated with the operation of sexual selection.
-
-\begin{table}[H]
-
-\caption{\label{tab:lit}Previous studies testing the association between sexual selection and speciation}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{8}{10}\selectfont
-\begin{threeparttable}
-\begin{tabular}{ll>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{3em}>{\raggedright\arraybackslash}p{27em}}
-\toprule
-Study & Taxa studied & Proxy for sexual selection & Support? & Outcome\\
-\midrule
- &  & Plumage dichromatism & Yes & Across all birds, evidence in 4/6 studies\\
-\cmidrule{3-5}
- &  & Mating system & Yes & Across all birds, evidence in 4/4 studies\\
-\cmidrule{3-5}
-\multirow{-3}{*}{\raggedright\arraybackslash Kraaijeveld et al.
-(2011)} & \multirow{-3}{*}{\raggedright\arraybackslash Meta-analysis across all animals} & Size dimorphism & Mixed & Across all birds, evidence in 1/2 studies\\
-\cmidrule{1-5}
-Maia et al. (2013) & Starlings (Sturnidae), 113 species & Ornamental innovations & Yes & Lineages with derived melanosomes (an ornamental innovation) diversify faster\\
-\cmidrule{1-5}
-Huang \& Rabosky (2014) & Across birds, \textasciitilde{}1000 species & Plumage dichromatism & No & No association between different measures of dichromatism and diversification\\
-\cmidrule{1-5}
-Gomes et al. (2016) & Estrildid finches, 134 species & Colour ornamentation & No & More ornamented lineages do not speciate more (but ornaments do evolve faster)\\
-\cmidrule{1-5}
-Cooney et al.
-(2017) & Across birds, 1306 pairs of species & Plumage dichromatism & No & Plumage dichromatism does not predict diversification rates, but might reduce the rate of fusion of lineages after secondary contact\\
-\cmidrule{1-5}
-Janicke et al. (2018) & Meta-analysis across all animals & Bateman gradient & Yes & Steepness of Bateman gradient in males predicts species richness\\
-\cmidrule{1-5}
-Mason et al. (2017) & Thraupids and Furnariids, 581 species & Vocal evolution & Yes & Bursts of speciation and song evolution are coincident\\
-\cmidrule{1-5}
-Iglesias-Carrasco et al. (2019) & Across birds, 954 species & Degree of polygyny & Yes & A higher degree of polygyny and rapid molecular evolution are linked with rate of diversification\\
-\cmidrule{1-5}
-Hosner et al. (2020) & Gallopheasants, 22 species & Sexual dimorphisn (range of traits) & No & No role of sexual selection in relation to diversification\\
-\cmidrule{1-5}
-Price-Waldman et al. (2020) & Thraupidae, 355 species & Plumage complexity & Yes & Elevated rates of plumage complexity evolution are associated with higher speciation rates\\
-\cmidrule{1-5}
- &  & Size dimorphism & Yes & Sexual size dimorphism predicts two out of three measures of speciation rates\\
-\cmidrule{3-5}
-\multirow{-2}{*}{\raggedright\arraybackslash This study} & \multirow{-2}{*}{\raggedright\arraybackslash Across passerines, 5812 species} & Plumage dichromatism & No & There was no link between plumage dichromatism (measured from spectral info or RGB values) and any speciation rate\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}
-\item Studies were obtained by searching 'Web of Science' for articles published from 2011 for terms containing ‘speciation’, ‘diversification’ and ‘sexual selection’.
-We summarised all the studies we found relevant and comparable to our study.
-\end{tablenotes}
-\end{threeparttable}}
-\end{table}
-
-\phantomsection
-
-\hypertarget{materials-and-methods}{%
-\section{Materials and methods}\label{materials-and-methods}}
-
-We examined the effect of sexual selection on speciation and extinction
-rate in 97\% of passerines (\emph{n} = 5,812 species; 58\% of all
-birds). Specifically, we (\emph{i}) compiled datasets for sexual
-dichromatism/selection strength and environmental variability,
-(\emph{ii}) obtained estimates of speciation and extinction rates across
-passerines, and (\emph{iii}) conducted phylogenetic generalized
-least-squares (PGLS) regressions. Analyses are documented with
-reproducible code in the
-\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}.
-\phantomsection
-
-\hypertarget{compiling-data-for-sexual-selection-and-environmental-stress}{%
-\subsection{Compiling data for sexual selection and environmental
-stress}\label{compiling-data-for-sexual-selection-and-environmental-stress}}
-
-\phantomsection
-
-\hypertarget{sexual-dichromatism}{%
-\subsubsection{Sexual dichromatism}\label{sexual-dichromatism}}
-
-We used a previously-published measure of sexual dichromatism for 5,983
-species of passerines (Dale et al. 2015). Briefly, Dale et al. (Dale et
-al. 2015) obtained sex-specific RGB (red-green-blue) values across six
-body patches (nape, crown, forehead, throat, upper breast, and lower
-breast) from \emph{Handbook of the Birds of the World} (Del Hoyo et al.
-2011). The relative contribution of male and female RGB colour values
-were averaged across body patches and provide `male-like' and
-`female-like' plumage scores. Here we use the absolute difference
-between male and female plumage scores as an estimate of sexual
-dichromatism. Technically, this measures differences in the `degree of
-male-ness' between males and females, rather than sex differences in
-colour \emph{per se} (i.e.~dichromatism in the strict sense). For
-example, the metric would fail to capture dichromatism when both the
-male and female possess a single, but differently coloured `male-like'
-patch. However, the metric is highly correlated with dichromatism
-measured from spectral data (see below).
-
-Additionally, we used another measure of dichromatism corresponding to
-colour distance in avian colour space derived from spectral data
-(Armenta et al. 2008). These measurements include variation in the
-ultraviolet and bird-visible range and --- unlike the RGB measures ---
-are sourced from museum specimens as opposed to illustrations. The
-spectrophotometry data covers only 581 passerine species (10-fold fewer
-than the RGB data), although there was a substantial correlation between
-the two dichromatism measures (\(r = 0.79\);
-\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S10}).
-\phantomsection
-
-\hypertarget{male-biased-sexual-selection}{%
-\subsubsection{Male-biased sexual
-selection}\label{male-biased-sexual-selection}}
-
-Sexual dichromatism is likely to be imperfectly correlated with
-variation in the strength of sexual selection across taxa. For this
-reason, we sourced an additional measure of sexual selection (Dale et
-al. 2015), referred to here as the `index of male-biased sexual
-selection'. This index is the first principal component from a
-phylogenetic principal component analysis (PPCA) of three
-characteristics possitively associated with sexual selection (sexual
-size dimorphism, social polygyny and {[}lack of{]} paternal care). The
-variables included in this index have all been positively linked to the
-intensity of sexual selection, and are usually correlated (Björklund
-1990; Owens and Hartley 1998; Dunn et al. 2001), which is why they were
-combined into a single metric in previous studies (Dale et al. 2015).
-This measure of male-biased sexual selection is available for only 2,465
-species, and shows a moderate correlation with the RGB measure of sexual
-dichromatism (\(r = 0.34\);
-\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Figure S12}).
-\phantomsection
-
-\hypertarget{environmental-variables}{%
-\subsubsection{Environmental variables}\label{environmental-variables}}
-
-We obtained estimates of species range size using expert range maps
-(BirdLife International and Handbook of the Birds of the World 2017).
-The names of 1,230 species in the Birdlife database (Hoyo and Collar
-2016) have been recently changed, so we manually matched these taxa with
-the names used in the sexual dichromatism dataset (Hoyo and Collar
-2016). For each species' range, we obtained estimates of climatic
-conditions by extracting 1,000 random point samples of each bioclimatic
-variable. We extracted 19 present-day bioclimatic variables
-(representing a variety of biologically relevant annual trends in
-temperature and precipitation) with 30-second (\textasciitilde{}1
-km\textsuperscript{2}) spatial resolution (Fick and Hijmans 2017). From
-the 1000 values of each bioclimatic variable, we obtained means and
-standard deviations for each species. Using the same spatial sampling,
-we extracted means and standard deviations of bioclimatic variables from
-the paleoclimate during the last interglacial (LIG; 120,000 - 140,000
-years ago) (Otto-Bliesner et al. 2006). To estimate variability in the
-energy available to species, we obtained the mean and standard deviation
-of net primary productivity (NPP) values between 2000 - 2015 across each
-species distribution. Estimates of NPP had 30-second resolution and were
-obtained through MODIS (Moderate Resolution Imaging Spectroradiometer)
-primary production products stage 3 (MOD17A3) (Zhao et al. 2005). We
-provide these data as a potentially useful data resource (see
-\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}).
-\phantomsection
-
-\hypertarget{generating-biologically-relevant-predictors-for-environmental-stress}{%
-\subsubsection{Generating biologically relevant predictors for
-environmental
-stress}\label{generating-biologically-relevant-predictors-for-environmental-stress}}
-
-Given that stressful environments are expected to interact with sexual
-selection and have a positive effect on adaptation (Cally et al. 2019),
-we used the extracted environmental variables from each species range
-size to create predictors of environmental variation/stress. We used
-(\emph{i}) the average NPP in each species' range and (\emph{ii}) the
-log-transformed range size as potentially informative predictors of
-speciation rates. We also used three environmental predictors derived
-from bioclimatic data. These predictors relate to seasonal climate
-variation, spatial climate variation and long-term climate variation. To
-obtain seasonal climate variation we used (\emph{iii}) mean values of
-temperature seasonality (BIO4) for each range. (\emph{iv}) To estimate
-levels of spatial environmental variation a species may endure, we used
-the first principle component (PC1) from a PCA on standard deviations
-from all bioclimatic variables, excluding temperature and precipitation
-seasonality (BIO4 and BIO15). PC1 was heavily loaded towards bioclimatic
-variables relating to temperature, thus PC1 largely reflects the
-variation in temperature across a species' range
-(\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Table S1}).
-A taxon's range size often correlates with speciation and extinction
-rates (Rosenzweig 1995; Castiglione et al. 2017), so we controlled for
-the correlation between environmental spatial variation and species'
-range sizes --- where larger ranges have larger variation in PC1 --- by
-using the residuals of a fitted general additive model (GAM;
-\href{https://justincally.github.io/SexualSelection_Speciation/#generating_biologically_relevent_predictors}{Figure S1})
-as a predictor. To obtain long-term variation in climates for each
-species range, we took (\emph{v}) the first principal component of the
-absolute difference in the bioclimatic variables between the LIG and
-current values. Similarly to spatial variation, the long-term climate
-variation is primarily loaded to temperature differences between the LIG
-and current climates
-(\href{https://justincally.github.io/SexualSelection_Speciation/#long_term_climate_variability_(lig_anomaly)}{Table S2, Figure S2}).
-The five predictors of environmental variability are not strongly
-correlated
-(\href{https://justincally.github.io/SexualSelection_Speciation/#correlations_between_environmental_predictors}{Figure S3}).
-Details and \texttt{R} code to generate these predictors can be found
-within the
-\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}.
-\phantomsection
-
-\hypertarget{estimating-extinction-and-speciation}{%
-\subsection{Estimating extinction and
-speciation}\label{estimating-extinction-and-speciation}}
-
-Phylogenetic information was obtained from www.birdtree.org (Jetz et al.
-2012). We used a maximum clade credibility (MCC) tree from 2,500 samples
-of the posterior distribution (\emph{n} = 5,965) as the main
-phylogenetic hypothesis in our comparative analysis. Additionally, a
-random draw of full trees (including species without genetic data) from
-the posterior distribution of phylogenetic trees was used for
-diversification analyses using tip-rate measures (1,000 trees) and BAMM
-(100 trees) (Rabosky 2014). These trees had crown clades with a topology
-that was heavily constrained on the basis of a previously published
-study (``Hacket backbone''; Hackett et al. 2008) and were constructed
-using a pure birth (Yule) model. We calculated three different tip-rate
-metrics of speciation and one of extinction across all trees.
-
-Diversification is the result of two processes, speciation and
-extinction through time. To estimate speciation rates, we first obtained
-two tip-rate metrics of speciation using statistics derived from the
-properties of the nodes and branches along root-to-tip paths of the
-phylogeny. Node density (ND) is a simple statistic calculating the
-density of nodes from the phylogenetic root to the tip, while the
-log-transformed equal splits (logES; also referred to as diversification
-rate/DR) is derived from the sum of edge lengths from each tip towards
-the root, with each edge towards the root having the length
-down-weighted (Jetz et al. 2012; Quintero and Jetz 2018; Rabosky et al.
-2018). Crucially, studies have suggested that DR and ND (henceforth
-referred to as \(\lambda_{DR}\) and \(\lambda_{ND}\)) are more
-reflective estimates of speciation than diversification. Because
-\(\lambda_{DR}\) and \(\lambda_{ND}\) cannot account for whole-clade
-extinctions and thus under-estimate extinction rate, which makes the
-composite measure of diversification more dependent on speciation
-(Belmaker and Jetz 2015; Title and Rabosky 2018). Therefore,
-\(\lambda_{DR}\) is a measure of speciation rate more heavily weighted
-to recent speciation events while \(\lambda_{ND}\) measures speciation
-across the root-to-tip path. These tip-rate measures are alternatives to
-state-dependent diversification models such as Quantitative State
-Speciation-Extinction (QuaSSE); but, based on previous simulation
-studies, \(\lambda_{DR}\) and \(\lambda_{ND}\) are robust and intuitive
-measures that provide high power and low false discovery rate with large
-phylogenies when incorporated into Phylogenetic Generalized Least
-Squares (PGLS) models (Harvey Michael et al. 2017).
-
-We used BAMM to model the dynamics of speciation and extinction across
-the 101 phylogenetic trees (one MCC tree and 100 random draws of the
-posterior). This software uses a Bayesian approach (reversible-jump
-Markov Chain Monte Carlo) to generate probability distributions of
-evolutionary rate-shift configurations with variable speciation and
-extinction rates (Rabosky 2014). These models provide tip-rate estimates
-of speciation and extinction rate that can be easily used in comparative
-analyses. The parameters of the 100 BAMM runs are detailed in full in
-the
-\href{https://justincally.github.io/SexualSelection_Speciation/#set_up_bamm_parameters}{Supplementary Information};
-briefly, we used a time-variable model with the prior expected number of
-evolutionary rate shifts set at 100 and prior rates set from the initial
-tip-level estimates of speciation and extinction using the
-\texttt{BAMMtools} R package (Rabosky et al. 2014). BAMM models were run
-independently for the 101 phylogenetic trees for 100 million
-generations. Given the computationally intensive nature of BAMM, runs
-were conducted across multiple CPUs. Important BAMM parameters
-(log-likelihood and number of rate shifts) reached convergence with
-effective sample size (ESS) of MCMC (Markov Chain Monte Carlo) samples
-surpassing 200; an arbitrary value, above which posterior distributions
-can often be accurately inferred
-(\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S3},
-\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Table S4}).
-Further details of BAMM parameters and output are available in the
-\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information},
-with tip-rate means and variances provided. Additionally, given the
-variability in BAMM estimates, we also provide analysis of BAMM shift
-configurations and tip-rate estimates from our run on the MCC tree and
-within a BAMM run on the MCC tree from a genetic-only phylogeny across
-all birds (Harvey et al. 2017). All analyses were conducted on
-log-rates. \phantomsection
-
-\hypertarget{phylogenetic-comparative-analysis}{%
-\subsection{Phylogenetic comparative
-analysis}\label{phylogenetic-comparative-analysis}}
-
-To test the association between speciation/extinction and sexual
-selection, environmental variability and their interaction, we used
-phylogenetic least squares (PGLS) models in the \texttt{nlme} package
-(Pinheiro et al. 2018). Firstly, we conducted model selection to compare
-models in which \(\lambda_{DR}\), \(\lambda_{ND}\), \(\lambda_{BAMM}\)
-or \(\mu_{BAMM}\) were the response variable: these tip-rate estimates
-all came from the same MCC tree (derived from 2,500 draws of the
-posterior distribution (Jetz et al. 2012)). For models of
-\(\lambda_{BAMM}\) and \(\mu_{BAMM}\), we used the inverse of the
-variance associated with each tip rate estimate as weights, to account
-for the variable precision of the estimates provided by BAMM. For each
-response variable, we conducted model selection to compare models with
-different combinations of predictor variables. The most complex model in
-each set under comparison contained one of the measures of sexual
-selection (sexual dichromatism or the index of male-biased sexual
-selection), all of the environmental measures (i.e.~log-transformed
-range size, seasonal temperature variation, spatial temperature
-variation, long-term temperature variation, and NPP), and all of the
-2-way interactions between sexual selection and each of the
-environmental measures. The simpler models contained all of the same
-main effects, but had fewer 2-way interaction terms (potentially none).
-Model selection was done in \texttt{MuMIn} using the \texttt{dredge}
-function (Bartoń 2017). Using the terms from the top-ranked model
-(ranked by AICc), we ran the equivalent model for each of the 1,000
-phylogenetic trees used to derive \(\lambda_{DR}\), \(\lambda_{ND}\) and
-each of the 100 trees used to derive \(\lambda_{BAMM}\) and
-\(\mu_{BAMM}\). Additionally, we investigated the effect of the
-individual variables used to derive the index of male-biased sexual
-selection on speciation rate. For these pgls models we replaced the
-composite index score with the individual biological variable (sexual
-size dimorphism, social polygyny and {[}lack of{]} paternal care) and
-ran the equivalent model for 300 phylogenetic trees used to derive
-\(\lambda_{DR}\), \(\lambda_{ND}\) and 100 trees used to derive
-\(\lambda_{BAMM}\).
-
-Across all our analyses we corrected for the phylogenetic signal. Our
-models used the unique response variables and correlation structure for
-a given phylogenetic tree. Specifically, for models using tip-rate
-metrics (\(\lambda_{DR}\), \(\lambda_{ND}\)), we estimated the
-phylogenetic signal independently for each of the 1,000 trees/models.
-Phylogenetic signal was estimated as Pagel's \(\lambda\) (Pagel 1999)
-using the \texttt{corPagel} function in the \texttt{ape} package
-(Paradis et al. 2004). Alternatively, for models using speciation and
-extinction estimates derived using BAMM (\(\lambda_{BAMM}\) and
-\(\mu_{BAMM}\)), we found that \(\lambda\) was consistently estimated at
-1 and hence assumed Brownian motion (using the \texttt{corBrownian}
-function) to estimate the correlation structure. This method enabled us
-to present model estimates for an MCC tree alongside 1,000/100 trees
-from the posterior distribution of trees to account for phylogenetic
-uncertainty. This approach was repeated on three datasets corresponding
-to each measure of sexual selection: dichromatism derived from RGB
-values of images (\emph{n} = 5,812); dichromatism from spectrophotometry
-(\emph{n} = 581) and the index of male-biased sexual selection (\emph{n}
-= 2,465).
-
-Finally, using the subset of species with an index of male-biased sexual
-selection, we conducted a phylogenetic path analysis using the
-\texttt{phylopath\ R} package (Bijl 2018). The phylogenetic path
-analysis was used to assess causal paths between variables unable to be
-modelled within the univariate response of PGLS. That is, a phylogenetic
-path analysis allowed us to model relationships between the predictor
-variables used in our PGLS analysis as we anticipate environmental
-variability, sexual dichromatism/selection, and range size to have
-effects on each other and not just on speciation rate. To minimise path
-complexity, we used temperature seasonality (BIO4) as the single measure
-for environmental variability, \(\lambda_{DR}\) as the single measure of
-speciation, and the tip-rates from the MCC tree. Further details of the
-path analysis, including our rationale for each path's directions, can
-be found within the
-\href{https://justincally.github.io/SexualSelection_Speciation/}{Supplementary Information}
-along with all other analyses and the relevant \texttt{R} code to
-reproduce results. \phantomsection
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\phantomsection
-
-\hypertarget{male-biased-sexual-selection-but-not-sexual-dichromatism-affects-speciation}{%
-\subsection{Male-biased sexual selection, but not sexual dichromatism,
-affects
-speciation}\label{male-biased-sexual-selection-but-not-sexual-dichromatism-affects-speciation}}
-
-We examined the effect of sexual selection on speciation and extinction
-rate in 97\% of passerines (\emph{n} = 5,812 species; 58\% of all birds;
-\autoref{Speciation_Figure}). We found a significant positive
-association between the index of male-biased sexual selection (\emph{n}
-= 2,465) and \(\lambda_{DR}\) from the maximum clade credibility (MCC)
-tree (\(\beta\) = \(3.89 \times 10^{-2}\), p = 0.01;
-\autoref{model_results}b). However, this association was not significant
-for the other two measures of speciation rate (\(\lambda_{ND}\):
-\(\beta\) = \(4.38 \times 10^{-4}\), p = 0.35; \(\lambda_{BAMM}\):
-\(\beta\) = \(9.42 \times 10^{-4}\), p = 0.76;
-\autoref{model_results}b). When we took into account phylogenetic
-uncertainty by running the models using 1,000 trees, the distribution of
-estimates from PGLS models was similar to the estimate from the MCC
-tree: among the 1,000 trees there was a positive association between
-sexual selection and \(\lambda_{DR}\) (highest posterior density (HPD)
-Interval = \(4.51 \times 10^{-3}\), \(5.72 \times 10^{-2}\)), and the
-distribution skewed towards a positive association between sexual
-selection and \(\lambda_{ND}\) (HPD Interval = \(-5.04 \times 10^{-4}\),
-\(1.58 \times 10^{-3}\)) as well as the 100 models using
-\(\lambda_{BAMM}\) (HPD Interval = \(-1.30 \times 10^{-2}\),
-\(3.09 \times 10^{-2}\);
-\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15}).
-
-We investigated which of the three variables comprising the index of
-male-biased sexual selection was driving the association observed with
-\(\lambda_{DR}\). Our results over 300 trees showed that this pattern is
-mainly driven by the sexual size dimorphism component (HPD Interval =
-\(8.53 \times 10^{-1}\), 3.11), with the effects of other components
-overlapping zero; paternal care (HPD Interval =
-\(-1.78 \times 10^{-1}\), \(7.90 \times 10^{-3}\)) and mating system
-(HPD Interval= \(-7.35 \times 10^{-2}\), \(4.32 \times 10^{-2}\)).
-Importantly, the association between sexual size dimorphism and
-speciation rates is also present when using \(\lambda_{ND}\) (HPD
-Interval = \(1.80 \times 10^{-1}\), \(6.38 \times 10^{-1}\)), but not
-when using \(\lambda_{BAMM}\) (HPD Interval = -1.49,
-\(7.45 \times 10^{-1}\), \autoref{SS_components}).
-
-In contrast to male-biased sexual selection, we found no evidence that
-species with increased sexual dichromatism have higher or lower rates of
-speciation. Sexual dichromatism showed no association with
-\(\lambda_{DR}\) (\(\beta\) = \(-1.28 \times 10^{-3}\), p = 0.15;
-\autoref{model_results}a, \autoref{Speciation_Figure}), \(\lambda_{ND}\)
-(\(\beta\) = \(-5.75 \times 10^{-5}\), p = 0.08;
-\autoref{model_results}a) or \(\lambda_{BAMM}\) (\(\beta\) =
-\(-1.43 \times 10^{-5}\), p = 0.87; \autoref{model_results}a). PGLS
-analyses using sexual dichromatism (\emph{n} = 581) measured by
-spectrophotometry (Armenta et al. 2008) yielded results concordant with
-the full dataset; i.e.~no association between sexual dichromatism and
-speciation
-(\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
-Our results from models based on the MCC tree are largely corroborated
-by model estimates from PGLS analyses of the rates and correlation
-structures from 1,000 trees (for \(\lambda_{DR}\), \(\lambda_{ND}\)) and
-100 trees for \(\lambda_{BAMM}\). The HPD intervals show model estimates
-are distributed around zero when using complete taxon sampling models
-and RGB measures of sexual dichromatism (\(\lambda_{DR}\): HPD Interval
-= \(-1.63 \times 10^{-3}\), \(1.66 \times 10^{-3}\), \(\lambda_{ND}\):
-HPD Interval = \(-4.26 \times 10^{-5}\), \(5.50 \times 10^{-5}\),
-\autoref{model_results}a,
-\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}).
-For PGLS models using spectrophotometry-based measures of sexual
-dichromatism, the estimates from the 100 trees in the \(\lambda_{DR}\)
-models are positively skewed (HPD Interval = \(-1.78 \times 10^{-2}\),
-\(3.49 \times 10^{-2}\)) but normally distributed around zero for
-\(\lambda_{ND}\) and \(\lambda_{BAMM}\)
-(\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).
-
-Our analyses also show that the differences in results between sexual
-dichromatism and male-biased sexual selection (i.e.~association with
-speciation rates only for the latter) were not due to differences in the
-size of the datasets used (5,812 species vs.~2,465,
-\href{https://justincally.github.io/SexualSelection_Speciation/#testing_the_sexual_dichromatism_dataset_using_the_same_data_points_as_the_sexual_selection_dataset}{Figure S17}).
-No interaction terms were present in the top models (\(\Delta\) AICc
-\textgreater{} 4) for any measure of speciation (\(\lambda_{DR}\),
-\(\lambda_{ND}\), \(\lambda_{BAMM}\)) or sexual selection (RGB values,
-spectrophotometry and the index of male-biased sexual selection;
-\(\Delta\) AICc \textgreater{} 4;
-\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5},
-\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6},
-\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11},
-\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S14}).
-Thus we found no evidence that the effect of sexual selection on
-speciation is dependent on our measures of environmental variation or
-range size. Furthermore, we found no evidence that these environmental
-factors --- seasonal temperature variation, long-term temperature
-variation, spatial temperature variation, and Net Primary Productivity
-(NPP) --- predict speciation independently from sexual
-dichromatism/selection (\autoref{model_results},
-\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).
-\phantomsection
-
-\hypertarget{species-with-smaller-ranges-have-increased-rates-of-speciation}{%
-\subsection{Species with smaller ranges have increased rates of
-speciation}\label{species-with-smaller-ranges-have-increased-rates-of-speciation}}
-
-Based on \(\lambda_{DR}\) and \(\lambda_{ND}\) tip-rate metrics of
-speciation, we found a negative association between range size and
-speciation; that is, species with smaller ranges show marginally higher
-values for \(\lambda_{DR}\) and \(\lambda_{ND}\). This negative
-association was small but significant for models using the MCC tree
-(\(\lambda_{DR}\): \(\beta\) = \(-6.58 \times 10^{-3}\), p =
-\(1.48 \times 10^{-3}\); \(\lambda_{ND}\): \(\beta\) =
-\(-1.46 \times 10^{-4}\), p = 0.03; \autoref{model_results}a,
-\autoref{Speciation_Figure}). This association was also evident across
-the estimates from models using the 1,000 trees (\(\lambda_{DR}\): HPD
-Interval = \(-8.87 \times 10^{-3}\), \(-6.61 \times 10^{-4}\);
-\(\lambda_{ND}\): HPD Interval = \(-1.51 \times 10^{-4}\),
-\(1.72 \times 10^{-5}\); \autoref{model_results}a). Subset models with
-reduced sample size and different measures of sexual selection --- but
-the same measure of range size --- showed equivocal evidence that range
-size is negatively associated with speciation. Range size is
-significantly associated with \(\lambda_{DR}\)
-(\autoref{model_results}b) using data subset for species with an index
-of male-biased sexual selection (\emph{n} = 2,465) but not
-\(\lambda_{ND}\) or \(\lambda_{BAMM}\). Models using data subset for
-spectrophotometry-based dichromatism (\emph{n} = 581) gave
-non-significant estimates for the effect of range size on all measures
-of speciation
-(\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11},
-\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12},
-\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S13}).
-Because the range size dataset is the same across the three data
-subsets, we draw our conclusions from the models with the highest power
-using near-complete taxon sampling (\emph{n} = 5,812). \phantomsection
-
-\hypertarget{phylogenetic-path-analysis}{%
-\subsection{Phylogenetic path
-analysis}\label{phylogenetic-path-analysis}}
-
-Using a phylogenetic path analysis, we found multiple significant paths
-between variables used in the PGLS (\autoref{Path_Analysis};
-\href{https://justincally.github.io/SexualSelection_Speciation/#phylogenetic_path_analysis}{Figure S14}).
-There was a modest effect of male-biased sexual selection on sexual
-dimorphism (\(\beta\) = 0.22). Additionally, temperature seasonality
-weakly affected sexual dimorphism (\(\beta\) = 0.07) and strongly
-affected range size (\(\beta\) = 0.52). This suggests an indirect effect
-of temperature seasonality on \(\lambda_{DR}\) (\(\beta_{indirect}\) =
--0.02; \autoref{Path_Analysis}), given the negative association we
-identified between \(\lambda_{DR}\) and range size in PGLS models.
-\phantomsection
-
-\hypertarget{extinction-rate}{%
-\subsection{Extinction rate}\label{extinction-rate}}
-
-We found no evidence that extinction (\(\mu_{BAMM}\)) was impacted by
-the extent of sexual dichromatism for full-taxon sampling (\(\beta\) =
-\(2.38 \times 10^{-5}\), p = 0.93; \autoref{model_results}a), nor
-spectrophotometry-based measures of sexual dichromatism
-(\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11},
-\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12},
-\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S13})
-or male-biased sexual selection (\autoref{model_results}b,
-\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S15},
-\href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S16}).
-\phantomsection
-
-\hypertarget{variability-across-phylogenetic-trees-and-speciation-rate-measures}{%
-\subsection{Variability across phylogenetic trees and speciation rate
-measures}\label{variability-across-phylogenetic-trees-and-speciation-rate-measures}}
-
-Estimates of the effect of predictor variables on speciation rates
-varied across phylogenetic trees, especially in the BAMM rates
-(\(\lambda_{BAMM}\) and \(\mu_{BAMM}\)), where the 95 \% HPD interval
-across PGLS model estimates from 100 trees was often \textgreater{} 20
-times larger than the 95 \% confidence interval for estimates from a
-single PGLS model using the MCC tree. This contrasts with variation
-across trees for the other rate estimates (\(\lambda_{DR}\) and
-\(\lambda_{ND}\)), where the 95 \% HPD interval of model estimates for
-pgls models using 1,000 trees was near-equivelant to the 95 \%
-confidence interval calculated for PGLS model estimates of the MCC tree
-(\href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S9}).
-The great majority of earlier studies have based their estimates on a
-single consensus tree due to the computational requirements of BAMM.
-However, our results suggest that BAMM estimates between alternative,
-similarly plausible phylogenies vary substantially. Mean measures of
-speciation rate across 100 trees were positively correlated between
-measures (\(\lambda_{DR}\) - \(\lambda_{BAMM}\): \emph{r}=0.75,
-\(\lambda_{DR}\) - \(\lambda_{ND}\): \emph{r}=0.65, \(\lambda_{ND}\) -
-\(\lambda_{BAMM}\): \emph{r}=0.51;
-\href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}).
-The calculation of BAMM rates can be affected by the settings of the run
-and the use of different priors. We therefore compared the estimate of
-our MCC tree with that of previously published analyses on birds and
-found a high correlation (\emph{r}=0.81,
-\href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S6, Figure S8},
-Harvey et al. (2017)). Full details of the BAMM results are presented as
-supplementary materials.
-
-\begin{figure}
-\centering
-\includegraphics{Figures/DRSpeciationFigure.pdf}
-\caption{Speciation rate (\(\lambda_{DR}\)) across all passerine birds
-(\emph{n} = 5,965) with estimates of sexual dichromatism, range size
-available for 5,812 species and an index of male-biased sexual selection
-available for 2,465 species. Across these species there was a small but
-significant negative association between \(\lambda_{DR}\) and log-range
-size as well as a significant positive association between
-\(\lambda_{DR}\) and male-biased sexual selection but no significant
-association between \(\lambda_{DR}\) and sexual dichromatism based on
-RGB measures. \(\lambda_{DR}\) are those from the MCC tree and images of
-birds are from the \emph{Handbook of the Birds of the World}. Clockwise
-the six species are: \emph{Sporophila bouvronides}, \emph{Euplectes
-franciscanus}, \emph{Phainopepla nitens}, \emph{Paradisaea rubra},
-\emph{Malurus pulcherrimus}, \emph{Lepidothrix coeruleocapilla}. Edge
-colours for the terminal branch correspond to \(\lambda_{DR}\) but all
-precluding branches have been generated for graphical purposes using
-ancestral character state estimation (Revell 2012) and should not be
-interpreted. Illustrations reproduced by permission of Lynx Editions.
-\label{Speciation_Figure}}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Figures/Estimates_and_scatter_edit.pdf}
-\caption{Model estimates (a) showing the effect size (i.e.~slope) of
-log-range size and sexual dichromatism on speciation and extinction
-rates using PGLS analyses with the sexual dichromatism dataset (RGB
-values, \emph{n} = 5,812). (b) depicts the scatter plot of speciation
-rate (\(\lambda_{DR}\)) and log-range size with the model estimate
-presented as a dashed line. (c) shows the scatter plot of speciation
-rate (\(\lambda_{DR}\)) and male-biased sexual selection (\emph{n} =
-2,465). Similar to (a), (d) presents model estimates for PGLS analyses
-using a restricted dataset with measures of an index of male-biased
-sexual selection (\emph{n} = 2,465) and shows the effect of log-range
-size and male-biased sexual selection on speciation and extinction
-rates. Both datasets were used for analyses with three measures of
-speciation (\(\lambda_{DR}\), \(\lambda_{ND}\), \(\lambda_{BAMM}\)) and
-one measure of extinction (\(\mu_{BAMM}\)) as response variables. The
-numerical values for the model estimates using the MCC tree and HPD
-intervals of estimates from 1,000 randomly sampled trees (for
-\(\lambda_{DR}\) and \(\lambda_{ND}\)) or 100 randomly sampled trees for
-\(\lambda_{BAMM}\) and \(\mu_{BAMM}\) can be found in the Supplementary
-Information. Density curves are based on model estimates from 1,000/100
-trees and the circle below with error bars is the estimate and 95\% CIs
-from the MCC tree. For this figure we removed outliers from estimates
-coming from the 100 randomly sampled trees for BAMM models in order to
-be able to visualise the MCC 95\% CIs. \label{model_results}}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Figures/Individual_SS_Estimates.pdf}
-\caption{Estimates of the effect of individual sexual selection
-components included in the PPCA (paternal care, sexual size dimorphism
-and mating system) on three measures of speciation rate
-(\(\lambda_{DR}\), \(\lambda_{ND}\) and \(\lambda_{BAMM}\)). Estimates
-are presented as density intervals from pgls models on 300 phylogentic
-trees that used species with available data for these sexual selection
-measures (\emph{n} = 2,465). The bar under each density ridge is the 95
-\% Highest Posterior Density Interval. Given that the mating system is a
-categorical variable, model estimates for three polygynous mating system
-levels are in reference to a striclty monogamous mating system (0\%
-polygyny). \label{SS_components}}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Figures/Path_Analysis.pdf}
-\caption{Path analysis of evolutionary and ecological variables. Arrows
-represent direct effects with the direction of effect corresponding to
-colours (blue = positive, red = negative). The numeric values are
-standardised regression slopes and the asterisks indicates that the 95\%
-confidence intervals of this estimate do not overlap with zero. The
-confidence intervals were obtained from 500 bootstrapped iterations and
-the data used in this analysis was subset to species with both sexual
-dichromatism and an index of male-biased sexual selection measures
-(\emph{n} = 2,465). \label{Path_Analysis}}
-\end{figure}
-
-\newpage\phantomsection
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-We found evidence that the composite index of male-biased sexual
-selection, but not measures of sexual dichromatism, is correlated with
-the rate of speciation in passerine birds. The absence of a detectable
-correlation between sexual dichromatism and speciation rate was
-consistent across different measures of speciation (\(\lambda_{DR}\),
-\(\lambda_{ND}\) and \(\lambda_{BAMM}\)) and both measures of
-dichromatism (spectral and RGB), and it cannot be explained by a
-difference in statistical power or sampling. These findings reaffirm the
-conclusions of previous, smaller studies in which sexual dichromatism
-was measured using spectrophotometry (Huang and Rabosky 2014) or human
-observers (Cooney et al. 2017). The correlation between speciation rate
-and the index of male-biased sexual selection (which encapsulates
-variation in sexual size dimorphism, social polygyny, and paternal care)
-was statistically significant for \(\lambda_{DR}\), but not for
-\(\lambda_{ND}\) and \(\lambda_{BAMM}\). This pattern seems to be mainly
-driven by an association between sexual size dimorphism and speciation.
-Interestingly, we also found a consistent negative relationship between
-range size and speciation rate, at least when this rate was quantified
-using \(\lambda_{DR}\) and \(\lambda_{ND}\). None of the bioclimatic
-measures of environmental variability that we investigated (i.e.,
-temperature seasonality, long-term temperature variation, and spatial
-temperature variation) were significantly associated speciation rate,
-nor mediated the relationship between sexual selection and
-diversification.
-
-The difference in findings between the analyses of sexual dichromatism
-\emph{versus} the index of male-biased sexual selection is noteworthy,
-because the majority of earlier studies used dichromatism alone as their
-proxy for sexual selection (e.g., Barraclough et al. 1995; Owens et al.
-1999; Morrow et al. 2003; Seddon et al. 2013; Huang and Rabosky 2014).
-Given our findings, and the modest correlation between dichromatism and
-the sexual selection index (\(r = 0.34\); Dale et al. 2015), we suggest
-that sexual dichromatism may not be a robust proxy for sexual selection
-(Cooney et al. 2018). Although dichromatism almost certainly provides
-some insight into the operation of sexual selection, it may be too
-indirect a measure to detect any association with speciation rate, even
-with large sample size. There are several reasons why the use of sexual
-dichromatism as a proxy for sexual selection is problematic. Sexual
-dichromatism can evolve for reasons other than sexual selection, such as
-when males and females occupy different ecological niches (Wallace 1889;
-Kottler 1980; Slatkin 1984; Shine 1989) or experience different
-selective pressures in contexts other than competition for mates (Price
-and Eaton 2014). For example, in superb fairy-wrens (\emph{Malurus
-cyaneus}) female colouration has probably evolved in response to spatial
-variation in predation pressure, increasing dichromatism (Medina et al.
-2017). In fact, our path analysis detected a weak relationship between
-environment and sexual dichromatism, where sexual dichromatism was
-positively predicted by temperature seasonality (a measure of
-environmental variation).
-
-In line with some theoretical predictions and previous studies
-(Kraaijeveld et al. 2011), we found that male-biased sexual selection
-increases speciation rate, at least when speciation is measured by
-\(\lambda_{DR}\). Many of the species that have both high scores of
-male-biased sexual selection and high diversification rates belong to
-the genera \emph{Ploceus}, \emph{Euplectes} (Ploceidae) and
-\emph{Paradisaea} (Paradiaseidae). Multiple weaver species (Ploceidae)
-are polygynous and lack paternal care, and both weavers and birds of
-paradise have strong size dimorphism. The association between speciation
-rates and principal component scores that we report seems to be mainly
-driven by sexual size dimorphism and, to a lesser extent, paternal care.
-Speciation rates (both \(\lambda_{DR}\) and \(\lambda_{ND}\)) are higher
-in species with larger sexual dimorphism and \(\lambda_{DR}\) also has a
-tendency to be higher in species with no paternal care. Size dimorphism
-is often thought to arise as a consequence of intrasexual competition,
-where one of the sexes (males in most birds) has to compete for access
-to the other sex, leading to selection for larger body sizes and thus
-greater dimorphism (Björklund 1990; Owens and Hartley 1998). Therefore,
-competition between males could be the underlying driver of the high
-speciation rates that we detect in some clades.
-
-Sexual dimorphism due to competition within sexes contrasts with the
-drivers of sexual dichromatism. Plumage dichromatism can evolve as a
-consequence of female cryptic choice and be related to extra-pair
-fertilizations, but not necessarily paternal care or mating system
-(Owens and Hartley 1998). It can also arise as a result of selection on
-the level of crypsis of the sex that cares for offspring (Dale et al.
-2015). The fact that traits linked with competition (such as size
-dimorphism) are the ones associated with higher \(\lambda_{DR}\) values
--- rather than sexual dichromatism -- supports the general view that
-antagonistic interactions and sexual conflict can lead to increased
-diversity (Bonduriansky 2011; Qvarnström et al. 2012; Tinghitella et al.
-2018; Tsuji and Fukami 2020). Moreover, body size is a trait that
-influences multiple aspects of the physiology and ecology of a species.
-Differences in body size (as a result of sexual selection) could be
-linked to changes in diet, vulnerability to predators or environmental
-tolerance (Damuth 1993; Liow et al. 2008; Bonduriansky 2011), and such
-differences could ultimately increase the likelihood of divergence
-between young lineages. In mammals, sexual selection is suggested to
-have driven the evolution of large body size which in turn has allowed
-diversification of ecological strategies in the clade, and higher
-speciation rates (McLain 1993; Bonduriansky 2011).
-
-We also found that the association between sexual selection and
-speciation appears to be independent of net primary productivity and
-spatiotemporal variation in the environment. The lack of an effect of
-these environmental variables on speciation rate has several possible
-interpretations. Firstly, the effects of sexual selection on adaptation
-and speciation may depend on the type of environmental variability under
-which the species is evolving. Specifically, speciation rates might be
-impacted by genetic constraints on adaptation, that vary across
-environments. Theory suggests that sexual antagonism (which is often
-exacerbated in species with strong sexual selection) may be lower in
-habitats experiencing cyclical environmental variation
-(e.g.~seasonality), relative to those experiencing directional change in
-the environment (Connallon and Hall 2016). Another possibility is that
-the environmental predictors we chose may not account for the key
-ecological sources of selection that interact with sexual selection to
-drive speciation. For example, our study does not include direct measure
-of food availability or the severity of predation and parasitism, which
-are both hypothesised to affect sexual selection and speciation
-(reviewed in Maan and Seehausen 2011). Finally, it is possible that
-environmental variability genuinely has little effect on speciation
-rates, at least in the taxa investigated here.
-
-We found that species with smaller ranges have elevated speciation
-rates. This result is similar to a study of 329 amphibian genera, which
-found higher diversification rates in taxa with smaller range size
-(Greenberg and Mooers 2017). Intuitively, large range size should
-promote speciation by creating more opportunities for geographic
-barriers to form (Rosenzweig 1995; Castiglione et al. 2017). However,
-the opposite pattern is also plausible because birds with limited
-dispersal or more specialised niches can have more fragmented
-populations, which would promote vicariant divergence and higher
-speciation rates (Jablonski and Roy 2003; Birand et al. 2012; Claramunt
-et al. 2012). Moreover, species that have recently split as a
-consequence of vicariant divergence might have smaller ranges as a
-result of the split of the ancestral lineage, leading to a link between
-smaller ranges and shorter divergence times. It is also possible that
-high speciation rates cause smaller range sizes, rather than the other
-way around, for example because repeatedly-speciating lineages tend to
-fill niches in ways that hinder the geographical expansion of new
-species (Rosenzweig 1995; Weir and Price 2011; Price and Eaton 2014).
-However, species undergoing adaptive radiation in new habitats are
-unlikely to be limited by competition for resources from existing taxa.
-One further explanation for the negative association between range size
-and sexual dichromatism/sexual selection is the potential bias of
-taxonomic classification, whereby over-splitting of species in clades
-with large ranges leads to increased recent phylogenetic branching as
-well as smaller ranges.
-
-In addition to speciation, sexual selection is hypothesised to affect
-extinction. Using the model-based approach of BAMM, we found no
-association between the estimated extinction rate and sexual
-dichromatism, male-biased sexual selection, or our measures of
-environmental variability. However, these extinction results should not
-be regarded as definitive because extinction is notoriously difficult to
-estimate accurately from phylogenies, principally because different
-combinations of speciation and extinction rates can give rise to similar
-patterns of diversity (see Rabosky 2016). Phylogenetic methods such as
-BAMM allow for speciation and extinction rates to be estimated using
-moderately-sized phylogenies, although the ability of BAMM to model
-evolutionary rate shifts and extinction rates is debated (see, Beaulieu
-and O'Meara 2015; Rabosky 2016; Moore et al. 2016; Rabosky et al. 2017).
-Additionally, while several tip-rate estimates exist for speciation rate
-(e.g., \(\lambda_{DR}\) and \(\lambda_{ND}\)), tip-rate estimates of
-extinction rate are difficult to obtain without complex Bayesian models
-which are sensitive to sampling bias (Davis et al. 2013). Although
-extinction rates can be inferred from alternative sources, such as the
-fossil record (Martins et al. 2018), direct observation extinction, or
-IUCN red list status (Greenberg and Mooers 2017), each approach has its
-limitations. Across the passerine bird phylogeny, we found that BAMM
-often produced homogeneous speciation and extinction rates for smaller
-clades showing few rate shifts, which might reduce our power to detect
-small differences in extinction rates among closely-related taxa
-(Rabosky et al. 2017; Title and Rabosky 2018). Thus, this methodological
-constraint likely decreases our ability to accurately measure the
-correlation between metrics of sexual selection and the probability of
-extinction.
-
-One outcome of our analyses was that different measures of speciation
-rates presented different results. This is not completely surprising,
-because each of the rates is calculated differently (Title and Rabosky
-2018). For instance, \(\lambda_{DR}\) is weighted more towards
-speciation events close to the tips and allows more rate heterogeneity
-compared to \(\lambda_{BAMM}\) estimates. Rate shifts are unlikely to be
-detected in smaller clades in BAMM, meaning that it is not uncommon for
-whole genera to have the same rate. Using the \(\lambda_{DR}\) metric,
-only sister species are guaranteed to have the same rate. This leads to
-greater variation in \(\lambda_{DR}\) relative to the \(\lambda_{BAMM}\)
-estimates, which is suggested to be an advantage when studying
-diversification patterns (Quintero et al. 2015). Additionally,
-\(\lambda_{BAMM}\) estimates were more sensitive to phylogenetic
-uncertainty and were 20 times more variable across trees compared to
-\(\lambda_{DR}\) estimates. We cannot completely reject the idea that
-the lack of association between \(\lambda_{BAMM}\) and sexual selection
-could be the result of low statistical power, due to the combination of
-both low variation across species in the speciation rates and high
-levels of variation in the estimates across trees.
-
-To summarise, we have shown that sexual size dimorphism (a putative
-proxy for male-biased sexual selection), but not sexual dichromatism,
-predicts speciation in passerines, that the magnitude of this effect is
-modest, and that this relationship is not markedly affected by
-environmental variability. We have also shown that there is no evidence
-of an association between sexual selection and extinction rates.
-Overall, our findings imply that male-male competition could be the
-mechanism driving increased speciation rates in birds, that sexual
-dichromatism may not be a reliable proxy for sexual selection, and that
-alternative measures of sexual selection are more directly related to
-diversification.
 
@@
 and as an attached HTML document (Supplementary Material) for reviewers.
-Data is archived in the adjoined GitHub repository and the environmental
-dataset collated for this manuscript is available on Dryad:
-\url{https://doi.org/10.5061/dryad.573n5tb6n}.
-
-\newpage\phantomsection
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Armenta_2008}{}%
-Armenta, J. K., P. O. Dunn, and L. A. Whittingham. 2008. Quantifying
-avian sexual dichromatism: A comparison of methods. Journal of
-Experimental Biology 211:2423.
-
-\leavevmode\hypertarget{ref-Barraclough_1995}{}%
-Barraclough, T. G., P. H. Harvey, and S. Nee. 1995. Sexual selection and
-taxonomic diversity in passerine birds. Proceedings of the Royal Society
-B-Biological Sciences 259:211--215.
-
-\leavevmode\hypertarget{ref-MuMIn_2017}{}%
-Bartoń, K. 2017. MuMIn: Multi-model inference.
-
-\leavevmode\hypertarget{ref-Beaulieu_2015}{}%
-Beaulieu, J. M., and B. C. O'Meara. 2015. Extinction can be estimated
-from moderately sized molecular phylogenies. Evolution 69:1036--1043.
-
-\leavevmode\hypertarget{ref-Belmaker_2015}{}%
-Belmaker, J., and W. Jetz. 2015. Relative roles of ecological and
-energetic constraints, diversification rates and region history on
-global species richness gradients. Ecology Letters 18:563--571.
-
-\leavevmode\hypertarget{ref-Berger_2014}{}%
-Berger, D., K. Grieshop, M. I. Lind, J. Goenaga, A. A. Maklakov, and G.
-Arnqvist. 2014. Intralocus sexual conflict and environmental stress.
-Evolution 68:2184--2196.
-
-\leavevmode\hypertarget{ref-Wouter_2018}{}%
-Bijl, W. van der. 2018. \emph{Phylopath}: Easy phylogenetic path
-analysis in R. PeerJ 6:e4718.
-
-\leavevmode\hypertarget{ref-Birand_2012}{}%
-Birand, A., A. Vose, and S. Gavrilets. 2012. Patterns of species ranges,
-speciation, and extinction. American Naturalist 179:1--21.
-
-\leavevmode\hypertarget{ref-birdlife_2017}{}%
-BirdLife International and Handbook of the Birds of the World. 2017.
-Bird species distribution maps of the world.
-\url{http://datazone.birdlife.org/species/requestdis}.
-
-\leavevmode\hypertarget{ref-bjorklund_1990}{}%
-Björklund, M. 1990. A phylogenetic interpretation of sexual dimorphism
-in body size and ornament in relation to mating system in birds. Journal
-of Evolutionary Biology 3:171--183. Wiley Online Library.
-
-\leavevmode\hypertarget{ref-Bonduriansky_2011}{}%
-Bonduriansky, R. 2011. Sexual selection and conflict as engines of
-ecological diversification. American Naturalist 178:729--745.
-
-\leavevmode\hypertarget{ref-Bonduriansky_Chenoweth_2009}{}%
-Bonduriansky, R., and S. F. Chenoweth. 2009. Intralocus sexual conflict.
-Trends in Ecology \& Evolution 24:280--8.
-
-\leavevmode\hypertarget{ref-Cally_2019}{}%
-Cally, J. G., D. Stuart-Fox, and L. Holman. 2019. Meta-analytic evidence
-that sexual selection improves population fitness. Nature communications
-10:2017. Nature Publishing Group.
-
-\leavevmode\hypertarget{ref-Candolin_2008}{}%
-Candolin, U., and J. Heuschele. 2008. Is sexual selection beneficial
-during adaptation to environmental change? Trends in Ecology \&
-Evolution 23:446--452.
-
-\leavevmode\hypertarget{ref-Castiglione_2017}{}%
-Castiglione, S., A. Mondanaro, M. Melchionna, C. Serio, M. Di Febbraro,
-F. Carotenuto, and P. Raia. 2017. Diversification rates and the
-evolution of species range size frequency distribution. Frontiers in
-Ecology and Evolution 5:147. Frontiers.
-
-\leavevmode\hypertarget{ref-Claramunt_2012}{}%
-Claramunt, S., E. P. Derryberry, J. V. Remsen, and R. T. Brumfield.
-2012. High dispersal ability inhibits speciation in a continental
-radiation of passerine birds. Proceedings of the Royal Society B:
-Biological Sciences 279:1567.
-
-\leavevmode\hypertarget{ref-Connallon_Hall_2016}{}%
-Connallon, T., and M. D. Hall. 2016. Genetic correlations and
-sex-specific adaptation in changing environments. 70:2198.
-
-\leavevmode\hypertarget{ref-Cooney_2018}{}%
-Cooney, C. R., H. E. A. MacGregor, N. Seddon, and J. A. Tobias. 2018.
-Multi-modal signal evolution in birds: Re-examining a standard proxy for
-sexual selection. Proceedings of the Royal Society of London B:
-Biological Sciences 285.
-
-\leavevmode\hypertarget{ref-Cooney_2017}{}%
-Cooney, C. R., J. A. Tobias, J. T. Weir, C. A. Botero, and N. Seddon.
-2017. Sexual selection, speciation and constraints on geographical range
-overlap in birds. Ecology Letters 20:863--871.
-
-\leavevmode\hypertarget{ref-Dale_2015}{}%
-Dale, J., C. J. Dey, K. Delhey, B. Kempenaers, and M. Valcu. 2015. The
-effects of life history and sexual selection on male and female plumage
-colouration. Nature 527:367--370.
-
-\leavevmode\hypertarget{ref-Damuth_1993}{}%
-Damuth, J. 1993. Cope's rule, the island rule and the scaling of
-mammalian population density. Nature 365:748--750. Nature Publishing
-Group.
-
-\leavevmode\hypertarget{ref-davies1986re}{}%
-Davies, N., and A. Houston. 1986. Reproductive success of dunnocks,
-\emph{Prunella modularis}, in a variable mating system. II. Conflicts of
-interest among breeding adults. Journal of Animal Ecology 55:139--154.
-
-\leavevmode\hypertarget{ref-Davis_2013}{}%
-Davis, M. P., P. E. Midford, and W. Maddison. 2013. Exploring power and
-parameter estimation of the BiSSE method for analyzing species
-diversification. BMC Evolutionary Biology 13:38.
-
-\leavevmode\hypertarget{ref-delhoyo2011}{}%
-Del Hoyo, J., A. Elliott, and D. Christie. 2011. Handbook of the birds
-of the world. Lynx Edicions 2003-2011.
-
-\leavevmode\hypertarget{ref-dunn_2001}{}%
-Dunn, P. O., L. A. Whittingham, and T. E. Pitcher. 2001. Mating systems,
-sperm competition, and the evolution of sexual dimorphism in birds.
-Evolution 55:161--175. Wiley Online Library.
-
-\leavevmode\hypertarget{ref-Fick_2017}{}%
-Fick, S. E., and R. J. Hijmans. 2017. WorldClim 2: New 1-km spatial
-resolution climate surfaces for global land areas. International Journal
-of Climatology 37:4302--4315.
-
-\leavevmode\hypertarget{ref-Fromhage_2016}{}%
-Fromhage, L., and M. D. Jennions. 2016. Coevolution of parental
-investment and sexually selected traits drives sex-role divergence.
-Nature Communications 7:12517.
-
-\leavevmode\hypertarget{ref-Gomes_2016}{}%
-Gomes, A. C. R., M. D. Sorenson, and G. C. Cardoso. 2016. Speciation is
-associated with changing ornamentation rather than stronger sexual
-selection. Evolution 70:2823--2838.
-
-\leavevmode\hypertarget{ref-Greenberg_2017}{}%
-Greenberg, D. A., and A. Ø. Mooers. 2017. Linking speciation to
-extinction: Diversification raises contemporary extinction risk in
-amphibians. Evolution Letters 1:40--48.
-
-\leavevmode\hypertarget{ref-Hackett_2008}{}%
-Hackett, S. J., R. T. Kimball, S. Reddy, R. C. K. Bowie, E. L. Braun, M.
-J. Braun, J. L. Chojnowski, W. A. Cox, K.-L. Han, J. Harshman, C. J.
-Huddleston, B. D. Marks, K. J. Miglia, W. S. Moore, F. H. Sheldon, D. W.
-Steadman, C. C. Witt, and T. Yuri. 2008. A phylogenomic study of birds
-reveals their evolutionary history. Science 320:1763--1768.
-
-\leavevmode\hypertarget{ref-harano2010intralocus}{}%
-Harano, T., K. Okada, S. Nakayama, T. Miyatake, and D. J. Hosken. 2010.
-Intralocus sexual conflict unresolved by sex-limited trait expression.
-Current Biology 20:2036--2039.
-
-\leavevmode\hypertarget{ref-Harvey_2017}{}%
-Harvey, M. G., G. F. Seeholzer, B. T. Smith, D. L. Rabosky, A. M.
-Cuervo, and R. T. Brumfield. 2017. Positive association between
-population genetic differentiation and speciation rates in new world
-birds. Proceedings of the National Academy of Sciences 114:6328--6333.
-
-\leavevmode\hypertarget{ref-Harvey_essim_2017}{}%
-Harvey Michael, G., L. Rabosky Daniel, and N. Cooper. 2017. Continuous
-traits and speciation rates: Alternatives to state‐dependent
-diversification models. Methods in Ecology and Evolution 9:984--993.
-
-\leavevmode\hypertarget{ref-Holman_Kokko_2013}{}%
-Holman, L., and H. Kokko. 2013. The consequences of polyandry for
-population viability, extinction risk and conservation. Philosophical
-Transactions of the Royal Society B-Biological Sciences 368.
-
-\leavevmode\hypertarget{ref-Hosner_2020}{}%
-Hosner, P. A., H. L. Owens, E. L. Braun, and R. T. Kimball. 2020.
-Phylogeny and diversification of the gallopheasants (aves: Galliformes):
-Testing roles of sexual selection and environmental niche divergence.
-Zoologica Scripta 49:549--562.
-
-\leavevmode\hypertarget{ref-delhoyo_2016}{}%
-Hoyo, J. del, and N. J. Collar. 2016. HBW and birdlife international
-illustrated checklist of the birds of the world. Lynx Edicions; BirdLife
-International.
-
-\leavevmode\hypertarget{ref-Huang_2014}{}%
-Huang, H. T., and D. L. Rabosky. 2014. Sexual selection and
-diversification: Reexamining the correlation between dichromatism and
-speciation rate in birds. American Naturalist 184:E101--E114.
-
-\leavevmode\hypertarget{ref-Huber_2007}{}%
-Huber, S. K., L. F. De Leon, A. P. Hendry, E. Bermingham, and J. Podos.
-2007. Reproductive isolation of sympatric morphs in a population of
-darwin's finches. Proceedings of the Royal Society of London B:
-Biological Sciences 274:1709--1714.
-
-\leavevmode\hypertarget{ref-Iglesias-Carrasco_2019}{}%
-Iglesias-Carrasco, M., M. D. Jennions, S. Y. W. Ho, and D. A. Duchêne.
-2019. Sexual selection, body mass and molecular evolution interact to
-predict diversification in birds. Proceedings of the Royal Society B:
-Biological Sciences 286:20190172.
-
-\leavevmode\hypertarget{ref-Jablonski_2003}{}%
-Jablonski, D., and K. Roy. 2003. Geographical range and speciation in
-fossil and living molluscs. Proceedings of the Royal Society of London.
-Series B: Biological Sciences 270:401--406. The Royal Society.
-
-\leavevmode\hypertarget{ref-Janicke_2018}{}%
-Janicke, T., M. G. Ritchie, E. H. Morrow, and L. Marie-Orleach. 2018.
-Sexual selection predicts species richness across the animal kingdom.
-Proceedings of the Royal Society B: Biological Sciences 285:20180173.
-The Royal Society.
-
-\leavevmode\hypertarget{ref-Jetz_2012}{}%
-Jetz, W., G. H. Thomas, J. B. Joy, K. Hartmann, and A. O. Mooers. 2012.
-The global diversity of birds in space and time. Nature 491:444--448.
-
-\leavevmode\hypertarget{ref-Kokko_Brooks_2003}{}%
-Kokko, H., and R. Brooks. 2003. Sexy to die for? Sexual selection and
-the risk of extinction. Annales Zoologici Fennici 40:207--219.
-
-\leavevmode\hypertarget{ref-Kokko_Jennions2008}{}%
-Kokko, H., and M. D. Jennions. 2008. Parental investment, sexual
-selection and sex ratios. Journal of Evolutionary Biology 21:919--948.
-
-\leavevmode\hypertarget{ref-Kottler_1980}{}%
-Kottler, M. J. 1980. Darwin, Wallace, and the origin of sexual
-dimorphism. Proceedings of the American Philosophical Society
-124:203--226.
-
-\leavevmode\hypertarget{ref-Kraaijeveld_2011}{}%
-Kraaijeveld, K., F. J. L. Kraaijeveld-Smit, and M. E. Maan. 2011. Sexual
-selection and speciation: The comparative evidence revisited. Biological
-Reviews 86:367--377.
-
-\leavevmode\hypertarget{ref-Lande_1981}{}%
-Lande, R. 1981. Models of speciation by sexual selection on polygenic
-traits. Proceedings of the National Academy of Sciences 78:3721--3725.
-
-\leavevmode\hypertarget{ref-Lande_1982}{}%
-Lande, R. 1982. Rapid origin of sexual isolation and character
-divergence in a cline. Evolution 36:213--223.
-
-\leavevmode\hypertarget{ref-Liow_2008}{}%
-Liow, L. H., M. Fortelius, E. Bingham, K. Lintulaakso, H. Mannila, L.
-Flynn, and N. C. Stenseth. 2008. Higher origination and extinction rates
-in larger mammals. Proceedings of the National Academy of Sciences
-105:6097--6102. National Acad Sciences.
-
-\leavevmode\hypertarget{ref-Long_2012}{}%
-Long, T. A. F., A. F. Agrawal, and L. Rowe. 2012. The effect of sexual
-selection on offspring fitness depends on the nature of genetic
-variation. Current Biology 22:204--208.
-
-\leavevmode\hypertarget{ref-Lorch_2003}{}%
-Lorch, P. D., S. Proulx, L. Rowe, and T. Day. 2003. Condition-dependent
-sexual selection can accelerate adaptation. Evolutionary Ecology
-Research 5:867--881.
-
-\leavevmode\hypertarget{ref-Maan_2011}{}%
-Maan, M. E., and O. Seehausen. 2011. Ecology, sexual selection and
-speciation. Ecology Letters 14:591--602.
-
-\leavevmode\hypertarget{ref-Maia_2013}{}%
-Maia, R., D. R. Rubenstein, and M. D. Shawkey. 2013. Key ornamental
-innovations facilitate diversification in an avian radiation.
-Proceedings of the National Academy of Sciences 110:10687--10692.
-National Acad Sciences.
-
-\leavevmode\hypertarget{ref-Martins_2018}{}%
-Martins, M. J. F., T. M. Puckett, R. Lockwood, J. P. Swaddle, and G.
-Hunt. 2018. High male sexual investment as a driver of extinction in
-fossil ostracods. Nature 556:366.
-
-\leavevmode\hypertarget{ref-Mason_2017}{}%
-Mason, N. A., K. J. Burns, J. A. Tobias, S. Claramunt, N. Seddon, and E.
-P. Derryberry. 2017. Song evolution, speciation, and vocal learning in
-passerine birds. Evolution 71:786--796. Wiley Online Library.
-
-\leavevmode\hypertarget{ref-Mclain_1993}{}%
-McLain, D. K. 1993. Cope's rules, sexual selection, and the loss of
-ecological plasticity. Oikos 490--500. JSTOR.
-
-\leavevmode\hypertarget{ref-Medina_2017}{}%
-Medina, I., K. Delhey, A. Peters, K. E. Cain, M. L. Hall, R. A. Mulder,
-and N. E. Langmore. 2017. Habitat structure is linked to the evolution
-of plumage colour in female, but not male, fairy-wrens. BMC evolutionary
-biology 17:35.
-
-\leavevmode\hypertarget{ref-Moore_2016}{}%
-Moore, B. R., S. Hohna, M. R. May, B. Rannala, and J. P. Huelsenbeck.
-2016. Critically evaluating the theory and performance of bayesian
-analysis of macroevolutionary mixtures. Proceedings of the National
-Academy of Sciences 113:9569--9574.
-
-\leavevmode\hypertarget{ref-Morrow_Speciation_2003}{}%
-Morrow, E. H., T. E. Pitcher, and G. Arnqvist. 2003. No evidence that
-sexual selection is an 'engine of speciation' in birds. Ecology Letters
-6:228--234.
-
-\leavevmode\hypertarget{ref-Otto-Bliesner_2006}{}%
-Otto-Bliesner, B. L., S. J. Marshall, J. T. Overpeck, G. H. Miller, A.
-Hu, and. 2006. Simulating arctic climate warmth and icefield retreat in
-the last interglaciation. Science 311:1751--1753.
-
-\leavevmode\hypertarget{ref-Owens_1999}{}%
-Owens, I. P. F., P. M. Bennett, and P. H. Harvey. 1999. Species richness
-among birds: Body size, life history, sexual selection or ecology?
-Proceedings of the Royal Society B-Biological Sciences 266:933--939.
-
-\leavevmode\hypertarget{ref-owens_1998}{}%
-Owens, I. P., and I. R. Hartley. 1998. Sexual dimorphism in birds: Why
-are there so many different forms of dimorphism? Proceedings of the
-Royal Society of London. Series B: Biological Sciences 265:397--407. The
-Royal Society.
-
-\leavevmode\hypertarget{ref-Pagel_1999}{}%
-Pagel, M. 1999. Inferring the historical patterns of biological
-evolution. Nature 401:877.
-
-\leavevmode\hypertarget{ref-ape_2004}{}%
-Paradis, E., J. Claude, and K. Strimmer. 2004. APE: Analyses of
-phylogenetics and evolution in R language. Bioinformatics 20:289--290.
-
-\leavevmode\hypertarget{ref-Pennell_2013}{}%
-Pennell, T. M., and E. H. Morrow. 2013. Two sexes, one genome: The
-evolutionary dynamics of intralocus sexual conflict. Ecology and
-Evolution 3:1819--1834.
-
-\leavevmode\hypertarget{ref-nlme_2018}{}%
-Pinheiro, J., D. Bates, S. DebRoy, D. Sarkar, and R Core Team. 2018.
-nlme: Linear and nonlinear mixed effects models.
-
-\leavevmode\hypertarget{ref-Pischedda_2006}{}%
-Pischedda, A., and A. K. Chippindale. 2006. Intralocus sexual conflict
-diminishes the benefits of sexual selection. PLOS Biology 4:e356.
-
-\leavevmode\hypertarget{ref-Price_2014}{}%
-Price, J. J., and M. D. Eaton. 2014. Reconstructing the evolution of
-sexual dichromatism: Current color diversity does not reflect past rates
-of male and female change. Evolution 68:2026--2037.
-
-\leavevmode\hypertarget{ref-Price-Waldman_2020}{}%
-Price-Waldman, R. M., A. J. Shultz, and K. J. Burns. 2020. Speciation
-rates are correlated with changes in plumage color complexity in the
-largest family of songbirds. Evolution 74:1155--1169.
-
-\leavevmode\hypertarget{ref-quintero_2018}{}%
-Quintero, I., and W. Jetz. 2018. Global elevational diversity and
-diversification of birds. Nature 555:246.
-
-\leavevmode\hypertarget{ref-Quintero_2015}{}%
-Quintero, I., P. Keil, W. Jetz, and F. W. Crawford. 2015. Historical
-biogeography using species geographical ranges. Systematic Biology
-64:1059--1073.
-
-\leavevmode\hypertarget{ref-Qvarnstrom_2012}{}%
-Qvarnström, A., N. Vallin, and A. Rudh. 2012. The role of male contest
-competition over mates in speciation. Current Zoology 58:493--509.
-Oxford University Press Oxford, Uk.
-
-\leavevmode\hypertarget{ref-bammtools_2014}{}%
-Rabosky, D., M. Grundler, C. Anderson, P. Title, J. Shi, J. Brown, H.
-Huang, and J. Larson. 2014. BAMMtools: An R package for the analysis of
-evolutionary dynamics on phylogenetic trees. Methods in Ecology and
-Evolution 5:701--707.
-
-\leavevmode\hypertarget{ref-Rabosky_2014}{}%
-Rabosky, D. L. 2014. Automatic detection of key innovations, rate
-shifts, and diversity-dependence on phylogenetic trees. PLOS ONE
-9:e89543.
-
-\leavevmode\hypertarget{ref-Rabosky_Extinction_2016}{}%
-Rabosky, D. L. 2016. Challenges in the estimation of extinction from
-molecular phylogenies: A response to Beaulieu and O'Meara. Evolution
-70:218--228.
-
-\leavevmode\hypertarget{ref-Rabosky_fish_2018}{}%
-Rabosky, D. L., J. Chang, P. O. Title, P. F. Cowman, L. Sallan, M.
-Friedman, K. Kaschner, C. Garilao, T. J. Near, M. Coll, and others.
-2018. An inverse latitudinal gradient in speciation rate for marine
-fishes. Nature 559:392.
-
-\leavevmode\hypertarget{ref-Rabosky_BAMM_flawed_2017}{}%
-Rabosky, D. L., J. S. Mitchell, and J. Chang. 2017. Is BAMM flawed?
-Theoretical and practical concerns in the analysis of multi-rate
-diversification models. Systematic biology 66:477--498.
-
-\leavevmode\hypertarget{ref-Rankin_2011}{}%
-Rankin, D. J., U. Dieckmann, and H. Kokko. 2011. Sexual conflict and the
-tragedy of the commons. American Naturalist 177:780--791.
-
-\leavevmode\hypertarget{ref-phytools_2012}{}%
-Revell, L. J. 2012. Phytools: An r package for phylogenetic comparative
-biology (and other things). Methods in Ecology and Evolution 3:217--223.
-
-\leavevmode\hypertarget{ref-Rosenzweig_1995}{}%
-Rosenzweig, M. L. 1995. Species diversity in space and time. Cambridge
-University Press.
-
-\leavevmode\hypertarget{ref-Safran_2013}{}%
-Safran, R. J., E. S. Scordato, L. B. Symes, R. L. Rodrı'guez, and T. C.
-Mendelson. 2013. Contributions of natural and sexual selection to the
-evolution of premating reproductive isolation: A research agenda. Trends
-in Ecology \& Evolution 28:643--650.
-
-\leavevmode\hypertarget{ref-Seddon_2013}{}%
-Seddon, N., C. A. Botero, J. A. Tobias, P. O. Dunn, H. E. A. MacGregor,
-D. R. Rubenstein, J. A. C. Uy, J. T. Weir, L. A. Whittingham, and R. J.
-Safran. 2013. Sexual selection accelerates signal evolution during
-speciation in birds. Proceedings of the Royal Society B: Biological
-Sciences 280.
-
-\leavevmode\hypertarget{ref-Seddon_2008}{}%
-Seddon, N., R. M. Merrill, and J. A. Tobias. 2008. Sexually selected
-traits predict patterns of species richness in a diverse clade of
-suboscine birds. American Naturalist 171:620--631.
-
-\leavevmode\hypertarget{ref-Shine_1989}{}%
-Shine, R. 1989. Ecological causes for the evolution of sexual
-dimorphism: A review of the evidence. The Quarterly Review of Biology
-64:419--461.
-
-\leavevmode\hypertarget{ref-Slatkin_1984}{}%
-Slatkin, M. 1984. Ecological causes of sexual dimorphism. Evolution
-38:622--630.
-
-\leavevmode\hypertarget{ref-sloan2019genitalia}{}%
-Sloan, N. S., and L. W. Simmons. 2019. The evolution of female
-genitalia. Journal of Evolutionary Biology in press.
-
-\leavevmode\hypertarget{ref-swanson1998abalone}{}%
-Swanson, W. J., and V. D. Vacquier. 1998. Concerted evolution in an egg
-receptor for a rapidly evolving abalone sperm protein. Science
-281:710--712.
-
-\leavevmode\hypertarget{ref-Tinghitella_2018}{}%
-Tinghitella, R. M., A. C. Lackey, M. Martin, P. D. Dijkstra, J. P.
-Drury, R. Heathcote, J. Keagy, E. S. Scordato, and A. M. Tyers. 2018. On
-the role of male competition in speciation: A review and research
-agenda. Behavioral Ecology 29:783--797. Oxford University Press UK.
-
-\leavevmode\hypertarget{ref-Rabosky_diversification_2018}{}%
-Title, P. O., and D. L. Rabosky. 2018. Diversification rates and
-phylogenies: What are we estimating, and how good are the estimates?
-bioRxiv 369124.
-
-\leavevmode\hypertarget{ref-tsuji2020sexual}{}%
-Tsuji, K., and T. Fukami. 2020. Sexual dimorphism and species diversity:
-From clades to sites. Trends in Ecology \& Evolution 35:105--114.
-Elsevier.
-
-\leavevmode\hypertarget{ref-Wallace_1889}{}%
-Wallace, A. R. 1889. Colours and ornaments characteristic of sex.
-\emph{in} Darwinism, an exponent of the theory of natural selection,
-with some of its applications. Macmillan; Company.
-
-\leavevmode\hypertarget{ref-Weir_2011}{}%
-Weir, J. T., and T. D. Price. 2011. Limits to speciation inferred from
-times to secondary sympatry and ages of hybridizing species along a
-latitudinal gradient. American Naturalist 177:462--469.
-
-\leavevmode\hypertarget{ref-west1983sexual}{}%
-West-Eberhard, M. J. 1983. Sexual selection, social competition, and
-speciation. Quarterly Review of Biology 58:155--183.
-
-\leavevmode\hypertarget{ref-Whitlock_2000}{}%
-Whitlock, M. C. 2000. Fixation of new alleles and the extinction of
-small populations: Drift load, beneficial alleles, and sexual selection.
-Evolution 54:1855--1861.
-
-\leavevmode\hypertarget{ref-Whitlock_Agrawal_2009}{}%
-Whitlock, M. C., and A. F. Agrawal. 2009. Purging the genome with sexual
-selection: Reducing mutation load through selection on males. Evolution
-63:569--582.
-
-\leavevmode\hypertarget{ref-Zhao_2005}{}%
-Zhao, M., F. A. Heinsch, R. R. Nemani, and S. W. Running. 2005.
-Improvements of the modis terrestrial gross and net primary production
-global data set. Remote Sensing of Environment 95:164--176.
+Data is archived in the adjoined GitHub repository and will be made
+publicly available on acceptance.
 

### 2021-01-18 Manuscript_Publication.Rmd
typo

@@
 The bar under each density ridge is the 95 % Highest Posterior Density Interval.
-Given that the mating system is a categorical variable, model estimates for three polygynous mating system levels are in reference to a striclty monogamous mating system (0% polygyny).
+Given that the mating system is a categorical variable, model estimates for three polygynous mating system levels are in reference to a strictly monogamous mating system (0% polygyny).
 \label{SS_components}](Figures/Individual_SS_Estimates.pdf)

### 2021-01-18 Manuscript_Publication.tex
typo

@@
 categorical variable, model estimates for three polygynous mating system
-levels are in reference to a striclty monogamous mating system (0\%
+levels are in reference to a strictly monogamous mating system (0\%
 polygyny). \label{SS_components}}

### 2021-03-02 Sexual Selection and Speciation.Rmd
update supp mat with names

@@
 title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds
-author: 
-# "Justin G. Cally, Devi Stuart-Fox, Luke Holman, Iliana Medina Guzmán"
+author: "Justin G. Cally, Devi Stuart-Fox, Luke Holman, James Dale and Iliana Medina"
 csl: References/evolution.csl
